### 2019-10-28 scripts/laggedScatterplot.Rmd
Lagged scatterplot .Rmd

---
title: "NLA17 Dissolved Gas Spatial Autocorrelation"
author: "J. Beaulieu"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

During the 2017 National Lakes Assessment, dissolved gas samples and air samples were collected from the index site at each waterbody.  Samples were analyzed for CO~2~, CH~4~, and N~2~O concentration via gas chromatography.  The following is an analysis of spatial autocorrelation via lagged scatterplots

## Analysis

This .rmd assumes 'DissolvedGasNla.prj' working environment.  Data is available as an sf and sp object.
```{r}
# Unique samples are identified by site.id and visit.no.  Some sites were sampled twice during the NLA17 index period.
# sample.source column discriminates between 'AIR' and dissolved gas 'DG" samples.  "AIR" samples were removed in
# readDissolvedGas.R.  Dissolved gas concentrations are reported in mol/L.

# Inspect sf object
dg.sf[ , c("site.id", "visit.no", "sample.source", "dissolved.co2", "dissolved.ch4", "dissolved.n2o")]

# Inspect sp object
dg.sp[ , c("site.id", "visit.no", "sample.source", "dissolved.co2", "dissolved.ch4", "dissolved.n2o")]
```

Bubble plots are a useful tool for visualizing spatial patterns.  Dissolved CH4 and N2O don't display obvious spatial patterns, but dissolved CO2 seems to greater along east coast, upper midwest, and a SW by NE transect between Texas and Wisconsin.  Pretty interesting.

```{r}
# bubble function requires sp object and can't handle missing values.
bubble(dg.sp[!is.na(dg.sp$dissolved.ch4), "dissolved.ch4"]) #
bubble(dg.sp[!is.na(dg.sp$dissolved.co2), "dissolved.co2"])
bubble(dg.sp[!is.na(dg.sp$dissolved.n2o), "dissolved.n2o"])
```

Lagged scatterplots are a useful tool for detecting spatial autocorrelation.  Evidence for spatial autocorrelation is stronger correltation among close vs. distant pairs of observations.  This pattern is not evident for any of the gases.

```{r}
# Need to select appropriate distance bins.
# What are distance between points in the data?
gdis <- spDists(dg.sp)
max(gdis) # 5617, this must be km.  5617 km = 3490 miles, AK to FL

hscat(dissolved.ch4 ~ 1,  
      dg.sp[!is.na(dg.sp$dissolved.ch4), ], # remove missing
      # 5617275m = 3490 miles. AK to FL
      c(0, 1000, 2000, 3000, 4000, 5000, 6000))

hscat(dissolved.co2 ~ 1,  
      dg.sp[!is.na(dg.sp$dissolved.co2), ], # remove missing
      # 5617275m = 3490 miles. AK to FL
      c(0, 1000, 2000, 3000, 4000, 5000, 6000))

hscat(dissolved.n2o ~ 1,  
      dg.sp[!is.na(dg.sp$dissolved.n2o), ], # remove missing
      # 5617275m = 3490 miles. AK to FL
      c(0, 1000, 2000, 3000, 4000, 5000, 6000))

```

Perhaps spatial autocoreltation is more apparent within particular ecoregions.  Can we do lagged scatterplots by ecoregion?
```{r}

# Loop to create lagged scatterplot by ecoregion x gas.  Not working yet....
# Need to customize size bins per ecoregion, I think.
# for(i in 1:length(unique(dg.sp@data$WSA9_NAME))) {
#   WSA9.i <- unique(dg.sp@data$WSA9_NAME)[i] # unique ecoregion
#   dg.sp.i <- filter(dg.sp, WSA9_NAME == WSA9.i) # subset spatial object for ecoregion of interest
#   gDis.1 <- spDists(dg.sp.i) # matrix of distances
#   gDis.2 <- seq(from = 0, to = max(gDis.1), by = max(gDis.1) / 6) # input to breaks argument in hscat()
#   
#   for(j in 1:3) { # for each gas
#     gas.j <- c("dissolved.ch4", "dissolved.co2", "dissolved.n2o")[j] # gas of interest
#     dg.sp.i.j <- filter(dg.sp.i, !is.na(gas.j)) # remove missing
#     
#     hscat(gas.j ~ 1, # not interpretting 'gas.j' correctly.  Works if "dissolved.ch4' is specified.
#           dg.sp.i, 
#           gDis.2) 
#   }
#   
# }
```

### 2019-10-29 scripts/laggedScatterplot.Rmd
Added self_contained to YAML, but still producing *_files folder.

@@
 author: "J. Beaulieu"
-date: "10/28/2019"
-output: html_document
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: 
+  html_document:
+    self_contained: yes
+    mode: selfcontained
 ---
@@
 
-hscat(dissolved.co2 ~ 1,  
-      dg.sp[!is.na(dg.sp$dissolved.co2), ], # remove missing
-      # 5617275m = 3490 miles. AK to FL
-      c(0, 1000, 2000, 3000, 4000, 5000, 6000))
-
-hscat(dissolved.n2o ~ 1,  
-      dg.sp[!is.na(dg.sp$dissolved.n2o), ], # remove missing
-      # 5617275m = 3490 miles. AK to FL
-      c(0, 1000, 2000, 3000, 4000, 5000, 6000))
+# hscat(dissolved.co2 ~ 1,  
+#       dg.sp[!is.na(dg.sp$dissolved.co2), ], # remove missing
+#       # 5617275m = 3490 miles. AK to FL
+#       c(0, 1000, 2000, 3000, 4000, 5000, 6000)) 
+# 
+# hscat(dissolved.n2o ~ 1,  
+#       dg.sp[!is.na(dg.sp$dissolved.n2o), ], # remove missing
+#       # 5617275m = 3490 miles. AK to FL
+#       c(0, 1000, 2000, 3000, 4000, 5000, 6000))
 

### 2019-10-29 identifyLocalPath.R
Adding .Rprofile to customize directory paths per user.

@@ -1,0 +1,9 @@
+if (grepl("JBEAULIE", getwd())) {
+  localPath = "C:/Users/JBEAULIE/"
+}
+
+
+
+if (grepl("MCMAN", getwd())) {
+  localPath = "C:/Users/MMCMANUS/"
+}

### 2019-10-29 scripts/readDissolvedGas.R

@@ -1,11 +1,18 @@
 # READ IN DISSOLVED GAS DATA
 # This file contains the aggregated (by duplicates) dissolved gas data
 # from NLA17.  See data dictionary in documents library for details.
+# 
+# dg <- read.table(file = paste0("C:/Users/JBEAULIE/Environmental Protection Agency (EPA)/",
+#                                "ORD NLA17 Dissolved Gas - Documents/",
+#                                "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
+#                  header = TRUE, sep = "\t", as.is = TRUE)
 
-dg <- read.table(file = paste0("C:/Users/JBEAULIE/Environmental Protection Agency (EPA)/",
+dg <- read.table(file = paste0(localPath, 
+                               "Environmental Protection Agency (EPA)/",
                                "ORD NLA17 Dissolved Gas - Documents/",
                                "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
                  header = TRUE, sep = "\t", as.is = TRUE)
+
 
 # sp
 # Define coordinates

### 2019-10-30 identifyLocalPath.R
New solution for identifying local path.  Removed from .Rprofile to improve reproducibility.

@@ -1,9 +1,0 @@
-if (grepl("JBEAULIE", getwd())) {
-  localPath = "C:/Users/JBEAULIE/"
-}
-
-
-
-if (grepl("MCMAN", getwd())) {
-  localPath = "C:/Users/MMCMANUS/"
-}

### 2019-10-30 scripts/identifyLocalPath.R

@@ -1,0 +1,1 @@
+localPath <- Sys.getenv("USERPROFILE") 

### 2019-10-30 scripts/masterScript.R

@@ -1,5 +1,6 @@
-source("scripts/masterLibrary.R")
-source("scripts/readDissolvedGas.R")
+source("scripts/identifyLocalPath.R")  # set local path for each user
+source("scripts/masterLibrary.R") # libraries
+source("scripts/readDissolvedGas.R") # read data
 
 # probably can't source this because it depends on larger environment.  Not sure.
 source("scripts/laggedScatterplot.Rmd") 

### 2019-10-30 scripts/readDissolvedGas.R

@@ -8,7 +8,7 @@
 #                  header = TRUE, sep = "\t", as.is = TRUE)
 
 dg <- read.table(file = paste0(localPath, 
-                               "Environmental Protection Agency (EPA)/",
+                               "/Environmental Protection Agency (EPA)/",
                                "ORD NLA17 Dissolved Gas - Documents/",
                                "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
                  header = TRUE, sep = "\t", as.is = TRUE)

### 2019-11-12 scripts/dgIndicatorAnalysis.Rmd
Adding .rmd that reads in raw data and begins exploratory data analysis.  Having trouble knitting.

---
title: "Analysis of NLA17 Dissolved Gas Indicator"
author: "J. Beaulieu"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: false
    fig_caption: yes
    depth: 2
    number_sections: true
---

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(rgdal) # spatial data
library(sf) # spatial data
library(gstat) # lagged scatterplot
library(spdplyr) # apply dplyr verbs to sp objects
library(rmarkdown)
library(tidyverse)
```

# Background
During the 2017 National Lakes Assesment, duplicate dissolved gas samples were collected at a depth of ~0.1m at the index site at each waterbody.  Gas samples were analyzed for CO~2~, CH~4~, and N~2~O concentration via gas chromatography and $\delta^{13}CO_{2}$ and $\delta^{13}CH_{4}$ via cavity ring-down spectroscopy.  The following analysis pertains to the concentration data.

# Data

The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.  A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

```{r}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Read data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
                 header = TRUE, sep = "\t", as.is = TRUE)
```
Concentrations of CO~2~, CH~4~, and N~2~O in ambient air over the waterbodies were measured and used in the dissolved gas concentrations.  These data are not needed in subsequent analysis and will be removed from the dataframe.
```{r}
dg <- dg %>% filter(sample.source == "DG") # filter out dissolved gas 'DG' samples
```
The dataframe also contains observations from site in Alaska.  These sites were not an official part of the NLA and will not be included in this analysis.
```{r}
dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
```

The data file contains `r dg %>% distinct(site.id, visit.no) %>% summarize(n = n()) %>% pull()` unique sampling trips defined by unique combinations of 'site.id' and 'visit.no'.  There were `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies sampled, `r dg %>% filter(visit.no == 2) %>% summarize(n = n()) %>% pull()` of which were sampled twice. The primary response variables are the concentrations of CO~2~, CH~4~, and N~2~O dissolved in the surface water of each waterbody.  Concentration is expressed as both mol/L (dissolved.co2, dissolved.ch4, dissolved.n2o) and as a saturation ratio (co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio).  Saturation ratio is defined as the ratio of the measured concentration to that expected if the waterbody were in equilibrium with the atmosphere.  A value of 1 indicates equilibrium, values < 1 indicate undersaturation and values > 1 indicate supersaturation.  Undersaturated waterbodies function as a sink for the atmospheric gas whereas supersaturated waterbodies function as a source.

```{r}
dg %>% select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
  {summary(.)}
```

To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
```{r}
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

#  Exploratory spatial analysis

Bubble plots offer a simple method for visualizing spatial patterns in the data.  For some reason, this isn't working.  WTF....

```{r}
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "Aggr_Ecoregions9_2015")

# "Temperate" is misspelled.
ecoR <- ecoR %>% 
  mutate(WSA9_NAME = as.character(WSA9_NAME), # conv to char
         WSA9_NAME = ifelse(WSA9_NAME == "Temporate Plains",
                            "Temperate Plains", # correct sp
                            WSA9_NAME),
         WSA9_NAME = as.factor(WSA9_NAME)) # back to factor

# Check CRS
st_crs(ecoR) # no EPSG code
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
st_crs(ecoR) # 5070

# Bubble plots
ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf, aes(size = dissolved.co2)) +
  ggtitle("Dissolved CO2")

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf, aes(size = dissolved.ch4)) +
  ggtitle("Dissolved CH4")

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf, aes(size = dissolved.n2o)) +
  ggtitle("Dissolved N2O")
```

### 2019-11-12 scripts/masterScript.R

@@ -1,6 +1,8 @@
-source("scripts/identifyLocalPath.R")  # set local path for each user
-source("scripts/masterLibrary.R") # libraries
-source("scripts/readDissolvedGas.R") # read data
+# THESE CAN BE SKIPPED IF WORKING IN dgIndicatorAnalysis.Rmd
+# source("scripts/identifyLocalPath.R")  # set local path for each user
+# source("scripts/masterLibrary.R") # libraries
+# source("scripts/readDissolvedGas.R") # read data
 
-# probably can't source this because it depends on larger environment.  Not sure.
-source("scripts/laggedScatterplot.Rmd") 
+# # probably can't source this because it depends on larger environment.  Not sure.
+# # This will likely be replaced with dgIndicatorAnalysis.Rmd
+# source("scripts/laggedScatterplot.Rmd") 

### 2019-11-12 scripts/readDissolvedGas.R

@@ -28,4 +28,4 @@ dg.sp@proj4string <- CRS('+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs')
 dg.sp <- filter(dg.sp, sample.source == "DG")
 
 # sf
-dg.sf <- st_as_sf(dg.sp, coords = c("map.lon.dd", "map.lat.dd"), crs = 4269)
+adg.sf <- st_as_sf(dg.sp, coords = c("map.lon.dd", "map.lat.dd"), crs = 4269)

### 2019-11-15 scripts/dgIndicatorAnalysis.Rmd
Added population estimate figures and variograms to .Rmd

@@
     toc: true
-    toc_float: false
+    toc_float: true
     fig_caption: yes
@@
     number_sections: true
+    code_folding: hide
+editor_options: 
+  chunk_output_type: console
 ---
@@
 # load libraries
-library(rgdal) # spatial data
 library(sf) # spatial data
-library(gstat) # lagged scatterplot
-library(spdplyr) # apply dplyr verbs to sp objects
-library(rmarkdown)
-library(tidyverse)
+library(gstat) # lagged scatterplot, variogram
+#library(spdplyr) # apply dplyr verbs to sp objects.  gstat functions require sp objects
+library(tidyverse) # dplyr, ggplot
+
+# Identify local path for each user
+localPath <- Sys.getenv("USERPROFILE")
+
+# Define helper functions
+# standardized formatting for column names
+toEPA <- function(X1){
+  names(X1) = tolower(names(X1))
+  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
+  X1
+}
 ```
@@
 
-The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.
+The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
 After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.
-A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
-When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
+A copy of this data file is stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
+When synced to a local computer, the documents library can be read directly with R.
 
 ```{r}
-# Identify local path for each user
-localPath <- Sys.getenv("USERPROFILE")
-
 # Read data file
@@
                                "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
-                 header = TRUE, sep = "\t", as.is = TRUE)
+                 header = TRUE, sep = "\t", as.is = TRUE) %>%
+  # unit conversion
+  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
+         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
+         dissolved.n20 = dissolved.n2o * 1000000000, # mol/L -> nmol/L
+         # add source/sink column
+         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
+         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
+         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
 ```
@@
 ```{r}
-dg <- dg %>% filter(sample.source == "DG") # filter out dissolved gas 'DG' samples
+dg <- dg %>% filter(sample.source == "DG") # select dissolved gas 'DG' samples
 ```
@@
 The primary response variables are the concentrations of CO~2~, CH~4~, and N~2~O dissolved in the surface water of each waterbody.
-Concentration is expressed as both mol/L (dissolved.co2, dissolved.ch4, dissolved.n2o) and as a saturation ratio (co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio).
+Concentration is expressed in both mass based units (umol/L, nmol/L: dissolved.co2, dissolved.ch4, dissolved.n2o) and as a saturation ratio (co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio).
 Saturation ratio is defined as the ratio of the measured concentration to that expected if the waterbody were in equilibrium with the atmosphere.
@@
 
-To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
+#  Exploratory spatial analysis
+## Population estimates by ecoregion
+The survey design allows for population estimates of central tendency and variance by ecoregion (and other subpopulations), which can provide insight into spatial patterning.
+There is broad overlap in mean +/- 95% CI estimates for dissolved CH4 concentation.
+The Northern Plains, Temperate Plains, and Western Mountains have low dissolved CO2 concentration.
+
+```{r, fig.cap= "Fig. caption: Mean +/- 95% CI for the dissolved gas concentration indicator"}
+# Read population estimates provided by Steve Paulsen (11/7/2019)
+pop <- read.csv(file = paste0(localPath, 
+                              "/Environmental Protection Agency (EPA)/",
+                              "ORD NLA17 Dissolved Gas - Documents/",
+                              "inputData/populationEstimates/",
+                              "NLA_2017_Percentile_Estimates_20191107.csv"),
+                as.is = TRUE) %>%
+  toEPA()  # enforce formatting convention
+
+# Convert units for dissolved gas concentration indicator
+pop <- pop %>%
+  mutate(estimate = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
+                            estimate * 1000000, # mol/L -> umol/L for co2 and ch4
+                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
+                            estimate * 1000000000, # mol/L -> nmol/L for n2o
+                            estimate)), # all others
+         lcb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
+                            lcb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
+                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
+                            lcb95pct * 1000000000, # mol/L -> nmol/L for n2o
+                            lcb95pct)), # all others
+         ucb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
+                            ucb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
+                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
+                            ucb95pct * 1000000000, # mol/L -> nmol/L for n2o
+                            ucb95pct))) # all others
+
+# extract population estimates for ecoregion and nation
+pop.eco <- pop %>% 
+  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
+         statistic == "50Pct") # extract mean estimate
+
+# extract dissolved gas indicator, order by dissolved CH4
+pop.eco.diss <- pop.eco %>%
+  filter(grepl("dissolved", indicator, ignore.case = TRUE)) %>%
+  mutate(subpopulation = factor(subpopulation, 
+                             levels = filter(pop.eco, indicator == "DISSOLVED_CH4", subpopulation != "National") %>%
+                               arrange(estimate) %>%
+                               select(subpopulation) %>%
+                               pull() %>%
+                               c("National", .)))
+
+# look at dissolved concentration indicator
+ggplot(pop.eco.diss, aes(estimate, subpopulation)) +
+  geom_point() +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct)) +
+  facet_wrap(~indicator, scales = "free_x") +
+  theme(axis.title = element_blank())
+
+```
+
+## Bubble plots
+Bubble plots can be used to visualize patterns in geospatial data.
+In the plots below, bubble size reflects the gas concentration and color reflects saturation status (i.e.
+source or sink).
+The code immediately below converts the dataframe to a spatial object and reads in the ecoregion ploygons.
+The first bubble plot below is of dissolved CO2 which seems to show a pattern of higher concentrations along the eastern seaboard.
+
 ```{r}
+# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
 # Define coordinates
@@
   st_transform(5070) # project to CONUS ALBERS for plotting
-```
 
-#  Exploratory spatial analysis
-
-Bubble plots offer a simple method for visualizing spatial patterns in the data.  For some reason, this isn't working.  WTF....
-
-```{r}
 # read in ecoregion polygons
@@
 st_crs(ecoR) # 5070
+```
 
-# Bubble plots
+```{r}
+# CO2 Bubble plots
 ggplot() +
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.co2)) +
-  ggtitle("Dissolved CO2")
+  geom_sf(data = dg.sf, aes(size = dissolved.co2, color = co2.src.snk), 
+          show.legend = "point") + # improve legend
+  ggtitle("Dissolved CO2 (uM)") +
+  scale_color_manual(values = c("white", "black"), name = "source/sink") +
+  scale_size(name = "dissolved CO2 (uM)",
+             range = c(0.1, 5), # custom size range
+             breaks = c(10, 100, 300, 600)) # custom breaks
+```
 
+Spatial patterns in dissolved CH4 are not apparent, although all but two sites are supersaturated.
+
+```{r}
 ggplot() +
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.ch4)) +
-  ggtitle("Dissolved CH4")
+  geom_sf(data = dg.sf, aes(size = dissolved.ch4, color = ch4.src.snk), 
+          show.legend = "point") +
+  ggtitle("Dissolved CH4 (uM)") +
+  scale_color_manual(values = c("white", "black"), name = "source/sink") +
+  scale_size(name = "dissolved CH4 (uM)",
+             range = c(0.1, 10), # custom size range
+             breaks = c(0.1, 1, 10, 50, 100)) # custom breaks
+```
 
+The prevalance of undersaturated dissolved N2O concentrations is an important finding.
+The IPCC assumes surface waters are an important source of N2O to the atmosphere.
+Webb et al.
+recently published a paper in PNAS reporting that 70 out of 101 reservoirs sampled in Saskatchewan were N2O sinks.
+This paper made quite a splash.
+In our study, `r sum(dg$n2o.src.snk == "sink")` out of `r sum(!is.na(dg$dissolved.n2o))` sampled waterbodies were an N2O sink.
+It is somewhat interesting to note that the some of the highest concentrations were observed in Indiana/Ohio, where I conducted much of my dissertation research on aquatic N2O emissions.
+
+```{r}
 ggplot() +
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.n2o)) +
-  ggtitle("Dissolved N2O")
+  geom_sf(data = dg.sf, aes(size = dissolved.n2o, color = n2o.src.snk),
+          show.legend = "point") +
+  ggtitle("Dissolved N2O (nM)") +
+  scale_color_manual(values = c("white", "black"), name = "source/sink") +
+  scale_size(name = "dissolved N2O (nM)",
+             range = c(0.1, 10), # custom size range
+             breaks = c(1, 10, 25, 50, 100)) # custom breaks
 ```
+
+We can also look at population estimates of N2O saturation ratio by ecoregion.
+Wow, the mean N2O saturation is less than one for all ecoregions and the nation!
+95% CI overlap one in only 3 levels of the stratification.
+This is pretty big news.
+```{r}
+
+# extract saturation indicator, order by n2o
+pop.eco.sat <- pop.eco %>%
+  filter(grepl("sat", indicator, ignore.case = TRUE)) %>%
+  mutate(subpopulation = factor(subpopulation, 
+                             levels = filter(pop.eco, indicator == "N2O_SAT_RATIO", subpopulation != "National") %>%
+                               arrange(estimate) %>%
+                               select(subpopulation) %>%
+                               pull() %>%
+                               c("National", .)))
+
+# look at dissolved concentration indicator
+ggplot(pop.eco.sat, aes(estimate, subpopulation)) +
+  geom_point() +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct)) +
+  facet_wrap(~indicator, scales = "free_x") +
+  theme(axis.title = element_blank())
+
+```
+
+## Lagged scatterplots
+Mike, can you integrate scatterplot code here?
+
+## Semivariograms
+Variograms display the variability among pairs of points at various separation distances.
+I was unable to fit a model to the dissolved gas concentration variograms using the default fitting parameters in gstat.
+The variograms suggest a lack of spatial structure, but I suspect the model parameterization can be improved.
+
+```{r, warning=FALSE, message=FALSE}
+# Variogram function requires sp object
+dg.sp <- as(dg.sf, Class = "Spatial")
+
+# CO2 variogram
+co2.v <- variogram(dissolved.co2~1, dg.sp)
+co2.v.fit <- fit.variogram(co2.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
+# co2.v.fit # chose sperical
+co2.v.fit <- fit.variogram(co2.v, vgm("Sph")) # refit spherical, several warnings (?)
+
+# CH4 variogram
+ch4.v <- variogram(dissolved.ch4~1, dg.sp)
+ch4.v.fit <- fit.variogram(ch4.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
+# ch4.v.fit # chose exp
+ch4.v.fit <- fit.variogram(ch4.v, vgm("Exp")) # refit exponential, no convergence
+
+# N2O variogram
+n2o.v <- variogram(dissolved.n2o~1, dg.sp)
+n2o.v.fit <- fit.variogram(n2o.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
+# n2o.v.fit # chose exp
+n2o.v.fit <- fit.variogram(n2o.v, vgm("Exp")) # refit exponential, no convergence
+
+# Plot variograms and models (if fit)
+plot(co2.v, model = co2.v.fit, main = "CO2 semivariogram")
+plot(ch4.v, main = "CH4 semivariogram")
+plot(n2o.v, main = "N2O semivariogram")
+
+```

### 2019-11-15 scripts/masterLibrary.R

@@ -3,3 +3,11 @@ library(sf) # spatial data
 library(gstat) # lagged scatterplot
 library(spdplyr) # apply dplyr verbs to sp objects
 library(rmarkdown)
+
+
+# ENFORCE EPA FORMAT ON NLA NAMES
+toEPA <- function(X1){
+  names(X1) = tolower(names(X1))
+  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
+  X1
+}

### 2019-12-10 scripts/dgIndicatorAnalysis.Rmd
Distances calculated and lagged scatter plots made.

--- 2019-11-12 scripts/dgIndicatorAnalysis.Rmd
@@
 title: "Analysis of NLA17 Dissolved Gas Indicator"
-author: "J. Beaulieu"
+authors: "J. Beaulieu, R. Martin, and M. McManus"
 date: "`r Sys.Date()`"
@@
     number_sections: true
+editor_options: 
+  chunk_output_type: console
 ---
@@
 When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
-
+## Getting data and filtering
 ```{r}
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
+                               "inputData/nla17gasDataAggregated_2019-11-12.txt"), 
                  header = TRUE, sep = "\t", as.is = TRUE)
@@
 ```{r}
+# are these the main response variables we'll be looking at?
 dg %>% select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
@@
 
+# head(ecoR, 9) throws Error: C stack usage  30746119 is too close to the limit
+# eventuall get the head output
 # Bubble plots
@@
 ```
+
+## Summarizing distances
+Getting distances between sites overall and by ecoregion to help set distances for lagged scatterplots.
+First, Worked out steps on a sample of NLA sites.
+Could only get parts of Anna Springsteen's code to work.
+```{r, distances}
+names(dg.sf)
+head(dg.sf$WSA9_NAME)
+distinct(dg.sf, WSA9_NAME) # is distinct on sf object only returning first 10 distinct coordinates, and not WSA9_NAME
+distinct.Spatial(dg.sf, WSA9_NAME) # this doesn't work either.
+distinct(dg, WSA9_NAME)
+
+# take a small sample of dg.sf to test getting distances
+set.seed(1859)
+# st_sample only returns a list of coordinates
+nla_samp <- st_sample(dg.sf, 10, type = "random", exact = TRUE)
+plot(nla_samp)
+class(nla_samp)
+head(nla_samp)
+
+# sample_n keeps attributes and geometries
+nla_samp1 <- sample_n(dg.sf, 10)
+plot(nla_samp1, max.plot=1)
+class(nla_samp1)
+head(nla_samp1$site.id, 10)
+distmat <- st_distance(nla_samp1,nla_samp1,by_element = FALSE)
+class(distmat) # thought this would return matrix
+max(distmat)
+min(distmat)
+summary(distmat) # think this returns summary of all pairwise distances on each obs, v1-v10
+dim(distmat)
+
+rdistmat <- distmat[1:10, ]
+print(rdistmat)
+rdistmat1 <- as.vector(rdistmat)
+round(min(rdistmat1[rdistmat1!=0]))
+round(quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[2])
+round(median(rdistmat1[rdistmat1!=0]))
+round(quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[4])
+round(max(rdistmat1[rdistmat1!=0]))
+
+names(ecoR)
+class(ecoR)
+# contrast ecoRname <- as.data.frame(select(ecoR,"WSA9_NAME"))
+# geom is sticky with select but not pull
+# ecoRname <- as.data.frame(pull(ecoR,"WSA9_NAME"))
+ecoRname1 <- as.character(pull(ecoR,"WSA9_NAME")) #Springsteen code needs to be character to populate columns of matrix below
+class(ecoRname1)
+head(ecoRname1, 9)
+# initialize summary tables
+euc.summary <- matrix(NA, nrow=5, ncol=length(ecoRname1)+1) # initializing blank matrices to fill in the loop
+dim(euc.summary)
+colnames(euc.summary) <- c("ALL",ecoRname1) #indicates same assignment in each object
+rownames(euc.summary) <- c("Min","Q1","Median","Q3","Max")
+euc.summary
+
+# Add all distances summary statistics to column 1 of the summary table (i.e. not by region)
+euc.summary[,1] <- round(c( min(rdistmat1[rdistmat1!=0]), quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[2], median(rdistmat1[rdistmat1!=0]), quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[4], max(rdistmat1[rdistmat1!=0]) ),1)
+
+print(euc.summary)
+
+# Unsure that Springsteen function below will work as euc is an spDists created object that I think retained ecoregional & not sure that's so with rdistmat1
+# Add within-region summary statistics to summary table 
+for (r in 1:length(ecoRname1)){ # loop through regions and fill in the matrices
+  euc.reg <- euc[grep(ecoRname1[r], rownames(euc)), grep(ecoRname1[r],colnames(euc))] # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
+  # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
+  euc.summary[, (r+1)] <- round(c( min(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[2], median(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[4], max(euc.reg[euc.reg!=0]) ),1)
+}
+
+# Distances for all NLA sites
+distmat_dg <- st_distance(dg.sf,dg.sf,by_element = FALSE)
+rdistmat_dg <- distmat_dg[1:1185, ]
+rdistmat1_dg <- as.vector(rdistmat_dg)
+
+# initialize summary tables
+euc.summary_dg <- matrix(NA, nrow=5, ncol=length(ecoRname1)+1) # initializing blank matrices to fill in the loop
+dim(euc.summary_dg)
+colnames(euc.summary_dg) <- c("ALL",ecoRname1) #indicates same assignment in each object
+rownames(euc.summary_dg) <- c("Min","Q1","Median","Q3","Max")
+
+euc.summary_dg[,1] <- round(c( min(rdistmat1_dg[rdistmat1_dg!=0]), quantile(rdistmat1_dg[rdistmat1_dg!=0], probs=seq(0,1, 0.25), names=F)[2], median(rdistmat1_dg[rdistmat1_dg!=0]), quantile(rdistmat1_dg[rdistmat1_dg!=0], probs=seq(0,1, 0.25), names=F)[4], max(rdistmat1_dg[rdistmat1_dg!=0]) ),1)
+
+print(euc.summary_dg)
+
+# brute force to get 5-number summary of ecoregions cut-n-paste ecoregions to get specific data frames:  cpl, nap, npl, sap, spl, tpl, umw, wmt, xer
+names(dg.sf)
+
+count(dg.sf, WSA9_NAME) %>% print(n = Inf)
+
+# Get ecoregion-specific NLA sites
+cpl <- dg.sf %>% filter(WSA9_NAME == "Coastal Plains")
+
+distmat_cpl <- st_distance(cpl, cpl,by_element = FALSE)
+dim(distmat_cpl)
+rdistmat_cpl <- distmat_cpl[1:150, ] # change to match # sites/rows
+rdistmat1_cpl <- as.vector(rdistmat_cpl)
+
+round(min(rdistmat1_cpl[rdistmat1_cpl!=0]))
+round(quantile(rdistmat1_cpl[rdistmat1_cpl!=0], probs=seq(0,1, 0.25), names=F)[2])
+round(median(rdistmat1_cpl[rdistmat1_cpl!=0]))
+round(quantile(rdistmat1_cpl[rdistmat1_cpl!=0], probs=seq(0,1, 0.25), names=F)[4])
+round(max(rdistmat1_cpl[rdistmat1_cpl!=0]))
+
+# GitHub question:  if I save a workspace, .RData, will that be pushed when I push to GitHub?  Or, should data frames go in .gitignore?
+```
+
+## Lagged scatterplots
+Note that can use simple feature object, such as cpl, for histogram, but spatial object needed for lagged scatterplots.  I took log10 transform to  get data somewhat unimodal and symmetric.
+```{r, laggedscatterplot}
+# cpl
+cpl_sp <- as(cpl, "Spatial")
+class(cpl_sp)
+ggplot(cpl,aes(dissolved.ch4)) + geom_histogram()
+ggplot(cpl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.ch4~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.ch4~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
+
+ggplot(cpl,aes(dissolved.co2)) + geom_histogram()
+ggplot(cpl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.co2~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(log10(dissolved.co2)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
+
+ggplot(cpl,aes(dissolved.n2o)) + geom_histogram()
+ggplot(cpl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.n2o~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(log10(dissolved.n2o)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
+
+# nap
+nap_sp <- as(nap, "Spatial")
+class(nap_sp)
+ggplot(nap,aes(dissolved.ch4)) + geom_histogram()
+ggplot(nap,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.ch4~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.ch4)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(nap,aes(dissolved.co2)) + geom_histogram()
+ggplot(nap,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.co2~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.co2)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(nap,aes(dissolved.n2o)) + geom_histogram()
+ggplot(nap,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.n2o~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.n2o)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+# npl
+npl_sp <- as(npl, "Spatial")
+class(npl_sp)
+ggplot(npl,aes(dissolved.ch4)) + geom_histogram()
+ggplot(npl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.ch4~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.ch4)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(npl,aes(dissolved.co2)) + geom_histogram()
+ggplot(npl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.co2~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.co2)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(npl,aes(dissolved.n2o)) + geom_histogram()
+ggplot(npl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.n2o~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.n2o)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+# sap
+sap_sp <- as(sap, "Spatial")
+class(sap_sp)
+ggplot(sap,aes(dissolved.ch4)) + geom_histogram()
+ggplot(sap,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(dissolved.ch4~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(log10(dissolved.ch4)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(sap,aes(dissolved.co2)) + geom_histogram()
+ggplot(sap,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(dissolved.co2~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(log10(dissolved.co2)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(sap,aes(dissolved.n2o)) + geom_histogram()
+ggplot(sap,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(dissolved.n2o~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(log10(dissolved.n2o)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+# spl
+spl_sp <- as(spl, "Spatial")
+class(spl_sp)
+ggplot(spl,aes(dissolved.ch4)) + geom_histogram()
+ggplot(spl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(dissolved.ch4~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(log10(dissolved.ch4)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(spl,aes(dissolved.co2)) + geom_histogram()
+ggplot(spl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(dissolved.co2~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(log10(dissolved.co2)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(spl,aes(dissolved.n2o)) + geom_histogram()
+ggplot(spl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(dissolved.n2o~1, spl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.n2o)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+# tpl
+tpl_sp <- as(tpl, "Spatial")
+class(tpl_sp)
+ggplot(tpl,aes(dissolved.ch4)) + geom_histogram()
+ggplot(tpl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(dissolved.ch4~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(log10(dissolved.ch4)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(tpl,aes(dissolved.co2)) + geom_histogram()
+ggplot(tpl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(dissolved.co2~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(log10(dissolved.co2)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(tpl,aes(dissolved.n2o)) + geom_histogram()
+ggplot(tpl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(dissolved.n2o~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(log10(dissolved.n2o)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+# umw
+umw_sp <- as(umw, "Spatial")
+class(umw_sp)
+ggplot(umw,aes(dissolved.ch4)) + geom_histogram()
+ggplot(umw,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
+
+hscat(dissolved.ch4~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
+
+hscat(log10(dissolved.ch4)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+ggplot(umw,aes(dissolved.co2)) + geom_histogram()
+ggplot(umw,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
+
+hscat(dissolved.co2~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+hscat(log10(dissolved.co2)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(umw,aes(dissolved.n2o)) + geom_histogram()
+ggplot(umw,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+hscat(dissolved.n2o~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+hscat(log10(dissolved.n2o)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+# wmt
+wmt_sp <- as(wmt, "Spatial")
+class(wmt_sp)
+ggplot(wmt,aes(dissolved.ch4)) + geom_histogram()
+ggplot(wmt,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(dissolved.ch4~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(log10(dissolved.ch4)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
+
+ggplot(wmt,aes(dissolved.co2)) + geom_histogram()
+ggplot(wmt,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(dissolved.co2~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(log10(dissolved.co2)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
+
+ggplot(wmt,aes(dissolved.n2o)) + geom_histogram()
+ggplot(wmt,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(dissolved.n2o~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(log10(dissolved.n2o)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
+
+# xer
+xer_sp <- as(xer, "Spatial")
+class(xer_sp)
+ggplot(xer,aes(dissolved.ch4)) + geom_histogram()
+ggplot(xer,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+# 10 km bin/lag only has 2 points plotted so start with 50 km
+hscat(log10(dissolved.ch4)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(dissolved.ch4~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(log10(dissolved.ch4)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
+
+ggplot(xer,aes(dissolved.co2)) + geom_histogram()
+ggplot(xer,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(dissolved.co2~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(log10(dissolved.co2)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
+
+ggplot(xer,aes(dissolved.n2o)) + geom_histogram()
+ggplot(xer,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(dissolved.n2o~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(log10(dissolved.n2o)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
+
+# all
+dg_sp <- as(dg.sf, "Spatial")
+class(dg_sp)
+ggplot(dg.sf,aes(dissolved.ch4)) + geom_histogram()
+ggplot(dg.sf,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(dissolved.ch4~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(log10(dissolved.ch4)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(dg.sf,aes(dissolved.co2)) + geom_histogram()
+ggplot(dg.sf,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(dissolved.co2~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(log10(dissolved.co2)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(dg.sf,aes(dissolved.n2o)) + geom_histogram()
+ggplot(dg.sf,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(dissolved.n2o~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(log10(dissolved.n2o)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
+
+````

### 2019-12-10 scripts/dgIndicatorAnalysis.Rmd

@@
     number_sections: true
+    code_folding:  hide
 editor_options: 
@@
 ## Getting data and filtering
+
+The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.
+The code can be found at a private github repository (https://github.com/USEPA/NLA).
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.
+A copy of this data file is stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
+When synced to a local computer, the documents library can be read directly with R.
+
 ```{r}
-# Identify local path for each user
-localPath <- Sys.getenv("USERPROFILE")
-
 # Read data file
@@
                            "inputData/nla17gasDataAggregated_2019-11-12.txt"), 
-                 header = TRUE, sep = "\t", as.is = TRUE)
+                 header = TRUE, sep = "\t", as.is = TRUE) %>%
+  # unit conversion
+  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
+         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
+         dissolved.n20 = dissolved.n2o * 1000000000, # mol/L -> nmol/L
+         # add source/sink column
+         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
+         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
+         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
+
 ```
@@
 st_crs(ecoR) # 5070
+```
 
-# head(ecoR, 9) throws Error: C stack usage  30746119 is too close to the limit
-# eventuall get the head output
 # Bubble plots
-ggplot() +
-  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.co2)) +
-  ggtitle("Dissolved CO2")
+
+```{r}
+# CO2 Bubble plots
 
@@
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.ch4)) +
-  ggtitle("Dissolved CH4")
+  geom_sf(data = dg.sf, aes(size = dissolved.co2, color = co2.src.snk), 
+          show.legend = "point") + # improve legend
+  ggtitle("Dissolved CO2 (uM)") +
+  scale_color_manual(values = c("white", "black"), name = "source/sink") +
+  scale_size(name = "dissolved CO2 (uM)",
+             range = c(0.1, 5), # custom size range
+             breaks = c(10, 100, 300, 600)) # custom breaks
+```
 
+Spatial patterns in dissolved CH4 are not apparent, although all but two sites are supersaturated.
+
+```{r}
 ggplot() +
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.n2o)) +
-  ggtitle("Dissolved N2O")
+  geom_sf(data = dg.sf, aes(size = dissolved.ch4, color = ch4.src.snk), 
+          show.legend = "point") +
+  ggtitle("Dissolved CH4 (uM)") +
+  scale_color_manual(values = c("white", "black"), name = "source/sink") +
+  scale_size(name = "dissolved CH4 (uM)",
+             range = c(0.1, 10), # custom size range
+             breaks = c(0.1, 1, 10, 50, 100)) # custom breaks
+```
+
+The prevalance of undersaturated dissolved N2O concentrations is an important finding.
+The IPCC assumes surface waters are an important source of N2O to the atmosphere.
+Webb et al.
+recently published a paper in PNAS reporting that 70 out of 101 reservoirs sampled in Saskatchewan were N2O sinks.
+This paper made quite a splash.
+In our study, `r sum(dg$n2o.src.snk == "sink")` out of `r sum(!is.na(dg$dissolved.n2o))` sampled waterbodies were an N2O sink.
+It is somewhat interesting to note that the some of the highest concentrations were observed in Indiana/Ohio, where I conducted much of my dissertation research on aquatic N2O emissions.
+
+```{r}
+ggplot() +
+  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+  geom_sf(data = dg.sf, aes(size = dissolved.n2o, color = n2o.src.snk),
+          show.legend = "point") +
+  ggtitle("Dissolved N2O (nM)") +
+  scale_color_manual(values = c("white", "black"), name = "source/sink") +
+  scale_size(name = "dissolved N2O (nM)",
+             range = c(0.1, 10), # custom size range
+             breaks = c(1, 10, 25, 50, 100)) # custom breaks
+```
+
+We can also look at population estimates of N2O saturation ratio by ecoregion.
+Wow, the mean N2O saturation is less than one for all ecoregions and the nation!
+95% CI overlap one in only 3 levels of the stratification.
+This is pretty big news.
+```{r}
+
+# extract saturation indicator, order by n2o
+pop.eco.sat <- pop.eco %>%
+  filter(grepl("sat", indicator, ignore.case = TRUE)) %>%
+  mutate(subpopulation = factor(subpopulation, 
+                             levels = filter(pop.eco, indicator == "N2O_SAT_RATIO", subpopulation != "National") %>%
+                               arrange(estimate) %>%
+                               select(subpopulation) %>%
+                               pull() %>%
+                               c("National", .)))
+
+# look at dissolved concentration indicator
+ggplot(pop.eco.sat, aes(estimate, subpopulation)) +
+  geom_point() +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct)) +
+  facet_wrap(~indicator, scales = "free_x") +
+  theme(axis.title = element_blank())
+
 ```
@@
 ````
+
+## Lagged scatterplots
+Mike, can you integrate scatterplot code here?
+
+## Semivariograms
+Variograms display the variability among pairs of points at various separation distances.
+I was unable to fit a model to the dissolved gas concentration variograms using the default fitting parameters in gstat.
+The variograms suggest a lack of spatial structure, but I suspect the model parameterization can be improved.
+
+```{r, warning=FALSE, message=FALSE}
+# Variogram function requires sp object
+dg.sp <- as(dg.sf, Class = "Spatial")
+
+# CO2 variogram
+co2.v <- variogram(dissolved.co2~1, dg.sp)
+co2.v.fit <- fit.variogram(co2.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
+# co2.v.fit # chose sperical
+co2.v.fit <- fit.variogram(co2.v, vgm("Sph")) # refit spherical, several warnings (?)
+
+# CH4 variogram
+ch4.v <- variogram(dissolved.ch4~1, dg.sp)
+ch4.v.fit <- fit.variogram(ch4.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
+# ch4.v.fit # chose exp
+ch4.v.fit <- fit.variogram(ch4.v, vgm("Exp")) # refit exponential, no convergence
+
+# N2O variogram
+n2o.v <- variogram(dissolved.n2o~1, dg.sp)
+n2o.v.fit <- fit.variogram(n2o.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
+# n2o.v.fit # chose exp
+n2o.v.fit <- fit.variogram(n2o.v, vgm("Exp")) # refit exponential, no convergence
+
+# Plot variograms and models (if fit)
+plot(co2.v, model = co2.v.fit, main = "CO2 semivariogram")
+plot(ch4.v, main = "CH4 semivariogram")
+plot(n2o.v, main = "N2O semivariogram")
+
+```

### 2019-12-10 scripts/readDissolvedGas.R
Likely garbage

@@ -13,6 +13,7 @@ dg <- read.table(file = paste0(localPath,
                                "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
                  header = TRUE, sep = "\t", as.is = TRUE)
 
+# need to drop WSA9 = AK, WSA9_NAME = Alaska
 
 # sp
 # Define coordinates
@@ -20,12 +21,19 @@ coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
 
 # to spatial points df
 dg.sp <- SpatialPointsDataFrame(coords, dg)
-
+class(dg.sp)
+summary(dg.sp)
+names(dg.sp@data)
 # projection
 dg.sp@proj4string <- CRS('+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs')
+dg.sp@bbox
 
-# remove air data
-dg.sp <- filter(dg.sp, sample.source == "DG")
-
+dg.sf <- st_as_sf(dg.sp)
+class(dg.sf)
+# remove air data, 
+# get error message as trying to sf-like function, filter, on an sp object
+# convert dg.sp to sf
+dg.sf <- filter(dg.sf, sample.source == "DG")
+dim(dg.sf)
 # sf
 adg.sf <- st_as_sf(dg.sp, coords = c("map.lon.dd", "map.lat.dd"), crs = 4269)

### 2019-12-10 scripts/dgIndicatorAnalysis.Rmd
Simple text change.
File management to sync with remote.

--- 2019-11-15 scripts/dgIndicatorAnalysis.Rmd
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
+                               "inputData/nla17gasDataAggregated_2019-11-12.txt"), 
                  header = TRUE, sep = "\t", as.is = TRUE) %>%
@@
          dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
-         dissolved.n20 = dissolved.n2o * 1000000000, # mol/L -> nmol/L
+         dissolved.n2o = dissolved.n2o * 1000000000, # mol/L -> nmol/L
          # add source/sink column
@@
 source or sink).
-The code immediately below converts the dataframe to a spatial object and reads in the ecoregion ploygons.
+The code immediately below converts the dataframe to a spatial object and reads in the ecoregion polygons.
 The first bubble plot below is of dissolved CO2 which seems to show a pattern of higher concentrations along the eastern seaboard.
@@
 ```
+
+# Exploratory analysis of N2O data
+The mean N2O saturation ratio for the country is `r pop.eco.sat %>% filter(type == "National", subpopulation == "National", indicator == "N2O_SAT_RATIO") %>% select(estimate) %>% pull() %>% round(2)` (95% CI = `r pop.eco.sat %>% filter(type == "National", subpopulation == "National", indicator == "N2O_SAT_RATIO") %>% select(lcb95pct) %>% pull() %>% round(2)` - `r pop.eco.sat %>% filter(type == "National", subpopulation == "National", indicator == "N2O_SAT_RATIO") %>% select(ucb95pct) %>% pull() %>% round(2)`), suggesting that US waterbodies are a new N2O sink during the NLA index period.
+This finding is consistent with a recent report that 67% of 101 agricultural ponds in Canada were undersaturated in N2O during the summer (Webb et al.
+2019, PNAS).
+
+There are several sources of N2O in surface waters.
+The overlying atmosphere contains trace quantities of N2O (~300 ppb) that dissolve into surface waters as a function of barometric pressure and water temperature.
+N2O is also produced and consumed in aquatic ecosystems via denitrification.
+Denitrification is the reduction of oxidized forms of N (i.e.
+NO3) to more reduced forms (i.e.
+N2).
+It is a form of anaerobic respiration where oxidized forms of N serve as the electron acceptor in the absence of more energetically favorable alternatives, such as oxygen.
+N2O concentrations/fluxes are therefore somewhat dependent upon the availability of oxidized N and low dissolved oxygen.
+N2O can also be produced via nitrification.
+Nitrification is the aerobic oxidation of NH4 to more oxidized forms.
+N2O can be produced as a byproduct during the process and rates are somewhat dependent upon the availability of ammonium.
+See Beaulieu et al.
+2011, 2014, and 2015 for more information.
+
+The effect of denitrification and nitrification on dissolved N2O concentrations near the air-water interface is somewhat mediated by hydrology.
+Denitrification is an anaerobic process that occurs predominantly in the sediment or anoxic hypolimnion.
+Therefore, N2O must advect or diffuse from the sediment/hyoplimion to the water surface before it can affect dissolved N2O concentration near the air/water interface (where NLA samples were collected).
+Previous invesitgations have interpretted correlations between N2O and lake size and degree of thermal stratification as indicative of hydrologic control (DelSontro et al.
+2019, Webb et al.
+2019).
+
+DelSontrol et al.
+(2019) reported that N2O emission rates were positively related to chl a.
+The authors (inlcuding me, I'm one of the 'authors'!) speculate that chla is a surrogate for labile carbon.
+Denitrification requires carbon, so perhaps high rates of denitrification-N2O production occur in the presence of abundant carbon availability.
+
+We do not yet have depth profiles of DO or temperature.
+When these data become available, we can include bottom water DO and strength of thermal stratification in the model.
+For now, we can explore correlations with chemistry variables.
+A correlation matrix of dissolved N2O concentration and potential controlling variables doesn't show any striking relationships, though the strongest relationship is with nitrate.
+
+```{r}
+library(corrr)
+dg %>%
+  select(dissolved.n2o, MeanDUsed, ph.result, nitrate.n.result, ntl.lab.result, chla.result.vol, ptl.lab.result, surftemp) %>%
+  correlate()
+```
+
+Categorization and regression analysis (CART), which can accomodate non-linear and interactive relationships, also identified nitrate as the most important predictor variable.
+The CART analysis did not identify splits associated with ecoregion.
+
+```{r}
+library(tree)
+model <- tree(dissolved.n2o ~ MeanDUsed + nitrate.n.result + ntl.lab.result + chla.result.vol + ptl.lab.result + surftemp + WSA9, data = dg)
+
+plot(model)
+text(model)
+
+```

### 2019-12-10 scripts/readDissolvedGas.R

@@ -1,16 +1,12 @@
 # READ IN DISSOLVED GAS DATA
 # This file contains the aggregated (by duplicates) dissolved gas data
 # from NLA17.  See data dictionary in documents library for details.
-# 
-# dg <- read.table(file = paste0("C:/Users/JBEAULIE/Environmental Protection Agency (EPA)/",
-#                                "ORD NLA17 Dissolved Gas - Documents/",
-#                                "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
-#                  header = TRUE, sep = "\t", as.is = TRUE)
+
 
 dg <- read.table(file = paste0(localPath, 
                                "/Environmental Protection Agency (EPA)/",
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2019-10-25.txt"), 
+                               "inputData/nla17gasDataAggregated_2019-11-12.txt"), 
                  header = TRUE, sep = "\t", as.is = TRUE)
 
 

### 2019-12-10 scripts/dgIndicatorAnalysis.Rmd

--- 2019-12-10 scripts/dgIndicatorAnalysis.Rmd
@@
 
-# Data
-
+## Getting data and filtering
 The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.
@@
 When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
-## Getting data and filtering
-
-The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.
-The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.
-A copy of this data file is stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
-When synced to a local computer, the documents library can be read directly with R.
 
@@
 ```{r}
-# are these the main response variables we'll be looking at?
+# These are the main response variables.
 dg %>% select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
@@
 The first bubble plot below is of dissolved CO2 which seems to show a pattern of higher concentrations along the eastern seaboard.
-
 ```{r}
@@
 ```
-
-# Bubble plots
 
@@
 
-````
-
-## Lagged scatterplots
-Mike, can you integrate scatterplot code here?
+```
 

### 2019-12-10 scripts/identifyLocalPath.R
Delete scripts that have been integrated into .rmd.

@@ -1,1 +1,0 @@
-localPath <- Sys.getenv("USERPROFILE") 

### 2019-12-10 scripts/masterLibrary.R

@@ -1,13 +1,0 @@
-library(rgdal) # spatial data
-library(sf) # spatial data
-library(gstat) # lagged scatterplot
-library(spdplyr) # apply dplyr verbs to sp objects
-library(rmarkdown)
-
-
-# ENFORCE EPA FORMAT ON NLA NAMES
-toEPA <- function(X1){
-  names(X1) = tolower(names(X1))
-  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
-  X1
-}

### 2019-12-10 scripts/masterScript.R

@@ -1,8 +1,0 @@
-# THESE CAN BE SKIPPED IF WORKING IN dgIndicatorAnalysis.Rmd
-# source("scripts/identifyLocalPath.R")  # set local path for each user
-# source("scripts/masterLibrary.R") # libraries
-# source("scripts/readDissolvedGas.R") # read data
-
-# # probably can't source this because it depends on larger environment.  Not sure.
-# # This will likely be replaced with dgIndicatorAnalysis.Rmd
-# source("scripts/laggedScatterplot.Rmd") 

### 2019-12-10 scripts/readDissolvedGas.R

@@ -1,35 +1,0 @@
-# READ IN DISSOLVED GAS DATA
-# This file contains the aggregated (by duplicates) dissolved gas data
-# from NLA17.  See data dictionary in documents library for details.
-
-
-dg <- read.table(file = paste0(localPath, 
-                               "/Environmental Protection Agency (EPA)/",
-                               "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2019-11-12.txt"), 
-                 header = TRUE, sep = "\t", as.is = TRUE)
-
-# need to drop WSA9 = AK, WSA9_NAME = Alaska
-
-# sp
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-# to spatial points df
-dg.sp <- SpatialPointsDataFrame(coords, dg)
-class(dg.sp)
-summary(dg.sp)
-names(dg.sp@data)
-# projection
-dg.sp@proj4string <- CRS('+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs')
-dg.sp@bbox
-
-dg.sf <- st_as_sf(dg.sp)
-class(dg.sf)
-# remove air data, 
-# get error message as trying to sf-like function, filter, on an sp object
-# convert dg.sp to sf
-dg.sf <- filter(dg.sf, sample.source == "DG")
-dim(dg.sf)
-# sf
-adg.sf <- st_as_sf(dg.sp, coords = c("map.lon.dd", "map.lat.dd"), crs = 4269)

### 2019-12-10 scripts/dgIndicatorAnalysis.Rmd
Added exploratory N2O section to dgIndicatorAnalysis.rmd

@@
          ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
-         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
+         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"),
+         # add state
+         state = substr(site.id, 7, 8))
 
@@
 Could only get parts of Anna Springsteen's code to work.
-```{r, distances}
+```{r, eval=FALSE, echo=FALSE, distances}
 names(dg.sf)
 head(dg.sf$WSA9_NAME)
-distinct(dg.sf, WSA9_NAME) # is distinct on sf object only returning first 10 distinct coordinates, and not WSA9_NAME
-distinct.Spatial(dg.sf, WSA9_NAME) # this doesn't work either.
+# distinct returns all unique geometry + attribute combinations.  convert to df if not interested in geometry
+dg.sf %>% as.data.frame() %>% distinct(WSA9_NAME) 
+# distinct.Spatial(dg.sf, WSA9_NAME) # this doesn't work either.
 distinct(dg, WSA9_NAME)
@@
 Note that can use simple feature object, such as cpl, for histogram, but spatial object needed for lagged scatterplots.  I took log10 transform to  get data somewhat unimodal and symmetric.
-```{r, laggedscatterplot}
+```{r, eval=FALSE, echo=FALSE, laggedscatterplot}
 # cpl
@@
 When these data become available, we can include bottom water DO and strength of thermal stratification in the model.
-For now, we can explore correlations with chemistry variables.
+For now, we can explore correlations with the following variables.
+
+* MeanDUsed - mean depth  
+* ph.result - pH  
+* nitrate.n.result - NO3-N (mg/L)  
+* ntl.lab.result - total nitrogen (mg/L)  
+* chla.result.vol - chlorophyll a (ug/L)  
+* ptl.lab.result - total phosphorus (mg/L)  
+* surftemp - surface water temperature (C)
+
 A correlation matrix of dissolved N2O concentration and potential controlling variables doesn't show any striking relationships, though the strongest relationship is with nitrate.
@@
 ```
+
+We may also want to include design variables in the analysis.
+The specifics of the NLA17 design have not been disseminated, but are likely very similar to the NLA12 design.
+Per the NLA12 Design Documentation (https://www.epa.gov/sites/production/files/2014-03/documents/nla2012_design_documentation_20110320.pdf):
+
+> Stratification  
+The survey design was stratified by state and by NLA12_CLS.
+NLA12_CLS has three classes: (1) NLA07RVT – defined as all NLA 2007 lakes that were target and sampled, (2) NLA12NEW – remaining lakes in NHD-Plus that are included in the sample frame, and (3) Exclude – lakes in NHD-Plus that are excluded from the sample frame (see Sample Frame section above).
+Each state design has two strata, ST_ NLA07RVT and ST_ NLA12NEW (where ST is replaced by two letter state abbreviation.
+The total number of strata is 96 (two for each state).
+
+> Unequal Probability Categories  
+The 48 state strata for lakes from the NLA 2007 that would be visited again in 2012 was an equal probability design within each stratum.
+The 48 state strata NLA12NEW was an unequal probability design within each state stratum.
+The unequal probability categories were defined based on lake area: 1 to 4 ha, 4 to 10 ha, 10 to 20 ha, 20 to 50 ha and greater than 50 ha.
+
+As you can see, the design is actually pretty complicated.  The file currently contains the following design elements:
+
+* state - two digit state abbreviation.  See line 62 above.
+* AREASQKM - waterbody area (km2)
+
+Corvallis has not yet shared the other design elements (i.e.
+NLA17_CLS) and sample weights.
+Presumably this file will be delivered with the code to execute the population estimates via spSurvey.
+Steve Paulsen will speak to Tom Kincaid about this during the week of Dec.
+16.

### 2019-12-12 scripts/dataMunge.Rmd
separating data prep from analysis

---
title: "Data Munging"
authors: "J. Beaulieu, R. Martin, and M. McManus"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
editor_options: 
  chunk_output_type: console
---
```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
# load libraries
library(tidyverse) # dplyr, ggplot

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")
```

# Purpose
The purpose of this .rmd is to prepare the data set for subsequent modeling.  The data will be written as .RData object and stored at the shared documents library associated with this project: https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas

The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-11-12.txt.  A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

Below we read in the .txt file, perform a few manipulations, then write out a .Rdata object.

```{r}
# Read data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2019-11-12.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE) %>%
  # unit conversion
  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
         dissolved.n2o = dissolved.n2o * 1000000000, # mol/L -> nmol/L
         # add source/sink column
         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
```

The survey design is stratified by State.  Within each state, an unequal probability category based on 5 lake size classes is used.  As of 12/12/2019, NLA has not provided data on waterbody size.  Size was estimated by doing a spatial overlay of the waterbody lat/lon with NHDPlusV2.  Surface area (AREASQKM) was then extracted from NHDPlusV2.  The data currently contains `r dg %>% as.data.frame() %>% filter(is.na(AREASQKM)) %>% {nrow(.)}` missing waterbody size values.  For now, we'll use these data to add the size categories to the data file.
```{r}
dg <- dg %>%
  mutate(
    # add state
    state = substr(site.id, 7, 8),
    AREA_HA = AREASQKM / 100, # 100Ha in 1 km2 
    size_cat = ifelse(AREA_HA <= 4,
                      "LE4",
                      ifelse(AREA_HA > 4 & AREA_HA <= 10,
                             "G4_LE10",
                             ifelse(AREA_HA > 10 & AREA_HA <= 20,
                                    "G10_LE20",
                                    ifelse(AREA_HA > 20 & AREA_HA <= 50,
                                           "G20_LE50",
                                           ifelse(AREA_HA > 50,
                                                  "G50",
                                                  NA))))))

```

Concentrations of CO~2~, CH~4~, and N~2~O in ambient air over the waterbodies were measured and used to calculate the equilibrium dissolved gas concentration (variables sat.co2, sat.ch4, and sat.n2o).  These data are not needed in subsequent analysis and will be removed from the dataframe.
```{r}
dg <- dg %>% filter(sample.source == "DG") # select dissolved gas 'DG' samples
```

The dataframe also contains observations from site in Alaska.  These sites were not an official part of the NLA and will not be included in this analysis.
```{r}
dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
```

Write out final .RData object.
```{r}
save(dg, file = paste0(localPath, 
                       "/Environmental Protection Agency (EPA)/",
                       "ORD NLA17 Dissolved Gas - Documents/",
                       "inputData/dg.RData"))
```

### 2019-12-12 scripts/dgIndicatorAnalysis.Rmd
separating data prep from analysis

@@
 A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
-When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
+This file was futher manipulated in "C:\Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" and written out as an .RData object which is imported below.
 
@@
 # Read data file
-dg <- read.table(file = paste0(localPath, 
+load(paste0(localPath, 
             "/Environmental Protection Agency (EPA)/",
             "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2019-11-12.txt"),
-                 header = TRUE, sep = "\t", as.is = TRUE) %>%
-  # unit conversion
-  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
-         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
-         dissolved.n2o = dissolved.n2o * 1000000000, # mol/L -> nmol/L
-         # add source/sink column
-         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
-         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
-         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"),
-         # add state
-         state = substr(site.id, 7, 8))
+            "inputData/dg.RData"))
 
-```
-Concentrations of CO~2~, CH~4~, and N~2~O in ambient air over the waterbodies were measured and used in the dissolved gas concentrations.
-These data are not needed in subsequent analysis and will be removed from the dataframe.
-```{r}
-dg <- dg %>% filter(sample.source == "DG") # select dissolved gas 'DG' samples
-```
-The dataframe also contains observations from site in Alaska.  These sites were not an official part of the NLA and will not be included in this analysis.
-```{r}
-dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
 ```

### 2019-12-13 scripts/dataMunge.Rmd
(1) Fixed error in AREA_HA calculation. (2) Used more efficient dplyr cut function for size_cat factor.

@@
     state = substr(site.id, 7, 8),
-    AREA_HA = AREASQKM / 100, # 100Ha in 1 km2 
-    size_cat = ifelse(AREA_HA <= 4,
-                      "LE4",
-                      ifelse(AREA_HA > 4 & AREA_HA <= 10,
-                             "G4_LE10",
-                             ifelse(AREA_HA > 10 & AREA_HA <= 20,
-                                    "G10_LE20",
-                                    ifelse(AREA_HA > 20 & AREA_HA <= 50,
-                                           "G20_LE50",
-                                           ifelse(AREA_HA > 50,
-                                                  "G50",
-                                                  NA))))))
+    AREA_HA = AREASQKM * 100, # 100Ha in 1 km2 
+    size_cat = factor( cut( AREASQKM,
+                            breaks = c( -Inf, 0.04, 0.1, 0.2, 0.5, Inf ),
+                            labels = c( "min_4", "4_10", "10_20", "20_50", "50_max") ) )
+    )
 
-```
-
-Concentrations of CO~2~, CH~4~, and N~2~O in ambient air over the waterbodies were measured and used to calculate the equilibrium dissolved gas concentration (variables sat.co2, sat.ch4, and sat.n2o).
-These data are not needed in subsequent analysis and will be removed from the dataframe.
-```{r}
-dg <- dg %>% filter(sample.source == "DG") # select dissolved gas 'DG' samples
 ```

### 2019-12-13 scripts/dataMunge.Rmd
added lake area and survey weights data from Karen Blocksom

@@
 library(tidyverse) # dplyr, ggplot
+library(forcats) # fct_explicit_na()
 
@@
 After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-11-12.txt.
-A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
+Data on waterbody surface area and design weights were provided by Karen Blocksom on 12/13/2019.
+These data files are stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
 When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
 
-Below we read in the .txt file, perform a few manipulations, then write out a .Rdata object.
+# Data
+## Read and merge
+Below we read in the data files.
 
 ```{r}
-# Read data file
+# Read dissolved gas data file
 dg <- read.table(file = paste0(localPath, 
@@
                                "inputData/nla17gasDataAggregated_2019-11-12.txt"),
-                 header = TRUE, sep = "\t", as.is = TRUE) %>%
+                 header = TRUE, sep = "\t", as.is = TRUE)
+
+wt.area <- read.csv(file = paste0(localPath, 
+                               "/Environmental Protection Agency (EPA)/",
+                               "ORD NLA17 Dissolved Gas - Documents/",
+                               "inputData/NLA17_SiteWeights_Areas.csv"),
+                    as.is = TRUE)
+```
+
+Unique waterbodies are identified by 'site id'.  Some waterbodies were visited twice, therefore each unique sampling trip is identified by a combination of 'site id' and visit number.
+
+``` {r}
+dg %>% distinct(site.id, visit.no)
+```
+
+Survey weights and lake area are invariant across site visits, therefore the weights + lake area file does not contain a visit number field:
+``` {r}
+# no visit number in this file
+names(wt.area)
+```
+
+The two data sets will be merged on the 'site id' fields.
+The dissolved gas file contains data from sites in Alaska that are not present in the weights + lake area file.
+Alaska sites are not an offical component of the NLA and will removed from this analysis.
+``` {r}
+# Will merge datasets on SITE_ID field.  Are all site.id values in dg present in wt.area?
+dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id) # no.
+
+# Remove AK sites
+dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
+```
+
+This leaves one site in WI that is in the dissolved gas file, but not in the weights + lake area file.
+I have the Chain of Custody form for this site and will forward to Karen Blocksom for clarification.
+``` {r}
+# Will merge datasets on SITE_ID field.  Are all site.id values in dg present in wt.area?
+dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id) # Just a single WI site
+```
+
+There are 23 observations in the weights + lake area file that are absent from the dissolved gas file.
+One observation is from Arizona.
+This corresponds to notes in the chain of custody form indicating the vials arrived unlabeled and were not run.
+The remaining sites are from OR.
+I can't find any record of these samples in chain of custody form.
+I'll check with Karen Blocksom regarding these samples.
+
+``` {r}
+# Any site id values in wt.area, but not present in dg?
+wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # hmm, 23 sites w/out dissolved gas data?  One from AZ, the rest from OR.
+```
+
+Finally, lets merge these two files in site id.  Use left_join to retain all dissolved gas data, including the WI site that doesn't have a matching record in the weights + lake area file.
+``` {r}
+dim(dg) # 2369, 364
+
+dg <- left_join(dg, wt.area, by = c("site.id" = "SITE_ID"))
+
+dim(dg) # 2369, 369;  good retained all DG observations, added a few columns.
+```
+
+## Manipulate
+Impliment unit conversion and create source/sink column:
+``` {r}
+dg <- dg %>%
   # unit conversion
@@
 Within each state, an unequal probability category based on 5 lake size classes is used.
-As of 12/12/2019, NLA has not provided data on waterbody size.
-Size was estimated by doing a spatial overlay of the waterbody lat/lon with NHDPlusV2.
-Surface area (AREASQKM) was then extracted from NHDPlusV2.
-The data currently contains `r dg %>% as.data.frame() %>% filter(is.na(AREASQKM)) %>% {nrow(.)}` missing waterbody size values.
-For now, we'll use these data to add the size categories to the data file.
+Use lake area data provided by Karen to define lake area categories.
 ```{r}
@@
     state = substr(site.id, 7, 8),
-    AREA_HA = AREASQKM * 100, # 100Ha in 1 km2 
-    size_cat = factor( cut( AREASQKM,
-                            breaks = c( -Inf, 0.04, 0.1, 0.2, 0.5, Inf ),
+    size_cat = factor( cut( AREA_HA,
+                            breaks = c( -Inf, 4, 10, 20, 50, Inf ),
                             labels = c( "min_4", "4_10", "10_20", "20_50", "50_max") ) )
@@
 ```
+NLA17_WI-10018 does not have a AREA_HA value, therefore does not have size_cat value.
 
-The dataframe also contains observations from site in Alaska.  These sites were not an official part of the NLA and will not be included in this analysis.
-```{r}
-dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
+``` {r}
+dg %>% filter(is.na(size_cat)) %>% distinct(site.id)
+```
+We do have an area estimate of 2.6 Ha derived from NHDPlusV2, however.  For now, lets use the NHD value to assign a size_cat value.
+
+``` {r}
+dg %>% filter(site.id == "NLA17_WI-10018") %>% distinct(AREASQKM)
+
+dg <- dg %>%
+  mutate(size_cat = fct_explicit_na(size_cat, "min_4")) # < 4HA, therefore belongs to "min_4" category.
+
+table(dg$size_cat, useNA = "ifany")
 ```
@@
                        "ORD NLA17 Dissolved Gas - Documents/",
-                       "inputData/dg.RData"))
+                       "inputData/dg.", Sys.Date(), ".RData"))
 ```

### 2019-12-19 scripts/dataMunge.Rmd
Added buoyancy frequency to data file

@@
 # load libraries
+library(plyr) # for ddply, load before tidyverse
 library(tidyverse) # dplyr, ggplot
 library(forcats) # fct_explicit_na()
+library(rLakeAnalyzer) # for buoyancy frequency
 
@@
 # Data
-## Read and merge
+## Read data
 Below we read in the data files.
@@
 
+# Read site weights and waterbody area data
 wt.area <- read.csv(file = paste0(localPath, 
@@
                     as.is = TRUE)
+
+# Read in temp profile data
+temp <- read.csv(file = paste0(localPath, 
+                               "/Environmental Protection Agency (EPA)/",
+                               "ORD NLA17 Dissolved Gas - Documents/",
+                               "inputData/NLA17_Profile_Data.csv"),
+                    as.is = TRUE)
 ```
 
+## Merge dissolved gas with survey weights + area
 Unique waterbodies are identified by 'site id'.  Some waterbodies were visited twice, therefore each unique sampling trip is identified by a combination of 'site id' and visit number.
@@
 This leaves one site in WI that is in the dissolved gas file, but not in the weights + lake area file.
-I have the Chain of Custody form for this site and will forward to Karen Blocksom for clarification.
+The Chain of Custody form verifies receipt of this sample.
+Per Karen Blocksom, this should be SITE)ID NLA17_WI-10008, visit 1.
 ``` {r}
 # Will merge datasets on SITE_ID field.  Are all site.id values in dg present in wt.area?
-dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id) # Just a single WI site
+dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id, visit.no) # Just a single WI site
+
+# Correct site code
+dg <- dg %>%
+  mutate(site.id = replace(site.id, site.id == "NLA17_WI-10018" & visit.no == 1, "NLA17_WI-10008"))
 ```
@@
 The remaining sites are from OR.
-I can't find any record of these samples in chain of custody form.
-I'll check with Karen Blocksom regarding these samples.
+Karen Blocksom confirmed that these sites were part of the OR state survey, but collected 'most' of the NLA measurements.
+They were included in the final NLA dataset for 'the purpose of adjusting weights'.
+They did not collect gas samples and can be disregarded for this analysis.
 
@@
 # Any site id values in wt.area, but not present in dg?
-wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # hmm, 23 sites w/out dissolved gas data?  One from AZ, the rest from OR.
+wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # h23 sites w/out dissolved gas data
 ```
 
-Finally, lets merge these two files in site id.  Use left_join to retain all dissolved gas data, including the WI site that doesn't have a matching record in the weights + lake area file.
+Finally, lets merge these two files on site id.  Use left_join to retain all dissolved gas data, but disregard OR and AZ sites where we don't have gas data.
 ``` {r}
@@
 ```
-NLA17_WI-10018 does not have a AREA_HA value, therefore does not have size_cat value.
+All observations have a value for the size_cat design variable.
 
 ``` {r}
-dg %>% filter(is.na(size_cat)) %>% distinct(site.id)
+table(dg$size_cat, useNA = "ifany")
 ```
-We do have an area estimate of 2.6 Ha derived from NHDPlusV2, however.  For now, lets use the NHD value to assign a size_cat value.
+
+## Calculate stratification indices from temperature profile data
+Water temperature profiles were measured at the index site in each waterbody.
+A variety of metrics describing the degree of thermal stratification can be extracted from these data.
+Webb et al.
+used the Brunt–Väisälä buoyancy frequency as an indicator of thermal stratification strength.
+We will do the same here.
+
+Three unique site visits contain only NA for temperature.
+Two of the incidents include notes that sonde was forgotten or malfunctioning.
+Two of the three sites were revisited later in the study and temperature data were collected.
+Lets assign data from the second site visit to the first visit when no temperature data were collected.
+This leaves only NLA17_CA-10082, visit 1, without temperature data.
+
+```{r}
+# Sites that only contain NAs.
+missingTemp <- temp %>% group_by(SITE_ID, VISIT_NO) %>% 
+  summarise(all.na = all(is.na(TEMPERATURE))) %>% 
+  filter(all.na == TRUE) %>%
+  select(SITE_ID, VISIT_NO)
+
+# View sites with missing temp.
+missingTemp
+
+# Two of the sites have data from second site visit
+temp %>% filter(SITE_ID %in% missingTemp$SITE_ID) %>%
+  select(SITE_ID, VISIT_NO, TEMPERATURE)
+
+# Extract data from second visit to assign to first visit
+replacementData <- temp %>% filter(SITE_ID %in% missingTemp$SITE_ID, VISIT_NO == 2) %>%
+  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>%
+  mutate(VISIT_NO = 1) # change visit no to one for join
+
+# Join the data
+temp1 <- full_join(temp, replacementData)
+
+# Inspect results of join.  Looks good.  Only one site missing temperature data.
+temp1 %>% filter(SITE_ID %in% missingTemp$SITE_ID) %>%
+  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>% 
+  arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE)
+
+```
+
+The field crews measured temperature starting near the water surface, then at increasing depths.
+A second measurement was made at the water surface before the sonde was removed from the water, therefore the file contains duplicate measurements at the shallowest depths.
+Duplicate measurements corrupt the buoyancy frequency function and will therefore be aggregated.
+``` {r}
+# Aggregate across SITE_ID, VISIT_NO, and DEPTH.
+I don't like the dplyr solution for this (group_map) and I'm not familiar with the purr::map_ functions.
+I'm going old-school and busting out the trusty plyr::ddply.
+temp2 <- temp1 %>% select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>%
+  # exclude NA, NaN, Inf...
+  filter(is.finite(TEMPERATURE)) %>% # This completely removes the CA site with only NA for temp data.
+  plyr::ddply(., .(SITE_ID, VISIT_NO, DEPTH), summarize, 
+              TEMPERATURE.M = mean(TEMPERATURE, na.rm = TRUE))
+dim(temp1) # 11003
+dim(temp2) # 10071, nearly 1000 observations aggregated or stripped out due to NA
+```
+
+Finally, calculate the buoyancy frequency depth profile for each waterbody and print to buoyancyFrequency.pdf for inspection.
 
 ``` {r}
-dg %>% filter(site.id == "NLA17_WI-10018") %>% distinct(AREASQKM)
+# The buoyancy frequency is calculated at multiple depths, but at fewer depths than the where 
+# temperature was measured.  The output of the buoyancy frequency function is shorter 
+# than the input, which is incompatible with dplyr mutate and summarize.  The dplyr alternative
+# is group_map which produces a list, but the grouping variables are stripped from the list
+# elements.  I really don't like this.  I think this can be addressed by combining group_map
+# with purr::map_ functions, but I'm not familiar with purr.  Rather, I'll use the old school
+# split -lapply - do(rbind) approach.
+bf <- temp2 %>%
+  group_split(SITE_ID, VISIT_NO) %>% # split into list of df
+  lapply(function(x) {
+    data.frame(BF = buoyancy.freq(wtr = x$TEMPERATURE.M, depths = x$DEPTH), # calculate buoyance frequency
+               SITE_ID = unique(x$SITE_ID), # add site ID
+               VISIT_NO = unique(x$VISIT_NO), # add visit no
+               stringsAsFactors = FALSE)}) %>% 
+  lapply(function(x) x %>% mutate(DEPTHS.BF = attr(BF, "depths"))) %>% # depths stored as attribute, assign to column
+  do.call("rbind", .) # collapse to df
 
-dg <- dg %>%
-  mutate(size_cat = fct_explicit_na(size_cat, "min_4")) # < 4HA, therefore belongs to "min_4" category.
+# there are three non-finite bf values, lets have a look
+bf.na <- bf %>% filter(!is.finite(BF)) %>%
+  select(SITE_ID, VISIT_NO)
 
-table(dg$size_cat, useNA = "ifany")
+bf.na
+
+# ONe of the three values comes from a site with other non-zero measurements.  Probably OK.
+# two sites have only a single NA for BF becasue they only have a single temp measurement, 
+# presumably these are shallow sites, therefore unlikely to be stratified.  Assign a bf
+# very close to 0.
+bf %>% filter(SITE_ID %in% bf.na$SITE_ID) 
+temp2 %>% filter(SITE_ID %in% bf.na$SITE_ID)
+
+# plot buoyancy frequency depth profile
+# (http://stackoverflow.com/questions/8018961/connect-points-in-qplot-by-adjacent-y-value-not-x-value)
+# https://stackoverflow.com/questions/29034863/apply-a-ggplot-function-per-group-with-dplyr-and-set-title-per-group
+bf.plot <- bf %>%
+  group_by(SITE_ID, VISIT_NO) %>%
+  do(plots = ggplot(data = .) +
+       aes(DEPTHS.BF, BF) +
+    geom_point() +
+    geom_line() +
+    scale_x_reverse() +
+    coord_flip() +
+    ggtitle(paste(unique(.$SITE_ID), "VISIT_NO = ", unique(.$VISIT_NO))))
+
+# pdf("output/figures/buoyancyFrequency.pdf", onefile = TRUE)
+# bf.plot$plots
+# dev.off()
+```
+
+Webb et al. used the maximum buoyancy frequency value as an index of thermal stratification.  The maximum value should occur near the thermocline.  Here is an example of a classical pattern:
+```{r}
+bf.plot$plots[92]
+```
+
+Extract the max buoyancy frequency value for each waterbody. 
+``` {r}
+# Finally, extract max buoyancy for each SITE_ID x VISIT_NO
+bf.max <- bf %>% group_by(SITE_ID, VISIT_NO) %>%
+  summarise(MAX.BF = max(BF, na.rm = TRUE)) %>%
+  ungroup()
+```
+
+`r bf.max %>% summarize(n.na = sum(is.na(bf.max$MAX.BF))) %>% pull()` of the profiles do not have a maximum buoyancy frequency value.  Very good.
+
+Finally, merge buoyancy frequency data with dg.
+The buoyancy frequency data contains information from the extra Oregon sites (see above) and the AZ site where the gas samples were unlabeled and not analyzed.
+This is as expected.
+The buoyancy frequency data do not contain observations for the California site (CA-10082) where the sonde was reported missing/malfunctioning.
+```{r}
+# are all SITE_ID values in buoyancy frequency also present in dg?
+bf.max %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007
+print(n= Inf)
+
+# are all site.id values in dg also present in bf.max?
+# Just CA site, as expected.
+dg %>% filter(!(site.id %in% bf.max$SITE_ID)) %>% # no
+  select(site.id, visit.no)
+
+# dim(bf.max) # 1209, 3
+# dim(dg) # 2369, 374
+
+# Left join: retain all dg values, exclude non-matching bf values (i.e. OR and AZ)
+dg <- left_join(dg, bf.max, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
+# dim(dg) # 2369, 375, good!
+```
+
+The California site with the broken sonde has no bf data.  Nothing we can do about it.
+```{r}
+dg %>% filter(is.na(MAX.BF)) %>%
+  select(site.id, visit.no)
 ```

### 2019-12-19 scripts/dataMunge.Rmd
Estimated missing max buouancy frequency data

@@
 bf.na <- bf %>% filter(!is.finite(BF)) %>%
-  select(SITE_ID, VISIT_NO)
+  select(SITE_ID, VISIT_NO) %>%
+  mutate(SITE.VISIT = paste0(SITE_ID, VISIT_NO))
 
@@
 
-# ONe of the three values comes from a site with other non-zero measurements.  Probably OK.
-# two sites have only a single NA for BF becasue they only have a single temp measurement, 
-# presumably these are shallow sites, therefore unlikely to be stratified.  Assign a bf
-# very close to 0.
-bf %>% filter(SITE_ID %in% bf.na$SITE_ID) 
-temp2 %>% filter(SITE_ID %in% bf.na$SITE_ID)
+# These site have a single temp measurement made at shallow depths (<1m).  Presumably these are 
+# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0.
+bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
+temp2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)
 
@@
 
-Extract the max buoyancy frequency value for each waterbody. 
+Extract the max buoyancy frequency value for each waterbody after converting negative BF values to 0, per Webb.
 ``` {r}
 # Finally, extract max buoyancy for each SITE_ID x VISIT_NO
-bf.max <- bf %>% group_by(SITE_ID, VISIT_NO) %>%
+bf.max <- bf %>% 
+  mutate(BF = ifelse(BF < 0,
+                     0,
+                     BF)) %>%
+  group_by(SITE_ID, VISIT_NO) %>%
   summarise(MAX.BF = max(BF, na.rm = TRUE)) %>%
@@
 
-`r bf.max %>% summarize(n.na = sum(is.na(bf.max$MAX.BF))) %>% pull()` of the profiles do not have a maximum buoyancy frequency value.  Very good.
+`r bf.max %>% summarize(n.na = sum(!is.finite(bf.max$MAX.BF))) %>% pull()` of the site have a non-finite value for maximum buoyancy frequency value.
+These are the three shallow sites with only a single temperature measurement.
+Assign these the minimum bf.max value.
+``` {r}
+bf.max %>% filter(!is.finite(MAX.BF))
+
+min.max.bf <- bf.max %>% 
+  filter(is.finite(MAX.BF)) %>%
+  summarize(min.max.bf = min(MAX.BF))
+
+bf.max <- bf.max %>%
+  mutate(MAX.BF = replace(MAX.BF, 
+                          !is.finite(MAX.BF), # if not finite
+                          min.max.bf$min.max.bf)) # then min
+
+# that took care of it
+bf.max %>% filter(!is.finite(MAX.BF))
+```
 

### 2020-01-16 scripts/dgIndicatorAnalysis.Rmd
Mike created a new section and code chunk, called Geopackage.  This creates a geodatabase, nladg_v1.gpkg, that can be opened in ArcMap.

@@
 
+```
+
+## Geopackage
+Create a simple feature on entire dg data frame, then split out a simple feature, df_sf2, by filtering sample_source = DG.
+The geodatabase will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
+```{r, geopackage}
+library(janitor) # clean_names function
+names(dg)
+
+dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2, dissolved.ch4, dissolved.n2o, WSA9, WSA9_NAME)
+names(dg_1)
+dg_1 <- clean_names(dg_1) # underscores are "safer" names used in GIS
+names(dg_1)
+# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)
+
+dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(4326) # 4326 is just the EPSG identifier of WGS84.
+WGS84 comprises a standard coordinate frame for the Earth, a datum/reference ellipsoid for raw altitude data, and a gravitational equipotential surface (the geoid) that defines the nominal sea level.
+st_crs(dg_sf1)
+
+dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1185 obs.
+st_crs(dg_sf1)
+
+# Bryan suggested EPSG 4326
+# Bryan Chastain as GeoServices suggested reducing the number of fields, and I tried that but was still unable to open the geodatabase in ArcMap 
+# Based on gis stackexchange I think the "." in variable names is causing the problem
+# https://gis.stackexchange.com/questions/319611/arcgis-geopackage-arcmap-drawing-error?noredirect=1
+# Even with change to _ still get error message about invalid coordinate system when st_transform(5070) used
+
+# Check CRS
+st_crs(ecoR)
+names(ecoR)
+
+ecoR1 <- st_transform(ecoR, 4326)
+st_crs(ecoR1)
+
+# write out dg.sf and ecoR as two layers in geopackage, gpkg, geodatabase
+# 74. Creating a geopackage geodatabase, then adding more layers
+# https://github.com/r-spatial/sf/issues/392
+st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE)
+st_layers("nladg_v1.gpkg")
+
+# https://gis.stackexchange.com/questions/223240/writing-multiple-layers-to-geopackage-using-writeogr-in-r
+st_write(ecoR1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "ecoR1", driver = "GPKG", update = TRUE, quiet = FALSE)
+st_layers("nladg_v1.gpkg")
+
+names(dg_sf1)
+summary(dg_sf1$dissolved_n2o) # matches Roy's stats
 ```

### 2020-01-29 scripts/dgIndicatorAnalysis.Rmd
Mike created a new section and code chunk call k-NN=1 and Moran plots to examine spatial autocorrelation in dissolved.n2o for each of the ecoregions and the entire NLA.

@@
 library(tidyverse) # dplyr, ggplot
+library(janitor) # clean names
+library(spdep) # k-NN and Moran plot
 
@@
 
+## k-NN = 1 and Moran plots
+As with the lagged scatter plots, I had to take a simple feature object and convert it to a spatial object so I could use the nearest-neighbor and Moran scatter plot functions from the spdep package.
+A Moran plot is based on plotting the N~2~O variable, dissolved.n2o, of a focal point on the x-axis and the average dissolved.n2o of the neighboring points on the y-axis.
+Moran scatter plots were made for each of the 9 ecoregions and 1 for all of the NLA sites.
+For each plot, a neighborhood of points was defined by the maximum distance to first nearest neighbor.
+A distance-based spatial relationship was used because many of the ecoregions have several non-contiguous polygons or areas, which prevents a contiguity-based spatial relationship from being used.
+The slope of the line through the scatter plot of points measures spatial autocorrelation, with a positive slope indicating positive spatial autocorrelation.
+
+The code below is to be run separately for each ecoregion.
+The CPL, TPL, and UMW ecoregions show strong positive spatial autocorrelation.
+Negative spatial autocorrelation occurs in the NPL ecoregion because of the maximum outlier at a South Dakota site.
+The SPL and XER ecoregions show no evidence of spatial autocorrelation.
+Interestingly, it appears that the ecoregions with the highest dissolved.n2o concentrations, CPL, TPL, and UMW, drive the national pattern.
+```{r, knn_moran}
+names(dg)
+# also includ WSA9, WSA9_NAME feature
+dg_1 <- select(dg,site.id:dissolved.n2o, WSA9, WSA9_NAME)
+names(dg_1)
+dg_1 <- clean_names(dg_1)
+names(dg_1)
+distinct(dg_1, wsa9)
+# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)
+
+dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"),crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+st_crs(dg_sf1)
+glimpse(dg_sf1)
+# Get ecoregion-specific NLA sites
+## CPL
+cpl <- dg_sf1 %>% filter(wsa9 == "CPL" & visit_no == 1 & sample_source == "DG")
+
+cpl_sp <- as(cpl, "Spatial")
+class(cpl_sp)
+
+ds1coords <- coordinates(cpl_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = cpl_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=134761, row.names=cpl_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(cpl_sp)
+plot(ds1nb1, ds1coords, labels = cpl_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(cpl)
+# cpl_df <- st_drop_geometry(cpl)
+# names(cpl_df)
+summary(cpl$dissolved_n2o)
+
+moran.plot(log(cpl$dissolved_n2o), listw = dsnb1.listw, labels = cpl$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## NAP
+nap <- dg_sf1 %>% filter(wsa9 == "NAP" & visit_no == 1 & sample_source == "DG")
+
+nap_sp <- as(nap, "Spatial")
+class(nap_sp)
+
+ds1coords <- coordinates(nap_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = nap_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=112630, row.names=nap_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(nap_sp)
+plot(ds1nb1, ds1coords, labels = nap_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(nap)
+# nap_df <- st_drop_geometry(nap)
+# names(nap_df)
+summary(nap$dissolved_n2o)
+
+moran.plot(log(nap$dissolved_n2o), listw = dsnb1.listw, labels = nap$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## NPL
+npl <- dg_sf1 %>% filter(wsa9 == "NPL" & visit_no == 1 & sample_source == "DG")
+
+npl_sp <- as(npl, "Spatial")
+class(npl_sp)
+
+ds1coords <- coordinates(npl_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = npl_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=215395, row.names=npl_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(npl_sp)
+plot(ds1nb1, ds1coords, labels = npl_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(npl)
+# npl_df <- st_drop_geometry(npl)
+# names(npl_df)
+summary(npl$dissolved_n2o)
+
+moran.plot(log(npl$dissolved_n2o), listw = dsnb1.listw, labels = npl$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## SAP
+sap <- dg_sf1 %>% filter(wsa9 == "SAP" & visit_no == 1 & sample_source == "DG")
+
+sap_sp <- as(sap, "Spatial")
+class(sap_sp)
+
+ds1coords <- coordinates(sap_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = sap_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=128285, row.names=sap_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(sap_sp)
+plot(ds1nb1, ds1coords, labels = sap_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(sap)
+# sap_df <- st_drop_geometry(sap)
+# names(sap_df)
+summary(sap$dissolved_n2o)
+
+moran.plot(log(sap$dissolved_n2o), listw = dsnb1.listw, labels = sap$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## SPL
+spl <- dg_sf1 %>% filter(wsa9 == "SPL" & visit_no == 1 & sample_source == "DG")
+
+spl_sp <- as(spl, "Spatial")
+class(spl_sp)
+
+ds1coords <- coordinates(spl_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = spl_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=173395, row.names=spl_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(spl_sp)
+plot(ds1nb1, ds1coords, labels = spl_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(spl)
+# spl_df <- st_drop_geometry(spl)
+# names(spl_df)
+summary(spl$dissolved_n2o)
+
+moran.plot(log(spl$dissolved_n2o), listw = dsnb1.listw, labels = spl$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## TPL
+tpl <- dg_sf1 %>% filter(wsa9 == "TPL" & visit_no == 1 & sample_source == "DG")
+
+tpl_sp <- as(tpl, "Spatial")
+class(tpl_sp)
+nrow(zerodist(tpl_sp)) # indicates 1 pair of identical coordinates
+
+(zerodist(tpl_sp)) # this gives the row numbers of the pair 30 & 36
+View(tpl) # to view rows 30, site_id NLA17_IN-10002, & 36, NLA17_IN-10047
+
+# Warning message from code below:
+# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found
+
+ds1coords <- coordinates(tpl_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = tpl_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=223893, row.names=tpl_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(tpl_sp)
+plot(ds1nb1, ds1coords, labels = tpl_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(tpl)
+# tpl_df <- st_drop_geometry(tpl)
+# names(tpl_df)
+summary(tpl$dissolved_n2o)
+
+moran.plot(log(tpl$dissolved_n2o), listw = dsnb1.listw, labels = tpl$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## UMW
+umw <- dg_sf1 %>% filter(wsa9 == "UMW" & visit_no == 1 & sample_source == "DG")
+
+umw_sp <- as(umw, "Spatial")
+class(umw_sp)
+nrow(zerodist(umw_sp)) # indicates 1 pair of identical coordinates
+
+(zerodist(umw_sp)) # this gives the row numbers of the pair 136 & 158
+View(umw) # to view rows 136, site_id NLA17_WI-10027, & 158, NLA17_WI-10127
+
+# Warning message from code below:
+# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found
+
+ds1coords <- coordinates(umw_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = umw_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=99590, row.names=umw_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(umw_sp)
+plot(ds1nb1, ds1coords, labels = umw_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(umw)
+# umw_df <- st_drop_geometry(umw)
+# names(umw_df)
+summary(umw$dissolved_n2o)
+
+moran.plot(log(umw$dissolved_n2o), listw = dsnb1.listw, labels = umw$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## WMT
+wmt <- dg_sf1 %>% filter(wsa9 == "WMT" & visit_no == 1 & sample_source == "DG")
+
+wmt_sp <- as(wmt, "Spatial")
+class(wmt_sp)
+
+# nrow(zerodist(wmt_sp)) # indicates 1 pair of identical coordinates
+# (zerodist(wmt_sp)) # this gives the row numbers of the pair 136 & 158
+# View(wmt) # to view rows 136, site_id NLA17_WI-10027, & 158, NLA17_WI-10127
+
+# Warning message from code below:
+# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found
+
+ds1coords <- coordinates(wmt_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = wmt_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=295987, row.names=wmt_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(wmt_sp)
+plot(ds1nb1, ds1coords, labels = wmt_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(wmt)
+# wmt_df <- st_drop_geometry(wmt)
+# names(wmt_df)
+summary(wmt$dissolved_n2o)
+
+moran.plot(log(wmt$dissolved_n2o), listw = dsnb1.listw, labels = wmt$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## XER
+xer <- dg_sf1 %>% filter(wsa9 == "XER" & visit_no == 1 & sample_source == "DG")
+
+xer_sp <- as(xer, "Spatial")
+class(xer_sp)
+
+# nrow(zerodist(xer_sp)) # indicates 1 pair of identical coordinates
+# (zerodist(xer_sp)) # this gives the row numbers of the pair 136 & 158
+# View(xer) # to view rows 136, site_id NLA17_WI-10027, & 158, NLA17_WI-10127
+
+# Warning message from code below:
+# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found
+
+ds1coords <- coordinates(xer_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = xer_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=259280, row.names=xer_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(xer_sp)
+plot(ds1nb1, ds1coords, labels = xer_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(xer)
+# xer_df <- st_drop_geometry(xer)
+# names(xer_df)
+summary(xer$dissolved_n2o)
+
+moran.plot(log(xer$dissolved_n2o), listw = dsnb1.listw, labels = xer$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+
+## NLA
+nla <- dg_sf1 %>% filter(visit_no == 1 & sample_source == "DG")
+
+nla_sp <- as(nla, "Spatial")
+class(nla_sp)
+
+nrow(zerodist(nla_sp)) # indicates 2 pairs of identical coordinates
+(zerodist(nla_sp)) # this gives the row numbers of the pair (192 & 210:  NLA17_IN-10002 & NLA17_IN-10047) and the pair (1028 & 1054:  NLA17_WI-10027, & 158, NLA17_WI-10127)
+# View(nla)
+
+# Warning message from code below:
+# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found
+
+ds1coords <- coordinates(nla_sp)
+ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = nla_sp$site_id)
+# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
+dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
+summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors
+
+dsnb1 <- dnearneigh(ds1coords, d1=0, d2=259280, row.names=nla_sp$site_id, longlat=FALSE)
+summary(dsnb1)
+
+plot(nla_sp)
+plot(ds1nb1, ds1coords, labels = nla_sp$site_id, add = TRUE)
+
+dsnb1.listw <- nb2listw(dsnb1, style = "W")
+summary(dsnb1.listw)
+
+class(nla)
+# nla_df <- st_drop_geometry(nla)
+# names(nla_df)
+summary(nla$dissolved_n2o)
+
+moran.plot(log(nla$dissolved_n2o), listw = dsnb1.listw, labels = nla$site_id)
+rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
+```
+
 ## Semivariograms

### 2020-02-25 scripts/dataMunge.Rmd
Updated dataMunge and dgIndicatorAnalysis with improved data set (better BP data, a few fixes to coordinates).

@@
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-11-12.txt.
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-02-24.txt.
 Data on waterbody surface area and design weights were provided by Karen Blocksom on 12/13/2019.
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2019-11-12.txt"),
+                               "inputData/nla17gasDataAggregated_2020-02-24.txt"),
                  header = TRUE, sep = "\t", as.is = TRUE)
@@
 # Read in temp profile data
-temp <- read.csv(file = paste0(localPath, 
+temp.o2 <- read.csv(file = paste0(localPath, 
                                "/Environmental Protection Agency (EPA)/",
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/NLA17_Profile_Data.csv"),
+                               "inputData/NLA17_Profile_Data_Updated_11Feb2020.csv"),
                     as.is = TRUE)
@@
 
-This leaves one site in WI that is in the dissolved gas file, but not in the weights + lake area file.
-The Chain of Custody form verifies receipt of this sample.
-Per Karen Blocksom, this should be SITE)ID NLA17_WI-10008, visit 1.
-``` {r}
-# Will merge datasets on SITE_ID field.  Are all site.id values in dg present in wt.area?
-dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id, visit.no) # Just a single WI site
-
-# Correct site code
-dg <- dg %>%
-  mutate(site.id = replace(site.id, site.id == "NLA17_WI-10018" & visit.no == 1, "NLA17_WI-10008"))
-```
-
 There are 23 observations in the weights + lake area file that are absent from the dissolved gas file.
@@
 # Any site id values in wt.area, but not present in dg?
-wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # h23 sites w/out dissolved gas data
+wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # 23 sites w/out dissolved gas data
 ```
@@
 
-## Calculate stratification indices from temperature profile data
-Water temperature profiles were measured at the index site in each waterbody.
+## Calculate dissolved oxygen and stratification indices from profile data
+Water temperature and dissolved oxygen profiles were measured at the index site in each waterbody.
 A variety of metrics describing the degree of thermal stratification can be extracted from these data.
@@
 
-Three unique site visits contain only NA for temperature.
+Three site visits contain only NA for temperature.
+Four site visits contain only NA for do.
 Two of the incidents include notes that sonde was forgotten or malfunctioning.
@@
 This leaves only NLA17_CA-10082, visit 1, without temperature data.
+NLA17_ID-10084 has no DO data and was not resampled.
+Data sheet comment "Probe not measuring LDO for all depths".
 
@@
 # Sites that only contain NAs.
-missingTemp <- temp %>% group_by(SITE_ID, VISIT_NO) %>% 
-  summarise(all.na = all(is.na(TEMPERATURE))) %>% 
-  filter(all.na == TRUE) %>%
-  select(SITE_ID, VISIT_NO)
+missing <- temp.o2 %>% group_by(SITE_ID, VISIT_NO) %>% 
+  summarise(all.temp = all(is.na(TEMPERATURE)),
+            all.do = all(is.na(OXYGEN))) %>% 
+  filter(all.temp == TRUE | all.do == TRUE) %>%
+  select(SITE_ID, VISIT_NO, all.temp, all.do)
 
-# View sites with missing temp.
-missingTemp
+# View sites with missing temp/do
+missing
 
 # Two of the sites have data from second site visit
-temp %>% filter(SITE_ID %in% missingTemp$SITE_ID) %>%
-  select(SITE_ID, VISIT_NO, TEMPERATURE)
+temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
+  select(SITE_ID, VISIT_NO, TEMPERATURE, OXYGEN)
 
 # Extract data from second visit to assign to first visit
-replacementData <- temp %>% filter(SITE_ID %in% missingTemp$SITE_ID, VISIT_NO == 2) %>%
-  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>%
+replacementData <- temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID, VISIT_NO == 2) %>%
+  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
   mutate(VISIT_NO = 1) # change visit no to one for join
@@
 # Join the data
-temp1 <- full_join(temp, replacementData)
+temp.o21 <- full_join(temp.o2, replacementData)
 
-# Inspect results of join.  Looks good.  Only one site missing temperature data.
-temp1 %>% filter(SITE_ID %in% missingTemp$SITE_ID) %>%
-  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>% 
-  arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE)
+# Inspect results of join.  Looks good.  
+# Only one site missing temperature and DO data. (NLA17_CA-10082)
+# One site missing all DO data (NLA17_ID-10084)
+temp.o21 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
+  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>% 
+  arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN)
 
@@
 
-The field crews measured temperature starting near the water surface, then at increasing depths.
+The field crews measured temperature/DO starting near the water surface, then at increasing depths.
 A second measurement was made at the water surface before the sonde was removed from the water, therefore the file contains duplicate measurements at the shallowest depths.
@@
 I'm going old-school and busting out the trusty plyr::ddply.
-temp2 <- temp1 %>% select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>%
-  # exclude NA, NaN, Inf...
-  filter(is.finite(TEMPERATURE)) %>% # This completely removes the CA site with only NA for temp data.
+temp.o2.2 <- temp.o21 %>% select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
   plyr::ddply(., .(SITE_ID, VISIT_NO, DEPTH), summarize, 
-              TEMPERATURE.M = mean(TEMPERATURE, na.rm = TRUE))
-dim(temp1) # 11003
-dim(temp2) # 10071, nearly 1000 observations aggregated or stripped out due to NA
+              TEMPERATURE.M = mean(TEMPERATURE, na.rm = TRUE),
+              OXYGEN.M = mean(OXYGEN, na.rm = TRUE))
+
+dim(temp.o21) # 11003
+dim(temp.o2.2) # 10133, 870 observations aggregated
 ```
@@
 # split -lapply - do(rbind) approach.
-bf <- temp2 %>%
+
+# First, strip out any NA, NaN, Inf...etc
+
+bf <- temp.o2.2 %>%
+  filter(is.finite(TEMPERATURE.M)) %>% # This completely removes the CA site with only NA for temp data.
   group_split(SITE_ID, VISIT_NO) %>% # split into list of df
@@
 # These site have a single temp measurement made at shallow depths (<1m).  Presumably these are 
-# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0.
+# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0 (see below)
 bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
-temp2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)
+temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)
 
@@
 
+Webb et al.
+used the 'surface' and 'bottom' water DO percent saturation and temperature as predictors, but didn't give more details on how these indices were calculated.
+We will do something similar and define the 'shallowest' and 'deepest' measurements as 'surface' and 'bottom', respectively, but will use mg.l for dissolved oxygen.
+```{r}
+# o2 surface (s) and bottom (b)
+o2.s.b <- 
+temp.o2.2 %>% 
+  group_by(SITE_ID, VISIT_NO) %>%
+  slice(c(which.min(DEPTH), which.max(DEPTH))) %>% # min and max depth by group
+  arrange(SITE_ID, VISIT_NO, DEPTH) %>%
+    mutate(DEPTH = ifelse(row_number() == 1,
+                        "surface",
+                        "bottom")) %>%
+  pivot_wider(id_cols = c(SITE_ID, VISIT_NO), 
+              names_from = DEPTH, 
+              values_from = c(TEMPERATURE.M, OXYGEN.M)) %>%
+  rename(temp.surf = TEMPERATURE.M_surface,
+         temp.bottom = TEMPERATURE.M_bottom,
+         o2.surf = OXYGEN.M_surface,
+         o2.bottom = OXYGEN.M_bottom)
+
+o2.s.b
+
+```
+
+These calculations produced 8 site visits with at least one NA for surface/bottom temp/DO.
+
+```{r}
+o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
+```
+
+o2.surf at NLA17_FL-10029 is a case with no DO at shallowest depth, but good measurements below.  Replace with next shallowest value.
+``` {r}
+# Inspect raw data
+temp.o2.2 %>% filter(SITE_ID == "NLA17_FL-10029", VISIT_NO == 1)
+
+# update data
+o2.s.b <- o2.s.b %>%
+  mutate(o2.surf = replace(o2.surf, 
+                           SITE_ID == "NLA17_FL-10029" & VISIT_NO == 1,
+                           temp.o2 %>% filter(SITE_ID=="NLA17_FL-10029",
+                                              VISIT_NO == 1,
+                                              DEPTH == 1) %>%
+                             select(OXYGEN) %>% pull()))
+
+```
+No bottom water DO or temp at NLA17_ID-10001, visit 1 due to probe malfunction.
+
+```{r}
+temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10001", VISIT_NO == 1)
+```
+
+No DO NLA17_ID-10084, visit 1 due to probe malfunction.
+
+```{r}
+temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10084", VISIT_NO == 1)
+```
+
+No deep DO or temp at NLA17_MI-10073, visit 1 becuase no measurement taken at deepst recorded depth.  Replace with next shallowest measurement.
+
+```{r}
+# look at raw data
+temp.o2.2 %>% filter(SITE_ID == "NLA17_MI-10073", VISIT_NO == 1)
+
+# update aggregated data
+o2.s.b <- o2.s.b %>%
+  mutate(o2.bottom = replace(o2.bottom, 
+                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
+                                              VISIT_NO == 1,
+                                              DEPTH == 1.3) %>%
+                             select(OXYGEN.M) %>% pull()),
+         temp.bottom = replace(temp.bottom, 
+                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
+                                              VISIT_NO == 1,
+                                              DEPTH == 1.3) %>%
+                             select(TEMPERATURE.M) %>% pull()))
+```
+
+No surface temp at NLA17_NE-10060, visit 1 becuase no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
+
+```{r}
+# look at raw data
+temp.o2.2 %>% filter(SITE_ID == "NLA17_NE-10060", VISIT_NO == 1)
+
+o2.s.b <- o2.s.b %>%
+  mutate(temp.surf = replace(temp.surf, 
+                           SITE_ID == "NLA17_NE-10060" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NE-10060",
+                                              VISIT_NO == 1,
+                                              DEPTH == 0.1) %>%
+                             select(TEMPERATURE.M) %>% pull()))
+```
+
+No bottom temp or O2 at NLA17_NH-10063, visit 1 becuase no measurement taken at greatest depth (8m). Replace with next deeptest measurement (7.5m)
+
+```{r}
+# look at raw data
+temp.o2.2 %>% filter(SITE_ID == "NLA17_NH-10063", VISIT_NO == 1)
+
+o2.s.b <- o2.s.b %>%
+  mutate(temp.bottom = replace(temp.bottom, 
+                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
+                                              VISIT_NO == 1,
+                                              DEPTH == 7.5) %>%
+                             select(TEMPERATURE.M) %>% pull()),
+         o2.bottom = replace(o2.bottom, 
+                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
+                                              VISIT_NO == 1,
+                                              DEPTH == 7.5) %>%
+                             select(OXYGEN.M) %>% pull()))
+```
+
+No surface temp or O2 at NLA17_OR-10003, visit 1 becuase no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
+
+```{r}
+# look at raw data
+temp.o2.2 %>% filter(SITE_ID == "NLA17_OR-10003", VISIT_NO == 1)
+
+o2.s.b <- o2.s.b %>%
+  mutate(temp.surf = replace(temp.surf, 
+                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
+                                              VISIT_NO == 1,
+                                              DEPTH == 0.1) %>%
+                             select(TEMPERATURE.M) %>% pull()),
+         o2.surf = replace(o2.surf, 
+                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
+                                              VISIT_NO == 1,
+                                              DEPTH == 0.1) %>%
+                             select(OXYGEN.M) %>% pull()))
+```
+
+No bottom temp or O2 at NLA17_SD-10057, visit 1 becuase no measurement taken at deepest depth (4m). Replace with next deepest measurement (3m)
+
+```{r}
+# look at raw data
+temp.o2.2 %>% filter(SITE_ID == "NLA17_SD-10057", VISIT_NO == 1)
+
+o2.s.b <- o2.s.b %>%
+  mutate(temp.bottom = replace(temp.bottom, 
+                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
+                                              VISIT_NO == 1,
+                                              DEPTH == 3) %>%
+                             select(TEMPERATURE.M) %>% pull()),
+         o2.bottom = replace(o2.bottom, 
+                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
+                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
+                                              VISIT_NO == 1,
+                                              DEPTH == 3) %>%
+                             select(OXYGEN.M) %>% pull()))
+```
+
+Only remainining missing values are observations where field crews noted sonde malfunctions.
+
+```{r}
+o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
+```
+
+Merge bottom and surface DO and temperature with dg data.
+
+```{r}
+# are all SITE_ID values in o2 and temp data also present in dg?
+o2.s.b %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007
+print(n= Inf)
+
+# are all site.id values in dg also present in o2 and temp data?
+# Just CA site, as expected.
+dg %>% filter(!(site.id %in% o2.s.b$SITE_ID)) %>% # no
+  select(site.id, visit.no)
+
+# dim(o2.s.b) # 1209, 6
+# dim(dg) # 2369, 375
+
+# Left join: retain all dg values, exclude non-matching bf values (i.e. OR and AZ)
+dg <- left_join(dg, o2.s.b, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
+# dim(dg) # 2369, 379, good!
+```
+
 Write out final .RData object.

### 2020-02-25 scripts/dgIndicatorAnalysis.Rmd
Updated dataMunge and dgIndicatorAnalysis with improved data set (better BP data, a few fixes to coordinates).

@@
 library(gstat) # lagged scatterplot, variogram
-#library(spdplyr) # apply dplyr verbs to sp objects.  gstat functions require sp objects
 library(tidyverse) # dplyr, ggplot
@@
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-02-24.txt.
 A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
-This file was futher manipulated in "C:\Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" and written out as an .RData object which is imported below.
+This file was futher manipulated in "C://Users//JBEAULIE//OneDrive-Environmental Protection Agency (EPA)//gitRepository//DissolvedGasNla//scripts//dataMunge.Rmd" and written out as an .RData object, which is imported below.
 
@@
             "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.RData"))
+            "inputData/dg.2020-02-24.RData"))
+
+# Strip out air samples
+dg <- dg %>% filter(sample.source == "DG")
 
@@
 # These are the main response variables.
-dg %>% select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
+dg %>% 
+  select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
   {summary(.)}
@@
   filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
-         statistic == "50Pct") # extract mean estimate
+         statistic == "Mean") # extract mean estimate
 
@@
                              "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "Aggr_Ecoregions9_2015")
+                layer = "aggr_ecoregions_simple")
 
@@
 # Check CRS
-st_crs(ecoR) # no EPSG code
+st_crs(ecoR) # 3857
 ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
@@
 We can also look at population estimates of N2O saturation ratio by ecoregion.
-Wow, the mean N2O saturation is less than one for all ecoregions and the nation!
-95% CI overlap one in only 3 levels of the stratification.
-This is pretty big news.
+Wow, despite the prevalance of N2O undersaturated waterbodies, the 95% CI for the mean N2O saturation ratio encompasses 1 in all ecoregions except the Northern Plains and Northern Appalachians.
 ```{r}
@@
 The geodatabase will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
-```{r, geopackage}
+```{r, geopackage, eval=FALSE, echo=FALSE,}
 library(janitor) # clean_names function
@@
 
-ecoR1 <- st_transform(ecoR, 4326)
+ecoR1 <- st_transform(ecoR, 4326) %>% 
+  select(OBJECTID, WSA9, WSA9_NAME) # Grab variables of interest
 st_crs(ecoR1)

### 2020-02-25 scripts/dataMunge.Rmd
fixing missing nitrate values

@@
 
-## Calculate dissolved oxygen and stratification indices from profile data
+## Calculate buoyancy frequency
 Water temperature and dissolved oxygen profiles were measured at the index site in each waterbody.
@@
 
+## Calculate surface and bottom DO/temp
 Webb et al.
@@
 
+## Fix chemistry
+The file contains `r dg %>% dplyr::select(nitrate.n.result) %>% filter(is.na(nitrate.n.result)) %>% summarize(na.no3 = nrow(.)) %>% pull()` missing nitrate values.
+Inspection of raw data sheet indicates that:
+* North Dakota reported NITRATE_NITRITE_N_RESULT rather than NITRATE.N.RESULT.  Most of the ND sites were 'non detect (ND)' with a detection limit of 0.03 mg /L.
+* WI sites were ND with a detection limit of 0.019.
+
+Will replace these values with NITRATE_NITRITE_N_RESULT when available, else 1/2 detection limit.
+
+``` {r}
+dg <- dg %>% 
+  mutate(nitrate.n.result = replace(nitrate.n.result, 
+                                    grepl("WI", site.id) & is.na(nitrate.n.result), # WI na sites
+                                    0.019/2), #1/2 detection limit
+         # ND sites without nitrate.n and nitrate.nitrite
+         nitrate.n.result = replace(nitrate.n.result, 
+                                    grepl("ND", site.id) & 
+                                      is.na(nitrate.n.result) & 
+                                      is.na(nitrate.nitrite.n.result), 
+                                    0.03/2), # 1/2 detection limit
+# ND sites with a nitrate.nitrite.n.result
+         nitrate.n.result = replace(nitrate.n.result, 
+                                    grepl("ND", site.id) & 
+                                      is.na(nitrate.n.result) & 
+                                      !is.na(nitrate.nitrite.n.result), # ND na sites
+                                    0.03)) # two sites reporting 0.03 mg/L
+```
+
+The file contains `r dg %>% filter(is.na(chla.result.vol)) %>% summarize(na.chl = nrow(.)) %>% pull()` missing chlorophyll values.
+Inspection reveals that these are all from North Dakota sites.
+It looks like the lab failed to record the volume filtered, so we can't calculate ug/L chl from these sites.
+These will remain missing.
+
+## Write out data
 Write out final .RData object.

### 2020-03-04 scripts/dgIndicatorAnalysis.Rmd
I pointed to 2020-02-25.RData and I added a vertical reference line of 1 to the saturation ratio plots.

@@
             "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.2020-02-24.RData"))
+            "inputData/dg.2020-02-25.RData"))
 
@@
   facet_wrap(~indicator, scales = "free_x") +
-  theme(axis.title = element_blank())
+  theme(axis.title = element_blank()) +
+  geom_vline(xintercept = 1)
 

### 2020-06-01 scripts/dgIndicatorAnalysis.Rmd
I added two code chunks, Correlogram:  ncf and pgmirness packages and Correlograms by Ecoregions.  The first code chunk compares two packages to make correlograms, and I chose ncf.  I also did some outlier analysis.  The second chunk makes correlograms by the different ecoregions.

@@
 library(spdep) # k-NN and Moran plot
+library(ncf) # ncf::correlogram
 
@@
 
+## Correlograms: ncf and pgmirness packages
+```{r, correlogram}
+
+# Much of code below is based on example comparing correlograms among packages ncf, pgmirness, and spdep described in this chapter
+# https://link.springer.com/chapter/10.1007/978-3-030-01989-1_5
+
+coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)
+
+dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+st_crs(dg_sf1)
+names(dg_sf1)
+nap_visit1 <- dg_sf1 %>% filter(wsa9 == "NAP" & visit_no == 1)
+class(nap_visit1)
+st_crs(nap_visit1)
+
+nap_sp_visit1 <- as(nap_visit1, "Spatial")
+class(nap_sp_visit1)
+
+# trying ncf package which uses resampling to specify number of null permutations under the null to assess significance levels.
+See if significance persist after increas resamp from 100 to 1000, which is the default.
+# The x-intercept in the results is the distance at which objects are no more similar than expected by chance alone across the region.
+# Run on CPL as that is giving correlation greater than 1, which according to Fletcher and Fortin might be due to small number of pairs of points used in that first distance bin or to outliers.
+Perhaps, cluster of high Delaware sites (n visit 1 = 5) causes the correlation to be greater than 1.
+# Looks like 71 pairs of points being used.
+# With a lag, or bin distance, of 50 km only go out 10 bins, 500 km, and do a Bonferroni adjustment of 0.05/10 = 0.005
+names(dg_1)
+dg_cplvisit1 <- dg_1 %>% filter(wsa9 == "CPL" & visit_no == 1)
+# N.B. default is resamp = 1000 so that can take 10 minutes for n =1090 sites
+class(dg_cplvisit1)
+
+correlo_n2o_ncf <- ncf::correlog(x = dg_cplvisit1$map_lon_dd, y = dg_cplvisit1$map_lat_dd, z = dg_cplvisit1$log_n2o, increment = 50, resamp = 1000, latlon = TRUE)
+names(correlo_n2o_ncf)
+plot(correlo_n2o_ncf) # works when pgmiress commented out
+rm(dg_napvisit1, correlo_n2o_ncf)
+
+# B/c returns a list of list have to access with [[]]
+print(correlo_n2o_ncf[["n"]])
+print(correlo_n2o_ncf[["mean.of.class"]])
+print(correlo_n2o_ncf[["correlation"]])
+head(correlo_n2o_ncf[["mean.of.class"]])
+plot(correlo_n2o_ncf[["mean.of.class"]], correlo_n2o_ncf[["correlation"]])
+
+class(correlo_n2o_ncf)
+ncf::plot.correlog(correlo_n2o_ncf)
+
+# CPL outlier check by working through all 10 pairwise removal of cluster of 5 high Delaware sites
+out_cplvisit1 <- dg_1 %>% filter(wsa9 == "CPL" & visit_no == 1 & (site_id != "NLA17_DE-10003" & site_id != "NLA17_DE-10004"))
+
+out_cplcorrelo_n2o <- ncf::correlog(x = out_cplvisit1$map_lon_dd, y = out_cplvisit1$map_lat_dd, z = out_cplvisit1$log_n2o, increment = 50, resamp = 1000, latlon = TRUE)
+
+plot(out_cplcorrelo_n2o)
+rm(out_cplvisit1, out_cplcorrelo_n2o)
+
+# trying pgirmess package uses library spdep including moran.test and geary.test functions.
+Distances are euclidean and in the same units as the spatial coordinates.
+As run below, using geographical coordinates in decimal degrees so are distance classes retuned in decimal degrees?
+# what if run on a spatial points data frame?
+# Uses normality assumption, which is faster, to assess significance
+coords_nap <- cbind(dg_napvisit1$map_lon_dd, dg_napvisit1$map_lat_dd)
+correl_napn2o_pg <- pgirmess::correlog(coords_nap, dg_napvisit1$log_n2o, method = "Moran", nbclass = NULL, alternative = "two.sided")
+
+head(round(correl_napn2o_pg, 2))
+tail(round(correl_napn2o_pg, 2))
+plot(correl_napn2o_pg)
+
+# apply to SpatialPointsDataFrame that has projected coordinates and believe returning distance classes in meters
+coords_sp_nap <- coordinates(nap_sp_visit1)
+correl_spnapn2o_pg <- pgirmess::correlog(coords_sp_nap, nap_sp_visit1@data$log_n2o, method = "Moran", nbclass = 30, alternative = "two.sided")
+
+head(round(correl_spnapn2o_pg, 2), 30)
+tail(round(correl_spnapn2o_pg ,2))
+plot(correl_spnapn2o_pg)
+```
+
+## Correlograms by Ecoregions
+Doug Anderson wrote functions to deal with the list of list output from the correlog function in the ncf package.
+He also wrote functions so that the correlogram of dissolved NO2 gas could be plotted by ecoregion using base plot, cowplot, or ggplot2 code.
+```{r, correlogram_ecoregions}
+# Original code from Doug Anderson at Neptune from file:  Neptune_correlogram_TDS05_Anderson.Rmd
+
+# set the theme of the ggplot
+theme_set(theme_bw()) # the theme I prefer is theme_bw()
+
+## Can define a new function to plot
+plot_correlog <- function(ncf_correlog, 
+                          # names of list components
+                          n_col = "n", 
+                          mean_col = "mean.of.class",
+                          cor_col = "correlation",
+                          xint_col = "x.intercept",
+                          pval_col = "p",
+                          name_col = NULL, #can change to column that has names of interest
+                          alpha = 0.05, # two sided significance level
+                          # plot labels and adjustments
+                          plot_title = "Correlogram", #can change to NULL
+                          plot_subtitle = waiver(),
+                          y_label = "Correlation",
+                          x_label = "Distance (mean-of-class)",
+                          pt_size = I(2), pt_color = "black",
+                          ln_size = I(0.5), ln_color = "black", 
+                          ln_type = "solid",
+                          title_size = 14, subtitle_size = 14, 
+                          axis_title_size = 14, axis_text_size = 14, 
+                          title_color = "black", 
+                          subtitle_color = "black",
+                          axis_title_color = "black", 
+                          axis_text_color = "black",
+                          title_bold = TRUE, axis_bold = FALSE,
+                          xaxis_color = "gray", xaxis_type = "solid",
+                          theme_style = theme_bw()){
+  # ncf_correlog is a list object exported from ncf::correlog() but can also be a dataframe with named columns
+
+  # prepare a data set for plotting based on the list components or data frame columns
+  dat <- data.frame(
+    n = ncf_correlog[[n_col]],
+    mean = ncf_correlog[[mean_col]],
+    cor = ncf_correlog[[cor_col]],
+    # xint = ncf_correlog[[xint_col]],
+    pval = ncf_correlog[[pval_col]]
+  )
+  # create a variable to help plot the hollow and solid points
+  # If a permutation test was performed, values significant at a nominal (two-sided) 5%-level 
+  # will be represented by filled circles and non-significant values by open circles.
+  dat$sig <- ifelse(dat$pval * 2 < alpha, TRUE, FALSE)
+
+  ## Start the plot
+  g <- ggplot(dat) +
+    theme_style +
+    geom_hline(yintercept = 0, color = xaxis_color, linetype = xaxis_type) +
+    # geom_point(aes(x=mean, y=cor, shape=sig), size=pt_size, color=pt_color) +
+    # geom_line(aes(x=mean, y=cor, group=1), size=ln_size, color=ln_color, linetype=ln_type) +
+    lemon::geom_pointline(aes(x=mean, y=cor, shape=sig, group=1),
+                          size = pt_size, color = pt_color, 
+                          linetype = ln_type, linesize = ln_size, 
+                          linecolor = ln_color) +
+    scale_shape_manual(values=c("TRUE"=19, "FALSE"=1),
+                       labels=c("TRUE"="Sig.", "FALSE"="Not Sig."),
+                       guide=FALSE) +
+    labs(x=x_label, y=y_label)
+
+  ## add a title
+  if(!is.null(name_col)){#include a title if column for name is provided or list object has a list item for name
+    if(!is.null(plot_title)){ #if plot_title is a string, separate the plot_title with the other name by a ':'
+      nm <- unique(as.character(ncf_correlog[[name_col]]))
+      g <- g + 
+        ggtitle(label = paste(plot_title, ": ", nm, sep=""), subtitle = plot_subtitle)
+    } else{ #if plot_title is null, only include the name
+      nm <- unique(as.character(ncf_correlog[[name_col]]))
+      g <- g + 
+        ggtitle(label = nm, subtitle = plot_subtitle)
+    }
+  } else{ # if there is no name column, only include the plot_title of interest (can be NULL)
+    g <- g + 
+      ggtitle(label = plot_title, subtitle = plot_subtitle) 
+  }
+
+  ## modify theme
+  g <- g + 
+    theme(plot.title = element_text(size=title_size, 
+                                    hjust=0.5, #change to a value between 0 (left) and 1 (right)
+                                    face=ifelse(title_bold == TRUE, "bold", "plain"), 
+                                    color=title_color),
+          plot.subtitle = element_text(size=subtitle_size, 
+                                       hjust=0.5, 
+                                       color=subtitle_color),
+          axis.title = element_text(size=axis_title_size, 
+                                    face=ifelse(axis_bold == TRUE, "bold", "plain"), 
+                                    color=axis_title_color),
+          axis.text = element_text(size=axis_text_size, #?theme
+                                   color=axis_text_color), #?element_text
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank())
+  return(g)
+}
+
+dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o, WSA9, WSA9_NAME, ONOFFNET, nitrate.n.result,chla.result.vol, MAX.BF, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size_cat)
+names(dg_1)
+dg_1 <- clean_names(dg_1)
+names(dg_1)
+dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o),3))
+## Subset dg_1 by visit_no == 1
+dg_visit1 <- dg_1 %>% dplyr::filter(visit_no == 1)
+
+## Perform the same test for each wsa9 using Z = "log_n2o"
+## Start by defining a function
+ncf_correlog <- function(dat, # data to send to ncf::correlog
+                         z_col, # name of the column in the dataset for water quality parameter
+                         i = 50, # increment to be used in ncf::correlog
+                         N = 1000, # number of resamples in ncf::correlog
+                         x_col = "map_lon_dd", # name of the column in the dataset for longitude
+                         y_col = "map_lat_dd", # name of the column in the dataset for latitude
+                         lat_lon = TRUE, # if the x and y coordinates for ncf::correlog are lat and lon
+                         rem_missing = FALSE, # argument of ncf::correlog
+                         run_quietly = FALSE){ # argument of ncf::correlog
+  library(ncf)
+
+  X = dat[[x_col]] #grabs the x_col from data.frame
+  Y = dat[[y_col]] #grabs the y_col from data.frame
+  Z = dat[[z_col]] #grabs the z_col from data.frame
+
+  ## Use ncf::correlog
+  crlg <- ncf::correlog(x = X, y = Y, z = Z, 
+                        increment = i, resamp = N, latlon = lat_lon, 
+                        na.rm = rem_missing, quiet = run_quietly)
+  return(crlg)
+}
+
+## Split the dataframe by each wsa9: turns the data frame into a list 
+##   where each list item is a data frame filtered to that wsa9
+## lapply applies the function ncf_correlog on each list object (a data frame)
+## the final product is a list of correlog objects
+## this is similar to looping
+## setting seed so get same result from permutations ncf::correlog function
+set.seed(2035)
+log_n2o_correlog_list <- dg_visit1 %>%
+  split(.$wsa9) %>% # '.' means use the data frame that preceded this line
+  lapply(function(x) ncf_correlog(dat = x, z_col = "log_n2o")) 
+#ncf_correlog is the function name, 
+#dat is the paramater that takes a data.frame,
+#z_col is the parameter that takes a string name for a column of a water quality paramater
+
+## look at the structure: list of 9 items where each item is a 
+##   list containing 6 items: n, mean, correlation, x.int, p, call
+str(log_n2o_correlog_list)
+
+## Function to remove the list item 'call' which is of a different length 
+##   than the other items (n, mean, correlation, x.int, and p)
+## This function will also convert each list item to a dataframe
+crlg_to_dat <- function(ncf_correlog, # ncf_correlog is a list object that was output from ncf::correlog
+                        n_col = "n", 
+                        mean_col = "mean.of.class",
+                        cor_col = "correlation",
+                        xint_col = "x.intercept",
+                        pval_col = "p"){ 
+  dat <- data.frame(
+    n = ncf_correlog[[n_col]],
+    mean = ncf_correlog[[mean_col]],
+    cor = ncf_correlog[[cor_col]],
+    xint = ncf_correlog[[xint_col]],
+    pval = ncf_correlog[[pval_col]]
+  )
+  return(dat)
+}
+
+## log_n20_correlog_list is already a list so there is no need to split
+log_n2o_correlog_dats <- lapply(log_n2o_correlog_list, 
+                                function(x) crlg_to_dat(ncf_correlog = x))
+str(log_n2o_correlog_dats) # each item in the list is now a dataframe
+
+## combine into one dataframe to grab the wsa9 (notice the name no longer has an s at the end)
+
+log_n2o_correlog_dat <- log_n2o_correlog_dats %>%
+  dplyr::bind_rows(., .id = "wsa9")
+
+str(log_n2o_correlog_dat)
+head(log_n2o_correlog_dat)
+## Make individual plots, the following object is a list containing a ggplot for each
+log_n2o_correlog_plots <- log_n2o_correlog_dat %>%
+  split(.$wsa9) %>% # splits the data frame by wsa9, which creates a list with 9 data frames
+  lapply(function(x) plot_correlog(ncf_correlog = x, # use the data.frame for each wsa9 to make a plot
+                                   plot_title = NULL,
+                                   n_col = "n", 
+                                   mean_col = "mean",
+                                   cor_col = "cor",
+                                   xint_col = "xint",
+                                   pval_col = "pval",
+                                   name_col = "wsa9",
+                                   title_size = 14, 
+                                   subtitle_size = 14, 
+                                   axis_title_size = 14, 
+                                   axis_text_size = 14))
+## if hard to see writing run again but increase "title_size", "subtitle_size", "axis_title_size", "axis_text_size" as needed
+
+## Look at each piece individually or scroll through
+length(log_n2o_correlog_plots) # shows 9 items
+# log_n20_correlog_plots
+log_n2o_correlog_plots$CPL #or log_n20_correlog_plots[["CPL"]] or log_n2o_correlog_plots[[1]]
+# log_n2o_correlog_plots$NAP
+# log_n2o_correlog_plots$NPL
+# log_n2o_correlog_plots$SAP
+# log_n2o_correlog_plots$SPL
+# log_n2o_correlog_plots$TPL
+# log_n2o_correlog_plots$UMW
+# log_n2o_correlog_plots$WMT
+# log_n2o_correlog_plots$XER
+
+## cowplot package (or other) can be used to group them into a matrix plot
+log_n2o_correlog_plotgrid <- 
+  cowplot::plot_grid(plotlist=log_n2o_correlog_plots, ncol=3)
+print(log_n2o_correlog_plotgrid) # click 'Zoom' in plots quadrant, takes a few seconds to render
+## can save the plot using 
+# cowplot::ggsave(filename = "FILENAME.png", #saves a png version of the figure to the working folder
+#                 plot = log_n20_correlog_plotgrid)
+## can change .png to other forms
+## can play with scale, width, and height to adjust the output of the figure saved to the png file
+
+## alternatively, can make a facet plot since log_n20_correlog_dat is a dataframe in the correct format
+names(log_n2o_correlog_dat) #"n"    "mean" "cor"  "xint" "pval" "wsa9"
+
+## create column to shape point, sig == TRUE means two-sided test < 0.05
+log_n2o_correlog_dat$sig <- ifelse(log_n2o_correlog_dat$pval * 2 < 0.05,
+                                   TRUE, FALSE) 
+head(log_n2o_correlog_dat)
+## can change wsa9 to a factor to arrange what order the plots are shown, otherwise default arranges them alphabetically
+# log_n20_correlog_dat$wsa9 <- factor(log_n20_correlog_dat, 
+#                                     levels = c("CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER"))
+
+## make facet plot
+log_n2o_correlog_facetplot <- ggplot(log_n2o_correlog_dat) +
+  geom_hline(yintercept = 0, color = "gray", linetype = "solid") +
+  # geom_point(aes(x=mean, y=cor, shape=sig), size=I(2), color="black") +
+  # geom_line(aes(x=mean, y=cor, group=1), size=I(0.5), color="black", linetype="solid") +
+  lemon::geom_pointline(aes(x=mean, y=cor, shape=sig, group=1), 
+                        size=I(2), color = "black", 
+                        linetype = "solid", linesize = I(0.5), 
+                        linecolor = "black") +
+  scale_shape_manual(values=c("TRUE"=19, "FALSE"=1), 
+                     labels=c("TRUE"="Sig.", "FALSE"="Not Sig."),
+                     guide=FALSE) +
+  facet_wrap(~wsa9, ncol=3, scales="free") +
+  labs(x="Distance (mean-of-class)", y="Correlation") +
+  theme_bw() +
+  theme(plot.title = element_text(size=14, hjust=0.5, 
+                                  face="bold", color="black"),
+        plot.subtitle = element_text(size=14, hjust=0.5, color="black"),
+        axis.title = element_text(size=14, face="plain", color="black"),
+        axis.text = element_text(size=14, color="black"),
+        strip.text = element_text(size=14, color="black", face="bold"),
+        strip.background = element_rect(fill="gray", color="black"),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
+print(log_n2o_correlog_facetplot)
+## can save the plot using 
+# cowplot::ggsave(filename = "FILENAME2.png", #saves a png version of the figure to the working folder
+#                 plot = log_n20_correlog_facetplot) 
+## can change .png to other forms
+## can play with scale, width, and height to adjust the output of the figure saved to the png file
+
+```
+
+```
+
 ## Semivariograms

### 2020-09-22 scripts/dataMunge.Rmd
Resolving version conflicts in dgIndicatorAnalysis

@@
 The dissolved gas file contains data from sites in Alaska that are not present in the weights + lake area file.
-Alaska sites are not an offical component of the NLA and will removed from this analysis.
+Alaska sites are not an official component of the NLA and will removed from this analysis.
 ``` {r}
@@
 Three site visits contain only NA for temperature.
-Four site visits contain only NA for do.
+Four site visits contain only NA for DO.
 Two of the incidents include notes that sonde was forgotten or malfunctioning.
@@
 
-No deep DO or temp at NLA17_MI-10073, visit 1 becuase no measurement taken at deepst recorded depth.  Replace with next shallowest measurement.
+No deep DO or temp at NLA17_MI-10073, visit 1 because no measurement taken at deepest recorded depth.  Replace with next shallowest measurement.
 
@@
 
-No surface temp at NLA17_NE-10060, visit 1 becuase no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
+No surface temp at NLA17_NE-10060, visit 1 because no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
 
@@
 
-No bottom temp or O2 at NLA17_NH-10063, visit 1 becuase no measurement taken at greatest depth (8m). Replace with next deeptest measurement (7.5m)
+No bottom temp or O2 at NLA17_NH-10063, visit 1 because no measurement taken at greatest depth (8m). Replace with next deepest measurement (7.5m)
 
@@
 
-No surface temp or O2 at NLA17_OR-10003, visit 1 becuase no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
+No surface temp or O2 at NLA17_OR-10003, visit 1 because no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
 
@@
 
-No bottom temp or O2 at NLA17_SD-10057, visit 1 becuase no measurement taken at deepest depth (4m). Replace with next deepest measurement (3m)
+No bottom temp or O2 at NLA17_SD-10057, visit 1 because no measurement taken at deepest depth (4m). Replace with next deepest measurement (3m)
 
@@
 
-Only remainining missing values are observations where field crews noted sonde malfunctions.
+Only remaining missing values are observations where field crews noted sonde malfunctions.
 

### 2020-10-05 scripts/dataMunge.Rmd
Added code to calculate upper mixed layer DO and temp.

@@
 library(rLakeAnalyzer) # for buoyancy frequency
+library(zoo) # interpolate NA temp/DO
 
@@
 ## Manipulate
-Impliment unit conversion and create source/sink column:
+Implement unit conversion and create source/sink column:
 ``` {r}
@@
 # Inspect results of join.  Looks good.  
-# Only one site missing temperature and DO data. (NLA17_CA-10082)
+# Only one site missing all temperature and DO data. (NLA17_CA-10082)
 # One site missing all DO data (NLA17_ID-10084)
@@
   arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN)
+
+# NLA17_ID-10001 visit 1 only has temp/DO measured at surface depth.  Notes indicate probe malfunction.
+# Second site visit shows similar surface DO/temp and no thermal stratification.  Replace visit 1 with visit 2
+# data
+temp.o21[temp.o21$SITE_ID == "NLA17_ID-10001" & temp.o21$"VISIT_NO" == 1, c("OXYGEN", "TEMPERATURE")] = 
+  temp.o21[temp.o21$SITE_ID == "NLA17_ID-10001" & temp.o21$"VISIT_NO" == 2, c("OXYGEN", "TEMPERATURE", "DEPTH")] %>%
+  group_by(DEPTH) %>% # Need to aggregate duplicate measurements at depth = 0
+  summarise(OXYGEN = mean(OXYGEN), TEMPERATURE = mean(TEMPERATURE)) %>%  # mean by depth
+  select(-DEPTH) # remove depth
 
@@
 
+`r temp.o2.2 %>% filter(is.na(DEPTH)) %>% nrow()` observations report NA for DEPTH, DO, and TEMPERATURE.  These will be removed.
+
+``` {r}
+# Missing depth values
+# All also have NA for DO and Temp.  Remove from dataset
+temp.o2.2 %>% 
+  filter(is.na(DEPTH))
+
+dim(temp.o2.2) # 10133, 5
+temp.o2.2 <- temp.o2.2 %>% filter(!is.na(DEPTH))
+dim(temp.o2.2) # 10085, 5
+```
+
+`r temp.o2.2 %>% filter(is.na(TEMPERATURE.M) & is.na(OXYGEN.M)) %>% nrow()` observations report NA for temperature and DO .
+These are always at the shallowest (0m) or deepest depths.
+When at shallowest, there is a 0.1m measurement.
+When at deepest, I'm guessing no measurements were made at that depth.
+Remove these observations.
+
+``` {r}
+# NA for temp and DO
+fewMissing <- temp.o2.2 %>% 
+  filter(is.na(TEMPERATURE.M) & is.na(OXYGEN.M)) %>%
+  select(SITE_ID, VISIT_NO) %>%
+  distinct(SITE_ID, VISIT_NO)
+
+temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% paste0(fewMissing$SITE_ID, fewMissing$VISIT_NO))
+
+dim(temp.o2.2) # 10085, 5
+temp.o2.2 <- temp.o2.2 %>% filter(!(is.na(TEMPERATURE.M) & is.na(OXYGEN.M)))
+dim(temp.o2.2) # 10081, 5
+```
+
+`r temp.o2.2 %>% filter(is.na(TEMPERATURE.M) | is.na(OXYGEN.M)) %>% nrow()` observations report NA for depth or temperature.
+If these are fairly random within the depth profile, they can be imputed.
+No DO data for NLA17_ID-10002 (see above).
+The remaining cases are fairly random.
+Will impute via linear interpolation.
+
+```{r}
+fewMissing <- temp.o2.2 %>% filter(is.na(TEMPERATURE.M) | is.na(OXYGEN.M)) %>%
+  select(SITE_ID, VISIT_NO) %>%
+  distinct(SITE_ID, VISIT_NO)
+
+temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% paste0(fewMissing$SITE_ID, fewMissing$VISIT_NO))
+
+temp.o2.2 <- temp.o2.2 %>% 
+  group_by(SITE_ID, VISIT_NO) %>%
+  mutate(OXYGEN.M = na.approx(OXYGEN.M, DEPTH, na.rm = FALSE, rule = 2), #rule = 2 assigns value closest to boundary 
+         TEMPERATURE.M = na.approx(TEMPERATURE.M, DEPTH, na.rm = FALSE, rule = 2)) %>%
+  ungroup()
+
+dim(temp.o2.2) #10081, 5
+```
+
 Finally, calculate the buoyancy frequency depth profile for each waterbody and print to buoyancyFrequency.pdf for inspection.
@@
   lapply(function(x) {
-    data.frame(BF = buoyancy.freq(wtr = x$TEMPERATURE.M, depths = x$DEPTH), # calculate buoyance frequency
+    data.frame(BF = buoyancy.freq(wtr = x$TEMPERATURE.M, depths = x$DEPTH), # calculate buoyancy frequency
                SITE_ID = unique(x$SITE_ID), # add site ID
@@
 
-# there are three non-finite bf values, lets have a look
+# there are two non-finite bf values, lets have a look
 bf.na <- bf %>% filter(!is.finite(BF)) %>%
@@
 
-# These site have a single temp measurement made at shallow depths (<1m).  Presumably these are 
-# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0 (see below)
+# These site have a single temp measurement made at shallow depths (<=1m, NLA17_MD-10004, NLA17_ND-10189).
+# Presumably these are shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0 (see below).
 bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
@@
 `r bf.max %>% summarize(n.na = sum(!is.finite(bf.max$MAX.BF))) %>% pull()` of the site have a non-finite value for maximum buoyancy frequency value.
-These are the three shallow sites with only a single temperature measurement.
+These are the two shallow sites with only a single temperature measurement.
 Assign these the minimum bf.max value.
@@
 min.max.bf <- bf.max %>% 
-  filter(is.finite(MAX.BF)) %>%
+  filter(is.finite(MAX.BF), MAX.BF > 0) %>%
   summarize(min.max.bf = min(MAX.BF))
@@
 
+## Calculate thermocline depth
+
+Dissolved gas concentrations collected near the water surface may be more related to mean conditions in the upper mixed layer than conditions near the air-water interface (e.g.
+depth = 0.1).
+Below we calculate depth to thermocline.
+The thermocline data contains information from the extra Oregon sites (see above) and the AZ site where the gas samples were unlabeled and not analyzed.
+This is as expected.
+No thermocline data for the California site (CA-10082) where the sonde was reported missing/malfunctioning.
+
+``` {r}
+therm <- temp.o2.2 %>%
+  group_by(SITE_ID, VISIT_NO) %>%
+  summarise(thermD = thermo.depth(TEMPERATURE.M, DEPTH))
+
+summary(therm$thermD) # 395 sites with no evidence of stratification
+dim(therm) #1209, 3 collapsed temp.o2.2 into one observation per site x visit
+
+# Merge therm depth into temp.o2.2 for calcs below.  A single therm depth repeated for each row in unique site x visit
+dim(temp.o2.2) # 10081, 5
+dim(therm) #1209, 3
+temp.o2.2 <- left_join(temp.o2.2, therm)
+dim(temp.o2.2) # 10081, 6
+# ggplot(dg, aes(thermD, MAX.BF)) + geom_point() + ylim(0,0.04)
+```
+
+## Surface mixed layer DO and temperature
+
+Average all values between surface and thermocline.
+The surface mixed layer data contains information from the extra Oregon sites (see above) and the AZ site where the gas samples were unlabeled and not analyzed.
+This is as expected.
+No mixed layer data for the California site (CA-10082) where the sonde was reported missing/malfunctioning.
+
+```{r}
+mixed.layer.do.temp <- temp.o2.2 %>%
+  group_split(SITE_ID, VISIT_NO) %>% # split into list of dataframes
+  lapply(function(x) { # apply to each element of list
+    data.i <- if(any(is.na(x$thermD))) { # if not stratified (thermD is NA or NaN)
+      x # then all data
+    } else { 
+      filter(x, DEPTH <= thermD) # if stratified, then only data at depths <= thermD
+    }
+    OXYGEN.EPI <- mean(data.i$OXYGEN.M, na.rm = TRUE) # calculate mean DO
+    TEMP.EPI <- mean(data.i$TEMPERATURE.M, na.rm = TRUE) # calculate mean Temp
+    # dump into dataframe
+    data.frame(SITE_ID = unique(x$SITE_ID), # extract site id
+               VISIT_NO = unique(x$VISIT_NO), # extract visit no
+               OXYGEN.EPI = OXYGEN.EPI,
+               TEMP.EPI = TEMP.EPI,
+               thermD = unique(x$thermD), # identical values within each df, extract unique
+               stringsAsFactors = FALSE)
+  }
+  ) %>% bind_rows() # collapse to df.  dplyr implementation of do.call(rbind, list)
+
+# are all SITE_ID values in mixed.layer.do.temp also present in dg?
+mixed.layer.do.temp %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007, as expected
+distinct(SITE_ID, VISIT_NO)
+
+# are all site.id values in dg also present in mixed.layer.do.temp?
+# Just CA site, as expected.
+dg %>% filter(!(site.id %in% mixed.layer.do.temp$SITE_ID)) %>% # just CA site, as expected
+  select(site.id, visit.no)
+
+# merge with dg
+dim(dg) # 2369, 375
+dim(mixed.layer.do.temp) #1209, 5
+dg <- left_join(dg, mixed.layer.do.temp, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))
+dim(dg) # 2369, 378
+```
+
 ## Calculate surface and bottom DO/temp
@@
 
-These calculations produced 8 site visits with at least one NA for surface/bottom temp/DO.
-
-```{r}
-o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
-```
-
-o2.surf at NLA17_FL-10029 is a case with no DO at shallowest depth, but good measurements below.  Replace with next shallowest value.
-``` {r}
-# Inspect raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_FL-10029", VISIT_NO == 1)
-
-# update data
-o2.s.b <- o2.s.b %>%
-  mutate(o2.surf = replace(o2.surf, 
-                           SITE_ID == "NLA17_FL-10029" & VISIT_NO == 1,
-                           temp.o2 %>% filter(SITE_ID=="NLA17_FL-10029",
-                                              VISIT_NO == 1,
-                                              DEPTH == 1) %>%
-                             select(OXYGEN) %>% pull()))
-
-```
-No bottom water DO or temp at NLA17_ID-10001, visit 1 due to probe malfunction.
-
-```{r}
-temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10001", VISIT_NO == 1)
-```
-
-No DO NLA17_ID-10084, visit 1 due to probe malfunction.
-
-```{r}
-temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10084", VISIT_NO == 1)
-```
-
-No deep DO or temp at NLA17_MI-10073, visit 1 because no measurement taken at deepest recorded depth.  Replace with next shallowest measurement.
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_MI-10073", VISIT_NO == 1)
-
-# update aggregated data
-o2.s.b <- o2.s.b %>%
-  mutate(o2.bottom = replace(o2.bottom, 
-                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
-                                              VISIT_NO == 1,
-                                              DEPTH == 1.3) %>%
-                             select(OXYGEN.M) %>% pull()),
-         temp.bottom = replace(temp.bottom, 
-                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
-                                              VISIT_NO == 1,
-                                              DEPTH == 1.3) %>%
-                             select(TEMPERATURE.M) %>% pull()))
-```
-
-No surface temp at NLA17_NE-10060, visit 1 because no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_NE-10060", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.surf = replace(temp.surf, 
-                           SITE_ID == "NLA17_NE-10060" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NE-10060",
-                                              VISIT_NO == 1,
-                                              DEPTH == 0.1) %>%
-                             select(TEMPERATURE.M) %>% pull()))
-```
-
-No bottom temp or O2 at NLA17_NH-10063, visit 1 because no measurement taken at greatest depth (8m). Replace with next deepest measurement (7.5m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_NH-10063", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.bottom = replace(temp.bottom, 
-                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
-                                              VISIT_NO == 1,
-                                              DEPTH == 7.5) %>%
-                             select(TEMPERATURE.M) %>% pull()),
-         o2.bottom = replace(o2.bottom, 
-                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
-                                              VISIT_NO == 1,
-                                              DEPTH == 7.5) %>%
-                             select(OXYGEN.M) %>% pull()))
-```
-
-No surface temp or O2 at NLA17_OR-10003, visit 1 because no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_OR-10003", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.surf = replace(temp.surf, 
-                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
-                                              VISIT_NO == 1,
-                                              DEPTH == 0.1) %>%
-                             select(TEMPERATURE.M) %>% pull()),
-         o2.surf = replace(o2.surf, 
-                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
-                                              VISIT_NO == 1,
-                                              DEPTH == 0.1) %>%
-                             select(OXYGEN.M) %>% pull()))
-```
-
-No bottom temp or O2 at NLA17_SD-10057, visit 1 because no measurement taken at deepest depth (4m). Replace with next deepest measurement (3m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_SD-10057", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.bottom = replace(temp.bottom, 
-                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
-                                              VISIT_NO == 1,
-                                              DEPTH == 3) %>%
-                             select(TEMPERATURE.M) %>% pull()),
-         o2.bottom = replace(o2.bottom, 
-                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
-                                              VISIT_NO == 1,
-                                              DEPTH == 3) %>%
-                             select(OXYGEN.M) %>% pull()))
-```
-
-Only remaining missing values are observations where field crews noted sonde malfunctions.
+These calculations produced 1 site visit with at least one NA for surface/bottom temp/DO.  This is NL17_ID-10084 with no DO data (see above)
 
@@
 # are all SITE_ID values in o2 and temp data also present in dg?
-o2.s.b %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007
+o2.s.b %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007, as expected
 print(n= Inf)
@@
 # dim(o2.s.b) # 1209, 6
-# dim(dg) # 2369, 375
+# dim(dg) # 2369, 378
 
@@
 dg <- left_join(dg, o2.s.b, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
-# dim(dg) # 2369, 379, good!
+# dim(dg) # 2369, 382, good!
 ```

### 2020-10-07 scripts/dataMunge.Rmd
Transferred most data manipulation code from dataMunge.Rmd to NLA repo.

--- 2020-09-22 scripts/dataMunge.Rmd
@@
 # load libraries
-library(plyr) # for ddply, load before tidyverse
 library(tidyverse) # dplyr, ggplot
-library(forcats) # fct_explicit_na()
-library(rLakeAnalyzer) # for buoyancy frequency
 
@@
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-02-24.txt.
-Data on waterbody surface area and design weights were provided by Karen Blocksom on 12/13/2019.
-These data files are stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-10-07.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
 When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2020-02-24.txt"),
+                               "inputData/nla17gasDataAggregated_2020-10-07.txt"),
                  header = TRUE, sep = "\t", as.is = TRUE) 
-
-# Read site weights and waterbody area data
-wt.area <- read.csv(file = paste0(localPath, 
-                               "/Environmental Protection Agency (EPA)/",
-                               "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/NLA17_SiteWeights_Areas.csv"),
-                    as.is = TRUE)
-
-# Read in temp profile data
-temp.o2 <- read.csv(file = paste0(localPath, 
-                               "/Environmental Protection Agency (EPA)/",
-                               "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/NLA17_Profile_Data_Updated_11Feb2020.csv"),
-                    as.is = TRUE)
-```
-
-## Merge dissolved gas with survey weights + area
-Unique waterbodies are identified by 'site id'.  Some waterbodies were visited twice, therefore each unique sampling trip is identified by a combination of 'site id' and visit number.
-
-``` {r}
-dg %>% distinct(site.id, visit.no)
-```
-
-Survey weights and lake area are invariant across site visits, therefore the weights + lake area file does not contain a visit number field:
-``` {r}
-# no visit number in this file
-names(wt.area)
-```
-
-The two data sets will be merged on the 'site id' fields.
-The dissolved gas file contains data from sites in Alaska that are not present in the weights + lake area file.
-Alaska sites are not an official component of the NLA and will removed from this analysis.
-``` {r}
-# Will merge datasets on SITE_ID field.  Are all site.id values in dg present in wt.area?
-dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id) # no.
-
-# Remove AK sites
-dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
-```
-
-There are 23 observations in the weights + lake area file that are absent from the dissolved gas file.
-One observation is from Arizona.
-This corresponds to notes in the chain of custody form indicating the vials arrived unlabeled and were not run.
-The remaining sites are from OR.
-Karen Blocksom confirmed that these sites were part of the OR state survey, but collected 'most' of the NLA measurements.
-They were included in the final NLA dataset for 'the purpose of adjusting weights'.
-They did not collect gas samples and can be disregarded for this analysis.
-
-``` {r}
-# Any site id values in wt.area, but not present in dg?
-wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # 23 sites w/out dissolved gas data
-```
-
-Finally, lets merge these two files on site id.  Use left_join to retain all dissolved gas data, but disregard OR and AZ sites where we don't have gas data.
-``` {r}
-dim(dg) # 2369, 364
-
-dg <- left_join(dg, wt.area, by = c("site.id" = "SITE_ID"))
-
-dim(dg) # 2369, 369;  good retained all DG observations, added a few columns.
 ```
@@
 
-The survey design is stratified by State.
-Within each state, an unequal probability category based on 5 lake size classes is used.
-Use lake area data provided by Karen to define lake area categories.
-```{r}
-dg <- dg %>%
-  mutate(
-    # add state
-    state = substr(site.id, 7, 8),
-    size_cat = factor( cut( AREA_HA,
-                            breaks = c( -Inf, 4, 10, 20, 50, Inf ),
-                            labels = c( "min_4", "4_10", "10_20", "20_50", "50_max") ) )
-    )
-
-```
-All observations have a value for the size_cat design variable.
-
-``` {r}
-table(dg$size_cat, useNA = "ifany")
-```
-
-## Calculate buoyancy frequency
-Water temperature and dissolved oxygen profiles were measured at the index site in each waterbody.
-A variety of metrics describing the degree of thermal stratification can be extracted from these data.
-Webb et al.
-used the Brunt–Väisälä buoyancy frequency as an indicator of thermal stratification strength.
-We will do the same here.
-
-Three site visits contain only NA for temperature.
-Four site visits contain only NA for DO.
-Two of the incidents include notes that sonde was forgotten or malfunctioning.
-Two of the three sites were revisited later in the study and temperature data were collected.
-Lets assign data from the second site visit to the first visit when no temperature data were collected.
-This leaves only NLA17_CA-10082, visit 1, without temperature data.
-NLA17_ID-10084 has no DO data and was not resampled.
-Data sheet comment "Probe not measuring LDO for all depths".
-
-```{r}
-# Sites that only contain NAs.
-missing <- temp.o2 %>% group_by(SITE_ID, VISIT_NO) %>% 
-  summarise(all.temp = all(is.na(TEMPERATURE)),
-            all.do = all(is.na(OXYGEN))) %>% 
-  filter(all.temp == TRUE | all.do == TRUE) %>%
-  select(SITE_ID, VISIT_NO, all.temp, all.do)
-
-# View sites with missing temp/do
-missing
-
-# Two of the sites have data from second site visit
-temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
-  select(SITE_ID, VISIT_NO, TEMPERATURE, OXYGEN)
-
-# Extract data from second visit to assign to first visit
-replacementData <- temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID, VISIT_NO == 2) %>%
-  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
-  mutate(VISIT_NO = 1) # change visit no to one for join
-
-# Join the data
-temp.o21 <- full_join(temp.o2, replacementData)
-
-# Inspect results of join.  Looks good.  
-# Only one site missing temperature and DO data. (NLA17_CA-10082)
-# One site missing all DO data (NLA17_ID-10084)
-temp.o21 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
-  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>% 
-  arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN)
-
-```
-
-The field crews measured temperature/DO starting near the water surface, then at increasing depths.
-A second measurement was made at the water surface before the sonde was removed from the water, therefore the file contains duplicate measurements at the shallowest depths.
-Duplicate measurements corrupt the buoyancy frequency function and will therefore be aggregated.
-``` {r}
-# Aggregate across SITE_ID, VISIT_NO, and DEPTH.
-I don't like the dplyr solution for this (group_map) and I'm not familiar with the purr::map_ functions.
-I'm going old-school and busting out the trusty plyr::ddply.
-temp.o2.2 <- temp.o21 %>% select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
-  plyr::ddply(., .(SITE_ID, VISIT_NO, DEPTH), summarize, 
-              TEMPERATURE.M = mean(TEMPERATURE, na.rm = TRUE),
-              OXYGEN.M = mean(OXYGEN, na.rm = TRUE))
-
-dim(temp.o21) # 11003
-dim(temp.o2.2) # 10133, 870 observations aggregated
-```
-
-Finally, calculate the buoyancy frequency depth profile for each waterbody and print to buoyancyFrequency.pdf for inspection.
-
-``` {r}
-# The buoyancy frequency is calculated at multiple depths, but at fewer depths than the where 
-# temperature was measured.  The output of the buoyancy frequency function is shorter 
-# than the input, which is incompatible with dplyr mutate and summarize.  The dplyr alternative
-# is group_map which produces a list, but the grouping variables are stripped from the list
-# elements.  I really don't like this.  I think this can be addressed by combining group_map
-# with purr::map_ functions, but I'm not familiar with purr.  Rather, I'll use the old school
-# split -lapply - do(rbind) approach.
-
-# First, strip out any NA, NaN, Inf...etc
-
-bf <- temp.o2.2 %>%
-  filter(is.finite(TEMPERATURE.M)) %>% # This completely removes the CA site with only NA for temp data.
-  group_split(SITE_ID, VISIT_NO) %>% # split into list of df
-  lapply(function(x) {
-    data.frame(BF = buoyancy.freq(wtr = x$TEMPERATURE.M, depths = x$DEPTH), # calculate buoyance frequency
-               SITE_ID = unique(x$SITE_ID), # add site ID
-               VISIT_NO = unique(x$VISIT_NO), # add visit no
-               stringsAsFactors = FALSE)}) %>% 
-  lapply(function(x) x %>% mutate(DEPTHS.BF = attr(BF, "depths"))) %>% # depths stored as attribute, assign to column
-  do.call("rbind", .) # collapse to df
-
-# there are three non-finite bf values, lets have a look
-bf.na <- bf %>% filter(!is.finite(BF)) %>%
-  select(SITE_ID, VISIT_NO) %>%
-  mutate(SITE.VISIT = paste0(SITE_ID, VISIT_NO))
-
-bf.na
-
-# These site have a single temp measurement made at shallow depths (<1m).  Presumably these are 
-# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0 (see below)
-bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
-temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)
-
-# plot buoyancy frequency depth profile
-# (http://stackoverflow.com/questions/8018961/connect-points-in-qplot-by-adjacent-y-value-not-x-value)
-# https://stackoverflow.com/questions/29034863/apply-a-ggplot-function-per-group-with-dplyr-and-set-title-per-group
-bf.plot <- bf %>%
-  group_by(SITE_ID, VISIT_NO) %>%
-  do(plots = ggplot(data = .) +
-       aes(DEPTHS.BF, BF) +
-    geom_point() +
-    geom_line() +
-    scale_x_reverse() +
-    coord_flip() +
-    ggtitle(paste(unique(.$SITE_ID), "VISIT_NO = ", unique(.$VISIT_NO))))
-
-# pdf("output/figures/buoyancyFrequency.pdf", onefile = TRUE)
-# bf.plot$plots
-# dev.off()
-```
-
-Webb et al. used the maximum buoyancy frequency value as an index of thermal stratification.  The maximum value should occur near the thermocline.  Here is an example of a classical pattern:
-```{r}
-bf.plot$plots[92]
-```
-
-Extract the max buoyancy frequency value for each waterbody after converting negative BF values to 0, per Webb.
-``` {r}
-# Finally, extract max buoyancy for each SITE_ID x VISIT_NO
-bf.max <- bf %>% 
-  mutate(BF = ifelse(BF < 0,
-                     0,
-                     BF)) %>%
-  group_by(SITE_ID, VISIT_NO) %>%
-  summarise(MAX.BF = max(BF, na.rm = TRUE)) %>%
-  ungroup()
-```
-
-`r bf.max %>% summarize(n.na = sum(!is.finite(bf.max$MAX.BF))) %>% pull()` of the site have a non-finite value for maximum buoyancy frequency value.
-These are the three shallow sites with only a single temperature measurement.
-Assign these the minimum bf.max value.
-``` {r}
-bf.max %>% filter(!is.finite(MAX.BF))
-
-min.max.bf <- bf.max %>% 
-  filter(is.finite(MAX.BF)) %>%
-  summarize(min.max.bf = min(MAX.BF))
-
-bf.max <- bf.max %>%
-  mutate(MAX.BF = replace(MAX.BF, 
-                          !is.finite(MAX.BF), # if not finite
-                          min.max.bf$min.max.bf)) # then min
-
-# that took care of it
-bf.max %>% filter(!is.finite(MAX.BF))
-```
-
-Finally, merge buoyancy frequency data with dg.
-The buoyancy frequency data contains information from the extra Oregon sites (see above) and the AZ site where the gas samples were unlabeled and not analyzed.
-This is as expected.
-The buoyancy frequency data do not contain observations for the California site (CA-10082) where the sonde was reported missing/malfunctioning.
-```{r}
-# are all SITE_ID values in buoyancy frequency also present in dg?
-bf.max %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007
-print(n= Inf)
-
-# are all site.id values in dg also present in bf.max?
-# Just CA site, as expected.
-dg %>% filter(!(site.id %in% bf.max$SITE_ID)) %>% # no
-  select(site.id, visit.no)
-
-# dim(bf.max) # 1209, 3
-# dim(dg) # 2369, 374
-
-# Left join: retain all dg values, exclude non-matching bf values (i.e. OR and AZ)
-dg <- left_join(dg, bf.max, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
-# dim(dg) # 2369, 375, good!
-```
-
-The California site with the broken sonde has no bf data.  Nothing we can do about it.
-```{r}
-dg %>% filter(is.na(MAX.BF)) %>%
-  select(site.id, visit.no)
-```
-
-## Calculate surface and bottom DO/temp
-Webb et al.
-used the 'surface' and 'bottom' water DO percent saturation and temperature as predictors, but didn't give more details on how these indices were calculated.
-We will do something similar and define the 'shallowest' and 'deepest' measurements as 'surface' and 'bottom', respectively, but will use mg.l for dissolved oxygen.
-```{r}
-# o2 surface (s) and bottom (b)
-o2.s.b <- 
-temp.o2.2 %>% 
-  group_by(SITE_ID, VISIT_NO) %>%
-  slice(c(which.min(DEPTH), which.max(DEPTH))) %>% # min and max depth by group
-  arrange(SITE_ID, VISIT_NO, DEPTH) %>%
-    mutate(DEPTH = ifelse(row_number() == 1,
-                        "surface",
-                        "bottom")) %>%
-  pivot_wider(id_cols = c(SITE_ID, VISIT_NO), 
-              names_from = DEPTH, 
-              values_from = c(TEMPERATURE.M, OXYGEN.M)) %>%
-  rename(temp.surf = TEMPERATURE.M_surface,
-         temp.bottom = TEMPERATURE.M_bottom,
-         o2.surf = OXYGEN.M_surface,
-         o2.bottom = OXYGEN.M_bottom)
-
-o2.s.b
-
-```
-
-These calculations produced 8 site visits with at least one NA for surface/bottom temp/DO.
-
-```{r}
-o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
-```
-
-o2.surf at NLA17_FL-10029 is a case with no DO at shallowest depth, but good measurements below.  Replace with next shallowest value.
-``` {r}
-# Inspect raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_FL-10029", VISIT_NO == 1)
-
-# update data
-o2.s.b <- o2.s.b %>%
-  mutate(o2.surf = replace(o2.surf, 
-                           SITE_ID == "NLA17_FL-10029" & VISIT_NO == 1,
-                           temp.o2 %>% filter(SITE_ID=="NLA17_FL-10029",
-                                              VISIT_NO == 1,
-                                              DEPTH == 1) %>%
-                             select(OXYGEN) %>% pull()))
-
-```
-No bottom water DO or temp at NLA17_ID-10001, visit 1 due to probe malfunction.
-
-```{r}
-temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10001", VISIT_NO == 1)
-```
-
-No DO NLA17_ID-10084, visit 1 due to probe malfunction.
-
-```{r}
-temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10084", VISIT_NO == 1)
-```
-
-No deep DO or temp at NLA17_MI-10073, visit 1 because no measurement taken at deepest recorded depth.  Replace with next shallowest measurement.
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_MI-10073", VISIT_NO == 1)
-
-# update aggregated data
-o2.s.b <- o2.s.b %>%
-  mutate(o2.bottom = replace(o2.bottom, 
-                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
-                                              VISIT_NO == 1,
-                                              DEPTH == 1.3) %>%
-                             select(OXYGEN.M) %>% pull()),
-         temp.bottom = replace(temp.bottom, 
-                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
-                                              VISIT_NO == 1,
-                                              DEPTH == 1.3) %>%
-                             select(TEMPERATURE.M) %>% pull()))
-```
-
-No surface temp at NLA17_NE-10060, visit 1 because no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_NE-10060", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.surf = replace(temp.surf, 
-                           SITE_ID == "NLA17_NE-10060" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NE-10060",
-                                              VISIT_NO == 1,
-                                              DEPTH == 0.1) %>%
-                             select(TEMPERATURE.M) %>% pull()))
-```
-
-No bottom temp or O2 at NLA17_NH-10063, visit 1 because no measurement taken at greatest depth (8m). Replace with next deepest measurement (7.5m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_NH-10063", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.bottom = replace(temp.bottom, 
-                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
-                                              VISIT_NO == 1,
-                                              DEPTH == 7.5) %>%
-                             select(TEMPERATURE.M) %>% pull()),
-         o2.bottom = replace(o2.bottom, 
-                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
-                                              VISIT_NO == 1,
-                                              DEPTH == 7.5) %>%
-                             select(OXYGEN.M) %>% pull()))
-```
-
-No surface temp or O2 at NLA17_OR-10003, visit 1 because no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_OR-10003", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.surf = replace(temp.surf, 
-                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
-                                              VISIT_NO == 1,
-                                              DEPTH == 0.1) %>%
-                             select(TEMPERATURE.M) %>% pull()),
-         o2.surf = replace(o2.surf, 
-                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
-                                              VISIT_NO == 1,
-                                              DEPTH == 0.1) %>%
-                             select(OXYGEN.M) %>% pull()))
-```
-
-No bottom temp or O2 at NLA17_SD-10057, visit 1 because no measurement taken at deepest depth (4m). Replace with next deepest measurement (3m)
-
-```{r}
-# look at raw data
-temp.o2.2 %>% filter(SITE_ID == "NLA17_SD-10057", VISIT_NO == 1)
-
-o2.s.b <- o2.s.b %>%
-  mutate(temp.bottom = replace(temp.bottom, 
-                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
-                                              VISIT_NO == 1,
-                                              DEPTH == 3) %>%
-                             select(TEMPERATURE.M) %>% pull()),
-         o2.bottom = replace(o2.bottom, 
-                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
-                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
-                                              VISIT_NO == 1,
-                                              DEPTH == 3) %>%
-                             select(OXYGEN.M) %>% pull()))
-```
-
-Only remaining missing values are observations where field crews noted sonde malfunctions.
-
-```{r}
-o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
-```
-
-Merge bottom and surface DO and temperature with dg data.
-
-```{r}
-# are all SITE_ID values in o2 and temp data also present in dg?
-o2.s.b %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007
-print(n= Inf)
-
-# are all site.id values in dg also present in o2 and temp data?
-# Just CA site, as expected.
-dg %>% filter(!(site.id %in% o2.s.b$SITE_ID)) %>% # no
-  select(site.id, visit.no)
-
-# dim(o2.s.b) # 1209, 6
-# dim(dg) # 2369, 375
-
-# Left join: retain all dg values, exclude non-matching bf values (i.e. OR and AZ)
-dg <- left_join(dg, o2.s.b, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
-# dim(dg) # 2369, 379, good!
-```
-
-## Fix chemistry
-The file contains `r dg %>% dplyr::select(nitrate.n.result) %>% filter(is.na(nitrate.n.result)) %>% summarize(na.no3 = nrow(.)) %>% pull()` missing nitrate values.
-Inspection of raw data sheet indicates that:
-* North Dakota reported NITRATE_NITRITE_N_RESULT rather than NITRATE.N.RESULT.  Most of the ND sites were 'non detect (ND)' with a detection limit of 0.03 mg /L.
-* WI sites were ND with a detection limit of 0.019.
-
-Will replace these values with NITRATE_NITRITE_N_RESULT when available, else 1/2 detection limit.
-
-``` {r}
-dg <- dg %>% 
-  mutate(nitrate.n.result = replace(nitrate.n.result, 
-                                    grepl("WI", site.id) & is.na(nitrate.n.result), # WI na sites
-                                    0.019/2), #1/2 detection limit
-         # ND sites without nitrate.n and nitrate.nitrite
-         nitrate.n.result = replace(nitrate.n.result, 
-                                    grepl("ND", site.id) & 
-                                      is.na(nitrate.n.result) & 
-                                      is.na(nitrate.nitrite.n.result), 
-                                    0.03/2), # 1/2 detection limit
-# ND sites with a nitrate.nitrite.n.result
-         nitrate.n.result = replace(nitrate.n.result, 
-                                    grepl("ND", site.id) & 
-                                      is.na(nitrate.n.result) & 
-                                      !is.na(nitrate.nitrite.n.result), # ND na sites
-                                    0.03)) # two sites reporting 0.03 mg/L
-```
-
-The file contains `r dg %>% filter(is.na(chla.result.vol)) %>% summarize(na.chl = nrow(.)) %>% pull()` missing chlorophyll values.
-Inspection reveals that these are all from North Dakota sites.
-It looks like the lab failed to record the volume filtered, so we can't calculate ug/L chl from these sites.
-These will remain missing.
-
 ## Write out data
@@
                        "inputData/dg.", Sys.Date(), ".RData"))
+
 ```

### 2020-10-08 scripts/dataMunge.Rmd
Updated input data file.

@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2020-10-07.txt"),
+                               "inputData/nla17gasDataAggregated_2020-10-08.txt"),
                  header = TRUE, sep = "\t", as.is = TRUE) 

### 2020-10-15 scripts/dgIndicatorAnalysis.Rmd
I specified getting the latest dataframe.  Cleaned names and filtered out AK sites.  Specified projection for geopackage.

--- 2020-06-01 scripts/dgIndicatorAnalysis.Rmd
@@
 ```{r}
-# Read data file
-load(paste0(localPath, 
-            "/Environmental Protection Agency (EPA)/",
-            "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.2020-02-25.RData"))
+# Read data file commented out b/c dataMunge.Rmd creates dg assignment based on latest text file nla17gasDataAggregated_2020-10-08.txt
+ # load(paste0(localPath, 
+ #            "/Environmental Protection Agency (EPA)/",
+ #            "ORD NLA17 Dissolved Gas - Documents/",
+ #            "inputData/dg.2020-10-09.RData"))
 
@@
   {summary(.)}
+
+dg %>% filter(WSA9 != "AK") %>%
+  select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
+  {summary(.)}
+
+dg %>% filter(WSA9 != "AK") %>% 
+  ggplot(aes(as.factor(x = WSA9), y = log(dissolved.n2o))) + geom_boxplot()
+
+dg %>% filter(WSA9 != "AK") %>% 
+  ggplot(aes(as.factor(x = WSA9), y = (n2o.sat.ratio))) + geom_boxplot()
 ```
@@
 names(dg_1)
-dg_1 <- clean_names(dg_1) # underscores are "safer" names used in GIS
+dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") # underscores are "safer" names used in GIS
 names(dg_1)
@@
                   crs = 4269) %>% # standard for lat/lon
-  st_transform(4326) # 4326 is just the EPSG identifier of WGS84.
+  st_transform(5070) # 4326 is just the EPSG identifier of WGS84.
 WGS84 comprises a standard coordinate frame for the Earth, a datum/reference ellipsoid for raw altitude data, and a gravitational equipotential surface (the geoid) that defines the nominal sea level.
+5070 is for Albers equal Area.
 st_crs(dg_sf1)
 
-dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1185 obs.
+dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1219 obs vs. previously 1185 obs., why change in sample size
 st_crs(dg_sf1)
@@
 names(ecoR)
-
-ecoR1 <- st_transform(ecoR, 4326) %>% 
+# changed from 4326 to 5070
+ecoR1 <- st_transform(ecoR, 5070) %>% 
   select(OBJECTID, WSA9, WSA9_NAME) # Grab variables of interest
@@
 # https://github.com/r-spatial/sf/issues/392
+# creating a _v1.gpkg having large sample size and should be in Albers
 st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE)
@@
 # https://gis.stackexchange.com/questions/223240/writing-multiple-layers-to-geopackage-using-writeogr-in-r
-st_write(ecoR1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "ecoR1", driver = "GPKG", update = TRUE, quiet = FALSE)
+st_write(ecoR1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "ecoR1", driver = "GPKG", append = TRUE, quiet = FALSE)
 st_layers("nladg_v1.gpkg")
+#CPL polygon not appearing in geopackage
 

### 2020-11-30 scripts/CDF_Estimates_Jb.Rmd
Added N2O flux calculation to dataMunge

---
title: "Cumulative Distribution Function (CDF) Estimates for the NLA 2017 Survey"
author: "Tom Kincaid originally, modified by Jake Beaulieu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

## Purpose
Calculate CDF estimates for the NLA 2017 survey.

## History
Created: November 5, 2019

Revised: December 12, 2019 to calculate estimates using the SRS variance
         estimator in addition to the local mean variance estimator

         November 2020, expanded to include estimates of N2O emission rates.

```{r, echo=FALSE}

# Attach the Hmisc and spsurvey packages
library(Hmisc)
library(spsurvey)

# Print the date and session information
cat(date(), "\n\n")
cat("Session Information:\n\n")
sessionInfo()

# Read the file containing data for CDF estimates
#cdf <- read.csv("L:/Priv/ARM Data/NLA 2017/Data Prep/Dissolved Gases/NLA17_DissolvedGases_forPopEst.csv")

# modified file to include N2O emission rates.  See "Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" for details
# updated file path
cdf <- read.csv(file = paste0(localPath, 
                              "/Environmental Protection Agency (EPA)/",
                              "ORD NLA17 Dissolved Gas - Documents/",
                              "inputData/populationEstimates/",
                              "NLA17_DissolvedGases_forPopEst.JB.csv"))
nr <- nrow(cdf)

# Reorder levels for the lake size class (AREA_CAT6) variable
levels(cdf$AREA_CAT6) <- list(
   "(1,4]" = "(1,4]",
   "(4,10]" = "(4,10]",
   "(10,20]" = "(10,20]",
   "(20,50]" = "(20,50]",
   ">50" = ">50")

# Reorder levels for the lake U.S. Forest Service (FS_EW) variable
levels(cdf$FS_EW) <- list(
   "East" = "East",
   "West" = "West",
   "Not_USFS" = "Not_USFS")

# Print the initial six lines in the cdf data frame
cat("\n\nThe initial six lines in the cdf data frame follow:\n\n")
head(cdf)

# Print the final six lines in the survey cdf data frame
cat("\n\nThe final six lines in the cdf data frame follow:\n\n")
tail(cdf)

# Print summary information for the response variables
cat("\n\nSummary Information for Response Variables\n")
temp <- match(c("CH4_SAT_RATIO", "CO2_SAT_RATIO", "DELTA_13_CH4_C",
   "DELTA_13_CO2_C", "DISSOLVED_CH4", "DISSOLVED_CO2", "DISSOLVED_N2O",
    "N2O_SAT_RATIO", "F_N2O", "F_N2O_D"), names(cdf))
for(c in temp) {
   cat("\n")
   print(describe(cdf[, c], descript=names(cdf)[c]))
}

#
# Calculate CDF estimates using the local mean variance estimator
#

# Create the sites data frame, which identifies sites to use in the analysis
sites <- data.frame(siteID=cdf$SITE_ID,
                    Use=rep(TRUE, nr))

# Create the subpop data frame, which defines populations and subpopulations for
# which estimates are desired
subpop <- data.frame(siteID=cdf$SITE_ID,
                     National=rep("National", nr),
                     LAKE_ORGN=cdf$LAKE_ORGN,
                     AREA_CAT6=cdf$AREA_CAT6,
                     #  > 1 hour with all variables and subpopulations
                     # AG_ECO3_NM=cdf$AG_ECO3_NM,
                      AG_ECO9_NM=cdf$AG_ECO9_NM
                     # EPA_REG=cdf$EPA_REG,
                     # MAJ_BAS_NM=cdf$MAJ_BAS_NM,
                     # HUC2_NM=cdf$HUC2_NM,
                     # OWN_NARS=cdf$OWN_NARS,
                     # FS_EW=cdf$FS_EW
)

# Create the design data frame, which identifies the weight, x-coordinate, and
# y-coordinate for each site ID
design <- data.frame(siteID=cdf$SITE_ID,
                     wgt=cdf$WGT_TP,
                     xcoord=cdf$XCOORD,
                     ycoord=cdf$YCOORD)

# Create the data.cont data frame, which specifies the variables to use in the
# analysis.  > 1 hour with all variables.
data.cont <- data.frame(siteID=cdf$SITE_ID,
                        # CH4_SAT_RATIO=cdf$CH4_SAT_RATIO,
                        # CO2_SAT_RATIO=cdf$CO2_SAT_RATIO,
                        # DELTA_13_CH4_C=cdf$DELTA_13_CH4_C,
                        # DELTA_13_CO2_C=cdf$DELTA_13_CO2_C,
                        # DISSOLVED_CH4=cdf$DISSOLVED_CH4,
                        # DISSOLVED_CO2=cdf$DISSOLVED_CO2,
                        # DISSOLVED_N2O=cdf$DISSOLVED_N2O,
                        # N2O_SAT_RATIO=cdf$N2O_SAT_RATIO,
                        F_N2O = cdf$F_N2O) # diffusive N2O nmol N2O m-2 day-1

```

```{r, echo=FALSE, results="hide"}

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the local mean variance estimator\n")
if(exists("warn.df")) rm("warn.df")
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont) # < 1 hour with all variables. 1 minute with F_N2O

```

```{r, echo=FALSE}

# Check for warning messages and print them if any exist
if(exists("warn.df")) {
   cat("\nWarning messages generated during the call to cat.analysis:\n")
   warnprnt()
} else {
   cat("\nNo warning messages were generated during the call to cat.analysis.\n")
}

# # Write CDF estimates as a comma-separated value (csv) file
# eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_Local_Mean_",
#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
# 
# # Create a PDF file containing plots of the CDF estimates
# eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_Local_Mean_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))

# Write percentile estimates as a csv file
write.csv(CDF_Estimates$Pct, 
          paste0(localPath, 
                 "/Environmental Protection Agency (EPA)/",
                 "ORD NLA17 Dissolved Gas - Documents/",
                 "inputData/populationEstimates/",
                 "NLA_2017_Percentile_Estimates_Local_Mean_F_N2O_JB",
                 gsub("-", "", Sys.Date()), ".csv"),
          row.names=FALSE)

```

#
# Calculate CDF estimates using the SRS variance estimator
#

```{r, echo=FALSE, results="hide"}

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the SRS variance estimator\n")
if(exists("warn.df")) rm("warn.df")
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont,
   vartype = "SRS")

```

```{r, echo=FALSE}

# Check for warning messages and print them if any exist
if(exists("warn.df")) {
   cat("\nWarning messages generated during the call to cat.analysis:\n")
   warnprnt()
} else {
   cat("\nNo warning messages were generated during the call to cat.analysis.\n")
}

# Write CDF estimates as a comma-separated value (csv) file
eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_SRS_",
   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

# Create a PDF file containing plots of the CDF estimates
eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_SRS_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))

# Write percentile estimates as a csv file
eval(parse(text = paste0("write.csv(CDF_Estimates$Pct, 'NLA_2017_Percentile_Estimates_SRS_",
   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

```

### 2020-11-30 scripts/dataMunge.Rmd
Added N2O flux calculation to dataMunge

@@
 
-The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.
+The dissolved gas data were curated in the 'NLA' dissolved gas project in RStudio.
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-10-07.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-11-10.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
 When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2020-10-08.txt"),
-                 header = TRUE, sep = "\t", as.is = TRUE) 
+                               "inputData/nla17gasDataAggregated_2020-11-10.txt"),
+                 header = TRUE, sep = "\t", as.is = TRUE) %>%
+  filter(sample.source == "DG", # pull out dissolved gas and leave out gas samples
+         !grepl("AK", site.id)) # omit alaska sites
 ```
@@
 ## Manipulate
-Implement unit conversion and create source/sink column:
+### Implement unit conversion and create source/sink column:
 ``` {r}
@@
   # unit conversion
-  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
-         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
-         dissolved.n2o = dissolved.n2o * 1000000000, # mol/L -> nmol/L
+  mutate(dissolved.ch4.umol = dissolved.ch4 * 1000000, # mol/L -> umol/L
+         dissolved.co2.umol = dissolved.co2 * 1000000, # mol/L -> umol/L,
+         dissolved.n2o.nmol = dissolved.n2o * 1000000000, # mol/L -> nmol/L
          # add source/sink column
@@
          n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
+```
+
+### Calculate emission rate. 
+The emission rate is the mass flux of a gas into or out of a waterbody.
+Emission rates are typically expressed as mass m^-2^ day^-1^.
+The phrase 'emission rate' is used interchangeably with 'Flux rate' and is often reported in the literature as (F~gas~), where the subscript 'gas' refers to the gas of interest (i.e.
+N~2~O, CO~2~, CH~4~).
+Emission rate can be  calculated as the product of gas^\*^ and the gas transfer velocity (k):
+
+F~gas~ = gas^\*^ * k
+
+where gas^\*^ is the difference between the observed (gas~obs~) and equilibrium (gas~eq~) dissolved gas concentration:
+
+gas^\*\ = (gas~obs~) - (gas~eq~)
+
+where a positive value indicates the waterbody is a source of the gas and a negative value indicates the waterbody is a sink.
+
+The gas transfer velocity (k; cm h^-1^) is a measure of the physical interaction between the waterbody and the atmosphere.
+Wavy and turbulent waterbodies have high gas transfer velocities whereas calm waterbodies have low values.
+k was not measured in the 2017 National Lakes Assessment, but was estimated from wind speed and lake area (Vachon, D., and Y.
+T.
+Prairie (2013), The ecosystem size and shape dependence of gas transfer velocity versus wind speed relationships in lakes, Can.
+J.
+Fish.
+Aquat.
+Sci., 70(12), 1757-1764, doi:10.1139/cjfas-2013-0241):
+
+k~600~ = 2.51 + 1.48 * U~10~ + 0.39 * U~10~ * log~10~Lake area
+
+where k~600~ is the gas transfer velocity normalized to CO~2~ at 20 ^o^C, U10 is wind speed 10m above the water surface.
+Wind speed was obtained on a 7.5km grid from the ERA-5 Land database and overlaid on the NLA sampling points.
+U~10~ was averaged between sunrise and sunset for each site, assuming that daytime wind conditions are best matched with the dissolved gas measurements which were conducted during the day.
+
+k~600~ is a standardized value that must be corrected for the differences in diffusivity among gases and water temperature at the sampling sites.
+These corrections will be implemented at a later date and emission rates will be calculated using k600 for now.
+
+```{r}
+dg <- dg %>% 
+  mutate(
+    # f.ch4 umol CH4/m2/d 
+    f.ch4 = ((dissolved.ch4 - (sat.ch4 * 1000000))*1000) * # 1000000 mol -> umol.  1000 L to m3
+           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
+    # f.co2 umol CO2/m2/d
+    f.co2 = ((dissolved.co2 - (sat.co2 * 1000000))*1000) * # 1000000 mol -> umol.  1000 L to m3
+           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
+        # f.n2o nmol N2O/m2/d
+    f.n2o = ((dissolved.n2o - (sat.n2o * 1000000000))*1000) * # 1000000 mol -> nmol.  1000 L to m3
+           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
+    # total flux of CH4 per day kmol CH4 day-1
+    f.ch4.d = f.ch4 * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
+    # total flux of CO2 per day kmol CO2 day-1
+    f.co2.d = f.co2 * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
+    # total flux of N2O per day mol N2O day-1
+    f.n2o.d = f.n2o * area.ha * 10000 * (1/10^9)) %>% # 1ha = 10,000 m2. 10^9 nmol->umol->mmol->mol
+         filter(state != "AK", sample.source == "DG")
+
+```
+
+## Emission rate for population estimate
+Tom Kincaid created a file to be used for population estimates.  Lets add our emission rates estimates to this file.  We can then use Tom's code to construct population estimates of emission rates.
+```{r}
+cdf <- read.csv(paste0(localPath,
+            "/Environmental Protection Agency (EPA)/",
+            "ORD NLA17 Dissolved Gas - Documents/",
+            "inputData/populationEstimates/",
+            "NLA17_DissolvedGases_forPopEst.csv"))
+
+# although data I provided Tom (nla17gasAggregatedForPopEst2019-10-23.txt) included data from both visits,
+# data used for population estimates included only visit one.  I don't know rationale for this, but will
+# do the same with the revised file here.
+unique(cdf$VISIT_NO) # 1
+dg.1 <- dg %>% filter(visit.no == 1)
+
+# even after filtering, dg has 85 more observations than cdf
+dim(cdf) # 1005, 81
+dim(dg.1) # 1090, 396
+
+# which observations in filtered dg, not in cdf?
+# many of these are handpicked (HP) sites.  Others not sure
+# why are excluded
+dg.1 %>% filter(!(dg.1$site.id %in% cdf$SITE_ID)) %>%
+  select(site.id) %>%
+  as_tibble() %>% print(n=Inf)
+
+# using left_join to exclude sites from dg.1 that Tom did
+# not include in cdf (see above).  Presumably the excluded sites
+# are not probability sites
+cdf.dg <- left_join(cdf, 
+                    select(dg.1, site.id, visit.no, f.ch4, f.co2, 
+                           f.n2o, f.ch4.d, f.co2.d, f.n2o.d) %>%
+                      # rename for consistency with Tom's data
+                      rename_all(~stringr::str_to_upper(.)) %>% 
+                      rename_all(~stringr::str_replace_all(., '\\.', '_')))
+
+dim(cdf.dg) # 1005, 87 same as cdf above, so good
+
+write.csv(cdf.dg, 
+          paste0(localPath,
+            "/Environmental Protection Agency (EPA)/",
+            "ORD NLA17 Dissolved Gas - Documents/",
+            "inputData/populationEstimates/",
+            "NLA17_DissolvedGases_forPopEst.jb.csv"))
+
 ```

### 2020-11-30 scripts/dgIndicatorAnalysis.Rmd
Added N2O flux calculation to dataMunge

@@
 ```{r}
-# Read data file commented out b/c dataMunge.Rmd creates dg assignment based on latest text file nla17gasDataAggregated_2020-10-08.txt
- # load(paste0(localPath, 
- #            "/Environmental Protection Agency (EPA)/",
- #            "ORD NLA17 Dissolved Gas - Documents/",
- #            "inputData/dg.2020-10-09.RData"))
-
-# Strip out air samples
-dg <- dg %>% filter(sample.source == "DG")
+# Read data file 
+ load(paste0(localPath,
+            "/Environmental Protection Agency (EPA)/",
+            "ORD NLA17 Dissolved Gas - Documents/",
+            "inputData/dg.2020-11-25.RData"))
 
@@
 dg %>% 
-  select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
+  select(dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
   {summary(.)}
 
-dg %>% filter(WSA9 != "AK") %>%
-  select(dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
-  {summary(.)}
-
-dg %>% filter(WSA9 != "AK") %>% 
-  ggplot(aes(as.factor(x = WSA9), y = log(dissolved.n2o))) + geom_boxplot()
-
-dg %>% filter(WSA9 != "AK") %>% 
+dg %>% 
+  ggplot(aes(as.factor(x = WSA9), y = log(dissolved.n2o.nmol))) + geom_boxplot()
+
+dg %>% 
   ggplot(aes(as.factor(x = WSA9), y = (n2o.sat.ratio))) + geom_boxplot()
@@
 The survey design allows for population estimates of central tendency and variance by ecoregion (and other subpopulations), which can provide insight into spatial patterning.
-There is broad overlap in mean +/- 95% CI estimates for dissolved CH4 concentation.
-The Northern Plains, Temperate Plains, and Western Mountains have low dissolved CO2 concentration.
+There is broad overlap in mean +/- 95% CI estimates for dissolved CH~4~ concentration.
+The Northern Plains, Temperate Plains, and Western Mountains have low dissolved CO~2~ concentration.
 
 ```{r, fig.cap= "Fig. caption: Mean +/- 95% CI for the dissolved gas concentration indicator"}
-# Read population estimates provided by Steve Paulsen (11/7/2019)
+# Read population estimates calculated with local neighborhood variance estimate provided by Tom Kincaid (12/12/2019)
 pop <- read.csv(file = paste0(localPath, 
@@
                               "inputData/populationEstimates/",
-                              "NLA_2017_Percentile_Estimates_20191107.csv"),
+                              "NLA_2017_Percentile_Estimates_Local_Mean_20191212.csv"),
                 as.is = TRUE) %>%
@@
 
+# Read in file containing population estimates for N2O flux (nmol N2O m-2 day-1).
+# This file was generated by running Tom Kincaid's code in scripts/CDF_Estimates_jb.Rmd
+pop.n2o <- read.csv(file = paste0(localPath, 
+                              "/Environmental Protection Agency (EPA)/",
+                              "ORD NLA17 Dissolved Gas - Documents/",
+                              "inputData/populationEstimates/",
+                              "NLA_2017_Percentile_Estimates_Local_Mean_F_N2O_JB20201119.csv"),
+                as.is = TRUE) %>%
+  toEPA()  # enforce formatting convention
+
+# extract population estimates for ecoregion and nation
+pop.n2o <- pop.n2o %>% 
+  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
+         statistic == "Mean") %>% # extract mean estimate
+   mutate(subpopulation = factor(subpopulation, 
+                             levels = filter(., subpopulation != "National") %>%
+                               arrange(estimate) %>%
+                               select(subpopulation) %>%
+                               pull() %>%
+                               c("National", .)))
+
+# look at dissolved concentration indicator
+ggplot(pop.n2o, aes(estimate, subpopulation)) +
+  geom_point() +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct)) +
+  facet_wrap(~indicator, scales = "free_x") +
+  theme(axis.title = element_blank())
+
+```
+
+Population level estimates of variance can be calculated using a simple random sample or local neighborhood estimators.
+The local neighborhood estimator will yield lower variance estimates if there is spatial structure in the data.
+Below are plotted the 95% confidence intervals calculated using both methods.
+In nearly all cases the local neighborhood estimator yields a more restricted interval.
+
+```{r}
+# calculate 95% CI using local neighborhood estimator
+pop.eco.diss <- pop.eco.diss %>%
+  mutate(ci95.local = ucb95pct - lcb95pct)
+
+# read in SRS estimator population estimates provided by Tom Kincaid (12/12/2019)
+pop.srs <- read.csv(file = paste0(localPath, 
+                              "/Environmental Protection Agency (EPA)/",
+                              "ORD NLA17 Dissolved Gas - Documents/",
+                              "inputData/populationEstimates/",
+                              "NLA_2017_Percentile_Estimates_SRS_20191212.csv"),
+                as.is = TRUE) %>%
+  toEPA()  # enforce formatting convention
+
+# Convert units for dissolved gas concentration indicator
+pop.srs <- pop.srs %>%
+  mutate(estimate = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
+                            estimate * 1000000, # mol/L -> umol/L for co2 and ch4
+                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
+                            estimate * 1000000000, # mol/L -> nmol/L for n2o
+                            estimate)), # all others
+         lcb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
+                            lcb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
+                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
+                            lcb95pct * 1000000000, # mol/L -> nmol/L for n2o
+                            lcb95pct)), # all others
+         ucb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
+                            ucb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
+                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
+                            ucb95pct * 1000000000, # mol/L -> nmol/L for n2o
+                            ucb95pct))) # all others
+
+# extract population estimates for ecoregion and nation
+# dissolved gas indicator
+pop.srs.eco.diss <- pop.srs %>% 
+  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
+         statistic == "Mean") %>% # extract mean estimate
+  mutate(ci95.srs = ucb95pct - lcb95pct) %>%
+  filter(grepl("dissolved", indicator, ignore.case = TRUE))
+
+# merge local neighborhood and SRS variance estimates
+pop.local.srs <- dplyr::full_join(pop.eco.diss,
+                                  select(pop.srs.eco.diss, -nresp, -estimate, 
+                                         -stderror, -lcb95pct, -ucb95pct)) %>%
+  select(-nresp, -estimate, -stderror, -lcb95pct, -ucb95pct) %>%
+  pivot_longer(cols = starts_with("ci95"))
+
+# look at dissolved concentration indicator
+ggplot(pop.local.srs, aes(value, subpopulation)) +
+  geom_point(aes(color = name)) +
+  facet_wrap(~indicator, scales = "free_x") +
+  xlab("95% confidence interval of mean") +
+  theme(axis.title.y = element_blank())
 ```

### 2020-12-10 scripts/dgIndicatorAnalysis.Rmd
Amending layer in geopackage so need to use delete_layer = TRUE, append = TRUE

--- 2020-10-15 scripts/dgIndicatorAnalysis.Rmd
@@
 
-dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2, dissolved.ch4, dissolved.n2o, WSA9, WSA9_NAME)
+dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2, dissolved.ch4, dissolved.n2o, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio, WSA9, WSA9_NAME)
 names(dg_1)
-dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") # underscores are "safer" names used in GIS
+dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") %>% 
+  mutate(log_n2o = round(log(dissolved_n2o),3)) %>% 
+  mutate(log_n2osatratio = round(log(n2o_sat_ratio),3))
+# underscores are "safer" names used in GIS
+
 names(dg_1)
@@
 
-dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1219 obs vs. previously 1185 obs., why change in sample size
+dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1184 obs vs. previously 1185 obs., why change in sample size
 st_crs(dg_sf1)
@@
 st_layers("nladg_v1.gpkg")
+
+# N.B. b/c old layer with untransformed was already in geodatabase have to delete it and then append the new layer
+st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", delete_layer = TRUE, append = TRUE, quiet = FALSE)
+st_layers("nla_dg1.gpkg")
 

### 2020-12-11 scripts/dgIndicatorAnalysis.Rmd
Added units to names of some columns, adopted convention of e.gas. for areal emission rate and f.gas for total flux in/out of lake (mass gas/day), added script for spSurvey population estimates (inlcuding extent estimate of N2O undersaturation).

--- 2020-11-30 scripts/dgIndicatorAnalysis.Rmd
@@
             "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.2020-11-25.RData"))
+            "inputData/dg.2020-12-02.RData"))
 
@@
 ## Population estimates by ecoregion
+### Dissolved gas concentration
 The survey design allows for population estimates of central tendency and variance by ecoregion (and other subpopulations), which can provide insight into spatial patterning.
-There is broad overlap in mean +/- 95% CI estimates for dissolved CH~4~ concentration.
+There is broad overlap in mean +/- 95% CI estimates for dissolved CH~4~ and N~2~O concentration across ecoregions.
 The Northern Plains, Temperate Plains, and Western Mountains have low dissolved CO~2~ concentration.
@@
 ```{r, fig.cap= "Fig. caption: Mean +/- 95% CI for the dissolved gas concentration indicator"}
-# Read population estimates calculated with local neighborhood variance estimate provided by Tom Kincaid (12/12/2019)
+# Read population estimates calculated with local neighborhood variance estimate.  Code provided by Tom Kincaid 
+# (12/12/2019), then modified to accomdate the gas emission rate and flux variables.  See CDF_Estimates_Jb.Rmd.
 pop <- read.csv(file = paste0(localPath, 
@@
                               "inputData/populationEstimates/",
-                              "NLA_2017_Percentile_Estimates_Local_Mean_20191212.csv"),
+                              "NLA_2017_Percentile_Estimates_Local_Mean_JB20201202.csv"),
                 as.is = TRUE) %>%
@@
   theme(axis.title = element_blank())
-
-# Read in file containing population estimates for N2O flux (nmol N2O m-2 day-1).
-# This file was generated by running Tom Kincaid's code in scripts/CDF_Estimates_jb.Rmd
-pop.n2o <- read.csv(file = paste0(localPath, 
+```
+
+### Dissolved gas saturation ratios and source/sink status
+Dissolved gas saturation ratios are calculated as the ratio of the observed to equilibrium concentration.
+The equilibrium concentration (sometimes called the saturation concentration) is the concentration expected if the waterbody were in equilibrium with the atmosphere.
+It is calculated from the partial pressure of the gas in the atmosphere, water temperature, and the gas specific solubility.
+Saturation ratios >1 indicate the water body is supersaturated and is a source of the gas to the atmosphere.
+Saturation ratios <1 indicate that the waterbody is undersaturated and is consuming the gas from the atmosphere.
+The saturation ratio can therefore be used to categorize each waterbody as a 'sink' or a 'source'.
+
+``` {r}
+# read in source/sink population data
+# Read extent estimates generated with CDF_Estimates_Jb.Rmd.
+File provided by Tom Kincaid (12/12/2019), then modified to accomodate the gas emission rate, flux variables, and cat.analysis for N2O source/sink.
+n2o.extent <- read.csv(file = paste0(localPath, 
                                      "/Environmental Protection Agency (EPA)/",
@@
                                      "inputData/populationEstimates/",
-                              "NLA_2017_Percentile_Estimates_Local_Mean_F_N2O_JB20201119.csv"),
+                                     "NLA_2017_Extent_Estimates20201203.csv"),
                        as.is = TRUE) %>%
-  toEPA()  # enforce formatting convention
-
+  toEPA() %>% # enforce formatting convention
+  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
+         category == "sink") %>% # extract extent estimate for sink
+  mutate(subpopulation = factor(subpopulation, # arrange by extent estimate for sink
+                                levels = filter(., subpopulation != "National") %>%
+                                  arrange(estimate.p) %>%
+                                  select(subpopulation) %>%
+                                  pull() %>%
+                                  c("National", .)))
+```
+
+The binary 'source' or 'sink' values can be analyzed at the population level using the cat.analysis function in the spsurvey package.
+The function provides an 'extent' estimate, which is the proportion of waterbodies in each region that are functioning as a source or sink.
+Over half of the waterbodies are functioning as N2O sinks in all ecoregions.
+At the national scale, `r n2o.extent %>% filter(type == "National") %>% select(estimate.p) %>% pull() %>% round(., 1)`% (95% CI: `r n2o.extent %>% filter(type == "National") %>% select(lcb95pct.p) %>% pull() %>% round(., 1)` - `r n2o.extent %>% filter(type == "National") %>% select(ucb95pct.p) %>% pull() %>% round(., 1)`) of lakes and reservoirs were net N~2~O sinks during the index period, suggesting that US lakes and reservoirs are a net N~2~O sink.
+However, the 95% confidence intervals of the mean N~2~O saturation ratio include 1 for all ecoregions except the Northern plains which are undersaturated.
+At the national scale, the mean N~2~O saturation ratio indicates supersaturation (`r pop.eco %>% filter(type == "National", statistic == "Mean", indicator == "N2O_SAT_RATIO") %>% select(estimate) %>% pull() %>% round(., 2)`% (95% CI: `r pop.eco %>% filter(type == "National", statistic == "Mean", indicator == "N2O_SAT_RATIO") %>% select(lcb95pct) %>% pull() %>% round(., 2)` - `r pop.eco %>% filter(type == "National", statistic == "Mean", indicator == "N2O_SAT_RATIO") %>% select(ucb95pct) %>% pull() %>% round(., 2)`)), suggesting that US lakes and reservoirs function as a net N~2~O sink.
+These findings indicate that high levels of N~2~O saturation in a few waterbodies offset undersaturation in most waterbodies, yielding a mean saturation ratio > 1.
+
+```{r message=FALSE}
+
+## Bubble plot of observed N2O saturation index----------------
+# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple")
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+ggplot() +
+  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+  geom_sf(data = dg.sf, aes(size = n2o.sat.ratio, color = n2o.src.snk),
+          show.legend = "point") +
+  ggtitle("N2O saturation ratio") +
+  scale_color_manual(values = c("white", "black"), name = "source/sink") +
+  scale_size_binned(name = "N2O saturation ratio",
+             range = c(1, 8), # custom size range
+             breaks = c(0.1, 0.5, 1, 2))  # custom breaks
+
+## horizontal bar chart of N2O sink extent---------------
+ggplot(n2o.extent, aes(estimate.p, subpopulation)) +
+  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
+  geom_errorbarh(aes(xmax = ucb95pct.p, xmin = lcb95pct.p, color = subpopulation), height = 0, size = 1) +
+  xlab("Percent of waterbodies functioning as N2O sinks") +
+  theme(axis.title.y = element_blank())
+
+## horizontal bar chart of N2O saturation ratio-------------
+# extract saturation indicator, order by n2o extent (see above).  This enforces same order for both plots.
+pop.eco.sat <- pop.eco %>%
+  filter(grepl("sat", indicator, ignore.case = TRUE)) %>%
+  mutate(subpopulation = factor(subpopulation, 
+                             levels = filter(n2o.extent, subpopulation != "National") %>%
+                                  arrange(estimate.p) %>%
+                                  select(subpopulation) %>%
+                                  pull() %>%
+                               as.character() %>%
+                                  c("National", .)))
+
+# N2O saturation ratio; horizontal bar chart
+ggplot(filter(pop.eco.sat, indicator == "N2O_SAT_RATIO"), aes(estimate, subpopulation)) +
+  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
+  xlab("N2O saturation ratio") +
+  theme(axis.title.y = element_blank()) +
+  geom_vline(xintercept = 1)
+
+```
+
+### Emission rates and flux
+The emission rate (E~gas~) is the rate at which a gas crosses the air water interface and is expressed on an areal basis (i.e.
+mass m^-2^ day^-1^).
+It is calculated as a function of the dissolved gas concentration and gas transfer velocity (see dataMunge.Rmd).
+The 95% confidence interval of the mean N~2~O emission rate is quite large in several ecoregions, encompassing 0 in all but the Northen Plains and Xeric ecoregions.
+The mean N~2~O emission rate at the national scale is positive (`r  pop.eco %>% filter(indicator == "E_N2O_NMOL_D", subpopulation == "National", statistic == "Mean") %>% select(estimate) %>% pull() %>% round(., 1)/1000` umol N~2~O m^-2^ day^-1^), although the 95% confidence interval includes 0, similar to the N~2~O saturation ratio pattern.
+
+```{r, fig.cap= "Fig. caption: Mean +/- 95% CI for the emission rate (mass of GHG time-1 area-1)"}
 # extract population estimates for ecoregion and nation
-pop.n2o <- pop.n2o %>% 
-  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
-         statistic == "Mean") %>% # extract mean estimate
+pop.eco.emission <- pop.eco %>% 
+  filter(grepl("^E", indicator)) %>% # ^ means starts with.  Extract the emission rates 
+  # order by N2O
    mutate(subpopulation = factor(subpopulation, 
-                             levels = filter(., subpopulation != "National") %>%
+                             levels = filter(., indicator == "E_N2O_NMOL_D", subpopulation != "National") %>%
                                arrange(estimate) %>%
@@
 # look at dissolved concentration indicator
-ggplot(pop.n2o, aes(estimate, subpopulation)) +
-  geom_point() +
-  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct)) +
-  facet_wrap(~indicator, scales = "free_x") +
-  theme(axis.title = element_blank())
-
-```
-
+ggplot(filter(pop.eco.emission, indicator == "E_N2O_NMOL_D"), 
+       aes(estimate/1000, subpopulation)) + # convert to umol
+  geom_point(aes(color = subpopulation, fill = subpopulation), alpha = 0.25, size = 5) +
+  geom_errorbarh(aes(xmax = ucb95pct/1000, xmin = lcb95pct/1000, color = subpopulation), height = 0, size = 1) + # convert to umol
+  geom_vline(xintercept = 0) +
+  xlab("N2O emission rate (umol N2O m-2 day-1") + # converted from nmol to umol (see above)
+  theme(axis.title.y = element_blank())
+
+```
+
+The N~2~O flux at a waterbody is the total mass of N~2~O crossing the air-water interface per unit time (e.g.
+mass N~2~O day^-1^) and is calculated as the product of the emission rate (mass N~2~O m^-2^ day^-1^) and waterbody area (m^2^).
+Negative values indicate N2O invasion into the waterbody and positive values indicate N~2~O efflux from the waterbody to the atmosphere.
+An assessment of the role of lakes and reservoirs in the atmospheric N~2~O budget must be based on flux, rather than emission rates, because flux accounts for waterbody size.
+For example, N~2~O emissions from several small waterbodies could be completely offset by N~2~O uptake by a single large waterbody with a small negative emission rate (e.g.
+N~2~O sink), yet the mean emission rate for these waterbodies would be > 0.
+
+The mean N~2~O flux per waterbody is close to zero for most ecoregions and the 95% confidence interval includes zero for all ecoregions.
+The national scale mean N~2~O flux is slightly negative (`r  pop.eco %>% filter(indicator == "F_N2O_M_D", subpopulation == "National", statistic == "Mean") %>% select(estimate) %>% pull() %>% round(., 2)` mol N~2~O day^-1^), but the 95% CI includes 0.
+
+```{r, fig.cap= "Fig. caption: Mean +/- 95% CI of the GHG flux across the air-water interface (mass GHG / time)"}
+# extract population estimates for ecoregion and nation
+pop.eco.flux <- pop.eco %>% 
+  filter(grepl("^F", indicator)) %>% # ^ means starts with.  Extract the flux rates 
+  # order by N2O
+   mutate(subpopulation = factor(subpopulation, 
+                             levels = filter(., indicator == "F_N2O_M_D", subpopulation != "National") %>%
+                               arrange(estimate) %>%
+                               select(subpopulation) %>%
+                               pull() %>%
+                               c("National", .)))
+
+# look at dissolved concentration indicator
+ggplot(filter(pop.eco.flux, indicator == "F_N2O_M_D"),
+       aes(estimate, subpopulation)) +
+  geom_point(aes(color = subpopulation, fill = subpopulation), alpha = 0.25, size = 5) +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
+  geom_vline(xintercept = 0) +
+  xlab("N2O flux (mol N2O day^-1 waterbody^-1") +
+  theme(axis.title.y = element_blank())
+
+```
+
+### 13C paper
+```{r}
+# 13C; horizontal bar chart by ecoregion
+pop.eco.13c <- filter(pop.eco, indicator == "DELTA_13_CH4_C")
+ggplot(pop.eco.13c, aes(estimate, subpopulation)) +
+  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
+  theme(axis.title.y = element_blank())
+
+# 13C; horizontal bar chart by origin
+pop.origin.13c <- filter(pop, type == "LAKE_ORGN", statistic == "Mean", indicator == "DELTA_13_CH4_C")
+ggplot(pop.origin.13c, aes(estimate, subpopulation)) +
+  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
+  theme(axis.title.y = element_blank())
+
+# 13C; horizontal bar by size
+pop.size.ch4 <- filter(pop, type == "AREA_CAT6", statistic == "Mean", indicator == "DISSOLVED_CH4")
+ggplot(pop.size.ch4, aes(estimate, subpopulation)) +
+  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
+  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
+  theme(axis.title.y = element_blank())
+
+# 13CH4 Bubble plots
+
+ggplot() +
+  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+  geom_sf(data = dg.sf, aes(size = delta.13.ch4), 
+          show.legend = "point") + # improve legend
+  ggtitle("13CH4") +
+  scale_size(name = "13C-CH4",
+             range = c(0.05, 3), # custom size range
+             breaks = c(-80, -50, -25, 0)) # custom breaks
+
+```
+
+Similarly, we can look at population level estimates of GHG emission rates (mass of GHG time^-1^ area^-1^).
+
+We can also look at the GHG flux across the air-water interface (mass GHG time^-1^)
+
+## Simple random sample vs local neighborhood variance estimator
 Population level estimates of variance can be calculated using a simple random sample or local neighborhood estimators.
@@
 
-```{r}
+```{r, message=FALSE}
 # calculate 95% CI using local neighborhood estimator
@@
 The first bubble plot below is of dissolved CO2 which seems to show a pattern of higher concentrations along the eastern seaboard.
-```{r}
+```{r, message=FALSE}
 # To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
@@
 
-# "Temperate" is misspelled.
-ecoR <- ecoR %>% 
-  mutate(WSA9_NAME = as.character(WSA9_NAME), # conv to char
-         WSA9_NAME = ifelse(WSA9_NAME == "Temporate Plains",
-                            "Temperate Plains", # correct sp
-                            WSA9_NAME),
-         WSA9_NAME = as.factor(WSA9_NAME)) # back to factor
-
 # Check CRS
@@
 
-```{r}
+```{r, message=FALSE}
 # CO2 Bubble plots
@@
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.co2, color = co2.src.snk), 
+  geom_sf(data = dg.sf, aes(size = dissolved.co2.umol, color = co2.src.snk), 
           show.legend = "point") + # improve legend
@@
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.ch4, color = ch4.src.snk), 
+  geom_sf(data = dg.sf, aes(size = dissolved.ch4.umol, color = ch4.src.snk), 
           show.legend = "point") +
@@
 This paper made quite a splash.
-In our study, `r sum(dg$n2o.src.snk == "sink")` out of `r sum(!is.na(dg$dissolved.n2o))` sampled waterbodies were an N2O sink.
+In our study, `r sum(dg$n2o.src.snk == "sink")` out of `r sum(!is.na(dg$dissolved.n2o.nmol))` sampled waterbodies were an N2O sink.
 It is somewhat interesting to note that the some of the highest concentrations were observed in Indiana/Ohio, where I conducted much of my dissertation research on aquatic N2O emissions.
@@
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.n2o, color = n2o.src.snk),
+  geom_sf(data = dg.sf, aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
           show.legend = "point") +
@@
 
-We can also look at population estimates of N2O saturation ratio by ecoregion.
-Wow, despite the prevalance of N2O undersaturated waterbodies, the 95% CI for the mean N2O saturation ratio encompasses 1 in all ecoregions except the Northern Plains and Northern Appalachians.
-```{r}
-
-# extract saturation indicator, order by n2o
-pop.eco.sat <- pop.eco %>%
-  filter(grepl("sat", indicator, ignore.case = TRUE)) %>%
-  mutate(subpopulation = factor(subpopulation, 
-                             levels = filter(pop.eco, indicator == "N2O_SAT_RATIO", subpopulation != "National") %>%
-                               arrange(estimate) %>%
-                               select(subpopulation) %>%
-                               pull() %>%
-                               c("National", .)))
-
-# look at dissolved concentration indicator
-ggplot(pop.eco.sat, aes(estimate, subpopulation)) +
-  geom_point() +
-  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct)) +
-  facet_wrap(~indicator, scales = "free_x") +
-  theme(axis.title = element_blank()) +
-  geom_vline(xintercept = 1)
-
-```
-
 ## Geopackage
-Create a simple feature on entire dg data frame, then split out a simple feature, df_sf2, by filtering sample_source = DG.
-The geodatabase will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
+Create a simple feature on entire dg data frame.  The geopackage will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
 ```{r, geopackage, eval=FALSE, echo=FALSE,}
@@
 
-dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2, dissolved.ch4, dissolved.n2o, WSA9, WSA9_NAME)
+dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, WSA9, WSA9_NAME)
 names(dg_1)
@@
 st_crs(dg_sf1)
-
-dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1219 obs vs. previously 1185 obs., why change in sample size
-st_crs(dg_sf1)
+dim(dg_sf1) # 1184 observations
 
@@
 # creating a _v1.gpkg having large sample size and should be in Albers
-st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE)
+st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
+         delete_layer = TRUE) # needed, else function gives error if object already exists
 st_layers("nladg_v1.gpkg")
@@
 # https://gis.stackexchange.com/questions/223240/writing-multiple-layers-to-geopackage-using-writeogr-in-r
-st_write(ecoR1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "ecoR1", driver = "GPKG", append = TRUE, quiet = FALSE)
+st_write(ecoR1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "ecoR1", driver = "GPKG", append = TRUE, quiet = FALSE, delete_layer = TRUE) # needed, else function gives error if object already exists
 st_layers("nladg_v1.gpkg")
-#CPL polygon not appearing in geopackage
 
 names(dg_sf1)
-summary(dg_sf1$dissolved_n2o) # matches Roy's stats
+summary(dg_sf1$dissolved_n2o_nmol) # matches Roy's stats
 ```
@@
 # Add within-region summary statistics to summary table 
-for (r in 1:length(ecoRname1)){ # loop through regions and fill in the matrices
-  euc.reg <- euc[grep(ecoRname1[r], rownames(euc)), grep(ecoRname1[r],colnames(euc))] # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
-  # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
-  euc.summary[, (r+1)] <- round(c( min(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[2], median(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[4], max(euc.reg[euc.reg!=0]) ),1)
-}
+# for (r in 1:length(ecoRname1)){ # loop through regions and fill in the matrices
+#   euc.reg <- euc[grep(ecoRname1[r], rownames(euc)), grep(ecoRname1[r],colnames(euc))] # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
+#   # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
+#   euc.summary[, (r+1)] <- round(c( min(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[2], median(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[4], max(euc.reg[euc.reg!=0]) ),1)
+# }
 
@@
 
-# GitHub question:  if I save a workspace, .RData, will that be pushed when I push to GitHub?  Or, should data frames go in .gitignore?
+# GitHub question:  if I save a workspace, .RData, will that be pushed when I push to GitHub?  Or, should data frames go in #.gitignore?
+# Jake - we are using a public github repo under the EPA enterprise license.
+Anything we push to github can be downloaded by the public.
+We should use the shared documents library at our SharePoint site for data.
+Code can go to github.
 ```
@@
 class(cpl_sp)
-ggplot(cpl,aes(dissolved.ch4)) + geom_histogram()
-ggplot(cpl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
-
-hscat(dissolved.ch4~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
-
-hscat(dissolved.ch4~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
-
-ggplot(cpl,aes(dissolved.co2)) + geom_histogram()
-ggplot(cpl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
-
-hscat(dissolved.co2~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
-
-hscat(log10(dissolved.co2)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
-
-ggplot(cpl,aes(dissolved.n2o)) + geom_histogram()
-ggplot(cpl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
-
-hscat(dissolved.n2o~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
-
-hscat(log10(dissolved.n2o)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
-
-# nap
+ggplot(cpl,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(cpl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.ch4.umol~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.ch4.umol~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
+
+ggplot(cpl,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(cpl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.co2.umol~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(log10(dissolved.co2.umol)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
+
+ggplot(cpl,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(cpl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(dissolved.n2o.nmol~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))
+
+hscat(log10(dissolved.n2o.nmol)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))
+
+# nap # nap object doesn't exist
+nap <- dg.sf %>% filter(WSA9_NAME == "Northern Appalachians")
 nap_sp <- as(nap, "Spatial")
 class(nap_sp)
-ggplot(nap,aes(dissolved.ch4)) + geom_histogram()
-ggplot(nap,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(dissolved.ch4~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(log10(dissolved.ch4)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
-
-ggplot(nap,aes(dissolved.co2)) + geom_histogram()
-ggplot(nap,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(dissolved.co2~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(log10(dissolved.co2)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
-
-ggplot(nap,aes(dissolved.n2o)) + geom_histogram()
-ggplot(nap,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(dissolved.n2o~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(log10(dissolved.n2o)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.ch4.umol~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.ch4.umol)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(nap,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(nap,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.co2.umol~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.co2.umol)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(nap,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(nap,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.n2o.nmol~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.n2o.nmol)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
 
 # npl 
+npl <- dg.sf %>% filter(WSA9_NAME == "Northern Plains")
 npl_sp <- as(npl, "Spatial")
 class(npl_sp)
-ggplot(npl,aes(dissolved.ch4)) + geom_histogram()
-ggplot(npl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(dissolved.ch4~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(log10(dissolved.ch4)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
-
-ggplot(npl,aes(dissolved.co2)) + geom_histogram()
-ggplot(npl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(dissolved.co2~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(log10(dissolved.co2)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
-
-ggplot(npl,aes(dissolved.n2o)) + geom_histogram()
-ggplot(npl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(dissolved.n2o~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(log10(dissolved.n2o)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.ch4.umol~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.ch4.umol)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(npl,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(npl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.co2.umol~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.co2.umol)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(npl,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(npl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(dissolved.n2o.nmol~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.n2o.nmol)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
 
 # sap
+sap <- dg.sf %>% filter(WSA9_NAME == "Southern Appalachians")
 sap_sp <- as(sap, "Spatial")
 class(sap_sp)
-ggplot(sap,aes(dissolved.ch4)) + geom_histogram()
-ggplot(sap,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
-
-hscat(dissolved.ch4~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
-
-hscat(log10(dissolved.ch4)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
-
-ggplot(sap,aes(dissolved.co2)) + geom_histogram()
-ggplot(sap,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
-
-hscat(dissolved.co2~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
-
-hscat(log10(dissolved.co2)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
-
-ggplot(sap,aes(dissolved.n2o)) + geom_histogram()
-ggplot(sap,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
-
-hscat(dissolved.n2o~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
-
-hscat(log10(dissolved.n2o)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(dissolved.ch4.umol~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(log10(dissolved.ch4.umol)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(sap,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(sap,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(dissolved.co2.umol~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(log10(dissolved.co2.umol)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(sap,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(sap,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(dissolved.n2o.nmol~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))
+
+hscat(log10(dissolved.n2o.nmol)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
 
 # spl
+spl <- dg.sf %>% filter(WSA9_NAME == "Southern Plains")
 spl_sp <- as(spl, "Spatial")
 class(spl_sp)
-ggplot(spl,aes(dissolved.ch4)) + geom_histogram()
-ggplot(spl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
-
-hscat(dissolved.ch4~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
-
-hscat(log10(dissolved.ch4)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
-
-ggplot(spl,aes(dissolved.co2)) + geom_histogram()
-ggplot(spl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
-
-hscat(dissolved.co2~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
-
-hscat(log10(dissolved.co2)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
-
-ggplot(spl,aes(dissolved.n2o)) + geom_histogram()
-ggplot(spl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
-
-hscat(dissolved.n2o~1, spl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
-
-hscat(log10(dissolved.n2o)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(dissolved.ch4.umol~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(log10(dissolved.ch4.umol)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(spl,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(spl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(dissolved.co2.umol~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(log10(dissolved.co2.umol)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(spl,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(spl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))
+
+hscat(dissolved.n2o.nmol~1, spl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))
+
+hscat(log10(dissolved.n2o.nmol)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
 
 # tpl
+tpl <- dg.sf %>% filter(WSA9_NAME == "Temperate Plains")
 tpl_sp <- as(tpl, "Spatial")
 class(tpl_sp)
-ggplot(tpl,aes(dissolved.ch4)) + geom_histogram()
-ggplot(tpl,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
-
-hscat(dissolved.ch4~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
-
-hscat(log10(dissolved.ch4)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
-
-ggplot(tpl,aes(dissolved.co2)) + geom_histogram()
-ggplot(tpl,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
-
-hscat(dissolved.co2~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
-
-hscat(log10(dissolved.co2)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
-
-ggplot(tpl,aes(dissolved.n2o)) + geom_histogram()
-ggplot(tpl,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
-
-hscat(dissolved.n2o~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
-
-hscat(log10(dissolved.n2o)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(dissolved.ch4.umol~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(log10(dissolved.ch4.umol)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(tpl,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(tpl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(dissolved.co2.umol~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(log10(dissolved.co2.umol)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
+
+ggplot(tpl,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(tpl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(dissolved.n2o.nmol~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))
+
+hscat(log10(dissolved.n2o.nmol)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))
 
 # umw
+umw <- dg.sf %>% filter(WSA9_NAME == "Upper Midwest")
 umw_sp <- as(umw, "Spatial")
 class(umw_sp)
-ggplot(umw,aes(dissolved.ch4)) + geom_histogram()
-ggplot(umw,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
-
-hscat(dissolved.ch4~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
-
-hscat(log10(dissolved.ch4)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
-
-ggplot(umw,aes(dissolved.co2)) + geom_histogram()
-ggplot(umw,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
-
-hscat(dissolved.co2~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
-
-hscat(log10(dissolved.co2)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
-
-ggplot(umw,aes(dissolved.n2o)) + geom_histogram()
-ggplot(umw,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
-
-hscat(dissolved.n2o~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
-
-hscat(log10(dissolved.n2o)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
+
+hscat(dissolved.ch4.umol~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
+
+hscat(log10(dissolved.ch4.umol)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+ggplot(umw,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(umw,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))
+
+hscat(dissolved.co2.umol~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+hscat(log10(dissolved.co2.umol)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(umw,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(umw,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+hscat(dissolved.n2o.nmol~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))
+
+hscat(log10(dissolved.n2o.nmol)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))
 
 # wmt
+wmt <- dg.sf %>% filter(WSA9_NAME == "Western Mountains")
 wmt_sp <- as(wmt, "Spatial")
 class(wmt_sp)
-ggplot(wmt,aes(dissolved.ch4)) + geom_histogram()
-ggplot(wmt,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
-
-hscat(dissolved.ch4~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
-
-hscat(log10(dissolved.ch4)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
-
-ggplot(wmt,aes(dissolved.co2)) + geom_histogram()
-ggplot(wmt,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
-
-hscat(dissolved.co2~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
-
-hscat(log10(dissolved.co2)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
-
-ggplot(wmt,aes(dissolved.n2o)) + geom_histogram()
-ggplot(wmt,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
-
-hscat(dissolved.n2o~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
-
-hscat(log10(dissolved.n2o)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
+ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(dissolved.ch4.umol~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(log10(dissolved.ch4.umol)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
+
+ggplot(wmt,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(wmt,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(dissolved.co2.umol~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(log10(dissolved.co2.umol)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
+
+ggplot(wmt,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(wmt,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(dissolved.n2o.nmol~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))
+
+hscat(log10(dissolved.n2o.nmol)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))
 
 # xer
+xer <- dg.sf %>% filter(WSA9_NAME == "Xeric")
 xer_sp <- as(xer, "Spatial")
 class(xer_sp)
-ggplot(xer,aes(dissolved.ch4)) + geom_histogram()
-ggplot(xer,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
+ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
 # 10 km bin/lag only has 2 points plotted so start with 50 km
-hscat(log10(dissolved.ch4)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
-
-hscat(dissolved.ch4~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
-
-hscat(log10(dissolved.ch4)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
-
-ggplot(xer,aes(dissolved.co2)) + geom_histogram()
-ggplot(xer,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
-
-hscat(dissolved.co2~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
-
-hscat(log10(dissolved.co2)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
-
-ggplot(xer,aes(dissolved.n2o)) + geom_histogram()
-ggplot(xer,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
-
-hscat(dissolved.n2o~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
-
-hscat(log10(dissolved.n2o)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
+hscat(log10(dissolved.ch4.umol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(dissolved.ch4.umol~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(log10(dissolved.ch4.umol)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
+
+ggplot(xer,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(xer,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(dissolved.co2.umol~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(log10(dissolved.co2.umol)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
+
+ggplot(xer,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(xer,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(dissolved.n2o.nmol~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))
+
+hscat(log10(dissolved.n2o.nmol)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))
 
@@
 class(dg_sp)
-ggplot(dg.sf,aes(dissolved.ch4)) + geom_histogram()
-ggplot(dg.sf,aes(dissolved.ch4)) + geom_histogram() + scale_x_log10()
-hscat(log10(dissolved.ch4)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
-
-hscat(dissolved.ch4~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
-
-hscat(log10(dissolved.ch4)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
-
-ggplot(dg.sf,aes(dissolved.co2)) + geom_histogram()
-ggplot(dg.sf,aes(dissolved.co2)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.co2)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
-
-hscat(dissolved.co2~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
-
-hscat(log10(dissolved.co2)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
-
-ggplot(dg.sf,aes(dissolved.n2o)) + geom_histogram()
-ggplot(dg.sf,aes(dissolved.n2o)) + geom_histogram() + scale_x_log10()
-
-hscat(log10(dissolved.n2o)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
-
-hscat(dissolved.n2o~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
-
-hscat(log10(dissolved.n2o)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
+ggplot(dg.sf,aes(dissolved.ch4.umol)) + geom_histogram()
+ggplot(dg.sf,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
+hscat(log10(dissolved.ch4.umol)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(dissolved.ch4.umol~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(log10(dissolved.ch4.umol)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(dg.sf,aes(dissolved.co2.umol)) + geom_histogram()
+ggplot(dg.sf,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.co2.umol)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(dissolved.co2.umol~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(log10(dissolved.co2.umol)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
+
+ggplot(dg.sf,aes(dissolved.n2o.nmol)) + geom_histogram()
+ggplot(dg.sf,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()
+
+hscat(log10(dissolved.n2o.nmol)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(dissolved.n2o.nmol~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))
+
+hscat(log10(dissolved.n2o.nmol)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))
 
@@
 # also includ WSA9, WSA9_NAME feature
-dg_1 <- select(dg,site.id:dissolved.n2o, WSA9, WSA9_NAME)
+dg_1 <- select(dg,site.id:dissolved.n2o.nmol, WSA9, WSA9_NAME)
 names(dg_1)
@@
 # names(cpl_df)
-summary(cpl$dissolved_n2o)
-
-moran.plot(log(cpl$dissolved_n2o), listw = dsnb1.listw, labels = cpl$site_id)
+summary(cpl$dissolved_n2o_nmol)
+
+moran.plot(log(cpl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = cpl$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(nap_df)
-summary(nap$dissolved_n2o)
-
-moran.plot(log(nap$dissolved_n2o), listw = dsnb1.listw, labels = nap$site_id)
+summary(nap$dissolved_n2o_nmol)
+
+moran.plot(log(nap$dissolved_n2o_nmol), listw = dsnb1.listw, labels = nap$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(npl_df)
-summary(npl$dissolved_n2o)
-
-moran.plot(log(npl$dissolved_n2o), listw = dsnb1.listw, labels = npl$site_id)
+summary(npl$dissolved_n2o_nmol)
+
+moran.plot(log(npl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = npl$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(sap_df)
-summary(sap$dissolved_n2o)
-
-moran.plot(log(sap$dissolved_n2o), listw = dsnb1.listw, labels = sap$site_id)
+summary(sap$dissolved_n2o_nmol)
+
+moran.plot(log(sap$dissolved_n2o_nmol), listw = dsnb1.listw, labels = sap$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(spl_df)
-summary(spl$dissolved_n2o)
-
-moran.plot(log(spl$dissolved_n2o), listw = dsnb1.listw, labels = spl$site_id)
+summary(spl$dissolved_n2o_nmol)
+
+moran.plot(log(spl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = spl$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(tpl_df)
-summary(tpl$dissolved_n2o)
-
-moran.plot(log(tpl$dissolved_n2o), listw = dsnb1.listw, labels = tpl$site_id)
+summary(tpl$dissolved_n2o_nmol)
+
+moran.plot(log(tpl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = tpl$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(umw_df)
-summary(umw$dissolved_n2o)
-
-moran.plot(log(umw$dissolved_n2o), listw = dsnb1.listw, labels = umw$site_id)
+summary(umw$dissolved_n2o_nmol)
+
+moran.plot(log(umw$dissolved_n2o_nmol), listw = dsnb1.listw, labels = umw$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(wmt_df)
-summary(wmt$dissolved_n2o)
-
-moran.plot(log(wmt$dissolved_n2o), listw = dsnb1.listw, labels = wmt$site_id)
+summary(wmt$dissolved_n2o_nmol)
+
+moran.plot(log(wmt$dissolved_n2o_nmol), listw = dsnb1.listw, labels = wmt$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(xer_df)
-summary(xer$dissolved_n2o)
-
-moran.plot(log(xer$dissolved_n2o), listw = dsnb1.listw, labels = xer$site_id)
+summary(xer$dissolved_n2o_nmol)
+
+moran.plot(log(xer$dissolved_n2o_nmol), listw = dsnb1.listw, labels = xer$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 # names(nla_df)
-summary(nla$dissolved_n2o)
-
-moran.plot(log(nla$dissolved_n2o), listw = dsnb1.listw, labels = nla$site_id)
+summary(nla$dissolved_n2o_nmol)
+
+moran.plot(log(nla$dissolved_n2o_nmol), listw = dsnb1.listw, labels = nla$site_id)
 rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
@@
 
-correlo_n2o_ncf <- ncf::correlog(x = dg_cplvisit1$map_lon_dd, y = dg_cplvisit1$map_lat_dd, z = dg_cplvisit1$log_n2o, increment = 50, resamp = 1000, latlon = TRUE)
+correlo_n2o_ncf <- ncf::correlog(x = dg_cplvisit1$map_lon_dd, y = dg_cplvisit1$map_lat_dd, z = log(dg_cplvisit1$dissolved_n2o_nmol), increment = 50, resamp = 1000, latlon = TRUE)
 names(correlo_n2o_ncf)
-plot(correlo_n2o_ncf) # works when pgmiress commented out
-rm(dg_napvisit1, correlo_n2o_ncf)
+#plot(correlo_n2o_ncf) # works when pgmiress commented out
 
@@
 class(correlo_n2o_ncf)
-ncf::plot.correlog(correlo_n2o_ncf)
+#ncf::plot.correlog(correlo_n2o_ncf)
 
@@
 
-out_cplcorrelo_n2o <- ncf::correlog(x = out_cplvisit1$map_lon_dd, y = out_cplvisit1$map_lat_dd, z = out_cplvisit1$log_n2o, increment = 50, resamp = 1000, latlon = TRUE)
-
-plot(out_cplcorrelo_n2o)
-rm(out_cplvisit1, out_cplcorrelo_n2o)
+out_cplcorrelo_n2o <- ncf::correlog(x = out_cplvisit1$map_lon_dd, y = out_cplvisit1$map_lat_dd, z = log(out_cplvisit1$dissolved_n2o_nmol), increment = 50, resamp = 1000, latlon = TRUE)
+
+#plot(out_cplcorrelo_n2o)
+#rm(out_cplvisit1, out_cplcorrelo_n2o)
 
@@
 # Uses normality assumption, which is faster, to assess significance
+dg_napvisit1 <- dg_1 %>% filter(wsa9 == "NAP" & visit_no == 1)
 coords_nap <- cbind(dg_napvisit1$map_lon_dd, dg_napvisit1$map_lat_dd)
-correl_napn2o_pg <- pgirmess::correlog(coords_nap, dg_napvisit1$log_n2o, method = "Moran", nbclass = NULL, alternative = "two.sided")
+correl_napn2o_pg <- pgirmess::correlog(coords_nap, log(dg_napvisit1$dissolved_n2o_nmol), method = "Moran", nbclass = NULL, alternative = "two.sided")
 
@@
 coords_sp_nap <- coordinates(nap_sp_visit1)
-correl_spnapn2o_pg <- pgirmess::correlog(coords_sp_nap, nap_sp_visit1@data$log_n2o, method = "Moran", nbclass = 30, alternative = "two.sided")
+correl_spnapn2o_pg <- pgirmess::correlog(coords_sp_nap, log(nap_sp_visit1@data$dissolved_n2o_nmol), method = "Moran", nbclass = 30, alternative = "two.sided")
 
@@
 
-dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o, WSA9, WSA9_NAME, ONOFFNET, nitrate.n.result,chla.result.vol, MAX.BF, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size_cat)
+dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o.nmol, WSA9, WSA9_NAME, ONOFFNET, nitrate.n.result,chla.result.vol, max.bf, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size.cat)
 names(dg_1)
@@
 names(dg_1)
-dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o),3))
+dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3))
 ## Subset dg_1 by visit_no == 1
@@
 
-```
-
 ## Semivariograms
@@
 ```{r, warning=FALSE, message=FALSE}
+
 # Variogram function requires sp object
@@
 # CO2 variogram
-co2.v <- variogram(dissolved.co2~1, dg.sp)
+co2.v <- variogram(dissolved.co2.umol~1, dg.sp)
 co2.v.fit <- fit.variogram(co2.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
@@
 # CH4 variogram
-ch4.v <- variogram(dissolved.ch4~1, dg.sp)
+ch4.v <- variogram(dissolved.ch4.umol~1, dg.sp)
 ch4.v.fit <- fit.variogram(ch4.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
@@
 # N2O variogram
-n2o.v <- variogram(dissolved.n2o~1, dg.sp)
+n2o.v <- variogram(dissolved.n2o.nmol~1, dg.sp)
 n2o.v.fit <- fit.variogram(n2o.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
@@
 dg %>%
-  select(dissolved.n2o, MeanDUsed, ph.result, nitrate.n.result, ntl.lab.result, chla.result.vol, ptl.lab.result, surftemp) %>%
+  select(dissolved.n2o.nmol, MeanDUsed, ph.result, nitrate.n.result, ntl.lab.result, chla.result.vol, ptl.lab.result, surftemp) %>%
   correlate()
@@
 library(tree)
-model <- tree(dissolved.n2o ~ MeanDUsed + nitrate.n.result + ntl.lab.result + chla.result.vol + ptl.lab.result + surftemp + WSA9, data = dg)
+model <- tree(dissolved.n2o.nmol ~ MeanDUsed + nitrate.n.result + ntl.lab.result + chla.result.vol + ptl.lab.result + surftemp + WSA9, data = dg)
 
@@
 text(model)
-
-```
-
-We may also want to include design variables in the analysis.
-The specifics of the NLA17 design have not been disseminated, but are likely very similar to the NLA12 design.
-Per the NLA12 Design Documentation (https://www.epa.gov/sites/production/files/2014-03/documents/nla2012_design_documentation_20110320.pdf):
-
-> Stratification  
-The survey design was stratified by state and by NLA12_CLS.
-NLA12_CLS has three classes: (1) NLA07RVT – defined as all NLA 2007 lakes that were target and sampled, (2) NLA12NEW – remaining lakes in NHD-Plus that are included in the sample frame, and (3) Exclude – lakes in NHD-Plus that are excluded from the sample frame (see Sample Frame section above).
-Each state design has two strata, ST_ NLA07RVT and ST_ NLA12NEW (where ST is replaced by two letter state abbreviation.
-The total number of strata is 96 (two for each state).
-
-> Unequal Probability Categories  
-The 48 state strata for lakes from the NLA 2007 that would be visited again in 2012 was an equal probability design within each stratum.
-The 48 state strata NLA12NEW was an unequal probability design within each state stratum.
-The unequal probability categories were defined based on lake area: 1 to 4 ha, 4 to 10 ha, 10 to 20 ha, 20 to 50 ha and greater than 50 ha.
-
-As you can see, the design is actually pretty complicated.  The file currently contains the following design elements:
-
-* state - two digit state abbreviation.  See line 62 above.
-* AREASQKM - waterbody area (km2)
-
-Corvallis has not yet shared the other design elements (i.e.
-NLA17_CLS) and sample weights.
-Presumably this file will be delivered with the code to execute the population estimates via spSurvey.
-Steve Paulsen will speak to Tom Kincaid about this during the week of Dec.
-16.
+```

### 2020-12-16 scripts/dgIndicatorAnalysis.Rmd
Added Random Semivariograms code chunk to dgIndicatorAnalysis

--- 2020-12-11 scripts/dgIndicatorAnalysis.Rmd
@@
 library(ncf) # ncf::correlogram
+library(lattice) # xyplot
 
@@
 
+## Random Semivariograms
+Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.
+Downloaded Chapter 8 code from https://asdar-book.org/ for generating random variograms.
+An initial empirical, or sample, variogram is calculated and plotted for pH, with that object saved.
+Then pH values are randomly sampled and assigned to the coordinates so that a random semivariogram can be calculated.
+That is done 100 times, and the gray lines of the random semivariograms are plotted with the blue line of the empirical semivariogram.
+
+The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation.
+```{r, random_semivar}
+names(dg)
+dg_1 <- clean_names(dg)
+# also includ WSA9, WSA9_NAME feature plus variables from Jake's 02/25/2020 email
+# added ntl.lab.result, ptl.lab.result on 06/18/2020
+# added physical variables, temperature and depth, on 06/23/2020
+# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
+# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
+# added ph_result 12/14/2020
+names(dg_1)
+dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, wsa9, wsa9_name, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & wsa9 != "AK")
+
+names(dg_1)
+
+#filter in View returns 514 nitrate_n_result equal to zero
+dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
+  mutate(log_ntl = round(log(ntl_lab_result),3)) %>% 
+  mutate(log_ptl = round(log(ptl_lab_result),3)) %>% 
+  mutate(log_chla = round(log(chla_result_vol),1)) %>% 
+  mutate(log_si = round(log(silica_result),2)) %>% 
+  mutate(log_ca = round(log(calcium_result),2)) %>% 
+  mutate(ph_result = round(ph_result,2)) %>% 
+  mutate(nitrate_n_b = case_when(
+    nitrate_n_result <= 0.0001 ~ "low",
+    nitrate_n_result > 0.0001 ~ "high"))
+
+dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
+summary(dg_1$nitrate_n_b)
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)
+
+dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+st_crs(dg_sf1)
+
+z_ph_sf1 = variogram(ph_result~1, dg_sf1)
+plot(z_ph_sf1)
+# plot(z_ph_sf1, plot.numbers = TRUE)
+# plot(z_ph_sf1, ylab = "NLA ph Semivariance")
+# plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters)", ylab = "NLA ph semivariance")
+# z_ph_plot1 <- plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters) ")
+
+print(xyplot(gamma ~ dist, z_ph_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$ph_result)
+                     v = variogram(random ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
+))
+
+z_cressie_ph <- variogram(ph_result~1, dg_sf1, cressie = TRUE)
+plot(variogram(ph_result~1, dg_sf1, cressie = TRUE))
+
+print(xyplot(gamma ~ dist, z_cressie_ph, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$ph_result)
+                     v = variogram(random ~ 1, dg_sf1, cressie = TRUE)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
+))
+
+z_do_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
+plot(z_do_sf1)
+# plot(z_ph_sf1, plot.numbers = TRUE)
+plot(z_do_sf1, ylab = "NLA Surface DO Semivariance")
+
+```
+
 # Exploratory analysis of N2O data

### 2021-01-14 scripts/dataMunge.Rmd
added air data back into generated .RData object

@@
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-11-10.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2021-01-13.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
 When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2020-11-10.txt"),
+                               "inputData/nla17gasDataAggregated_2021-01-13.txt"),
                  header = TRUE, sep = "\t", as.is = TRUE) %>%
-  filter(sample.source == "DG", # pull out dissolved gas and leave out gas samples
-         !grepl("AK", site.id)) # omit alaska sites
+  filter(!grepl("AK", site.id)) # omit Alaska sites
 ```
@@
          dissolved.n2o.nmol = dissolved.n2o * 1000000000, # mol/L -> nmol/L
+         sat.ch4.umol = sat.ch4 * 1000000, # mol/L -> umol/L
+         sat.co2.umol = sat.co2 * 1000000, # mol/L -> umol/L
+         sat.n2o.nmol =  sat.n2o * 1000000000, # mol/L -> nmol/L
          # add source/sink column
@@
          ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
-         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
+         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink")) %>%
+  # remove fields no longer needed
+  select(-dissolved.ch4, -dissolved.co2, -dissolved.n2o,
+         -sat.ch4, -sat.co2, - sat.n2o)
 ```
@@
 ### Calculate emission rate. 
-The emission rate is the mass flux of a gas into or out of a waterbody.
-Emission rates are typically expressed as mass m^-2^ day^-1^.
-The phrase 'emission rate' is used interchangeably with 'Flux rate' and is often reported in the literature as (F~gas~), where the subscript 'gas' refers to the gas of interest (i.e.
-N~2~O, CO~2~, CH~4~).
-Emission rate can be  calculated as the product of gas^\*^ and the gas transfer velocity (k):
+The emission rate (E~gas~) is the rate at which a gas crosses the air water interface and is expressed on an areal basis (i.e.
+mass m^-2^ day^-1^).
+The emission rate can be converted to a flux (F~gas~; mass per unit time; mass d^-1^).
+In the context of emissions from waterbodies, the flux (F~gas~) is calculated as the product of the emission rate (E~gas~) and waterbody area.
 
-F~gas~ = gas^\*^ * k
+E~gas~ can be calculated as the product of gas^\*^ and the gas transfer velocity (k):
+
+E~gas~ = gas^\*^ * k
 
@@
 k~600~ is a standardized value that must be corrected for the differences in diffusivity among gases and water temperature at the sampling sites.
-These corrections will be implemented at a later date and emission rates will be calculated using k600 for now.
+These corrections will be implemented at a later date and emission rates will be calculated using k~600~ for now.
 
@@
   mutate(
-    # f.ch4 umol CH4/m2/d 
-    f.ch4 = ((dissolved.ch4 - (sat.ch4 * 1000000))*1000) * # 1000000 mol -> umol.  1000 L to m3
+    # e.ch4 umol CH4/m2/d 
+    e.ch4.umol.d = ((dissolved.ch4.umol - sat.ch4.umol) * 1000) * # 1000 L to m3
            (k600.day * (24/100)), # 24 hour to day.  100 cm to m
-    # f.co2 umol CO2/m2/d
-    f.co2 = ((dissolved.co2 - (sat.co2 * 1000000))*1000) * # 1000000 mol -> umol.  1000 L to m3
+    # e.co2 umol CO2/m2/d
+    e.co2.umol.d = ((dissolved.co2.umol - sat.co2.umol) * 1000) * # 1000 L to m3
            (k600.day * (24/100)), # 24 hour to day.  100 cm to m
-        # f.n2o nmol N2O/m2/d
-    f.n2o = ((dissolved.n2o - (sat.n2o * 1000000000))*1000) * # 1000000 mol -> nmol.  1000 L to m3
+        # e.n2o nmol N2O/m2/d
+    e.n2o.nmol.d = ((dissolved.n2o.nmol - sat.n2o.nmol) * 1000) * # 1000 L to m3
            (k600.day * (24/100)), # 24 hour to day.  100 cm to m
     # total flux of CH4 per day kmol CH4 day-1
-    f.ch4.d = f.ch4 * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
+    f.ch4.km.d = e.ch4.umol.d * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
     # total flux of CO2 per day kmol CO2 day-1
-    f.co2.d = f.co2 * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
+    f.co2.km.d = e.co2.umol.d * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
     # total flux of N2O per day mol N2O day-1
-    f.n2o.d = f.n2o * area.ha * 10000 * (1/10^9)) %>% # 1ha = 10,000 m2. 10^9 nmol->umol->mmol->mol
-         filter(state != "AK", sample.source == "DG")
+    f.n2o.m.d = e.n2o.nmol.d * area.ha * 10000 * (1/10^9))  # 1ha = 10,000 m2. 10^9 nmol->umol->mmol->mol
 
@@
 # data used for population estimates included only visit one.  I don't know rationale for this, but will
-# do the same with the revised file here.
+# do the same with the revised file here.  Also need to strip out air samples.
 unique(cdf$VISIT_NO) # 1
-dg.1 <- dg %>% filter(visit.no == 1)
+dg.1 <- dg %>% filter(visit.no == 1, sample.source == "DG")
 
@@
 dim(cdf) # 1005, 81
-dim(dg.1) # 1090, 396
+dim(dg.1) # 1090, 306
 
@@
 cdf.dg <- left_join(cdf, 
-                    select(dg.1, site.id, visit.no, f.ch4, f.co2, 
-                           f.n2o, f.ch4.d, f.co2.d, f.n2o.d) %>%
+                    select(dg.1, site.id, visit.no, 
+                           e.ch4.umol.d , e.co2.umol.d, e.n2o.nmol.d, 
+                           f.ch4.km.d, f.co2.km.d, f.n2o.m.d) %>%
                       # rename for consistency with Tom's data

### 2021-01-21 scripts/CDF_Estimates_Jb.Rmd

@@
 
-Revised: December 12, 2019 to calculate estimates using the SRS variance
-         estimator in addition to the local mean variance estimator
+Revision history:
+December 12, 2019 to calculate estimates using the SRS variance estimator in addition to the local mean variance estimator.
 
 November 2020, expanded to include estimates of N2O emission rates.
+
+December 3, 2020 added extent estimate of N~2~O source/sink
 
@@
 sessionInfo()
+
+# Setup file path
+localPath <- Sys.getenv("USERPROFILE")
 
@@
 cat("\n\nSummary Information for Response Variables\n")
-temp <- match(c("CH4_SAT_RATIO", "CO2_SAT_RATIO", "DELTA_13_CH4_C",
-   "DELTA_13_CO2_C", "DISSOLVED_CH4", "DISSOLVED_CO2", "DISSOLVED_N2O",
-    "N2O_SAT_RATIO", "F_N2O", "F_N2O_D"), names(cdf))
+temp <- match(c("CH4_SAT_RATIO", "DISSOLVED_CH4", "DELTA_13_CH4_C", "E_CH4_UMOL_D", "F_CH4_KM_D",
+                "CO2_SAT_RATIO", "DISSOLVED_CO2", "DELTA_13_CO2_C", "E_CO2_UMOL_D", "F_CO2_KM_D",
+                "N2O_SAT_RATIO", "DISSOLVED_N2O", "E_N2O_NMOL_D", "F_N2O_M_D"), names(cdf))
 for(c in temp) {
@@
                      AREA_CAT6=cdf$AREA_CAT6,
-                     #  > 1 hour with all variables and subpopulations
-                     # AG_ECO3_NM=cdf$AG_ECO3_NM,
+                     #AG_ECO3_NM=cdf$AG_ECO3_NM,
                      AG_ECO9_NM=cdf$AG_ECO9_NM
-                     # EPA_REG=cdf$EPA_REG,
-                     # MAJ_BAS_NM=cdf$MAJ_BAS_NM,
-                     # HUC2_NM=cdf$HUC2_NM,
-                     # OWN_NARS=cdf$OWN_NARS,
-                     # FS_EW=cdf$FS_EW
+                     #EPA_REG=cdf$EPA_REG,
+                     #MAJ_BAS_NM=cdf$MAJ_BAS_NM,
+                     #HUC2_NM=cdf$HUC2_NM,
+                     #OWN_NARS=cdf$OWN_NARS,
+                     #FS_EW=cdf$FS_EW
                      )
@@
 # Create the data.cont data frame, which specifies the variables to use in the
-# analysis.  > 1 hour with all variables.
+# analysis.  30 minutes on VM.
 data.cont <- data.frame(siteID=cdf$SITE_ID,
-                        # CH4_SAT_RATIO=cdf$CH4_SAT_RATIO,
-                        # CO2_SAT_RATIO=cdf$CO2_SAT_RATIO,
-                        # DELTA_13_CH4_C=cdf$DELTA_13_CH4_C,
-                        # DELTA_13_CO2_C=cdf$DELTA_13_CO2_C,
-                        # DISSOLVED_CH4=cdf$DISSOLVED_CH4,
-                        # DISSOLVED_CO2=cdf$DISSOLVED_CO2,
-                        # DISSOLVED_N2O=cdf$DISSOLVED_N2O,
-                        # N2O_SAT_RATIO=cdf$N2O_SAT_RATIO,
-                        F_N2O = cdf$F_N2O) # diffusive N2O nmol N2O m-2 day-1
+                        CH4_SAT_RATIO=cdf$CH4_SAT_RATIO,
+                        CO2_SAT_RATIO=cdf$CO2_SAT_RATIO,
+                        N2O_SAT_RATIO=cdf$N2O_SAT_RATIO,
+                        DELTA_13_CH4_C=cdf$DELTA_13_CH4_C,
+                        DELTA_13_CO2_C=cdf$DELTA_13_CO2_C,
+                        DISSOLVED_CH4=cdf$DISSOLVED_CH4,
+                        DISSOLVED_CO2=cdf$DISSOLVED_CO2,
+                        DISSOLVED_N2O=cdf$DISSOLVED_N2O,
+                        E_CH4_UMOL_D = cdf$E_CH4_UMOL_D, # umol CH4 m-2 day-1
+                        F_CH4_KM_D = cdf$F_CH4_KM_D, # kmol CH4 day-1
+                        E_CO2_UMOL_D = cdf$E_CO2_UMOL_D, # umol CO2 m-2 day-1
+                        F_CO2_KM_D = cdf$F_CO2_KM_D, # kmol CO2 day-1
+                        E_N2O_NMOL_D = cdf$E_N2O_NMOL_D, # nmol N2O m-2 day-1
+                        F_N2O_M_D = cdf$F_N2O_M_D) # mol N2O day-1
 
@@
 if(exists("warn.df")) rm("warn.df")
-CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont) # < 1 hour with all variables. 1 minute with F_N2O
-
+time.initial <- Sys.time()
+CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont) # 30 minutes on VM
+Sys.time() - time.initial
 ```
@@
 if(exists("warn.df")) {
-   cat("\nWarning messages generated during the call to cat.analysis:\n")
+   cat("\nWarning messages generated during the call to cont.analysis:\n")
    warnprnt()
 } else {
-   cat("\nNo warning messages were generated during the call to cat.analysis.\n")
+   cat("\nNo warning messages were generated during the call to cont.analysis.\n")
 }
@@
                  "inputData/populationEstimates/",
-                 "NLA_2017_Percentile_Estimates_Local_Mean_F_N2O_JB",
+                 "NLA_2017_Percentile_Estimates_Local_Mean_JB",
                  gsub("-", "", Sys.Date()), ".csv"),
@@
 
-#
-# Calculate CDF estimates using the SRS variance estimator
-#
+# Calculate extent estimate of N2O source/sink
+```{r}
+# Can use same sites, subpop, and design frames from above. 
+# Add N2O source/sink column to dg
+cdf$N2O_SRC_SNK = with(cdf, ifelse(N2O_SAT_RATIO > 1, "source", "sink"))
+data.cat <-  data.frame(siteID=cdf$SITE_ID,
+                        N2O_SRC_SNK = cdf$N2O_SRC_SNK)
 
-```{r, echo=FALSE, results="hide"}
+Extent_Estimate <- cat.analysis(sites, subpop, design, data.cat)
 
-# Calculate the estimates
-cat("\n\nCalculate CDF estimates using the SRS variance estimator\n")
-if(exists("warn.df")) rm("warn.df")
-CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont,
-   vartype = "SRS")
-
+print(Extent_Estimate)
 ```
@@
 
-# Write CDF estimates as a comma-separated value (csv) file
-eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_SRS_",
-   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
-
-# Create a PDF file containing plots of the CDF estimates
-eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_SRS_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))
-
-# Write percentile estimates as a csv file
-eval(parse(text = paste0("write.csv(CDF_Estimates$Pct, 'NLA_2017_Percentile_Estimates_SRS_",
-   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
+# # Write extent estimates as a comma-separated value (csv) file
+write.csv(Extent_Estimate, 
+          paste0(localPath, 
+                 "/Environmental Protection Agency (EPA)/",
+                 "ORD NLA17 Dissolved Gas - Documents/",
+                 "inputData/populationEstimates/",
+                 "NLA_2017_Extent_Estimates",
+                 gsub("-", "", Sys.Date()), ".csv"),
+          row.names=FALSE)
 
 ```
+
+# Calculate CDF estimates using the SRS variance estimator
+#
+
+```{r, echo=FALSE, results="hide"}
+
+# Calculate the estimates
+# cat("\n\nCalculate CDF estimates using the SRS variance estimator\n")
+# if(exists("warn.df")) rm("warn.df")
+# CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont,
+#    vartype = "SRS")
+
+```
+
+```{r, echo=FALSE}
+
+# # Check for warning messages and print them if any exist
+# if(exists("warn.df")) {
+#    cat("\nWarning messages generated during the call to cat.analysis:\n")
+#    warnprnt()
+# } else {
+#    cat("\nNo warning messages were generated during the call to cat.analysis.\n")
+# }
+# 
+# # Write CDF estimates as a comma-separated value (csv) file
+# eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_SRS_",
+#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
+# 
+# # Create a PDF file containing plots of the CDF estimates
+# eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_SRS_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))
+# 
+# # Write percentile estimates as a csv file
+# eval(parse(text = paste0("write.csv(CDF_Estimates$Pct, 'NLA_2017_Percentile_Estimates_SRS_",
+#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
+
+```

### 2021-01-21 scripts/dataMunge.Rmd

@@
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2021-01-13.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-11-10.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
 When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2021-01-13.txt"),
+                               "inputData/nla17gasDataAggregated_2020-11-10.txt"),
                  header = TRUE, sep = "\t", as.is = TRUE) %>%
-  filter(!grepl("AK", site.id)) # omit Alaska sites
+  filter(sample.source == "DG", # pull out dissolved gas and leave out gas samples
+         !grepl("AK", site.id)) # omit alaska sites
 ```
@@
 k~600~ is a standardized value that must be corrected for the differences in diffusivity among gases and water temperature at the sampling sites.
-These corrections will be implemented at a later date and emission rates will be calculated using k~600~ for now.
+These corrections will be implemented at a later date and emission rates will be calculated using k600 for now.
 
@@
 # data used for population estimates included only visit one.  I don't know rationale for this, but will
-# do the same with the revised file here.  Also need to strip out air samples.
+# do the same with the revised file here.
 unique(cdf$VISIT_NO) # 1
-dg.1 <- dg %>% filter(visit.no == 1, sample.source == "DG")
+dg.1 <- dg %>% filter(visit.no == 1)
 
@@
 dim(cdf) # 1005, 81
-dim(dg.1) # 1090, 306
+dim(dg.1) # 1090, 399
 
@@
 
-dim(cdf.dg) # 1005, 87 same as cdf above, so good
+dim(cdf.dg) # 1005, 86 same as cdf above, so good
 

### 2021-01-21 scripts/DG_sensitivity_to_measurement_error.Rmd

---
title: "Evaluating sensitivity to measurement error in dissolved gas measurements"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook: 
    depth: 4
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
#bibliography: RWM_Endnote_Library.bib
link-citations: yes
---

<style>
.vscroll-plot {
    width: 800px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

# Background
In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).

# Setup R
```{r setup_1, include=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")

#set directories and R package library
setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
.libPaths("C:/Users/rmartin/Desktop/R/library")
```

```{r load_packages, warning=FALSE, message=FALSE}
library(brms)
library(bayesplot)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(gridExtra)

options(mc.cores = parallel::detectCores())
options( max.print = 2000 )
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}

#Calc geometric mean
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

gm_sd = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(sd(log(x), na.rm = na.rm))
  } else {
    exp(sd(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

#calculate skew
MySkew <- function(x) {
     n <- length(x)
     v <- var(x)
     m <- mean(x)
     third.moment <- (1/(n - 2)) * sum((x - m)^3)
     third.moment/(var(x)^(3/2))
 }

#Center and scale by (1) sd
MyStd <- function(x){ (x-mean(x))/(sd(x)*2)}

#First letter in string to uppercase
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x}
```

# Dissolved gas concentration calculation

The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water is:

$$C = 10^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}$$

# Set up model for simulating true value

```{r setvars_and_check}
B <- 99 # barometric pressure (kPa)
V_g <- 20 # volume of reference gas in headspace
V_w <- 40 # volume of water below headspace
T_c <- 23 # temperature ( C )
C_e <- 0.055 # gas mixture ratio in headspace
C_a <- 0.091 # gas mixture ratio in reference
H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant

C <- 10e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )

C <- C * 1e9 #convert mol to nmol

round( C, 1 )

```

# Simulate measurement error for gas measured in water
```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 )

C_e_me <- rnorm( 1e4, C_e, 0.001 ) #m.e. in measured gas in headspace

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

qplot( x = C_e_me ) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_e , color = 'red' ) + 
  ggtitle( "Simulated measurement error in water gas concentration") +
  xlab( "concentration (?)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error for gas measured in air
```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

C_a_me <- rnorm( 1e4, C_a, 0.001 ) #m.e. in measured gas in air

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = C_a_me ) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (?)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in $C_e$ and $C_a$
```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in water temperature
```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

T_c_me <- rnorm( 1e4, T_c, 0.1 ) #m.e. in measured water temp

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = T_c_me ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = T_c , color = 'red' ) + 
  ggtitle( "Simulated measurement error in temperature") +
  xlab( "temperature (C)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in barometric pressure
```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured water temp

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = B_me ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in water volume
```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

V_w_me <- rnorm( 1e4, V_w, 1 ) #m.e. in measured water volume
V_g_me <- 60 - V_w_me

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = V_w_me ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = V_w , color = 'red' ) + 
  ggtitle( "Simulated measurement error in water volume") +
  xlab( "water vol (mL)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error for all observables
```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

### 2021-01-21 scripts/dgIndicatorAnalysis.Rmd
correcting merge conflict flags

@@
             "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.2020-12-02.RData"))
+            "inputData/dg.2021-01-14.RData"))
 
@@
 dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, WSA9, WSA9_NAME)
+
 names(dg_1)
-dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") # underscores are "safer" names used in GIS
+dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") %>% 
+  mutate(log_n2o = round(log(dissolved_n2o),3)) %>% 
+  mutate(log_n2osatratio = round(log(n2o_sat_ratio),3))
+# underscores are "safer" names used in GIS
+
 names(dg_1)
@@
 
+dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1184 obs vs. previously 1185 obs., why change in sample size
+st_crs(dg_sf1)
+
 # Bryan suggested EPSG 4326

### 2021-02-01 scripts/dgIndicatorAnalysis.Rmd
updated path to current data object

@@
             "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.2021-01-14.RData"))
+            "inputData/dg.2021-02-01.RData"))
 

### 2021-02-16 scripts/dgIndicatorAnalysis.Rmd
Using dg.2021-02-21.RData re-ran random semivariogram analysis.

--- 2020-12-16 scripts/dgIndicatorAnalysis.Rmd
@@
 ```{r}
-# Read data file 
+# Had used dg.2021-01-14 now try dg.2021-02-01.RData to see if results change
  load(paste0(localPath,
@@
             "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.2020-12-02.RData"))
+            "inputData/dg.2021-02-01.RData"))
 
@@
 
+names(dg)
+
 dg %>% 
-  ggplot(aes(as.factor(x = WSA9), y = log(dissolved.n2o.nmol))) + geom_boxplot()
+  ggplot(aes(as.factor(x = ag.eco9), y = log(dissolved.n2o.nmol))) + geom_boxplot()
+# shows a boxplot for NA category
 
 dg %>% 
-  ggplot(aes(as.factor(x = WSA9), y = (n2o.sat.ratio))) + geom_boxplot()
+  ggplot(aes(as.factor(x = ag.eco9), y = (n2o.sat.ratio))) + geom_boxplot()
+# shows a boxplot for NA category
 ```
@@
 ```{r, geopackage, eval=FALSE, echo=FALSE,}
-library(janitor) # clean_names function
+
 names(dg)
-
-dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, WSA9, WSA9_NAME)
+dg_1 <- clean_names(dg)
 names(dg_1)
-dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") # underscores are "safer" names used in GIS
+# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
+dg_1 <- filter( dg_1, sample_source == "DG")
+
+dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")
+
 names(dg_1)
+str(dg_1)
+
+dg_1 %>% 
+  ggplot(aes(as.factor(x = ag_eco9), y = log(dissolved_n2o_nmol))) + geom_boxplot()
+# now no longer show a boxplot for NA category
+
+dg_1 %>% 
+  ggplot(aes(as.factor(x = ag_eco9), y = (n2o_sat_ratio))) + geom_boxplot()
+# now no longer show a boxplot for NA category
+
+#filter in View returns 514 nitrate_n_result equal to zero
+dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
+  mutate(log_ntl = round(log(ntl_result),3)) %>% 
+  mutate(log_ptl = round(log(ptl_result),3)) %>% 
+  mutate(log_chla = round(log(chla_result),1)) %>% 
+  mutate(log_si = round(log(silica_result),2)) %>% 
+  mutate(log_ca = round(log(calcium_result),2)) %>% 
+  mutate(ph_result = round(ph_result,2))
+# %>% 
+#   mutate(nitrate_n_b = case_when(
+#     nitrate_n_result <= 0.0001 ~ "low",
+#     nitrate_n_result > 0.0001 ~ "high"))  
+# 
+# dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
+# summary(dg_1$nitrate_n_b)
+
 # To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
@@
                   crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # 4326 is just the EPSG identifier of WGS84.
-WGS84 comprises a standard coordinate frame for the Earth, a datum/reference ellipsoid for raw altitude data, and a gravitational equipotential surface (the geoid) that defines the nominal sea level.
-5070 is for Albers equal Area.
+  st_transform(5070) # project to CONUS ALBERS for plotting 
+dim(dg_sf1) # 1184 observations
+
+dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1082 obs vs. 1184 obs vs. previously 1185 obs., why change in sample size
 st_crs(dg_sf1)
-dim(dg_sf1) # 1184 observations
 
@@
 # creating a _v1.gpkg having large sample size and should be in Albers
-st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
-         delete_layer = TRUE) # needed, else function gives error if object already exists
+# st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
+#          delete_layer = TRUE) # needed, else function gives error if object already exists
+st_layers("nladg_v1.gpkg")
+
+# N.B. b/c old layer with untransformed was already in geodatabase have to delete it and then append the new layer
+st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", delete_layer = TRUE, append = TRUE, quiet = FALSE)
 st_layers("nladg_v1.gpkg")
@@
 ## Random Semivariograms
-Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.
-Downloaded Chapter 8 code from https://asdar-book.org/ for generating random variograms.
+Empirical semivariograms were calculated for:  pH, total nitrogen (ntl), total phosphorous (ptl), and chloropyll a (chla).
+In addition, semivariograms were calculated for the dissolved gasses:  DO, N2O, CH4, and CO2.
+We downloaded Chapter 8 code from https://asdar-book.org/ for generating random variograms.
 An initial empirical, or sample, variogram is calculated and plotted for pH, with that object saved.
@@
 That is done 100 times, and the gray lines of the random semivariograms are plotted with the blue line of the empirical semivariogram.
-
-The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation.
+Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.
+
+The water quality variables of pH, tnl, tpl, and chla showe strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.
+The dissolved gasses, DO, N2O, CH4, and CO2 show none to modest spatial autocorrelation.
+The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation over the Euclidean distances that the lakes were sampled.
+CH4 shows some autocorrelation to ~ 1000 km, but with no sill or leveling off the semivariance.
 ```{r, random_semivar}
@@
 dg_1 <- clean_names(dg)
-# also includ WSA9, WSA9_NAME feature plus variables from Jake's 02/25/2020 email
-# added ntl.lab.result, ptl.lab.result on 06/18/2020
-# added physical variables, temperature and depth, on 06/23/2020
-# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
-# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
-# added ph_result 12/14/2020
 names(dg_1)
-dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, wsa9, wsa9_name, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & wsa9 != "AK")
+# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
+dg_1 <- filter( dg_1, sample_source == "DG")
+
+dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")
 
 names(dg_1)
-
-#filter in View returns 514 nitrate_n_result equal to zero
-dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
-  mutate(log_ntl = round(log(ntl_lab_result),3)) %>% 
-  mutate(log_ptl = round(log(ptl_lab_result),3)) %>% 
-  mutate(log_chla = round(log(chla_result_vol),1)) %>% 
-  mutate(log_si = round(log(silica_result),2)) %>% 
-  mutate(log_ca = round(log(calcium_result),2)) %>% 
-  mutate(ph_result = round(ph_result,2)) %>% 
-  mutate(nitrate_n_b = case_when(
-    nitrate_n_result <= 0.0001 ~ "low",
-    nitrate_n_result > 0.0001 ~ "high"))
-
-dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
-summary(dg_1$nitrate_n_b)
+str(dg_1)
 
@@
 st_crs(dg_sf1)
+## pH
+hist(dg_sf1$ph_result)
 
@@
              },
-             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
+             ylim = c(0, 0.8), xlab = 'distance', ylab = 'pH semivariance'
 ))
 
+# used Cressie's robust variogram b/c skewness and not wanting to transform pH as it is already on a log scale
 z_cressie_ph <- variogram(ph_result~1, dg_sf1, cressie = TRUE)
@@
              },
-             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
+             ylim = c(0, 0.8), xlab = 'distance', ylab = 'pH semivariance'
 ))
 
-z_do_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
-plot(z_do_sf1)
+## NTL
+summary(dg_sf1$ntl_result) #no NA's
+
+filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(ntl_result)) + geom_histogram()
+
+# log transformation ~ unimodal & symmetric
+filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(log(ntl_result))) + geom_histogram()
+
+# hist(dg_sf1$ntl_result)
+# hist(log(dg_sf1$ntl_result))
+
+# default cutoff and width and no transformation
+z_lntl_sf1 = variogram(log(ntl_result) ~ 1, dg_sf1)
+plot(z_lntl_sf1, ylab = "NLA NTL Semivariance")
+# set ylim for random semivariograms based on this plot
+
+# cutoff and width from King et al. paper still not working even with tranform
+z1_ntl_king = variogram(log(ntl_result) ~1, dg_sf1, cutoff=2500, width=20)
+plot(z1_ntl_king)
+
 # plot(z_ph_sf1, plot.numbers = TRUE)
-plot(z_do_sf1, ylab = "NLA Surface DO Semivariance")
+# plot(z_ph_sf1, ylab = "NLA ph Semivariance")
+# plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters)", ylab = "NLA ph semivariance")
+# z_ph_plot1 <- plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters) ")
+
+print(xyplot(gamma ~ dist, z_lntl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$ntl_result)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NTL semivariance'
+))
+
+## PTL
+summary(dg_sf1$ptl_result) # no NA's
+
+filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(ptl_result)) + geom_histogram()
+
+# log transformation ~ unimodal & symmetric
+filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(log(ptl_result))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_lptl_sf1 = filter(dg_sf1, (!is.na(ptl_result))) %>% variogram(log(ptl_result)~1, .)
+plot(z_lptl_sf1, ylab = "NLA TP Semivariance")
+# set ylim of random semivariograms based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's
+# dg_sf2 <- filter(dg_sf1, (!is.na(ptl_result)))
+
+print(xyplot(gamma ~ dist, z_lptl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$ptl_result)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 2), xlab = 'distance', ylab = 'TP semivariance'
+))
+
+## CHLA
+summary(dg_sf1$chla_result) # no NA's
+
+filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(chla_result)) + geom_histogram()
+
+# log transformation not that great but use for now
+filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(log(chla_result))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_lchla_sf1 = filter(dg_sf1, (!is.na(chla_result))) %>% variogram(log(chla_result)~1, .)
+plot(z_lchla_sf1, ylab = "NLA CHLA Semivariance")
+# set ylim of random semivariograms based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's
+# dg_sf2 <- filter(dg_sf1, (!is.na(chla_result)))
+
+print(xyplot(gamma ~ dist, z_lchla_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$chla_result)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 2.5), xlab = 'distance', ylab = 'CHLa semivariance'
+))
+
+## DO
+summary(dg_sf1$o2_surf) # have 4 NA's
+
+filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(o2_surf)) + geom_histogram()
+
+# log probably too strong of transformation but use for now
+filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(log(o2_surf))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_ldo_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
+# plot(z_ldo_sf1)
+# plot(z_ph_sf1, plot.numbers = TRUE)
+plot(z_ldo_sf1, ylab = "NLA Surface DO Semivariance")
+
+# data frame with no NA's
+dg_sf2 <- filter(dg_sf1, (!is.na(o2_surf)))
+
+print(xyplot(gamma ~ dist, z_ldo_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf2$random = sample(dg_sf2$o2_surf)
+                     v = variogram(log(random) ~ 1, dg_sf2)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 0.12), xlab = 'distance', ylab = 'NLA DO semivariance'
+))
+
+## N2O
+summary(dg_sf1$dissolved_n2o_nmol) # no NA's
+
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()
+
+# unimodal, but not symmetrical
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()
+
+z_n2o_sf1 <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol) ~1, .)
+plot(z_n2o_sf1, ylab = "N2O Semivariance")
+
+# change cutoff and width
+z_n2o = filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol)~1, ., cutoff = 750000, width = 25000)
+# plot(z_n2o, ylab = "N2O Semivariance", plot.numbers = TRUE)
+plot(z_n2o, ylab = "NLA N2O Semivariance")
+
+print(xyplot(gamma ~ dist, z_n2o, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+    panel = function(x, y, ...) {
+        for (i in 1:100) {
+            dg_sf1$random = sample(dg_sf1$dissolved_n2o_nmol)
+            v = variogram(log(random) ~ 1, dg_sf1, cutoff = 750000, width = 25000)
+            llines(v$dist, v$gamma, col = 'grey')
+        }
+        panel.xyplot(x, y, ...)
+    },
+    ylim = c(0, 0.35), xlab = 'distance', ylab = 'N2O semivariance'
+))
+
+## CH4
+summary(dg_1$dissolved_ch4_umo) # no NA's
+
+filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(dissolved_ch4_umol)) + geom_histogram()
+
+# log transformation looks good, ~ unimodal & symmetric
+filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(log(dissolved_ch4_umol))) + geom_histogram()
+
+z_lch4_sf1 = filter(dg_sf1, (!is.na(dissolved_ch4_umol))) %>% variogram(log(dissolved_ch4_umol)~1, .)
+plot(z_lch4_sf1, ylab = "NLA CH4 Semivariance")
+# change ylim below in random semivariogram based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's for random sampling below
+# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_ch4_umol)))
+
+print(xyplot(gamma ~ dist, z_lch4_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$dissolved_ch4_umol)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 2), xlab = 'distance', ylab = 'NLA CH4 semivariance'
+))
+
+## CO2
+summary(dg_1$dissolved_co2_umo) # no NA's
+
+filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(dissolved_co2_umol)) + geom_histogram()
+
+# log transformation, unimodal and symmetric
+filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(log(dissolved_co2_umol))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_lco2_sf1 = filter(dg_sf1, (!is.na(dissolved_co2_umol))) %>% variogram(log(dissolved_co2_umol)~1, .)
+plot(z_lco2_sf1, ylab = "NLA CO2 Semivariance")
+# change ylim in random semivariogram based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's
+# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_co2_umol)))
+
+print(xyplot(gamma ~ dist, z_lco2_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$dissolved_co2_umol)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NLA CO2 semivariance'
+))
 
@@
 library(corrr)
-dg %>%
-  select(dissolved.n2o.nmol, MeanDUsed, ph.result, nitrate.n.result, ntl.lab.result, chla.result.vol, ptl.lab.result, surftemp) %>%
+dg_1 %>%
+  select(index_site_depth, dissolved_n2o_nmol, ph_result, nitrate_n_result, ntl_result, chla_result, ptl_result, temp_surf) %>%
   correlate()

### 2021-02-16 scripts/dataMunge.Rmd

--- 2021-01-14 scripts/dataMunge.Rmd
@@
                                "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2021-01-13.txt"),
+                               "inputData/nla17gasDataAggregated_2021-02-01.txt"),
                  header = TRUE, sep = "\t", as.is = TRUE) %>%

### 2021-03-09 scripts/dgIndicatorAnalysis.Rmd
Added text describing random semivariogram and did random semivariogram by ecoregion.

@@
 
-The water quality variables of pH, tnl, tpl, and chla showe strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.
+The water quality variables of pH, tnl, tpl, and chla show strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.
 The dissolved gasses, DO, N2O, CH4, and CO2 show none to modest spatial autocorrelation.
@@
 # dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
+
 dg_1 <- filter( dg_1, sample_source == "DG")
-
 dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")
@@
              ylim = c(0, 1.2), xlab = 'distance', ylab = 'NLA CO2 semivariance'
+))
+
+```
+
+## Random Ecoregional Semivariogram N2O for ASLO Abstract
+I used ecoregional-specific data frames to feed into random semivariogram function.  
+```{r, ecoregion_semivar}
+
+## N2O
+summary(dg_sf1$dissolved_n2o_nmol) # no NA's
+
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()
+
+# unimodal, but not symmetrical
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER") %>%  ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()
+
+# zcpl <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="CPL")
+# zcpl_variog <- variogram(log(dissolved_n2o_nmol) ~1, zcpl)
+# plot(zcpl_variog, ylab = "N2O zCPL Semivariance")
+# the 2-step process above matches the piped process below
+xer_n2o_sf1 <- filter(dg_sf1,!is.na(dissolved_n2o_nmol)) %>%
+  filter(ag_eco9 =="XER") %>% 
+  variogram(log(dissolved_n2o_nmol) ~1, .)
+plot(xer_n2o_sf1, ylab = "N2O XER Semivariance")
+
+xer <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER")
+
+print(xyplot(gamma ~ dist, wmt_n2o_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+    panel = function(x, y, ...) {
+        for (i in 1:100) {
+            xer$random = sample(xer$dissolved_n2o_nmol)
+            v = variogram(log(random) ~ 1, xer)
+            llines(v$dist, v$gamma, col = 'grey')
+        }
+        panel.xyplot(x, y, ...)
+    },
+    ylim = c(0, 0.35), xlab = 'distance', ylab = 'XER N2O semivariance'
 ))

### 2021-04-05 scripts/Mung_NLA_Sample_Frame.Rmd
Import and munge NLA sample frame

---
title: "Create NLA Sample Frame"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook: 
    toc: yes
    toc_depth: 4
    toc_float: true
    code_folding: show
    font_size: 14
    number_sections: true
    theme: simplex
---
# Background
Steps to import, mung, and save NLA sample frame.

# Setup R
```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}
```

# Mung and save to file
```{r import_sample_frame, echo=TRUE, paged.print=TRUE}
sframe <- read.csv( file = paste0( localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/NLA_Sample_Frame.csv" ), header = T )

sframe <- sframe %>%
  filter(nla17_sf != "Exclude2017") %>%
  filter(nla17_sf != "Exclude2017_Include2017NH") %>%
  mutate(WSA9 = factor(ag_eco9),
          state = factor(state),
          size_cat = factor(area_cat6)) %>% 
  select(WSA9, state, size_cat, area_ha, perim_km) %>% 
  filter(state != "DC") %>%
  filter(state != "HI") %>%
  droplevels() %>%
  mutate(WSA9 = fct_drop(WSA9)) %>% # remove NA level
  #filter( size_cat != "(0,1]" ) %>%
  mutate(size_cat = recode(size_cat, 
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  mutate(size_cat = factor(size_cat, 
                           levels = c("min_4", "4_10", "10_20", "20_50", "50_max"),
                           ordered = TRUE))

save(sframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

print(sframe)
```

### 2021-05-21 scripts/dgIndicatorAnalysis.Rmd
reconciling merge conflict

--- 2021-02-01 scripts/dgIndicatorAnalysis.Rmd
@@
 dg %>% 
-  ggplot(aes(as.factor(x = WSA9), y = log(dissolved.n2o.nmol))) + geom_boxplot()
+  ggplot(aes(as.factor(x = ag.eco9.nm), y = log(dissolved.n2o.nmol))) + geom_boxplot()
 
 dg %>% 
-  ggplot(aes(as.factor(x = WSA9), y = (n2o.sat.ratio))) + geom_boxplot()
+  ggplot(aes(as.factor(x = ag.eco9.nm), y = (n2o.sat.ratio))) + geom_boxplot()
 ```
@@
 
-dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, WSA9, WSA9_NAME)
+dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, ag.eco9.nm, ag.eco9.nm)
 
 names(dg_1)
-dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") %>% 
+dg_1 <- clean_names(dg_1) %>% filter(ag.eco9.nm != "AK") %>% 
   mutate(log_n2o = round(log(dissolved_n2o),3)) %>% 
@@
 ecoR1 <- st_transform(ecoR, 5070) %>% 
-  select(OBJECTID, WSA9, WSA9_NAME) # Grab variables of interest
+  select(OBJECTID, ag.eco9.nm, ag.eco9.nm) # Grab variables of interest
 st_crs(ecoR1)
@@
 names(dg.sf)
-head(dg.sf$WSA9_NAME)
+head(dg.sf$ag.eco9.nm)
 # distinct returns all unique geometry + attribute combinations.  convert to df if not interested in geometry
-dg.sf %>% as.data.frame() %>% distinct(WSA9_NAME) 
-# distinct.Spatial(dg.sf, WSA9_NAME) # this doesn't work either.
-distinct(dg, WSA9_NAME)
+dg.sf %>% as.data.frame() %>% distinct(ag.eco9.nm) 
+# distinct.Spatial(dg.sf, ag.eco9.nm) # this doesn't work either.
+distinct(dg, ag.eco9.nm)
 
@@
 class(ecoR)
-# contrast ecoRname <- as.data.frame(select(ecoR,"WSA9_NAME"))
+# contrast ecoRname <- as.data.frame(select(ecoR,"ag.eco9.nm"))
 # geom is sticky with select but not pull
-# ecoRname <- as.data.frame(pull(ecoR,"WSA9_NAME"))
-ecoRname1 <- as.character(pull(ecoR,"WSA9_NAME")) #Springsteen code needs to be character to populate columns of matrix below
+# ecoRname <- as.data.frame(pull(ecoR,"ag.eco9.nm"))
+ecoRname1 <- as.character(pull(ecoR,"ag.eco9.nm")) #Springsteen code needs to be character to populate columns of matrix below
 class(ecoRname1)
@@
 
-count(dg.sf, WSA9_NAME) %>% print(n = Inf)
+count(dg.sf, ag.eco9.nm) %>% print(n = Inf)
 
 # Get ecoregion-specific NLA sites
-cpl <- dg.sf %>% filter(WSA9_NAME == "Coastal Plains")
+cpl <- dg.sf %>% filter(ag.eco9.nm == "Coastal Plains")
 
@@
 # nap # nap object doesn't exist
-nap <- dg.sf %>% filter(WSA9_NAME == "Northern Appalachians")
+nap <- dg.sf %>% filter(ag.eco9.nm == "Northern Appalachians")
 nap_sp <- as(nap, "Spatial")
@@
 # npl 
-npl <- dg.sf %>% filter(WSA9_NAME == "Northern Plains")
+npl <- dg.sf %>% filter(ag.eco9.nm == "Northern Plains")
 npl_sp <- as(npl, "Spatial")
@@
 # sap
-sap <- dg.sf %>% filter(WSA9_NAME == "Southern Appalachians")
+sap <- dg.sf %>% filter(ag.eco9.nm == "Southern Appalachians")
 sap_sp <- as(sap, "Spatial")
@@
 # spl
-spl <- dg.sf %>% filter(WSA9_NAME == "Southern Plains")
+spl <- dg.sf %>% filter(ag.eco9.nm == "Southern Plains")
 spl_sp <- as(spl, "Spatial")
@@
 # tpl
-tpl <- dg.sf %>% filter(WSA9_NAME == "Temperate Plains")
+tpl <- dg.sf %>% filter(ag.eco9.nm == "Temperate Plains")
 tpl_sp <- as(tpl, "Spatial")
@@
 # umw
-umw <- dg.sf %>% filter(WSA9_NAME == "Upper Midwest")
+umw <- dg.sf %>% filter(ag.eco9.nm == "Upper Midwest")
 umw_sp <- as(umw, "Spatial")
@@
 # wmt
-wmt <- dg.sf %>% filter(WSA9_NAME == "Western Mountains")
+wmt <- dg.sf %>% filter(ag.eco9.nm == "Western Mountains")
 wmt_sp <- as(wmt, "Spatial")
@@
 # xer
-xer <- dg.sf %>% filter(WSA9_NAME == "Xeric")
+xer <- dg.sf %>% filter(ag.eco9.nm == "Xeric")
 xer_sp <- as(xer, "Spatial")
@@
 names(dg)
-# also includ WSA9, WSA9_NAME feature
-dg_1 <- select(dg,site.id:dissolved.n2o.nmol, WSA9, WSA9_NAME)
+# also includ ag.eco9.nm, ag.eco9.nm feature
+dg_1 <- dg
 names(dg_1)
@@
 names(dg_1)
-distinct(dg_1, wsa9)
+distinct(dg_1, ag_eco9_nm)
 # To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
@@
 ## CPL
-cpl <- dg_sf1 %>% filter(wsa9 == "CPL" & visit_no == 1 & sample_source == "DG")
+cpl <- dg_sf1 %>% filter(ag_eco9 == "CPL" & visit_no == 1 & sample_source == "DG")
 
@@
 ## NAP
-nap <- dg_sf1 %>% filter(wsa9 == "NAP" & visit_no == 1 & sample_source == "DG")
+nap <- dg_sf1 %>% filter(ag_eco9_nm == "NAP" & visit_no == 1 & sample_source == "DG")
 
@@
 ## NPL
-npl <- dg_sf1 %>% filter(wsa9 == "NPL" & visit_no == 1 & sample_source == "DG")
+npl <- dg_sf1 %>% filter(ag_eco9_nm == "NPL" & visit_no == 1 & sample_source == "DG")
 
@@
 ## SAP
-sap <- dg_sf1 %>% filter(wsa9 == "SAP" & visit_no == 1 & sample_source == "DG")
+sap <- dg_sf1 %>% filter(ag_eco9_nm == "SAP" & visit_no == 1 & sample_source == "DG")
 
@@
 ## SPL
-spl <- dg_sf1 %>% filter(wsa9 == "SPL" & visit_no == 1 & sample_source == "DG")
+spl <- dg_sf1 %>% filter(ag_eco9_nm == "SPL" & visit_no == 1 & sample_source == "DG")
 
@@
 ## TPL
-tpl <- dg_sf1 %>% filter(wsa9 == "TPL" & visit_no == 1 & sample_source == "DG")
+tpl <- dg_sf1 %>% filter(ag_eco9_nm == "TPL" & visit_no == 1 & sample_source == "DG")
 
@@
 ## UMW
-umw <- dg_sf1 %>% filter(wsa9 == "UMW" & visit_no == 1 & sample_source == "DG")
+umw <- dg_sf1 %>% filter(ag_eco9_nm == "UMW" & visit_no == 1 & sample_source == "DG")
 
@@
 ## WMT
-wmt <- dg_sf1 %>% filter(wsa9 == "WMT" & visit_no == 1 & sample_source == "DG")
+wmt <- dg_sf1 %>% filter(ag_eco9_nm == "WMT" & visit_no == 1 & sample_source == "DG")
 
@@
 ## XER
-xer <- dg_sf1 %>% filter(wsa9 == "XER" & visit_no == 1 & sample_source == "DG")
+xer <- dg_sf1 %>% filter(ag_eco9_nm == "XER" & visit_no == 1 & sample_source == "DG")
 
@@
 names(dg_sf1)
-nap_visit1 <- dg_sf1 %>% filter(wsa9 == "NAP" & visit_no == 1)
+nap_visit1 <- dg_sf1 %>% filter(ag_eco9_nm == "NAP" & visit_no == 1)
 class(nap_visit1)
@@
 names(dg_1)
-dg_cplvisit1 <- dg_1 %>% filter(wsa9 == "CPL" & visit_no == 1)
+dg_cplvisit1 <- dg_1 %>% filter(ag_eco9 == "CPL" & visit_no == 1)
 # N.B. default is resamp = 1000 so that can take 10 minutes for n =1090 sites
@@
 # CPL outlier check by working through all 10 pairwise removal of cluster of 5 high Delaware sites
-out_cplvisit1 <- dg_1 %>% filter(wsa9 == "CPL" & visit_no == 1 & (site_id != "NLA17_DE-10003" & site_id != "NLA17_DE-10004"))
+out_cplvisit1 <- dg_1 %>% filter(ag_eco9 == "CPL" & visit_no == 1 & (site_id != "NLA17_DE-10003" & site_id != "NLA17_DE-10004"))
 
@@
 # Uses normality assumption, which is faster, to assess significance
-dg_napvisit1 <- dg_1 %>% filter(wsa9 == "NAP" & visit_no == 1)
+dg_napvisit1 <- dg_1 %>% filter(ag_eco9 == "NAP" & visit_no == 1)
 coords_nap <- cbind(dg_napvisit1$map_lon_dd, dg_napvisit1$map_lat_dd)
@@
 
-dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o.nmol, WSA9, WSA9_NAME, ONOFFNET, nitrate.n.result,chla.result.vol, max.bf, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size.cat)
+dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o.nmol, ag.eco9.nm, ag.eco9.nm, ONOFFNET, nitrate.n.result,chla.result.vol, max.bf, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size.cat)
 names(dg_1)
@@
 
-## Perform the same test for each wsa9 using Z = "log_n2o"
+## Perform the same test for each ag.eco9.nm using Z = "log_n2o"
 ## Start by defining a function
@@
 
-## Split the dataframe by each wsa9: turns the data frame into a list 
-##   where each list item is a data frame filtered to that wsa9
+## Split the dataframe by each ag.eco9.nm: turns the data frame into a list 
+##   where each list item is a data frame filtered to that ag.eco9.nm
 ## lapply applies the function ncf_correlog on each list object (a data frame)
@@
 log_n2o_correlog_list <- dg_visit1 %>%
-  split(.$wsa9) %>% # '.' means use the data frame that preceded this line
+  split(.$ag_eco9_nm) %>% # '.' means use the data frame that preceded this line
   lapply(function(x) ncf_correlog(dat = x, z_col = "log_n2o")) 
@@
 
-## combine into one dataframe to grab the wsa9 (notice the name no longer has an s at the end)
+## combine into one dataframe to grab the ag.eco9.nm (notice the name no longer has an s at the end)
 
 log_n2o_correlog_dat <- log_n2o_correlog_dats %>%
-  dplyr::bind_rows(., .id = "wsa9")
+  dplyr::bind_rows(., .id = "ag.eco9.nm")
 
@@
 log_n2o_correlog_plots <- log_n2o_correlog_dat %>%
-  split(.$wsa9) %>% # splits the data frame by wsa9, which creates a list with 9 data frames
-  lapply(function(x) plot_correlog(ncf_correlog = x, # use the data.frame for each wsa9 to make a plot
+  split(.$ag.eco9.nm) %>% # splits the data frame by ag.eco9.nm, which creates a list with 9 data frames
+  lapply(function(x) plot_correlog(ncf_correlog = x, # use the data.frame for each ag.eco9.nm to make a plot
                                    plot_title = NULL,
@@
                                    pval_col = "pval",
-                                   name_col = "wsa9",
+                                   name_col = "ag.eco9.nm",
                                    title_size = 14, 
@@
 ## alternatively, can make a facet plot since log_n20_correlog_dat is a dataframe in the correct format
-names(log_n2o_correlog_dat) #"n"    "mean" "cor"  "xint" "pval" "wsa9"
+names(log_n2o_correlog_dat) #"n"    "mean" "cor"  "xint" "pval" "ag.eco9.nm"
 
@@
 head(log_n2o_correlog_dat)
-## can change wsa9 to a factor to arrange what order the plots are shown, otherwise default arranges them alphabetically
-# log_n20_correlog_dat$wsa9 <- factor(log_n20_correlog_dat, 
+## can change ag.eco9.nm to a factor to arrange what order the plots are shown, otherwise default arranges them alphabetically
+# log_n20_correlog_dat$ag.eco9.nm <- factor(log_n20_correlog_dat, 
 #                                     levels = c("CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER"))
@@
                      guide=FALSE) +
-  facet_wrap(~wsa9, ncol=3, scales="free") +
+  facet_wrap(~ag.eco9.nm, ncol=3, scales="free") +
   labs(x="Distance (mean-of-class)", y="Correlation") +
@@
 dg_1 <- clean_names(dg)
-# also includ WSA9, WSA9_NAME feature plus variables from Jake's 02/25/2020 email
+# also includ ag.eco9.nm, ag.eco9.nm feature plus variables from Jake's 02/25/2020 email
 # added ntl.lab.result, ptl.lab.result on 06/18/2020
@@
 names(dg_1)
-dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, wsa9, wsa9_name, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & wsa9 != "AK")
+dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag.eco9.nm, ag.eco9.nm, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag.eco9.nm != "AK")
 
@@
 library(tree)
-model <- tree(dissolved.n2o.nmol ~ MeanDUsed + nitrate.n.result + ntl.lab.result + chla.result.vol + ptl.lab.result + surftemp + WSA9, data = dg)
+model <- tree(dissolved.n2o.nmol ~ MeanDUsed + nitrate.n.result + ntl.lab.result + chla.result.vol + ptl.lab.result + surftemp + ag.eco9.nm, data = dg)
 

### 2021-05-21 scripts/dgIndicatorAnalysis.Rmd

@@
 ```{r}
-# Read data file 
+# Load data
  load(paste0(localPath,
@@
 
+names(dg)
+
 dg %>% 
@@
   ggplot(aes(as.factor(x = ag.eco9.nm), y = (n2o.sat.ratio))) + geom_boxplot()
+
 ```
@@
 ```{r, geopackage, eval=FALSE, echo=FALSE,}
-library(janitor) # clean_names function
+
 names(dg)
-
-dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, ag.eco9.nm, ag.eco9.nm)
-
+dg_1 <- clean_names(dg)
 names(dg_1)
-dg_1 <- clean_names(dg_1) %>% filter(ag.eco9.nm != "AK") %>% 
-  mutate(log_n2o = round(log(dissolved_n2o),3)) %>% 
-  mutate(log_n2osatratio = round(log(n2o_sat_ratio),3))
-# underscores are "safer" names used in GIS
+# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
+dg_1 <- filter( dg_1, sample_source == "DG")
+
+dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")
 
 names(dg_1)
+str(dg_1)
+
+dg_1 %>% 
+  ggplot(aes(as.factor(x = ag_eco9), y = log(dissolved_n2o_nmol))) + geom_boxplot()
+# now no longer show a boxplot for NA category
+
+dg_1 %>% 
+  ggplot(aes(as.factor(x = ag_eco9), y = (n2o_sat_ratio))) + geom_boxplot()
+# now no longer show a boxplot for NA category
+
+#filter in View returns 514 nitrate_n_result equal to zero
+dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
+  mutate(log_ntl = round(log(ntl_result),3)) %>% 
+  mutate(log_ptl = round(log(ptl_result),3)) %>% 
+  mutate(log_chla = round(log(chla_result),1)) %>% 
+  mutate(log_si = round(log(silica_result),2)) %>% 
+  mutate(log_ca = round(log(calcium_result),2)) %>% 
+  mutate(ph_result = round(ph_result,2))
+# %>% 
+#   mutate(nitrate_n_b = case_when(
+#     nitrate_n_result <= 0.0001 ~ "low",
+#     nitrate_n_result > 0.0001 ~ "high"))  
+# 
+# dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
+# summary(dg_1$nitrate_n_b)
+
 # To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
@@
                   crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # 4326 is just the EPSG identifier of WGS84.
-WGS84 comprises a standard coordinate frame for the Earth, a datum/reference ellipsoid for raw altitude data, and a gravitational equipotential surface (the geoid) that defines the nominal sea level.
-5070 is for Albers equal Area.
-st_crs(dg_sf1)
+  st_transform(5070) # project to CONUS ALBERS for plotting 
 dim(dg_sf1) # 1184 observations
 
-dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1184 obs vs. previously 1185 obs., why change in sample size
+dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1082 obs vs. 1184 obs vs. previously 1185 obs., why change in sample size
 st_crs(dg_sf1)
@@
 # creating a _v1.gpkg having large sample size and should be in Albers
-st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
-         delete_layer = TRUE) # needed, else function gives error if object already exists
+# st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
+#          delete_layer = TRUE) # needed, else function gives error if object already exists
 st_layers("nladg_v1.gpkg")
@@
 st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", delete_layer = TRUE, append = TRUE, quiet = FALSE)
-st_layers("nla_dg1.gpkg")
+st_layers("nladg_v1.gpkg")
 
@@
 That is done 100 times, and the gray lines of the random semivariograms are plotted with the blue line of the empirical semivariogram.
-
-The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation.
+Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.
+
+The water quality variables of pH, tnl, tpl, and chla show strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.
+The dissolved gasses, DO, N2O, CH4, and CO2 show none to modest spatial autocorrelation.
+The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation over the Euclidean distances that the lakes were sampled.
+CH4 shows some autocorrelation to ~ 1000 km, but with no sill or leveling off the semivariance.
 ```{r, random_semivar}
@@
 dg_1 <- clean_names(dg)
+
 # also includ ag.eco9.nm, ag.eco9.nm feature plus variables from Jake's 02/25/2020 email
@@
 names(dg_1)
-dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag.eco9.nm, ag.eco9.nm, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag.eco9.nm != "AK")
+# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
+
+dg_1 <- filter( dg_1, sample_source == "DG")
+dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")
 
 names(dg_1)
-
-#filter in View returns 514 nitrate_n_result equal to zero
-dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
-  mutate(log_ntl = round(log(ntl_lab_result),3)) %>% 
-  mutate(log_ptl = round(log(ptl_lab_result),3)) %>% 
-  mutate(log_chla = round(log(chla_result_vol),1)) %>% 
-  mutate(log_si = round(log(silica_result),2)) %>% 
-  mutate(log_ca = round(log(calcium_result),2)) %>% 
-  mutate(ph_result = round(ph_result,2)) %>% 
-  mutate(nitrate_n_b = case_when(
-    nitrate_n_result <= 0.0001 ~ "low",
-    nitrate_n_result > 0.0001 ~ "high"))
-
-dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
-summary(dg_1$nitrate_n_b)
+str(dg_1)
 
@@
              },
-             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
+             ylim = c(0, 0.8), xlab = 'distance', ylab = 'pH semivariance'
 ))
 
-z_do_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
-plot(z_do_sf1)
+## NTL
+summary(dg_sf1$ntl_result) #no NA's
+
+filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(ntl_result)) + geom_histogram()
+
+# log transformation ~ unimodal & symmetric
+filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(log(ntl_result))) + geom_histogram()
+
+# hist(dg_sf1$ntl_result)
+# hist(log(dg_sf1$ntl_result))
+
+# default cutoff and width and no transformation
+z_lntl_sf1 = variogram(log(ntl_result) ~ 1, dg_sf1)
+plot(z_lntl_sf1, ylab = "NLA NTL Semivariance")
+# set ylim for random semivariograms based on this plot
+
+# cutoff and width from King et al. paper still not working even with tranform
+z1_ntl_king = variogram(log(ntl_result) ~1, dg_sf1, cutoff=2500, width=20)
+plot(z1_ntl_king)
+
 # plot(z_ph_sf1, plot.numbers = TRUE)
-plot(z_do_sf1, ylab = "NLA Surface DO Semivariance")
+# plot(z_ph_sf1, ylab = "NLA ph Semivariance")
+# plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters)", ylab = "NLA ph semivariance")
+# z_ph_plot1 <- plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters) ")
+
+print(xyplot(gamma ~ dist, z_lntl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$ntl_result)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NTL semivariance'
+))
+
+## PTL
+summary(dg_sf1$ptl_result) # no NA's
+
+filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(ptl_result)) + geom_histogram()
+
+# log transformation ~ unimodal & symmetric
+filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(log(ptl_result))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_lptl_sf1 = filter(dg_sf1, (!is.na(ptl_result))) %>% variogram(log(ptl_result)~1, .)
+plot(z_lptl_sf1, ylab = "NLA TP Semivariance")
+# set ylim of random semivariograms based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's
+# dg_sf2 <- filter(dg_sf1, (!is.na(ptl_result)))
+
+print(xyplot(gamma ~ dist, z_lptl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$ptl_result)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 2), xlab = 'distance', ylab = 'TP semivariance'
+))
+
+## CHLA
+summary(dg_sf1$chla_result) # no NA's
+
+filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(chla_result)) + geom_histogram()
+
+# log transformation not that great but use for now
+filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(log(chla_result))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_lchla_sf1 = filter(dg_sf1, (!is.na(chla_result))) %>% variogram(log(chla_result)~1, .)
+plot(z_lchla_sf1, ylab = "NLA CHLA Semivariance")
+# set ylim of random semivariograms based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's
+# dg_sf2 <- filter(dg_sf1, (!is.na(chla_result)))
+
+print(xyplot(gamma ~ dist, z_lchla_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$chla_result)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 2.5), xlab = 'distance', ylab = 'CHLa semivariance'
+))
+
+## DO
+summary(dg_sf1$o2_surf) # have 4 NA's
+
+filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(o2_surf)) + geom_histogram()
+
+# log probably too strong of transformation but use for now
+filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(log(o2_surf))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_ldo_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
+# plot(z_ldo_sf1)
+# plot(z_ph_sf1, plot.numbers = TRUE)
+plot(z_ldo_sf1, ylab = "NLA Surface DO Semivariance")
+
+# data frame with no NA's
+dg_sf2 <- filter(dg_sf1, (!is.na(o2_surf)))
+
+print(xyplot(gamma ~ dist, z_ldo_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf2$random = sample(dg_sf2$o2_surf)
+                     v = variogram(log(random) ~ 1, dg_sf2)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 0.12), xlab = 'distance', ylab = 'NLA DO semivariance'
+))
+
+## N2O
+summary(dg_sf1$dissolved_n2o_nmol) # no NA's
+
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()
+
+# unimodal, but not symmetrical
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()
+
+z_n2o_sf1 <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol) ~1, .)
+plot(z_n2o_sf1, ylab = "N2O Semivariance")
+
+# change cutoff and width
+z_n2o = filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol)~1, ., cutoff = 750000, width = 25000)
+# plot(z_n2o, ylab = "N2O Semivariance", plot.numbers = TRUE)
+plot(z_n2o, ylab = "NLA N2O Semivariance")
+
+print(xyplot(gamma ~ dist, z_n2o, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+    panel = function(x, y, ...) {
+        for (i in 1:100) {
+            dg_sf1$random = sample(dg_sf1$dissolved_n2o_nmol)
+            v = variogram(log(random) ~ 1, dg_sf1, cutoff = 750000, width = 25000)
+            llines(v$dist, v$gamma, col = 'grey')
+        }
+        panel.xyplot(x, y, ...)
+    },
+    ylim = c(0, 0.35), xlab = 'distance', ylab = 'N2O semivariance'
+))
+
+## CH4
+summary(dg_1$dissolved_ch4_umo) # no NA's
+
+filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(dissolved_ch4_umol)) + geom_histogram()
+
+# log transformation looks good, ~ unimodal & symmetric
+filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(log(dissolved_ch4_umol))) + geom_histogram()
+
+z_lch4_sf1 = filter(dg_sf1, (!is.na(dissolved_ch4_umol))) %>% variogram(log(dissolved_ch4_umol)~1, .)
+plot(z_lch4_sf1, ylab = "NLA CH4 Semivariance")
+# change ylim below in random semivariogram based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's for random sampling below
+# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_ch4_umol)))
+
+print(xyplot(gamma ~ dist, z_lch4_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$dissolved_ch4_umol)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 2), xlab = 'distance', ylab = 'NLA CH4 semivariance'
+))
+
+## CO2
+summary(dg_1$dissolved_co2_umo) # no NA's
+
+filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(dissolved_co2_umol)) + geom_histogram()
+
+# log transformation, unimodal and symmetric
+filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(log(dissolved_co2_umol))) + geom_histogram()
+
+# hist(log(dg_sf1$ntl_result))
+
+z_lco2_sf1 = filter(dg_sf1, (!is.na(dissolved_co2_umol))) %>% variogram(log(dissolved_co2_umol)~1, .)
+plot(z_lco2_sf1, ylab = "NLA CO2 Semivariance")
+# change ylim in random semivariogram based on this plot
+# plot(z_ph_sf1, plot.numbers = TRUE)
+
+# data frame with no NA's
+# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_co2_umol)))
+
+print(xyplot(gamma ~ dist, z_lco2_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+             panel = function(x, y, ...) {
+                 for (i in 1:100) {
+                     dg_sf1$random = sample(dg_sf1$dissolved_co2_umol)
+                     v = variogram(log(random) ~ 1, dg_sf1)
+                     llines(v$dist, v$gamma, col = 'grey')
+                 }
+                 panel.xyplot(x, y, ...)
+             },
+             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NLA CO2 semivariance'
+))
+
+```
+
+## Random Ecoregional Semivariogram N2O for ASLO Abstract
+I used ecoregional-specific data frames to feed into random semivariogram function.  
+```{r, ecoregion_semivar}
+
+## N2O
+summary(dg_sf1$dissolved_n2o_nmol) # no NA's
+
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()
+
+# unimodal, but not symmetrical
+filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER") %>%  ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()
+
+# zcpl <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="CPL")
+# zcpl_variog <- variogram(log(dissolved_n2o_nmol) ~1, zcpl)
+# plot(zcpl_variog, ylab = "N2O zCPL Semivariance")
+# the 2-step process above matches the piped process below
+xer_n2o_sf1 <- filter(dg_sf1,!is.na(dissolved_n2o_nmol)) %>%
+  filter(ag_eco9 =="XER") %>% 
+  variogram(log(dissolved_n2o_nmol) ~1, .)
+plot(xer_n2o_sf1, ylab = "N2O XER Semivariance")
+
+xer <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER")
+
+print(xyplot(gamma ~ dist, wmt_n2o_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
+    panel = function(x, y, ...) {
+        for (i in 1:100) {
+            xer$random = sample(xer$dissolved_n2o_nmol)
+            v = variogram(log(random) ~ 1, xer)
+            llines(v$dist, v$gamma, col = 'grey')
+        }
+        panel.xyplot(x, y, ...)
+    },
+    ylim = c(0, 0.35), xlab = 'distance', ylab = 'XER N2O semivariance'
+))
 

### 2021-06-09 scripts/dgIndicatorAnalysis.Rmd
Added micromap package and linked micromap code

@@
 library(lattice) # xyplot
+library(micromap) # linked micromaps
 
@@
 
+## Linked micromaps
+The ecoreg_stats summary is for population estimates.
+The code at the bottom, ecoreg_pred, I believe gives summary stats for the predicted distributions, but I would double-check with Roy.
+There is a quick-and-dirty micromap code used to look trying linked micromap plots on different variables.
+When you settle on the variables you want to use then the full micromap code produces the pretty linked micromap plots.
+```{r, linked_micromap}
+
+# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
+# this is analogous to spsurvey "population estimates"
+ecoreg_stats <- all_predictions %>%
+  group_by(WSA9, .draw) %>%
+  summarise( mean_n2o = mean(preds_n2o),
+             mean_n2oeq = mean(preds_n2oeq),
+             mean_sat = mean(preds_sat))%>%
+  summarise( post_med_n2o = round(median(mean_n2o), 1),
+    LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
+    UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
+     post_med_n2oeq = round(median(mean_n2oeq), 2),
+    LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
+    UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
+     post_med_sat = round(median(mean_sat), 2),
+    LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
+    UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
+    n = dplyr::n())
+
+ecoreg_stats <- as.data.frame(ecoreg_stats)
+class(ecoreg_stats)
+str(ecoreg_stats) # WSA9 is a factor, but some of the stats have attribute of Named num
+
+ecoreg_stats$WSA9 <- as.character(ecoreg_stats$WSA9)
+str(ecoreg_stats)
+
+ecoreg_stats
+
+ecoregions <- st_read("aggr_ecoregions_simple.shp")
+wsa9names <- st_set_geometry(ecoregions, NULL) %>%
+  select(WSA9, WSA9_NAME)
+
+ecoreg_stats <- left_join(ecoreg_stats, wsa9names)
+
+ecoregions <- st_transform(ecoregions, 5070)
+# make spatial polygon dataframe need for micromap
+ecoregions_sp1 <- as(ecoregions, "Spatial")
+
+# convert to map_table for linked micromap
+ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
+class(ecor.polys)
+str(ecor.polys, max =2) # ID is chr "CPL"...
+
+# Quick & Dirty micromap code:  use this to test out micromap plot of different summary statistics to get their x-axis values
+mmplot(stat.data=ecoreg_stats,
+       map.data=ecor.polys,
+       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
+       panel.data=list(
+         NA,
+         'WSA9_NAME',
+         list('post_med_n2o', 'LCI_n2o', 'UCI_n2o'),
+         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
+         NA),
+       ord.by='post_med_n2o', grouping=3,
+       median.row=FALSE,
+       map.link=c('WSA9', 'ID'))
+
+# Full code:  Use this to produce final linked micromaps of the variables you want
+# export as pdf US letter landscape from Plots pane
+# export as PNG Width 900 Height 968, maintain aspect ratio
+mmplot(stat.data=ecoreg_stats,
+       map.data=ecor.polys,
+       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
+       panel.data=list(
+         NA,
+         'WSA9_NAME',
+         list('post_med_n2o', 'LCI_n2o', 'UCI_n2o'),
+         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
+         NA),
+       ord.by = 'post_med_n2o', rev.ord = TRUE, grouping=3,
+       median.row=FALSE,
+       map.link=c('WSA9', 'ID'),
+
+       plot.height=7, plot.width=7, #for poster use plot.height=6, plot.width=9,
+       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
+       map.color2='lightgray',
+
+       panel.att=list(list(1, point.type=20, point.size=2,
+                           graph.border.color='white',
+                           xaxis.text.display=F, xaxis.line.display=F,
+                           graph.grid.major=F),
+
+                      list(2, header='NARS Reporting\n Regions', panel.width=1.25,#increased panel.width to 1.1 from .8
+                           align='left', text.size=1),#increased text.size to .8 from .6
+
+                      list(3, header='N2O Concentration\n with 95% CI',
+                           graph.bgcolor='lightgray',
+                           xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           xaxis.title='N2O mg/L', panel.width=1.25),
+
+                      list(4, header='N2O Saturation\n with 95% CI',
+                           graph.bgcolor='lightgray',
+                           xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
+                           xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
+                           xaxis.title='N2O Saturation, %',panel.width=1.25),
+
+                      list(5, header='Light Gray Means\n Highlighted Above',
+                           inactive.border.color=gray(.7), inactive.border.size=2,
+                           panel.width=2)))
+
+# To get ecoregional distributions of the predictions I think this is the code we would use based on our conversation with Roy, but I would double-check with him.
+
+ecoreg_preds <- all_predictions %>%
+  group_by(WSA9, .draw) %>%
+  summarise( post_med_n2o = round(median(mean_n2o), 1),
+    LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
+    UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
+     post_med_n2oeq = round(median(mean_n2oeq), 2),
+    LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
+    UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
+     post_med_sat = round(median(mean_sat), 2),
+    LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
+    UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
+    n = dplyr::n())
+```
 # Exploratory analysis of N2O data

### 2021-06-09 scripts/ASLO_2021.Rmd
adding scripts addressing wind data, population estimates, and ASLO presentation

---
title: "ASLO 2021"
author: "J. Beaulieu"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
# load libraries
library(sf) # spatial data
#library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
#library(janitor) # clean names
#library(spdep) # k-NN and Moran plot
#library(ncf) # ncf::correlogram
#library(lattice) # xyplot
library(USAboundaries)

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}
```

## ASLO 2021

This is a convenient place to generate the stats and figures included in the ASLO 2021 presentation.

## Site Map
```{r}
# Load data
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

# To enable mapping the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# Load state boundaries
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf) +
  geom_sf(data = states, fill = NA, color = "white", size = 0.1) +
  theme_void() +
  theme(legend.title = element_blank())

ggsave("output/figures/siteMap.tiff")

```

## Population Data

```{r}
# about 15 minutes to load first time
# 6/7/21 <2 minutes
Sys.time()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
Sys.time()
```

## Explore structure of all_predictions object
```{r}
head(all_predictions)
```
.row uniquely identifies each waterbody.  There are `r all_predictions %>% distinct(.row) %>% nrow()` waterbodies in the sample frame.  `.draw` reflects the simulation number for each waterbody. I think this is a 'draw from the posterior distribution'.  The data contains `r all_predictions %>% distinct(.draw) %>% nrow()` draws for each waterbody.

## N2O saturation ratio: continuous variable
```{r}
# PLOT N2O SAT RATIO BY WSA9

# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_sat = mean(preds_sat)) %>%
  summarise( estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()

```

## N2O saturation ratio: extent estimate
```{r}
# EXTENT OF WATERBODIES FUNCTIONING AS SINKS BY ECOREGION

# density plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  # below summarise generates 500 proportions per WSA
  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  ggplot(aes(x=prop_sat)) + 
  geom_density() +
  facet_wrap(~WSA9) +
  xlab("Proportion of waterbodies functioning as N2O sinks") +
  theme_bw()
ggsave("output/figures/extentDensityPlot.tiff")

# point and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/extentPointLinerange.tiff")

# bar and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_bar(stat = "identity", aes(fill = ecoregion)) +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()  
ggsave("output/figures/extentBar.tiff")
```

### 2021-06-09 scripts/nlaSampleFrameWind.R

@@ -1,0 +1,167 @@
+# THIS SCRIPT IS TO CALCULATE WIND DATA FOR ALL ~600,000 WATERBODIES IN THE 
+# NLA SAMPLE FRAME.  THESE DATA WILL BE USED TO CALCULATE A K VALUE FOR
+# EACH SITE.  WHEN COMBINED WITH ROY'S MODEL PREDICTIONS OF DISSOLVED N2O 
+# WILL ENABLE EMISSION RATE PREDICTION FOR EACH SITE.  THESE PREDICTIONS CAN
+# BE USED TO ESTIMATE POPULATION LEVEL MEAN AND VARIANCE.  
+# THIS CODE SHOULD BE RUN ON VM.  SCRIPT IS NOT YET COMPLETE.
+# SEE 'readWind.R' IN THE NLA PROJECT FOR A VERSION TAILORED TO THE 1200
+# NLA OBSERVATIONS.  THE NLA VERSION UTILIZES LARGE DATA FRAMES, BUT CAN
+# BE RUN ON LAPTOP.  VERSION BELOW HAS A STRONGER RELIANCE ON LISTS, BUT
+# THE DATAFRAME APPROACH MIGHT WORK ON VM.  NEED TO DECIDE HOW WE ARE GOING
+# TO AGGREGATE DATA.  OPTIONS INCLUDE 1) INDEX PERIOD MEAN PER SITE, 2) 
+# INDEX PERIOD MEAN OVER DAYLIGHT HOURS, AND 3) MEAN DAYLIGHT HOURS PER 
+# DAY.  OPTION 3 WOULD ENABLE CALCULATING A K AND EMISSION RATE PER DAY
+# FOR EACH SITE.
+
+
+#R 4.0.3
+library(tidyverse)
+library(sf)
+library(ncdf4)
+library(ncdf4.helpers)
+library(suncalc)
+library(lubridate)
+library(pbapply)
+
+# ENFORCE EPA FORMAT ON NLA NAMES
+toEPA <- function(X1){
+  names(X1) = tolower(names(X1))
+  names(X1) = gsub(pattern = "_", replacement = ".", x = names(X1))
+  X1
+}
+
+
+# NLA17 sample frame--------------------------
+sampFrm <- read.csv(file = paste0(Sys.getenv("USERPROFILE"), 
+                                    "/Environmental Protection Agency (EPA)/",
+                                    "ORD NLA17 Dissolved Gas - Documents/",
[…]

### 2021-06-09 scripts/dgIndicatorAnalysis.Rmd
Minor tweaks to micromap code

@@
 ```{r, linked_micromap}
+# Read data object containing predictions for population
+Sys.time()
+load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
+Sys.time()
 
@@
 
-ecoregions <- st_read("aggr_ecoregions_simple.shp")
+ecoregions <- st_read(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
+                             "ORD NLA17 Dissolved Gas - Documents\\inputData\\aggr_ecoregions_simple.shp"))
+
 wsa9names <- st_set_geometry(ecoregions, NULL) %>%
@@
 # export as PNG Width 900 Height 968, maintain aspect ratio
+tiff(filename = "output/figures/mmplot.tiff",
+     res = 800, width = 9.00, height = 9.68, units = "in" )
+
 mmplot(stat.data=ecoreg_stats,
@@
                            panel.width=2)))
+
+dev.off()
 

### 2021-09-22 scripts/ASLO_2021_2.Rmd
syncing up after ASLO 2021 slide preparation

---
title: "ASLO 2021"
authors: "J. Beaulieu"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
    code_folding:  hide
editor_options: 
  chunk_output_type: console
---

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
library(janitor) # clean names
library(spdep) # k-NN and Moran plot
library(ncf) # ncf::correlogram
library(lattice) # xyplot
library(micromap) # linked micromaps

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}
```

## ASLO 2021

This is a convenient place to generate the stats and figures included in the ASLO 2021 presentation.

## Site Map
```{r}
# Load data
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

# To enable mapping the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# Load state boundaries
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf) +
  geom_sf(data = states, fill = NA, color = "white", size = 0.1) +
  theme_void() +
  theme(legend.title = element_blank())

ggsave("output/figures/siteMap.tiff")

```

## Population Data

```{r}
# about 15 minutes to load first time
# 6/7/21 <2 minutes
Sys.time()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
Sys.time()
```

## Explore structure of all_predictions object
```{r}
head(all_predictions)
```
.row uniquely identifies each waterbody.  There are `r all_predictions %>% distinct(.row) %>% nrow()` waterbodies in the sample frame.  `.draw` reflects the simulation number for each waterbody. I think this is a 'draw from the posterior distribution'.  The data contains `r all_predictions %>% distinct(.draw) %>% nrow()` draws for each waterbody.

## N2O saturation ratio: continuous variable
N2O saturation ratio in the sample data has a long right tail, with a maximum value of ~30.  The distribution of the values predicted to the population also has a long right tail, but features more extreme values (>600).
```{r}
# PLOT N2O SAT RATIO BY WSA9

# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_sat = mean(preds_sat)) %>%
  summarise( estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()

# density plots of predicted distribution
# huge right tail
p <- all_predictions %>%
  filter(.draw %in% 1:10) %>% # subset for practice
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
  ggplot( aes( x = preds_sat, group = .draw, color = .draw ) ) +
  geom_density(show.legend = FALSE) +
  geom_vline(xintercept = 1, color = "red") +
  xlab("N2O saturation ratio") +
  xlim(0,5) + # data range is 0-600, but sample max is 30
  facet_wrap(~ecoregion) +
  theme_bw()

#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oSatDensitySmallRange.tiff", g)

summary(dg$n2o.sat.ratio) # max value in sample is 33

# rplot, but show full x-axis range
# include RUG to better show extreme values
p <- all_predictions %>%
  filter(.draw == 1) %>% # need only one draw, otherwise RUG is misleading
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
  ggplot( aes( x = preds_sat ) ) +
  geom_density(color = "blue") +
  geom_vline(xintercept = 1, color = "red") +
  geom_rug() +
  xlab("N2O saturation ratio") +
  #xlim(0,35) + # data range is 0-600, but sample max is 30
  facet_wrap(~ecoregion, scales = "free") +
  theme_bw()

#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oSatDensityFullRange.tiff", g)

all_predictions %>%
    filter(preds_sat > 40, .draw == 1) %>% # look at one draw
  {table(.$WSA9)} # >40 is uncommon outside of CPL

# # density plots of observed distribution
# huge right tail
dg %>%
  mutate(ecoregion = factor(ag.eco9.nm)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, n2o.sat.ratio)) %>%
  ggplot( aes( x = n2o.sat.ratio ) ) +
  geom_density() +
  xlab("mean N2O saturation ratio") +
  facet_wrap(~ecoregion) +
  theme_bw()

```

## N2O saturation ratio: extent estimate
```{r}
# EXTENT OF WATERBODIES FUNCTIONING AS SINKS BY ECOREGION

# density plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  # below summarise generates 500 proportions per WSA
  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  ggplot(aes(x=prop_sat)) + 
  geom_density() +
  facet_wrap(~WSA9) +
  xlab("Proportion of waterbodies functioning as N2O sinks") +
  theme_bw()
ggsave("output/figures/extentDensityPlot.tiff")

# point and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/extentPointLinerange.tiff")

# bar and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_bar(stat = "identity", aes(fill = ecoregion)) +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()  
ggsave("output/figures/extentBar.tiff")
```

# N2O saturation ratio: continuous and extent micromap
```{r}
# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
# this is analogous to spsurvey "population estimates"
ecoreg_stats <- all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(preds_n2o),
             mean_n2oeq = mean(preds_n2oeq),
             mean_sat = mean(preds_sat),
             prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  summarise( post_med_n2o = round(mean(mean_n2o), 1),
             LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
             UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
             post_med_n2oeq = round(mean(mean_n2oeq), 2),
             LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
             UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
             post_med_sat = round(mean(mean_sat), 2),
             LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
             UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
             estimate = round(mean(prop_sat), 3),
             LCL = round(quantile(prop_sat, probs = 0.025), 3),
             UCL = round(quantile(prop_sat, probs = 0.975), 3),
             n = dplyr::n())

ecoreg_stats <- as.data.frame(ecoreg_stats)
class(ecoreg_stats)
str(ecoreg_stats) # WSA9 is a factor, but some of the stats have attribute of Named num

ecoreg_stats$WSA9 <- as.character(ecoreg_stats$WSA9)
str(ecoreg_stats)

ecoreg_stats

ecoregions <- st_read(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
                             "ORD NLA17 Dissolved Gas - Documents\\inputData\\aggr_ecoregions_simple.shp"))

wsa9names <- st_set_geometry(ecoregions, NULL) %>%
  select(WSA9, WSA9_NAME)

ecoreg_stats <- left_join(ecoreg_stats, wsa9names)

ecoregions <- st_transform(ecoregions, 5070)
# make spatial polygon dataframe need for micromap
ecoregions_sp1 <- as(ecoregions, "Spatial")

# convert to map_table for linked micromap
ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
class(ecor.polys)
str(ecor.polys, max =2) # ID is chr "CPL"...

# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

mmplot(stat.data=ecoreg_stats,
       map.data=ecor.polys,
       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
       panel.data=list(
         NA,
         'WSA9_NAME',
         list('estimate', 'LCL', 'UCL'),
         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
         NA),
       ord.by = 'estimate', rev.ord = TRUE, grouping=3,
       median.row=FALSE,
       map.link=c('WSA9', 'ID'),

       plot.height=5, # what does this do?
       plot.width=7, #what does this do? 
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
       map.color2='lightgray',

       panel.att=list(list(1, point.type=20, point.size=2,
                           graph.border.color='white',
                           xaxis.text.display=F, xaxis.line.display=F,
                           graph.grid.major=F, 
                           panel.width = 2), # added this argument to ensure the full point is shown

                      list(2,  panel.width=2,#increased panel.width to 1.1 from .8, removed header='NARS Reporting\n Regions',
                           align='left', text.size=0.8),#increased text.size to .8 from .6

                      list(3, header= "Proportion undersaturated\n with 95% CI", 
                           graph.bgcolor='lightgray',
                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           xaxis.title='Proportion undersaturated', panel.width=2.3),

                      list(4, header="N2O Saturation\n with 95% CI", #N2O Saturation\n with 95% CI
                           graph.bgcolor='lightgray',
                           xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.title='N2O Saturation, %',panel.width=2.3),

                      list(5, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                           panel.width=2)),

       print.file = "output/figures/mmapSatExtent.tiff", print.res = 300)

```

# N2O emission rate and flux
Roy predicted N2O equilibrium and concentration to the population.  I used ERA5 to calculate mean wind speed during the index period for each lake in the population.  I need to match wind speed to Roy's predictions.  Waterbodies in Roy's predictions are identified by '.row'.  There are `r all_predictions %>% distinct(.row) %>% summarize(population = n()) %>% pull()` rows.

```{r}

# Read in sample frame
sampFrm <- read.csv(file = paste0(Sys.getenv("USERPROFILE"), 
                                    "/Environmental Protection Agency (EPA)/",
                                    "ORD NLA17 Dissolved Gas - Documents/",
                                    "inputData/NLA_Sample_Frame.csv"), header = TRUE,
                    na.strings = "") %>%
  toEPA()

dim(sampFrm) # 614730, 71

sampFrm <- sampFrm %>%
  filter(nla17.sf != "Exclude2017") %>%
  filter(nla17.sf != "Exclude2017_Include2017NH") %>%
  filter(state != "DC") %>%
  filter(state != "HI")

dim(sampFrm)

# print first 6 lines
sampFrm %>% select(ag.eco9, area.cat6, area.ha) %>% head()

# print first 6 .row values of all_predictions
all_predictions %>%
  filter(.row %in% 1:6) %>%
  group_by(.row) %>%
  slice(1)

# OK, .row values in all_predictions match with row number in sampFrm.
# This same order was preserved in wind data.

```

### 2021-09-22 scripts/dgIndicatorAnalysis.Rmd
syncing up after ASLO 2021 slide preparation

@@
 ```
+# ASLO 2021
+
+## Site Map
+```{r}
+# Load data
+ load(paste0(localPath,
+            "/Environmental Protection Agency (EPA)/",
+            "ORD NLA17 Dissolved Gas - Documents/",
+            "inputData/dg.2021-02-01.RData"))
+
+# To enable mapping the dataframe will be converted to a 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple")
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# Load state boundaries
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+ggplot() +
+  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+  geom_sf(data = dg.sf) +
+  geom_sf(data = states, fill = NA, color = "white", size = 0.1) +
+  theme_void() +
+  theme(legend.title = element_blank())
+
+ggsave("output/figures/siteMap.tiff")
+
+```
+
+## Population Data
+
+```{r}
+# about 15 minutes to load first time
+# 6/7/21 <2 minutes
+Sys.time()
+load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
+Sys.time()
+```
+
+## Explore structure of all_predictions object
+```{r}
+head(all_predictions)
+```
+.row uniquely identifies each waterbody.
+There are `r all_predictions %>% distinct(.row) %>% nrow()` waterbodies in the sample frame.
+`.draw` reflects the simulation number for each waterbody.
+I think this is a 'draw from the posterior distribution'.
+The data contains `r all_predictions %>% distinct(.draw) %>% nrow()` draws for each waterbody.
+
+The `size_cat` variable largely uses the NLA size category definitions.  This may not be the best for showing N2O fluxes by lake size bins, however.  The >50 ha (0.5km2) has most of the surface area
+
+```{r}
+# total area by size_cat bar chart
+all_predictions %>%
+  filter(.draw == 1) %>%
+  group_by(WSA9, size_cat) %>%
+  summarise(tot_area = sum(area_ha)) %>%
+  ggplot(aes(x=size_cat, y=tot_area)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~WSA9)
+
+```
+
+Lets create a new size category with more even surface area distributions.
+
+```{r}
+# 11 minutes on VM!
+all_predictions <- all_predictions %>%
+  #filter(.draw == 1) %>% # testing
+  mutate(area_km2 = area_ha / 100, # this works
+         size_cat2 = ifelse(area_km2 < 0.1,
+                            "<0.1",
+                            ifelse(area_km2 >= 0.1 & area_km2 < 1,
+                                   "0.1 - 1",
+                                   ifelse(area_km2 >=1 & area_km2 <10,
+                                          "0.1 - 1",
+                                          ifelse(area_km2 >=10 & area_km2 <100,
+                                                 "10 - 100",
+                                                 ifelse(area_km2 >=100, 
+                                                        ">100",
+                                                        NA))))),
+         size_cat2 = factor(size_cat2, ordered = TRUE, 
+                            levels = c("<0.1", "0.1 - 1", "1 - 10", "10 - 100", ">100"))) %>%
+  select(-area_km2)
+
+# total area by size_cat2 bar chart
+all_predictions %>%
+  filter(.draw == 1) %>%
+  group_by(WSA9, size_cat2) %>%
+  summarise(tot_area = sum(area_ha)) %>%
+  ggplot(aes(x=size_cat2, y=tot_area)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~WSA9)
+
+```
+
+## Reorder WSA9
+Reorder WSA9 to correspond with order of first micromap produced below, which is ordered by extent of N2O saturation.
+```{r}
+all_predictions <- all_predictions %>%
+  mutate(WSA9 = factor(WSA9,
+                       levels = c("NPL", "TPL", "UMW", 
+                                  "CPL", "SPL", "SAP",
+                                  "XER", "NAP", "WMT"),
+                       labels = c("Northern Plains", "Temperate Plains",
+                                  "Upper Midwest", "Coastal Plains", "Southern Plains", "Southern Appalachians", "Xeric", "Northern Appalachians", "Western Mountains")))
+
+```
+
+## N2O saturation ratio: continuous variable
+N2O saturation ratio in the sample data has a long right tail, with a maximum value of ~30.
+The distribution of the values predicted to the population also has a long right tail, but features more extreme values (>600).
+```{r}
+# PLOT N2O SAT RATIO BY WSA9
+
+# point and linerange
+all_predictions %>%
+  group_by(WSA9, .draw) %>%
+  #group_by(.draw) %>%
+  summarise( mean_sat = mean(preds_sat)) %>%
+  summarise( estimate = round(median(mean_sat), 3),
+    LCL = round(quantile(mean_sat, probs = 0.025), 3),
+    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = ecoregion, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  geom_hline(yintercept = 1, color='blue') +
+  ylab("mean N2O saturation ratio") +
+  coord_flip() + 
+  theme_bw()
+
+# density plots of predicted distribution
+# using xlim to zoom in on values close to 1
+p <- all_predictions %>%
+  filter(.draw %in% 1:10) %>% # subset for practice
+  #mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
+  ggplot( aes( x = preds_sat, group = .draw, color = .draw ) ) +
+  geom_density(show.legend = FALSE) +
+  geom_vline(xintercept = 1, color = "red") +
+  xlab(expression(N[2]*O~saturation~ratio)) +
+  xlim(0,5) + # data range is 0-600, but sample max is 30
+  facet_wrap(~WSA9) +
+  theme_bw()
+
+# this bit fills strip with same colors used in micromap (see below)
+#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
+g <- ggplot_gtable(ggplot_build(p))
+stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
+fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
+k <- 1
+for (i in stripr) {
+  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
+  k <- k+1
+}
+grid::grid.draw(g)
+
+ggsave("output/figures/n2oSatDensitySmallRange.tiff", g)
+
+summary(dg$n2o.sat.ratio) # max value in sample is 33
+
+# rplot, but show full x-axis range
+# include RUG to better show extreme values
+p <- all_predictions %>%
+  filter(.draw == 1) %>% # need only one draw, otherwise RUG is misleading
+  #mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
+  ggplot( aes( x = preds_sat ) ) +
+  geom_density(color = "blue") +
+  geom_vline(xintercept = 1, color = "red") +
+  geom_rug() +
+  xlab(expression(N[2]*O~saturation~ratio)) +
+  #xlim(0,35) + # data range is 0-600, but sample max is 30
+  facet_wrap(~WSA9, scales = "free") +
+  theme_bw()
+
+# this bit fills strip with same colors used in micromap (see below)
+#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
+g <- ggplot_gtable(ggplot_build(p))
+stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
+fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
+k <- 1
+for (i in stripr) {
+  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
+  k <- k+1
+}
+grid::grid.draw(g)
+
+ggsave("output/figures/n2oSatDensityFullRange.tiff", g)
+
+# Take a peak at values >40 in predictions
+all_predictions %>%
+    filter(preds_sat > 40, .draw == 1) %>% # look at one draw
+  {table(.$WSA9)} # >40 is uncommon outside of CPL
+
+# # density plots of observed distribution
+# huge right tail
+dg %>%
+  mutate(ecoregion = factor(ag.eco9.nm)) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, n2o.sat.ratio)) %>%
+  ggplot( aes( x = n2o.sat.ratio ) ) +
+  geom_density() +
+  xlab("mean N2O saturation ratio") +
+  facet_wrap(~ecoregion) +
+  theme_bw()
+
+```
+
+## N2O saturation ratio: extent estimate
+```{r}
+# EXTENT OF WATERBODIES FUNCTIONING AS SINKS BY ECOREGION
+
+# density plot
+all_predictions %>%
+  group_by(WSA9, .draw) %>%
+  # below summarise generates 500 proportions per WSA
+  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
+  ggplot(aes(x=prop_sat)) + 
+  geom_density() +
+  facet_wrap(~WSA9) +
+  xlab("Proportion of waterbodies functioning as N2O sinks") +
+  theme_bw()
+ggsave("output/figures/extentDensityPlot.tiff")
+
+# point and line plot
+all_predictions %>%
+  group_by(WSA9, .draw) %>%
+  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
+  summarise(estimate = round(median(prop_sat), 3),
+    LCL = round(quantile(prop_sat, probs = 0.025), 3),
+    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = ecoregion, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
+  ylab("Proportion of waterbodies functioning as N2O sinks") +
+  coord_flip() + 
+  theme_bw()
+ggsave("output/figures/extentPointLinerange.tiff")
+
+# bar and line plot
+all_predictions %>%
+  group_by(WSA9, .draw) %>%
+  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
+  summarise(estimate = round(median(prop_sat), 3),
+    LCL = round(quantile(prop_sat, probs = 0.025), 3),
+    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = ecoregion, y = estimate ) ) +
+  geom_bar(stat = "identity", aes(fill = ecoregion)) +
+  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
+  ylab("Proportion of waterbodies functioning as N2O sinks") +
+  coord_flip() + 
+  theme_bw()  
+ggsave("output/figures/extentBar.tiff")
+```
+
+## N2O emission rate and flux
+### Calculations
+Roy predicted N2O equilibrium and concentration to the population.
+I used ERA5 to calculate mean wind speed during the index period for each lake in the population.
+I need to match wind speed to Roy's predictions.
+Waterbodies in Roy's predictions are identified by '.row'.
+There are `r all_predictions %>% distinct(.row) %>% summarize(population = n()) %>% pull()` rows.
+
+```{r}
+
+# Read in wind data
+# Mean index period wind speed was calculated on VM via 
+# nlaSampleFrameWind.R.  The resulting data object was written out and read in below.
+wind <- read.table(paste0(Sys.getenv("USERPROFILE"), 
+                                    "/Environmental Protection Agency (EPA)/",
+                                    "ORD NLA17 Dissolved Gas - Documents/",
+                                    "inputData/windSampleFrame.txt"))
+# print first 6 lines
+wind %>% head()
+
+# print first 6 .row values of all_predictions
+all_predictions %>%
+  filter(.row %in% 1:6) %>% # each row is a unique waterbody
+  group_by(.row) %>%
+  slice(1) # grabs first record
+
+# Row order in wind and all_predictions is inherited from sample frame.  Thus row 1
+# of wind corresponds to .row == 1 in all_predictions.  This is a sketchy way to merge
+# things, but will do for now.
+
+# add .id to wind which is simply row number
+wind <- wind %>% mutate(.id = row_number())
+
+# Merge wind and all_predictions
+all_predictions.w <- full_join(all_predictions, wind, by = c(".row" = ".id")) # too much for laptop,30 seconds on VM
+dim(wind) # 465897
+dim(all_predictions) #232948500, 500 observations for each lake
+465897*500 #232948500
+dim(all_predictions.w) #wind.sampFr. good
+
+# ESTIMATE GAS EXCHANGE RATE FOR POPULATION
+
+# Using gas exchange model from:
+# Vachon, D., and Y. T. Prairie (2013), The ecosystem size and shape 
+# dependence of gas transfer velocity versus wind speed relationships in 
+# lakes, Can. J. Fish. Aquat. Sci., 70(12), 1757-1764, doi:10.1139/cjfas-2013-0241.
+# Table 2, model B.
+
+# define k600 prediction function
+pred.k600 <- function(ws, area) { 
+  # predict k600 in cm h-1
+  # ws = wind speed (m s-1)
+  # area = lake area (km2)
+  2.51 + (1.48 * ws) + (0.39 * ws * log10(area))
+}
+
+all_predictions.w <- all_predictions.w %>%
+  mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
+         # 1000 L to m3. 44ng to 1nmol. ng->ug. ug->mg
+         e.n2o.mg.m2.d = (((preds_n2o - preds_n2oeq) * 1000 * 44) / (1000 * 1000)) * 
+           (k600.cm.h * (24/100)), #h->day. cm->m
+         f.n2o.Mg.y = (e.n2o.mg.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
+           (1000*1000*1000)) #mg-g, g-kg, kg-Mg
+
+# write reduced file to disk for less computationally intensive wrangling
+#all_predictions.w.small <- all_predictions.w %>% filter(.draw %in% 1:5)
+
+#save(all_predictions.w.small,
+#     file = paste0(Sys.getenv("USERPROFILE"), 
+#                   "/Environmental Protection Agency (EPA)/",
+#                   "ORD NLA17 Dissolved Gas - Documents/",
+#                   "inputData/all_predictions_w_small.rda"))
+```
+### compared to literature and US GHG inventory
+The distribution of N2O emission rates (mg N2O m-2 d-1) for one draw is very similar to that derived from a literature review of lake/reservoir emission rates (DelSontro et al), except for CPL which shows some exceedingly high rates.
+Overall, our results look reasonable.
+
+The total N2O flux from US lakes and reservoirs is minor compared to other anthropogenic N2O sources
+```{r}
+all_predictions.w %>% 
+  filter(.draw == 1) %>% 
+  ggplot(aes(x=e.n2o.mg.m2.d)) + 
+  geom_density() +
+  geom_rug() + 
+  facet_wrap(~WSA9, scales = "free")
+
+knitr::include_graphics((paste0(Sys.getenv("USERPROFILE"), 
+                                    "/Environmental Protection Agency (EPA)/",
+                                    "ORD NLA17 Dissolved Gas - Documents/",
+                                    "EDA/en2oLiterature.tiff")))
+
+## Comparison to US GHG inventory
+# top sources from inventory report in kt / year
+n2o <- data.frame(kt = c(1156,88,84,66,60,34,18,14,7),
+                     source = c("Agricultural Soil Management",
+                                "Wastewater Treatment",
+                                "Stationary Combustion",
+                                "Manure Management",
+                                "Mobile Combustion",
+                                "Nitric Acid Production",
+                                "Adipid Acid Production",
+                                "N2O from Product Uses",
+                                "Composting"))
+
+n2o <- n2o %>% mutate(Mg = (kt * 1000 * 1000) / 1000,  # 1000 metric tons in a kt, 1000 kg in a metric ton, 1000 kg in a Mg
+                      Gg = Mg / 1000)
+
+# National scale number for lakess/reservoirs (this is is Gg)
+national.lakes <- all_predictions.w %>% 
+  group_by(.draw) %>% 
+  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
+  summarise(Gg = round(mean(total.n2o), 3) / 1000, # divide by 1000 for Mg to Gg
+            Gg_LCL = round(quantile(total.n2o, probs = 0.025), 3) / 1000, # divide by 1000 for Mg to Gg
+            Gg_UCL = round(quantile(total.n2o, probs = 0.975), 3) / 1000) %>% # divide by 1000 for Mg to Gg
+  mutate(source = "Lakes and Reservoirs")
+
+# Merge lakes and inventory category
+inventory <- full_join(n2o, national.lakes)
+inventory
+
+inventory %>% mutate(source = as.factor(source),
+                     source = reorder(source, Gg)) %>%
+  ggplot(aes(source, Gg)) +
+  geom_bar(stat = "identity", color = "orange", fill = "orange") +
+  geom_errorbar(aes(x=source, ymin = Gg_LCL, ymax = Gg_UCL), width = 0.5) +
+  coord_flip() +
+  ylab(expression(N[2]*O~flux~(Gg~year^{-1}))) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggsave(filename = "output/figures/n2oInventory.tiff")
+
+tiff("output/figures/n2oInventoryPie.tiff")
+pie.labels <- c("Agricultural Soil Management",
+                "Wastewater Treatment",
+                "Stationary Combustion",
+                "Manure Management",
+                "Mobile Combustion",
+                "Nitric Acid Production",
+                NA,
+                NA,
+                NA,
+                "Lakes and Reservoirs")
+pie(inventory$Gg, labels = pie.labels, col = brewer.pal(8, "Set2"), radius = 0.6, cex = 0.9)
+dev.off()
+
+(national.lakes$Gg/sum(inventory$Gg)) * 1000 #1.6%
+```
+
+### Figures and tables
+Below is the distribution of total annual fluxes from the lakes.  This is calculated as lake area * emission rate.
+
+```{r}
+# density plot with RUG for one draw
+all_predictions.w %>% filter(.draw == 1) %>% ggplot(aes(x=f.n2o.Mg.y)) + geom_density() +geom_rug() + facet_wrap(~WSA9, scales = "free")
+
+# flux vs area for one draw
+all_predictions.w %>% filter(.draw == 1) %>% ggplot(aes(y=f.n2o.Mg.y, x=area_ha)) + geom_point() + facet_wrap(~WSA9, scales = "free")
+
+# bar plot by size category and ecoregion  
+all_predictions.w %>% filter(.draw == 1) %>% group_by(WSA9, size_cat2) %>% summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
+  ggplot(aes(x=size_cat2, y=total.n2o)) + geom_bar(stat = "identity") + facet_wrap(~WSA9, scales = "free")
+
+# bar plot by size category with error
+p <- all_predictions.w %>% 
+  group_by(WSA9, .draw, size_cat2) %>% 
+  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
+  group_by(WSA9, size_cat2) %>%
+  summarise(post_med_fn2o = round(mean(total.n2o), 3), 
+            LCL_fn2o = round(quantile(total.n2o, probs = 0.025), 3),
+            UCL_fn2o = round(quantile(total.n2o, probs = 0.975), 3)) %>%
+  ggplot(aes(x=size_cat2, y=post_med_fn2o)) + 
+  geom_bar(stat = "identity") + 
+  geom_errorbar(aes(x=size_cat2, ymin = LCL_fn2o, ymax = UCL_fn2o)) +
+    xlab(expression(Lake~Size~Bin~(km^{2}))) +
+  ylab(expression(Mg~N[2]*O~year^{-1})) +
+  facet_wrap(~WSA9, scales = "free") +
+  theme_bw()
+
+# this bit fills strip with same colors used in micromap (see below)
+#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
+g <- ggplot_gtable(ggplot_build(p))
+stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
+fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
+k <- 1
+for (i in stripr) {
+  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
+  k <- k+1
+}
+grid::grid.draw(g)
+
+ggsave("output/figures/n2oFluxByLakeSizeWithError.tiff", g)
+
+# bar plot by size category without error
+p <- all_predictions.w %>% 
+  group_by(WSA9, .draw, size_cat2) %>% 
+  # tot x size x draw
+  summarise(total.n2o = sum(f.n2o.Mg.y),
+            obs = n()) %>% # number of lakes per WSA x size
+  group_by(WSA9, size_cat2) %>%
+  summarise(post_med_fn2o = round(mean(total.n2o), 3), # across draws
+            LCL_fn2o = round(quantile(total.n2o, probs = 0.025), 3),
+            UCL_fn2o = round(quantile(total.n2o, probs = 0.975), 3),
+            obs = mean(obs)) %>% # obs is repeated across draws.  mean gives identity
+  ggplot(aes(x=size_cat2, y=post_med_fn2o)) + 
+  geom_bar(stat = "identity") + 
+  xlab(expression(Lake~Size~Bin~(km^{2}))) +
+  ylab(expression(Mg~N[2]*O~year^{-1})) +
+  #scale_x_discrete(labels = c("min to 4", "4-10", "10-20", "20-50", ">50")) +
+  #geom_errorbar(aes(x=size_cat, ymin = LCL_fn2o, ymax = UCL_fn2o)) +
+  facet_wrap(~WSA9, scales = "free_y") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
+  geom_text(aes(label = scales::comma(obs, accuracy = 1)), # comma for thousands separator, accurary for decimal points
+            vjust = -0.5, size = 3) +
+  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2)))
+
+# this bit fills strip with same colors used in micromap (see below)
+#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
+g <- ggplot_gtable(ggplot_build(p))
+stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
+fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
+k <- 1
+for (i in stripr) {
+  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
+  k <- k+1
+}
+grid::grid.draw(g)
+
+ggsave("output/figures/n2oFluxByLakeSizeNoError.tiff", g)
+
+```
+
+## micromap: N2O saturation ratio continuous, extent, and emission rate
+```{r}
+# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
+# this is analogous to spsurvey "population estimates"
+ecoreg_stats <- all_predictions.w %>%
+  group_by(WSA9, .draw) %>%
+  summarise( mean_n2o = mean(preds_n2o),
+             mean_n2oeq = mean(preds_n2oeq),
+             mean_sat = mean(preds_sat),
+             prop_sat = sum(preds_sat < 1) / length(unique(.row)),
+             mean_en2o = mean(e.n2o.mg.m2.d),
+             fn2o_Mg = sum(f.n2o.Mg.y)) %>%
+  summarise( post_med_n2o = round(mean(mean_n2o), 1),
+             LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
+             UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
+             post_med_n2oeq = round(mean(mean_n2oeq), 2),
+             LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
+             UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
+             post_med_sat = round(mean(mean_sat), 2),
+             LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
+             UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
+             estimate = round(mean(prop_sat), 3),
+             LCL = round(quantile(prop_sat, probs = 0.025), 3),
+             UCL = round(quantile(prop_sat, probs = 0.975), 3),
+             post_med_en2o = round(mean(mean_en2o), 3),
+             LCL_en2o = round(quantile(mean_en2o, probs = 0.025), 3),
+             UCL_en2o = round(quantile(mean_en2o, probs = 0.975), 3),
+             post_med_fn2o = round(mean(fn2o_Mg), 3), 
+             LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
+             UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
+             n = dplyr::n())
+
+ecoreg_stats <- as.data.frame(ecoreg_stats)
+class(ecoreg_stats)
+str(ecoreg_stats) # WSA9 is a factor, but some of the stats have attribute of Named num
+
+ecoreg_stats$WSA9_NAME <- as.character(ecoreg_stats$WSA9)
+ecoreg_stats <- ecoreg_stats %>% select(-WSA9)
+str(ecoreg_stats)
+
+ecoreg_stats
+
+ecoregions <- st_read(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
+                             "ORD NLA17 Dissolved Gas - Documents\\inputData\\aggr_ecoregions_simple.shp"))
+
+wsa9names <- st_set_geometry(ecoregions, NULL) %>%
+  select(WSA9, WSA9_NAME)
+
+ecoreg_stats <- left_join(ecoreg_stats, wsa9names)
+
+ecoregions <- st_transform(ecoregions, 5070)
+# make spatial polygon dataframe need for micromap
+ecoregions_sp1 <- as(ecoregions, "Spatial")
+
+# convert to map_table for linked micromap
+ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
+class(ecor.polys)
+str(ecor.polys, max =2) # ID is chr "CPL"...
+
+# N2O SATURATION AND UNDERSATURATION EXTENT
+# Full code:  Use this to produce final linked micromaps of the variables you want
+# export as pdf US letter landscape from Plots pane
+# export as PNG Width 900 Height 968, maintain aspect ratio
+
+mmplot(stat.data=ecoreg_stats,
+       map.data=ecor.polys,
+       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
+       panel.data=list(
+         NA,
+         'WSA9_NAME',
+         list('estimate', 'LCL', 'UCL'),
+         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
+         NA),
+       ord.by = 'estimate', rev.ord = TRUE, grouping=3,
+       median.row=FALSE,
+       map.link=c('WSA9', 'ID'),
+
+       plot.height=5, # what does this do?
+       plot.width=7, #what does this do? 
+       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
+       map.color2='lightgray',
+
+       panel.att=list(list(1, point.type=20, point.size=2,
+                           graph.border.color='white',
+                           xaxis.text.display=F, xaxis.line.display=F,
+                           graph.grid.major=F, 
+                           panel.width = 2), # added this argument to ensure the full point is shown
+
+                      list(2,  panel.width=2,#increased panel.width to 1.1 from .8, removed header='NARS Reporting\n Regions',
+                           align='left', text.size=0.8),#increased text.size to .8 from .6
+
+                      list(3, header= "Proportion undersaturated\n with 95% CI", 
+                           graph.bgcolor='lightgray',
+                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           xaxis.title='Proportion undersaturated', panel.width=2.3),
+
+                      list(4, header="N2O Saturation\n with 95% CI", #N2O Saturation\n with 95% CI
+                           graph.bgcolor='lightgray',
+                           xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
+                           xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
+                           xaxis.title='N2O Saturation, %',panel.width=2.3),
+
+                      list(5, header='Light Gray Means\n Highlighted Above',
+                           inactive.border.color=gray(.7), inactive.border.size=2,
+                           panel.width=2)),
+
+       print.file = "output/figures/mmapSatExtent.tiff", print.res = 300)
+
+# N2O EMISSION AND FLUX
+# Full code:  Use this to produce final linked micromaps of the variables you want
+# export as pdf US letter landscape from Plots pane
+# export as PNG Width 900 Height 968, maintain aspect ratio
+
+mmplot(stat.data=ecoreg_stats,
+       map.data=ecor.polys,
+       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
+       panel.data=list(
+         NA,
+         'WSA9_NAME',
+         list('post_med_en2o', 'LCL_en2o', 'UCL_en2o'),
+         list('post_med_fn2o', 'LCL_fn2o', 'UCL_fn2o'),
+         NA),
+       ord.by = 'post_med_en2o', rev.ord = TRUE, grouping=3,
+       median.row=FALSE,
+       map.link=c('WSA9', 'ID'),
+
+       plot.height=5, # what does this do?
+       plot.width=7, #what does this do? 
+       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
+       map.color2='lightgray',
+
+       panel.att=list(list(1, point.type=20, point.size=2,
+                           graph.border.color='white',
+                           xaxis.text.display=F, xaxis.line.display=F,
+                           graph.grid.major=F, 
+                           panel.width = 2), # added this argument to ensure the full point is shown
+
+                      list(2,  panel.width=2,
+                           align='left', text.size=0.8),
+
+                      list(3, header= "Emission rate\n with 95% CI", 
+                           graph.bgcolor='lightgray',
+                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           xaxis.title='Emission rate (mg m-2 d-1)', panel.width=2.3),
+
+                      list(4, header="N2O flux\n with 95% CI", #N2O Saturation\n with 95% CI
+                           graph.bgcolor='lightgray',
+                           #xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
+                           #xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
+                           xaxis.title="N2O flux (Mg y-1)", panel.width=2.3),
+
+                      list(5, header='Light Gray Means\n Highlighted Above',
+                           inactive.border.color=gray(.7), inactive.border.size=2,
+                           panel.width=2)),
+
+       print.file = "output/figures/mmapEmisFlux.tiff", print.res = 300)
+```

### 2021-09-22 scripts/nlaSampleFrameWind.R

@@ -54,6 +54,7 @@ dim(sampFrm) # 465,897, 71
 # https://cran.r-project.org/web/packages/futureheatwaves/vignettes/starting_from_netcdf.html#:~:text=You%20can%20read%20netCDF%20data,connection%20to%20a%20netCDF%20file.
 # ncdf file downloaded  from Climate Data Center contains hourly wind and lake temperature
 # for US.
+# 10u10v24hour.nc file is too big.  crashed system after 32 hours
 wind <- nc_open(paste0(Sys.getenv("USERPROFILE"), 
                        "/Environmental Protection Agency (EPA)/",
                        "ORD NLA17 Dissolved Gas - Documents/",

### 2021-10-12 scripts/ASLO_2021_2.Rmd
cleaned up ASLO_2021_2.Rmd

@@
 ## Population Data
-
+Roy Martin predicted dissolved N2O concentration (preds_n2O), dissolved equilibrium N2O concentration (preds_n2oeq), and N2O saturation ratio (preds_sat) for all waterbodies in the population.
+See his .rmd for more details.
 ```{r}
@@
 ## Explore structure of all_predictions object
-```{r}
-head(all_predictions)
-```
 .row uniquely identifies each waterbody.
@@
 The data contains `r all_predictions %>% distinct(.draw) %>% nrow()` draws for each waterbody.
+```{r}
+head(all_predictions)
+```
 
@@
   coord_flip() + 
-  theme_bw()
+  theme_bw() +
+  ggtitle("Mean (+/- 95% CI) predicted N2O saturation ratio")
 

### 2022-12-14 scripts/dgIndicatorAnalysis.Rmd
Jake syncing local and remote

@@
   theme(axis.title = element_blank())
+
+# prep object for linked micromap
+pop.eco_stats <- pop.eco.diss %>%
+  mutate(subpopulation = as.character(subpopulation)) %>%
+  rename(WSA9_NAME = subpopulation)
+
+# read in ecoregion polygons
+ecoregions <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple")
+
+wsa9names <- st_set_geometry(ecoregions, NULL) %>%
+  select(WSA9, WSA9_NAME)
+
+pop.eco_stats <- left_join(pop.eco_stats, wsa9names)
+
+ecoregions <- st_transform(ecoregions, 5070)
+# make spatial polygon dataframe need for micromap
+ecoregions_sp1 <- as(ecoregions, "Spatial")
+
+# convert to map_table for linked micromap
+ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
+class(ecor.polys)
+str(ecor.polys, max =2) # ID is chr "CPL"...
+
+# Full code:  Use this to produce final linked micromaps of the variables you want
+# export as pdf US letter landscape from Plots pane
+# export as PNG Width 900 Height 968, maintain aspect ratio
+
+mmplot(stat.data=pop.eco_stats %>% 
+         filter(WSA9_NAME != "National",
+                indicator == "DISSOLVED_CH4"),
+       map.data=ecor.polys,
+       panel.types=c('dot_legend', 'labels', 'dot_cl', 'map'),
+       panel.data=list(
+         NA,
+         'WSA9_NAME',
+         list('estimate', 'lcb95pct', 'ucb95pct'),
+         NA),
+       ord.by = 'estimate', rev.ord = TRUE, grouping=3,
+       median.row=FALSE,
+       map.link=c('WSA9', 'ID'),
+
+       plot.height=5, # what does this do?
+       plot.width=5.5, #what does this do? 
+       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
+       map.color2='lightgray',
+
+       panel.att=list(list(1, point.type=20, point.size=2,
+                           graph.border.color='white',
+                           xaxis.text.display=F, xaxis.line.display=F,
+                           graph.grid.major=F, 
+                           panel.width = 2), # added this argument to ensure the full point is shown
+
+                      list(2,  panel.width=2,#increased panel.width to 1.1 from .8, removed header='NARS Reporting\n Regions',
+                           align='left', text.size=0.8),#increased text.size to .8 from .6
+
+                      list(3, header= "CH4 concentration\n with 95% CI", 
+                           graph.bgcolor='lightgray',
+                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           xaxis.title='CH4 concentration (uM)', panel.width=2.3),
+
+                      list(4, header='Light Gray Means\n Highlighted Above',
+                           inactive.border.color=gray(.7), inactive.border.size=2,
+                           panel.width=2)),
+
+       print.file = "output/figures/mmapCh4concentration.tiff", print.res = 300)
+
 ```
@@
   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf, aes(size = dissolved.ch4.umol, color = ch4.src.snk), 
+  geom_sf(data = dg.sf %>% filter(!is.na(ch4.src.snk)), 
+          aes(size = dissolved.ch4.umol, color = ch4.src.snk), 
           show.legend = "point") +
@@
              range = c(0.1, 10), # custom size range
-             breaks = c(0.1, 1, 10, 50, 100)) # custom breaks
+             breaks = c(0.1, 10, 50, 100)) # custom breaks
+
+ggsave("output/figures/ch4BubblePlot.tiff" )
 ```

### 2022-12-20 scripts/ASLO_2021_2.Rmd

@@
 library(micromap) # linked micromaps
+library(tictoc) # time operations
 
@@
 ```{r}
-# Load data
+# Load data.  Loaded object is named dg.
  load(paste0(localPath,
@@
 ```{r}
-# about 15 minutes to load first time
-# 6/7/21 <2 minutes
-Sys.time()
+# 12/14/22 <2 minutes
+tic()
 load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
-Sys.time()
+toc()
 ```
@@
   #group_by(.draw) %>%
-  summarise( mean_sat = mean(preds_sat)) %>%
-  summarise( estimate = round(median(mean_sat), 3),
+  summarise(mean_sat = mean(n2osat)) %>%
+  summarise(estimate = round(median(mean_sat), 3),
     LCL = round(quantile(mean_sat, probs = 0.025), 3),
@@
 p <- all_predictions %>%
-  filter(.draw %in% 1:10) %>% # subset for practice
+  #filter(.draw %in% 1:10) %>% # subset for practice
   mutate(ecoregion = factor(WSA9)) %>%
-  mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
-  ggplot( aes( x = preds_sat, group = .draw, color = .draw ) ) +
+  mutate(ecoregion = fct_reorder(ecoregion, n2osat)) %>%
+  ggplot( aes( x = n2osat, group = .draw, color = .draw ) ) +
   geom_density(show.legend = FALSE) +
@@
   xlab("N2O saturation ratio") +
-  xlim(0,5) + # data range is 0-600, but sample max is 30
+  xlim(0,5) + # data range is 0-300, but sample max is 30
   facet_wrap(~ecoregion) +

### 2023-02-08 scripts/dgIndicatorAnalysis.Rmd
reconciling merge conflicts

--- 2021-06-09 scripts/dgIndicatorAnalysis.Rmd
@@
 library(micromap) # linked micromaps
+library(spmodel) # esv function
 
@@
 ```{r, linked_micromap}
-# Read data object containing predictions for population
-Sys.time()
-load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
-Sys.time()
 
@@
 
-ecoregions <- st_read(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
-                             "ORD NLA17 Dissolved Gas - Documents\\inputData\\aggr_ecoregions_simple.shp"))
-
+ecoregions <- st_read("aggr_ecoregions_simple.shp")
 wsa9names <- st_set_geometry(ecoregions, NULL) %>%
@@
 # export as PNG Width 900 Height 968, maintain aspect ratio
-tiff(filename = "output/figures/mmplot.tiff",
-     res = 800, width = 9.00, height = 9.68, units = "in" )
-
 mmplot(stat.data=ecoreg_stats,
@@
                            panel.width=2)))
-
-dev.off()
 

### 2023-02-08 scripts/ASLO_2021_2.Rmd

@@
 ## Population Data
-Roy Martin predicted dissolved N2O concentration (preds_n2O), dissolved equilibrium N2O concentration (preds_n2oeq), and N2O saturation ratio (preds_sat) for all waterbodies in the population.
+Roy Martin predicted dissolved N2O concentration (n2o), dissolved equilibrium N2O concentration (n2oeq), and N2O saturation ratio (n2osat) for all waterbodies in the population.
 See his .rmd for more details.
@@
   mutate(ecoregion = factor(WSA9)) %>%
-  mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
-  ggplot( aes( x = preds_sat ) ) +
+  mutate(ecoregion = fct_reorder(ecoregion, n2osat)) %>%
+  ggplot( aes( x = n2osat ) ) +
   geom_density(color = "blue") +
@@
 all_predictions %>%
-    filter(preds_sat > 40, .draw == 1) %>% # look at one draw
+    filter(n2osat > 40, .draw == 1) %>% # look at one draw
   {table(.$WSA9)} # >40 is uncommon outside of CPL
@@
   # below summarise generates 500 proportions per WSA
-  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
+  summarise(prop_sat = sum(n2osat < 1) / length(unique(.row))) %>%
   ggplot(aes(x=prop_sat)) + 
@@
   group_by(WSA9, .draw) %>%
-  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
+  summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
   summarise(estimate = round(median(prop_sat), 3),
@@
 
-# bar and line plot
-all_predictions %>%
-  group_by(WSA9, .draw) %>%
-  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
-  summarise(estimate = round(median(prop_sat), 3),
-    LCL = round(quantile(prop_sat, probs = 0.025), 3),
-    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
-  mutate(ecoregion = factor(WSA9)) %>%
-  select(ecoregion, estimate, LCL, UCL) %>%
-  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
-  ggplot( aes( x = ecoregion, y = estimate ) ) +
-  geom_bar(stat = "identity", aes(fill = ecoregion)) +
-  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
-  ylab("Proportion of waterbodies functioning as N2O sinks") +
-  coord_flip() + 
-  theme_bw()  
-ggsave("output/figures/extentBar.tiff")
 ```
@@
   group_by(WSA9, .draw) %>%
-  summarise( mean_n2o = mean(preds_n2o),
-             mean_n2oeq = mean(preds_n2oeq),
-             mean_sat = mean(preds_sat),
-             prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
+  summarise( mean_n2o = mean(n2o),
+             mean_n2oeq = mean(n2oeq),
+             mean_sat = mean(n2osat),
+             prop_sat = sum(n2osat < 1) / length(unique(.row))) %>%
   summarise( post_med_n2o = round(mean(mean_n2o), 1),
@@
                            graph.bgcolor='lightgray',
-                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
-                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           xaxis.ticks=list(0.6, 0.65, 0.7, 0.75, 0.8),
+                           xaxis.labels=list(0.6, 0.65, 0.7, 0.75, 0.8),
                            xaxis.title='Proportion undersaturated', panel.width=2.3),
@@
                            xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
-                           xaxis.title='N2O Saturation, %',panel.width=2.3),
+                           xaxis.title='N2O Saturation',panel.width=2.3),
 

### 2023-02-08 scripts/N2O_MV_MRP_modeling.Rmd

---
title: "MRP with NLA 2017 $N_2O$ Data"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook: 
    toc: yes
    toc_depth: 4
    toc_float: true
    code_folding: show
    font_size: 14
    number_sections: true
    theme: simplex
bibliography: RWM_Endnote_Library.bib
link-citations: yes
---
# Setup R
```{r load_packages, message=FALSE, warning=FALSE}
library(brms)
library(bayesplot)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(tidybayes)
library(ggrepel)

options(mc.cores = parallel::detectCores(logical = FALSE))
options( max.print = 2000 )
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}

#Calc geometric mean
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

gm_sd = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(sd(log(x), na.rm = na.rm))
  } else {
    exp(sd(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

#calculate skew
skew <- function(x) {
  xdev <- x - mean(x)
  n <- length(x)
  r <- sum(xdev^3) / sum(xdev^2)^1.5
  return(r * sqrt(n) * (1 - 1/n)^1.5)
 }
```

# Background
In this document we use multilevel regression with post-stratification (MRP) to adjust population estimates of dissolved N2O gas concentration derived from sample data from the 2017 National Lakes Assessment (NLA).

As explained in the related document,
(https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html), duplicate gas samples were collected at a depth of ~0.1$\text{m}$ at designated index sites distributed across 1091 waterbodies nationwide, of which 95 were sampled twice. Gas samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L.

The data were collected under a stratified, unequal probability design. The observations can be thought of as hierarchically organized, such that each sample ($n \in n=1,...,N = 1185$) was situated within an index water body ($i \in i=1,...,I = 1091$), which were selected, with unequal probability from different size categories ($l \in l=1,...,L = 5$; < 4$\text{ha}$, 4-10$\text{ha}$, 10-20$\text{ha}$, 20-50$\text{ha}$, and >50$\text{ha}$), within a state ($k \in k=1,...,K = 48$), and an aggregated, Omernik ecoregion ($j \in j=1,...,J = 9$). All 9 ecoregions were represented in the sample, including Xeric (XER), Western Mountain (WMT),  Northern Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and Southern Appalachian (SAP) regions.

# Data
The first step is to read in the .RData file containing the observations, previously munged here: 
https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html

```{r import_data, eval=FALSE, include=TRUE}
load( file = paste0( localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/dg.2021-02-01.RData")
      )

save(dg, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda") 
```

## Observational data
The next step involves creating a new dataframe object including only variables of interest for this modeling exercise: the dissolved N2O observations; the design variables indexed to the observations; and the NO3 covariate observations. This step also includes: renaming some of these variables for convenience; rounding observations to the stated measurement precision above; and recoding the left-censored NO3 covariate in two different ways. The first treats the left-censored data as continuous, assigning to observations below the noted detection limit a value of 1/2 the limit. The second converts the NO3 data to an ordered factor. 
```{r mod1_data_import, paged.print=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda")

# number of observations before filtering probability samples
dg %>%
  filter(!is.na(dissolved.n2o.nmol)) %>% # remove obs with missing response measurements
  nrow()

df_model <- dg %>%
  filter(!is.na(dissolved.n2o.nmol)) %>%
  filter(sitetype == "PROB") %>% # probability samples only
  filter(visit.no == 1) %>%
  mutate(n2o = dissolved.n2o.nmol,
         n2o_eq = sat.n2o.nmol,
         n2o_sat = n2o.sat.ratio,
         n2o_em = e.n2o.nmol.d,
         n2o_flux = f.n2o.m.d,
         WSA9 = factor(ag.eco9),
         state = factor(state.abb[match(state.nm, state.name)]),
         size_cat = recode(area.cat6, 
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  mutate(size_cat = factor(size_cat,
                           levels = c("min_4", "4_10", "10_20", "20_50", "50_max"),
                           ordered = TRUE)) %>%
  mutate(no3 = ifelse(nitrate.n.result <= 0.005, 0.01, round(nitrate.n.result, 2))) %>% # 1/2 mdl
  mutate(no3_mis = ifelse( nitrate.n.result < 0.02, -1, 0)) %>%
  #mutate(no3_cat = cut(lognitrate.n.result, 
  #                     breaks = c(-Inf, 0.01, 0.05, 0.1, 0.25, 0.5, 1.0, Inf),
  #                     labels = c("1", "2", "3", "4", "5", "6", "7"))) %>%
  #                     #breaks = c(-Inf, 0.01, 0.10, 0.2, 0.5, Inf),
  #                     #labels = c("vL", "L", "M", "H", "vH"))) %>%
  mutate(no3_cat = cut(log(no3), 
                       breaks = c(-Inf, -4, -3, -2, -1, 1.0, Inf),
                       labels = c("1", "2", "3", "4", "5", "6"))) %>%
                       #breaks = c(-Inf, 0.01, 0.10, 0.2, 0.5, Inf),
                       #labels = c("vL", "L", "M", "H", "vH"))) %>%
  mutate(no3_cat = factor(no3_cat,
                          levels = c("1", "2", "3", "4", "5", "6"),
                          ordered = TRUE)) %>%
  select(WSA9,
         state,
         size_cat,
         n2o,
         n2o_eq,
         n2o_sat,
         n2o_em,
         n2o_flux,
         no3,
         no3_mis,
         no3_cat
         )

save(df_model, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

nrow(df_model) # number of obs after filtering for probability samples and removing one site missing ecoregion (WSA9) info.
print(df_model)
```

## Population data
The next step is to import the sample frame, which will be used in the post-stratification step to adjust the sample inferences to the population of interest.
```{r import_sample_frame, eval=FALSE, include=TRUE}
sframe <- read.csv( file = paste0( localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/NLA_Sample_Frame.csv" ), header = T )

sframe <- sframe %>%
  filter(nla17_sf != "Exclude2017") %>%
  filter(nla17_sf != "Exclude2017_Include2017NH") %>%
  mutate(WSA9 = factor(ag_eco9),
          state = factor(state),
          size_cat = factor(area_cat6)) %>% 
  select(WSA9, state, size_cat, area_ha, perim_km) %>% 
  filter(state != "DC") %>%
  filter(state != "HI") %>%
  droplevels() %>%
  mutate(WSA9 = fct_drop(WSA9)) %>% # remove NA level
  #filter( size_cat != "(0,1]" ) %>%
  mutate(size_cat = recode(size_cat, 
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  mutate(size_cat = factor(size_cat, 
                           levels = c("min_4", "4_10", "10_20", "20_50", "50_max"),
                           ordered = TRUE))

save(sframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

print(sframe)
```

Next, filter and select the frame down to only the observations and design variables needed for post-stratification. Some of the variables are again renamed to match the naming conventions used in the observational data above. DC and HI are also filtered out because the population of interest includes only the lower 48 states.

The post-stratification cells are summarized next.
```{r filter_frame, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

pframe <- sframe %>%
  select(WSA9, state, size_cat) %>% 
  mutate( obs = 1 ) %>%
  group_by( WSA9, state, size_cat ) %>%
  summarise( n_cell = sum( obs ) ) %>%
  ungroup() %>%
  mutate(prop_cell = n_cell/sum(n_cell)) %>%
  mutate(type = "population")

save(pframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

print(pframe)
```
Above there are 536 rows indicating that, with respect to the probabilistic sampling design based on ecoregion (WSA9), state, and size category, there are 536 types of sites in the population of interest; and the counts and proportions of those types relative to the population are indicated for each row.

## Sample vs. population
The cell distributions in the population of interest are compared to the proportions in the sample below.
```{r sample_cell_counts, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

samp_props <- df_model %>%
  mutate(obs = 1) %>%
  group_by(WSA9, state, size_cat) %>%
  summarize(n_cell = sum(obs)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_cell / sum(n_cell), 7)) %>%
  mutate(type = "sample")

save(samp_props, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

print(samp_props)
```
Above, it is show that there are 352 site types (rows) in the sample compared to 536 in the population. The multilevel model will enable inferences to the missing types.

Below is a graphical comparison of the distribution of cells in the population of interest vs. the sample.
```{r compare_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

pframe %>%
  bind_rows(samp_props) %>%
  ggplot(aes(x = interaction(WSA9, state, size_cat), y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes( shape = type, color = type)) +
  geom_line() +
  theme_tidybayes() +
  theme(axis.text.x = element_blank()) +
  xlab("WSA9:state:size") +
  ylab("proportion in cell")
```
In the above graphic, it can be seen that the sample proportions don't match the proportions in the population of interest. This was by design, since this was a stratified probabilistic sample.

Below is a comparison of population vs. sample distributions by ecoregion.
```{r eco_props_pop, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

pframe_eco <- pframe %>%
  group_by(WSA9) %>%
  summarise(n_cell = sum(n_cell)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_cell/sum(n_cell), 7)) %>%
  ungroup() %>%
  mutate(type = 'population')

save(pframe_eco, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_eco.rda")
```

```{r eco_props_sample, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

samp_props_eco <- samp_props %>%
  group_by(WSA9) %>%
  summarise(n_cell = sum(n_cell)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_cell/sum(n_cell), 7)) %>%
  ungroup() %>%
  mutate(type = 'sample')

save(samp_props_eco, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_eco.rda")
```

```{r compare_eco_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=6}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_eco.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_eco.rda")

pframe_eco %>%
  bind_rows(samp_props_eco) %>%
  ggplot(mapping = aes(x = WSA9, y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes( shape = type, color = type), size = 3) +
  geom_line() +
  theme_tidybayes() +
  xlab("Ecoregion") +
  ylab("proportion in cell") + 
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  theme(text = element_text(size = 12))
```
The graphic above clearly indicates undersampling in the CPL relative to the population and oversampling in the WMT and XER ecoregions.

The comparison can also be done by state.
```{r state_props_pop, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

pframe_state <- pframe %>%
  group_by(state) %>%
  summarise(n_cell = sum(n_cell)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_cell/sum(n_cell), 7)) %>%
  ungroup() %>%
  mutate(type = 'population')

save(pframe_state, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_state.rda")
```

```{r state_props_sample, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

samp_props_state <- samp_props %>%
  group_by(state) %>%
  summarise(n_cell = sum(n_cell)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_cell/sum(n_cell), 7)) %>%
  ungroup() %>%
  mutate(type = 'sample')

save(samp_props_state, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_state.rda")
```

```{r compare_state_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_state.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_state.rda")

pframe_state %>%
  bind_rows(samp_props_state) %>%
  ggplot(mapping = aes(x = state, y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes(shape = type, color = type)) +
  geom_line() +
  theme_tidybayes() +
  theme(axis.text.x = element_text(angle = 45)) +
  xlab("State") +
  ylab("proportion in cell")
```
The graphic above indicates clear undersampling, relative to the population of interest, in some states and oversampling in others.

The comparison by size class is next.
```{r size_props_pop, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

pframe_size <- pframe %>%
  group_by(size_cat) %>%
  summarise(n_cell = sum(n_cell)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_cell/sum(n_cell), 7)) %>%
  ungroup() %>%
  mutate(type = 'population')

save(pframe_size, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_size.rda")
```

```{r size_props_sample, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

samp_props_size <- samp_props %>%
  group_by(size_cat) %>%
  summarise(n_cell = sum(n_cell)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_cell/sum(n_cell), 7)) %>%
  ungroup() %>%
  mutate(type = 'sample')

save(samp_props_size, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_size.rda")
```

```{r compare_size_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=6}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_size.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_size.rda")

pframe_size %>%
  bind_rows(samp_props_size) %>%
  ggplot(mapping = aes(x = size_cat, y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes( shape = type, color = type)) +
  geom_line() +
  theme_tidybayes() +
  xlab("Size category") +
  ylab("proportion in cell")
```
Clearly, small streams were undersampled relative to the population of interest; and larger streams were oversampled.

## Sample-based estimates
Naive estimates using the sample for inference about the population of interest are presented below. First, the overall mean and standard deviation for N2O in the sample.
```{r sample_summary}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  summarise(mean = mean(n2o),
             sd = sd(n2o)) %>%
  print()
```

Next, a plot of the overall mean (dashed line) and ecoregion-specific means (black circles) based only on the sample. The shaded areas represent +/- 1 standard deviation.
```{r sample_summary_eco, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  group_by(WSA9) %>%
  summarise(mean = mean( n2o),
             sd = sd( n2o)) %>%
  ggplot(aes(x = WSA9, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = WSA9), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = WSA9, y = mean))+
  geom_point()+
  geom_hline(yintercept = 8.72, linetype = 'dashed') +
  theme_tidybayes() + 
  xlab("Ecoregion") +
  ylab("Sample mean")
```
The same summary by state is below.
```{r sample_summary_state, echo=FALSE, fig.align='center', fig.height=4, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  group_by(state) %>%
  summarise(mean = mean(n2o),
             sd = sd(n2o)) %>%
  ggplot(aes(x = state, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = state), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = state, y = mean))+
  geom_point()+
  geom_hline(yintercept = 8.72, linetype = 'dashed') +
  theme_tidybayes() +
  theme(axis.text.x = element_text( angle = 45)) +
  xlab("State") +
  ylab("Sample mean")
```
And finally by size class.
```{r sample_summary_size, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  group_by(size_cat) %>%
  summarise(mean = mean(n2o),
             sd = sd(n2o)) %>%
  ggplot(aes(x = size_cat, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = size_cat), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = size_cat, y = mean))+
  geom_point()+
  geom_hline(yintercept = 8.72, linetype = 'dashed') +
  theme_tidybayes() +
  theme(axis.text.x = element_text(angle = 45)) +
  xlab("Size category") +
  ylab("Sample mean")
```

## Additional graphical exploration
Below, the N2O observations are graphed according to the survey design factors. First, a plot of log(N2O) by WSA9 ecoregion.
```{r summary_N2O_vs_ecoregion, echo=FALSE, fig.align='center', fig.height=10, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = WSA9, y = log(n2o), group = WSA9, color = WSA9)) +
  geom_boxplot(outlier.shape = NA, color = "black" ,alpha = 0.7) + 
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  theme(text = element_text(size=12)) +
  coord_flip() +
  theme_bw()
```
Note the strong right-skew in the figure above, even with the N2O observations presented on the natural log scale. There is also some left-skew. Environmental concentrations and N2O, in particular, are typically modeled according to a lognormal or Gamma data generating process [e.g., @Webb_etal_2019]; however, the right tails associated with these N2O observations could prove tough to replicate even with a data generating process enforcing multiplicative errors.

Next, a plot of log(N2O) by states in ecoregions. Note that interacting state and ecoregion (i.e., state:ecoregion) reflects the nested nature of the data. That is, each ecoregion does not contain every state. Thus, state is nested in ecoregion.
```{r summary_N2O_vs_state, echo=FALSE, fig.align='center', fig.height=10, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = WSA9:state, y = log(n2o), group = WSA9:state, color = WSA9:state)) +
  geom_boxplot(outlier.shape = NA, color = "black" ,alpha = 0.7) + 
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  theme(text = element_text(size=12)) +
  coord_flip() +
  theme_bw()
```
In the figure above, strong left and right skew in log(N2O) remain. Both the means and variances may vary by grouping.

Next, a plot of size categories in each state in each ecoregion. Size category is also nested in state and ecoregion, since not every ecoregion:state grouping contains all size classes of lakes.  
```{r summary_N2O_vs_all, echo=FALSE, fig.align='center', fig.height=14, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = WSA9:state:size_cat, y = log(n2o), group = WSA9:state:size_cat, color = WSA9:state:size_cat)) +
  geom_boxplot(outlier.shape = NA, color = "black" ,alpha = 0.7) + 
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  theme_bw() +
  theme(text = element_text(size=12)) +
  theme(axis.text.y = element_blank()) +
  coord_flip()
```
In the figure above, the axis labels containing the ecoregion:state:size labels have been removed because the labels are messy and unecessary to convey the primary message: some clear skew remains in log(N2O) after considering all of the design factors. Again, the means and variances appear likely to differ across groupings.

Next, a look at NO3 vs. the survey variables under the assumption that wide variation in NO3 may be driving some of the skew in the N2O distributions. The first boxplot below shows just log(NO3), where NO3 observations below the detection limit have been replaced with a value 1/2 the detection limit.
```{r summary_NO3_vs_ecoregion, echo=FALSE, fig.align='center', fig.height=10, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = WSA9, y = log(no3), group = WSA9, color = WSA9)) +
  geom_boxplot(outlier.shape = NA, color = "black" ,alpha = 0.7) + 
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  theme(text = element_text(size=12)) +
  coord_flip() +
  theme_bw()
```
Note, above, the wide and skewed variation in log(NO3).

Next, a plot of log(N2O) vs. log(NO3).
```{r summary_N2O_vs_NO3, echo=FALSE, fig.align='center', fig.height=8, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = log(no3), y = log(n2o))) +
  geom_point(show.legend = F) +
  geom_smooth(method = "lm") +
  theme(text = element_text(size=12)) +
  theme_bw()
```

Next, a plot of log(N2O) vs. log(NO3) by ecoregion.
```{r summary_N2O_vs_NO3_ecoregion, echo=FALSE, fig.align='center', fig.height=8, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = log(no3), y = log(n2o), group = WSA9, color = WSA9)) +
  geom_point(show.legend = F) +
  geom_smooth(method = "lm") +
  facet_wrap(~ WSA9) +
  theme(text = element_text(size=12)) +
  theme_bw()
```
The above scatterplots indicate consistent, positive, linear associations between log(N2O) and log(NO3). The slope of the trend may also vary according to ecoregion. Note also the left bounds on log(no3) due to the left-censored nature of that variable. This figure suggests that NO3 could be a useful covariate to use in the multilevel regression model to help account for potential overdispersion (i.e., unmodeled variability in log(N2O)).

Similar plots are shown below, but with NO3 expressed as an ordered categorical variable with 6 levels.
```{r summary_N2O_vs_NO3cat, echo=FALSE, fig.align='center', fig.height=8, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = no3_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(outlier.shape = NA, color = "black" ,alpha = 0.7) +
  theme(text = element_text(size=12)) +
  theme_bw()
```

```{r summary_N2O_vs_NO3cat_ecoregion, echo=FALSE, fig.align='center', fig.height=8, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = no3_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(aes(x = no3_cat, y = log(n2o)), 
               outlier.shape = NA, color = "black" ,alpha = 0.7) + 
  facet_wrap(~ WSA9) +
  theme(text = element_text(size=12)) +
  theme_bw()
```
The trend is still apparent with the ordered categorical version of NO3.

Below, similar plots are shown for log(N2O) vs. lake size category.
```{r summary_N2O_vs_size, echo=FALSE, fig.align='center', fig.height=8, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = size_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(outlier.shape = NA, color = "black" ,alpha = 0.7) + 
  theme(text = element_text(size=12)) +
  theme_bw()
```

```{r summary_N2O_vs_size_ecoregion, echo=FALSE, fig.align='center', fig.height=8, fig.width=10, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = size_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(outlier.shape = NA, color = "black" ,alpha = 0.7) + 
  facet_wrap(~ WSA9) +
  theme(text = element_text(size=12)) +
  theme_bw()
```
In the two figures above, if there is any relationship between log(N2O) and size category, it appears likely to swamped by variation in log(N2O).

# Executing MRP
The survey had sampling bias by design, so the sample-based inferences above are also likely biased with respect to the population of interest. Therefore, the estimates should be adjusted to minimize inferential bias. To do that, a regression model can be fit to the observations to estimate expectations conditional on the design variables. Then, the resulting estimates can be adjusted based on the cell weighting in the population of interest [@Gelman_etal_2020 Ch. 17].

With a Bayesian multilevel regression model [e.g., @Gelman_etal_2014; @Gelman_etal_2020; @McElreath_2020], uncertainties are naturally propagated through the post-stratification step. A multilevel/hierarchical regression model also enables partial pooling of information across the groupings, which can regularize estimates and lead to more robust out-of-sample inferences. The degree of regularization itself is conditional on the observed data, influenced by both the size of the data (i.e., number of groupings and number of observations within each group) and the observed variation within and between groups. The multilevel model also lends itself naturally to estimates for missing groups. Recall from the summaries above that only 352 of the 536 cell types in the target population were represented in the sample. The missing cells are readily estimated using the multilevel parameterization because unobserved/missing levels of such hierarchical groupings above the observation level (e.g., a missing ecoregion) are treated as replicates from a defined population with its own distributional parameters (e.g., hierarchical mean and variance terms) that are estimated from the data. Thus, expectations and uncertainties for the unobserved levels are directly defined by the model. Estimates for missing poststratification cells is much less straightforward with a traditional "fixed effects" parameterization, which treats the individual levels of a grouping factor as independent with infinite group-level variance.

## Fitting a model
The first model estimates N2O concentration in a sample conditional on only the design factors.

### Model 1
```{r n2o_mod_mv_1, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(n2o ~ (1 | WSA9) + (1 | WSA9:state) + (1 | WSA9:state:size_cat),
             family = Gamma(link = "log"))
n2o_mod1 <- brm(bf_n2o,
                data = df_model, 
                prior = c(
                  prior(normal(2, 1), class = "Intercept"),
                  prior(normal(0, 0.5), class = "sd"),
                  prior(gamma(0.1, 0.1), class = "shape")
                  ),
  control = list(adapt_delta = 0.99, max_treedepth = 14),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 86541,
  chains=4, 
  iter=2000, 
  cores=4)

save(n2o_mod1, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
```

#### Summarize fit
We call a summary of the fitted parameters.

```{r print_mod1, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")

print(n2o_mod1, prior = T)
```

#### Model checks
A quick posterior predictive check to ensure the fit is a reasonable approximation of the true data generating process.

```{r ppc_dens_overlay_mod_n2o1, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
pp_check(n2o_mod1, type = "dens_overlay", nsamples = 100, cores = 1) + 
  theme_tidybayes() + xlim(0, 50) + xlab(expression(paste("N"[2],"O"))) + ylab("density")
```
In the figure above, replicated datasets (n = 100) from the fitted model are shown as blue lines, indicating the estimated density curve for each dataset. The black line is a density estimate for the observed dataset. Clearly the model replicates don't closely resemble the observed dataset. The fitted model generates datasets with much broader density and a lighter right tail. The observed data are much more concentrated between 0 and 10 mg/L N2O, and have a much longer right tail. The model is not capturing some important details about the observed data. In the next model below, more structure is introduced in an attempt to better capture the variation in the observations.

### Model 2
In the next model, a distributional term is included. This is a model specifying potential heterogeneous variances due to the design factors. As seen in the graphical exploration in the previous section, both the mean and variance of log(N2O) look to potentially vary by the survey factors.
```{r n2o_mod_mv_2, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(n2o ~ (1 | WSA9) + (1 | WSA9:state) + (1 | WSA9:state:size_cat),
             shape ~ (1 | WSA9) + (1 | WSA9:state),
             family = Gamma(link = "log"))
n2o_mod2 <- brm(bf_n2o,
                data = df_model, 
                prior = c(
                  prior(normal(1.9, 0.5), class = "Intercept"),
                  prior(normal(0, 1), class = "sd"),
                  prior(normal(0, 1), class = "sd", dpar= "shape")
                  ),
  #control = list(adapt_delta = 0.90, max_treedepth = 12),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 8531,
  chains=4, 
  iter=3000, 
  cores=4)

save(n2o_mod2, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
```

### Model 2

```{r n2o_mod_mv_2, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(n2o ~ mo(no3_cat) + mo(size_cat) +
               (mo(no3_cat) | a | WSA9) +
               (1 | b | WSA9:state) +
               (1 | c | WSA9:state:size_cat),
             shape ~ mo(no3_cat) +
               (1 | a | WSA9) +
               (1 | b | WSA9:state),
             family = Gamma(link = "log"))
bf_n2oeq <- bf(n2o_eq ~ 1 +
               (1 | a | WSA9) +
               (1 | b | WSA9:state) +
               (1 | c | WSA9:state:size_cat),
               shape ~ (1 | a | WSA9) +
               (1 | b | WSA9:state),
               family = Gamma(link = "log"))
bf_no3 <- bf(no3_cat ~ mo(size_cat) + 
               (1 | a | WSA9) +
               (1 | b | WSA9:state) +
               (1 | c | WSA9:state:size_cat),
             family = cumulative(link = "logit", threshold="flexible"))

n2o_mod <- brm(bf_n2o + 
                 bf_n2oeq + 
                 bf_no3 + 
                 set_rescor(FALSE),
  data = df_model, 
  prior = c(
      prior(normal(1.9, 0.5), class = "Intercept", resp ="n2o"),
      prior(normal(1.9, 0.5), class = "Intercept", resp = "n2oeq"),
      prior(normal(0, 2.5), class = "Intercept", resp = "no3cat"),

      #prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "n2o"),
      #prior(normal(0, 3), class = "Intercept", dpar = "shape", resp = "n2o"),

      #prior(normal(0, 1), class = "sigma", resp = "n2oeq"),
      #prior(normal(0, 3), class = "Intercept", dpar = "shape", resp = "n2oeq"),

      prior(normal(0, 1), class = "sd", resp = "n2o"),
      prior(normal(0, 1), class = "sd", resp = "n2oeq"),
      prior(normal(0, 1), class = "sd", resp = "no3cat"),

      prior(normal(0, 1), class = "sd", dpar= "shape", resp = "n2o"),
      prior(normal(0, 1), class = "sd", dpar= "shape", resp = "n2oeq")
      ),
  control = list(adapt_delta = 0.90, max_treedepth = 12),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 15224,
  chains=4, 
  iter=3000, 
  cores=4)

save(n2o_mod, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")
```

## Summarize fit
We call a summary of the fitted parameters.

```{r print_mod2, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")

print(n2o_mod, prior = T)
```

## Model checks
Before we get too far with this model, we'll want to do extensive checking to ensure it is a reasonable approximation of the true data generating process.

```{r ppc_dens_overlay_mod_n2o, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")
pp_check(n2o_mod, type = "dens_overlay", resp = "n2o", nsamples = 100, cores = 1) + 
  theme_tidybayes() + xlim(0, 50) + ggtitle(expression(paste("N"[2],"O")))#+ scale_x_continuous(trans = "log")

pp_check(n2o_mod, type = "dens_overlay", resp = "n2oeq", nsamples = 100, cores = 1) + 
  theme_tidybayes() + ggtitle(expression(paste("eqilibrium","N"[2],"O")))

pp_check(n2o_mod, type = "bars", resp = "no3cat", nsamples = 100, cores = 1) + 
  theme_tidybayes() +
  ggtitle(expression(paste("NO"[3], "\n category")))
```

## Leave-one-out evaluations
Next we'll conduct a Bayesian leave-one-out cross validation ('LOO-CV') assessment using the $\textbf{loo}$ package in $\textbf{R}$, which implements fast and stable computations for approximate LOO-CV (and WAIC) [@Vehtari_etal_2017]. These computation methods estimate pointwise, out-of-sample prediction accuracy from a fitted model by using the log-likelihood evaluated at the posterior simulations of the parameter values. Conveniently, the $\textbf{brms}$ package automatically calculates the log-likelihood for us for all relevant models, making it very simple to use in conjunction with the $\textbf{loo}$ package. For more information on $\textbf{loo}$ and LOO-CV in general, check out the page on the $\textbf{Stan}$ website: http://mc-stan.org/loo/.

### Calculate LOO-CV
```{r loo_overlay, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")
pp_check(n2o_mod, "loo_pit_overlay", resp = "n2o", cores = 1)
pp_check(n2o_mod, "loo_pit_overlay", resp = "n2oeq", cores = 1)
```

The distribution of LOO-PIT values (thick black line) is not clearly a plausible draw from the uniform. Note the rise around 0.25 (x-axis) and dip around 0.6. The calibration looks like it is maybe slightly off, perhaps with a bias induced by compensation for extremes noted in the measured vs. predicted check. The checks below will investigate this potential mis-calibration further.

Next is a plot of the LOO marginal predictive distribution for each of the observations in the study, divided among the different methods and ordered by the median.

Clear pattern of mismatches between the observations ($y$) and the LOO prediction intervals ($y_{rep}$) and can indicate over-fitting, bias, under- or overdispersion, _etc._. The broad expectation is roughly even over- and under-prediction with roughly 50% of the observations falling outside the 50% intervals and 95% to outside the 95% intervals.
```{r obs_v_fitted, eval=FALSE, include=TRUE, fig.align='center', fig.height=5, fig.width=5}
load("full-analysis-files/n2o_mod.rda")

pp_check(n2o_mod, resp = "n2o", "scatter_avg", cores = 1)

pp_check(n2o_mod, resp = "n2oeq", "scatter_avg", cores = 1)
```

```{r loo_intervals, eval=FALSE, include=TRUE, fig.align='center', fig.height=20, fig.width=10}
load("full-analysis-files/n2o_mod.rda")

pp_check(n2o_mod, resp = "n2o", "loo_intervals", order = "median", prob = 0.5, prob_outer = 0.95, cores = 1) +
  facet_wrap( ~ df_model$WSA9, ncol = 1, scales = "free") + scale_y_continuous(trans = "log")

pp_check(n2o_mod, resp = "n2oeq", "loo_intervals", order = "median", prob = 0.5, prob_outer = 0.95, cores = 1) +
  facet_wrap( ~ df_model$WSA9, ncol = 1, scales = "free")
```

# Population estimates
Make a random sample from the population, proportional to the estimated cell sizes in the population. 
```{r psframe}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

ID <- sample(1:536, # rows in pframe
                  10000, # size of the sample / any size could be used, depending on desired precision
                  prob = pframe$n_cell, # probability set proportional to cell size
                  replace = T)

psframe <- pframe[ID, 1:3] %>%
  print()

save(psframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/psframe.rda")
```

Next, we need to post-stratify the no3 predictions, in order to carry that uncertainty through and predict n2o.
```{r no3_predict_psframe}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/psframe.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")

predict_no3 <- psframe %>%
  add_predicted_draws(model = n2o_mod, resp="no3cat", allow_new_levels = TRUE, cores =1, n = 500) %>% # 500 draws reduce size
  mutate(no3_cat = .prediction) %>%
  print()

save(predict_no3, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
```

```{r loop_predictions}
predict_n2o <- predict_no3[,c(1:4, 7, 9)] %>%
  mutate(n2o_pred = 0,
         n2oeq_pred = 0)

predict_n2o$pr_grp <- rep(1:100, each = 50000)

system.time(
for(i in 1:max(predict_n2o$.draw)){
  predict_n2o[which(predict_n2o$pr_grp == i), 7:8] <- t(apply(posterior_predict(n2o_mod, newdata = predict_n2o[which(predict_n2o$pr_grp == i),], resp = c("n2o", "n2oeq"), allow_new_levels = T, nsamples = 500, cores = 1), 2, diag))
}
)
```

```{r loop_parallel_predict}
require(foreach)
require(itertools)

cl <- parallel::makeCluster(6) # set number of cores to use for emb. parallel predictions
doSNOW::registerDoSNOW(cl) # register the workers

predict_X <- predict_no3[,c(1:3, 11)] %>%
  ungroup()

predict_X$pr_grp <- as.integer(c(rep(1:2675, each = 87000), rep(2676, 223500)))

system.time(
preds <- foreach(i = 1:max(predict_X$pr_grp),
                 .combine = rbind,
                 .packages = c("brms")) %dopar% {
  t(apply(posterior_predict(n2o_mod, 
                            newdata = predict_X[which(predict_X$pr_grp == i),], 
                            resp = c("n2o", "n2oeq"), 
                            allow_new_levels = T, 
                            nsamples = 500, 
                            cores = 1),
          2, diag))
                 })

parallel::stopCluster(cl)

colnames(preds) <- c("preds_n2o", "preds_n2oeq")

save(preds, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/preds.rda")

predict_n2o <- predict_no3 %>% 
  ungroup() %>%
  cbind(preds) %>% 
  as_tibble() %>%
  mutate(preds_sat = preds_n2o / preds_n2oeq)

all_predictions <- predict_n2o[,c(1:6, 9, 11:14),] # remove unnecessary columns

save(all_predictions, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

rm(preds) # no longer needed
rm(predict_X)
rm(predict_no3)
rm(predict_n2o)
```

Lets first summarize at national level.
## National
### $N_2O$
#### Mean
```{r group_mean}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(preds_n2o)) %>%
  summarise( posterior_median = round(median(mean_n2o), 1),
    LCI = round(quantile(mean_n2o, probs = 0.025), 1),
    UCI = round(quantile(mean_n2o, probs = 0.975), 1))
```

```{r graph wsa9_n2o_estimates, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_n2o = mean(preds_n2o)) %>%
  summarise( estimate = round(median(mean_n2o), 1),
    LCL = round(quantile(mean_n2o, probs = 0.025), 1),
    UCL = round(quantile(mean_n2o, probs = 0.975), 1)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  bind_rows(cbind(survey_ests[-10,], type = rep("survey", 9))) %>%
  ggplot( aes( x = ecoregion, y = estimate, group = type ) ) +
  geom_point( position=position_dodge(width=0.5) ) +
  geom_linerange( aes( ymin = LCL, ymax = UCL ) , position=position_dodge(width=0.5) ) +
  scale_colour_manual( values = c( "black", "grey" ) ) +
  #scale_y_continuous(limits = c(0.8, 2), breaks = seq(0.8, 2, 0.1)) +
  ylab("mean N2O concentration") +
  coord_flip() + 
  theme_bw()
```

```{r graph wsa9_sat_estimates, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_sat = mean(preds_sat)) %>%
  summarise( estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  #bind_rows(cbind(survey_ests, type = rep("survey", 10))) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point( position=position_dodge(width=0.5) ) +
  geom_linerange( aes( ymin = LCL, ymax = UCL ) , position=position_dodge(width=0.5) ) +
  #geom_point(aes( x = ecoregion, y = sample_mean), shape = 8, color='red') +
  #scale_colour_manual( values = c( "black", "grey" ) ) +
  #scale_y_continuous(limits = c(0.8, 2), breaks = seq(0.8, 2, 0.1)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()
```
```{r graph wsa9_prop_undersat, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point( position=position_dodge(width=0.5) ) +
  geom_linerange( aes( ymin = LCL, ymax = UCL ) , position=position_dodge(width=0.5) ) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()
```

```{r postrat_mean_n2o, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")

posterior_mean <- posterior_epred(n2o_mod, 
                                  newdata = predict_no3, 
                                  resp = c("n2o", "n2oeq"),
                                  allow_new_levels = TRUE, 
                                  cores = 1,
                                  nsamples = 1000)

poststrat_mean_n2o <- (posterior_mean[,,1] %*% predict_no3$n_cell) / sum(predict_no3$n_cell) # div by nsims pstrat_no3
poststrat_mean_n2oeq <- (posterior_mean[,,2] %*% predict_no3$n_cell) / sum(predict_no3$n_cell)

popn_mean_summary_n2o <- c(posterior_median = median(poststrat_mean_n2o),
                           posterior_sd = sd(poststrat_mean_n2o), 
                           LCI = quantile(poststrat_mean_n2o, 0.025),
                           UCI = quantile(poststrat_mean_n2o, 0.975))

popn_mean_summary_n2oeq <- c(posterior_median = median(poststrat_mean_n2oeq),
                           posterior_sd = sd(poststrat_mean_n2oeq), 
                           LCI = quantile(poststrat_mean_n2oeq, 0.025),
                           UCI = quantile(poststrat_mean_n2oeq, 0.975))

save(posterior_mean, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/posterior_mean.rda")

save(poststrat_mean_n2o, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/poststrat_mean_n2o.rda")
save(poststrat_mean_n2oeq, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/poststrat_mean_n2oeq.rda")

round(popn_mean_summary_n2o, 2)
round(popn_mean_summary_n2oeq, 2)
```

Compare to sample means.

```{r sample_mean_n2o}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

round(mean(df_model$n2o), 2)
round(mean(df_model$n2o_eq), 2)
```

We can compare the n2o estimate to the survey estimate as well.
```{r survey_ests}
#make a dataframe object

survey_ests <- data.frame( 
  ecoregion = c( "CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER", "US" ), 
  estimate = c( 8.4, 7.7, 6.9, 7.1, 6.5, 7.8, 10.8, 7.8, 10.6, 8.1 ),
  LCL = c( 5.1, 7.2, 6.4, 5.9, 4.9, 5.9, 6.4, 7.1, 7.5, 7.0 ),
  UCL = c( 11.7, 8.1, 7.4, 8.2, 8.1, 9.6, 15.2, 8.4, 13.7, 9.1 )
) %>%
  mutate( ecoregion = factor( ecoregion, 
                              levels = c( "CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER", "US" ) ) ) #relevel for ggplot so not alphabetical, but original order

save(survey_ests, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/survey_ests.rda")

print( survey_ests[10,] )
```

## Saturation ratio
```{r postrat_mean_n2o_eq, message=FALSE}
posterior_mean_sat <- posterior_mean[,,1] / posterior_mean[,,2]

poststrat_mean_sat  <- (posterior_mean_sat %*% predict_no3$n_cell) / sum(predict_no3$n_cell)
popn_mean_summary_sat <- c(posterior_median = median(poststrat_mean_sat), 
                     posterior_sd = sd(poststrat_mean_sat), 
                     LCI = quantile(poststrat_mean_sat, 0.025), 
                     UCI = quantile(poststrat_mean_sat, 0.975))

round(popn_mean_summary_sat, 2)
```

We can compare that to the sample estimate.
```{r sample_mean_again}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

round(mean(df_model$n2o / df_model$n2o_eq), 2)
```

#### Predictive distribution
```{r posterior_pr_n2o, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")

posterior_pr <- posterior_predict(n2o_mod, 
                                  newdata = predict_no3, 
                                  resp = c("n2o", "n2oeq"), 
                                  allow_new_levels = TRUE, 
                                  cores = 1,
                                  nsamples = 500)

save(posterior_pr, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/posterior_pr.rda")

#pp <- posterior_pr[,rbinom(536000, 1, prob = predict_no3$prop_cell*16) == 1]
```

```{r poststrat_nat_n2o}
poststrat_pr_n2o <- (posterior_pr[,,1] * predict_no3$n_cell) / (predict_no3$n_cell)
poststrat_pr_n2oeq <- (posterior_pr[,,2] * predict_no3$n_cell) / (predict_no3$n_cell)

popn_pr_summary_n2o <- c(posterior_mean = mean(poststrat_pr_n2o), 
                     posterior_sd = sd(poststrat_pr_n2o), 
                     LCI = quantile(poststrat_pr_n2o, 0.025), 
                     UCI = quantile(poststrat_pr_n2o, 0.975))

popn_pr_summary_n2oeq <- c(posterior_mean = mean(poststrat_pr_n2oeq), 
                     posterior_sd = sd(poststrat_pr_n2oeq), 
                     LCI = quantile(poststrat_pr_n2oeq, 0.025), 
                     UCI = quantile(poststrat_pr_n2oeq, 0.975))

save(poststrat_pr_n2o, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/poststrat_pr_n2o.rda")
save(poststrat_pr_n2oeq, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/poststrat_pr_n2oeq.rda")

round(popn_pr_summary_n2o, 2)
round(popn_pr_summary_n2oeq, 2)
```

### Predictive distribution
```{r postrat_pred_sat_ratio, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")

posterior_pr_n2o <- posterior_predict(n2o_mod, 
                                  newdata = pstrat_no3, 
                                  resp = "n2o",
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

posterior_pr_n2oeq <- posterior_predict(n2o_mod, 
                                  newdata = pstrat_no3, 
                                  resp = "n2oeq",
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 3000)

posterior_pr_n2o <- exp(posterior_pr_n2o)
posterior_pr_n2o_eq <- exp(posterior_pr_n2oeq)

posterior_pr_sat <- posterior_pr_n2o / posterior_pr_n2o_eq

poststrat_pr_sat <- (posterior_pr_sat %*% pstrat_no3$prop_cell) / 1000 # div by nsims pstrat_no3

model_popn_pref <- c(posterior_mean = mean(poststrat_pr_sat), 
                     posterior_sd = sd(poststrat_pr_sat), 
                     LCI = quantile(poststrat_pr_sat, 0.025), 
                     UCI = quantile(poststrat_pr_sat, 0.975))

save(poststrat_pr_sat, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/poststrat_pr_sat.rda")

round(model_popn_pref, 2)
```

# Ecoregion estimates
## Mean
We can also post-stratify and make estimates at the ecoregion level.
```{r poststrat_eco_estimates}
#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency #(EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pstrat_no3.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")

eco_df <- data.frame(
  ecoregion = levels(as.factor(pstrat_no3$WSA9)),
  posterior_median = rep(-1, 9),
  sd = rep(-1, 9),
  LCI = rep(-1, 9),
  UCI = rep(-1, 9),
  sample_mean = rep(-1, 9),
  N = rep(-1, 9)
)

for(i in 1:length(levels(as.factor(pstrat_no3$WSA9)))) {
  poststrat_eco <- pstrat_no3[pstrat_no3$WSA9 == levels(as.factor(pstrat_no3$WSA9))[[i]], ]

  posterior_mu <- posterior_epred(n2o_mod, 
                                  newdata = poststrat_eco, 
                                  dpar = "mu", 
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

  posterior_sigma <- posterior_epred(n2o_mod, 
                                  newdata = poststrat_eco, 
                                  resp = "logn2o", 
                                  dpar = "sigma", 
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

  posterior_sigma_eq <- sample(posterior_samples(n2o_mod, resp = "logn2oeq")$sigma_logn2oeq, 1000, replace = FALSE)

  posterior_mean <- exp(posterior_mu[, , 1] + 0.5 * posterior_sigma ^ 2)

  posterior_mean_eq <- exp(posterior_mu[, , 2] + 0.5 * posterior_sigma_eq ^ 2)

  posterior_mean_sat <- posterior_mean / posterior_mean_eq

  poststrat_mean_sat <- (posterior_mean_sat %*% poststrat_eco$prop_cell) / 1000

  poststrat_prob_eco <- ((posterior_mean_sat %*% poststrat_eco$n_cell) / sum(poststrat_eco$n_cell))
  #This is the estimate for popn in ecoregion:
  eco_df$posterior_median[i] <- round(median(poststrat_prob_eco), 3)
  eco_df$sd[i] <- round(sd(poststrat_prob_eco), 3)
  eco_df$LCI[i] <- round(quantile(poststrat_prob_eco, 0.025), 3)
  eco_df$UCI[i] <- round(quantile(poststrat_prob_eco, 0.975), 3)
  #This is the estimate for sample
  eco_df$sample_mean[i] <- round(mean(df_model$n2o[df_model$WSA9 == levels(as.factor(pframe$WSA9))[[i]]] / 
                                        df_model$n2o_eq[df_model$WSA9 == levels(as.factor(pframe$WSA9))[[i]]]), 3)
  eco_df$N[i] <- length(df_model$n2o[df_model$WSA9 == levels(as.factor(pframe$WSA9))[[i]]])
}

save(eco_df, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/eco_sat_df.rda")

print(eco_df)
```
We can compare those to the survey estimates as well.
```{r survey_mean_again_eco}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/survey_ests.rda")

print(survey_ests[-10, ])
```
Lets do a graphical comparison at the ecoregion scale. We'll include the sample means as well (red asterisk).
```{r graph eco_estimates, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/eco_sat_df.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/survey_ests.rda")

eco_df %>%
  select(ecoregion, posterior_median, LCI, UCI, sample_mean) %>%
  rename(estimate = posterior_median,
         LCL = LCI,
         UCL = UCI) %>%
  mutate(ecoregion = factor(ecoregion)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point( position=position_dodge(width=0.5) ) +
  geom_linerange( aes( ymin = LCL, ymax = UCL ) , position=position_dodge(width=0.5) ) +
  geom_point(aes( x = ecoregion, y = sample_mean), shape = 8, color='red') +
  scale_colour_manual( values = c( "black", "grey" ) ) +
  scale_y_continuous(limits = c(0.8, 2), breaks = seq(0.8, 2, 0.1)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()
```
## Predictive distribution
We can also post-stratify and make estimates at the ecoregion level.
```{r poststrat_eco_estimates}
#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency #(EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pstrat_no3.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod.rda")

eco_df <- data.frame(
  ecoregion = levels(as.factor(pstrat_no3$WSA9)),
  posterior_median = rep(-1, 9),
  sd = rep(-1, 9),
  LCI = rep(-1, 9),
  UCI = rep(-1, 9),
  sample_mean = rep(-1, 9),
  N = rep(-1, 9)
)

for(i in 1:length(levels(as.factor(pstrat_no3$WSA9)))) {
  poststrat_eco <- pstrat_no3[pstrat_no3$WSA9 == levels(as.factor(pstrat_no3$WSA9))[[i]], ]

  posterior_pr <- posterior_epred(n2o_mod, 
                                  newdata = poststrat_eco, 
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

  posterior_pr_n2o <- exp(posterior_pr[,,1])

  posterior_pr_n2o_eq <- exp(posterior_pr[, , 2])

  posterior_pr_sat <- posterior_pr_n2o / posterior_pr_n2o_eq

  poststrat_pr_sat <- (posterior_pr_sat %*% poststrat_eco$prop_cell) / 1000

  #This is the estimate for popn in ecoregion:
  eco_df$posterior_median[i] <- round(median(poststrat_pr_sat), 3)
  eco_df$sd[i] <- round(sd(poststrat_pr_sat), 3)
  eco_df$LCI[i] <- round(quantile(poststrat_pr_sat, 0.025), 3)
  eco_df$UCI[i] <- round(quantile(poststrat_pr_sat, 0.975), 3)
  #This is the estimate for sample
  eco_df$sample_mean[i] <- round(mean(df_model$n2o[df_model$WSA9 == levels(as.factor(pframe$WSA9))[[i]]] / 
                                        df_model$n2o_eq[df_model$WSA9 == levels(as.factor(pframe$WSA9))[[i]]]), 3)
  eco_df$N[i] <- length(df_model$n2o[df_model$WSA9 == levels(as.factor(pframe$WSA9))[[i]]])
}

save(eco_df, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/eco_sat_pr_df.rda")

print(eco_df)
```

# State estimates
Lets post-stratify at the state level now.
```{r poststrat_state_estimates}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/mod.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

state_df <- data.frame(
  state = levels(as.factor(df_model$state)),
  posterior_median = rep(-1, 48),
  sd = rep(-1, 48),
  LCI = rep(-1, 48),
  UCI = rep(-1, 48),
  sample_mean = rep(-1, 48),
  N = rep(-1, 48)
)

for(i in 1:length(levels(factor(df_model$state)))) {
  poststrat_state <- pstrat_no3[pstrat_no3$state == levels(factor(pstrat_no3$state))[[i]], ]
    posterior_mu <- posterior_epred(n2o_mod, 
                                  newdata = poststrat_state, 
                                  dpar = "mu", 
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

  posterior_sigma <- posterior_epred(n2o_mod, 
                                  newdata = poststrat_state, 
                                  resp = "logn2o", 
                                  dpar = "sigma", 
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

  posterior_sigma_eq <- sample(posterior_samples(n2o_mod, resp = "logn2oeq")$sigma_logn2oeq, 1000, replace = FALSE)

  posterior_mean <- exp(posterior_mu[, , 1] + 0.5 * posterior_sigma ^ 2)

  posterior_mean_eq <- exp(posterior_mu[, , 2] + 0.5 * posterior_sigma_eq ^ 2)

  posterior_mean_sat <- posterior_mean / posterior_mean_eq

  poststrat_mean_sat <- (posterior_mean_sat %*% poststrat_state$prop_cell) / 1000

  poststrat_prob_state <- ((posterior_mean_sat %*% poststrat_state$n_cell) / sum(poststrat_state$n_cell))

  #This is the estimate for popn in state:
  state_df$posterior_median[i] <- round(median(poststrat_prob_state), 2)
  state_df$sd[i] <- round(sd(poststrat_prob_state), 2)
  state_df$LCI[i] <- round(quantile(poststrat_prob_state, 0.025), 2)
  state_df$UCI[i] <- round(quantile(poststrat_prob_state, 0.975), 2)
  #This is the estimate for sample
  state_df$sample_mean[i] <- round(mean(df_model$n2o[df_model$state == levels(as.factor(df_model$state))[[i]]]), 2)
  state_df$N[i] <- length(df_model$n2o[df_model$state == levels(as.factor(df_model$state))[[i]]])
}

save(state_df, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/state_sat_df.rda")

print(state_df)
```

We don't yet have the survey design-based estimates for the state level, but lets plot the MRP state estimates (black) vs. the sample means (red asterisk).
```{r graph state_estimates, fig.align='center', fig.height=8, fig.width=8, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/state_sat_df.rda")

state_df %>%
  mutate(state = factor(state)) %>%
  select(state, sample_mean, posterior_median, LCI, UCI) %>%
  rename(estimate = posterior_median,
         LCL = LCI,
         UCL = UCI,
         sample_mean = sample_mean) %>%
  mutate( state = fct_reorder( state, estimate)) %>%
  ggplot(aes(x = state, y = estimate)) +
  geom_point() +
  geom_point(aes( x = state, y = sample_mean), shape = 8, color='red') +
  geom_linerange(aes(ymin = LCL, ymax = UCL)) +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() +
  theme_tidybayes()
```

# Size category estimates
Finally, we'll look at the stream size level.
```{r poststrat_size_estimates}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/mod.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

size_df <- data.frame(
  size = levels(as.factor(df_model$size)),
  posterior_median = rep(-1, 5),
  sd = rep(-1, 5),
  LCI = rep(-1, 5),
  UCI = rep(-1, 5),
  sample_mean = rep(-1, 5),
  N = rep(-1, 5)
)

for(i in 1:length(levels(factor(df_model$size_cat)))) {
  poststrat_size <- pstrat_no3[pstrat_no3$size_cat == levels(factor(pstrat_no3$size_cat))[[i]], ]
    posterior_mu <- posterior_epred(n2o_mod, 
                                  newdata = poststrat_size, 
                                  dpar = "mu", 
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

  posterior_sigma <- posterior_epred(n2o_mod, 
                                  newdata = poststrat_size, 
                                  resp = "logn2o", 
                                  dpar = "sigma", 
                                  allow_new_levels = TRUE, 
                                  cores = 1, 
                                  nsamples = 1000)

  posterior_sigma_eq <- sample(posterior_samples(n2o_mod, resp = "logn2oeq")$sigma_logn2oeq, 1000, replace = FALSE)

  posterior_mean <- exp(posterior_mu[, , 1] + 0.5 * posterior_sigma ^ 2)

  posterior_mean_eq <- exp(posterior_mu[, , 2] + 0.5 * posterior_sigma_eq ^ 2)

  posterior_mean_sat <- posterior_mean / posterior_mean_eq

  poststrat_mean_sat <- (posterior_mean_sat %*% poststrat_size$prop_cell) / 1000

  poststrat_prob_size <- ((posterior_mean_sat %*% poststrat_size$n_cell) / sum(poststrat_size$n_cell))
  #This is the estimate for popn in size:
  size_df$posterior_median[i] <- round(median(poststrat_prob_size), 2)
  size_df$sd[i] <- round(sd(poststrat_prob_size), 2)
  size_df$LCI[i] <- round(quantile(poststrat_prob_size, 0.025), 2)
  size_df$UCI[i] <- round(quantile(poststrat_prob_size, 0.975), 2)
  #This is the estimate for sample
  size_df$sample_mean[i] <- round(mean(df_model$n2o[df_model$size_cat == levels(as.factor(df_model$size_cat))[[i]]] / 
                                         df_model$n2o_eq[df_model$size_cat == levels(as.factor(df_model$size_cat))[[i]]]), 2)
  size_df$N[i] <- length(df_model$n2o[df_model$size_cat == levels(as.factor(df_model$size_cat))[[i]]])
}

save(size_df, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/size_sat_df.rda")

print(size_df)
```

And plot them.
```{r graph size_estimates, fig.align='center', fig.height=4, fig.width=6, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/size_sat_df.rda")

size_df %>%
  mutate(size = factor(size)) %>%
  select(size, sample_mean, posterior_median, LCI, UCI) %>%
  rename(estimate = posterior_median,
         LCL = LCI,
         UCL = UCI) %>%
  mutate(size = fct_reorder( size, estimate)) %>%
  ggplot(aes(x = size, y = estimate)) +
  geom_point() +
  geom_point(aes( x = size, y = sample_mean), shape = 8, color='red')+
  geom_linerange(aes(ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() +
  theme_tidybayes()
```

# References

### 2023-02-08 scripts/population.Rmd

---
title: "Explore model results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load the packages
library(tidyverse)
library(ggplot2)

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")
```

## Population Data

```{r}
# about 15 minutes to load
Sys.time()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
Sys.time()
```

## Exploratory analysis
```{r}

# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(preds_n2o)) %>%
  summarise( posterior_median = round(median(mean_n2o), 1),
    LCI = round(quantile(mean_n2o, probs = 0.025), 1),
    UCI = round(quantile(mean_n2o, probs = 0.975), 1))

# plot N2O conc means by wsa9
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_n2o = mean(preds_n2o)) %>%
  summarise( estimate = round(median(mean_n2o), 1),
    LCL = round(quantile(mean_n2o, probs = 0.025), 1),
    UCL = round(quantile(mean_n2o, probs = 0.975), 1)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  #scale_y_continuous(limits = c(0.8, 2), breaks = seq(0.8, 2, 0.1)) +
  ylab("mean N2O concentration") +
  coord_flip() + 
  theme_bw()

# plot N2O sat ratio means by WSA9
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_sat = mean(preds_sat)) %>%
  summarise( estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()

# plot proportion waterbodies functioning as N2O sinks by WSA9
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()

```

### 2023-02-08 scripts/dgIndicatorAnalysis.Rmd
Commiting changes to facilitate merge with github

--- 2022-12-14 scripts/dgIndicatorAnalysis.Rmd
@@
 library(micromap) # linked micromaps
+library(tictoc)
 
@@
 # 11 minutes on VM!
+tic() # 7 minutes on laptop
 all_predictions <- all_predictions %>%
@@
   select(-area_km2)
+toc()
+gc() # this pushed memory up to 57.7 GB!  gc drops it to 32GB
 
@@
                        labels = c("Northern Plains", "Temperate Plains",
-                                  "Upper Midwest", "Coastal Plains", "Southern Plains", "Southern Appalachians", "Xeric", "Northern Appalachians", "Western Mountains")))
-
+                                  "Upper Midwest", "Coastal Plains", "Southern Plains", 
+                                  "Southern Appalachians", "Xeric", "Northern Appalachians", 
+                                  "Western Mountains")))
+
+```
+
+## N2O Concentration
+I think we should start with some plots of the primary measurement value.
+```{r}
+
+# PLOT N2O CONCENTRATION BY WSA9 AND LAKE SIZE
+
+# N2O CONCENTRATION BY ECOREGION AND SIZE
+# point and linerange
+all_predictions %>%
+  group_by(WSA9, size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_n2o = mean(n2o)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_n2o), 3), 
+    LCL = round(quantile(mean_n2o, probs = 0.025), 3),
+    UCL = round(quantile(mean_n2o, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  select(-WSA9) %>% # necessary?
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = size_cat, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  ylab("mean N2O concentration (nmol/L)") +
+  coord_flip() + 
+  theme_bw() +
+  facet_wrap(~ecoregion)
+ggsave("output/figures/n2oConcentrationByEcoSize.tiff")
+
+# N2O CONCENTRATION BY ECOREGION
+# point and linerange
+all_predictions %>%
+  group_by(WSA9, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_n2o = mean(n2o)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_n2o), 3), 
+    LCL = round(quantile(mean_n2o, probs = 0.025), 3),
+    UCL = round(quantile(mean_n2o, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = ecoregion, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  ylab("mean N2O concentration (nmol/L)") +
+  coord_flip() + 
+  theme_bw()
+ggsave("output/figures/n2oConcentrationByEco.tiff")
 ```
@@
 ```{r}
-# PLOT N2O SAT RATIO BY WSA9
+# PLOT N2O SAT RATIO BY WSA9------------------
 
@@
 all_predictions %>%
-  group_by(WSA9, .draw) %>%
+  group_by(WSA9, .draw) %>% # group by iteration
   #group_by(.draw) %>%
-  summarise( mean_sat = mean(preds_sat)) %>%
+  summarise(mean_sat = mean(n2osat)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
   summarise( estimate = round(median(mean_sat), 3), 
@@
   theme_bw()
+ggsave("output/figures/n2oSatRatioByEco.tiff")
 
@@
   #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
-  ggplot( aes( x = preds_sat, group = .draw, color = .draw ) ) +
+  ggplot(aes(x = n2osat, group = .draw, color = .draw ) ) +
   geom_density(show.legend = FALSE) +
@@
 
-ggsave("output/figures/n2oSatDensitySmallRange.tiff", g)
+ggsave("output/figures/n2oSatDensitySmallRange.tiff", g, width = 7.93, height = 4.56)
 
@@
   #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
-  ggplot( aes( x = preds_sat ) ) +
+  ggplot( aes( x = n2osat ) ) +
   geom_density(color = "blue") +
@@
 
+# PLOT N2O SAT RATIO BY SIZE------------------
+# point and linerange
+all_predictions %>%
+  group_by(size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_sat = mean(n2osat)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_sat), 3), 
+    LCL = round(quantile(mean_sat, probs = 0.025), 3),
+    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
+  ggplot( aes( x = size_cat, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  geom_hline(yintercept = 1, color='blue') +
+  ylab("mean N2O saturation ratio") +
+  coord_flip() + 
+  theme_bw() 
+ggsave("output/figures/n2oSatRatioByEco.tiff")
+
+# PLOT N2O* BY SIZE------------------
+# point and linerange
+all_predictions %>%
+  mutate(n2ostar = abs(n2o - n2oeq)) %>%
+  group_by(size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_star = mean(n2ostar)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_star), 3), 
+    LCL = round(quantile(mean_star, probs = 0.025), 3),
+    UCL = round(quantile(mean_star, probs = 0.975), 3)) %>% 
+  ggplot( aes( x = size_cat, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  ylab("absolute[mean N2O - equilibrium N2O] (nmol/L)") +
+  coord_flip() + 
+  theme_bw() 
+ggsave("output/figures/n2oStarBySize.tiff")
+
+# PLOT N2O SAT RATIO BY WSA9 AND SIZE------------------
+
+# point and linerange
+all_predictions %>%
+  group_by(WSA9, size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_sat = mean(n2osat)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_sat), 3), 
+    LCL = round(quantile(mean_sat, probs = 0.025), 3),
+    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  #select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = size_cat, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  geom_hline(yintercept = 1, color='blue') +
+  ylab("mean N2O saturation ratio") +
+  coord_flip() + 
+  theme_bw() +
+  facet_wrap(~ecoregion)
+ggsave("output/figures/n2oSatRatioByEcoAndSize.tiff")
+
 ```
@@
 # Merge wind and all_predictions
-all_predictions.w <- full_join(all_predictions, wind, by = c(".row" = ".id")) # too much for laptop,30 seconds on VM
+tic() # 33 seconds
+all_predictions <- full_join(all_predictions, wind, by = c(".row" = ".id")) # 40 seconds on laptop
+toc()
 dim(wind) # 465897
@@
 465897*500 #232948500
-dim(all_predictions.w) #wind.sampFr. good
 
@@
 
-all_predictions.w <- all_predictions.w %>%
+all_predictions <- all_predictions %>%
   mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
          # 1000 L to m3. 44ng to 1nmol. ng->ug. ug->mg
-         e.n2o.mg.m2.d = (((preds_n2o - preds_n2oeq) * 1000 * 44) / (1000 * 1000)) * 
+         e.n2o.mg.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000 * 1000)) * 
            (k600.cm.h * (24/100)), #h->day. cm->m
@@
 # write reduced file to disk for less computationally intensive wrangling
-#all_predictions.w.small <- all_predictions.w %>% filter(.draw %in% 1:5)
-
-#save(all_predictions.w.small,
-#     file = paste0(Sys.getenv("USERPROFILE"), 
-#                   "/Environmental Protection Agency (EPA)/",
-#                   "ORD NLA17 Dissolved Gas - Documents/",
-#                   "inputData/all_predictions_w_small.rda"))
+all_predictions.w.small <- all_predictions.w %>% filter(.draw %in% 1:5)
+
+save(all_predictions.w.small,
+     file = paste0(Sys.getenv("USERPROFILE"),
+                   "/Environmental Protection Agency (EPA)/",
+                   "ORD NLA17 Dissolved Gas - Documents/",
+                   "inputData/all_predictions_w_small.rda"))
 ```
@@
 # National scale number for lakess/reservoirs (this is is Gg)
-national.lakes <- all_predictions.w %>% 
+national.lakes <- all_predictions %>% 
   group_by(.draw) %>% 
@@
 # Merge lakes and inventory category
-inventory <- full_join(n2o, national.lakes)
+inventory <- full_join(n2o, national.lakes) %>%
+  arrange(desc(Gg))
 inventory
@@
   theme(axis.title.y = element_blank())
-ggsave(filename = "output/figures/n2oInventory.tiff")
-
-tiff("output/figures/n2oInventoryPie.tiff")
+ggsave(filename = "output/figures/n2oInventory.tiff", width = 9, height = 5.5)
+
+tiff("output/figures/n2oInventoryPie.tiff", 
+     type = "cairo", bg = "transparent") # for transparent background
 pie.labels <- c("Agricultural Soil Management",
@@
 
-(national.lakes$Gg/sum(inventory$Gg)) * 1000 #1.6%
+(national.lakes$Gg/sum(inventory$Gg)) * 1000 #0.7%
 ```
@@
 # density plot with RUG for one draw
-all_predictions.w %>% filter(.draw == 1) %>% ggplot(aes(x=f.n2o.Mg.y)) + geom_density() +geom_rug() + facet_wrap(~WSA9, scales = "free")
+all_predictions %>% filter(.draw == 1) %>% ggplot(aes(x=f.n2o.Mg.y)) + geom_density() +geom_rug() + facet_wrap(~WSA9, scales = "free")
 
@@
 # bar plot by size category with error
-p <- all_predictions.w %>% 
+p <- all_predictions %>% 
   group_by(WSA9, .draw, size_cat2) %>% 
@@
 # this is analogous to spsurvey "population estimates"
-ecoreg_stats <- all_predictions.w %>%
+ecoreg_stats <- all_predictions %>%
   group_by(WSA9, .draw) %>%
-  summarise( mean_n2o = mean(preds_n2o),
-             mean_n2oeq = mean(preds_n2oeq),
-             mean_sat = mean(preds_sat),
-             prop_sat = sum(preds_sat < 1) / length(unique(.row)),
+  summarise( mean_n2o = mean(n2o),
+             mean_n2oeq = mean(n2oeq),
+             mean_sat = mean(n2osat),
+             prop_sat = sum(n2osat < 1) / length(unique(.row)),
              mean_en2o = mean(e.n2o.mg.m2.d),
@@
                            graph.bgcolor='lightgray',
-                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
-                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
+                           xaxis.ticks=list(.65, 0.7, 0.75),
+                           xaxis.labels=list(.65, 0.7, 0.75),
                            xaxis.title='Proportion undersaturated', panel.width=2.3),

### 2023-02-08 scripts/dgIndicatorAnalysis.Rmd

@@
 library(tictoc)
+library(spmodel) # esv function
 

### 2023-02-08 scripts/n2o_Multilevel_Analysis.Rmd

---
title: "Modeling NLA 2017 $N_2O$ Data"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook: 
    depth: 4
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
bibliography: RWM_Endnote_Library.bib
link-citations: yes
---

<style>
.vscroll-plot {
    width: 800px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

# Background
In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).

## Data description and structure
As explained in the related document,
(https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html), duplicate gas samples were collected at a depth of ~0.1$\text{m}$ at designated index sites distributed across 1091 waterbodies nationwide, of which 95 were sampled twice. Gas samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L.

The data were collected under a stratified, unequal probability design. The observations can be thought of as hierarchically organized, such that the $N=1185$ samples ($N = \sum1,...,n$) were situated within $I=1091$ index waterbodies ($I=\sum 1,...,i$), which were situated, with unequal probability, in one of $L=5$ size categories ($L=\sum 1,...,l$; < 4$\text{ha}$, 4-10$\text{ha}$, 10-20$\text{ha}$, 20-50$\text{ha}$, and >50$\text{ha}$), which were within one of $K=48$ states ($K=\sum 1,...,k$) in the interior U.S., within one of $J=9$ aggregated, Omernik ecoregions ($J=\sum 1,...,j$; WSA9 = Xeric (XER), Western Mountain (WMT),  Northern Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and Southern Appalachian (SAP) regions).

## Goals for modeling
Our goals for this particular excercise include model-based inference via hierarchical modeling [e.g., @Park_etal_2004; @Ghitza_Gelman_2013] with regard to estimation of $N_2O$ concentrations from probabilistic survey. Our approach will consider the sampling design by including as predictors in our models the design factors that affected probability of inclusion [@Rubin_1976; @Rubin_1983; @Gelman_2007], which included size category, state, and ecoregion. This follows the concept that the inclusion mechanism, $I$, may be considered ignorable if its distribution, conditional on $y$, is independent of $y$, conditional on the weighting variables. All inferences will be conducted in a fully Bayesian fashion.

# Setup R
```{r setup_1, include=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")

#set directories and R package library
setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
.libPaths("C:/Users/rmartin/Desktop/R/library")
```

```{r load_packages, warning=FALSE, message=FALSE}
library(brms)
library(bayesplot)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(gridExtra)

options(mc.cores = parallel::detectCores())
options( max.print = 2000 )
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}

#Calc geometric mean
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

gm_sd = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(sd(log(x), na.rm = na.rm))
  } else {
    exp(sd(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

#Center and scale by (1) sd
MyStd <- function(x){ (x-mean(x))/(sd(x)*2)}

#First letter in string to uppercase
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x}
```

# Import and organize data

Lets now read in the .RData file containing the NLA data, previously munged here: 
https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html

```{r import_data, echo=TRUE}
load( file = paste0( localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/dg.2019-12-19.RData")
      )
```

Then, we'll create a new dataframe, including only variables we are interestd in for this exercise.

```{r mod1_data_import, paged.print=TRUE}
dg %>%
  filter( !is.na( dissolved.n2o ) ) %>%
  nrow() #number of observations in original n2o data

model_dat <- dg %>%
  filter( !is.na( dissolved.n2o ) ) %>%
  mutate( WSA9 = factor( WSA9 ),
          state = factor( state ),
          area = AREA_HA
          ) %>%
  select( WSA9,
          state,
          area,
          size_cat,
          dissolved.n2o
          )

nrow( model_dat ) #number of observations in model data after filtering NA for various covariates

#print top 50 rows
top_n( model_dat, n = 20 )
```

# A general framework for model specification

As we consider our own model, we'll follow the lead of others [e.g., @Webb_etal_2019] and consider a mixed model and a gamma or lognormal likelihood. In future iterations, we'll potentially consider some linear and/or smooth covariate functions as well [@Webb_etal_2019]. The general hierarchical structure will be samples (level-1) from lakes of particular size classes (level-2), within states (level-3), within ecoregions (level-4).

Specifically, considering the Gamma likelihood, the model may have the general form:

$$y \sim Gamma( \mu, \phi)$$
where $y$ represents the observed samples, which are assumed to be Gamma distributed according location parameter $\mu$ and shape parameter $\phi$.

The location parameter is further parameterized, such that:

$$\mu_{i,j,k,l} = \alpha + f_d(X_{i,d}) + \gamma_{i,j} + \delta_{i,j,k} + \epsilon_{i,j,k,l}$$
In this explicit notation, each sample, $i$, within lake size class, $l$, in state, $k$, within each ecoregion, $j$, is indexed to the location parameter, $\mu_{i,j,k,l}$, and mapped to the linear predictor on the right-hand side, where $\alpha$ indexes the global intercept parameter, $f_d$ indicates the smooth or regular linear functions determining sample-level (i.e., level-1) covariate effects of $d=1,...,D$ independent variables, $x_{i,d}$; and $\gamma_{i,j}$, $\delta_{i,j,k}$, and $\epsilon_{i,j,k,l}$ are the latent, random effects coding for ecoregion, state, and size class, respectively.

# Exploration of data structure
Before we get into our specific model parameterizations, lets explore the structure of the data a bit. First, we'll look at the overall distribution of the national dataset.

```{r summary_n2o}
summary( model_dat$dissolved.n2o )
```

From this basica summary, we can see that the data has a strong right-skew. The mean is relatively small compared to the max. However, these numbers are somewhat inline with those observed by [@Webb_etal_2019] over 101 constructed agricultural reservoirs across Saskatewan, Canada. They reported a mean of 6.55 and a range [1.14 - 110].

Lets now take a look at the data graphically.

```{r violin, fig.align='center', fig.height= 6, fig.width= 3}
ggplot( model_dat, aes( x = NA, y = dissolved.n2o ) ) +
  geom_violin( fill = "grey",
               color = "black" ) +
  geom_boxplot( width = 0.1 ) +
  theme_bw()
```

This violin plot captures the skew pretty well. There are quite a few extreme values, though the bulk of the data seems to be in the < 15 nmol/L range.

Lets look now at the same data after a log (base $e$) transformation.

```{r violin_2, fig.align='center', fig.height= 6, fig.width= 3}
ggplot( model_dat, aes( x = NA, y = log( dissolved.n2o ) ) ) +
  geom_violin( fill = "grey",
               color = "black" ) +
  geom_boxplot( width = 0.1 ) +
  theme_bw()
```
We see that the overall mean is just under 2 on the log scale and there remains a good deal of left- and right-skew.

Lets look now at some of the hierarchical structure.

## WSA9 ecoregion

We start with examining the structure by ecoregion. Here we see again that, even on the log scale, there is a good deal of right-skew in distribution of values observed within the ecoregions. The boxplots here (within the violins) suggest that the means may be slightly different across the ecoregions, if considering the data were generated from a lognormal likelihod, but the skew should probably caution against this interpretation.

```{r wsa9_violin, fig.align='center', fig.height= 6, fig.width= 8}
ggplot( model_dat, aes( x = WSA9, y = log( dissolved.n2o ) ) ) +
  geom_violin( fill = "grey",
               color = "black" ) +
  geom_boxplot( width = 0.1 ) +
  theme_bw()
```

## States in WSA9 ecoregions

Lets look now at the distribution within states (further divided within ecoregions ). Note: This is a long plot that you can scroll through on the right side.

<div class="vscroll-plot">
```{r violin_states, fig.align='center', fig.height= 80, fig.width= 8}
ggplot( model_dat, aes( x = WSA9, y = log( dissolved.n2o ) ) ) +
  facet_wrap( ~ state, ncol= 1 ) +
  geom_violin( fill = "grey",
               color = "black" ) +
  geom_boxplot( width = 0.1 ) +
  theme_bw()
```
</div>

There's not a lot to gather from this long figure, except that there is still a good bit of right-skew within most groups.

## Sizes classes in states

Lets look now at the distribution within size clases (within states).

<div class="vscroll-plot">
```{r violin_sizes, fig.align='center', fig.height= 40, fig.width= 8}
ggplot( model_dat, aes( x = size_cat, y = log( dissolved.n2o ) ) ) +
  facet_wrap( ~ state, ncol= 2 ) +
  geom_boxplot( width = 0.4 ) +
  theme_bw()
```
</div>

Here, aside from several states (e.g., IN, WI), we _maybe_ see a little less skew when broken down into size classes within groups. However, the sample sizes are also pretty small when broken down to this level.

# Model 1
For our first model, we'll fit a GLMM with with a Gamma likelihood. We won't include any covariates. We are looking here to see whether we can model the data based only on the sampling design and likelihood. Specifically, we will eventually want to compare the estimates from these models without covariates to those made using the traditional survey method, as a sort of reality check.

Because the model and data are fairly straightforward, we should be able to efficiently conduct the analysis using the $\textbf{brms}$ package [@Burkner_2017], which affords fully Bayesian inference via its interface with $\textbf{Stan}$ [@Carpenter_etal_2017], which is a $\textbf{C++}$ code package for obtaining full Bayesian inference [@Carpenter_etal_2015].

Below, we fit our first model of the form:

$$y \sim Gamma( \mu, \phi)$$
where $y$ represents the observed samples, which are assumed to be Gamma distributed according location parameter $\mu$ and shape parameter $\phi$.

The location parameter is further parameterized, such that:

$$\mu_{i,j,k,l} = \alpha + \gamma_{i,j} + \delta_{i,j,k} + \epsilon_{i,j,k,l}$$
In this explicit notation, each sample, $i$, within lake size class, $l$, in state, $k$, within each ecoregion, $j$, is indexed to the location parameter, $\mu_{i,j,k,l}$, and mapped to the linear predictor on the right-hand side, where $\alpha$ indexes the global intercept parameter, and $$\gamma_{i,j}$$, $\delta_{i,j,k}$, and $\epsilon_{i,j,k,l}$ are the latent, random effects coding for ecoregion, state, and size class, respectively. The random effects are specified as normally distributed about 0, with scale parameter $\tau$ learned from the data, such that $$\gamma_{i,j} \sim N( 0, \tau_{\gamma} )$$ 
$$\delta_{i,j,k} \sim N( 0, \tau_{\delta} )$$ 
$$\epsilon_{i,j,k,l} \sim N( 0, \tau_{\epsilon} )$$

## Fitting

```{r mod1}
mod1 <- brm( dissolved.n2o ~ 
               1 + ( 1 | WSA9 ) + ( 1 | WSA9:state ) + ( 1 | WSA9:state:size_cat ),
             data = model_dat, 
             family = Gamma( link = 'log' ),
             prior = c( prior( normal( 2, 2 ), class = "Intercept" ),
                        prior( gamma( 0.01, 0.01 ), class = "shape" ),
                        prior( normal( 0, 1 ), class = 'sd' )
                        ),
             control = list( adapt_delta = 0.99, max_treedepth = 14 ),
             chains=4, 
             iter=2000, 
             cores=4 )
```

## Fit summary
Here, we call a summary of the fitted parameters.

```{r print_mod1}
print( mod1, prior = T )
```

We see that the intercept parameter is estimated to be about 2.1, which is coincidentally the same as estimated in [@Webb_etal_2019] ( 2.11 +/- 0.05). We also see that the variability between size classes (within states and ecoregions; $\tau_{\epsilon} = 0.29$) is estimated close to the estimate for variability between states (within ecoregions; $\tau_{\delta} = 0.28$). The variability between ecoregions is esimated to be quite small by comparison (i.e., $\tau_{\gamma} = 0.07$.

Before we think too much about this summary, however, we'll do some checks to see how well the fitted model does at describing the data.

## Posterior predictive checks

First, we'll extract the posterior samples and assign the empirical data to an object
```{r ppc_check_extract_mod1}
mod1_post <- posterior_samples( mod1 )
yrep_1 <- posterior_predict( mod1 )
y <- model_dat$dissolved.n2o
```

Next, we'll look at density overalap check. Here, we're looking to see if the parameters fitted by the model can generate replicate datasets that look something like the empirical data.

```{r ppc_dens_overlay_mod1}
ppc_dens_overlay( y = log( y ), yrep = log( yrep_1[ sample( seq(1 , 1000, 1 ), 50 ), ] ) )
```

Clearly, the fitted model is generating replicate datasets (blue lines) where the bulk of the data are too broadly distributed relative to the observed data (black line). There appears to be a clear misfit to the data with this model, but we'll still look at a couple of other checks.

Next, we'll look at the scatter of the log of the average of the posterior predictions vs measured; that is, the observed, $log( y_i)$ _vs_ the average of the replicate datasets at each $i$ or $log(y_{rep_i})$. Here, we see a diagonal cloud with a 1:1 slope near the mean, with increasing error as you move into the more extreme values.

```{r ppc_error_scatter_avg_mod1}
ppc_scatter_avg( y = log( y ), yrep = log( yrep_1[ sample( seq(1 , 1000, 1 ), 12 ), ] ) )
```

Next, lets see how the model does with a simple stat like the median. Here, we are checking whether replicate datasets from the fitted model generate median estimates (blue histogram) similar to the observed median (black vertical line).

```{r ppc_stat_median_mod1}
ppc_stat( y = y, yrep = yrep_1, stat = median, binwidth = 0.05 )
```

It looks like the model consistently generates datasets with medians higher than that calculated for the observed data. This is suspect, but at least in the ball park (barely). It is also to be expected after seeing the replicated densities in the figure above, where clearly the bulk of the densities from the replicated datasets are right of the observed.

Lets now do this same check, but by ecoregion groupings.

```{r ppc_stat_grouped_median_mod1}
ppc_stat_grouped( y = y, yrep = yrep_1, group = model_dat$WSA9, stat = median, binwidth = 0.05 )
```

Here, we see that the model doesn't do a great job, particularly with the CPL or TPL groups when it comes to replicating the empirical medians. Aside form the NAP, NPL, and WMT, the medians are also consistently right of the observed.

Lets look at another aspect of the data: skew.

```{r ppc_stat_skew_mod1}
ppc_stat( y = y, 
          yrep = yrep_1, 
          stat = function(x) 3 * ( mean( x ) - median( x ) ) / sd( x ), 
          binwidth = 0.01 )
```

Here we see that the model consistently generates replicate datasets with more right-skew than the observed data.

Lets look by ecoregion next.

```{r ppc_stat_grouped_skew_mod1}
ppc_stat_grouped( y = y, 
                  yrep = yrep_1, 
                  stat = function(x) 3 * ( mean( x ) - median( x ) ) / sd( x ), 
                  group = model_dat$WSA9,
                  binwidth = 0.05 )
```

Here, it looks like the UMW skew is clearly not well replicated, but the others aren't as far off.

## Leave-one-out evaluations
Though it is obvious at this point that our model doesn't generate replicate datasets with features that resemble important aspects of the observed data, indicating that it isn't a very good approximation of the true data generating process, we'll continue with the checking process to get a better understanding of some of these methods before we move forward with different model parmeterizations.

Here we'll conduct a Bayesian leave-one-out cross validation ('LOO-CV') assessment using the $\textbf{loo}$ package in $\textbf{R}$, which implements fast and stable computations for approximate LOO-CV (and WAIC) [@Vehtari_etal_2017]. These computation methods estimate pointwise out-of-sample prediction accuracy from a fitted model by using the log-likelihood evaluated at the posterior simulations of the parameter values. Conveniently, the $\textbf{brms}$ package automatically calculates the log-likelihood for us for all relevant models, making it very convenient to use in conjunction with the $\textbf{loo}$ package. For more information on $\textbf{loo}$ and LOO-CV in general, check out the page on the $\textbf{Stan}$ website: http://mc-stan.org/loo/. For further explanation through worked examples, see: https://avehtari.github.io/modelselection/slides_model_assesment_selection.pdf.

```{r mod1_loo}
loo_1 <- loo( mod1, save_psis = TRUE, cores = 4 )
psis_1 <- loo_1$psis_object
lw_1 <- weights( psis_1 )

print( loo_1 )
```

Here we have a diagnostic summary of the pareto $k$ values calculated via the `loo()` call for our model and data. Pareto-smoothed importance sampling is the method used in $\textbf{loo}$ for approximating the true leave-one-out posterior. In essesnce, we fit the model once, save the log-likelihood for each data point, and `loo()` utilizes that to re-weight and approximate the leave-one-out posterior. This is convenient because it allows us to check the model using leave-one-out methods without fitting the model $N$ times. However, the approximation can fail, and the $\textbf{loo}$ package provides some diagnostics that indicate when failure is more likely. In particular, each data point is assigned an importance ratio, $k$, which indicates how "influential" it may be. As such, too many data points with high (pareto-k > 0.7) or very high (pareto-k > 1) influence may indicate problems with the approximation. Such influential data points could have infinite variance, making smoothing problematic. When there are too many such points, it is generally recommended to either re-fit the model to each of the problematic datapoints, independently (via `reloo()` explained later), or with a more traditional leave-k-out method, such as k-fold-CV.

The LOO-CV evaluation for this model found a large number of "influential" or "problematic" observations (i.e., "bad" = 20 + "very bad" = 15). Note that observations with high $k$ are only problematic in the sense of trying to approximate the leave-one-out posterior, and that there isn't necessarily anyting inherently wrong with them. Although, with this many problematic observations, it _could_ be an indication of mispecification (and likely is, as we saw from our previous checking excercises). At this point we might have wanted to try a 'reloo' or 'k-fold' cross-validation, but we are already pretty sure the model is a poor fit to important aspects of the data, or potentially mis-specified, so we're not going to waste any more time with cross-validation on this particular parameterization and fit.

With regard to the other outputs from the `summary(loo)` call, the "elpd_loo" is a way to compare models. For each observation, a calculation of how "surprised" the model is to see the left-out datapoint is made based on the density at the observed value. The elpd_loo is just a sum of these individual contributions (https://mc-stan.org/loo/reference/loo-glossary.html). The p_loo, by comparision, is the difference between the elpd_loo and the non-CV log posterior predictive density, and may be interpreted as the effective number of parameters. The "looic", finally, is just $-2^{elpd_{loo}}$ to provide output on the conventional scale of "deviance" or AIC.

One interesting aspect of 'p_loo' is that, when $k$ is large, as in this particular case, we can look to the p_loo estimate for additional information about the problem (https://mc-stan.org/loo/reference/loo-glossary.html). More specifically, it may be helpful to compare p_loo to the actual number of model parameters to get some sense of potential for misspecification. In our case p_loo was calculated to be 255.6, and the total number of parameters in our model, including latent parameters, was about 450. In this case, p_loo = 255.6 is considerably less than p = 450 and the model may be misspecified (e.g., likelihood) or need more structural information (e.g., covariates).

# Model 2
Since our first model didn't seem to work out well at all, we'll try a second model. This time we'll fit a similar GLMM, but this time with with a lognormal likelihood. The lognormal probably won't produce wildly different results, but we'll give it a try to see if there are any interesting differences. Again, we won't include any covariates.

Below, we fit our second model of the form:

$$y \sim lognormal( \mu, \sigma)$$
where $y$ represents the observed samples, which are assumed to be lognormally distributed according location parameter $\mu$ and scale parameter $\sigma$.

The location parameter is again  parameterized, such that:

$$\mu_{i,j,k,l} = \alpha + \gamma_{i,j} + \delta_{i,j,k} + \epsilon_{i,j,k,l}$$

with the same parameterization for the random effects as before.

## Fitting

```{r mod2}
mod2 <- brm( dissolved.n2o ~ 
               1 + ( 1 | WSA9 ) + ( 1 | WSA9:state ) + ( 1 | WSA9:state:size_cat ),
             data = model_dat, 
             family = lognormal(),
             prior = c( prior( normal( 2, 2 ), class = "Intercept" ),
                        prior( normal( 0, 1 ), class = "sigma" ),
                        prior( normal( 0, 1 ), class = 'sd' )
                        ),
             control = list( adapt_delta = 0.99, max_treedepth = 12 ),
             chains=4, 
             iter=2000, 
             cores=4 )
```

## Fit summary
Here, we call a summary of the fitted parameters.

```{r print_mod2}
print( mod2, prior = T )
```

We see that the intercept parameter is estimated to be about 2.03. In contrast with our first model, we see that the variability between size classes (within states and ecoregions) is estimated to be a good deal smaller than variability between states (within ecoregions). The variability between ecoregions is, again, esimated to be very small by comparison. In this lognormal model, we also have an estimate of within size-class (within states and ecoregions) variability (also residual error), which is estimated to be larger than variability between classes within any of the levels above.

Again, we'll do some checks to see how well the fitted model does at describing the data.

## Posterior predictive checks

First, we'll extract the posterior samples and assign the empirical data to an object
```{r ppc_check_extract_mod2}
mod2_post <- posterior_samples( mod2 )
yrep_2 <- posterior_predict( mod2 )
```

Next, we'll look at a density overaly.

```{r ppc_dens_overlay_mod2}
ppc_dens_overlay( y = log( y ), yrep = log( yrep_2[ sample( seq(1 , 1000, 1 ), 100 ), ] ) )
```

Again, the fitted model is generating replicate datasets (blue lines) where the bulk of the data are too broadly distributed relative to the observed data (black line). The extremes in the right tail are particularly well replicated either. So, again, there appears to be a clear misfit to the data with this model, but we'll still look at a couple of other checks.

Next, we'll look at the scatter of the (log) average predictions _vs._ log(observed).

```{r ppc_error_scatter_avg_mod2}
ppc_scatter_avg( y = log( y ), yrep = log( yrep_2[ sample( seq(1 , 1000, 1 ), 12 ), ] ) )
```

Next, lets see how the model does with the median.

```{r ppc_stat_median_mod2}
ppc_stat( y = y, yrep = yrep_2, stat = median, binwidth = 0.02 )
```

Again, it looks like the model consistently generates datasets with medians higher than that calculated for the observed data.

Lets now do this check, but by ecoregion groupings.

```{r ppc_stat_grouped_median_mod2}
ppc_stat_grouped( y = y, yrep = yrep_2, group = model_dat$WSA9, stat = median, binwidth = 0.05 )
```

Here, we see that the model maybe doesn't do a great job; with the CPL or TPL groups, especially. Overall, the median again seem to be consistently overestimated across groups, excepting maybe two or three (e.g., WMT, NPL).

Lets look at the skew.

```{r ppc_stat_skew_mod2}
ppc_stat( y = y, 
          yrep = yrep_2, 
          stat = function(x) 3 * ( mean( x ) - median( x ) ) / sd( x ), 
          binwidth = 0.01 )
```

Here we see that the model consistently generates replicate datasets with more right-skew than the observed data.

Lets look by ecoregion next.

```{r ppc_stat_grouped_skew_mod2}
ppc_stat_grouped( y = y, 
                  yrep = yrep_2, 
                  stat = function(x) 3 * ( mean( x ) - median( x ) ) / sd( x ), 
                  group = model_dat$WSA9,
                  binwidth = 0.05 )
```

Here, it looks like the skew for some groups is maybe not well replicated, but we're at least somewhere in the ballpark, by comparison to the first model.

## Leave-one-out evaluations
We'll continue with the LOO-CV assessment.

```{r mod2_loo}
loo_2 <- loo( mod2, save_psis = TRUE, cores = 4 )
psis_2 <- loo_2$psis_object
lw_2 <- weights( psis_2 )

print( loo_2 )
```

Compared to our first model, the LOO-CV evaluation for this model resulted in fewer "problematic" observations (i.e., "bad/very bad" = 7 ). Since this isn't a very large number, we'll take the advice noted in the warning, and use the `reloo()` function, which calculates the "true" leave-one-out posterior for the problematic points only, and then uses the approximation for the remaining observations.

```{r mod2_reloo}
reloo_2 <- reloo( mod2, loo_2 )#loo( mod2, reloo = TRUE, reloo_args = list( save_psis = TRUE), cores = 4 )
print( reloo_2 )
```

The `reloo()` result suggests that the issue with potentially problematic observations can be cautiously ignored; and we could proceed forward with this "reloo" result for model checking or model comparisons. However, our other posterior predictive checks, again, suggested a poor fit, so there isn't much gained in moving forward with this model for now. We will utilize the 'loo' output to do one more check below though, just to further illustrate the potential utility.

To examine the calibration of predictions, we can look at the LOO-CV predictive cumulative density function values, which are asymptotically uniform if the model is well calibrated. The below plot compares the density of the computed LOO probability integral transforms (dark line) _vs._ 100 simulated datasets from the standard uniform distribution (blue lines).

```{r mod2_ppc_loo_pit_overlay}
ppc_loo_pit_overlay( y = y,
                     yrep = yrep_2,
                     lw = weights( reloo_2$psis_object ),
                     samples = 100 )
```

From this check, it is fairly clear that there is some miscalibration, in that the hump/frown-shape indicates that the univariate predictive distributions are too broad compared to the observed data (see also earlier PPCs), suggesting that further modeling (e.g., including additional structure/covariates) will be needed to more accurately reflect uncertainty [@Gabry_etal_2019].

## Check _vs._ survey estimates 
Although our model appears to be pretty poor again, lets go ahead and check some estimates against those provided by the traditional survey method, which have already been applied and discussed in a previous document: https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html. The estimates we'll compare are the mean estimates (and uncertainty) for dissolved $N_2O$ by ecoregion and nationally.

First, we'll manually enter the survey estimates (means, lower and upper 95% CI by ecoregion) into an object.

```{r survey_ests}
#make a dataframe object

survey_ests <- data.frame( 
  ecoregion = c( "CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER", "US" ), 
  estimate = c( 8.4, 7.7, 6.9, 7.1, 6.5, 7.8, 10.8, 7.8, 10.6, 8.1 ),
  LCL = c( 5.1, 7.2, 6.4, 5.9, 4.9, 5.9, 6.4, 7.1, 7.5, 7.0 ),
  UCL = c( 11.7, 8.1, 7.4, 8.2, 8.1, 9.6, 15.2, 8.4, 13.7, 9.1 )
) %>%
  mutate( ecoregion = factor( ecoregion, 
                              levels = c( "CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER", "US" ) ) ) #relevel for ggplot so not alphabetical, but original order

print( survey_ests )
```

Lets make a plot as well, if only as a check on our data entry.

```{r survey_ests_plot, fig.align = "center"}
ggplot( survey_ests, aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL ) ) +
  coord_flip()
```

This plot for the survey estimates looks about right, so we'll move to constructing the estimates for the means from our model.

With our hierarchical model specification, we should be able to get the ecoregion mean estimates by utilizing the intercept parameter estimate, $\alpha$, along with the latent hierarchical effects estimates for the ecoregion intercepts, $\gamma_{i, j}$. However, because we have a lognormal likelihood, we'l have to do some transformations. If we add the intercept estimate to the random effects estimates, we get an estimate for the log-scale means. If we exponentiate that, we get an estimate for the geometric means on the data scale. To get the mean estimates on the data scale, we have to apply a transformation, such that: $$E( \theta ) = exp( \mu + 0.5 \sigma^2)$$

So, lets get our model estimates for the means on the data scale. We'll utilize the full posterior extracted above, and create an object for each ecoregion ( and US ) containing the posterior distributions for the means.

```{r mod2_ests}
CPL <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[CPL,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
NAP <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[NAP,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
NPL <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[NPL,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
SAP <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[SAP,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
SPL <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[SPL,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
TPL <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[TPL,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
UMW <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[UMW,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
WMT <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[WMT,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
XER <- exp( mod2_post$b_Intercept + mod2_post$`r_WSA9[XER,Intercept]` + ( 0.5 * mod2_post$sigma ^ 2 ) )
US <- exp( mod2_post$b_Intercept + ( 0.5 * mod2_post$sigma ^ 2 ) )

mod2_ests <- data.frame( 
  ecoregion = c( "CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER", "US" ),
  estimate = c( mean( CPL ), mean( NAP ), mean( NPL ), mean( SAP ), mean( SPL ), 
                mean( TPL ), mean( UMW ), mean( WMT ), mean( XER ), mean( US ) ),
  LCL = c( quantile( CPL , 0.025 ), quantile( NAP , 0.025 ), quantile( NPL , 0.025 ), 
           quantile( SAP , 0.025 ), quantile( SPL , 0.025 ), quantile( TPL , 0.025 ), 
           quantile( UMW , 0.025 ), quantile( WMT , 0.025 ), quantile( XER , 0.025 ), 
           quantile( US , 0.025 ) ),
  UCL = c( quantile( CPL , 0.975 ), quantile( NAP , 0.975 ), quantile( NPL , 0.975 ), 
           quantile( SAP , 0.975 ), quantile( SPL , 0.975 ), quantile( TPL , 0.975 ), 
           quantile( UMW , 0.975 ), quantile( WMT , 0.975 ), quantile( XER , 0.975 ), 
           quantile( US , 0.975 ) )
  ) %>%
  mutate( estimate = round( estimate, 1 ), # round to nearest tenth
          LCL = round( LCL, 1 ),
          UCL = round( UCL, 1 )
  )  %>%
  mutate( ecoregion = factor( ecoregion, #relevel for ggplot so not alphabetical, but original order
                              levels = c( "CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER", "US" ) ) )

rm( CPL, NAP, NPL, SAP, SPL, TPL, UMW, WMT, XER, US ) #clean up workspace
print( mod2_ests )
```

Lets combine our estimates into one dataframe now.

```{r combine_ests_mod2}
mod2_ests <- data.frame( rbind( survey_ests, mod2_ests ) ) %>%
  mutate( model = factor( c( rep( "survey", 10 ), rep( "mod2", 10 ) ) ) )

print( mod2_ests )
```

Now lets plot these estimates alongside the ones from the survey.

```{r survey_mod2_ests_plot, fig.align = "center"}
dodge <- position_dodge(width=0.5) # so lines and points don't overlap on graph

ggplot( mod2_ests, aes( x = ecoregion, y = estimate, group = model ) ) +
  geom_point( aes(colour = model ) , position=dodge ) +
  geom_linerange( aes( ymin = LCL, ymax = UCL , color = model ) , position=dodge ) +
  scale_colour_manual(values = c("black", "darkgray") ) +
  coord_flip() +
  theme_bw()
```

Though our model doesn't look like a good fit at all, the means estimates maybe aren't _terribly_ far off. In all cases, the uncertainty intervals at least overlap with the 95% confidence intervals on the survey estimates. We can also see that, for the most part, the hierarchical model estimates have much greater certainty, as expected due to partial pooling of information across groups. Likewise, as the estimate for between-ecoregion variance indicated for this model, these estimates don't suggest much variability in the means from one ecoregion to the next.

Lets move on and see if we can get a better fitting model now.

# Model 3
It is maybe a bit surprising that neither of the relatively straightforward Gamma or lognormal models seemed to provide a good fit. Here we'll try a third model that should provide more flexibility. Specifically, we'll fit another lognormal GLMM, but this time we're going to incorporate a model for the residual variance. The model will again be hierarchical, such that we model it as a function of the size_class within states stucture.

Below, we fit our third model of the form:

$$y \sim log-N( \mu, \sigma)$$
The location parameter is again  parameterized, such that:

$$\mu_{i,j,k,l} = \alpha + \gamma_{i,j} + \delta_{i,j,k} + \epsilon_{i,j,k,l}$$

This time, however, we have a model for $\sigma$, such that:

$$\sigma \sim log-N( \mu_{sigma}, \nu_{sigma})$$

where $\mu_{sigma}$ is an intercept parameter capturing the (log-scale) mean residual variation and $\nu_{sigma}$ is the scale parameter describing variation in $\sigma$ by size-class, which is estimated from the data.

## Fitting

```{r mod3}
mod3 <- brm( bf( dissolved.n2o ~ 
               1 + ( 1 | WSA9 ) + ( 1 | WSA9:state ) + ( 1 | WSA9:state:size_cat ), 
               sigma ~ ( 1 | WSA9:state:size_cat ) ),
             data = model_dat, 
             family = lognormal(),
             prior = c( prior( normal( 2, 2 ), class = "Intercept" ),
                        prior( normal( 0, 1 ), class = "Intercept", dpar = "sigma" ),
                        prior( normal( -0.5, 1 ), class = "sd", dpar = "sigma" ),
                        prior( normal( 0, 1 ), class = "sd" )
                       ),
             control = list( adapt_delta = 0.90, max_treedepth = 12 ),
             #sample_prior = "only",
             chains=4, 
             iter=2000, 
             cores=4 )
```

## Fit summary
Here, we call a summary of the fitted parameters.

```{r print_mod3}
print( mod3, prior = T )
```

We see that the intercept parameter is estimated to be about 1.95. In contrast with our last model, we see that the variability between size classes (within states and ecoregions) is estimated to be a great deal larger than variability between states (within ecoregions), which is estimated to be simlar to variability within ecoregions. In this model, our residual error estimate is a little trickier because we've actually allowed it to vary, in a hierarchical fashion (i.e., partially pooled ) among size-categories within states and ecoregions. The intercept for this parameter, "sigma_Intercept", indicates that the (geometric) mean of the residual variation was around $exp( -1.41) = 0.24 ), but that this varied considerably among groups (i.e., "sd(sigma_Intercept)" = 0.82). To put into perspective, a rough calculation of the central 95% interval on a lognormal variate with location = -1.4 and scale = 0.82 results in a range of standard deviations from 0.05 to 1.2. This is a pretty wide range and seems like a pretty nutty parameterization that would be prone to overfitting, but we'll go ahead and continue to analyze this model below.

## Posterior predictive checks

First, we'll extract the posterior samples and assign the empirical data to an object
```{r ppc_check_extract_mod3}
mod3_post <- posterior_samples( mod3 )
yrep_3 <- posterior_predict( mod3 )
```

Next, we'll look at a density overaly. Here, we're looking to see if the parameters fitted by the model can generate replicate datasets that look something like the empirical data.

```{r ppc_dens_overlay_mod3}
ppc_dens_overlay( y = log( y ), yrep = log( yrep_3[ sample( seq(1 , 1000, 1 ), 100 ), ] ) ) 
```

This is actually the closest we've gotten so far with this check. It looks like the replicates are still a little too broadly distributed relative to the data, but this a far sight better than the other two models.

Next, we'll look at the scatter of the (log) average predictive errors.

```{r ppc_scatter_avg_mod3}
ppc_scatter_avg( y = log( y ), yrep = log( yrep_3[ sample( seq(1 , 1000, 1 ), 12 ), ] ) )
```

Next, lets see how the model does with the median.

```{r ppc_stat_median_mod3}
ppc_stat( y = y, yrep = yrep_3, stat = median, binwidth = 0.02 )
```

It looks like the model consistently generates datasets with medians similar to the observed data. Again, this is looking much better.

Lets now do this check, but by ecoregion groupings.

```{r ppc_stat_grouped_median_mod3}
ppc_stat_grouped( y = y,
                  yrep = yrep_3, 
                  group = model_dat$WSA9, 
                  stat = median, 
                  binwidth = 0.1)
```

Again, looking much better this time. All of the observed medians appear to be pretty well replicated for each group.

Lets look at the skew.

```{r ppc_stat_skew_mod3}
ppc_stat( y = log( y ), 
          yrep = yrep_3, 
          stat = function(x) 3 * ( mean( x ) - median( x ) ) / sd( x ), 
          binwidth = 0.01 )
```

This is a fairly strange looking plot with the abrupt shoulder on the lower end, but it does seem like we're capturing the skew pretty well overall.

Lets look by ecoregion next.

```{r ppc_stat_grouped_skew_mod3}
ppc_stat_grouped( y = log( y ), 
                  yrep = yrep_3, 
                  stat = function(x) 3 * ( mean( x ) - median( x ) ) / sd( x ), 
                  group = model_dat$WSA9,
                  binwidth = 0.05 )
```

Here, it looks like the CPL skew is pretty far off, and that this ecoregion is maybe the source of the weird left tail. Maybe there is something wrong with my calculation. For now, though, we'll just move forward, but we should check this eventually.

## Leave-one-out evaluations
Here' we'll again conduct a Bayesian leave-one-out cross validation ('LOO-CV') assessment.

```{r mod3_loo}
loo_3 <- loo( mod3, save_psis = TRUE, cores = 4 )
psis_3 <- loo_3$psis_object
lw_3 <- weights( psis_3 )

print( loo_3 )
```

Compared to our last models, the LOO-CV evaluation for this model resulted in _many_ "problematic" observations. Since we have such a very large number, we'll not use `reloo()` this time, but we'll take the advise from the generated warning and try a kfold-CV.

```{r mod3_reloo}
kfold_3 <- kfold( mod3, K = 10, save_fits = TRUE )
print( kfold_3 )
```

We can now use this k_fold estimate of the elpd to compare to the other models.

```{r mod3_compare}
loo_compare( loo_1, loo_2, kfold_3 )
```

This model is clearly better than the other two by this metric. And the posterior predictive checks above were also much better. Perhaps we're getting closer.

## Check _vs._ survey estimates 
Lets go ahead and check some estimates against those provided by the traditional survey method and our last model.

So, lets get our model estimates for the means on the data scale. We'll again utilize the full posterior extracted above, and create an object for each ecoregion ( and US ) containing the posterior distributions for the means.

```{r mod3_ests}
CPL <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[CPL,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
NAP <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[NAP,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
NPL <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[NPL,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
SAP <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[SAP,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
SPL <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[SPL,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
TPL <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[TPL,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
UMW <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[UMW,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
WMT <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[WMT,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
XER <- exp( mod3_post$b_Intercept + mod3_post$`r_WSA9[XER,Intercept]` + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )
US <- exp( mod3_post$b_Intercept + ( 0.5 * exp( mod3_post$b_sigma_Intercept ) ^ 2 ) )

mod3_ests <- data.frame( 
  ecoregion = c( "CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER", "US" ),
  estimate = c( mean( CPL ), mean( NAP ), mean( NPL ), mean( SAP ), mean( SPL ), 
                mean( TPL ), mean( UMW ), mean( WMT ), mean( XER ), mean( US ) ),
  LCL = c( quantile( CPL , 0.025 ), quantile( NAP , 0.025 ), quantile( NPL , 0.025 ), 
           quantile( SAP , 0.025 ), quantile( SPL , 0.025 ), quantile( TPL , 0.025 ), 
           quantile( UMW , 0.025 ), quantile( WMT , 0.025 ), quantile( XER , 0.025 ), 
           quantile( US , 0.025 ) ),
  UCL = c( quantile( CPL , 0.975 ), quantile( NAP , 0.975 ), quantile( NPL , 0.975 ), 
           quantile( SAP , 0.975 ), quantile( SPL , 0.975 ), quantile( TPL , 0.975 ), 
           quantile( UMW , 0.975 ), quantile( WMT , 0.975 ), quantile( XER , 0.975 ), 
           quantile( US , 0.975 ) )
  ) %>%
  mutate( estimate = round( estimate, 1 ), # round to nearest tenth
          LCL = round( LCL, 1 ),
          UCL = round( UCL, 1 )
  )  %>%
  mutate( ecoregion = factor( ecoregion, #relevel for ggplot so not alphabetical, but original order
                              levels = c( "CPL", "NAP", "NPL", "SAP", "SPL",
                                          "TPL", "UMW", "WMT", "XER", "US" ) ) ) %>%
  mutate( model = "mod3" )

rm( CPL, NAP, NPL, SAP, SPL, TPL, UMW, WMT, XER, US ) #clean up workspace
print( mod3_ests )
```

Lets combine our estimates from the survey, mod2, and now mod3 into one dataframe.

```{r combine_ests_mod3}
mod3_ests <- data.frame( rbind( mod2_ests, mod3_ests ) ) 
print( mod3_ests )
```

Finally, lets plot these estimates alongside the ones from the survey and mod2.

```{r survey_mod3_ests_plot, fig.align = "center"}
ggplot( mod3_ests, aes( x = ecoregion, y = estimate, group = model ) ) +
  geom_point( aes(colour = model ) , position=dodge ) +
  geom_linerange( aes( ymin = LCL, ymax = UCL , color = model ) , position=dodge ) +
  scale_colour_manual( values = c( "gray55", "black", "gray85" ) ) +
  coord_flip() +
  theme_bw()
```

In all cases, the uncertainty intervals for mod3 are much narrower than those for mod2. They are also narrower than the estimates from the survey-based method, except for the NPL where the intervals were still pretty similar in width. Likewise, these estimates don't suggest a ton of variability in the means from one ecoregion to the next. However, because they have greater certainty, we might be able to distinguish some differences in means for a few ecoregions (e.g., CPL _vs._ NAP ).

# References

### 2023-03-02 scripts/NLA_N2O_models.Rmd
Initial commit of multivariate N2O modeling workflow.

---
title: "Bayesian models for NLA 2017 $N_2O$ survey data"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_depth: 4
    toc_float: yes
    code_folding: show
    font_size: 14
    number_sections: yes
    theme: simplex
bibliography: RWM_Endnote_Library.bib
link-citations: yes
editor_options:
  chunk_output_type: inline
---

```{r begin, eval=TRUE, include=FALSE}
library(ggpubr)
library(moments)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(kableExtra)
library(ggrepel)
library(dplyr)
library(tidyverse)
library(tidyr)
library(future)
library(foreach)
library(itertools)
library(bayesplot)
library(tidybayes)
library(brms)

options(mc.cores = parallel::detectCores(logical = FALSE))
options( max.print = 1000 )

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}

# stat: skew 
skew <- function(x) {
  xdev <- x - mean(x)
  n <- length(x)
  r <- sum(xdev^3) / sum(xdev^2)^1.5
  return(r * sqrt(n) * (1 - 1/n)^1.5)
}

# function for DHARMa residual analysis
check_brms <- function(model,             # brms model
                       integer = FALSE,   # integer response? (TRUE/FALSE)
                       plot = TRUE,       # make plot?
                       resp = NULL,
                       ...                # further arguments for DHARMa::plotResiduals 
) {

  mdata <- brms::standata(model)
  if(!"Y" %in% names(mdata))
    oResp <- mdata[[paste0("Y_", resp)]]
  else
    oResp <- mdata[["Y"]]
  #  stop("Cannot extract the required information from this brms model")

  dharma.obj <- DHARMa::createDHARMa(
    simulatedResponse = t(brms::posterior_predict(model, resp = resp, ndraws = 1000)),
    observedResponse = oResp, 
    fittedPredictedResponse = apply(
      t(brms::posterior_epred(model, resp = resp, ndraws = 1000, re.form = NA)),
      1,
      median),
    integerResponse = integer)

  if (isTRUE(plot)) {
    plot(dharma.obj, ...)
  }

  invisible(dharma.obj)

}
```

# Rationale and Objectives
This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O concentrations and saturation ratios using the 2017 Nation Lakes Assessment (NLA) survey data. The NLA sampling sites were distributed among the target population of US lakes (in the lower 48 states) according to a probabilistic survey design with samples stratified among categories of lake surface area, WSA9 ecoregion, and US state (excluding AK and HI). Due to the stratification scheme, some types of lakes in the sample population were intentionally over-represented (e.g., large lakes) and some were under-represented (e.g., small lakes) relative to the target population. Due to the unequal probability design, inferences from the sample had to be adjusted for inferences on the broader populations of interest (e.g, National-, state-, ecoregion-, and size class-specific estimates).

The concept of the "complete data likelihood" is useful for conceptualizing biases arising from sampling design [@Zachmann_etal_2022; @Gelman_etal_2014 Ch. 8; @Link_Barker_2010]. For the NLA survey data, the population of US lakes in the lower 48 states larger than 4 hectares was considered the complete data and the probabilistic samples were considered a subset of that complete data. The portion of US lakes not included in the sample were considered "missing" from the complete data _not_ at random, but conditional on the pre-specified design (stratification) variables. This non-random missingness was not ignorable for the purpose of making inferences from the sample to the target population. In a model-based framework, however, including the design parameters as predictors in a regression model is one way to adjust for the missingness. For a thorough and recent treatment of this concept in the context of national surveys of environmental resources, refer to [@Zachmann_etal_2022]. This concept is a key motivator for the increasingly popular mulitilevel regression with poststratification (MRP) approach to model-based inference [@Gelman_etal_2014; Gelmant_etal_2020 Ch. 17].

The following workflow illustrates our model-based approach, based largely on the logic of MRP, but with an elaboration on the poststratification step to enable eventual estimates of total gas flux at the population level, which required scaling up from lake-level estimates. The typical MRP process is carried out in two steps. The first step is to fit regression models for the response variables of interest (e.g., dissolved N2O, equilibrium N2O) conditional on the survey design variables ┼(i.e., ecoregion, state, lake size). The second step is post-stratfication, wherein the posterior parameter estimates from the regression model for the sample population are weighted based on their known or assumed distribution in the population of interest [i.e., post-stratification table; @Gelman_etal_2020 Ch. 17]. The poststratification table in our case, for example, would be a population summary of lakes among the design variables: ecoregion, state, and size category. However, because we eventually needed lake-level estmates, instead of predicting to a postratification table, we predicted to each individual lake in the population of interest. This meant predicting to the full target population of 465,897 natural and man made US lakes larger than 4 hectares in the lower 48 states. These predictions were assumed relevant to average conditions during the "index period" for each lake in 2017. Details about the sampling frame as well as the target population are further clarified in the workbook below with data summaries and code.

For the regressions, we used multilevel models fit in a fully Bayesian fashion Multilevel models are thought to work well in this context because they provide regularized estimates along the design groupings, which can improve out-of-sample inferences [@McElreath_2020]. Inferences for lake types that may be missing from the sample, but are part of the population of interest are also straightforward using this approach [@Gelman_etal_2020 Ch. 17; @McElreath_2020]. More information these models, their specific parameters, R code, fit evaluations, and resulting inferences are presented in this document.

The overriding objective of the modeling effort was to provide population level estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1). The estimates would also be used to later estimate the total flux of N2O gas attributable to the target population of lakes over the index period. The saturation ratio estimates were calculated as a derived quantity based on the ratio of modeled dissolved to equilibrium N2O. Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), we developed models for estimating their joint distribution. The response variable in the models was, therefore, multivariate to account for potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography. Although point predictions of the mean marginal probabilities from separate models could be comparable, a joint model allowing correlated observation-level errors (i.e., residuals) was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021]. All of the models fit were constructed using the `brms` package [@Burkner_2017] in `R` [@R_Core_Team_2021] as an interface to Stan, a software package for fitting fully Bayesian models via Hamiltonian Monte Carlo [HMC; @Stan_Development_Team_2018_a; @Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c].

# Data
As explained in a previous data munging document document
(https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html), duplicate dissolved gas samples were collected at a depth of ~0.1m at designated index sites distributed across 1091 lakes nationwide, of which 95 were sampled twice as repeat visits. This randomly selected subset of revisit sites was used as a test set for assessing model fit and out-of-sample performance.

Gas samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L. The samples were collected under a stratified, unequal probability design and each gas observation was indexed to an individual lake selected with unequal probability from 5 different lake size categories, $j \in j=1,...,J = 5$, according to surface area (ha), and from within a state, $k \in k=1,...,K = 48$, situated within an aggregated, WSA9 or Omernik ecoregion, $l \in l=1,...,L = 9$. All 9 WSA9 ecoregions were represented in the sample, including Xeric (XER), Western Mountain (WMT),  Northern Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and Southern Appalachian (SAP) regions. As shown below, the data from the initial and revisit samples were separately compiled into data frame objects in $\textbf{R}$, with $n=984$ and $n=95$ rows, respectively, of gas observations indexed to the survey design variables and several potentially relevant covariates.

## Import
The gas data and covariates were previously described and munged at 
https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html. That dataset was imported below. 
```{r import_data, eval=FALSE, include=TRUE}
load( file = paste0( localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/dg.2021-02-01.RData")
      )

save(dg, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda") 
```

From the imported dataset, a new data frame for modeling was constructed from the original file including only the variables of interest: (1) the N2O gas observations; (2) the survey design variables indexed to those observations; and (3) additional covariates considered potentially useful for improving the fit of the model. The data frame below excluded the second-visit observations, which would later be used for model checking. Some variables from the imported data were renamed for convenience. In addition, the NO3 covariate was rounded according to the documented measurement precision. An alternative version of the NO3 covariate was also created in this step by log-transforming and re-coding it as an ordered factor with five levels at hand-drawn cut points. The left-most cut point separated observations below the detection limit from the completely observed samples. The remaining cut points in the positive direction were drawn at approximately equal distances along the log scale. Finally, it should be noted that one lake that was sampled was missing information on the N2O gas measurements and it was removed from the data frame.
```{r model_data, echo=TRUE, paged.print=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda")

dg %>%
  filter(!is.na(dissolved.n2o.nmol)) %>% # 1 obs with missing measurement
  nrow() # number of observations before filtering

df_model <- dg %>%
  filter(!is.na(dissolved.n2o.nmol)) %>%
  filter(sitetype == "PROB") %>% # probability samples only
  filter(visit.no == 1) %>%
  mutate(n2o = round(dissolved.n2o.nmol, 2),
         n2o_eq = round(sat.n2o.nmol, 2),
         n2o_sat = n2o.sat.ratio,
         n2o_em = e.n2o.nmol.d,
         n2o_flux = f.n2o.m.d,
         WSA9 = factor(ag.eco9),
         state = factor(state.abb[match(state.nm, state.name)]),
         area_ha = area.ha,
         log_area = log(area_ha),
         chla = chla.result,
         log_chla = log(chla),
         elev = elevation,
         log_elev = log(elev + 1),
         do_surf = o2.surf,
         log_do = log(do_surf),
         bf_max = max.bf,
         sqrt_bf = sqrt(bf_max),
         size_cat = recode(area.cat6, 
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  mutate(size_cat = factor(size_cat,
                           levels = c("min_4", "4_10", "10_20", "20_50", "50_max"),
                           ordered = TRUE)) %>%
  mutate(no3 = ifelse(nitrate.n.result <= 0.0005, 0.0005, round(nitrate.n.result, 4))) %>%# 1/2 mdl 0.01
  mutate(no3_cat = cut(log(no3), # convert no3 to ordered factor with 5 levels
                       breaks = c(-Inf, -7.5, -5.5, -3.5, -1.5, Inf),
                       labels =seq(1, 5, 1))) %>%
  mutate(no3_cat = factor(no3_cat,
                          levels = seq(1, 5, 1),
                          ordered = TRUE)) %>%
  mutate(date = as.Date(date.col)) %>%
  mutate(jdate = as.numeric(format(date, "%j"))) %>% 
  mutate(lat = map.lat.dd,
         lon = map.lon.dd) %>% # longitude
  mutate(surftemp = surftemp,
         log_surftemp = log(surftemp)) %>% 
  select(WSA9,
         state,
         size_cat,
         site.id,
         lat,
         lon,
         date,
         jdate,
         surftemp,
         log_surftemp,
         area_ha,
         log_area,
         elev,
         log_elev,
         chla,
         log_chla,
         do_surf,
         log_do,
         bf_max,
         sqrt_bf,
         n2o,
         n2o_eq,
         no3,
         no3_cat
         )

save(df_model, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

nrow(df_model) # number of obs after filtering

print(df_model)
```

A second dataframe, including only the second visit observations, was constructed below. These data were later used as a "test set" to assess the out-of-sample fit of the model developed on the first-visit or training data.
```{r test_data, echo=TRUE, paged.print=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda")

# number of observations before filtering probability samples
dg %>%
  filter(!is.na(dissolved.n2o.nmol)) %>% # remove obs with missing response measurements
  nrow()

df_test <- dg %>%
  filter(!is.na(dissolved.n2o.nmol)) %>%
  filter(sitetype == "PROB") %>% # probability samples only
  filter(visit.no == 2) %>%
  mutate(n2o = round(dissolved.n2o.nmol, 2),
         n2o_eq = round(sat.n2o.nmol, 2),
         n2o_sat = n2o.sat.ratio,
         n2o_em = e.n2o.nmol.d,
         n2o_flux = f.n2o.m.d,
         WSA9 = factor(ag.eco9),
         state = factor(state.abb[match(state.nm, state.name)]),
         area_ha = area.ha,
         log_area = log(area_ha),
         chla = chla.result,
         log_chla = log(chla),
         elev = elevation,
         log_elev = log(elev + 1),
         do_surf = o2.surf,
         log_do = log(do_surf),
         bf_max = max.bf,
         sqrt_bf = sqrt(bf_max),
         size_cat = recode(area.cat6, 
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  mutate(size_cat = factor(size_cat,
                           levels = c("min_4", "4_10", "10_20", "20_50", "50_max"),
                           ordered = TRUE)) %>%
  mutate(no3 = ifelse(nitrate.n.result <= 0.0005, 0.0005, round(nitrate.n.result, 4))) %>%# 1/2 mdl 0.01
  mutate(no3_cat = cut(log(no3), # convert no3 to ordered factor with 5 levels
                       breaks = c(-Inf, -7.5, -5.5, -3.5, -1.5, Inf),
                       labels =seq(1, 5, 1))) %>%
  mutate(no3_cat = factor(no3_cat,
                          levels = seq(1, 5, 1),
                          ordered = TRUE)) %>%
  mutate(date = as.Date(date.col)) %>%
  mutate(jdate = as.numeric(format(date, "%j"))) %>% 
  mutate(lat = map.lat.dd,
         lon = map.lon.dd) %>% # longitude
  mutate(surftemp = surftemp,
         log_surftemp = log(surftemp)) %>% 
  select(WSA9,
         state,
         size_cat,
         site.id,
         lat,
         lon,
         date,
         jdate,
         surftemp,
         log_surftemp,
         area_ha,
         log_area,
         elev,
         log_elev,
         chla,
         log_chla,
         do_surf,
         log_do,
         bf_max,
         sqrt_bf,
         n2o,
         n2o_eq,
         no3,
         no3_cat
         )

save(df_test, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")

nrow(df_test) # number of obs after filtering for probability samples, first visits, and removing one site missing ecoregion (WSA9) info.

print(df_test)
```

## Target population
Below. the NLA sampling frame was imported and then filtered to include only the target population or sampling frame for this project.
```{r import_sample_frame, echo=TRUE}
df_pop <- read.csv(file = paste0(localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/NLA_Sample_Frame.csv"), header = T)

sframe <- df_pop %>%
  filter(nla17_sf != "Exclude2017") %>%
  filter(nla17_sf != "Exclude2017_Include2017NH") %>%
  filter(state != "DC") %>%
  filter(state != "HI") %>%
  droplevels() %>%
  mutate(WSA9 = factor(ag_eco9),
         WSA9 = forcats::fct_drop(WSA9), # remove NA level
         state = factor(state),
         size_cat = factor(area_cat6),
         lat = lat_dd83,
         lon = lon_dd83,
         log_area = log(area_ha),
         elev = elevation,
         log_elev = ifelse(elev <= 0, 0, elev), # assumed elev < 0 to be elev = 0
         log_elev = log(log_elev + 1)
         ) %>% 
  mutate(size_cat = recode(size_cat, 
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  mutate(size_cat = factor(size_cat, 
                           levels = c("min_4", "4_10", "10_20", "20_50", "50_max"),
                           ordered = TRUE)) %>%
  select(WSA9, state, size_cat, lat, lon, area_ha, log_area, elev, log_elev)

rm(df_pop)

save(sframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

print(sframe)
```

The resulting target population above included a total of 465,897 waterbodies.

Cross tabulations below describe the structure of the target population with respect to the design variables. The cross-tabulation makes it clear that each ecoregion does not contain each state. Therefore, in the statistical sense, states were nested in ecoregions.
```{r frame_dimensions_1, echo=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

sframe %>%
  group_by(WSA9, state) %>%
  summarise(n = n(), .groups = "drop") %>%
  spread(state, n) %>%
  print()
```

Likewise, lake size category was nested in state (which was nested in ecoregion). That is, not every ecoregion:state in the population of interest contained every size category (below).
```{r frame_dimensions_4, echo=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

sframe %>%
  group_by(WSA9, state, size_cat) %>%
  summarise(n = n(), .groups = "drop") %>%
  spread(size_cat, n) %>%
  print()
```

Below, the sampling frame was selected down to create a post-stratification table. Some of the variables were renamed to match the naming conventions used in the observational data above. There were 536 types of lakes in the population of interest with respect to the sampling design. The counts of those lake types (n_lakes) and their proportions relative to the total population of lakes in the sampling frame (prop_cell) are indicated below.
```{r filter_frame, echo=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

pframe <- sframe %>%
  mutate(obs = 1) %>%
  group_by(WSA9, state, size_cat) %>%
  summarise(n_lakes = sum(obs), .groups = "drop") %>%
  ungroup() %>%
  mutate(prop_cell = n_lakes/sum(n_lakes)) %>%
  mutate(type = "population")

save(pframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

print(pframe)
```

## Sample vs. population
Below, the lake distributions in the population of interest were compared to the proportions in the observed sample. There were 352 lake types in the sample compared to the 536 in the population of of interest. In total, there were 984 observations distributed across these 352 lake types in the sample; and the number of samples was not distributed evenly across the types. Some cells were represented by as few as 1 lake. In total, 536-352 = 184 lake types in the population of interest were not represented in the sample.
```{r sample_cell_counts, echo=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

samp_props <- df_model %>%
  mutate(obs = 1) %>%
  group_by(WSA9, state, size_cat) %>%
  summarize(n_lakes = sum(obs), .groups = "drop") %>%
  ungroup() %>%
  mutate(prop_cell = round(n_lakes / sum(n_lakes), 7)) %>%
  mutate(type = "sample")

save(samp_props, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

print(samp_props)
```

Below, a graphical comparison was constructed to depict the distribution of cells in the population of interest _versus_ those in the sample.
```{r compare_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

pframe %>%
  bind_rows(samp_props) %>%
  ggplot(aes(x = interaction(WSA9, state, size_cat), y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes( shape = type, color = type)) +
  geom_line() +
  theme_tidybayes() +
  theme(axis.text.x = element_blank()) +
  xlab("WSA9:state:size") +
  ylab("proportion in cell")
```

Another comparison between population and sample was constructed below by ecoregion. The samples were not balanced across ecoregions. Lakes in the Coastal Plains (CPL) ecoregion, for example, were clearly undersampled relative to their proportion of the population.
```{r eco_props_pop, eval=FALSE, include=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

pframe_eco <- pframe %>%
  group_by(WSA9) %>%
  summarise(n_lakes = sum(n_lakes)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_lakes/sum(n_lakes), 7)) %>%
  ungroup() %>%
  mutate(type = 'population')

save(pframe_eco, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_eco.rda")
```

```{r eco_props_sample, eval=FALSE, include=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

samp_props_eco <- samp_props %>%
  group_by(WSA9) %>%
  summarise(n_lakes = sum(n_lakes)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_lakes/sum(n_lakes), 7)) %>%
  ungroup() %>%
  mutate(type = 'sample')

save(samp_props_eco, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_eco.rda")
```

```{r compare_eco_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=6}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_eco.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_eco.rda")

pframe_eco %>%
  bind_rows(samp_props_eco) %>%
  ggplot(mapping = aes(x = WSA9, y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes( shape = type, color = type), size = 3) +
  geom_line() +
  theme_tidybayes() +
  xlab("Ecoregion") +
  ylab("proportion in cell") + 
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  theme(text = element_text(size = 12))
```

A similar comparison by state was constructed below.
```{r state_props_pop, eval=FALSE, include=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

pframe_state <- pframe %>%
  group_by(state) %>%
  summarise(n_lakes = sum(n_lakes)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_lakes/sum(n_lakes), 7)) %>%
  ungroup() %>%
  mutate(type = 'population')

save(pframe_state, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_state.rda")
```

```{r state_props_sample, eval=FALSE, include=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

samp_props_state <- samp_props %>%
  group_by(state) %>%
  summarise(n_lakes = sum(n_lakes)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_lakes/sum(n_lakes), 7)) %>%
  ungroup() %>%
  mutate(type = 'sample')

save(samp_props_state, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_state.rda")
```

```{r compare_state_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_state.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_state.rda")

pframe_state %>%
  bind_rows(samp_props_state) %>%
  ggplot(mapping = aes(x = state, y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes(shape = type, color = type)) +
  geom_line() +
  theme_tidybayes() +
  theme(axis.text.x = element_text(angle = 45)) +
  xlab("State") +
  ylab("proportion in cell")
```

Finally, a comparison by lake size category is shown below. Note that small lakes were under-sampled relative to larger lakes by design.
```{r size_props_pop, eval=FALSE, include=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")

pframe_size <- pframe %>%
  group_by(size_cat) %>%
  summarise(n_lakes = sum(n_lakes)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_lakes/sum(n_lakes), 7)) %>%
  ungroup() %>%
  mutate(type = 'population')

save(pframe_size, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_size.rda")
```

```{r size_props_sample, eval=FALSE, include=TRUE, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")

samp_props_size <- samp_props %>%
  group_by(size_cat) %>%
  summarise(n_lakes = sum(n_lakes)) %>%
  ungroup() %>%
  mutate(prop_cell = round(n_lakes/sum(n_lakes), 7)) %>%
  ungroup() %>%
  mutate(type = 'sample')

save(samp_props_size, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_size.rda")
```

```{r compare_size_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=6}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_size.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_size.rda")

pframe_size %>%
  bind_rows(samp_props_size) %>%
  ggplot(mapping = aes(x = size_cat, y = prop_cell, group = type, linetype = type)) +
  geom_point(stat = "identity", aes( shape = type, color = type)) +
  geom_line() +
  theme_tidybayes() +
  xlab("Size category") +
  ylab("proportion in cell")
```

## Sample-based estimates
The overall mean and standard deviation for N2O in the sample:
```{r sample_summary_n2o}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  summarise(mean = mean(n2o),
             sd = sd(n2o)) %>%
  print()
```

The same summary for equilibrium N2O:
```{r sample_summary_n2oeq}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  summarise(mean = mean(n2o_eq),
             sd = sd(n2o_eq)) %>%
  print()
```

The saturation ratio (i.e., N2O / N2O-eq):
```{r sample_summary_n2osat}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  summarise(mean = mean(n2o / n2o_eq),
             sd = sd(n2o / n2o_eq)) %>%
  print()
```

Finally, roughly 67% of lakes in the sample were undersaturated (i.e., saturation ratio < 1):
```{r sample_summary_propsat}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  summarise(prop_undersat = sum((n2o / n2o_eq) < 1) / 984) %>%
  print()
```

Using only the sample observations, a plot was constructed below of the overall mean (dashed line) along with the ecoregion-specific means (black circles). The shaded areas indicate +/- 1 standard deviation. Neither dissolved N2O nor the saturation ratio were clearly structured by ecoregion in the sample, but there did appear to be some structure in the equilibrium N2O observations.
```{r sample_summary_eco, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

p1 <- df_model %>%
  group_by(WSA9) %>%
  summarise(mean = mean(n2o),
             sd = sd( n2o)) %>%
  ggplot(aes(x = WSA9, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = WSA9), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = WSA9, y = mean))+
  geom_point()+
  geom_hline(yintercept = 8.72, linetype = 'dashed') +
  theme_tidybayes() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) + 
  ylab("") +
  ggtitle("N2O")

p2 <- df_model %>%
  group_by(WSA9) %>%
  summarise(mean = mean(n2o_eq),
             sd = sd(n2o_eq)) %>%
  ggplot(aes(x = WSA9, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = WSA9), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = WSA9, y = mean))+
  geom_point()+
  geom_hline(yintercept = 7.48, linetype = 'dashed') +
  theme_tidybayes() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  ylab("Sample mean") +
  ggtitle("N2O equilibrium")

p3 <- df_model %>%
  group_by(WSA9) %>%
  summarise(mean = mean(n2o / n2o_eq),
             sd = sd(n2o / n2o_eq)) %>%
  ggplot(aes(x = WSA9, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = WSA9), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = WSA9, y = mean))+
  geom_point()+
  geom_hline(yintercept = 1.17, linetype = 'dashed') +
  theme_tidybayes() +  
  xlab("Ecoregion") +
  ylab("") +
  ggtitle("N2O saturation ratio")

grid.arrange(p1, p2, p3)
```

The same summary by state is below.
```{r sample_summary_state, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

p1 <- df_model %>%
  group_by(state) %>%
  summarise(mean = mean(n2o),
             sd = sd( n2o)) %>%
  ggplot(aes(x = state, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = state), 
              fill = 'lightgrey', 
              alpha = .7) +
  geom_line(aes(x = state, y = mean))+
  geom_point()+
  geom_hline(yintercept = 8.72, linetype = 'dashed') + 
  theme_tidybayes() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) + 
  ylab("") +
  ggtitle("N2O")

p2 <- df_model %>%
  group_by(state) %>%
  summarise(mean = mean(n2o_eq),
             sd = sd(n2o_eq)) %>%
  ggplot(aes(x = state, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = state), 
              fill = 'lightgrey', 
              alpha = .7) +
  geom_line(aes(x = state, y = mean))+
  geom_point()+
  geom_hline(yintercept = 7.48, linetype = 'dashed') +
  theme_tidybayes() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  ylab("Sample mean") +
  ggtitle("N2O equilibrium")

p3 <- df_model %>%
  group_by(state) %>%
  summarise(mean = mean(n2o / n2o_eq),
             sd = sd(n2o / n2o_eq)) %>%
  ggplot(aes(x = state, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = state), 
              fill = 'lightgrey', 
              alpha = .7)+
  geom_line(aes(x = state, y = mean))+
  geom_point()+
  geom_hline(yintercept = 1.17, linetype = 'dashed') +
  theme_tidybayes() + 
  theme(axis.text.x = element_text(angle = 45)) + 
  xlab("State") +
  ylab("") +
  ggtitle("N2O saturation ratio")

grid.arrange(p1, p2, p3)
```

Finally, the same summary by size category:
```{r sample_summary_size, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

p1 <- df_model %>%
  group_by(size_cat) %>%
  summarise(mean = mean(n2o),
             sd = sd( n2o)) %>%
  ggplot(aes(x = size_cat, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = size_cat), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = size_cat, y = mean))+
  geom_point()+
  geom_hline(yintercept = 8.72, linetype = 'dashed') +
  theme_tidybayes() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) + 
  ylab("") +
  ggtitle("N2O")

p2 <- df_model %>%
  group_by(size_cat) %>%
  summarise(mean = mean(n2o_eq),
             sd = sd(n2o_eq)) %>%
  ggplot(aes(x = size_cat, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = size_cat), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = size_cat, y = mean))+
  geom_point()+
  geom_hline(yintercept = 7.48, linetype = 'dashed') +
  theme_tidybayes() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  ylab("Sample mean") +
  ggtitle("N2O equilibrium")

p3 <- df_model %>%
  group_by(size_cat) %>%
  summarise(mean = mean(n2o / n2o_eq),
             sd = sd(n2o / n2o_eq)) %>%
  ggplot(aes(x = size_cat, y = mean, group = 1)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, x = size_cat), fill = 'lightgrey', alpha = .7)+
  geom_line(aes(x = size_cat, y = mean))+
  geom_point()+
  geom_hline(yintercept = 1.17, linetype = 'dashed') +
  theme_tidybayes() + 
  theme(axis.text.x = element_text(angle = 45)) + 
  xlab("Size class") +
  ylab("") +
  ggtitle("N2O saturation ratio")

grid.arrange(p1, p2, p3)
```

## Sample data exploration
Below, the empirical distribution of N2O observations in the sample was summarized using a density and rug plot below. Note the natural log scale of the x axis. Both the N2O and equilibrium N2O data had considerable right skew even after the log transformation, which was not unexpected and has been noted in other studies [@Webb_etal_2019]. The saturation ratio was also skewed since it was derived from the other two observed variables (i.e., sat_ratio = n2o / n2o_eq).
```{r summary_N2O, echo=FALSE, fig.align='center', fig.height=8, fig.width=5, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

p1 <- df_model %>%
  ggplot(aes(x = n2o)) +
  geom_boxplot(aes(x = n2o, y = -0.5), outlier.shape = NA, alpha = 0.7) + 
  geom_density(aes(x = n2o)) +
  geom_rug(aes(x = n2o), show.legend = F ) +
  theme(text = element_text(size=12)) +
  scale_x_continuous(trans = "log", breaks = c(0, 5, 10, 25, 50, 150)) +
  ylab("density") +
  xlab("N2O (nmol/L)") +
  theme_tidybayes() +
  theme(axis.text.y = element_blank())

p2 <- df_model %>%
  ggplot(aes(x = n2o_eq)) +
  geom_boxplot(aes(x = n2o_eq, y = -0.5), outlier.shape = NA, alpha = 0.7) + 
  geom_density(aes(x = n2o_eq)) +
  geom_rug(aes(x = n2o_eq), show.legend = F ) +
  theme(text = element_text(size=12)) +
  scale_x_continuous(trans = "log") +
  ylab("density") +
  xlab("Equilibrium N2O (nmol/L)") +
  theme_tidybayes() +
  theme(axis.text.y = element_blank())

p3 <- df_model %>%
  ggplot(aes(x = n2o / n2o_eq)) +
  geom_boxplot(aes(x = n2o / n2o_eq, y = -0.5), outlier.shape = NA, alpha = 0.7) + 
  geom_density(aes(x = n2o / n2o_eq)) +
  geom_rug(aes(x = n2o / n2o_eq), show.legend = F ) +
  theme(text = element_text(size=12)) +
  scale_x_continuous(trans = "log", breaks = c(0, 1, 5, 10, 20)) +
  ylab("density") +
  xlab("N2O saturation ratio") +
  theme_tidybayes() +
  theme(axis.text.y = element_blank())

grid.arrange(p1, p2, p3)
```

Below are plots of N2O vs. NO3. The first plot shows log(N2O) vs. log(NO3), as well as the ordinal categories assigned to NO3 (vertical lines). The leftmost vertical line is dashed and separates the NO3 observations below the detection limit.
```{r summary_N2O_vs_NO3, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = log(no3), y = log(n2o))) +
  geom_point(show.legend = F) +
  geom_smooth(method = "loess", span = 2) +
  geom_vline(xintercept = -7.5, linetype = "dashed") +
  geom_vline(xintercept = -5.5) +
  geom_vline(xintercept = -3.5) +
  geom_vline(xintercept = -1.5) +
  theme(text = element_text(size=12)) +
  theme_bw()
```

In the plot above, the trend is increasing and nonlinear on the log scale. The increasing variance in N2O along the NO3 gradeient suggested a potential mediator of the relationship between NO3 on N2O. Below are plots of N2O vs. NO3 for 6 quantiles of the surface temperature measurements (quantiles increasing from 1 to 6). This plot below suggested that the NO3 effect on N2O may have been stronger in lakes with higher observed temperatures.
```{r summary_N2O_vs_NO3_surftemp, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = log(no3), y = log(n2o))) +
  geom_point(show.legend = F) +
  geom_smooth(method = "loess", span = 2) +
  theme(text = element_text(size=12)) +
  facet_wrap(~ as.factor(ntile(surftemp, 6))) +
  theme_bw()
```

The next plot below shows the relationship between N2O and NO3 at 6 different quantiles (increasing 1 to 6) of the log-scaled lake surface area estimates.
```{r summary_N2O_vs_NO3_logarea, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = log(no3), y = log(n2o))) +
  geom_point(show.legend = F) +
  geom_smooth(method = "loess", span = 2) +
  theme(text = element_text(size=12)) +
  facet_wrap(~ as.factor(ntile(log_area, 6))) +
  theme_bw()
```

Similar plots are below, but with NO3 expressed as an ordered categorical variable with 5 levels. The positive and monotonic trends area similar to the previous plots where NO3 was treated as continuous. Note the large number of observations in the first NO3 category (no3_cat = 1). This category represented all of the censored observations for NO3, which was most of the data.
```{r summary_N2O_vs_NO3cat, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = no3_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(outlier.shape = NA, notch = TRUE, color = "black", alpha = 0.7) +
  theme(text = element_text(size=12)) +
  theme_bw()
```

```{r summary_N2O_vs_NO3cat_surftemp, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = no3_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(outlier.shape = NA, color = "black", alpha = 0.7) +
  theme(text = element_text(size=12)) +
  facet_wrap(~ as.factor(ntile(surftemp, 6))) +
  theme_bw()
```

```{r summary_N2O_vs_NO3cat_logarea, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = no3_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(outlier.shape = NA, color = "black", alpha = 0.7) +
  theme(text = element_text(size=12)) +
  facet_wrap(~ as.factor(ntile(log_area, 6))) +
  theme_bw()
```

Below is a plot of log(N2O) vs. log(NO3) by ecoregion, which suggested that the NO3 effect on N2O may have varied by ecoregion.
```{r summary_N2O_vs_NO3_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = log(no3), y = log(n2o))) +
  geom_point(show.legend = F) +
  geom_smooth(method = "loess", span = 2) +
  geom_vline(xintercept = -7.5, linetype = "dashed") +
  geom_vline(xintercept = -5.5) +
  geom_vline(xintercept = -3.5) +
  geom_vline(xintercept = -1.5) +
  theme(text = element_text(size=12)) +
  facet_wrap(~ WSA9) +
  theme_bw()
```

Below is the same plot as above but for the ordered categorical version of NO3.
```{r summary_N2O_vs_NO3cat_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  ggplot(aes(x = no3_cat, y = log(n2o), color = 1)) +
  geom_point( position = position_jitterdodge(), show.legend = F ) +
  geom_boxplot(aes(x = no3_cat, y = log(n2o)), 
               outlier.shape = NA, 
               color = "black", 
               alpha = 0.7) + 
  facet_wrap(~ WSA9) +
  theme(text = element_text(size=12)) +
  theme_bw()
```

A plot below shows trends by state within just the Temperate Plains (TPL) ecoregion. Within states, the number of observations were relatively small, but the trends appeared closer to linear.
```{r summary_N2O_vs_NO3_wsa9state3, echo=FALSE, fig.align='center', fig.height=4, fig.width=8, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

df_model %>%
  filter(WSA9 == "TPL") %>%
  ggplot(aes(x = log(no3), y = log(n2o), group = state, color = state)) +
  geom_point(show.legend = F) +
  geom_smooth(method = "lm", span = 2, alpha = 0.1) +
  theme(text = element_text(size=12)) +
  theme_bw()
```

# Model fitting
The first regression model was constructed to estimate the joint distribution of log-transformed N2O and equilibrium N2O conditional on the the design factors. Each log-transformed observation, $i \in 1,..,N=984$, for each response, $p \in 1:P=2$, was assumed to be drawn from a multivariate normal distribution with the parameters $\nu$ and $\Sigma$, where $\nu$ is the multivariate mean estimated conditional on the design effects and $\Sigma$ is a covariance matrix containing the observation-level variances and residual correlation:
$$Y \sim MVN(\nu, \Sigma)$$

The multivariate mean is a vector of mean parameters, $\nu:[\mu_{p=1}, \mu_{p=2}]$, for each response. Each mean is further defined by a linear combination of parameters where, for each response $p$ and observation $i$:

$$\mu_{pi} = \alpha_{0(pi)} + \alpha_{1(pij)} + \alpha_{2(pijk)} + \alpha_{3(pijkl)} \\
\alpha_1 \sim MVN(0, \Lambda_1) \\
\alpha_2 \sim MVN(0, \Lambda_2) \\
\alpha_3 \sim MVN(0, \Lambda_3)$$

The linear combination of parameters defining $\mu$ above include a fixed global intercept, $a_0$, that is estimated directly from the data, and three separate, latent group-level effects matrices, $\alpha_1, \alpha_2, \alpha_3$. The group effects were assumed to be multivariate normal and are centered on zero in multivariate space. The spread of the effects around zero are determined by a covariance matrix, $\Lambda_1, \Lambda_2, \text{or } \Lambda_3$, which are estimated directly from the data. These covariance terms are further defined where:

$$\Lambda = \begin{pmatrix} 1 & \tau_{p=1} \\ \tau_{p=2} & 1 \end{pmatrix} \chi \begin{pmatrix} 1 & \tau_{p=1} \\ \tau_{p=2} & 1 \end{pmatrix}$$

The $\tau$ parameters are the group-level scale parameters, which constrain the spread of effects for each response, and $\chi$ comprises the group-level residual correlation matrix:

$$\chi = \begin{pmatrix} 1 & \varrho \\ \varrho & 1 \end{pmatrix}$$

wherein $\varrho$ is the group-level residual correlation between responses.

The explicit indexing in the notation above conveys the relationship between the parameters and each observation, $i$, and emphasizes the nested structure of the observations within the group effects. Specifically, every observation, $i$, was nested in a lake size category, $l$, which was nested in a state, $k$, and ecoregion, $j$. The parameter $\alpha_1$, therefore, accounted for ecoregion-scale group effects or deviations from the global mean; $\alpha_2$ accounted for state-level group effects nested in ecoregions; and $\alpha_3$ accounted for lake size group effects within states and ecoregions.

Finally, the observation-level covariance term, $\Sigma$, was parameterized as:
$$\Sigma = \begin{pmatrix} 1 & \sigma_{p=1} \\ \sigma_{p=2} & 1 \end{pmatrix} \Omega \begin{pmatrix} 1 & \sigma_{p=1} \\ \sigma_{p=2} & 1 \end{pmatrix}$$

wherein the $\sigma$ parameters are the observation-level standard deviations for each response and $\Omega$ comprises the observation-level residual correlation matrix:
$$\Omega = \begin{pmatrix} 1 & \rho \\ \rho & 1 \end{pmatrix}$$
wherein $\rho$ is the residual correlation between responses.

For model fitting, priors were needed for all parameters conditioned directly on the data, which included the global intercept, the scale parameters, and the correlation matrices. A normal or Gaussian prior, $N(\mu = 2, \sigma = 1)$ centered near the (log-scale) data means, was used for the global intercept parameter for each response. This prior was considered minimally informative as it placed most (~80%) of the prior mass over values between approximately 2 and 27 ng/L for median N2O or N2O equilibrium concentration and included support in the tails for values approaching 0 ng/L on the lower end and 80 ng/L on the high end. We placed $Exp(2)$ priors over all scale parameters, which placed most of the support between values very close to 0 and values near 1 (central 80% density interval from approximately 0.005 to 1.15). Finally, for the correlation matrices, an $LKJ(\eta =2)$ prior was used, which, for a 2-dimensional response, placed most support for correlations between approximately -0.9 and 0.9. This prior seemed reasonable as there was no clear causal mechanisms that were thought to ensure a strong direct correlation between the N2O measures. Any potential residual dependence was expected to be indirect due to, for example, a common causal factor (e.g., elevation, temperature). For more information on prior choice recommendations in Stan, see: https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations

The $\textbf{brms}$ package [@Burkner_2017] for $\textbf{R}$ [@R_Core_Team_2021] was used to fit all of the models in a fully Bayesian setting. The formula syntax of the $\textbf{brms}$ package is similar to the syntax used in the $\textbf{lme4}$ package that is widely used to fit mixed effects models in frequentist settings In either package, the linear predictor for $\mu$ described above could be expressed as:

$$\sim 1 + (1|WSA9) + (1|WSA9:state) + (1|WSA9:state:size)$$

In the $\textbf{brms}$ package, there is additionaly functionality and syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:

$$ N_2O_{dissolved}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\ 
N_2O_{equilibrium}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$

The above syntax would indicate that the linear predictor for both responses in the multivariate model have the same group-level varying effects, and that each of those effects are allowed to be correlated between responses.

For the remainder of this document, only this simplified syntax is presented to describe the model parameterizations. For more information on $\textbf{brms}$ functionality and syntax with multivariate response models, the package vignette may be helpful, and can be found at: https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html.

## Model 1
The first model fit was the one described above.
```{r n2o_mod_mv_1, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(log(n2o) ~ 1 + 
               (1 | a | WSA9) + 
               (1 | b | WSA9:state) + 
               (1 | c | WSA9:state:size_cat),
             family = gaussian())

bf_n2oeq <- bf(log(n2o_eq) ~ 1 + 
               (1 | a | WSA9) + 
               (1 | b | WSA9:state) +
               (1 | c | WSA9:state:size_cat),
             family = gaussian())

priors <- c(
  prior(normal(2, 1), class = "Intercept", resp = "logn2o"), # centered near data mean
  prior(exponential(2), class = "sd", resp = "logn2o"),
  prior(exponential(2), class = "sigma", resp = "logn2o"),
  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), # centered near data mean
  prior(exponential(2), class = "sd", resp = "logn2oeq"),
  prior(exponential(2), class = "sigma", resp = "logn2oeq"),
  prior(lkj(2), class = "rescor"),
  prior(lkj(2), class = "cor")
  )

n2o_mod1 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
                data = df_model,
                prior = priors,
                control = list(adapt_delta = 0.99, max_treedepth = 14),
                #sample_prior = "only",
                save_pars = save_pars(all = TRUE),
                seed = 145,
                chains=4, 
                iter=5000, 
                cores=4)

save(n2o_mod1, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
```

### Summarize fit
The summaries of the estimated parameters and key HMC convergence diagnostics for the fitted model are printed below. There were no obvious issues with the HMC sampling. All $\hat{R}$ values were less than 1.01 and effective sample size ($ESS$) calculations suggested that the posterior contained a sufficient number of effective samples for conducting inference.
```{r print_mod1, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")

print(n2o_mod1, prior = T)
```
In the summary above, the estimated standard deviations for the varying group effects on the mean behavior of the dissolved N2O response suggested fairly low, but non-zero variability across each of the three levels. The standard deviations estimated for the same varying effects for equilibrium N2O were also relatively small. Finally, note the relatively small, but positive residual correlation between the two N2O responses.

Before investing too much into the interpretation of this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC; @Gelman_etal_2014; @Gelman_etal_2020, Ch. 11].

### Model checks
#### Dissolved N2O
Below are a series of panels illustrating graphical PPCs for the log(N2O) component of the model. The top left panel compares a density plot of the observed data (black line) to density lines drawn for 200 samples from the posterior predictive distribution (PPD; blue lines) of the fitted model. The top right panel similarly compares the cumulative density distributions. The left middle panel simulataneously compares means _vs._ standard deviations for 1000 draws from the PPD (blue dots) to the sample mean and standard deviation (black dot). The right middle panel compares skewness _vs._ kurtosis for 1000 draws from the PPD to the skewness and kurtosis values calculated for the observed data. The bottom left panel compares max _vs._ min values for 1000 draws from the PPD to the max and min values of the sample data. Finally, the bottom right panel shows the observed _vs._ average predicted values for each observation in the sample. The average predicted values were calculated as the mean prediction for each observation in the PPD based on 1000 draws.
```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "scatter_avg",
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
The general takeaway from the PPCs above was that the model replicated the central tendency of the observed data fairly well, but failed to sufficiently replicate other important aspects of the distribution, such as skewness and kurtosis. The observed _vs._ average predictions scatterplot suggested substantial heteroscedasticity in the errors.

The same checks were run below, but for the test set of 95 held-out, second-visit data points. 
```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
grid.arrange(
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "scatter_avg",
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
The patterns in misfit indicated above for the re-visit data were similar to the patterns indicated in the PPCs with the training data.

#### Equilibrium N2O
Below are PPCs for the equilibrium N2O component of the model. As with the dissolved N2O response above, the model did an OK job at replicating the central tendency, but performed less well at replicating some important aspects of the overall distribution. 
```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

Below are the same PPCs for equilibrium N2O in the re-visit sites.
```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1,
         newdata = df_test,
         resp = "logn2oeq",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
#### Bivariate
The graphical check below compares bivariate density contours estimated from the observed data (black lines) to density contours estimated for each of 20 draws from the PPD. The model appeared to do a good job of replicating the bivariate mean, but was poor at representing the overall joint distribution.
```{r ppc_biv1, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 20) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o), y = log(n2o_eq))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

The same bivariate check is shown below for the re-visit data.
```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 20) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o), y = log(n2o_eq))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

#### Saturation
The graphical PPCs below were aimed at evaluating how well the multivariate model did at representing the observed saturation ratio: 
$$N_2O_{dissolved}:N_2O_{equilibrium}$$
This quantity was estimated as a derived variable by simply dividing the N2O PPD by the equilibrium N2O PPD. Likewise, the proportion of under-saturated lakes in the sample was estimated by summing the number of lakes from each posterior predictive draw wherein the ratio was < 1 and dividing that number by the total number of lakes in the sample, which was 984.
Overall, these checks indicated that properly representing the tails of the N2O and N2O-eq observations would likely be necessary in order to better replicate the observed saturation metrics. For example, the model did a poor job replicating the observed proportion of under-saturated lakes, underestimating it by more than 10 percentage points, on average.
```{r ppc_sat1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(logn2o < logn2oeq, 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 984) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```

The top left panel, above, is a density plot of the observed saturation ratio (black line) compared to an estimate using 50 draws from the model (blue lines). The top right panel shows the observed proportion of under-saturated lakes compared to a model estimate based on 1000 draws from the PPD. The left middle panel shows the mean _vs._ standard deviation of the saturation ratio for the observed data compared to the same estimates for 500 posterior draws from the model's PPD. The right middle panel shows the max _vs._ min for the sample compared to 500 draws from the model's PPD. Finally, the bottom left panel shows the observed _vs._ average predicted saturation ratio for all 984 lakes sampled in the dataset.

The same PPCs are show below for the revisit data.
```{r ppc_sat1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(logn2o < logn2oeq, 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 95) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/95)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```
The checks above indicated that the model did a similarly underwhelming job of replicating some key properties of the saturation metrics calculated from the re-visit data.

#### R-square
Below, the Bayesian $R^2$ values are reported for each reasponse in the model. 
```{r r2_1, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
round(bayes_R2(n2o_mod1, resp = "logn2o", cores = 1), 3)
round(bayes_R2(n2o_mod1, resp = "logn2oeq", cores = 1), 3)
```

The $R^2$ were also estimated for the re-visit data.
```{r r2_1_test, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
round(bayes_R2(n2o_mod1, resp = "logn2o", newdata = df_test, cores = 1), 3)
round(bayes_R2(n2o_mod1, resp = "logn2oeq", newdata = df_test, cores = 1), 3)
```

## Model 2
In an attempt to better fit the observed data, the next model included distributional sub-models to allow for heterogeneous variances for each response conditional on the survey design structure.
```{r n2o_mod_2, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(log(n2o) ~ 1 +
               (1 | a | WSA9) + 
               (1 | b | WSA9:state) + 
               (1 | c | WSA9:state:size_cat),
             sigma ~ 1 +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat), 
             family = gaussian())

bf_n2oeq <- bf(log(n2o_eq) ~ 1 +
                 (1 | a | WSA9) + 
                 (1 | b | WSA9:state) +
                 (1 | c | WSA9:state:size_cat),
             sigma ~ 1 +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat),
             family = gaussian())

priors <- c(
  prior(normal(2, 1), class = "Intercept", resp = "logn2o"),
  prior(exponential(2), class = "sd", resp = "logn2o"),
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2o"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2o"),
  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), 
  prior(exponential(2), class = "sd", resp = "logn2oeq"),
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2oeq"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2oeq"),

  prior(lkj(2), class = "rescor"),
  prior(lkj(2), class = "cor")
  )

n2o_mod2 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
                data = df_model, 
                prior = priors,
  control = list(adapt_delta = 0.975, max_treedepth = 12),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 84512,
  chains=4, 
  iter=5000, 
  cores=4)

save(n2o_mod2, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
```

### Summarize fit
The summaries of the estimated parameters and key HMC convergence diagnostics for the fitted model are printed below. 
```{r print_mod2, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")

print(n2o_mod2, prior = T)
```
From the summary above, note the moderate and positive residual correlation between the two N2O responses. The estimated standard deviations for the varying group effects on the mean behavior of the dissolved N2O response suggested fairly low, but non-zero variability across each of the three levels. The standard deviations estimated for the same varying effects for equilibrium N2O were also relatively small. However, before investing too much into the interpretation of these results, the model fit was further evaluated below using a series of graphical posterior predictive checks (PPCs).

### Model checks
Below the same PPCs were performed as with the initial model (see above for more details on each panel). 
##### Dissolved N2O
Though the checks below suggest some improvement in replicating the tails of the observed data, this model did a poorer job at replicating central tendency.
```{r ppc_n2o2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
grid.arrange(
pp_check(n2o_mod2, 
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlim(-5, 5) +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod2, 
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlim(-5, 5) +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod2, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod2, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod2, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod2, 
         resp = "logn2o",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Equilibrium N2O
The checks below suggest this model offered no improvement upon the initial model for equilibrium N2O. This model also appeared to do a poorer job of replicating the mean and overall standard deviation compared to the initial model.
```{r ppc_n2oeq2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
grid.arrange(
pp_check(n2o_mod2, 
         resp = "logn2oeq",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
  ylab("density")
,
pp_check(n2o_mod2, 
         resp = "logn2oeq",
         type = "ecdf_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod2, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         ndraws = 1000,
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod2, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),
         ndraws = 1000,
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod2, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod2, 
         resp = "logn2oeq",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Bivariate
This check perhaps suggested an improvement with regard to replicating the joint density. However, the predictions were still clearly over-dispersed relative to the observations.
```{r ppc_biv2, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
df_model %>%
  add_predicted_draws(n2o_mod2, 
                      ndraws = 20) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

#### Saturation
The PPCs for the saturation metrics below indicated that including the distributional models was perhaps an improvement on the initial model in some aspects; in particular, the bias in the predicted proportion of under-saturated lakes was substantially decreased. However, there appeared to still be issues in replicating the tails as well as issues with central tendency.
```{r ppc_sat2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
grid.arrange(
df_model %>%
  add_predicted_draws(n2o_mod2, 
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod2, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(logn2o < logn2oeq, 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 984) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod2, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod2, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod2, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```

#### R-square
Relative to model 1, there was a substantial decrease in the $R^2$ estimate for the dissolved N2O component of this model. The estimate for the equilibrium N2O-eq component was similar to the model 1. 
```{r r2_2, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
round(bayes_R2(n2o_mod2, resp = "logn2o", cores = 1), 3) 
round(bayes_R2(n2o_mod2, resp = "logn2oeq", cores = 1), 3)
```

## Model 3
In the next model, we used covariates to try to improve the fit. The categorical version of the NO3 covariate was used as a monotonic ordinal predictor in the dissolved N2O component of the modl. For the equlibrium N2O component, we included surface temperature and log-transformed elevation, along with their interaction. The models also retained the distributional specifications included in model 2 above.
```{r n2o_mod_3, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
               surftemp +
               (mo(no3_cat) | a | WSA9) + 
               (mo(no3_cat) | b | WSA9:state) + 
               (1 | c | WSA9:state:size_cat),
             sigma ~ 1 +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat), 
             family = gaussian())

bf_n2oeq <- bf(log(n2o_eq) ~ surftemp +
                 log_elev +
                 surftemp:log_elev +
                 (1 | a | WSA9) + 
                 (1 | b | WSA9:state) +
                 (1 | c | WSA9:state:size_cat),
             sigma ~ 1 +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat),
             family = gaussian())

priors <- c(
  prior(normal(2, 1), class = "Intercept", resp = "logn2o"),
  prior(normal(0, 1), class = "b", resp = "logn2o"),
  prior(exponential(2), class = "sd", resp = "logn2o"),
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2o"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2o"),
  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), 
  prior(normal(0, 1), class = "b", resp = "logn2oeq"), 
  prior(exponential(2), class = "sd", resp = "logn2oeq"), 
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2oeq"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2oeq"),

  prior(lkj(2), class = "rescor"),
  prior(lkj(2), class = "cor")
  )

n2o_mod3 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
                data = df_model, 
                prior = priors,
  control = list(adapt_delta = 0.975, max_treedepth = 12),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 98456,
  chains=4, 
  iter=5000, 
  cores=4)

save(n2o_mod3, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
```

### Summarize fit
The fitted parameters and MCMC diagnostics are below.
```{r print_mod3, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")

print(n2o_mod3, prior = T)
```

### Model checks
#### Dissolved N2O
The PPCs below indicated a better fit compared to the previous models. The central tendency and tail behavior looked to be reasonably replicated by comparison. However, the observed _vs._ predicted plot suggested that larger overserved values were being systematically underestimated.
```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
grid.arrange(
pp_check(n2o_mod3, 
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlim(0, 5) +
  xlab(expression(paste("log(N"[2],"O)"))) + 
  ylab("density")
,
pp_check(n2o_mod3, 
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlim(0, 5) +
  xlab(expression(paste("log(N"[2],"O)"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod3, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         ndraws = 1000,
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod3, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),  
         ndraws = 1000,
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod3, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"),  
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod3, 
         resp = "logn2o",
         type = "scatter_avg",  
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Equilibrium N2O
The PPCs below indicated that this model appeared to be an improvement for equilibrium N2O as well. However, some checks (e.g., skewness) suggested some room for additional improvement.
```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
grid.arrange(
pp_check(n2o_mod3, 
         resp = "logn2oeq",
         type = "dens_overlay", 
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
  ylab("density")
,
pp_check(n2o_mod3, 
         resp = "logn2oeq",
         type = "ecdf_overlay", 
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod3, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"),  
         ndraws = 1000,
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod3, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000, 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod3, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"),  
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod3, 
         resp = "logn2oeq",
         type = "scatter_avg",  
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Bivariate
The check for the joint distribution below also suggested an improvement up the previous models.
```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
df_model %>%
  add_predicted_draws(n2o_mod3, 
                      ndraws = 20) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

#### Saturation
This model looked to be an improvement with regard to the PPCs for the saturation metrics. However, the proportion of under-saturated lakes remained biased low and other checks indicated that further improvements would be ideal.
```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
grid.arrange(
df_model %>%
  add_predicted_draws(n2o_mod3, 
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod3, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 984) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod3, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod3, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod3, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```

#### R-square
The $R^2$ estimates for this model are below and suggested substantial improvements on the previous models.
```{r r2_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
round(bayes_R2(n2o_mod3, resp = "logn2o", cores = 1), 3) 
round(bayes_R2(n2o_mod3, resp = "logn2oeq", cores = 1), 3)
```

### Covariate effects
Below are plots illustrating the modeled effects of covariates on both N2O and equilibrium N2O.
#### N2O
The conditional effects plots below for N2O illustrate a positive, monotonic, and non-linear relationship between NO3 and N2O; and a negative, linear relationship between surface temperature and N2O.
```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")

p1 <- conditional_effects(n2o_mod3, 
                          resp = "logn2o", 
                          effects = c("no3_cat"), 
                          plot = F)
p2 <- conditional_effects(n2o_mod3, 
                          resp = "logn2o", 
                          effects = c("surftemp"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]], 
          plot(p2, plot = F)[[1]],
          ncol = 2)

rm(p1, p2)

annotate_figure(plt, top = text_grob("N2O: conditional effects", 
               color = "black", face = "bold", size = 14))
```

#### Equilibrium N2O
The modeled effects below for the equilibrium N2O component of the model illustrated a negative relationship between equilibrium N2O and both predictors and an interaction such that the surface temperature effect became slightly steeper at lower elevations.
```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
p1 <- conditional_effects(n2o_mod3, 
                          resp = "logn2oeq", 
                          effects = c("log_elev:surftemp"), 
                          plot = F)
p2 <- conditional_effects(n2o_mod3, 
                          resp = "logn2oeq", 
                          effects = c("surftemp:log_elev"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, top = text_grob("Equilibrium N2O: conditional effects", 
               color = "black", face = "bold", size = 14))
```

## Model 4
In the next model, covariate terms were also included in the $\sigma$ components of both models in order to try to better capture remaining heterogeneity in the variances of both N2O and N2O-eq.
```{r n2o_mod_mv_4, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
               surftemp +
               (mo(no3_cat) | a | WSA9) + 
               (mo(no3_cat) | b | WSA9:state) + 
               (1 | c | WSA9:state:size_cat),
             sigma ~ mo(no3_cat) +
               surftemp +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat), 
             family = gaussian())

bf_n2oeq <- bf(log(n2o_eq) ~ surftemp +
                 log_elev +
                 surftemp:log_elev +
                 (1 | a | WSA9) + 
                 (1 | b | WSA9:state) +
                 (1 | c | WSA9:state:size_cat),
             sigma ~ surftemp +
               log_elev +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat),
             family = gaussian())

priors <- c(
  prior(normal(2, 1), class = "Intercept", resp = "logn2o"),
  prior(normal(0, 1), class = "b", resp = "logn2o"),
  prior(exponential(2), class = "sd", resp = "logn2o"),
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2o"),
  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2o"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2o"),
  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), 
  prior(normal(0, 1), class = "b", resp = "logn2oeq"), 
  prior(exponential(2), class = "sd", resp = "logn2oeq"), 
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2oeq"),
  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2oeq"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2oeq"),

  prior(lkj(2), class = "rescor"),
  prior(lkj(2), class = "cor")
  )

n2o_mod4 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
                data = df_model, 
                prior = priors,
  control = list(adapt_delta = 0.975, max_treedepth = 12),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 15851,
  chains=4, 
  iter=5000, 
  cores=4)

save(n2o_mod4, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
```

### Summarize fit
Below is a summary of the fitted parameters along with some convergence diagnostics.
```{r print_mod4, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")

print(n2o_mod4, prior = T, digits = 3)
```

### Model checks
Again, the same PPCs were employed for this model as above.
#### Dissolved N2O
Again, this model appeared to be an improvement on the previous model, particularly with regard to the more constant variance indicated in the observed _vs._ predicted plot (bottom, right panel).
```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
grid.arrange(
pp_check(n2o_mod4, 
         ndraws = 200,
         resp = "logn2o",
         type = "dens_overlay",
         nsamples = 40,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O)"))) + 
  ylab("density")
,
pp_check(n2o_mod4, 
         ndraws = 200,
         resp = "logn2o",
         type = "ecdf_overlay",
         nsamples = 40, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O)"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod4, 
         ndraws = 1000,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod4, 
         ndraws = 1000,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod4, 
         ndraws = 1000,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod4, 
         resp = "logn2o",
         type = "scatter_avg", 
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Equilibrium N2O
This component of the model also seemed to be an improvement over model 3, with better representation in the tails as indicated in the skewness _vs._ kurtosis PPC. 
```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
grid.arrange(
pp_check(n2o_mod4, 
         ndraws = 200,
         resp = "logn2oeq",
         type = "dens_overlay",
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
  ylab("density")
,
pp_check(n2o_mod4, 
         ndraws = 200,
         resp = "logn2oeq",
         type = "ecdf_overlay",
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod4, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod4, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod4, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod4, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "scatter_avg", 
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Bivariate
Again, an improvement over the previous model with a tighter fit of the PPC to the observed bivariate density.
```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
df_model %>%
  add_predicted_draws(n2o_mod4, 
                      ndraws = 20) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

#### Saturation
This check also suggested an improvement over the previous models, with better tail behavior and less bias in the proportion under-saturated measure.
```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
grid.arrange(
df_model %>%
  add_predicted_draws(n2o_mod4, 
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod4, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 984) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod4, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod4, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod4, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```

#### R-square
The Bayesian $R^2$ estimates below indicated an improvement from the previous models.
```{r r2_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
round(bayes_R2(n2o_mod4, resp = "logn2o", cores = 1), 3) 
round(bayes_R2(n2o_mod4, resp = "logn2oeq", cores = 1), 3)
```

### Covariate effects
#### N2O
The conditional effects plots for the covariate effects on N2O remained largely unchanged from the previous model.
```{r conditional_effects_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")

p1 <- conditional_effects(n2o_mod4, 
                          resp = "logn2o", 
                          effects = c("no3_cat"), 
                          plot = F)
p2 <- conditional_effects(n2o_mod4, 
                          resp = "logn2o", 
                          effects = c("surftemp"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]], 
          plot(p2, plot = F)[[1]],
          ncol = 2)

rm(p1, p2)

annotate_figure(plt, top = text_grob(expression(paste("N2O: covariate effects on ", mu)), 
               color = "black", face = "bold", size = 14))
```

Below are estimates of the conditional effects of the covariates on $\sigma$ for N2O. These plots suggested a large effect of NO3 on the variance of N2O, but little to no effect of surface temperature.
```{r conditional_effects_sigma_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")

p1 <- conditional_effects(n2o_mod4, 
                          resp = "logn2o",
                          dpar = "sigma",
                          effects = c("no3_cat"), 
                          plot = F)
p2 <- conditional_effects(n2o_mod4, 
                          resp = "logn2o",
                          dpar = "sigma",
                          effects = c("surftemp"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, 
                top = text_grob(expression(paste("N2O: covariate effects on ", sigma)),
                                color = "black",
                                face = "bold",
                                size = 14))
```

#### Equilibrium N2O
The covariate effects on N2O remained largely the same as for the previous model.
```{r conditional_effects_mod_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")

p1 <- conditional_effects(n2o_mod4, 
                          resp = "logn2oeq", 
                          effects = c("log_elev:surftemp"), 
                          plot = F)
p2 <- conditional_effects(n2o_mod4, 
                          resp = "logn2oeq", 
                          effects = c("surftemp:log_elev"), 
                          plot = F)
plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", mu)), 
               color = "black", face = "bold", size = 14))
```

The covariate effects on $\sigma$ for N2O-eq suggested an negative effect of surface temperature and litte to no effect of elevation.
```{r conditional_effects_sigma_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")

p1 <- conditional_effects(n2o_mod4, 
                          resp = "logn2oeq",
                          dpar = "sigma",
                          effects = c("surftemp"), 
                          plot = F)
p2 <- conditional_effects(n2o_mod4, 
                          resp = "logn2oeq",
                          dpar = "sigma",
                          effects = c("log_elev"), 
                          plot = F)
plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", sigma)), 
               color = "black", face = "bold", size = 14))
```

## Model 5
In the next model, more complexity is added to the N2O component by including a covariate for lake surface area (log scale) as well as interactions between NO3 and log(area) and surface temperature. 
```{r n2o_mod_mv_5, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
               log_area +
               surftemp + 
               mo(no3_cat):log_area +
               mo(no3_cat):surftemp +
               (mo(no3_cat) | a | WSA9) + 
               (mo(no3_cat) | b | WSA9:state) + 
               (1 | c | WSA9:state:size_cat),
             sigma ~ log_area +
               mo(no3_cat) +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat), 
             family = gaussian())

bf_n2oeq <- bf(log(n2o_eq) ~ surftemp +
                 log_elev +
                 surftemp:log_elev +
                 (1 | a | WSA9) + 
                 (1 | b | WSA9:state) +
                 (1 | c | WSA9:state:size_cat),
             sigma ~ surftemp +
               log_elev +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat),
             family = gaussian())

priors <- c(
  prior(normal(2, 1), class = "Intercept", resp = "logn2o"),
  prior(normal(0, 1), class = "b", resp = "logn2o"),
  prior(exponential(2), class = "sd", resp = "logn2o"),
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2o"),
  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2o"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2o"),

  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), 
  prior(normal(0, 1), class = "b", resp = "logn2oeq"), 
  prior(exponential(2), class = "sd", resp = "logn2oeq"), 
  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2oeq"),
  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2oeq"),
  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2oeq"),

  prior(lkj(2), class = "rescor"),
  prior(lkj(2), class = "cor")
  )

n2o_mod5 <- brm(bf_n2o + 
                  bf_n2oeq +
                  set_rescor(rescor = TRUE),
                data = df_model, 
                prior = priors,
  control = list(adapt_delta = 0.975, max_treedepth = 12),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 54741,
  chains=4, 
  iter=5000, 
  cores=4)

save(n2o_mod5, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
```

### Summarize fit
Below is a summary of the fitted parameters along with MCMC convergence diagnostics.
```{r print_mod5, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")

print(n2o_mod5, prior = T, digits = 3)
```

### Model checks
Again, the same PPCs as above were performed for this model.
#### N2O PPC
This PPC for N2O looked similar to the previous model.
```{r ppc_full_checks_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
grid.arrange(
pp_check(n2o_mod5, 
         ndraws = 200,
         resp = "logn2o",
         type = "dens_overlay",
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod5, 
         ndraws = 200,
         resp = "logn2o",
         type = "ecdf_overlay",
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2o",
         type = "scatter_avg", 
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Equilibrium N2O
Again, the PPCs for this model were similar to the previous model, which was unsurprising given that it was the same model for N2O-eq.
```{r ppc_full_checks_mod_n2oeq5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
grid.arrange(
pp_check(n2o_mod5, 
         ndraws = 200,
         resp = "logn2oeq",
         type = "dens_overlay",
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod5, 
         ndraws = 200,
         resp = "logn2oeq",
         type = "ecdf_overlay",
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod5, 
         ndraws = 1000,
         resp = "logn2oeq",
         type = "scatter_avg", 
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Bivariate
This PPC was also similar to the previous model.
```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
df_model %>%
  add_predicted_draws(n2o_mod5, 
                      ndraws = 20) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

#### Saturation
This check was also similar to the prevoius model, with perhaps slightly less bias in the proportion unsaturated estimates. There is also a potentially concerning extreme prediction in the observed _vs_ predicted PPC.
```{r ppc_sat_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
grid.arrange(
df_model %>%
  add_predicted_draws(n2o_mod5, 
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod5, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 984) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod5, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod5, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod5, 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```

#### R-square
```{r r2_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
round(bayes_R2(n2o_mod5, resp = "logn2o", cores = 1), 3) 
round(bayes_R2(n2o_mod5, resp = "logn2oeq", cores = 1), 3)
```

### Covariate effects
#### N2O
The conditional effects plot for the covariate effects N2O suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature. For lake area, the effect was estimated to be larger and more negative at the highest levels of NO3; and slightly negative at the lowest level of NO3. For surface temperature, the effect was estimated to be largest and positive at the highest level of NO3; and negative at the lowest level of NO3.
```{r conditional_effects_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")

p1 <- conditional_effects(n2o_mod5, 
                          resp = "logn2o", 
                          effects = c("no3_cat"), 
                          plot = F)

p2 <- conditional_effects(n2o_mod5, 
                          resp = "logn2o", 
                          effects = c("log_area:no3_cat"), 
                          plot = F)
p3 <- conditional_effects(n2o_mod5, 
                          resp = "logn2o", 
                          effects = c("surftemp:no3_cat"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]], 
                 plot(p2, plot = F)[[1]],
                 plot(p3, plot = F)[[1]])

rm(p1, p2, p3)

annotate_figure(plt, top = text_grob(expression(paste("N2O: covariate effects on ", mu)), 
               color = "black", face = "bold", size = 14))
```

The estimated covariate effects on $\sigma$ suggested a negative relationship with log(area) and a positive relationship, again, with NO3.
```{r conditional_effects_sigma_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")

p1 <- conditional_effects(n2o_mod5, 
                          resp = "logn2o",
                          dpar = "sigma",
                          effects = c("log_area"),
                          plot = F)

p2 <- conditional_effects(n2o_mod5, 
                          resp = "logn2o",
                          dpar = "sigma",
                          effects = c("no3_cat"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, 
                top = text_grob(expression(paste("N2O: covariate effects on ", sigma)),
                                color = "black",
                                face = "bold",
                                size = 14))
```

#### Equilibrium N2O
The estimated covariate effect on N2O remained largely the same as estimated in the previous model.
```{r conditional_effects_mod_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")

p1 <- conditional_effects(n2o_mod5, 
                          resp = "logn2oeq", 
                          effects = c("surftemp:log_elev"), 
                          plot = F)

p2 <- conditional_effects(n2o_mod5, 
                          resp = "logn2oeq", 
                          effects = c("log_elev:surftemp"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", mu)), 
               color = "black", face = "bold", size = 14))
```

```{r conditional_effects_sigma_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")

p1 <- conditional_effects(n2o_mod5, 
                          resp = "logn2oeq",
                          dpar = "sigma",
                          effects = c("surftemp"), 
                          plot = F)

p2 <- conditional_effects(n2o_mod5, 
                          resp = "logn2oeq",
                          dpar = "sigma",
                          effects = c("log_elev"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", sigma)), 
               color = "black", face = "bold", size = 14))
```

## A Final Model
As demonstrated above, models excluding the NO3 covariate consistently resulted in poorer fits to to the observed dissolved N2O data. Including surface temperature and elevation in the equilibrium N2O part of the model resulted in substantially improved replication of key aspects of the observed data. Likewise, added flexibility in the distributional terms for both dissolved and equilibrium N2O led to improvements.

To make inferences from this model for N2O in the population of interest, however, the included covariates needed to be (1) fully observed across that population or (2) their missingness needed to be modeled. For the lake area and elevation covariates, data _was_ available for all lakes from previously compiled geospatial databases. However, neither surface temperature or NO3 were observed for lakes outside of the sample. They were partially observed with respect to the target population. Their missingness needed to be accounted for in a model. Therefore, a more complex model was constructed below that included surface temperature and NO3 as additional responses conditioned on the survey design variables and fully observed covariates. This approach to inference for N2O was similar to a Bayesian structural equation model [@Merkle_etal_2021; @Merkle_Rosseel_2018]. The main details of the logical dependence structure could be characterized as:

$$\begin{align} 
\color{#1F449C}{\boldsymbol{N_2O_{diss}}} &=\sim Survey + Area + \color{#F05039}{\boldsymbol{NO_3}} + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
\color{#A8B6CC}{\boldsymbol{N_2O_{equil}}} &=\sim Survey + Elev  + \color{#EEBAB4}{\boldsymbol{Temp}}\\
\color{#F05039}{\boldsymbol{NO_3}} &=\sim Survey + Area + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
\color{#EEBAB4}{\boldsymbol{Temp}} &=\sim Survey + Lat + Elev + Day
\end{align}$$

Variables in color text above were treated as partially observed with respect to the population of interest (i.e., observed only in the sample), whereas variables in black text were considered fully observed. The partially observed variables, being dissolved and equilibrium N2O, NO3, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates. This structural equation approach requires a more complex set of post-processing steps compared to a typical MRP analysis. In order to propagate estimates and uncertainty through the dependency structure and make inferences, the fitted model was used to first predict surface temperature in the target population, since it depended only on the fully observed covariates. That predictive distribution was then used alongside the relevant fully observed covariates to predict NO3 in the target population. Finally, the predictive distributions for termperature and NO3 were used to predict the N2O responses. These steps were carried out in the "Predict to population" section to follow.

In the final model below, the submodel for surface temperature assumed a Gamma distributed error distribution and the linear predictor included the survey design variables, latitude, elevation, and julian date. The shape parameter was also modeled as a function of latitude to address increasing response variance along the latitudinal gradient. The NO3 submodel was a cumulative logit formulation and the linear predictor included all of the survey factors as well as surface temperature and lake area.

The N2O and N2O-eq responses were each modeled with Gamma distributed errors, but with the same covariate structure as in model 5. The same structure was also employed for the shape terms in these responses, corresponding to the $\sigma$ terms in the previous model. Though not shown in this document, the Gamma error structure appeared to result in slightly better performance in the predictive checks compared to the Gaussian errors in previous models. This was primarily apparent in the saturation ratio checks, which may have been more sensitive to model performance in the tails of the N2O responses. Others have also indicated that the Gamma error distribution can work well for dissolved N2O data [@Webb_etal_2019].

Note that there was no residual correlation term for this model, since the residuals are undefined for the Gamma and cumulative logit models. Dropping the observation-level residual correlation term was deemed a reasonable compromise that enabled modeling the missingness of NO3, in particular. Nevertheless, the random intercepts again allowed for potential correlations between responses at the group levels.

```{r n2o_mod_mv_6, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")

bf_n2o <- bf(n2o ~ mo(no3_cat) +
               log_area +
               surftemp + 
               mo(no3_cat):log_area +
               mo(no3_cat):surftemp +
               (mo(no3_cat) | a | WSA9) + 
               (mo(no3_cat) | b | WSA9:state) + 
               (1 | c | WSA9:state:size_cat),
             shape ~ log_area +
               mo(no3_cat) +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat),
             family = Gamma(link = "log"))

bf_n2oeq <- bf(n2o_eq ~ surftemp +
                 log_elev +
                 surftemp:log_elev +
                 (1 | a | WSA9) + 
                 (1 | b | WSA9:state) +
                 (1 | c | WSA9:state:size_cat),
             shape ~ surftemp +
               log_elev +
               (1 | WSA9) + 
               (1 | WSA9:state) + 
               (1 | WSA9:state:size_cat),
             family = Gamma(link = "log"))

bf_temp <- bf(surftemp ~ lat +
                s(log_elev) +
                s(jdate) +
                (1 | a | WSA9) + 
                (1 | b | WSA9:state) +
                (1 | c | WSA9:state:size_cat),
              shape ~ lat,
              family = Gamma(link = "log"))

bf_no3 <- bf(no3_cat ~ surftemp +
               log_area +
               (1 | a | WSA9) +
               (1 | b | WSA9:state) +
               (1 | c | WSA9:state:size_cat),
             family = cumulative(link = "logit", threshold="flexible"))

priors <- c(
  prior(normal(2, 1), class = "Intercept", resp = "n2o"),
  prior(normal(0, 1), class = "b", resp = "n2o"),
  prior(exponential(2), class = "sd", resp = "n2o"),
  prior(normal(5, 4), class = "Intercept", dpar = "shape", resp = "n2o"),
  prior(normal(0, 1), class = "b", dpar = "shape", resp = "n2o"),
  prior(exponential(2), class = "sd", dpar = "shape", resp = "n2o"),

  prior(normal(2, 1), class = "Intercept", resp = "n2oeq"), 
  prior(normal(0, 1), class = "b", resp = "n2oeq"),  
  prior(exponential(2), class = "sd", resp = "n2oeq"),
  prior(normal(5, 4), class = "Intercept", dpar = "shape", resp = "n2oeq"),
  prior(normal(0, 1), class = "b", dpar = "shape", resp = "n2oeq"),
  prior(exponential(2), class = "sd", dpar = "shape", resp = "n2oeq"),

  prior(normal(3, 1), class = "Intercept", resp = "surftemp"), 
  prior(normal(0, 1), class = "b", resp = "surftemp"), 
  prior(exponential(0.5), class = "sds", resp = "surftemp"),
  prior(exponential(2), class = "sd", resp = "surftemp"),
  prior(normal(5, 4), class = "Intercept", dpar = "shape", resp = "surftemp"),
  prior(normal(0, 1), class = "b", dpar = "shape", resp = "surftemp"),

  prior(normal(0, 3), class = "Intercept", resp = "no3cat"),
  prior(normal(0, 1), class = "b", resp = "no3cat"),
  prior(exponential(1), class = "sd", resp = "no3cat"),

  prior(lkj(2), class = "cor")
  )

n2o_mod6 <- brm(bf_n2o + 
                  bf_n2oeq + 
                  bf_temp + 
                  bf_no3 + 
                  set_rescor(rescor = FALSE),
                data = df_model, 
                prior = priors,
  control = list(adapt_delta = 0.975, max_treedepth = 14),
  #sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  seed = 85132,#14548,
  #init = my_inits,
  init_r = 0.5,
  chains=4, 
  iter=5000, 
  cores=4)

save(n2o_mod6, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
```

### Summarize fit
Below is a summary of the fitted parameters and MCMC diagnostics. 
```{r print_mod6, echo=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")

print(n2o_mod6, digits = 3, prior = T)
```

### Model checks
Below, the same PPCs for N2O and N2O-eq were employed as before.
#### N2O PPC
The PPCs for N2O from this model were similarly reasonable as for models 4 and 5 above.
```{r ppc_full_checks_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
grid.arrange(
pp_check(n2o_mod6, 
         resp = "n2o",
         type = "dens_overlay",
         nsamples = 40,
         cores = 1) + 
  theme_tidybayes() +
  scale_x_continuous(trans = "log") +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod6, 
         resp = "n2o",
         type = "ecdf_overlay",
         nsamples = 40, 
         cores = 1) + 
  theme_tidybayes() +
  scale_x_continuous(trans = "log") +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod6, 
         resp = "n2o",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod6, 
         resp = "n2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod6, 
         resp = "n2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod6, 
         resp = "n2o",
         type = "scatter_avg", 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
, ncol = 2)
```

#### Equilibrium N2O PPC
Again, the PPCs for N2O-eq in this model were similar to those for models 4 and 5.
```{r ppc_full_checks_mod_n2oeq6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
grid.arrange(
pp_check(n2o_mod6, 
         resp = "n2oeq",
         type = "dens_overlay",
         nsamples = 40,
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod6, 
         resp = "n2oeq",
         type = "ecdf_overlay",
         nsamples = 40, 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  theme_tidybayes() +
  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod6, 
         resp = "n2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod6, 
         resp = "n2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod6, 
         resp = "n2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
,
pp_check(n2o_mod6, 
         resp = "n2oeq",
         type = "scatter_avg", 
         cores = 1) + 
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  theme_tidybayes()
, ncol = 2)
```

#### Bivariate PPC
This model again provided a very reasonable representation of the bivariate relationship between N2O and N2O-eq (below).
```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
df_model %>%
  add_predicted_draws(n2o_mod6,
                      resp = c("n2o","n2oeq"),
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
  geom_density_2d(aes(x = log(n2o), 
                      y = log(n2oeq), 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

#### Saturation PPC
The saturation ratio PPCs below show similar behavior as with models 4 and 5 above, but with perhaps slightly less bias in the predictions for the proportion of undersaturated waterbodies and fewer extreme predictions for the means and standard deviations. The observed _vs._ predicted PPC also appears to have a better behaved variance and no extreme predictions, compared to models 4 and 5 with the lognormal errors.
```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
grid.arrange(
df_model %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"),
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(n2o < n2oeq, 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 984) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  theme_tidybayes()
,
df_model %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```

The plot below shows the same PPC, but for the "test" or second-vist data. Overall, the model looked to perform similarly as with the data used to fit it.
```{r ppc_sat_check_testdata_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
grid.arrange(
df_test %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"),
                      ndraws = 50) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  ggplot(aes(x = sat_ratio)) +
  geom_density(aes(x = sat_pred, group = .draw), 
               n = 1024, 
               adjust = 1,
               color = "lightblue") +
  geom_density(n = 1024, adjust = 2) +
  xlim(0, 5) +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(usat_pred = ifelse(n2o < n2oeq, 1, 0)) %>%
  group_by(.draw) %>%
  summarise(prop_pred = sum(usat_pred) / 95) %>%
  ggplot(aes(x = prop_pred)) +
  geom_histogram(binwidth = 0.001, fill = "lightblue") +
  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
  xlab("Proportion undersaturated waterbodies") +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  group_by(.draw) %>%
  mutate(mean_yrep = mean(sat_pred),
         sd_yrep = sd(sat_pred)) %>% 
  ungroup() %>%
  mutate(mean_y = mean(sat_ratio),
         sd_y = sd(sat_ratio)) %>%
  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  group_by(.draw) %>%
  mutate(min_yrep = min(sat_pred),
         max_yrep = max(sat_pred)) %>% 
  ungroup() %>%
  mutate(min_y = min(sat_ratio),
         max_y = max(sat_ratio)) %>%
  ggplot(aes(x = min_yrep, y = max_yrep)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
  theme_tidybayes()
,
df_test %>%
  add_predicted_draws(n2o_mod6, 
                      resp = c("n2o", "n2oeq"), 
                      ndraws = 500) %>%
  rename(n2o_obs = n2o,
         n2oeq_obs = n2o_eq) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  mutate(sat_ratio = n2o_obs / n2oeq_obs,
         sat_pred = n2o / n2oeq) %>%
  group_by(.row) %>%
  mutate(mean_yrep = mean(sat_pred)) %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
  geom_point(color = "lightblue") +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
, ncol = 2)
```

#### R-square
Below are estimates for the Bayesian $R^2$, which were largely similar for N2O and N2O-eq as with models 4 and 5 above. The $R^2$ for the surface temperature response also suggested a fairly good fit.
```{r r2_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
round(bayes_R2(n2o_mod6, resp = "n2o"), 3) 
round(bayes_R2(n2o_mod6, resp = "n2oeq"), 3)
round(bayes_R2(n2o_mod6, resp = "surftemp"), 3)
```

Below are the same $R^2$ estimates, but for the second-visit data. That these estimates are similar to those for the data used to fit the model, suggesting that the model may perform similarly well out-of-sample.
```{r r2_test_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
round(bayes_R2(n2o_mod6, resp = "n2o", newdata = df_test), 3) 
round(bayes_R2(n2o_mod6, resp = "n2oeq", newdata = df_test), 3)
round(bayes_R2(n2o_mod6, resp = "surftemp", newdata = df_test), 3)
```

### Covariate effects
#### N2O
The conditional effects plot for the covariate effects N2O suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature. For lake area, the effect was estimated to be larger and more negative at the highest levels of NO3; and slightly negative at the lowest level of NO3. For surface temperature, the effect was estimated to be largest and positive at the highest level of NO3; and negative at the lowest level of NO3.
```{r conditional_effects_mod_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")

p1 <- conditional_effects(n2o_mod6, 
                          resp = "n2o", 
                          effects = c("no3_cat"), 
                          plot = F)

p2 <- conditional_effects(n2o_mod6, 
                          resp = "n2o", 
                          effects = c("log_area:no3_cat"), 
                          plot = F)
p3 <- conditional_effects(n2o_mod6, 
                          resp = "n2o", 
                          effects = c("surftemp:no3_cat"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]], 
                 plot(p2, plot = F)[[1]],
                 plot(p3, plot = F)[[1]])

rm(p1, p2, p3)

annotate_figure(plt, top = text_grob(expression(paste("N2O: covariate effects on ", mu)), 
               color = "black", face = "bold", size = 14))
```

The estimated covariate effects on $\sigma$ suggested a negative relationship with log(area) and a positive relationship, again, with NO3.
```{r conditional_effects_sigma_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")

p1 <- conditional_effects(n2o_mod6, 
                          resp = "n2o",
                          dpar = "shape",
                          effects = c("log_area"),
                          plot = F)

p2 <- conditional_effects(n2o_mod6, 
                          resp = "n2o",
                          dpar = "shape",
                          effects = c("no3_cat"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, 
                top = text_grob(expression(paste("N2O: covariate effects on ", shape)),
                                color = "black",
                                face = "bold",
                                size = 14))
```

#### Equilibrium N2O
The estimated covariate effect on N2O remained largely the same as estimated in the previous model.
```{r conditional_effects_mod_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")

p1 <- conditional_effects(n2o_mod6, 
                          resp = "n2oeq", 
                          effects = c("surftemp:log_elev"), 
                          plot = F)

p2 <- conditional_effects(n2o_mod6, 
                          resp = "n2oeq", 
                          effects = c("log_elev:surftemp"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", mu)), 
               color = "black", face = "bold", size = 14))
```

```{r conditional_effects_sigma_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")

p1 <- conditional_effects(n2o_mod6, 
                          resp = "n2oeq",
                          dpar = "shape",
                          effects = c("surftemp"), 
                          plot = F)

p2 <- conditional_effects(n2o_mod6, 
                          resp = "n2oeq",
                          dpar = "shape",
                          effects = c("log_elev"), 
                          plot = F)

plt <- ggarrange(plot(p1, plot = F)[[1]],
                 plot(p2, plot = F)[[1]])

rm(p1, p2)

annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", shape)), 
               color = "black", face = "bold", size = 14))
```

# Predict to population
As previously described, in order to make inferences to the population of interest, the final model above was used to, first, predict surface temperature in the target population, since it depended only on the fully observed covariates. Next, the predictive distribution for surface temperature was used, along with the relevant fully observed covariates, to predict NO3 in the target population. Finally, the predictive distributions for temperature and NO3 were used to predict the N2O responses. The code for these steps is outlined in the following.

The first step used the final model to predict to the population:
```{r predict_obsframe, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")

predict_temp <- sframe %>%
  mutate(jdate = 205) %>%
  add_predicted_draws(n2o_mod6, resp=c("surftemp"), 
                      allow_new_levels = TRUE, 
                      cores =1, 
                      ndraws = 500) %>%
  mutate(surftemp = .prediction)

save(predict_temp, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")
```

NO3 was next predicted. Note that the posterior predictive distribution for NO3 was subsampled in order to minimize excess simulations
```{r parallel_predict_draws, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")

temp_X <- predict_temp %>% # select relevant columns as predictors
  ungroup() %>%
  select(WSA9,
         state,
         size_cat,
         log_area,
         .row,
         .draw,
         surftemp) %>%
  select(WSA9, state, size_cat, log_area, surftemp)

rm(predict_temp) # reduce memory
gc()

# set number of cores to use for parallel predictions
# and register the workers
cl <- parallel::makeCluster(5) 
doSNOW::registerDoSNOW(cl)

# make a progress bar
pb <- txtProgressBar(max = 1500, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

system.time( # approx 26 hrs with 5 workers & 500 draws from PPD
predict_no3 <- foreach(sub_X = isplitRows(temp_X, chunkSize = 155299), 
                       .combine = 'c',
                       .packages = c("brms"),
                       .options.snow = opts
                       ) %dopar% {
                         apply(brms::posterior_predict(n2o_mod6,
                                                 newdata = sub_X,
                                                 resp = "no3cat",
                                                 allow_new_levels = T,
                                                 ndraws = 500,
                                                 cores = 1), 2, sample, 1)
                         }
)

close(pb)
parallel::stopCluster(cl)

save(predict_no3, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
```

Finally, N2O and N2O-eq were predicted using the surface temperature and nitrate predictions along with the survey variables and known covariates. Again, the posterior was subsampled in order to reduce excess simulations.
```{r n2o_covariates_X, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")

# Assemble dataframe containing relevant covariates (known and predicted)
n2o_X <- predict_temp %>%
  ungroup() %>%
  mutate(no3_cat = predict_no3) %>%
  select(WSA9,
         state,
         size_cat,
         log_area,
         surftemp,
         log_elev,
         no3_cat)

# clear objects to reduce memory overhead
rm(predict_no3, predict_temp) 
gc()

# save the predictors for n2o and n2oeq
save(n2o_X, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_X.rda")
```

```{r parallel_predict, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_X.rda")

# set number of cores to use for parallel predictions
# and register the workers
cl <- parallel::makeCluster(6) 
doSNOW::registerDoSNOW(cl)

# make a progress bar
pb <- txtProgressBar(max = 1500, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

# make predictions in parallel
system.time(
predict_n2o <- foreach(sub_X = isplitRows(n2o_X, chunkSize = 155299),
                 .combine = rbind,
                 .options.snow = opts,
                 .packages = c("brms")) %dopar% {
  apply(posterior_predict(n2o_mod6,
                          newdata = sub_X,
                          resp = c("n2o", "n2oeq"),
                          allow_new_levels = T,
                          ndraws = 500,
                          cores = 1),
        2, sample, 1)
                   }
)

close(pb)
parallel::stopCluster(cl)

colnames(predict_n2o) <- c("n2o", "n2oeq")

save(predict_n2o, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_n2o.rda")
```

Finally, the predictions for all four partially observed responses were assembled into a new dataframe for use in inference.
```{r assemble_predictions, eval=FALSE, include=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_n2o.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")

all_predictions <- predict_temp %>%
  ungroup() %>%
  mutate(no3cat = predict_no3) %>%
  bind_cols(predict_n2o) %>%
  mutate(n2osat = n2o / n2oeq, # calculate saturation ratio
         .row = rep(1:465897, each = 500),
         .draw = rep(seq(1,500, 1), 465897)) %>%
  mutate(area_ha = exp(log_area)) %>% # include area on ha scale
  select(WSA9,
         state,
         size_cat,
         area_ha,
         lat,
         lon,
         .row,
         .draw,
         surftemp,
         no3cat,
         n2o,
         n2oeq,
         n2osat)

rm(predict_n2o, predict_temp, predict_no3) # clean up workspace for RAM
gc()

save(all_predictions, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
```

# Population estimates
A number of estimates for the target population were assembled and presented below. First, the full posterior predictive distributions for dissolved N2O, equilibrium N2o, and the saturation ratio were assessed. These distributions summarized the predicted distribution of concentrations or ratios for all lakes in the population of interest and included parameter uncertainty propagated through the model. Next, population means were assessed, followed by comparisons of some model-based estimates to previously calculated design-based estimates.

## Posterior predictive distributions
Below, a density plot summarized the posterior predictive distribution of N2O and N2O-eq concentrations across the target population of lakes, based on 500 draws from the posterior predictive distribution. Note that the x-axis was truncated at 50 nmol/L for a clearer visualization of the bulk of the predicted distribution. For reference, the max predicted value was 4403.2 nmol/L for dissolved N2O, 20.4 nmol/L for dissolved N2O, and 793.5 for the saturation ratio. 
```{r plot_n2o_posterior_preds, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>% 
  slice_sample(n=1e4) %>% # simple random sample 10k lakes
  ggplot(aes(x = n2o)) +
  stat_dist_slabinterval(.width = c(0.5, 0.95)) +
  scale_y_log10() +
  xlim(0, 50) +
  xlab("Dissolved N2O concentration") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_n2oeq_posterior_preds, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>% 
  slice_sample(n=1e4) %>% # simple random sample 10k lakes
  ggplot(aes(x = n2o)) +
  stat_dist_slabinterval(.width = c(0.5, 0.95)) +
  scale_y_log10() +
  xlim(0, 50) +
  xlab("Equilibrium N2O concentration") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_sat_posterior_preds, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>% 
  slice_sample(n=1e4) %>% # simple random sample 10k lakes
  ggplot(aes(x = n2osat)) +
  stat_dist_slabinterval(.width = c(0.5, 0.95)) +
  scale_y_log10() +
  xlim(0, 8) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  xlab("N2O saturation ratio") +
  ylab("density") +
  theme_tidybayes()
```

## Estimated means
### National
Below are density plots summarizing the posterior distribution of _means_ for N2O concentrations and the saturation ratio for the target population (i.e., all US lakes > 1ha in the lower 48 states).
```{r plot_n2o_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>% 
  summarise(mean_n2o = mean(n2o)) %>%
  ggplot(aes(x = mean_n2o)) + 
  stat_dist_slabinterval(.width = c(0.5, 0.95)) +
  xlab("mean dissolved N2O") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_n2oeq_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>% 
  summarise(mean_n2o = mean(n2oeq)) %>%
  ggplot(aes(x = mean_n2o)) + 
  stat_dist_slabinterval(.width = c(0.5, 0.95)) +
  xlab("mean equlilibrium N2O") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_sat_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>% 
  summarise(mean_sat = mean(n2o/n2oeq)) %>%
  ggplot(aes(x = mean_sat)) + 
  stat_dist_slabinterval(.width = c(0.5, 0.95)) +
  xlab("mean N2O saturation ratio") +
  ylab("density") +
  theme_tidybayes()
```

To illustrate the skewness in the predictive distribution for the saturation ratio, an estimate for the median ratio is shown below. The entire posterior distribution of the mean above is larger than 1, the ratio representing the boundary of under- _vs._ oversaturation. By comparison, the posterior estimate of the median below only included values less than one, suggesting that though the mean saturation ratio was greater than 1, most lakes in the national populaiton were undersaturated (i.e., ratio less than 1). In distributions with right-skew, the mean can often be considerably larger than the median.
```{r plot_sat_nat_posterior_median, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>% 
  summarise(median_sat = median(n2o/n2oeq)) %>%
  ggplot(aes(x = median_sat)) + 
  stat_dist_slabinterval(.width = c(0.5, 0.95)) +
  xlab("median N2O saturation ratio") +
  ylab("density") +
  theme_tidybayes()
```

Below is a plot of the posterior mean estimate for the proportion of unsaturated lakes at the national scale.
```{r plot_undersat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>%
  summarise(prop_sat = sum(n2osat < 1) / length(unique(.row))) %>%
  ggplot(aes(x = prop_sat)) + 
  stat_slabinterval(.width = c(0.5, 0.95)) +
  xlab("Proportion of undersaturated waterbodies") +
  ylab("density") +
  theme_tidybayes()
```

### Ecoregion
Below are posterior estimates of the means for dissolved and equilibrium N2O and the saturation ratio by WSA9 ecoregion.
```{r plot_n2o_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  mutate(ecoregion = WSA9) %>%
  ggplot() + 
  stat_slabinterval(aes(x = mean_n2o, 
                        y = reorder(ecoregion, mean_n2o)), 
                    quantiles = 100,
                    .width = c(0.5, 0.95)) +
  xlab("mean dissolved N2O") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_n2oeq_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2oeq = mean(n2oeq), .groups = "drop") %>%
  mutate(ecoregion = WSA9) %>%
  ggplot() + 
  stat_slabinterval(aes(x = mean_n2oeq, 
                        y = reorder(ecoregion, mean_n2oeq)), 
                    .width = c(0.5, 0.95)) +
  xlab("mean equilibrium N2O") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_sat_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_sat = mean(n2o/n2oeq), .groups = "drop") %>%
  mutate(ecoregion = WSA9) %>%
  ggplot() + 
  stat_slabinterval(aes(x = mean_sat,
                        y = reorder(ecoregion, mean_sat)),
                    .width = c(0.5, 0.95)) +
  xlab("mean N2O saturation ratio") +
  ylab("density") +
  theme_tidybayes()
```

A plot of the posterior estimates for the median saturation ratio below indicated, again, that most lakes in each ecoregion were undersaturated (i.e., median << 1).
```{r plot_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(median_sat = median(n2o/n2oeq), .groups = "drop") %>%
  mutate(ecoregion = WSA9) %>%
  ggplot() + 
  stat_slabinterval(aes(x = median_sat,
                        y = reorder(ecoregion, median_sat)),
                    .width = c(0.5, 0.95)) +
  xlab("median N2O saturation ratio") +
  ylab("density") +
  theme_tidybayes()
```

A plot of the estimates of the proportion of under-saturated lakes by ecoregion is below.
A plot of the posterior estimates for the median saturation ratio below indicated, again, that most lakes in each ecoregion were undersaturated (i.e., median << 1).
```{r plot_prop_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(n2osat < 1) / length(unique(.row)), .groups = "drop") %>%
  mutate(ecoregion = WSA9) %>%
  ggplot() + 
  stat_slabinterval(aes(x = prop_sat,
                        y = reorder(ecoregion, prop_sat)),
                    .width = c(0.5, 0.95)) +
  xlab("Proportion of undersaturated lakes") +
  ylab("density") +
  theme_tidybayes()
```

### State
Comparisons of mean estimates (posterior median, upper and lower 95th percentiles) by state are below. Density estimates were not included to minimize plot space.
```{r plot_state_mean_n2o, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(state, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(state) %>%
  summarise(estimate = round(median(mean_n2o), 1),
    LCL = round(quantile(mean_n2o, probs = 0.025), 1),
    UCL = round(quantile(mean_n2o, probs = 0.975), 1),
    .groups = "drop") %>% 
  select(state, estimate, LCL, UCL) %>%
  mutate(state = forcats::fct_reorder(state, estimate)) %>%
  ggplot(aes(x = state, y = estimate )) +
  geom_point(position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL),
                 position=position_dodge(width=0.5)) +
  ylab("mean dissolved N2O") +
  scale_y_continuous(position = "left") + 
  theme_tidybayes() +
  theme(axis.text.x = element_text(size=9, angle=45))
```

```{r plot_state_mean_n2oeq, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(state, .draw) %>%
  summarise(mean_n2oeq = mean(n2oeq), .groups = "drop") %>%
  group_by(state) %>%
  summarise(estimate = round(median(mean_n2oeq), 1),
    LCL = round(quantile(mean_n2oeq, probs = 0.025), 1),
    UCL = round(quantile(mean_n2oeq, probs = 0.975), 1),
    .groups = "drop") %>% 
  select(state, estimate, LCL, UCL) %>%
  mutate(state = forcats::fct_reorder(state, estimate)) %>%
  ggplot(aes(x = state, y = estimate )) +
  geom_point(position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL),
                 position=position_dodge(width=0.5)) +
  ylab("mean equilibrium N2O") +
  scale_y_continuous(position = "left") +
  theme_tidybayes() +
  theme(axis.text.x = element_text(size=9, angle=45))
```

Below, a plot of estimates for the mean (black circles) and median (grey circles) saturation ratio by state. A horizontal, dashed, black line is shown at ratio = 1, indicating the boundary for under- _vs._ oversaturation. Only a few states (e.g., NV, DE) had median estimates that were 1 or greater, suggesting that, for most states, most lakes were undersaturated.
```{r plot_state_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(state, .draw) %>%
  summarise(mean_sat = mean(n2osat), 
            median_sat = median(n2osat),
            .groups = "drop") %>%
  group_by(state) %>%
  summarise(estimate_mean = round(median(mean_sat), 4),
    LCL_mean = round(quantile(mean_sat, probs = 0.025), 4),
    UCL_mean = round(quantile(mean_sat, probs = 0.975), 4),
    estimate_median = round(median(median_sat), 4),
    LCL_median = round(quantile(median_sat, probs = 0.025), 4),
    UCL_median = round(quantile(median_sat, probs = 0.975), 4),
    .groups = "drop") %>% 
  select(state, 
         estimate_mean, 
         estimate_median, 
         LCL_mean,
         LCL_median,
         UCL_mean,
         UCL_median) %>%
  mutate(state = forcats::fct_reorder(state, estimate_mean)) %>%
  ggplot(aes(x = state, y = estimate_mean )) +
  geom_point(position=position_dodge(width=0.5),
             size = 2) +
  geom_linerange(aes(ymin = LCL_mean, ymax = UCL_mean),
                 position=position_dodge(width=0.5)) +
  geom_point(aes(x = state, y = estimate_median), 
             position=position_dodge(width=0.5),
             color = "grey",
             size = 2) +
  geom_linerange(aes(ymin = LCL_median, ymax = UCL_median),
                 position=position_dodge(width=0.5),
                 color = "gray") +
  ylab("mean and median N2O saturation ratio") +
  scale_y_continuous(position = "left") +
  geom_hline(yintercept = 1, color = "black", linetype = "dashed") +
  theme_tidybayes() +
  theme(axis.text.x = element_text(size=9, angle=45))
```

Finally, a plot of the estimated proportion of undersaturated lakes for each state in the target population. Point estimates are the posterior median of the proportion and bars are the upper and lower boundaries of the central 95th percentile of the posterior distributions of proportions.
```{r plot_state_prop_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(state, .draw) %>%
  summarise(prop_sat = sum(n2osat < 1) / length(unique(.row)),
            .groups = "drop") %>%
  group_by(state) %>%
  summarise(estimate = round(median(prop_sat), 4),
    LCL = round(quantile(prop_sat, probs = 0.025), 4),
    UCL = round(quantile(prop_sat, probs = 0.975), 4),
    .groups = "drop") %>% 
  select(state, 
         estimate, 
         LCL,
         UCL) %>%
  mutate(state = forcats::fct_reorder(state, estimate)) %>%
  ggplot(aes(x = state, y = estimate)) +
  geom_point(position=position_dodge(width=0.5),
             size = 2) +
  geom_linerange(aes(ymin = LCL, ymax = UCL),
                 position=position_dodge(width=0.5)) +
  ylab("Proportion of undersaturated lakes") +
  scale_y_continuous(position = "left") +
  theme_tidybayes() +
  theme(axis.text.x = element_text(size=9, angle=45))
```

### Size category
The estimated means by size category are below for dissolved and equilibrium N2O and the saturation ratio. Median estimates for the saturation ratio are also shown.
```{r plot_n2o_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2o),
            .groups = "drop") %>%
  ggplot(aes(x = mean_n2o, y = size_cat)) + 
  stat_slabinterval(.width = c(0.5, 0.95)) +
  xlab("mean dissolved N2O") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_n2oeq_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2oeq),
            .groups = "drop") %>%
  ggplot(aes(x = mean_n2o, y = size_cat)) + 
  stat_slabinterval(.width = c(0.5, 0.95)) +
  xlab("mean equilibrium N2O") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_sat_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2osat),
            .groups = "drop") %>%
  ggplot(aes(x = mean_n2o, y = size_cat)) + 
  stat_slabinterval(.width = c(0.5, 0.95)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  xlab("mean N2O saturation ratio") +
  ylab("density") +
  theme_tidybayes()
```

```{r plot_sat_size_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = median(n2osat),
            .groups = "drop") %>%
  ggplot(aes(x = mean_n2o, y = size_cat)) + 
  stat_slabinterval(.width = c(0.5, 0.95)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  xlab("median N2O saturation ratio") +
  ylab("density") +
  theme_tidybayes()
```

Mean _vs._ median below.
```{r plot_size_cat_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_sat = mean(n2osat), 
            median_sat = median(n2osat),
            .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate_mean = round(median(mean_sat), 2),
    LCL_mean = round(quantile(mean_sat, probs = 0.025), 2),
    UCL_mean = round(quantile(mean_sat, probs = 0.975), 2),
    estimate_median = round(median(median_sat), 2),
    LCL_median = round(quantile(median_sat, probs = 0.025), 2),
    UCL_median = round(quantile(median_sat, probs = 0.975), 2),
    .groups = "drop") %>% 
  select(size_cat, 
         estimate_mean, 
         estimate_median, 
         LCL_mean,
         LCL_median,
         UCL_mean,
         UCL_median) %>%
  mutate(state = forcats::fct_reorder(size_cat, estimate_mean)) %>%
  ggplot(aes(x = size_cat, y = estimate_mean )) +
  geom_point(position=position_dodge(width=0.5),
             size = 2) +
  geom_linerange(aes(ymin = LCL_mean, ymax = UCL_mean),
                 position=position_dodge(width=0.5)) +
  geom_point(aes(x = size_cat, y = estimate_median), 
             position=position_dodge(width=0.5),
             color = "grey",
             size = 2) +
  geom_linerange(aes(ymin = LCL_median, ymax = UCL_median),
                 position=position_dodge(width=0.5),
                 color = "gray") +
  ylab("mean and median N2O saturation ratio") +
  scale_y_continuous(position = "left") +
  geom_hline(yintercept = 1, color = "black", linetype = "dashed") +
  theme_tidybayes() +
  theme(axis.text.x = element_text(size=9, angle=45))
```

And, finally, the estimated proportion of undersaturated lakes in the target population by size category
```{r plot_prop_sat_size_cat_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(prop_sat = sum(n2osat < 1) / length(unique(.row)), .groups = "drop") %>%
  ggplot() + 
  stat_slabinterval(aes(x = prop_sat,
                        y = reorder(size_cat, prop_sat)),
                    .width = c(0.5, 0.95)) +
  xlab("Proportion of undersaturated lakes") +
  ylab("density") +
  theme_tidybayes()
```

## Model- _vs._ design-based
Below, estimates from the model-based approach are compared to design-based estimates. In general, the model estimates were similar to the design-based estimates. Model estimates were typically within the confidence bounds of the design-based estimates, but with much greater precision. Improved precision was expected due to the "shrinkage" induced by the multilevel parameterization, which allowed some "borrowing" of information across the various levels of the survey factors.

### Dissolved N2O
Below, National mean estimates for dissolved N2O from the model and design-based approaches were compared. The sample-based estimate was also included as a naive reference.
```{r n2o_means_national, message=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  summarise(estimate = round(median(mean_n2o), 2), # posterior median
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2)) %>% 
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
  add_row(estimate = round(mean(df_model$n2o), 2),
          type = "sample") %>%
  print()
```

The black, vertical, dashed line in the figure below represents the mean of the sample. 
```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")

all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  summarise(estimate = round(median(mean_n2o), 2), # posterior median
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2)) %>% 
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
  mutate(cl_width = round(UCL - LCL, 2)) %>%
  ggplot(aes(x = type, y = estimate, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  geom_hline(yintercept = round(mean(df_model$n2o), 2), 
             linetype = "dashed",
             color = "black") +
  ylab("mean N2O concentration") +
  ggtitle("National estimate comparison") +
  coord_flip() + 
  theme_tidybayes()
```

Below, estimates were compared by ecoregion.
```{r n2o_mean_wsa9, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  group_by(WSA9, .groups = "drop") %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "model") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
  bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
  arrange(ecoregion) %>%
  print()
```

```{r plot_mean_n2o_wsa9, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(WSA9) %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "model") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
  bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
  ggplot(aes(x = ecoregion, y = estimate, group = type, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  coord_flip() + 
  theme_tidybayes() +
  ylab("mean dissolved N2O") +
  ggtitle("Ecoregion estimates comparison")
```

Means were compared according to size categories below.
```{r table_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_n2o), 1),
    LCL = round(quantile(mean_n2o, probs = 0.025), 1),
    UCL = round(quantile(mean_n2o, probs = 0.975), 1),
    .groups = "drop") %>%
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests_size, type = rep("survey", 5))) %>%
  mutate(size = factor(size_cat)) %>%
  mutate(size = forcats::fct_reorder(size, estimate)) %>%
  mutate(cl_width = UCL - LCL) %>%
  arrange(size) %>%
  select(size, estimate, LCL, UCL, type) %>% 
  print()
```

```{r plot_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  #mutate(size_cat = factor(size_cat)) %>%
  mutate(type = "model") %>%
  select(size_cat, estimate, LCL, UCL, type) %>%
  bind_rows(cbind(n2o_survey_ests_size, type = rep("survey", 5))) %>%
  mutate(size_cat = forcats::fct_reorder(size_cat, estimate)) %>%
  ggplot(aes(x = size_cat, y = estimate, group = type, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  coord_flip() + 
  theme_tidybayes() +
  ylab("mean dissolved N2O") +
  ggtitle("Size category estimates comparison")
```

### Saturation
Below, the same comparisons were made for the saturation estimates.
```{r table_nat_sat_mean, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")

all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_sat = mean(n2osat), .groups = "drop") %>%
  summarise(estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3),
    .groups = "drop") %>% 
  mutate(type = "model") %>%
  bind_rows(cbind(sat_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
  add_row(estimate = round(mean(df_model$n2o / df_model$n2o_eq), 3),
          type = "sample") %>%
  print()
```

```{r plot_nat_sat_mean, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")

all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_sat = mean(n2osat)) %>%
  summarise(estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>%  
  mutate(type = "model") %>%
  bind_rows(cbind(sat_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
  ggplot(aes(x = type, y = estimate, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  geom_hline(yintercept = round(mean(df_model$n2o / df_model$n2o_eq), 3), 
             linetype = "dashed",
             color = "black") +
  ylab("mean N2O saturation ratio") +
  ggtitle("National estimates comparison") +
  coord_flip() + 
  theme_tidybayes()
```
```{r plot_wsa9_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")

all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_sat = mean(n2osat), .groups = "drop") %>%
  group_by(WSA9) %>%
  summarise( estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3),
    .groups = "drop") %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "model") %>%
  mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  bind_rows(cbind(sat_survey_ests[-10,], type = rep("survey", 9))) %>%
  mutate(cl_width = UCL - LCL) %>%
  ggplot(aes(x = ecoregion, y = estimate, group = type, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  geom_hline(yintercept = 1, color = "lightgrey") +
  ylab("mean N2O saturation ratio") +
  ggtitle("Ecoregion estimates comparison") +
  coord_flip() + 
  theme_tidybayes()
```

```{r table_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests_size.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_sat = mean(n2osat), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>%
  mutate(type = "model") %>%
  bind_rows(cbind(sat_survey_ests_size, type = rep("survey", 5))) %>%
  mutate(size = factor(size_cat)) %>%
  mutate(cl_width = UCL - LCL) %>%
  arrange(size) %>%
  select(size, estimate, LCL, UCL, type) %>% 
  print()
```

```{r plot_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")

all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_sat = mean(n2osat), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3),
    .groups = "drop") %>%
  mutate(type = "model") %>%
  bind_rows(cbind(sat_survey_ests_size, type = rep("survey", 5))) %>%
  mutate(size = factor(size_cat)) %>%
  select(size, estimate, LCL, UCL, type) %>% 
  ggplot(aes(x = size, y = estimate, group = type, color = type)) +
  geom_point(position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  ylab("mean N2O saturation ratio") +
  ggtitle("Size category estimates comparison") +
  coord_flip() + 
  theme_tidybayes()
```

# References

# Session Info
```{r session}
sessionInfo()
```

### 2023-03-03 scripts/NLA_N2O_models.Rmd

@@
 output:
+  pdf_document:
+    toc: yes
+    toc_depth: '4'
   html_notebook:

### 2023-03-08 scripts/NLA_N2O_models.Rmd
Revisions to rmd text and format. Adding github_document/.md version

--- 2023-03-02 scripts/NLA_N2O_models.Rmd
@@
 ---
-title: "Bayesian models for NLA 2017 $N_2O$ survey data"
+title: "Modeling workflow: Lake $N_2O$ survey data"
 author: "Roy Martin, Jake Beaulieu, Michael McManus"
@@
 output:
-  html_notebook:
-    toc: yes
+  github_document:
+    toc: TRUE
     toc_depth: 4
-    toc_float: yes
-    code_folding: show
-    font_size: 14
-    number_sections: yes
-    theme: simplex
+    number_sections: TRUE
+    df_print: "tibble"
+    math_method: 
+      engine: webtex
+      url: https://latex.codecogs.com/svg.image?
+    html_preview: TRUE
+    keep_html: TRUE
 bibliography: RWM_Endnote_Library.bib
@@
 
-# Rationale and Objectives
-This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O concentrations and saturation ratios using the 2017 Nation Lakes Assessment (NLA) survey data.
-The NLA sampling sites were distributed among the target population of US lakes (in the lower 48 states) according to a probabilistic survey design with samples stratified among categories of lake surface area, WSA9 ecoregion, and US state (excluding AK and HI).
-Due to the stratification scheme, some types of lakes in the sample population were intentionally over-represented (e.g., large lakes) and some were under-represented (e.g., small lakes) relative to the target population.
-Due to the unequal probability design, inferences from the sample had to be adjusted for inferences on the broader populations of interest (e.g, National-, state-, ecoregion-, and size class-specific estimates).
-
-The concept of the "complete data likelihood" is useful for conceptualizing biases arising from sampling design [@Zachmann_etal_2022; @Gelman_etal_2014 Ch.
-8; @Link_Barker_2010].
-For the NLA survey data, the population of US lakes in the lower 48 states larger than 4 hectares was considered the complete data and the probabilistic samples were considered a subset of that complete data.
-The portion of US lakes not included in the sample were considered "missing" from the complete data _not_ at random, but conditional on the pre-specified design (stratification) variables.
-This non-random missingness was not ignorable for the purpose of making inferences from the sample to the target population.
-In a model-based framework, however, including the design parameters as predictors in a regression model is one way to adjust for the missingness.
-For a thorough and recent treatment of this concept in the context of national surveys of environmental resources, refer to [@Zachmann_etal_2022].
-This concept is a key motivator for the increasingly popular mulitilevel regression with poststratification (MRP) approach to model-based inference [@Gelman_etal_2014; Gelmant_etal_2020 Ch.
+# Background and Objectives
+This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O gas concentrations and saturation ratios for a targeted population of US lakes, which included all freshwater lakes and reservoirs in the lower 48 US states larger than 4 ha.
+Data for these estimates came from the 2017 Nation Lakes Assessment (NLA) survey, wherein waterbodies were sampled according to a stratified, unequal probability design.
+Sampling was random within categories of lake size (surface area in hectares), WSA9 ecoregion, and US state (excluding AK and HI).
+In addition, larger lakes were intentionally over-sampled in proportion to smaller lakes.
+In order to make inferences from this sample to the population of interest, the estimates needed to be adjusted according to these factors.
+
+The "Complete Data Likelihood" is a useful conceptualization for describing how biases arise in sampling [@Link_Barker_2010; @Zachmann_etal_2022].
+For the 2017 NLA survey, the complete data, or population of interest, was all US lakes and reservoirs in the lower 48 states larger than 4 hectares.
+In this context, the survey samples were considered a subset of the complete data.
+Due to the pattern of selection assigned by the survey, some lakes of certain categories (e.g., small lakes) in the target population were missing from the sample not at random, but conditional on the unequal probability design.
+This type of non-random "missingness" is not ignorable when making inferences from a sample to a population of interest, as it can lead to biased estimates [@Gelman_etal_2014, Ch.
+8].
+
+A relatively simple, model-based strategy to account for this type of selection bias is to include the survey design variables as predictors in a regression model and then poststratify the estimates based on a known target population distribution; an adjustment procedure often referred to as multilevel regression and poststratification [MRP, @Park_etal_2004; @Gelman_Little_1997; @Gelman_etal_2020, Ch.
 17].
-
-The following workflow illustrates our model-based approach, based largely on the logic of MRP, but with an elaboration on the poststratification step to enable eventual estimates of total gas flux at the population level, which required scaling up from lake-level estimates.
-The typical MRP process is carried out in two steps.
-The first step is to fit regression models for the response variables of interest (e.g., dissolved N2O, equilibrium N2O) conditional on the survey design variables ┼(i.e., ecoregion, state, lake size).
-The second step is post-stratfication, wherein the posterior parameter estimates from the regression model for the sample population are weighted based on their known or assumed distribution in the population of interest [i.e., post-stratification table; @Gelman_etal_2020 Ch.
-17].
-The poststratification table in our case, for example, would be a population summary of lakes among the design variables: ecoregion, state, and size category.
-However, because we eventually needed lake-level estmates, instead of predicting to a postratification table, we predicted to each individual lake in the population of interest.
-This meant predicting to the full target population of 465,897 natural and man made US lakes larger than 4 hectares in the lower 48 states.
-These predictions were assumed relevant to average conditions during the "index period" for each lake in 2017.
-Details about the sampling frame as well as the target population are further clarified in the workbook below with data summaries and code.
-
-For the regressions, we used multilevel models fit in a fully Bayesian fashion Multilevel models are thought to work well in this context because they provide regularized estimates along the design groupings, which can improve out-of-sample inferences [@McElreath_2020].
-Inferences for lake types that may be missing from the sample, but are part of the population of interest are also straightforward using this approach [@Gelman_etal_2020 Ch.
+With a stratified design, such as the one used for the 2017 NLA, the selection scheme results in pre-specified groupings within which samples may be drawn at random.
+Within a particular grouping or stratification, the selected observations would be considered a simple random sample (SRS) from that group's population.
+A regression model conditioning the responses on that grouping structure would then result in group-specific estimates that are unbiased with respect to the groups.
+To make make unbiased inferences to a broader population of interest, those regression estimates may be re-weighted by the known distribution of units from those same groups in a broader target population.
+That distribution may be summarized in a poststratification table, which, for the 2017 NLA survey, could be a cross-tabulation of lakes in the target population along the design variables: ecoregion, state, and size category.
+Thus, a regression model conditioning on this design would provide estimates for groups of lakes described by a size classification within a particular state and ecoregion.
+To adjust the regression estimates to an estimate for the population of interest (e.g., all US lakes larger than 4 ha), those group-specific regression estimates from the sample would be re-weighted based on the proportional distribution of lakes in those groups in the target population.
+For a recent applied example and tutorial, see @Kennedy_Gelman_2021.
+For an example of a recent application in the context of national surveys of environmental resources, see @Zachmann_etal_2022.
+
+To make population-level estimates from the 2017 NLA dissolved gas data, a model-based approach, similar to those referenced above, was followed.
+The specific workflow, data, models, and code used to make these estimates are documented in the remainder of this document.
+First, Bayesian multilevel regression models were used to fit the sample data to the design variable structure.
+Multilevel models are often recommended in this type of survey estimation because they provide regularized estimates along the design groupings, which can improve out-of-sample inferences [e.g., @Gelman_Little_1997; @Kennedy_Gelman_2021].
+Inferences for group levels that may be missing from the sample, but are part of the population of interest are also straightforward using this approach [@Gelman_etal_2020 Ch.
 17; @McElreath_2020].
-More information these models, their specific parameters, R code, fit evaluations, and resulting inferences are presented in this document.
-
-The overriding objective of the modeling effort was to provide population level estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1).
-The estimates would also be used to later estimate the total flux of N2O gas attributable to the target population of lakes over the index period.
-The saturation ratio estimates were calculated as a derived quantity based on the ratio of modeled dissolved to equilibrium N2O.
-Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), we developed models for estimating their joint distribution.
-The response variable in the models was, therefore, multivariate to account for potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography.
-Although point predictions of the mean marginal probabilities from separate models could be comparable, a joint model allowing correlated observation-level errors (i.e., residuals) was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
-All of the models fit were constructed using the `brms` package [@Burkner_2017] in `R` [@R_Core_Team_2021] as an interface to Stan, a software package for fitting fully Bayesian models via Hamiltonian Monte Carlo [HMC; @Stan_Development_Team_2018_a; @Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c].
+The second step is typically a simple poststratification of the sample-based estimates over a table of target population weights.
+However, the estimates ultimately needed to be useful for estimating the total gas flux at the population level, which required lake level estimates.
+That is, flux estimates were dependent on lake-level measures of surface area.
+Therefore, instead of predicting to a postratification table, predictions were made to each individual lake in the population of interest.
+This meant predicting to 465,897 individual natural and man made US lakes larger than 4 hectares in the lower 48 states.
+These predictions were assumed relevant to average conditions during the biological index period for each lake in 2017.
+
+The specific objectives of the modeling effort were to provide population level estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1).
+The gas estimates would also be subsequently used to estimate the total flux of N2O gas attributable to the target population of lakes over the index period.
+The saturation ratio estimates were modeled as a derived quantity from the joint model of dissolved and equilibrium N2O and were calculated as the ratio of modeled dissolved to equilibrium N2O.
+Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), models for estimating their joint distribution were used.
+That is, the response variable was multivariate to account for any potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography.
+Although point predictions of the mean marginal probabilities from separate models could provide comparable estimates, a joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
+All of the models fit were constructed using the `brms` package [@Burkner_2017] in `R` [@R_Core_Team_2021] as an interface to Stan, a software package for fitting fully Bayesian models via Hamiltonian Monte Carlo [HMC, @Stan_Development_Team_2018_a; @Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c].
+More specific details on these models is provided in the "Model fitting" section below.
 
@@
 
-print(df_model)
+df_model %>%
+  print()
 ```
@@
 
-print(df_test)
+df_test %>%
+  print()
 ```
@@
 
-print(sframe)
+sframe %>%
+  print()
 ```
@@
 
-print(pframe)
+pframe %>%
+  print()
 ```
@@
 
-print(samp_props)
+samp_props %>%
+  print()
 ```
@@
 
-$$ N_2O_{dissolved}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\ 
-N_2O_{equilibrium}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$
+$$ N_2O_{diss}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\ 
+N_2O_{equi}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$
 
@@
 
-Before investing too much into the interpretation of this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC; @Gelman_etal_2014; @Gelman_etal_2020, Ch.
+Before investing too much into the interpretation of this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch.
 11].
@@
 The average predicted values were calculated as the mean prediction for each observation in the PPD based on 1000 draws.
-```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 The same checks were run below, but for the test set of 95 held-out, second-visit data points. 
-```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 As with the dissolved N2O response above, the model did an OK job at replicating the central tendency, but performed less well at replicating some important aspects of the overall distribution.
-```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 Below are the same PPCs for equilibrium N2O in the re-visit sites.
-```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 ```
+
 #### Bivariate
@@
 The model appeared to do a good job of replicating the bivariate mean, but was poor at representing the overall joint distribution.
-```{r ppc_biv1, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r ppc_biv1, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 The same bivariate check is shown below for the re-visit data.
-```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 The graphical PPCs below were aimed at evaluating how well the multivariate model did at representing the observed saturation ratio: 
-$$N_2O_{dissolved}:N_2O_{equilibrium}$$
+$$\dfrac{N_2O_{diss}} {N_2O_{equi}}$$
 This quantity was estimated as a derived variable by simply dividing the N2O PPD by the equilibrium N2O PPD.
@@
 For example, the model did a poor job replicating the observed proportion of under-saturated lakes, underestimating it by more than 10 percentage points, on average.
-```{r ppc_sat1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 The same PPCs are show below for the revisit data.
-```{r ppc_sat1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 Though the checks below suggest some improvement in replicating the tails of the observed data, this model did a poorer job at replicating central tendency.
-```{r ppc_n2o2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_n2o2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
@@
 This model also appeared to do a poorer job of replicating the mean and overall standard deviation compared to the initial model.
-```{r ppc_n2oeq2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_n2oeq2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
@@
 This check perhaps suggested an improvement with regard to replicating the joint density. However, the predictions were still clearly over-dispersed relative to the observations.
-```{r ppc_biv2, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r ppc_biv2, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
@@
 However, there appeared to still be issues in replicating the tails as well as issues with central tendency.
-```{r ppc_sat2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
@@
 However, the observed _vs._ predicted plot suggested that larger overserved values were being systematically underestimated.
-```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
@@
 The PPCs below indicated that this model appeared to be an improvement for equilibrium N2O as well. However, some checks (e.g., skewness) suggested some room for additional improvement.
-```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
@@
 The check for the joint distribution below also suggested an improvement up the previous models.
-```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
@@
 However, the proportion of under-saturated lakes remained biased low and other checks indicated that further improvements would be ideal.
-```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
@@
 Below are plots illustrating the modeled effects of covariates on both N2O and equilibrium N2O.
-#### N2O
+#### Dissolved N2O
 The conditional effects plots below for N2O illustrate a positive, monotonic, and non-linear relationship between NO3 and N2O; and a negative, linear relationship between surface temperature and N2O.
-```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3}
+```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
@@
 The modeled effects below for the equilibrium N2O component of the model illustrated a negative relationship between equilibrium N2O and both predictors and an interaction such that the surface temperature effect became slightly steeper at lower elevations.
-```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3}
+```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
@@
 Again, this model appeared to be an improvement on the previous model, particularly with regard to the more constant variance indicated in the observed _vs._ predicted plot (bottom, right panel).
-```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 This component of the model also seemed to be an improvement over model 3, with better representation in the tails as indicated in the skewness _vs._ kurtosis PPC. 
-```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 Again, an improvement over the previous model with a tighter fit of the PPC to the observed bivariate density.
-```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 This check also suggested an improvement over the previous models, with better tail behavior and less bias in the proportion under-saturated measure.
-```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 ### Covariate effects
-#### N2O
+#### Dissolved N2O
 The conditional effects plots for the covariate effects on N2O remained largely unchanged from the previous model.
-```{r conditional_effects_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 These plots suggested a large effect of NO3 on the variance of N2O, but little to no effect of surface temperature.
-```{r conditional_effects_sigma_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_sigma_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 The covariate effects on N2O remained largely the same as for the previous model.
-```{r conditional_effects_mod_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_mod_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 The covariate effects on $\sigma$ for N2O-eq suggested an negative effect of surface temperature and litte to no effect of elevation.
-```{r conditional_effects_sigma_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_sigma_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 Again, the same PPCs as above were performed for this model.
-#### N2O PPC
+#### Dissolved N2O
 This PPC for N2O looked similar to the previous model.
-```{r ppc_full_checks_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 Again, the PPCs for this model were similar to the previous model, which was unsurprising given that it was the same model for N2O-eq.
-```{r ppc_full_checks_mod_n2oeq5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2oeq5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 This PPC was also similar to the previous model.
-```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 There is also a potentially concerning extreme prediction in the observed _vs_ predicted PPC.
-```{r ppc_sat_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 ### Covariate effects
-#### N2O
+#### Dissolved N2O
 The conditional effects plot for the covariate effects N2O suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature.
@@
 For surface temperature, the effect was estimated to be largest and positive at the highest level of NO3; and negative at the lowest level of NO3.
-```{r conditional_effects_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4}
+```{r conditional_effects_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 The estimated covariate effects on $\sigma$ suggested a negative relationship with log(area) and a positive relationship, again, with NO3.
-```{r conditional_effects_sigma_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_sigma_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 The estimated covariate effect on N2O remained largely the same as estimated in the previous model.
-```{r conditional_effects_mod_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_mod_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 
-```{r conditional_effects_sigma_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_sigma_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 $$\begin{align} 
-\color{#1F449C}{\boldsymbol{N_2O_{diss}}} &=\sim Survey + Area + \color{#F05039}{\boldsymbol{NO_3}} + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
-\color{#A8B6CC}{\boldsymbol{N_2O_{equil}}} &=\sim Survey + Elev  + \color{#EEBAB4}{\boldsymbol{Temp}}\\
-\color{#F05039}{\boldsymbol{NO_3}} &=\sim Survey + Area + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
-\color{#EEBAB4}{\boldsymbol{Temp}} &=\sim Survey + Lat + Elev + Day
+\color{#1F449C}{\boldsymbol{N_2O_{diss}}} \sim Survey + Area + \color{#F05039}{\boldsymbol{NO_3}} + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
+\color{#A8B6CC}{\boldsymbol{N_2O_{equil}}} \sim Survey + Elev  + \color{#EEBAB4}{\boldsymbol{Temp}}\\
+\color{#F05039}{\boldsymbol{NO_3}} \sim Survey + Area + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
+\color{#EEBAB4}{\boldsymbol{Temp}} \sim Survey + Lat + Elev + Day
 \end{align}$$
@@
 Below, the same PPCs for N2O and N2O-eq were employed as before.
-#### N2O PPC
+#### Dissolved N2O
 The PPCs for N2O from this model were similarly reasonable as for models 4 and 5 above.
-```{r ppc_full_checks_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 
-#### Equilibrium N2O PPC
+#### Equilibrium N2O
 Again, the PPCs for N2O-eq in this model were similar to those for models 4 and 5.
-```{r ppc_full_checks_mod_n2oeq6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_full_checks_mod_n2oeq6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 
-#### Bivariate PPC
+#### Bivariate
 This model again provided a very reasonable representation of the bivariate relationship between N2O and N2O-eq (below).
-```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 
-#### Saturation PPC
+#### Saturation
 The saturation ratio PPCs below show similar behavior as with models 4 and 5 above, but with perhaps slightly less bias in the predictions for the proportion of undersaturated waterbodies and fewer extreme predictions for the means and standard deviations.
 The observed _vs._ predicted PPC also appears to have a better behaved variance and no extreme predictions, compared to models 4 and 5 with the lognormal errors.
-```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 The plot below shows the same PPC, but for the "test" or second-vist data. Overall, the model looked to perform similarly as with the data used to fit it.
-```{r ppc_sat_check_testdata_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+```{r ppc_sat_check_testdata_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 For surface temperature, the effect was estimated to be largest and positive at the highest level of NO3; and negative at the lowest level of NO3.
-```{r conditional_effects_mod_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4}
+```{r conditional_effects_mod_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 The estimated covariate effects on $\sigma$ suggested a negative relationship with log(area) and a positive relationship, again, with NO3.
-```{r conditional_effects_sigma_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_sigma_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 The estimated covariate effect on N2O remained largely the same as estimated in the previous model.
-```{r conditional_effects_mod_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_mod_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 
-```{r conditional_effects_sigma_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2}
+```{r conditional_effects_sigma_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 For reference, the max predicted value was 4403.2 nmol/L for dissolved N2O, 20.4 nmol/L for dissolved N2O, and 793.5 for the saturation ratio.
-```{r plot_n2o_posterior_preds, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+```{r plot_n2o_posterior_preds, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE, cache.lazy=FALSE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
@@
 
-```{r plot_n2oeq_posterior_preds, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_n2oeq_posterior_preds, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_sat_posterior_preds, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_sat_posterior_preds, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Below are density plots summarizing the posterior distribution of _means_ for N2O concentrations and the saturation ratio for the target population (i.e., all US lakes > 1ha in the lower 48 states).
-```{r plot_n2o_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_n2o_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_n2oeq_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_n2oeq_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_sat_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_sat_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 In distributions with right-skew, the mean can often be considerably larger than the median.
-```{r plot_sat_nat_posterior_median, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_sat_nat_posterior_median, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Below is a plot of the posterior mean estimate for the proportion of unsaturated lakes at the national scale.
-```{r plot_undersat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_undersat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Below are posterior estimates of the means for dissolved and equilibrium N2O and the saturation ratio by WSA9 ecoregion.
-```{r plot_n2o_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_n2o_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_n2oeq_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_n2oeq_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_sat_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_sat_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 A plot of the posterior estimates for the median saturation ratio below indicated, again, that most lakes in each ecoregion were undersaturated (i.e., median << 1).
-```{r plot_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 A plot of the posterior estimates for the median saturation ratio below indicated, again, that most lakes in each ecoregion were undersaturated (i.e., median << 1).
-```{r plot_prop_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_prop_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Comparisons of mean estimates (posterior median, upper and lower 95th percentiles) by state are below. Density estimates were not included to minimize plot space.
-```{r plot_state_mean_n2o, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_state_mean_n2o, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_state_mean_n2oeq, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_state_mean_n2oeq, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Only a few states (e.g., NV, DE) had median estimates that were 1 or greater, suggesting that, for most states, most lakes were undersaturated.
-```{r plot_state_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_state_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Point estimates are the posterior median of the proportion and bars are the upper and lower boundaries of the central 95th percentile of the posterior distributions of proportions.
-```{r plot_state_prop_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_state_prop_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 The estimated means by size category are below for dissolved and equilibrium N2O and the saturation ratio. Median estimates for the saturation ratio are also shown.
-```{r plot_n2o_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_n2o_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_n2oeq_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_n2oeq_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_sat_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_sat_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-```{r plot_sat_size_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_sat_size_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Mean _vs._ median below.
-```{r plot_size_cat_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_size_cat_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 And, finally, the estimated proportion of undersaturated lakes in the target population by size category
-```{r plot_prop_sat_size_cat_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_prop_sat_size_cat_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 Below, National mean estimates for dissolved N2O from the model and design-based approaches were compared. The sample-based estimate was also included as a naive reference.
-```{r n2o_means_national, message=FALSE}
+```{r n2o_means_national, message=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 The black, vertical, dashed line in the figure below represents the mean of the sample. 
-```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
+```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
 
@@
 Below, estimates were compared by ecoregion.
-```{r n2o_mean_wsa9, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
+```{r n2o_mean_wsa9, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
 
@@
 
-```{r plot_mean_n2o_wsa9, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
+```{r plot_mean_n2o_wsa9, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
 
@@
 Means were compared according to size categories below.
-```{r table_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r table_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda")
@@
 
-```{r plot_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda")
+```{r plot_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda")
 
@@
 Below, the same comparisons were made for the saturation estimates.
-```{r table_nat_sat_mean, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r table_nat_sat_mean, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
@@
 
-```{r plot_nat_sat_mean, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
+```{r plot_nat_sat_mean, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
 
@@
 ```
-```{r plot_wsa9_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
+```{r plot_wsa9_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
 
@@
 
-```{r table_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r table_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests_size.rda")
@@
 
-```{r plot_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+```{r plot_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
+#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
 
@@
 
-# References
-
 # Session Info
@@
 ```
+
+# References

### 2023-03-14 manuscript/manuscript_file.Rmd
Pushing infrastructure for writing manuscript in .Rmd

---
title: "Widespread nitrous oxide undersaturation in U.S. lakes and reservoirs"
output: word_document
date: "2023-03-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Specify paths-----
# local path for each user.
localPath <- Sys.getenv("USERPROFILE")

# Libraries-----
# load libraries needed to run scripts in this .Rmd
# libraries loaded by sourcing an external script don't seem to
# be loaded into the global knitr environment
library(tidyverse) # dplyr, ggplot

# Working directory------
# wd is set to root of RStudio project via options under Knitr drop-down
# menu above.

# source scripts------
# some libraries are called within this script.  Not sure if that is necessary
# if libraries are also called above.  Thus far I'm including necessary
# library calls in external script.
# see above for .Rmd working directory
source("manuscript/manuscript_support.R") 
```

## Introduction

## Results and Discussion

In this study we present the first national-scale survey **kasdflkjldfj** of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S. (CONUS). A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1). Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?). `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N2O sinks. We modeled the data using a bayesion hierchical approach and predicted the dissolved N2O concentration for all 224,916 CONUS waterbodies in the target population. At the population level, `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% pull()` percent (95% CI: `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% pull()` - `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% pull()`) of CONUS waterbodies were functioning as N2O sink. The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N2O sinks (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "WMT") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()`) and the Northern Plains had the greatest (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "NPL") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()`)).

```{r echo=FALSE, fig.cap= "Location of sampling sites.  Color of the points indicates whether the waterbody was functioning as a source or sink of nitrous oxide (N~2~O).  Point size indicates the dissolved N~2~O concentration"}

# [3/9/2023] knitr can only find the file if placed in the root of the RStudio project.  Why can't it pull the figure from the manuscript_figures folder?
knitr::include_graphics("figure1.png")
```

### 2023-03-14 scripts/dgIndicatorAnalysis.Rmd
Pushing infrastructure for writing manuscript in .Rmd

@@
 ```{r}
-# about 15 minutes to load first time
-# 6/7/21 <2 minutes
-Sys.time()
+# 2/8/2023 75sec on VM
+tic()
 load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
-Sys.time()
+toc()
 ```
@@
 ```{r}
-# 11 minutes on VM!
-tic() # 7 minutes on laptop
+# 4 minutes on VM. pushes memory up to 65GB [2/8/2023]
+tic()
 all_predictions <- all_predictions %>%
@@
                                    ifelse(area_km2 >=1 & area_km2 <10,
-                                          "0.1 - 1",
+                                          "1 - 10",
                                           ifelse(area_km2 >=10 & area_km2 <100,
@@
 toc()
-gc() # this pushed memory up to 57.7 GB!  gc drops it to 32GB
+gc()
 
@@
 
-# point and linerange
+# point and linerange: mean +/- 95%CI
 all_predictions %>%
@@
 
+# Box and whisker
+all_predictions %>%
+  filter(.draw == 1) %>%
+  ggplot( aes( x = WSA9, y = n2osat ) ) +
+  geom_boxplot() +
+  ylab("mean N2O saturation ratio") +
+  coord_flip() + 
+  theme_bw()
+ggsave("output/figures/n2oSatRatioByEcoBoxWhisker.tiff")
+
+# point and linerange: mean AND MEDIAN.+/- 95%CI
+all_predictions %>%
+  group_by(WSA9, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_sat = mean(n2osat),
+            median_sat = median(n2osat)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise(estimate = round(median(mean_sat), 3), 
+            LCL = round(quantile(mean_sat, probs = 0.025), 3),
+            UCL = round(quantile(mean_sat, probs = 0.975), 3),
+            estimate_md = round(median(median_sat), 3), 
+            LCL_md = round(quantile(median_sat, probs = 0.025), 3),
+            UCL_md = round(quantile(median_sat, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  #select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot(aes( x = ecoregion, y = estimate )) +
+  geom_point(color = "blue") +
+  geom_linerange( aes( ymin = LCL, ymax = UCL), color = "blue") +
+  geom_point(aes(x = ecoregion, y = estimate_md), color = "red") +
+  geom_linerange( aes( ymin = LCL_md, ymax = UCL_md), color = "red") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  ylab("mean N2O saturation ratio") +
+  coord_flip() + 
+  theme_bw()
+ggsave("output/figures/n2oSatRatioByEcoMedianAndMean.tiff")
+
+# density plots of predicted distribution, mean, and median
+# using xlim to zoom in on values close to 1
+
+dummy <- all_predictions %>%
+  group_by(WSA9) %>%
+  summarize(mean = mean(n2osat),
+            median = median(n2osat)) %>%
+  mutate(dens.med = c(2, 2.5, 2.3, 2.05, 1.75, 2.15, 2.2, 1.9, 1.9),
+         dens.mean = c(.4, 1.9, 2, .6, .65, .45, 1.5, 1.45, 0.9),
+         .draw = 1) # needed to match grouping aesthetic in ggplot call
+
+tic() # 40 seconds for 100 draws
+all_predictions %>%
+  filter(.draw %in% 1:100) %>% # subset for practice
+  #mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
+  ggplot(aes(x = n2osat, group = .draw, color = .draw ) ) +
+  geom_density(show.legend = FALSE) +
+  geom_vline(xintercept = 1, color = "red") +
+  geom_segment(data = dummy,
+               aes(x = mean, xend = mean, y = 0,
+                   yend = dens.mean),
+               linetype = "solid") +
+  geom_segment(data = dummy,
+               aes(x = median, xend = median, y = 0,
+                   yend = dens.med),
+               linetype = "dashed") +
+  ylab ("density") + 
+  xlab(expression(N[2]*O~saturation~ratio)) +
+  xlim(0,3) + # data range is 0-600, but sample max is 30
+  facet_wrap(~WSA9) +
+  theme_bw() +
+  theme(legend.position = "none")
+toc()
+ggsave("output/figures/n2oSatRatioDensityByEcoWithMeanMedian.tiff", width = 8.5, height = 5)
+
 # density plots of predicted distribution
@@
 
-# PLOT N2O* BY SIZE------------------
-# point and linerange
-all_predictions %>%
-  mutate(n2ostar = abs(n2o - n2oeq)) %>%
-  group_by(size_cat, .draw) %>% # group by iteration
-  #group_by(.draw) %>%
-  summarise(mean_star = mean(n2ostar)) %>% # 500 means for each WSA9
-  # now summarize to 1 statistic per WSA9
-  summarise( estimate = round(median(mean_star), 3), 
-    LCL = round(quantile(mean_star, probs = 0.025), 3),
-    UCL = round(quantile(mean_star, probs = 0.975), 3)) %>% 
-  ggplot( aes( x = size_cat, y = estimate ) ) +
-  geom_point() +
-  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
-  ylab("absolute[mean N2O - equilibrium N2O] (nmol/L)") +
-  coord_flip() + 
-  theme_bw() 
-ggsave("output/figures/n2oStarBySize.tiff")
-
 # PLOT N2O SAT RATIO BY WSA9 AND SIZE------------------
-
 # point and linerange
@@
 
+# PLOT N2O* BY SIZE------------------
+# point and linerange
+all_predictions %>%
+  mutate(n2ostar = abs(n2o - n2oeq)) %>%
+  group_by(size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_star = mean(n2ostar)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_star), 3), 
+    LCL = round(quantile(mean_star, probs = 0.025), 3),
+    UCL = round(quantile(mean_star, probs = 0.975), 3)) %>% 
+  ggplot( aes( x = size_cat, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  ylab("absolute[mean N2O - equilibrium N2O] (nmol/L)") +
+  coord_flip() + 
+  theme_bw() 
+ggsave("output/figures/n2oStarBySize.tiff")
+
 ```
@@
 # Merge wind and all_predictions
-tic() # 33 seconds
+tic() # 21 seconds on VM
 all_predictions <- full_join(all_predictions, wind, by = c(".row" = ".id")) # 40 seconds on laptop
@@
 
-# write reduced file to disk for less computationally intensive wrangling
-all_predictions.w.small <- all_predictions.w %>% filter(.draw %in% 1:5)
-
-save(all_predictions.w.small,
-     file = paste0(Sys.getenv("USERPROFILE"),
-                   "/Environmental Protection Agency (EPA)/",
-                   "ORD NLA17 Dissolved Gas - Documents/",
-                   "inputData/all_predictions_w_small.rda"))
+# # write reduced file to disk for less computationally intensive wrangling
+# all_predictions.w.small <- all_predictions %>% filter(.draw %in% 1:5)
+# 
+# save(all_predictions.w.small,
+#      file = paste0(Sys.getenv("USERPROFILE"),
+#                    "/Environmental Protection Agency (EPA)/",
+#                    "ORD NLA17 Dissolved Gas - Documents/",
+#                    "inputData/all_predictions_w_small.rda"))
 ```
@@
 ```{r}
-all_predictions.w %>% 
+all_predictions %>% 
   filter(.draw == 1) %>% 
@@
 
-# National scale number for lakess/reservoirs (this is is Gg)
+# National scale number for lakes/reservoirs (this is is Gg)
 national.lakes <- all_predictions %>% 
@@
 
-(national.lakes$Gg/sum(inventory$Gg)) * 1000 #0.7%
+(national.lakes$Gg/sum(inventory$Gg)) * 100 #0.07%
 ```
@@
 ```{r}
+# EMISSION RATES---------
+# density plot by ecoregion
+all_predictions %>%
+  filter(.draw %in% 1:10) %>% # subset for practice
+  #mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
+  ggplot(aes(x = e.n2o.mg.m2.d, group = .draw, color = .draw )) +
+  geom_density(show.legend = FALSE) +
+  #geom_vline(xintercept = 0, color = "red") +
+  xlab(expression("N"[2]*"O"~ "emission"~ "rate" ~ "(mg" ~ N[2]*O ~ m^-2 ~ d^-1 ~ ")")) +
+  #xlim(0,5) + # data range is 0-600, but sample max is 30
+  facet_wrap(~WSA9, scales = "free") +
+  theme_bw()
+
 # density plot with RUG for one draw
-all_predictions %>% filter(.draw == 1) %>% ggplot(aes(x=f.n2o.Mg.y)) + geom_density() +geom_rug() + facet_wrap(~WSA9, scales = "free")
-
+all_predictions %>% 
+  filter(.draw == 1) %>% 
+  ggplot(aes(x=f.n2o.Mg.y)) + 
+  geom_density() +
+  geom_rug() + 
+  facet_wrap(~WSA9, scales = "free")
+
+# point and line range by size and ecoregion
+all_predictions %>%
+  group_by(WSA9, size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_e.n2o.mg.m2.d = mean(e.n2o.mg.m2.d)) %>% # 500 means for each WSA9 x size cat
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_e.n2o.mg.m2.d), 3), 
+    LCL = round(quantile(mean_e.n2o.mg.m2.d, probs = 0.025), 3),
+    UCL = round(quantile(mean_e.n2o.mg.m2.d, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  #select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = size_cat, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  geom_hline(yintercept = 0, color='blue') +
+  ylab("mean N2O emission rate") +
+  coord_flip() + 
+  theme_bw() +
+  facet_wrap(~ecoregion)
+ggsave("output/figures/n2oEmissionRateBySizeEco.tiff", width = 7)
+
+# FLUX-------------
 # flux vs area for one draw
-all_predictions.w %>% filter(.draw == 1) %>% ggplot(aes(y=f.n2o.Mg.y, x=area_ha)) + geom_point() + facet_wrap(~WSA9, scales = "free")
+all_predictions %>% 
+  filter(.draw == 1) %>% 
+  ggplot(aes(y=f.n2o.Mg.y, x=area_ha)) + 
+  geom_point() + 
+  facet_wrap(~WSA9, scales = "free")
+
+# flux by ecoregion
+# point and linerange
+all_predictions %>%
+  group_by(WSA9, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_f = mean(f.n2o.Mg.y)) %>% # 500 means for each WSA9
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_f), 3), 
+    LCL = round(quantile(mean_f, probs = 0.025), 3),
+    UCL = round(quantile(mean_f, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = ecoregion, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  geom_hline(yintercept = 0, color='blue') +
+  ylab("mean N2O saturation ratio") +
+  coord_flip() + 
+  theme_bw()
+ggsave("output/figures/n2oSatRatioByEco.tiff")
+
+# point and line range by size and ecoregion
+all_predictions %>%
+  group_by(WSA9, size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_f.n2o.Mg.y = mean(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
+    LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
+    UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
+  mutate(ecoregion = factor(WSA9)) %>%
+  #mutate(type = "MR") %>%
+  #select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
+  ggplot( aes( x = size_cat, y = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
+  geom_hline(yintercept = 0, color='blue') +
+  ylab("mean N2O flux") +
+  coord_flip() + 
+  theme_bw() +
+  facet_wrap(~ecoregion, scales = "free")
+ggsave("output/figures/n2oFluxBySizeEco.tiff", height = 5.4, width = 9.6)
 
 # bar plot by size category and ecoregion  
-all_predictions.w %>% filter(.draw == 1) %>% group_by(WSA9, size_cat2) %>% summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
-  ggplot(aes(x=size_cat2, y=total.n2o)) + geom_bar(stat = "identity") + facet_wrap(~WSA9, scales = "free")
+all_predictions %>% 
+  filter(.draw == 1) %>% 
+  group_by(WSA9, size_cat2) %>% 
+  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
+  ggplot(aes(x=size_cat2, y=total.n2o)) + 
+  geom_bar(stat = "identity") + 
+  facet_wrap(~WSA9, scales = "free")
 
@@
   facet_wrap(~WSA9, scales = "free") +
-  theme_bw()
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90))
 
@@
 
-ggsave("output/figures/n2oFluxByLakeSizeWithError.tiff", g)
+ggsave("output/figures/n2oFluxByLakeSizeWithError.tiff", g, width = 9.6, height = 5.4)
 
 # bar plot by size category without error
-p <- all_predictions.w %>% 
+p <- all_predictions %>% 
   group_by(WSA9, .draw, size_cat2) %>% 
@@
 class(ecoreg_stats)
-str(ecoreg_stats) # WSA9 is a factor, but some of the stats have attribute of Named num
+str(ecoreg_stats) # WSA9 is a factor
 
@@
 ```
+
+# Manuscript
+
+Results and Discussion
+In this study we present the first national-scale survey of N2O concentrations and emissions from lakes,  reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S.
+(CONUS).
+A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` [weird, I have 1090, but NLA17 report says 1,005?] sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1).
+Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?).
+`r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N2O sinks.
+We modeled the data using a bayesion hierchical approach and predicted the dissolved N2O concentration for all  224,916 CONUS waterbodies in the target population.
+At the population level, `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% pull()` percent (95% CI: `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% pull()` - `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% pull()`) of CONUS waterbodies were functioning as N2O sink.
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N2O sinks (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "WMT") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()`) and the Northern Plains had the greatest (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "NPL") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()`)).
+
+This survey design allows for extrapolation of the measurement results to all 224,916 CONUS waterbodies (this is from the 2017 NLA web report) in the target population.

### 2023-03-14 manuscript/manuscript_support.R

@@ -1,0 +1,111 @@
+## This script generates data to be referenced in inline-code in the manuscript.
+## This makes the inline code more succint and the .Rmd easier to read.
+
+
+# SETUP----------------------
+# load libraries-----------------
+library(sf) # spatial data
+library(tidyverse) # dplyr, ggplot
+library(janitor) # clean names
+library(USAboundaries) # state boundaries
+library(tictoc) # processing time
+
+# Identify local path for each user
+localPath <- Sys.getenv("USERPROFILE")
+
+
+# DATA---------------
+#### load sample data (dg object)
+load(paste0(localPath,  # object name dg
+            "/Environmental Protection Agency (EPA)/",
+            "ORD NLA17 Dissolved Gas - Documents/",
+            "inputData/dg.2021-02-01.RData"))
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+
+##### load ecoregions
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple")
+
+# Check CRS
+st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS---------
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
[…]

### 2023-03-14 scripts/NLA_N2O_models.Rmd

@@
   github_document:
-    toc: TRUE
-    toc_depth: 4
     number_sections: TRUE
@@
 df_model %>%
-  print()
+  head(10)
 ```
@@
 df_test %>%
-  print()
+  head(10)
 ```
@@
 sframe %>%
-  print()
+  head(10)
 ```
@@
   spread(state, n) %>%
-  print()
+  head(10)
 ```
@@
   spread(size_cat, n) %>%
-  print()
+  head(10)
 ```
@@
 pframe %>%
-  print()
+  head(10)
 ```
@@
 samp_props %>%
-  print()
+  head(10)
 ```
@@
 
-In the $\textbf{brms}$ package, there is additionaly functionality and syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:
-
-$$ N_2O_{diss}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\ 
-N_2O_{equi}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$
+In the $\textbf{brms}$ package, there is functionality and syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:
+
+$$N_2O_{diss} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\\ 
+N_2O_{equi} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$
 
@@
 This approach to inference for N2O was similar to a Bayesian structural equation model [@Merkle_etal_2021; @Merkle_Rosseel_2018].
-The main details of the logical dependence structure could be characterized as:
+The logical dependence structure could be characterized as:
 
 $$\begin{align} 
-\color{#1F449C}{\boldsymbol{N_2O_{diss}}} \sim Survey + Area + \color{#F05039}{\boldsymbol{NO_3}} + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
-\color{#A8B6CC}{\boldsymbol{N_2O_{equil}}} \sim Survey + Elev  + \color{#EEBAB4}{\boldsymbol{Temp}}\\
-\color{#F05039}{\boldsymbol{NO_3}} \sim Survey + Area + \color{#EEBAB4}{\boldsymbol{Temp}} \\ 
-\color{#EEBAB4}{\boldsymbol{Temp}} \sim Survey + Lat + Elev + Day
+{\boldsymbol{N_2O_{diss}}} &\sim Survey + Area + {\boldsymbol{NO_3}} + {\boldsymbol{Temp}} \\ 
+{\boldsymbol{N_2O_{equil}}} &\sim Survey + Elev  + {\boldsymbol{Temp}}\\
+{\boldsymbol{NO_3}} &\sim Survey + Area + {\boldsymbol{Temp}} \\ 
+{\boldsymbol{Temp}} &\sim Survey + Lat + Elev + Day
 \end{align}$$
 
-Variables in color text above were treated as partially observed with respect to the population of interest (i.e., observed only in the sample), whereas variables in black text were considered fully observed.
+Variables in bold text above were treated as partially observed with respect to the population of interest (i.e., observed only in the sample), whereas variables not in bold were considered fully observed.
 The partially observed variables, being dissolved and equilibrium N2O, NO3, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.
@@
 That predictive distribution was then used alongside the relevant fully observed covariates to predict NO3 in the target population.
-Finally, the predictive distributions for termperature and NO3 were used to predict the N2O responses.
+Finally, the predictive distributions for temperature and NO3 were used to predict the N2O responses.
 These steps were carried out in the "Predict to population" section to follow.
 
-In the final model below, the submodel for surface temperature assumed a Gamma distributed error distribution and the linear predictor included the survey design variables, latitude, elevation, and julian date.
+In the final model below, the sub-model for surface temperature assumed a Gamma distributed error distribution and the linear predictor included the survey design variables, latitude, elevation, and Julian date.
 The shape parameter was also modeled as a function of latitude to address increasing response variance along the latitudinal gradient.
-The NO3 submodel was a cumulative logit formulation and the linear predictor included all of the survey factors as well as surface temperature and lake area.
+The NO3 sub-model was a cumulative logit formulation and the linear predictor included all of the survey factors as well as surface temperature and lake area.
 

### 2023-03-17 scripts/NLA_N2O_models.Rmd
Edits to math formatting and text

@@
 ---
-title: "Modeling workflow: Lake $N_2O$ survey data"
+title: "Modeling workflow: Lake N2O survey data"
 author: "Roy Martin, Jake Beaulieu, Michael McManus"
@@
 # Background and Objectives
-This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O gas concentrations and saturation ratios for a targeted population of US lakes, which included all freshwater lakes and reservoirs in the lower 48 US states larger than 4 ha.
-Data for these estimates came from the 2017 Nation Lakes Assessment (NLA) survey, wherein waterbodies were sampled according to a stratified, unequal probability design.
-Sampling was random within categories of lake size (surface area in hectares), WSA9 ecoregion, and US state (excluding AK and HI).
-In addition, larger lakes were intentionally over-sampled in proportion to smaller lakes.
-In order to make inferences from this sample to the population of interest, the estimates needed to be adjusted according to these factors.
-
-The "Complete Data Likelihood" is a useful conceptualization for describing how biases arise in sampling [@Link_Barker_2010; @Zachmann_etal_2022].
-For the 2017 NLA survey, the complete data, or population of interest, was all US lakes and reservoirs in the lower 48 states larger than 4 hectares.
-In this context, the survey samples were considered a subset of the complete data.
-Due to the pattern of selection assigned by the survey, some lakes of certain categories (e.g., small lakes) in the target population were missing from the sample not at random, but conditional on the unequal probability design.
-This type of non-random "missingness" is not ignorable when making inferences from a sample to a population of interest, as it can lead to biased estimates [@Gelman_etal_2014, Ch.
-8].
-
-A relatively simple, model-based strategy to account for this type of selection bias is to include the survey design variables as predictors in a regression model and then poststratify the estimates based on a known target population distribution; an adjustment procedure often referred to as multilevel regression and poststratification [MRP, @Park_etal_2004; @Gelman_Little_1997; @Gelman_etal_2020, Ch.
+This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O gas concentrations and saturation ratios for all freshwater lakes and reservoirs in the lower 48 US states larger than 4 ha.
+Data for these estimates came from the 2017 Nation Lakes Assessment (NLA) survey, wherein waterbodies were sampled according to a spatially balanced, stratified, and unequal probability design.
+Stratification was among categories of lake size (surface area in hectares), WSA9 ecoregion, and US state (excluding AK and HI); and larger lakes were intentionally over-sampled relative to smaller lakes.
+In order for this sample to be useful for making inferences relevant to the population of interest, any estimates derived from it needed to be adjusted for potential biases arising from these design factors.
+It can be useful to consider the complete data likelihood when imagining how such biases may arise in sampling designs[@Link_Barker_2010; @Zachmann_etal_2022].
+For the 2017 NLA survey, the complete data was considered to be all US lakes and reservoirs in the lower 48 states larger than 4 hectares.
+The survey samples, by comparison, are then considered a subset of the complete data with a known pattern of "missingess" assigned by the survey.
+For example, some lakes in the target population were missing from the sample not at random, but conditional on the design parameters.
+Their "missingness" is not random and, therefore, not ignorable when making inferences from the sample to the population of interest [@Gelman_etal_2014, Ch.
+8; @Zachmann_etal_2022].
+
+One model-based strategy to account for this selection bias is to include the survey design variables as predictors in a regression model and then adjust the resulting estimates based on a known target population distribution; a post-fitting adjustment referred to as poststratification [@Gelman_etal_2020, Ch.
 17].
-With a stratified design, such as the one used for the 2017 NLA, the selection scheme results in pre-specified groupings within which samples may be drawn at random.
-Within a particular grouping or stratification, the selected observations would be considered a simple random sample (SRS) from that group's population.
-A regression model conditioning the responses on that grouping structure would then result in group-specific estimates that are unbiased with respect to the groups.
-To make make unbiased inferences to a broader population of interest, those regression estimates may be re-weighted by the known distribution of units from those same groups in a broader target population.
-That distribution may be summarized in a poststratification table, which, for the 2017 NLA survey, could be a cross-tabulation of lakes in the target population along the design variables: ecoregion, state, and size category.
-Thus, a regression model conditioning on this design would provide estimates for groups of lakes described by a size classification within a particular state and ecoregion.
-To adjust the regression estimates to an estimate for the population of interest (e.g., all US lakes larger than 4 ha), those group-specific regression estimates from the sample would be re-weighted based on the proportional distribution of lakes in those groups in the target population.
-For a recent applied example and tutorial, see @Kennedy_Gelman_2021.
+A particularly popular version of this approach uses multilevel regression models and is aptly referred to as multilevel regression and poststratification, or MRP [@Park_etal_2004; @Gelman_Little_1997; @Gelman_etal_2020, Ch.
+17; @Kennedy_Gelman_2021].
+Multilevel models are often recommended because they can provide regularized estimates along the design groupings, which can improve out-of-sample inferences [e.g., @Gelman_Little_1997; @Kennedy_Gelman_2021].
+Estimates for group levels that may be missing from the sample, but are part of the population of interest, are also straightforward using the multilevel approach [@Gelman_etal_2020 Ch.
+17; @McElreath_2020].
+
+With a stratified, unequal probability design, such as the one used for the 2017 NLA, the selection design prescribes a number of random draws for each of a number of pre-specified groupings (i.e., sampling unit types).
+Within a particular stratification level or grouping, the draws may be considered a simple random sample (SRS) from that group type in the target population.
+A regression model conditioning the sample observations on the design grouping structure could, therefore, be parameterized to provide group-specific estimates that are unbiased with respect to the target population.
+To make inferences to the broader target population, however, those regression estimates would need to be re-weighted by the known proportions of units across groups in the target population.
+If the estimates are not adjusted in this manner, the estimates may be biased, particularly if the response of interest varies importantly across the grouping structure.
+For example, for the 2017 NLA survey, the proportion of samples prescribed to small lakes was much lower, relative to their proportion in the target population, compared to larger lakes, which were sampled in greater proportion relative to the target population.
+Small lakes are so numerous in the US that, if this unequal weighting by lake size weren't prescribed, nearly all the samples from the survey would be from small lakes and inferences regarding larger lakes could be highly uncertain for any reasonably achievable number of samples.
+On the other hand, if lake size had no real influence on the N2O responses, the stratification across size class would make little difference with regard to inferences.
+Sampling any sizes of lakes would result in the same marginal estimates and standard errors.
+
+For a recent applied example and helpful tutorial employing these concepts in a model-based approach using MRP, see @Kennedy_Gelman_2021.
 For an example of a recent application in the context of national surveys of environmental resources, see @Zachmann_etal_2022.
-
-To make population-level estimates from the 2017 NLA dissolved gas data, a model-based approach, similar to those referenced above, was followed.
+In this study, a similar model-based approach was used to make population-level estimates from the 2017 NLA dissolved gas data.
 The specific workflow, data, models, and code used to make these estimates are documented in the remainder of this document.
 First, Bayesian multilevel regression models were used to fit the sample data to the design variable structure.
-Multilevel models are often recommended in this type of survey estimation because they provide regularized estimates along the design groupings, which can improve out-of-sample inferences [e.g., @Gelman_Little_1997; @Kennedy_Gelman_2021].
-Inferences for group levels that may be missing from the sample, but are part of the population of interest are also straightforward using this approach [@Gelman_etal_2020 Ch.
-17; @McElreath_2020].
-The second step is typically a simple poststratification of the sample-based estimates over a table of target population weights.
-However, the estimates ultimately needed to be useful for estimating the total gas flux at the population level, which required lake level estimates.
-That is, flux estimates were dependent on lake-level measures of surface area.
-Therefore, instead of predicting to a postratification table, predictions were made to each individual lake in the population of interest.
+Next, because eventual flux estimates needed to be estimated from lake-level measures (i.e., surface area), instead of predicting to a typical postratification table, predictions were made to each individual lake in the population of interest.
 This meant predicting to 465,897 individual natural and man made US lakes larger than 4 hectares in the lower 48 states.
 These predictions were assumed relevant to average conditions during the biological index period for each lake in 2017.
-
-The specific objectives of the modeling effort were to provide population level estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1).
-The gas estimates would also be subsequently used to estimate the total flux of N2O gas attributable to the target population of lakes over the index period.
-The saturation ratio estimates were modeled as a derived quantity from the joint model of dissolved and equilibrium N2O and were calculated as the ratio of modeled dissolved to equilibrium N2O.
-Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), models for estimating their joint distribution were used.
-That is, the response variable was multivariate to account for any potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography.
-Although point predictions of the mean marginal probabilities from separate models could provide comparable estimates, a joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
+The specific objective of the modeling effort was to provide population estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1).
+As previously indicated, the gas estimates would also be used to estimate the total flux of N2O gas attributable to the target population of lakes over the index period.
+Saturation ratio was calculated as the ratio of dissolved to equilibrium N2O.
+Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), models were fit to their joint distribution and the ratio estimates were assembled as a derived quantity.
+The response variable was, therefore, multivariate in order to account for any potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography.
+Although predictions of the mean marginal probabilities from separate models may have provided comparable estimates, a joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
 All of the models fit were constructed using the `brms` package [@Burkner_2017] in `R` [@R_Core_Team_2021] as an interface to Stan, a software package for fitting fully Bayesian models via Hamiltonian Monte Carlo [HMC, @Stan_Development_Team_2018_a; @Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c].
-More specific details on these models is provided in the "Model fitting" section below.
+More specific details on the model structures is provided throughout the "Model fitting" section below.
 
@@
 (https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html), duplicate dissolved gas samples were collected at a depth of ~0.1m at designated index sites distributed across 1091 lakes nationwide, of which 95 were sampled twice as repeat visits.
-This randomly selected subset of revisit sites was used as a test set for assessing model fit and out-of-sample performance.
+This subset of revisit sites was used as a test set for assessing model fit and out-of-sample performance.
 
 Gas samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L.
-The samples were collected under a stratified, unequal probability design and each gas observation was indexed to an individual lake selected with unequal probability from 5 different lake size categories, $j \in j=1,...,J = 5$, according to surface area (ha), and from within a state, $k \in k=1,...,K = 48$, situated within an aggregated, WSA9 or Omernik ecoregion, $l \in l=1,...,L = 9$.
+As part of the survey design, each gas observation was indexed to an individual lake selected with unequal probability from 5 different lake size categories, $j \in j=1,...,J = 5$, according to surface area (ha), and from within a state, $k \in k=1,...,K = 48$, situated within an aggregated, WSA9 or Omernik ecoregion, $l \in l=1,...,L = 9$.
 All 9 WSA9 ecoregions were represented in the sample, including Xeric (XER), Western Mountain (WMT),  Northern Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and Southern Appalachian (SAP) regions.
@@
 https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html. That dataset was imported below.
+
 ```{r import_data, eval=FALSE, include=TRUE}
@@
 
-From the imported dataset, a new data frame for modeling was constructed from the original file including only the variables of interest: (1) the N2O gas observations; (2) the survey design variables indexed to those observations; and (3) additional covariates considered potentially useful for improving the fit of the model.
-The data frame below excluded the second-visit observations, which would later be used for model checking.
+A new data frame for modeling was constructed from the original file including only the variables of interest: (1) the N2O gas observations; (2) the survey design variables indexed to those observations; and (3) additional covariates considered potentially useful for improving the fit of the model.
+The modeling data frame below excluded the second-visit observations, which would later be used for model checking.
 Some variables from the imported data were renamed for convenience.
-In addition, the NO3 covariate was rounded according to the documented measurement precision.
-An alternative version of the NO3 covariate was also created in this step by log-transforming and re-coding it as an ordered factor with five levels at hand-drawn cut points.
+In addition, the NO3 covariate was rounded according to the documented measurement precision and an alternative version was also created by log-transforming and re-coding the variable as an ordered factor with five levels.
 The left-most cut point separated observations below the detection limit from the completely observed samples.
-The remaining cut points in the positive direction were drawn at approximately equal distances along the log scale.
-Finally, it should be noted that one lake that was sampled was missing information on the N2O gas measurements and it was removed from the data frame.
+The remaining cut points were drawn at approximately equal distances in the positive direction along the log scale.
+Finally, one lake had no N2O gas information and was removed from the data frame.
+
 ```{r model_data, echo=TRUE, paged.print=TRUE}
@@
 
-A second dataframe, including only the second visit observations, was constructed below.
-These data were later used as a "test set" to assess the out-of-sample fit of the model developed on the first-visit or training data.
+A second data frame including only the second visit observations was constructed below.
+These data were later used as a "test set" to assess the out-of-sample performance of the model developed on the first-visit or "training set".
+
 ```{r test_data, echo=TRUE, paged.print=TRUE}
@@
 ## Target population
-Below. the NLA sampling frame was imported and then filtered to include only the target population or sampling frame for this project.
+Below. the NLA sampling frame was imported and filtered to include all lakes the target population. The resulting target population above included a total of 465,897 waterbodies.
+
 ```{r import_sample_frame, echo=TRUE}
@@
 
-The resulting target population above included a total of 465,897 waterbodies.
-
-Cross tabulations below describe the structure of the target population with respect to the design variables.
-The cross-tabulation makes it clear that each ecoregion does not contain each state.
+Cross tabulations below describe the structure of the target population with respect to the survey design variables.
+The cross-tabulation shows that each ecoregion does not contain each state.
 Therefore, in the statistical sense, states were nested in ecoregions.
+
 ```{r frame_dimensions_1, echo=TRUE, message=FALSE}
@@
 
-Likewise, lake size category was nested in state (which was nested in ecoregion). That is, not every ecoregion:state in the population of interest contained every size category (below).
+The cross-tablulation below indicates that lake size category was nested in state (which was nested in ecoregion). That is, not every ecoregion:state contained every size category.
+
 ```{r frame_dimensions_4, echo=TRUE, message=FALSE}
@@
 
-Below, the sampling frame was selected down to create a post-stratification table.
-Some of the variables were renamed to match the naming conventions used in the observational data above.
-There were 536 types of lakes in the population of interest with respect to the sampling design.
-The counts of those lake types (n_lakes) and their proportions relative to the total population of lakes in the sampling frame (prop_cell) are indicated below.
-```{r filter_frame, echo=TRUE, message=FALSE}
+Below, the sampling frame was munged to create a post-stratification table.
+There were 536 "types" or groupings of lakes in the population of interest with respect to the sampling design.
+The total counts of those lake types (n_lakes) and their proportions (prop_cell) relative to the counts in the target population were tabulated.
+
+```{r filter_frame, echo=TRUE, message=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")
@@
 There were 352 lake types in the sample compared to the 536 in the population of of interest.
-In total, there were 984 observations distributed across these 352 lake types in the sample; and the number of samples was not distributed evenly across the types.
+There were 984 observations distributed across these 352 lake types in the sample; and the number of samples was not distributed evenly across the types.
 Some cells were represented by as few as 1 lake.
 In total, 536-352 = 184 lake types in the population of interest were not represented in the sample.
-```{r sample_cell_counts, echo=TRUE, message=FALSE}
+
+```{r sample_cell_counts, echo=TRUE, message=FALSE, cache=TRUE}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
@@
 
-Below, a graphical comparison was constructed to depict the distribution of cells in the population of interest _versus_ those in the sample.
+Below, a graphical comparison was constructed to depict the distribution of cells in the population of interest _vs._ those in the sample.
+
 ```{r compare_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
@@
 
-Another comparison between population and sample was constructed below by ecoregion.
+Another comparison between population and sample was constructed by ecoregion.
 The samples were not balanced across ecoregions.
 Lakes in the Coastal Plains (CPL) ecoregion, for example, were clearly undersampled relative to their proportion of the population.
+
 ```{r eco_props_pop, eval=FALSE, include=TRUE, message=FALSE}
@@
 A similar comparison by state was constructed below.
+
 ```{r state_props_pop, eval=FALSE, include=TRUE, message=FALSE}
@@
 
-Finally, a comparison by lake size category is shown below. Note that small lakes were under-sampled relative to larger lakes by design.
+Finally, a comparison by lake size category is shown below. Small lakes were under-sampled relative to larger lakes.
+
 ```{r size_props_pop, eval=FALSE, include=TRUE, message=FALSE}
@@
 ## Sample-based estimates
-The overall mean and standard deviation for N2O in the sample:
+Below are _naive_ estimates, based only on the sample, for national means for dissolved and equilibrium N2O and the saturation ratio.
+
 ```{r sample_summary_n2o}
@@
 
-The same summary for equilibrium N2O:
 ```{r sample_summary_n2oeq}
@@
 
-The saturation ratio (i.e., N2O / N2O-eq):
 ```{r sample_summary_n2osat}
@@
 
-Finally, roughly 67% of lakes in the sample were undersaturated (i.e., saturation ratio < 1):
+Roughly 67% of lakes in the sample were under-saturated (i.e., saturation ratio < 1):
+
 ```{r sample_summary_propsat}
@@
 
-Using only the sample observations, a plot was constructed below of the overall mean (dashed line) along with the ecoregion-specific means (black circles).
+Using only the sample observations again, a plot was constructed of the overall mean (dashed line) along with the ecoregion-specific means (black circles).
 The shaded areas indicate +/- 1 standard deviation.
-Neither dissolved N2O nor the saturation ratio were clearly structured by ecoregion in the sample, but there did appear to be some structure in the equilibrium N2O observations.
+Neither dissolved N2O nor the saturation ratio were clearly structured by ecoregion, but there did appear to be some structure along this variable in the equilibrium N2O observations.
+
 ```{r sample_summary_eco, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
@@
 The same summary by state is below.
+
 ```{r sample_summary_state, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
@@
 
-Finally, the same summary by size category:
+Finally, the same summary by size category.
+
 ```{r sample_summary_size, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
@@
 ## Sample data exploration
-Below, the empirical distribution of N2O observations in the sample was summarized using a density and rug plot below.
-Note the natural log scale of the x axis.
-Both the N2O and equilibrium N2O data had considerable right skew even after the log transformation, which was not unexpected and has been noted in other studies [@Webb_etal_2019].
+Below, the distribution N2O concentrations for the sample was summarized using a density and rug plot.
+Note the natural log scale of the x-axis.
+Both the dissolved and equilibrium N2O data had considerable right skew even after the log transformation.
+This was not unexpected and has been noted in other studies [@Webb_etal_2019].
 The saturation ratio was also skewed since it was derived from the other two observed variables (i.e., sat_ratio = n2o / n2o_eq).
+
 ```{r summary_N2O, echo=FALSE, fig.align='center', fig.height=8, fig.width=5, message=FALSE}
@@
 
-Below are plots of N2O vs.
+Below are plots of dissolved N2O vs.
 NO3.
@@
 The leftmost vertical line is dashed and separates the NO3 observations below the detection limit.
+The trend is increasing and nonlinear on the log scale, with increasing variance in N2O as NO3 increased.
+
 ```{r summary_N2O_vs_NO3, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
@@
 
-In the plot above, the trend is increasing and nonlinear on the log scale.
-The increasing variance in N2O along the NO3 gradeient suggested a potential mediator of the relationship between NO3 on N2O.
-Below are plots of N2O vs.
+Below are plots of dissolved N2O vs.
 NO3 for 6 quantiles of the surface temperature measurements (quantiles increasing from 1 to 6).
-This plot below suggested that the NO3 effect on N2O may have been stronger in lakes with higher observed temperatures.
+This plot suggested that the NO3 effect on N2O may have been stronger in lakes with higher observed temperatures.
+
 ```{r summary_N2O_vs_NO3_surftemp, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
@@
 
-The next plot below shows the relationship between N2O and NO3 at 6 different quantiles (increasing 1 to 6) of the log-scaled lake surface area estimates.
+The next plot below shows the relationship between dissolved N2O and NO3 at 6 different quantiles (increasing 1 to 6) of the log-scaled lake surface area estimates.
+
 ```{r summary_N2O_vs_NO3_logarea, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
@@
 Note the large number of observations in the first NO3 category (no3_cat = 1).
-This category represented all of the censored observations for NO3, which was most of the data.
+This category represented all of the observations for NO3 that were below the detection limit, which was most of the data.
+
 ```{r summary_N2O_vs_NO3cat, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE}
@@
 
-Below is a plot of log(N2O) vs. log(NO3) by ecoregion, which suggested that the NO3 effect on N2O may have varied by ecoregion.
+Below is a plot of log(dissolved N2O) vs. log(NO3) by ecoregion, which suggested that the NO3 effect on N2O may have varied by ecoregion.
+
 ```{r summary_N2O_vs_NO3_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
@@
 Below is the same plot as above but for the ordered categorical version of NO3.
+
 ```{r summary_N2O_vs_NO3cat_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
@@
 A plot below shows trends by state within just the Temperate Plains (TPL) ecoregion. Within states, the number of observations were relatively small, but the trends appeared closer to linear.
+
 ```{r summary_N2O_vs_NO3_wsa9state3, echo=FALSE, fig.align='center', fig.height=4, fig.width=8, message=FALSE}
@@
 # Model fitting
-The first regression model was constructed to estimate the joint distribution of log-transformed N2O and equilibrium N2O conditional on the the design factors.
+The first regression model was constructed to estimate the joint distribution of log-transformed dissolved and equilibrium N2O conditional on the the design factors.
 Each log-transformed observation, $i \in 1,..,N=984$, for each response, $p \in 1:P=2$, was assumed to be drawn from a multivariate normal distribution with the parameters $\nu$ and $\Sigma$, where $\nu$ is the multivariate mean estimated conditional on the design effects and $\Sigma$ is a covariance matrix containing the observation-level variances and residual correlation:
@@
 
-The multivariate mean is a vector of mean parameters, $\nu:[\mu_{p=1}, \mu_{p=2}]$, for each response.
-Each mean is further defined by a linear combination of parameters where, for each response $p$ and observation $i$:
+The multivariate mean is a vector of location parameters, $\nu:[\mu_{p=1}, \mu_{p=2}]$, for each response.
+Each location parameter was further defined by a linear combination of parameters where, for each response $p$ and observation $i$:
 
@@
 
-The linear combination of parameters defining $\mu$ above include a fixed global intercept, $a_0$, that is estimated directly from the data, and three separate, latent group-level effects matrices, $\alpha_1, \alpha_2, \alpha_3$.
-The group effects were assumed to be multivariate normal and are centered on zero in multivariate space.
-The spread of the effects around zero are determined by a covariance matrix, $\Lambda_1, \Lambda_2, \text{or } \Lambda_3$, which are estimated directly from the data.
-These covariance terms are further defined where:
-
-$$\Lambda = \begin{pmatrix} 1 & \tau_{p=1} \\ \tau_{p=2} & 1 \end{pmatrix} \chi \begin{pmatrix} 1 & \tau_{p=1} \\ \tau_{p=2} & 1 \end{pmatrix}$$
-
-The $\tau$ parameters are the group-level scale parameters, which constrain the spread of effects for each response, and $\chi$ comprises the group-level residual correlation matrix:
+The linear combination included a fixed global intercept, $a_0$, estimated directly from the data, and three separate, latent group-level effects matrices, $\alpha_1, \alpha_2, \alpha_3$.
+The group effects were assumed to be multivariate normal and were centered on zero in the two-dimensional multivariate space.
+The spread of the effects around zero were determined by a covariance matrix, $\Lambda_1, \Lambda_2, \text{or } \Lambda_3$, which were estimated directly from the data.
+The covariance terms were further defined where:
+
+$$\Lambda = \begin{pmatrix} 1 & \tau^2_{p=1} \\ \tau^2_{p=2} & 1 \end{pmatrix} \chi \begin{pmatrix} 1 & \tau^2_{p=1} \\ \tau^2_{p=2} & 1 \end{pmatrix}$$
+
+The $\tau$ parameters captured the group-level standard deviations, which constrain the spread of group-level effects for each response, and $\chi$ was the group-level residual correlation matrix:
 
@@
 
-wherein $\varrho$ is the group-level residual correlation between responses.
+wherein $\varrho$ captured the group-level residual correlation between effects.
 
 The explicit indexing in the notation above conveys the relationship between the parameters and each observation, $i$, and emphasizes the nested structure of the observations within the group effects.
-Specifically, every observation, $i$, was nested in a lake size category, $l$, which was nested in a state, $k$, and ecoregion, $j$.
+Specifically, each observation, $i$, was nested in a lake size category, $l$, which was nested in a state, $k$, and ecoregion, $j$.
 The parameter $\alpha_1$, therefore, accounted for ecoregion-scale group effects or deviations from the global mean; $\alpha_2$ accounted for state-level group effects nested in ecoregions; and $\alpha_3$ accounted for lake size group effects within states and ecoregions.
@@
 Finally, the observation-level covariance term, $\Sigma$, was parameterized as:
-$$\Sigma = \begin{pmatrix} 1 & \sigma_{p=1} \\ \sigma_{p=2} & 1 \end{pmatrix} \Omega \begin{pmatrix} 1 & \sigma_{p=1} \\ \sigma_{p=2} & 1 \end{pmatrix}$$
-
-wherein the $\sigma$ parameters are the observation-level standard deviations for each response and $\Omega$ comprises the observation-level residual correlation matrix:
+$$\Sigma = \begin{pmatrix} 1 & \sigma^2_{p=1} \\ \sigma^2_{p=2} & 1 \end{pmatrix} \Omega \begin{pmatrix} 1 & \sigma^2_{p=1} \\ \sigma^2_{p=2} & 1 \end{pmatrix}$$
+
+wherein the $\sigma$ parameters were the observation-level standard deviations for each response and $\Omega$ was the observation-level residual correlation matrix:
 $$\Omega = \begin{pmatrix} 1 & \rho \\ \rho & 1 \end{pmatrix}$$
-wherein $\rho$ is the residual correlation between responses.
+wherein $\rho$ captured the residual correlation between responses.
 
 For model fitting, priors were needed for all parameters conditioned directly on the data, which included the global intercept, the scale parameters, and the correlation matrices.
-A normal or Gaussian prior, $N(\mu = 2, \sigma = 1)$ centered near the (log-scale) data means, was used for the global intercept parameter for each response.
+A normal or Gaussian prior, $N(\mu = 2, \sigma = 1)$ centered near the (log-scale) data means, was used for the global mean parameter for each response.
 This prior was considered minimally informative as it placed most (~80%) of the prior mass over values between approximately 2 and 27 ng/L for median N2O or N2O equilibrium concentration and included support in the tails for values approaching 0 ng/L on the lower end and 80 ng/L on the high end.
@@
 Finally, for the correlation matrices, an $LKJ(\eta =2)$ prior was used, which, for a 2-dimensional response, placed most support for correlations between approximately -0.9 and 0.9.
-This prior seemed reasonable as there was no clear causal mechanisms that were thought to ensure a strong direct correlation between the N2O measures.
-Any potential residual dependence was expected to be indirect due to, for example, a common causal factor (e.g., elevation, temperature).
+This prior seemed reasonable as there were no apparant causal mechanisms to ensure a very strong correlation between the N2O measures.
+Any potential residual dependence was expected to be indirect due to, for example, a common correlate (e.g., elevation, temperature).
 For more information on prior choice recommendations in Stan, see: https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
@@
 The $\textbf{brms}$ package [@Burkner_2017] for $\textbf{R}$ [@R_Core_Team_2021] was used to fit all of the models in a fully Bayesian setting.
-The formula syntax of the $\textbf{brms}$ package is similar to the syntax used in the $\textbf{lme4}$ package that is widely used to fit mixed effects models in frequentist settings In either package, the linear predictor for $\mu$ described above could be expressed as:
+The formula syntax of the $\textbf{brms}$ package is similar to the syntax used in the $\textbf{lme4}$ package that is widely used to fit mixed effects models in frequentist settings.
+In either package, the linear predictor for $\mu$ described above could be expressed as:
 
@@
 
-In the $\textbf{brms}$ package, there is functionality and syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:
-
-$$N_2O_{diss} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\\ 
-N_2O_{equi} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$
-
-The above syntax would indicate that the linear predictor for both responses in the multivariate model have the same group-level varying effects, and that each of those effects are allowed to be correlated between responses.
-
-For the remainder of this document, only this simplified syntax is presented to describe the model parameterizations.
+In the $\textbf{brms}$ package, there is additional functionality and syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:
+
+$$
+\begin{aligned} 
+  N_2O_{diss} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\
+  N_2O_{equi} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) 
+\end{aligned}
+$$
+
+The above syntax would indicate that the linear predictor for both responses in the multivariate model have the same group-level varying effects, and that those effects may be correlated between responses.
+
+For the remainder of this document, only this simplified syntax is presented to describe the model structure.
 For more information on $\textbf{brms}$ functionality and syntax with multivariate response models, the package vignette may be helpful, and can be found at: https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html.
@@
 ## Model 1
-The first model fit was the one described above.
+The first model fit was as explained above.
+
 ```{r n2o_mod_mv_1, eval=FALSE, include=TRUE}
@@
 All $\hat{R}$ values were less than 1.01 and effective sample size ($ESS$) calculations suggested that the posterior contained a sufficient number of effective samples for conducting inference.
+
 ```{r print_mod1, echo=FALSE}
@@
 ```
+
 In the summary above, the estimated standard deviations for the varying group effects on the mean behavior of the dissolved N2O response suggested fairly low, but non-zero variability across each of the three levels.
@@
 
-Before investing too much into the interpretation of this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch.
+Before investing too much into the inferences from this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch.
 11].
@@
 #### Dissolved N2O
-Below are a series of panels illustrating graphical PPCs for the log(N2O) component of the model.
-The top left panel compares a density plot of the observed data (black line) to density lines drawn for 200 samples from the posterior predictive distribution (PPD; blue lines) of the fitted model.
-The top right panel similarly compares the cumulative density distributions.
-The left middle panel simulataneously compares means _vs._ standard deviations for 1000 draws from the PPD (blue dots) to the sample mean and standard deviation (black dot).
-The right middle panel compares skewness _vs._ kurtosis for 1000 draws from the PPD to the skewness and kurtosis values calculated for the observed data.
-The bottom left panel compares max _vs._ min values for 1000 draws from the PPD to the max and min values of the sample data.
+Below are a series of panels illustrating graphical PPCs for the dissolved N2O component of the multivariate model.
+The top left panel compares a density plot of the observed data (black line) to density estimates drawn for 200 samples from the posterior predictive distribution (PPD; blue lines) of the fitted model.
+The top right panel compares the cumulative density distributions in the same manner.
+The left middle panel compares means _vs._ standard deviations for 1000 draws from the PPD (blue dots) to the sample mean _vs._ standard deviation (black dot).
+The right middle panel compares skewness _vs._ kurtosis for 1000 draws from the PPD to the skewness _vs._ kurtosis estimates for the observed data.
+The bottom left panel compares max _vs._ min values for 1000 draws from the PPD to the max _vs._ min values of the sample data.
 Finally, the bottom right panel shows the observed _vs._ average predicted values for each observation in the sample.
 The average predicted values were calculated as the mean prediction for each observation in the PPD based on 1000 draws.
+
 ```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 ```
+
 The general takeaway from the PPCs above was that the model replicated the central tendency of the observed data fairly well, but failed to sufficiently replicate other important aspects of the distribution, such as skewness and kurtosis.
-The observed _vs._ average predictions scatterplot suggested substantial heteroscedasticity in the errors.
-
-The same checks were run below, but for the test set of 95 held-out, second-visit data points. 
+The observed _vs._ average predictions scatterplot also suggested substantial heteroscedasticity in the errors.
+
+The same checks were run below, but for the test set of 95 held-out, second-visit data points. The patterns in misfit were similar to the patterns indicated in the PPCs with the training data above.
+
 ```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 ```
-The patterns in misfit indicated above for the re-visit data were similar to the patterns indicated in the PPCs with the training data.
 
@@
 Below are PPCs for the equilibrium N2O component of the model.
-As with the dissolved N2O response above, the model did an OK job at replicating the central tendency, but performed less well at replicating some important aspects of the overall distribution.
+As with the dissolved N2O PPCs, the model seemed to do an OK job at replicating the central tendency, but performed less well at replicating other important aspects of the overall distribution.
+
 ```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Below are the same PPCs for equilibrium N2O in the re-visit sites.
+
 ```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### Bivariate
-The graphical check below compares bivariate density contours estimated from the observed data (black lines) to density contours estimated for each of 20 draws from the PPD.
+The graphical check below compared bivariate density contours estimated from the observed data (black lines) to density contours estimated for each of 20 draws from the PPD.
 The model appeared to do a good job of replicating the bivariate mean, but was poor at representing the overall joint distribution.
+
 ```{r ppc_biv1, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 The same bivariate check is shown below for the re-visit data.
+
 ```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 The graphical PPCs below were aimed at evaluating how well the multivariate model did at representing the observed saturation ratio:
+
 $$\dfrac{N_2O_{diss}} {N_2O_{equi}}$$
-This quantity was estimated as a derived variable by simply dividing the N2O PPD by the equilibrium N2O PPD.
-Likewise, the proportion of under-saturated lakes in the sample was estimated by summing the number of lakes from each posterior predictive draw wherein the ratio was < 1 and dividing that number by the total number of lakes in the sample, which was 984.
-Overall, these checks indicated that properly representing the tails of the N2O and N2O-eq observations would likely be necessary in order to better replicate the observed saturation metrics.
-For example, the model did a poor job replicating the observed proportion of under-saturated lakes, underestimating it by more than 10 percentage points, on average.
+
+This quantity was estimated as a derived variable by dividing the dissolved N2O PPD by the equilibrium N2O PPD.
+The proportion of under-saturated lakes in the sample was estimated by summing the number of lakes from each posterior predictive draw wherein the ratio was < 1 and dividing that number by the total number of lakes in the sample, which was 984.
+Overall, these checks indicated that properly representing the tails of the dissolved N2O and N2O-eq observations would likely be necessary in order to better replicate the observed saturation metrics.
+The observed proportion of under-saturated lakes was underestimated by more than 10 percentage points, on average.
+
+The top left panel, below, is a density plot of the observed saturation ratio (black line) compared to an estimate using 50 draws from the derived PPD (blue lines).
+The top right panel shows the observed proportion of under-saturated lakes compared to a model estimate based on 1000 draws from the PPD.
+The left middle panel shows the mean _vs._ standard deviation of the saturation ratio for the observed data compared to the same estimates for 500 draws from the PPD.
+The right middle panel shows the max _vs._ min for the sample compared to 500 draws from the PPD.
+Finally, the bottom left panel shows the observed _vs._ average predicted saturation ratio for all 984 lakes sampled in the dataset.
+
 ```{r ppc_sat1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 
-The top left panel, above, is a density plot of the observed saturation ratio (black line) compared to an estimate using 50 draws from the model (blue lines).
-The top right panel shows the observed proportion of under-saturated lakes compared to a model estimate based on 1000 draws from the PPD.
-The left middle panel shows the mean _vs._ standard deviation of the saturation ratio for the observed data compared to the same estimates for 500 posterior draws from the model's PPD.
-The right middle panel shows the max _vs._ min for the sample compared to 500 draws from the model's PPD.
-Finally, the bottom left panel shows the observed _vs._ average predicted saturation ratio for all 984 lakes sampled in the dataset.
-
-The same PPCs are show below for the revisit data.
+The same PPCs are shown below for the revisit data. These checks indicated that the model did a similarly underwhelming job of replicating the re-visit data.
+
 ```{r ppc_sat1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 ```
-The checks above indicated that the model did a similarly underwhelming job of replicating some key properties of the saturation metrics calculated from the re-visit data.
 
 #### R-square
-Below, the Bayesian $R^2$ values are reported for each reasponse in the model. 
-```{r r2_1, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+Below, the Bayesian $R^2$ values are reported for each response in the model.
+
+```{r r2_1, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 The $R^2$ were also estimated for the re-visit data.
-```{r r2_1_test, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+
+```{r r2_1_test, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
@@
 ## Model 2
-In an attempt to better fit the observed data, the next model included distributional sub-models to allow for heterogeneous variances for each response conditional on the survey design structure.
+In an attempt to better fit the observed data, the next model included a distributional model for each sub-model that allowed for heterogeneous variances.
+The distributional terms were each fit as a function of the survey design structure.
+The same structure as for the models for the mean components.
+
 ```{r n2o_mod_2, eval=FALSE, include=TRUE}
@@
 The summaries of the estimated parameters and key HMC convergence diagnostics for the fitted model are printed below.
+
 ```{r print_mod2, echo=FALSE}
@@
 ```
-From the summary above, note the moderate and positive residual correlation between the two N2O responses.
-The estimated standard deviations for the varying group effects on the mean behavior of the dissolved N2O response suggested fairly low, but non-zero variability across each of the three levels.
-The standard deviations estimated for the same varying effects for equilibrium N2O were also relatively small.
-However, before investing too much into the interpretation of these results, the model fit was further evaluated below using a series of graphical posterior predictive checks (PPCs).
 
@@
 Below the same PPCs were performed as with the initial model (see above for more details on each panel).
+Though the checks below suggested some improvement in replicating the tails of the observed data, the overall fit again suggested room for improvement.
+
 ##### Dissolved N2O
-Though the checks below suggest some improvement in replicating the tails of the observed data, this model did a poorer job at replicating central tendency.
+
 ```{r ppc_n2o2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### Equilibrium N2O
-The checks below suggest this model offered no improvement upon the initial model for equilibrium N2O.
-This model also appeared to do a poorer job of replicating the mean and overall standard deviation compared to the initial model.
+The checks below suggested that this model offered little no improvement upon the initial model for equilibrium N2O.
+
 ```{r ppc_n2oeq2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 This check perhaps suggested an improvement with regard to replicating the joint density. However, the predictions were still clearly over-dispersed relative to the observations.
+
 ```{r ppc_biv2, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 However, there appeared to still be issues in replicating the tails as well as issues with central tendency.
+
 ```{r ppc_sat2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### R-square
-Relative to model 1, there was a substantial decrease in the $R^2$ estimate for the dissolved N2O component of this model.
-The estimate for the equilibrium N2O-eq component was similar to the model 1.
-```{r r2_2, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+Relative to model 1, there was a substantial decrease in the $R^2$ estimate for the dissolved N2O component of this model. The estimate for the equilibrium N2O-eq component was similar to the model 1.
+
+```{r r2_2, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
@@
 ## Model 3
-In the next model, we used covariates to try to improve the fit.
-The categorical version of the NO3 covariate was used as a monotonic ordinal predictor in the dissolved N2O component of the modl.
-For the equlibrium N2O component, we included surface temperature and log-transformed elevation, along with their interaction.
-The models also retained the distributional specifications included in model 2 above.
+In the next model, the categorical version of the NO3 covariate and an surface temperature covariate were included to try to improve the fit.
+The ordinal NO3 variable was used as a monotonic, ordinal effect and only in the dissolved N2O component of the model.
+For the equlibrium N2O component, surface temperature and log-transformed elevation were used, along with their interaction.
+The model also retained the same distributional specifications included in model 2 above.
+
 ```{r n2o_mod_3, eval=FALSE, include=TRUE}
@@
 The fitted parameters and MCMC diagnostics are below.
+
 ```{r print_mod3, echo=FALSE}
@@
 The central tendency and tail behavior looked to be reasonably replicated by comparison.
-However, the observed _vs._ predicted plot suggested that larger overserved values were being systematically underestimated.
+However, the observed _vs._ predicted plot suggested that larger overserved values were likely being systematically underestimated.
+
 ```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 The PPCs below indicated that this model appeared to be an improvement for equilibrium N2O as well. However, some checks (e.g., skewness) suggested some room for additional improvement.
+
 ```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### Bivariate
-The check for the joint distribution below also suggested an improvement up the previous models.
+The check for the joint distribution below also suggested an improvement upon the previous models.
+
 ```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 However, the proportion of under-saturated lakes remained biased low and other checks indicated that further improvements would be ideal.
+
 ```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 The $R^2$ estimates for this model are below and suggested substantial improvements on the previous models.
-```{r r2_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+
+```{r r2_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
@@
 Below are plots illustrating the modeled effects of covariates on both N2O and equilibrium N2O.
+
 #### Dissolved N2O
-The conditional effects plots below for N2O illustrate a positive, monotonic, and non-linear relationship between NO3 and N2O; and a negative, linear relationship between surface temperature and N2O.
+The conditional effects plots below for dissolved N2O illustrated a positive, monotonic, and non-linear relationship with NO3; and a negative, linear relationship with surface temperature.
+
 ```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3, cache=TRUE}
@@
 #### Equilibrium N2O
-The modeled effects below for the equilibrium N2O component of the model illustrated a negative relationship between equilibrium N2O and both predictors and an interaction such that the surface temperature effect became slightly steeper at lower elevations.
+The modeled effects for the equilibrium N2O component illustrated a negative relationship with both predictors and an interaction such that the surface temperature effect became slightly steeper at lower elevations.
+
 ```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3, cache=TRUE}
@@
 In the next model, covariate terms were also included in the $\sigma$ components of both models in order to try to better capture remaining heterogeneity in the variances of both N2O and N2O-eq.
+
 ```{r n2o_mod_mv_4, eval=FALSE, include=TRUE}
@@
 Below is a summary of the fitted parameters along with some convergence diagnostics.
+
 ```{r print_mod4, echo=FALSE}
@@
 ### Model checks
-Again, the same PPCs were employed for this model as above.
+The same PPCs were employed for this model as above.
+
 #### Dissolved N2O
-Again, this model appeared to be an improvement on the previous model, particularly with regard to the more constant variance indicated in the observed _vs._ predicted plot (bottom, right panel).
+This model appeared to be an improvement on the previous model, particularly with regard to the more constant variance indicated in the observed _vs._ predicted plot (bottom, right panel).
+
 ```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 This component of the model also seemed to be an improvement over model 3, with better representation in the tails as indicated in the skewness _vs._ kurtosis PPC.
+
 ```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Again, an improvement over the previous model with a tighter fit of the PPC to the observed bivariate density.
+
 ```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 This check also suggested an improvement over the previous models, with better tail behavior and less bias in the proportion under-saturated measure.
+
 ```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 The Bayesian $R^2$ estimates below indicated an improvement from the previous models.
-```{r r2_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+
+```{r r2_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
@@
 #### Dissolved N2O
-The conditional effects plots for the covariate effects on N2O remained largely unchanged from the previous model.
+The conditional effects plots for the covariate effects on dissolved N2O remained largely unchanged from the previous model.
+
 ```{r conditional_effects_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 These plots suggested a large effect of NO3 on the variance of N2O, but little to no effect of surface temperature.
+
 ```{r conditional_effects_sigma_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 #### Equilibrium N2O
-The covariate effects on N2O remained largely the same as for the previous model.
+The covariate effects on equilibrium N2O remained largely the same as for the previous model.
+
 ```{r conditional_effects_mod_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 
-The covariate effects on $\sigma$ for N2O-eq suggested an negative effect of surface temperature and litte to no effect of elevation.
+The covariate effects on $\sigma$ for N2O-eq suggested a negative effect of surface temperature and litte to no effect of elevation.
+
 ```{r conditional_effects_sigma_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 ## Model 5
-In the next model, more complexity is added to the N2O component by including a covariate for lake surface area (log scale) as well as interactions between NO3 and log(area) and surface temperature. 
+In the next model, additional complexity is added to the dissolved N2O component by including a covariate for continuous lake surface area (log scale) as well as interactions between NO3 and log(surface area) and surface temperature.
+
 ```{r n2o_mod_mv_5, eval=FALSE, include=TRUE}
@@
 Below is a summary of the fitted parameters along with MCMC convergence diagnostics.
+
 ```{r print_mod5, echo=FALSE}
@@
 Again, the same PPCs as above were performed for this model.
+
 #### Dissolved N2O
-This PPC for N2O looked similar to the previous model.
+This PPCs for dissolved N2O looked similar to the previous model.
+
 ```{r ppc_full_checks_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Again, the PPCs for this model were similar to the previous model, which was unsurprising given that it was the same model for N2O-eq.
+
 ```{r ppc_full_checks_mod_n2oeq5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 This PPC was also similar to the previous model.
+
 ```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 This check was also similar to the prevoius model, with perhaps slightly less bias in the proportion unsaturated estimates.
-There is also a potentially concerning extreme prediction in the observed _vs_ predicted PPC.
+There was also a potentially concerning extreme prediction in the observed _vs_ predicted PPC.
+
 ```{r ppc_sat_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### R-square
-```{r r2_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+
+```{r r2_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
@@
 #### Dissolved N2O
-The conditional effects plot for the covariate effects N2O suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature.
+The conditional effects plots suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature.
 For lake area, the effect was estimated to be larger and more negative at the highest levels of NO3; and slightly negative at the lowest level of NO3.
 For surface temperature, the effect was estimated to be largest and positive at the highest level of NO3; and negative at the lowest level of NO3.
+
 ```{r conditional_effects_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4, cache=TRUE}
@@
 The estimated covariate effects on $\sigma$ suggested a negative relationship with log(area) and a positive relationship, again, with NO3.
+
 ```{r conditional_effects_sigma_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 #### Equilibrium N2O
-The estimated covariate effect on N2O remained largely the same as estimated in the previous model.
+The estimated covariate effects on equilibrium N2O remained largely the same as estimated in the previous model.
+
 ```{r conditional_effects_mod_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 ## A Final Model
-As demonstrated above, models excluding the NO3 covariate consistently resulted in poorer fits to to the observed dissolved N2O data.
-Including surface temperature and elevation in the equilibrium N2O part of the model resulted in substantially improved replication of key aspects of the observed data.
+As demonstrated above, models excluding the NO3 covariate consistently resulted in poorer fits to to the observed dissolved N2O data and potentially strongly biased estimates of the saturation ratio.
+Including surface temperature and elevation in the equilibrium N2O part of the model also resulted in substantially improved replication of key aspects of the observed data.
 Likewise, added flexibility in the distributional terms for both dissolved and equilibrium N2O led to improvements.
 
-To make inferences from this model for N2O in the population of interest, however, the included covariates needed to be (1) fully observed across that population or (2) their missingness needed to be modeled.
+To make inferences to the population of interest from a model including these covariates, however, however, the covariates needed to be (1) fully observed across that population or (2) their missingness needed to be modeled.
 For the lake area and elevation covariates, data _was_ available for all lakes from previously compiled geospatial databases.
 However, neither surface temperature or NO3 were observed for lakes outside of the sample.
-They were partially observed with respect to the target population.
-Their missingness needed to be accounted for in a model.
+That is, they were partially observed with respect to the target population.
 Therefore, a more complex model was constructed below that included surface temperature and NO3 as additional responses conditioned on the survey design variables and fully observed covariates.
-This approach to inference for N2O was similar to a Bayesian structural equation model [@Merkle_etal_2021; @Merkle_Rosseel_2018].
+This approach to inference for N2O was similar to Bayesian structural equation modeling approaches [@Merkle_etal_2021; @Merkle_Rosseel_2018].
 The logical dependence structure could be characterized as:
 
-$$\begin{align} 
+$$\begin{aligned} 
 {\boldsymbol{N_2O_{diss}}} &\sim Survey + Area + {\boldsymbol{NO_3}} + {\boldsymbol{Temp}} \\ 
@@
 {\boldsymbol{Temp}} &\sim Survey + Lat + Elev + Day
-\end{align}$$
+\end{aligned}$$
 
@@
 The partially observed variables, being dissolved and equilibrium N2O, NO3, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.
-This structural equation approach requires a more complex set of post-processing steps compared to a typical MRP analysis.
+This piece-wise approach required a more complex set of post-processing steps compared to a typical MRP analysis.
 In order to propagate estimates and uncertainty through the dependency structure and make inferences, the fitted model was used to first predict surface temperature in the target population, since it depended only on the fully observed covariates.
@@
 The shape parameter was also modeled as a function of latitude to address increasing response variance along the latitudinal gradient.
-The NO3 sub-model was a cumulative logit formulation and the linear predictor included all of the survey factors as well as surface temperature and lake area.
-
-The N2O and N2O-eq responses were each modeled with Gamma distributed errors, but with the same covariate structure as in model 5.
-The same structure was also employed for the shape terms in these responses, corresponding to the $\sigma$ terms in the previous model.
+The NO3 sub-model was a cumulative logit model and the linear predictor included all of the survey factors as well as surface temperature and lake area.
+
+The dissolved and equilibrium N2O responses were each modeled with Gamma distributed errors, but with the same covariate structure as in model 5.
+The same structure was also employed for the shape terms in these responses, corresponding to the $\sigma$ terms in the previous models.
 Though not shown in this document, the Gamma error structure appeared to result in slightly better performance in the predictive checks compared to the Gaussian errors in previous models.
@@
 
-Note that there was no residual correlation term for this model, since the residuals are undefined for the Gamma and cumulative logit models.
-Dropping the observation-level residual correlation term was deemed a reasonable compromise that enabled modeling the missingness of NO3, in particular.
-Nevertheless, the random intercepts again allowed for potential correlations between responses at the group levels.
+Note that there was no observation-level residual correlation term for this model, since the residuals are undefined for the Gamma and cumulative logit models.
+Dropping the observation-level residual correlation term was deemed a reasonable compromise that enabled the inclusion of NO3 as a covariate on dissolved N2O.
+The random intercepts, however, still allowed for potential correlations between the four responses at the group levels.
 
@@
 ### Model checks
-Below, the same PPCs for N2O and N2O-eq were employed as before.
+Below, the same PPCs for dissolved and equilibrium N2O were used as before.
+
 #### Dissolved N2O
-The PPCs for N2O from this model were similarly reasonable as for models 4 and 5 above.
+The PPCs for dissolved N2O were similar to those for models 4 and 5 above.
+
 ```{r ppc_full_checks_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### Equilibrium N2O
-Again, the PPCs for N2O-eq in this model were similar to those for models 4 and 5.
+The PPCs for eqilibrium N2O were also similar to the same checks in models 4 and 5.
+
 ```{r ppc_full_checks_mod_n2oeq6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### Bivariate
-This model again provided a very reasonable representation of the bivariate relationship between N2O and N2O-eq (below).
+This model provided a reasonable representation of the bivariate relationship between the two N2O responses.
+
 ```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### Saturation
-The saturation ratio PPCs below show similar behavior as with models 4 and 5 above, but with perhaps slightly less bias in the predictions for the proportion of undersaturated waterbodies and fewer extreme predictions for the means and standard deviations.
-The observed _vs._ predicted PPC also appears to have a better behaved variance and no extreme predictions, compared to models 4 and 5 with the lognormal errors.
+The saturation ratio PPCs below suggested similar behavior as with models 4 and 5 above, but with perhaps slightly less bias in the predictions for the proportion of undersaturated waterbodies and fewer extreme predictions for the means and standard deviations.
+The observed _vs._ predicted PPC also appears to have a better behaved variance and no extreme predictions, compared to models 4 and 5.
+
 ```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 
-The plot below shows the same PPC, but for the "test" or second-vist data. Overall, the model looked to perform similarly as with the data used to fit it.
+The plot below shows the same PPC, but for the second-vist data.
+
 ```{r ppc_sat_check_testdata_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 #### R-square
-Below are estimates for the Bayesian $R^2$, which were largely similar for N2O and N2O-eq as with models 4 and 5 above. The $R^2$ for the surface temperature response also suggested a fairly good fit.
-```{r r2_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+Below are estimates for the Bayesian $R^2$, which were largely similar for the N2O responses as with models 4 and 5 above. The $R^2$ for the surface temperature response also suggested a good fit.
+
+```{r r2_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 
-Below are the same $R^2$ estimates, but for the second-visit data.
-That these estimates are similar to those for the data used to fit the model, suggesting that the model may perform similarly well out-of-sample.
-```{r r2_test_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=4, fig.height=4}
+Below are the $R^2$ estimates for the second-visit data.
+That these estimates were similar to those for the data used to fit the model was encouraging and suggested that the model may perform reasonably well out-of-sample.
+
+```{r r2_test_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
 load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
@@
 #### Dissolved N2O
-The conditional effects plot for the covariate effects N2O suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature.
-For lake area, the effect was estimated to be larger and more negative at the highest levels of NO3; and slightly negative at the lowest level of NO3.
-For surface temperature, the effect was estimated to be largest and positive at the highest level of NO3; and negative at the lowest level of NO3.
+The conditional effects plot for the covariate effects on dissolved N2O suggested a similar effect of NO3 as in previous models, but with interesting potential interactions between NO3 and lake area and NO3 and surface temperature.
+The lake area effect was estimated to be larger and more negative at the highest levels of NO3 and slightly negative at the lowest level of NO3.
+The surface temperature effect was estimated to be largest and positive at the highest level of NO3 and negative at the lowest level of NO3.
+
 ```{r conditional_effects_mod_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4, cache=TRUE}
@@
 
-The estimated covariate effects on $\sigma$ suggested a negative relationship with log(area) and a positive relationship, again, with NO3.
+The estimated covariate effects on $\sigma$ for dissolved N2O suggested a negative relationship with log(area) and a positive relationship with NO3.
+
 ```{r conditional_effects_sigma_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 #### Equilibrium N2O
-The estimated covariate effect on N2O remained largely the same as estimated in the previous model.
+The estimated covariate effects on equilibrium N2O remained largely the same as estimated in the previous model.
+
 ```{r conditional_effects_mod_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
@@
 
-The first step used the final model to predict to the population:
+The first step used the final model to predict surface temperature to the population:
+
 ```{r predict_obsframe, eval=FALSE, include=TRUE}
@@
 
-NO3 was next predicted. Note that the posterior predictive distribution for NO3 was subsampled in order to minimize excess simulations
+NO3 was next predicted using the posterior predictions for surface temperature and the other fully observed covariates.
+Note that the posterior predictive distribution for NO3 was subsampled in order to minimize excess simulations.
+
 ```{r parallel_predict_draws, eval=FALSE, include=TRUE}
@@
 
-Finally, N2O and N2O-eq were predicted using the surface temperature and nitrate predictions along with the survey variables and known covariates.
+Finally, dissolved and equilibrium N2O were predicted using the surface temperature and NO3 predictions along with the survey variables and known covariates.
 Again, the posterior was subsampled in order to reduce excess simulations.
+
 ```{r n2o_covariates_X, eval=FALSE, include=TRUE}
@@
 
-Finally, the predictions for all four partially observed responses were assembled into a new dataframe for use in inference.
+Finally, the predictions for all four partially observed responses were assembled into a new dataframe for use in inference:
+
 ```{r assemble_predictions, eval=FALSE, include=TRUE}
@@
 A number of estimates for the target population were assembled and presented below.
-First, the full posterior predictive distributions for dissolved N2O, equilibrium N2o, and the saturation ratio were assessed.
+First, the full posterior predictive distributions for dissolved N2O, equilibrium N2O, and the saturation ratio were assessed.
 These distributions summarized the predicted distribution of concentrations or ratios for all lakes in the population of interest and included parameter uncertainty propagated through the model.
@@
 ## Posterior predictive distributions
-Below, a density plot summarized the posterior predictive distribution of N2O and N2O-eq concentrations across the target population of lakes, based on 500 draws from the posterior predictive distribution.
-Note that the x-axis was truncated at 50 nmol/L for a clearer visualization of the bulk of the predicted distribution.
+Below, a density plot summarized the posterior predictive distributions across the target population of lakes.
+The PPDs consisted of based on 500 simulations for each variable.
+Note that the x-axis was truncated at 50 nmol/L for a clearer visualization of the bulk of the predictive distribution.
 For reference, the max predicted value was 4403.2 nmol/L for dissolved N2O, 20.4 nmol/L for dissolved N2O, and 793.5 for the saturation ratio.
+
 ```{r plot_n2o_posterior_preds, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE, cache.lazy=FALSE}
@@
 ### National
-Below are density plots summarizing the posterior distribution of _means_ for N2O concentrations and the saturation ratio for the target population (i.e., all US lakes > 1ha in the lower 48 states).
+Below are density plots summarizing the posterior distribution of _means_ for N2O concentrations and the saturation ratio for all US lakes and reservoirs > 4ha in the lower 48 states).
+
 ```{r plot_n2o_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 To illustrate the skewness in the predictive distribution for the saturation ratio, an estimate for the median ratio is shown below.
-The entire posterior distribution of the mean above is larger than 1, the ratio representing the boundary of under- _vs._ oversaturation.
-By comparison, the posterior estimate of the median below only included values less than one, suggesting that though the mean saturation ratio was greater than 1, most lakes in the national populaiton were undersaturated (i.e., ratio less than 1).
-In distributions with right-skew, the mean can often be considerably larger than the median.
+The entire posterior distribution of the mean above was larger than 1, which represents the boundary of under- _vs._ oversaturation.
+By comparison, the posterior estimate of the median below only included values less than one, suggesting that though the mean saturation ratio was likely greater than 1, most lakes in the national population were undersaturated (i.e., ratio less than 1).
+In distributions with strong right-skew, the mean can often be considerably larger than the median.
+
 ```{r plot_sat_nat_posterior_median, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Below is a plot of the posterior mean estimate for the proportion of unsaturated lakes at the national scale.
+
 ```{r plot_undersat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Below are posterior estimates of the means for dissolved and equilibrium N2O and the saturation ratio by WSA9 ecoregion.
+
 ```{r plot_n2o_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 A plot of the posterior estimates for the median saturation ratio below indicated, again, that most lakes in each ecoregion were undersaturated (i.e., median << 1).
+
 ```{r plot_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 A plot of the estimates of the proportion of under-saturated lakes by ecoregion is below.
-A plot of the posterior estimates for the median saturation ratio below indicated, again, that most lakes in each ecoregion were undersaturated (i.e., median << 1).
+These summaries again suggested that most lakes in each ecoregion were likely undersaturated (i.e., median << 1).
+
 ```{r plot_prop_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 ### State
-Comparisons of mean estimates (posterior median, upper and lower 95th percentiles) by state are below. Density estimates were not included to minimize plot space.
+Comparisons of mean estimates by state are below. Density estimates were not included to minimize the vertical plot space.
+
 ```{r plot_state_mean_n2o, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 
-Below, a plot of estimates for the mean (black circles) and median (grey circles) saturation ratio by state.
+Below is a plot of estimates for the mean (black circles) and median (grey circles) saturation ratio by state.
 A horizontal, dashed, black line is shown at ratio = 1, indicating the boundary for under- _vs._ oversaturation.
 Only a few states (e.g., NV, DE) had median estimates that were 1 or greater, suggesting that, for most states, most lakes were undersaturated.
+
 ```{r plot_state_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Point estimates are the posterior median of the proportion and bars are the upper and lower boundaries of the central 95th percentile of the posterior distributions of proportions.
+
 ```{r plot_state_prop_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 ### Size category
-The estimated means by size category are below for dissolved and equilibrium N2O and the saturation ratio. Median estimates for the saturation ratio are also shown.
+The estimated means and medians by size category are below for dissolved and equilibrium N2O and the saturation ratio.
+
 ```{r plot_n2o_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Mean _vs._ median below.
+
 ```{r plot_size_cat_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
@@
 
-And, finally, the estimated proportion of undersaturated lakes in the target population by size category
+And, finally, the estimated proportion of undersaturated lakes in the target population by size category.
+
 ```{r plot_prop_sat_size_cat_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 ## Model- _vs._ design-based
-Below, estimates from the model-based approach are compared to design-based estimates.
-In general, the model estimates were similar to the design-based estimates.
-Model estimates were typically within the confidence bounds of the design-based estimates, but with much greater precision.
-Improved precision was expected due to the "shrinkage" induced by the multilevel parameterization, which allowed some "borrowing" of information across the various levels of the survey factors.
+Below, estimates from the model-based approach are compared to previously calculated, design-based estimates.
+In general, the model-based estimates were similar to the design-based estimates.
+The model-based estimates were typically within the confidence bounds of the design-based estimates, but with much greater precision.
+Improved precision was expected due to the "shrinkage" induced by the multilevel parameterization, which affords some "borrowing" of information across the various levels of the survey factors.
 
 ### Dissolved N2O
-Below, National mean estimates for dissolved N2O from the model and design-based approaches were compared. The sample-based estimate was also included as a naive reference.
+Below, National mean estimates for dissolved N2O from the model and design-based approaches were compared.
+The sample-based estimate was also included as a reference.
+The black, vertical, dashed line represents the mean of the sample.
+
 ```{r n2o_means_national, message=FALSE, cache=TRUE}
@@
 
-The black, vertical, dashed line in the figure below represents the mean of the sample. 
 ```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Below, estimates were compared by ecoregion.
+
 ```{r n2o_mean_wsa9, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Means were compared according to size categories below.
+
 ```{r table_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
@@
 Below, the same comparisons were made for the saturation estimates.
+
 ```{r table_nat_sat_mean, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8, cache=TRUE}
@@
 ```
+
 ```{r plot_wsa9_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
@@
 # Session Info
+
 ```{r session}

### 2023-03-22 manuscript/manuscript_file.Rmd
setting up manuscript infrastructure

@@
 output: word_document
-date: "2023-03-08"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 ---
@@
 
-In this study we present the first national-scale survey **kasdflkjldfj** of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S.
+In this study we present the first national-scale survey of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S.
 (CONUS).
@@
 Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?).
-`r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N2O sinks.
-We modeled the data using a bayesion hierchical approach and predicted the dissolved N2O concentration for all 224,916 CONUS waterbodies in the target population.
-At the population level, `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% pull()` percent (95% CI: `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% pull()` - `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% pull()`) of CONUS waterbodies were functioning as N2O sink.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N2O sinks (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "WMT") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()`) and the Northern Plains had the greatest (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "NPL") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()`)).
+`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N~2~O sinks (Figure 1).
+We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population.
+At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`).
 
@@
 # [3/9/2023] knitr can only find the file if placed in the root of the RStudio project.  Why can't it pull the figure from the manuscript_figures folder?
-knitr::include_graphics("figure1.png")
+#knitr::include_graphics("manuscript_figures/figure1.png")
 ```

### 2023-03-22 scripts/dgIndicatorAnalysis.Rmd
setting up manuscript infrastructure

@@
 
-## micromap: N2O saturation ratio continuous, extent, and emission rate
+## MICROMAP: N2O saturation ratio continuous, extent, and emission rate
+
+### Set up data
 ```{r}
@@
 
-ecoreg_stats$WSA9_NAME <- as.character(ecoreg_stats$WSA9)
-ecoreg_stats <- ecoreg_stats %>% select(-WSA9)
+ecoreg_stats$WSA9 <- as.character(ecoreg_stats$WSA9)
 str(ecoreg_stats)
@@
 
-# N2O SATURATION AND UNDERSATURATION EXTENT
+```
+
+### N2O SATURATION AND UNDERSATURATION EXTENT
+```{r}
 # Full code:  Use this to produce final linked micromaps of the variables you want
@@
        print.file = "output/figures/mmapSatExtent.tiff", print.res = 300)
-
-# N2O EMISSION AND FLUX
+```
+
+### N2O EMISSION AND FLUX
+```{r}
 # Full code:  Use this to produce final linked micromaps of the variables you want

### 2023-03-22 manuscript/manuscript_support.R

@@ -70,42 +70,105 @@ toc()
 
 
 # FIGURES-----------------
-# eliminate black points from with WSA9_NAME legend
-# rename WSA9_NAME legend to Ecoregions
 
 
-ggplot() +
-  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-  geom_sf(data = dg.sf %>% filter(!is.na(n2o.src.snk)), 
-          aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
-          show.legend = "point") +
-  geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-  # guide argument below removes points from the boxes in the ecoregion legend
-  # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-  scale_fill_manual("Ecoregion", values = cols, 
-                    guide = guide_legend(override.aes = list(shape = NA))) +
-  scale_color_manual(values = c("white", "black"), name = "source/sink") +
-  scale_size(name = expression(N[2]*O~(nM)),
-             range = c(0.1, 10), # custom size range
-             breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-  theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-        legend.title = element_text(size = 8))
-ggsave("manuscript/manuscript_figures/figure1.png", width = 8, height = 4, units = "in")
+# # Figure 1
+# # Commented to expedite script
+# ggplot() +
+#   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+#   geom_sf(data = dg.sf %>% filter(!is.na(n2o.src.snk)), 
+#           aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
+#           show.legend = "point") +
+#   geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+#   # guide argument below removes points from the boxes in the ecoregion legend
+#   # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+#   scale_fill_manual("Ecoregion", values = cols, 
+#                     guide = guide_legend(override.aes = list(shape = NA))) +
[…]

### 2023-03-23 manuscript/manuscript_file.Rmd
Resolved issue with fetching graphics. Added infrastructure for references.

@@
 title: "Widespread nitrous oxide undersaturation in U.S. lakes and reservoirs"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: word_document
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+bibliography: references.bib
+link-citations: yes
 ---
@@
 
-## Introduction
+# Introduction
 
-## Results and Discussion
+# Results and Discussion
 
@@
 
-```{r echo=FALSE, fig.cap= "Location of sampling sites.
+![Location of sampling sites.
 Color of the points indicates whether the waterbody was functioning as a source or sink of nitrous oxide (N~2~O).
-Point size indicates the dissolved N~2~O concentration"}
+Point size indicates the dissolved N~2~O concentration.](manuscript_figures/figure1.png)
 
-# [3/9/2023] knitr can only find the file if placed in the root of the RStudio project.  Why can't it pull the figure from the manuscript_figures folder?
-#knitr::include_graphics("manuscript_figures/figure1.png")
-```
+Here is a citation example: Multilevel models are useful for survey inference [@Kennedy_Gelman_2021]. This paper by @Kennedy_Gelman_2021 provides a nice worked example.
+
+# References

### 2023-03-23 manuscript/manuscript_support.R

@@ -74,23 +74,23 @@ toc()
 
 # # Figure 1
 # # Commented to expedite script
-# ggplot() +
-#   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-#   geom_sf(data = dg.sf %>% filter(!is.na(n2o.src.snk)), 
-#           aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
-#           show.legend = "point") +
-#   geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-#   # guide argument below removes points from the boxes in the ecoregion legend
-#   # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-#   scale_fill_manual("Ecoregion", values = cols, 
-#                     guide = guide_legend(override.aes = list(shape = NA))) +
-#   scale_color_manual(values = c("white", "black"), name = "source/sink") +
-#   scale_size(name = expression(N[2]*O~(nM)),
-#              range = c(0.1, 10), # custom size range
-#              breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-#   theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-#         legend.title = element_text(size = 8))
-# ggsave("manuscript/manuscript_figures/figure1.png", width = 8, height = 4, units = "in")
+ ggplot() +
+   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+   geom_sf(data = dg.sf %>% filter(!is.na(n2o.src.snk)), 
+           aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
+           show.legend = "point") +
+   geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+   # guide argument below removes points from the boxes in the ecoregion legend
+   # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+   scale_fill_manual("Ecoregion", values = cols, 
+                     guide = guide_legend(override.aes = list(shape = NA))) +
+   scale_color_manual(values = c("white", "black"), name = "source/sink") +
+   scale_size(name = expression(N[2]*O~(nM)),
+              range = c(0.1, 10), # custom size range
+              breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+   theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
[…]

### 2023-03-23 manuscript/manuscript_file.Rmd
Added figures and text to manuscript files

--- 2023-03-22 manuscript/manuscript_file.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-
 # Specify paths-----
@@
 localPath <- Sys.getenv("USERPROFILE")
+knitr::opts_knit$set(root.dir = "../")
 
@@
 
+```
+
+```{r source_scripts, echo=FALSE, warning=FALSE, message=FALSE}
+# currently knitting messages and warnings into .docx.  hopefully warning and
+# message setting above will fix
+
 # source scripts------
@@
 # see above for .Rmd working directory
-source("manuscript/manuscript_support.R") 
+source("manuscript/manuscript_support.R", local = knitr::knit_global())
 ```
@@
 
+### N~2~O undersaturation extent  
 In this study we present the first national-scale survey of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S.
@@
 Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?).
-`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N~2~O sinks (Figure 1).
+`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks (Figure 1).
 We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population.
 At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`).
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`), the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), and >65% of the waterbodies in every ecoregion were undersaturated (Table 1).
 
-```{r echo=FALSE, fig.cap= "Location of sampling sites.
+The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
+This finding  runs counter to the long standing paradigm that nitrogen rich surface waters are a source of N~2~O to the atmosphere (Mosier et al.
+1998) and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
+Webb et al.
+(2019) found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Soued et al (2015) report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks.
+While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
+This pattern holds across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use across the continental U.S..
+
+```{r figure1, echo=FALSE, fig.cap= "Location of sampling sites.
 Color of the points indicates whether the waterbody was functioning as a source or sink of nitrous oxide (N~2~O).
@@
 # [3/9/2023] knitr can only find the file if placed in the root of the RStudio project.  Why can't it pull the figure from the manuscript_figures folder?
-#knitr::include_graphics("manuscript_figures/figure1.png")
+knitr::include_graphics("manuscript/manuscript_figures/figure1.png")
 ```
+
+```{r table1, echo = FALSE}
+rbind(national_stats %>% 
+        select(WSA9, estimate, LCL, UCL, post_med_sat, LCI_sat, UCI_sat, post_med_en2o, LCL_en2o, UCL_en2o),
+      ecoreg_stats %>% 
+        select(WSA9, estimate, LCL, UCL, post_med_sat, LCI_sat, UCI_sat, post_med_en2o, LCL_en2o, UCL_en2o)) %>%
+  mutate('N2O undersaturation extent' = paste0(estimate*100, " (", LCL*100, "-", UCL*100, ")"),
+         'N2O saturation ratio' = paste0(post_med_sat, " (", LCI_sat, "-", UCI_sat, ")"),
+         'N2O emission rate' = paste0(post_med_en2o, " (", LCL_en2o, "-", UCL_en2o, ")")) %>% 
+  select(WSA9, 'N2O undersaturation extent', 'N2O saturation ratio', 'N2O emission rate') %>%
+  rename('spatial domain' = WSA9) %>%
+knitr::kable(caption = "Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, and emission rate at the national and ecoregional scales")
+
+```
+
+### N~2~O saturation ratio
+#### Distribution
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean N~2~O saturation ratio (ratio of measured to equilibrium concentration) was equal to or greater than 1 in seven of the nine ecoregions (Table 1, Figure 2), suggesting that, on average, waterbodies in these ecoregions were supersatured with N~2~O and therefore functioning as a source of N~2~O.
+The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations that exceed equilibrium in most ecoregions can be attributed to the distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies (Figure 2).
+Across the CONUS, and within each of the nine major ecoregions, the distribution of N~2~O saturation ratios had median values that were less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
+Thus most waterbodies in the CONUS are functioning as N~2~O sinks, but a few function as strong N~2~O sources.
+
+### Controls on N~2~O saturation
+N~2~O saturation was related to waterbody size in several ecoregions, but the direction of the relationship varied.
+For example, N~2~O saturation increased with waterbody size in the Northern Plains, Upper Midwest, Northern Appalachians, and Western Mountain ecoregions, but decreased with increasing size in the Coastal and Southern Plains (Figure 3).
+A pattern common to many of the ecoregions, regardless of the direction of the relationship between N~2~O saturation and waterbody size, is that N~2~O saturation deviates from equilibrium most strongly in the smallest waterbody size class.
+When pooled across all ecoregions, the absolute magnitude of the difference beween the measured and equilibrium N~2~O concentration, known as delta N~2~O

### 2023-03-23 scripts/dgIndicatorAnalysis.Rmd
Added figures and text to manuscript files

@@
 
-```{r}
+```{r calculate_emissions}
 
@@
 
-all_predictions <- all_predictions %>%
+all_data <- all_predictions %>%
   mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
@@
 
-# # write reduced file to disk for less computationally intensive wrangling
-# all_predictions.w.small <- all_predictions %>% filter(.draw %in% 1:5)
-# 
-# save(all_predictions.w.small,
-#      file = paste0(Sys.getenv("USERPROFILE"),
-#                    "/Environmental Protection Agency (EPA)/",
-#                    "ORD NLA17 Dissolved Gas - Documents/",
-#                    "inputData/all_predictions_w_small.rda"))
 ```

### 2023-03-23 manuscript/manuscript_support.R

@@ -3,25 +3,26 @@
 
 
 # SETUP----------------------
-# load libraries-----------------
+## load libraries
 library(sf) # spatial data
 library(tidyverse) # dplyr, ggplot
 library(janitor) # clean names
 library(USAboundaries) # state boundaries
 library(tictoc) # processing time
 
-# Identify local path for each user
+## Identify local path for each user
 localPath <- Sys.getenv("USERPROFILE")
 
 
 # DATA---------------
-#### load sample data (dg object)
+## load sample data (dg object)----
 load(paste0(localPath,  # object name dg
             "/Environmental Protection Agency (EPA)/",
             "ORD NLA17 Dissolved Gas - Documents/",
             "inputData/dg.2021-02-01.RData"))
 
-# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
 # Define coordinates
 coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
 
@@ -30,19 +31,20 @@ dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"),
   st_transform(5070) # project to CONUS ALBERS for plotting
 
 
-##### load ecoregions
+## load ecoregions----
 # read in ecoregion polygons
 ecoR <- st_read(dsn = paste0(localPath, 
                              "/Environmental Protection Agency (EPA)/",
                              "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple")
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
 
 # Check CRS
 st_crs(ecoR) # 3857
 ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
 st_crs(ecoR) # 5070
 
-# SET UP CUSTOM COLORS FOR ECOREGIONS---------
+# SET UP CUSTOM COLORS FOR ECOREGIONS
 # Custom color pallette for ecoregion polygons. Attempted to mirror
 # https://www.epa.gov/national-aquatic-resource-surveys/
 # ecoregional-results-national-lakes-assessment-2012
@@ -57,46 +59,229 @@ cols <- c("Coastal Plains" = "orange1",
           "Xeric" = "lightskyblue4")
 
 
[…]

### 2023-03-23 manuscript/manuscript_support.R
Minor improvements to manuscript files

@@ -207,7 +207,7 @@ national_stats <- all_predictions %>%
 
 
 # FIGURES-----------------
-<<<<<<< HEAD
+
 # The images are written to manuscript/manuscript_figures/...
 # the first time these scripts are run.  Unless the code has changed,
 # there is no need to run the scripts again.  Commenting out the script
@@ -278,30 +278,7 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   ggsave("manuscript/manuscript_figures/figure2.tiff", width = 8.5, height = 5)
   
 }
-=======
 
-
-# # Figure 1
-# # Commented to expedite script
- ggplot() +
-   geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-   geom_sf(data = dg.sf %>% filter(!is.na(n2o.src.snk)), 
-           aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
-           show.legend = "point") +
-   geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-   # guide argument below removes points from the boxes in the ecoregion legend
-   # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-   scale_fill_manual("Ecoregion", values = cols, 
-                     guide = guide_legend(override.aes = list(shape = NA))) +
-   scale_color_manual(values = c("white", "black"), name = "source/sink") +
-   scale_size(name = expression(N[2]*O~(nM)),
-              range = c(0.1, 10), # custom size range
-              breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-   theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-         legend.title = element_text(size = 8))
- ggsave("manuscript/manuscript_figures/figure1.png", width = 8, height = 4, units = "in")
-
->>>>>>> e42d262904b8409cff9d0f19a94d7534ecfdbf87
 
 
 # MANUSCRIPT DATA-----

### 2023-03-28 manuscript/manuscript_support.R
fixed issues with invalid characters in code

@@ -207,7 +207,7 @@ national_stats <- all_predictions %>%
 
 
 # FIGURES-----------------
-<<<<<<< HEAD
+
 # The images are written to manuscript/manuscript_figures/...
 # the first time these scripts are run.  Unless the code has changed,
 # there is no need to run the scripts again.  Commenting out the script
@@ -278,7 +278,7 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   ggsave("manuscript/manuscript_figures/figure2.tiff", width = 8.5, height = 5)
   
 }
-=======
+
 
 
 # # Figure 1
@@ -300,8 +300,6 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
  ggsave("manuscript/manuscript_figures/figure1.png", width = 8, height = 4, units = "in")
-
->>>>>>> e42d262904b8409cff9d0f19a94d7534ecfdbf87
 
 
 # MANUSCRIPT DATA-----

### 2023-04-12 manuscript/manuscript_file.Rmd
Incremental progress on manuscript.  Added discussion of N2O saturation vs lake size

@@
 localPath <- Sys.getenv("USERPROFILE")
-knitr::opts_knit$set(root.dir = "../")
+knitr::opts_knit$set(root.dir = "../") # set wd to root of RStudio project
 
@@
 
-# Working directory------
-# wd is set to root of RStudio project via options under Knitr drop-down
-# menu above.
-
 ```
@@
 ```{r source_scripts, echo=FALSE, warning=FALSE, message=FALSE}
-# currently knitting messages and warnings into .docx.  hopefully warning and
-# message setting above will fix
-
 # source scripts------
@@
 ```
+
+## NOTES
+* Table and figure captions should specify if displaying sample vs population estimates.
 
@@
 
+![Figure 1.
+Location of sampling sites.
+Color of the points indicates whether the waterbody was functioning as a source or sink of nitrous oxide (N~2~O).
+Point size indicates the dissolved N~2~O concentration.](manuscript_figures/figure1.png)
+
 The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
@@
 Webb et al.
-(2019) found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Soued et al (2015) report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks.
+(2019) found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Soued et al (2015) report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020)  but kortelainen and Soued are mostly pristine lakes]**.
 While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
 This pattern holds across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use across the continental U.S..
-
-```{r figure1, echo=FALSE, fig.cap= "Location of sampling sites.
-Color of the points indicates whether the waterbody was functioning as a source or sink of nitrous oxide (N~2~O).
-Point size indicates the dissolved N~2~O concentration"}
-
-# [3/9/2023] knitr can only find the file if placed in the root of the RStudio project.  Why can't it pull the figure from the manuscript_figures folder?
-knitr::include_graphics("manuscript/manuscript_figures/figure1.png")
-```
 
@@
   rename('spatial domain' = WSA9) %>%
-knitr::kable(caption = "Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, and emission rate at the national and ecoregional scales")
+knitr::kable(caption = "Table 1. Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, and emission rate at the national and ecoregional scales")
 
 ```
+
+Here is a citation example: Multilevel models are useful for survey inference [@Kennedy_Gelman_2021]. This paper by @Kennedy_Gelman_2021 provides a nice worked example.
 
@@
 
-### Controls on N~2~O saturation
+![Figure 2.
+Distribution of N~2~O saturation values for the population of waterbodies in the nine major ecoregions.
+Each blue line (n=500) represents the distribution of values from a single realization of the posterior prediction.
+Vertical black dashed and solid lines indicated median and mean, respectively.
+Vertical red line indicates a saturation ratio of 1.](manuscript_figures/figure2.tiff)
+
+#### Controls on N~2~O saturation
 N~2~O saturation was related to waterbody size in several ecoregions, but the direction of the relationship varied.
-For example, N~2~O saturation increased with waterbody size in the Northern Plains, Upper Midwest, Northern Appalachians, and Western Mountain ecoregions, but decreased with increasing size in the Coastal and Southern Plains (Figure 3).
-A pattern common to many of the ecoregions, regardless of the direction of the relationship between N~2~O saturation and waterbody size, is that N~2~O saturation deviates from equilibrium most strongly in the smallest waterbody size class.
-When pooled across all ecoregions, the absolute magnitude of the difference beween the measured and equilibrium N~2~O concentration, known as delta N~2~O
+For example, N~2~O saturation increased with waterbody size in the Northern Plains, Upper Midwest, Northern Appalachians, and Western Mountain ecoregions, but decreased with increasing size in the Coastal and Southern Plains (Figure 3A).
+The inconsistent relationship between N~2~O saturation and waterbody size may be attributable differences in nitrate concentrations across ecoregions.
+The model identified an interaction between waterbody surface area and nitrate concentration (expressed categorically) where N~2~O concentration decreases with increasing lake size at high NO3 concentrations, but increases, or remains constant at low nitrate concentrations.
+THIS IS BECAUSE THERE CAN'T BE MUCH N2O BIOGEOCHEMISTRY WITHOUT INORGANIC N LAYING AROUND.
+
+![Figure 3.
+Relationship between delta N~2~O and lake size category across A) the population of CONUS waterbodies and B) within each of the nine major CONUS ecoregions.
+Red arrows in panel A illustrate the magnitude of N~2~O disequilibrium in the smallest lake size class.
+Data are inferences made to the population of CONUS waterbodies.](manuscript_figures/figure3.tiff)
+
+![Figure 4.
+Partial dependence plot illustrating the relationship between N~2~O concentration and waterbody surface area as a function of nitrate concentration.](manuscript_figures/N2O vs surface area x NO3 partial dependence plot.png)
+
+Regardless of the direction of the relationship between N~2~O saturation and waterbody size, N~2~O saturation deviates from equilibrium most strongly in the smallest waterbody size class.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases linearly with increasing waterbody size class (Figure 3B).
+Waterbody size can affect dissolved gas concentration through several mechanisms.
+In lakes with low volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column.
+Since water volume to benthic surface area ratios tend to scale positively with lake surface area, benthic N~2~O biogeochemistry will have a greater affect on dissolved N~2~O concentration in small lakes, regardless of whether the sediments function as a net sink or source of N~2~O **[see Wetzel for references]**.
+Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes **[need a citation for N2O ~ temperature relationship]**.
+Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (Beaulieu et al N2O Harsha Lake and USACE N2O papers).
+
+Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequlibrim in smaller waterbodies.
+Air-water gas exchange will vent excess N~2~O from surface waters or replenish depleted N~2~O in undersaturated waterbodies by facilitating the invasion of atmospheric N~2~O.
+Regardless of whether a waterbody is superaturated or undersatured in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
+One of the primary drivers of air-water gas exchange in lentic waters is wind speed (Vachon and Prairie reference).
+Wind speeds to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
+Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the affect of biogeochemistry on dissolved N~2~O is more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
+
+**Not sure where I am going with this paragraph.  Maybe scratch it.  Mostly included as a placeholder for these references**
+While previous surveys have documented relationships between lake size and dissolved methane concentration (Pighini et al.
+2018, Juutinen et al 2009, Borges et al.
+2011), another greenhouse gas, to the best of our knowledge, this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
+Perhaps because previous surveys have included little variation in waterbody size (Webb et al.
+2019, Kortelainen et al) or water quality (Soued et al.
+2015).
+
+# References

### 2023-04-13 manuscript/manuscript_file.Rmd
Added methodology, references, reference sytle file csl

--- 2023-03-23 manuscript/manuscript_file.Rmd
@@
 link-citations: yes
+csl: nature-geoscience.csl
 ---
@@
 localPath <- Sys.getenv("USERPROFILE")
+knitr::opts_knit$set(root.dir = "../") # set wd to root of RStudio project
 
@@
 
-# Working directory------
-# wd is set to root of RStudio project via options under Knitr drop-down
-# menu above.
+```
 
+```{r source_scripts, echo=FALSE, warning=FALSE, message=FALSE}
 # source scripts------
@@
 
+## N~2~O undersaturation extent
+
 In this study we present the first national-scale survey of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S.
 (CONUS).
-A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1).
-Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?).
-`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N~2~O sinks (Figure 1).
-We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population.
+A generalized random-tessellation survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?]** sites from CONUS waterbodies that were least 1 hectare in area, at least 1 meter deep, had at least 0.1 hectare of open water, and had a minimum residence time of one week (Figure 1)[@USEPA_NLA17_2022].
+Sample sites were distributed across all nine major CONUS ecoregions, which vary widely in water chemistry, morphometry, watershed land-use, and climate [@Omernik_1987; @Omernik_Griffith_2014].
+`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks (Figure 1).
+We modeled the data using a Bayesian hierarchical approach and predicted the dissolved and equilibrium N~2~O concentrations for all 224,916 CONUS waterbodies in the target population.
 At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`).
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`), the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), and >65% of the waterbodies in every ecoregion were undersaturated (Table 1).
 
@@
 
-Here is a citation example: Multilevel models are useful for survey inference [@Kennedy_Gelman_2021]. This paper by @Kennedy_Gelman_2021 provides a nice worked example.
+The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
+This finding  runs counter to the long standing paradigm that nitrogen rich surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
+@Webb_etal_2019 found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and @Soued_etal_2016 report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks.
+While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
+This pattern holds across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use across the continental U.S..
+
+```{r table1, echo = FALSE, cache=TRUE}
+rbind(national_stats %>% 
+        select(WSA9, estimate, LCL, UCL, post_med_sat, LCI_sat, UCI_sat, post_med_en2o, LCL_en2o, UCL_en2o),
+      ecoreg_stats %>% 
+        select(WSA9, estimate, LCL, UCL, post_med_sat, LCI_sat, UCI_sat, post_med_en2o, LCL_en2o, UCL_en2o)) %>%
+  mutate('N2O undersaturation extent' = paste0(estimate*100, " (", LCL*100, "-", UCL*100, ")"),
+         'N2O saturation ratio' = paste0(post_med_sat, " (", LCI_sat, "-", UCI_sat, ")"),
+         'N2O emission rate' = paste0(post_med_en2o, " (", LCL_en2o, "-", UCL_en2o, ")")) %>% 
+  select(WSA9, 'N2O undersaturation extent', 'N2O saturation ratio', 'N2O emission rate') %>%
+  rename('spatial domain' = WSA9) %>%
+knitr::kable(caption = "Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, and emission rate at the national and ecoregional scales")
+
+```
+
+Here is a citation example: Multilevel models are useful for survey inference [@Gelman_Little_1997; @Park_etal_2004; @Kennedy_Gelman_2021; @ Zachmann_etal_2022].
+
+### N~2~O saturation ratio
+
+#### Distribution
+
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean N~2~O saturation ratio (ratio of measured to equilibrium concentration) was equal to or greater than 1 in seven of the nine ecoregions (Table 1, Figure 2), suggesting that, on average, waterbodies in these ecoregions were supersatured with N~2~O and therefore functioning as a source of N~2~O.
+The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations that exceed equilibrium in most ecoregions can be attributed to the distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies (Figure 2).
+Across the CONUS, and within each of the nine major ecoregions, the distribution of N~2~O saturation ratios had median values that were less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that were sufficiently high to drive the population mean to values close to, or even greater than 1.
+Thus most waterbodies in the CONUS are functioning as N~2~O sinks, but a few function as strong N~2~O sources.
+
+![Distribution of N~2~O saturation values for the population of waterbodies in the nine major ecoregions.
+Vertical black dashed and solid lines indicated median and mean, respectively.
+Vertical red line indicates a saturation ratio of 1.](manuscript_figures/figure2.tiff)
+
+#### Controls on N~2~O saturation
+
+N~2~O saturation was related to waterbody size in several ecoregions, but the direction of the relationship varied.
+For example, N~2~O saturation increased with waterbody size in the Northern Plains, Upper Midwest, Northern Appalachians, and Western Mountain ecoregions, but decreased with increasing size in the Coastal and Southern Plains (Figure 3).
+A pattern common to many of the ecoregions, regardless of the direction of the relationship between N~2~O saturation and waterbody size, is that N~2~O saturation deviates from equilibrium most strongly in the smallest waterbody size class.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as delta N~2~O, decreases linearly with increasing waterbody size class.
+
+NEED TO ADD THE POOLED FIGURE
 
 # References
+
+<div id="refs"></div>
+
+# Methods
+
+## Survey design
+
+The population of interest (POI) for the 2017 NLA survey was CONUS waterbodies of at least 1 hectare in area, at least 1 meter deep, with at least 0.1 hectare of open water, and a minimum residence time of one week (Figure 1)[@USEPA_NLA17_2022].
+The POI was sampled according to a generalized random-tessellation design with stratification [@Stevens_Olsen_2004; USEPA_NLA17_2020].
+The design prescribed a spatially balanced, probabilistic sample with stratification along three factors: surface area (hectares; 5 ordered levels), ecoregion [@Omernik_1987; @Herlihy_etal_2008; @Omernik_Griffith_2014] (9 levels), and US state (excluding AK and HI; 48 levels).
+The size categories were sampled with unequal probability, such that larger lakes were over-sampled relative to smaller lakes.
+
+## Data collection
+
+Samples were collected at each of 1091 index waterbodies across the US during the summer index period.
+In addition, 95 of those waterbodies were re-visited for a second sample.
+This subset of revisit sites was used for assessing model fit and out-of-sample model performance.
+
+### Gas samples
+
+Duplicate dissolved N~2~O samples were collected from each waterbody at a depth of ~0.1m. Gas samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L.
+
+## Statistical inference
+
+The survey samples were a randomly selected subset of the complete POI, but with a pattern of "missingess" prescribed by the survey selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
+Therefore, to make inferences from the sample to the POI, the sample was statistically adjusted in order to account for the selection mechanism.
+We used a model-based approach, described as multilevel regression with poststratification (MRP)[@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020], to make adjustments.
+This procedure prescribes conditioning the response variables (_i.e._, N~2~O) in the sample data on the survey design variables (_i.e._, ecoregion, state, size) in a multilevel regression model; and then "poststratifying" or re-weighting estimates based the known or assumed distribution of waterbodies within the survey design groupings in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+Multilevel models are often recommended because they can provide regularized estimates along the design groupings, which can improve out-of-sample inferences [e.g., @Gelman_Little_1997; @Kennedy_Gelman_2021].
+Estimates for group levels that may be missing from the sample, but are part of the population of interest, are also straightforward using this approach [@Gelman_etal_2020 Ch.
+17; @McElreath_2020].
+The poststratification step typically entails using the fitted model to predict to a corresponding summary of the design groupings in the POI and then simply normalizing the estimates by the proportional distribution of units in the POI [@Kennedy_Gelman_2021].
+However, because N~2~O flux was to be calculated in this study as a function of waterbody-level properties (_i.e._, surface area), instead of predicting to a poststratification table of summarized population groupings, predictions were instead made to each of 465,897 individual waterbodies in the POI.
+For inferences, the waterbody-level predictions across the POI were readily summarized along the relevant groupings (_e.g._, ecoregion).
+The multilevel regression models were implemented using the Bayesian mode of inference, which made propagation of uncertainty from the waterbody-level predictions to group level summaries straightforward.
+Predictions were assumed relevant to the POI for average conditions during the biological index period of 2017.
+
+The final Bayesian model resulted in predictions for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio; and (3) the proportion of under-saturated water bodies (_i.e._, saturation ratio < 1).
+Saturation ratio is calculated as the ratio of dissolved to equilibrium N~2~O.
+Because dissolved and equilibrium N~2~O were observed on the same sample units (lake sites), their joint distribution was modeled to account for any potential statistical dependencies between the N~2~O responses due to, for example, common dependencies on geography.
+The model response variable was, therefore, multivariate.
+Although predictions of the mean marginal probabilities from separate models may have provided comparable estimates, a joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
+Therefore, the saturation ratio was was not predicted directly, but as a _post hoc_ or "derived" parameter [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
+Similarly, the proportion of under-saturated waterbodies was derived as the proportion of the _post-hoc_ PPD for the ratio that was larger than 1.
+All of the models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models via Hamiltonian Monte Carlo.
+
+Models excluding the NO3 covariate consistently resulted in poorer fits to to the observed dissolved N2O data and potentially strongly biased estimates of the saturation ratio.
+Including surface temperature and elevation in the equilibrium N2O part of the model also resulted in substantially improved replication of key aspects of the observed data.
+Likewise, added flexibility in the distributional terms for both dissolved and equilibrium N2O led to improvements.
+
+To make inferences to the population of interest from a model including these covariates, however, the covariates needed to be (1) fully observed across that population or (2) their missingness needed to be modeled.
+For the lake area and elevation covariates, data was available for all lakes from previously compiled geospatial databases.
+However, neither surface temperature or NO3 were observed for lakes outside of the sample.
+That is, they were partially observed with respect to the target population.
+Therefore, a more complex model was constructed below that included surface temperature and NO3 as additional responses conditioned on the survey design variables and fully observed covariates.
+This approach to inference for N2O was similar to Bayesian structural equation modeling approaches [@Merkle_etal_2021; @Merkle_Rosseel_2018].
+The logical dependence structure could be characterized as:
+
+$$\begin{aligned} 
+{\boldsymbol{N_2O_{diss}}} &\sim Survey + Area + {\boldsymbol{NO_3}} + {\boldsymbol{Temp}} \\ 
+{\boldsymbol{N_2O_{equil}}} &\sim Survey + Elev  + {\boldsymbol{Temp}}\\
+{\boldsymbol{NO_3}} &\sim Survey + Area + {\boldsymbol{Temp}} \\ 
+{\boldsymbol{Temp}} &\sim Survey + Lat + Elev + Day
+\end{aligned}$$
+
+Variables in bold text above were treated as partially observed with respect to the population of interest (i.e., observed only in the sample), whereas variables not in bold were considered fully observed.
+The partially observed variables, being dissolved and equilibrium N2O, NO3, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.
+This piece-wise approach required a more complex set of post-processing steps compared to a typical MRP analysis.
+In order to propagate estimates and uncertainty through the dependency structure and make inferences, the fitted model was used to first predict surface temperature in the target population, since it depended only on the fully observed covariates.
+That predictive distribution was then used alongside the relevant fully observed covariates to predict NO3 in the target population.
+Finally, the predictive distributions for temperature and NO3 were used to predict the N2O responses.
+These steps were carried out in the "Predict to population" section to follow.
+
+In the final model below, the sub-model for surface temperature assumed a Gamma distributed error distribution and the linear predictor included the survey design variables, latitude, elevation, and Julian date.
+The shape parameter was also modeled as a function of latitude to address increasing response variance along the latitudinal gradient.
+The NO3 sub-model was a cumulative logit model and the linear predictor included all of the survey factors as well as surface temperature and lake area.
+
+The dissolved and equilibrium N2O responses were each modeled with Gamma distributed errors, but with the same covariate structure as in model 5.
+The same structure was also employed for the shape terms in these responses, corresponding to the $\sigma$ terms in the previous models.
+Though not shown in this document, the Gamma error structure appeared to result in slightly better performance in the predictive checks compared to the Gaussian errors in previous models.
+This was primarily apparent in the saturation ratio checks, which may have been more sensitive to model performance in the tails of the N2O responses.
+Others have also indicated that the Gamma error distribution can work well for dissolved N2O data [@Webb_etal_2019].
+
+Note that there was no observation-level residual correlation term for this model, since the residuals are undefined for the Gamma and cumulative logit models.
+Dropping the observation-level residual correlation term was deemed a reasonable compromise that enabled the inclusion of NO3 as a covariate on dissolved N2O.
+The random intercepts, however, still allowed for potential correlations between the four responses at the group levels.

### 2023-04-13 manuscript/manuscript_file.Rmd

@@
 
+## NOTES
+* Table and figure captions should specify if displaying sample vs population estimates.
+
 # Introduction
@@
 
-![Location of sampling sites.
+![Figure 1.
+Location of sampling sites.
 Color of the points indicates whether the waterbody was functioning as a source or sink of nitrous oxide (N~2~O).
@@
 This finding  runs counter to the long standing paradigm that nitrogen rich surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
-@Webb_etal_2019 found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and @Soued_etal_2016 report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks.
+Webb et al.
+[@Webb_etal_2019] found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Soued et al [@Soued_etal_2016] report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020)  but kortelainen and Soued are mostly pristine lakes]**.
 While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
@@
   rename('spatial domain' = WSA9) %>%
-knitr::kable(caption = "Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, and emission rate at the national and ecoregional scales")
+knitr::kable(caption = "Table 1. Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, and emission rate at the national and ecoregional scales")
 
@@
 
-![Distribution of N~2~O saturation values for the population of waterbodies in the nine major ecoregions.
+![Figure 2.
+Distribution of N~2~O saturation values for the population of waterbodies in the nine major ecoregions.
+Each blue line (n=500) represents the distribution of values from a single realization of the posterior prediction.
 Vertical black dashed and solid lines indicated median and mean, respectively.
@@
 #### Controls on N~2~O saturation
+N~2~O saturation was related to waterbody size in several ecoregions, but the direction of the relationship varied.
+For example, N~2~O saturation increased with waterbody size in the Northern Plains, Upper Midwest, Northern Appalachians, and Western Mountain ecoregions, but decreased with increasing size in the Coastal and Southern Plains (Figure 3A).
+The inconsistent relationship between N~2~O saturation and waterbody size may be attributable differences in nitrate concentrations across ecoregions.
+The model identified an interaction between waterbody surface area and nitrate concentration (expressed categorically) where N~2~O concentration decreases with increasing lake size at high NO3 concentrations, but increases, or remains constant at low nitrate concentrations.
+THIS IS BECAUSE THERE CAN'T BE MUCH N2O BIOGEOCHEMISTRY WITHOUT INORGANIC N LAYING AROUND.
 
-N~2~O saturation was related to waterbody size in several ecoregions, but the direction of the relationship varied.
-For example, N~2~O saturation increased with waterbody size in the Northern Plains, Upper Midwest, Northern Appalachians, and Western Mountain ecoregions, but decreased with increasing size in the Coastal and Southern Plains (Figure 3).
-A pattern common to many of the ecoregions, regardless of the direction of the relationship between N~2~O saturation and waterbody size, is that N~2~O saturation deviates from equilibrium most strongly in the smallest waterbody size class.
-When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as delta N~2~O, decreases linearly with increasing waterbody size class.
+![Figure 3.
+Relationship between delta N~2~O and lake size category across A) the population of CONUS waterbodies and B) within each of the nine major CONUS ecoregions.
+Red arrows in panel A illustrate the magnitude of N~2~O disequilibrium in the smallest lake size class.
+Data are inferences made to the population of CONUS waterbodies.](manuscript_figures/figure3.tiff)
 
-NEED TO ADD THE POOLED FIGURE
+![Figure 4.
+Partial dependence plot illustrating the relationship between N~2~O concentration and waterbody surface area as a function of nitrate concentration.](manuscript_figures/N2O vs surface area x NO3 partial dependence plot.png)
+
+Regardless of the direction of the relationship between N~2~O saturation and waterbody size, N~2~O saturation deviates from equilibrium most strongly in the smallest waterbody size class.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases linearly with increasing waterbody size class (Figure 3B).
+Waterbody size can affect dissolved gas concentration through several mechanisms.
+In lakes with low volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column.
+Since water volume to benthic surface area ratios tend to scale positively with lake surface area, benthic N~2~O biogeochemistry will have a greater affect on dissolved N~2~O concentration in small lakes, regardless of whether the sediments function as a net sink or source of N~2~O **[see Wetzel for references]**.
+Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes **[need a citation for N2O ~ temperature relationship]**.
+Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (Beaulieu et al N2O Harsha Lake and USACE N2O papers).
+
+Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequlibrim in smaller waterbodies.
+Air-water gas exchange will vent excess N~2~O from surface waters or replenish depleted N~2~O in undersaturated waterbodies by facilitating the invasion of atmospheric N~2~O.
+Regardless of whether a waterbody is superaturated or undersatured in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
+One of the primary drivers of air-water gas exchange in lentic waters is wind speed (Vachon and Prairie reference).
+Wind speeds to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
+Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the affect of biogeochemistry on dissolved N~2~O is more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
+
+**Not sure where I am going with this paragraph.  Maybe scratch it.  Mostly included as a placeholder for these references**
+While previous surveys have documented relationships between lake size and dissolved methane concentration (Pighini et al.
+2018, Juutinen et al 2009, Borges et al.
+2011), another greenhouse gas, to the best of our knowledge, this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
+Perhaps because previous surveys have included little variation in waterbody size (Webb et al.
+2019, Kortelainen et al) or water quality (Soued et al.
+2015).
 

### 2023-04-13 manuscript/manuscript_support.R
minor changes to manuscript_support.R

@@ -284,30 +284,11 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
 # if the image is already on computer, then nothing, else create image
 if(!("figure3.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   
-# PLOT N2O* BY SIZE
-# point and linerange
-p1 <- all_predictions %>%
-  mutate(n2ostar = abs(n2o - n2oeq)) %>%
-  group_by(size_cat, .draw) %>% # group by iteration
-  #group_by(.draw) %>%
-  summarise(mean_star = mean(n2ostar)) %>% # 500 means for each WSA9
-  # now summarize to 1 statistic per WSA9
-  summarise( estimate = round(median(mean_star), 3), 
-             LCL = round(quantile(mean_star, probs = 0.025), 3),
-             UCL = round(quantile(mean_star, probs = 0.975), 3)) %>% 
-  ggplot( aes( x = size_cat, y = estimate ) ) +
-  geom_point() +
-  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
-  ylab(expression("["*Delta~N[2]*O*"]")) +
-  xlab("size category (ha)") +
-  coord_flip() + 
-  theme_bw() 
-#ggsave("output/figures/n2oStarBySize.tiff")
 
 
 # PLOT N2O SAT RATIO BY WSA9 AND SIZE
 # point and linerange
-p2.data <- all_predictions %>%
+p1.data <- all_predictions %>%
   group_by(WSA9, size_cat, .draw) %>% # group by iteration
   summarise(mean_sat = mean(n2osat)) %>% # 500 means for each WSA9
   # now summarize to 1 statistic per WSA9
@@ -318,25 +299,45 @@ p2.data <- all_predictions %>%
   mutate(ecoregion = fct_reorder(ecoregion, estimate)) 
 
 # data fpr arrow segment.  Only smallest size category
-p2.data.arrow <- p2.data %>%
+p1.data.arrow <- p1.data %>%
   filter(size_cat == "min_4")
 
-p2 <- p2.data %>%
+p1 <- p1.data %>%
   ggplot(aes(x=estimate, y=size_cat)) +
   geom_point() +
   geom_linerange( aes( xmin = LCL, xmax = UCL)) +
   geom_vline(xintercept = 1, color='blue') +
-  geom_segment(data = p2.data.arrow, 
[…]

### 2023-04-14 manuscript/manuscript_file.Rmd
added discussion of NO3 and temperature partial dependence plots

--- 2023-04-12 manuscript/manuscript_file.Rmd
@@
 `r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks (Figure 1).
-We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population.
+We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897]**.
 At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
@@
 
-#### Controls on N~2~O saturation
-N~2~O saturation was related to waterbody size in several ecoregions, but the direction of the relationship varied.
-For example, N~2~O saturation increased with waterbody size in the Northern Plains, Upper Midwest, Northern Appalachians, and Western Mountain ecoregions, but decreased with increasing size in the Coastal and Southern Plains (Figure 3A).
-The inconsistent relationship between N~2~O saturation and waterbody size may be attributable differences in nitrate concentrations across ecoregions.
-The model identified an interaction between waterbody surface area and nitrate concentration (expressed categorically) where N~2~O concentration decreases with increasing lake size at high NO3 concentrations, but increases, or remains constant at low nitrate concentrations.
-THIS IS BECAUSE THERE CAN'T BE MUCH N2O BIOGEOCHEMISTRY WITHOUT INORGANIC N LAYING AROUND.
+#### Controls on N~2~O
+Dissolved N~2~O concentration increased with NO~3~ concentration **[need to pick a figure]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the  sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ (Knowles 1982; Richardson et al.
+2009).
+The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
+Incomplete denitrification often occurs when NO~3~ is relatively abundant, likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification (Koike and Hattori, 1975;Richardson et al., 2009).
+When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~ thereby functioning as an N~2~O sink.
+
+It is also possible that the correlation between N2O and NO3 reflects the simultaneous production of these solutes via nitrification.
+Nitrification, an aerobic process which can produce N2O as a byproduct, is the microbial oxidation of ammonium (NH~4~^+^) to NO~3~ **[citation]**.
+Nitrification might be expected to be more prevalent in the predominantly well oxygenated surface waters where the N~2~O samples were collected than denitrification, which is an anaeorbic process.
+However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column **[citation]** and can be transported from aoxic sediments to the water surface via advection and diffusion (Delsontro no longer a paradox).
+Thus the N~2~O - NO3 relationship may be reflective of some combination of N2O production/consumption via denitrification and N2O production via nitrification.
+
+The sensitivity of N2O to NO3 increased with water temperature **[need to pick a figure]**.
+This is consistent with N2O production and consumption occurring through a temperature sensitive biological process.
+This relationship is apparent at high NO3 concentrations, where nitrogen biogeochemistry is a net source of N2O, and at low NO3 concentrations where nitrogen biogeochemistry is a net sink for N2O.
+Thus NO3 concentration dictates whether a water body is a source or sink of N2O and elevated water temperature magnifies the effect.
+
+**[N2O vs NO3 with different lines representing quantiles of the temperature distribution]**
+
+#### Controls on N~2~O saturation **[this heading is to organize the .Rmd, not intended to be retained in final MS]**
+
+While NO3 availability determined whether N biogeochemistry was a source or sink of N2O and water temperature affected the rate of N2O production or consumption, waterbody morphology determined the extent to which N2O production or consumption affected dissolved N~2~O concentration at the air-water interface.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases linearly with increasing waterbody size class (Figure 3A).
+Waterbody size can affect dissolved gas concentration through several mechanisms.
+In lakes with low volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column.
+Since water volume to benthic surface area ratios tend to scale positively with lake surface area, benthic N~2~O biogeochemistry will have a greater affect on dissolved N~2~O concentration in small lakes, regardless of whether the sediments function as a net sink or source of N~2~O **[see Wetzel for references]**.
+Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes **[need a citation for N2O ~ temperature relationship]**.
+Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (Beaulieu et al N2O Harsha Lake and USACE N2O papers).
+
+Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequlibrim in smaller waterbodies.
+Air-water gas exchange will vent excess N~2~O from surface waters or facilitate the invasion of air that can replenish depleted N~2~O in undersaturated waterbodies.
+Regardless of whether a waterbody is superaturated or undersatured in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
+One of the primary drivers of air-water gas exchange in lentic waters is wind speed (Vachon and Prairie reference).
+Wind speeds tend to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
+Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the affect of biogeochemistry on dissolved N~2~O is more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
+
+The relationship between N~2~O and waterbody size is also evident in a few ecoregions (Figure 3B).
+For example, mean N2O saturation ratio is greater than 1 for all size classed in the Coastal Plains, consistent with the high NO3 concentrations in this ecoregion **[is this true?]**, but decreases with increasing lake size, consistent the effect of waterbody morphology on N2O production and gas exchange.
+By contrast, mean N2O saturation is less than one across all size classes in the low NO3 northern plains, but again, N2O saturation moves toward equilibrium with increasing lake size.
+These relationships are evident to lesser degrees in other ecoregions, likely dependent on the distribution on NO3 and water temperature across lake classes.
+**[I'm beginning to think that Fig.
+3B unnecessarily complicates an otherwise pretty clean story.
+Maybe we just show Fig.
+3B?]**
 
@@
 
-Regardless of the direction of the relationship between N~2~O saturation and waterbody size, N~2~O saturation deviates from equilibrium most strongly in the smallest waterbody size class.
-When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases linearly with increasing waterbody size class (Figure 3B).
-Waterbody size can affect dissolved gas concentration through several mechanisms.
-In lakes with low volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column.
-Since water volume to benthic surface area ratios tend to scale positively with lake surface area, benthic N~2~O biogeochemistry will have a greater affect on dissolved N~2~O concentration in small lakes, regardless of whether the sediments function as a net sink or source of N~2~O **[see Wetzel for references]**.
-Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes **[need a citation for N2O ~ temperature relationship]**.
-Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (Beaulieu et al N2O Harsha Lake and USACE N2O papers).
-
-Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequlibrim in smaller waterbodies.
-Air-water gas exchange will vent excess N~2~O from surface waters or replenish depleted N~2~O in undersaturated waterbodies by facilitating the invasion of atmospheric N~2~O.
-Regardless of whether a waterbody is superaturated or undersatured in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
-One of the primary drivers of air-water gas exchange in lentic waters is wind speed (Vachon and Prairie reference).
-Wind speeds to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
-Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the affect of biogeochemistry on dissolved N~2~O is more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
-
 **Not sure where I am going with this paragraph.  Maybe scratch it.  Mostly included as a placeholder for these references**
@@
 
+# EMISSION RATES
+
 # References

### 2023-04-14 manuscript/manuscript_support.R

@@ -298,9 +298,10 @@ p1.data <- all_predictions %>%
   mutate(ecoregion = factor(WSA9)) %>%
   mutate(ecoregion = fct_reorder(ecoregion, estimate)) 
 
-# data fpr arrow segment.  Only smallest size category
+# data fpr arrow segment.  Only smallest size category.  Only CPL and NPL.
 p1.data.arrow <- p1.data %>%
-  filter(size_cat == "min_4")
+  filter(size_cat == "min_4",
+         WSA9 %in% c("CPL", "NPL"))
 
 p1 <- p1.data %>%
   ggplot(aes(x=estimate, y=size_cat)) +
@@ -314,6 +315,7 @@ p1 <- p1.data %>%
   xlab(expression(mean~N[2]*O~saturation~ratio)) +
   ylab("waterbody size category (ha)") +
   theme_bw() +
+  theme(axis.title.y = element_blank()) +
   facet_wrap(~ecoregion)
 
 
@@ -332,12 +334,13 @@ p2 <- all_predictions %>%
   geom_point() +
   geom_linerange( aes( xmin = LCL, xmax = UCL)) +
   xlab(expression("["*Delta~N[2]*O*"]"~"("*nmol~L^{-1}*")")) +
+  ylab("waterbody size category (ha)") +
   #coord_flip() + 
-  theme_bw() +
-  theme(axis.title.y = element_blank())
+  theme_bw() 
+
 #ggsave("output/figures/n2oStarBySize.tiff")
 
-ggpubr::ggarrange(p1, p2, ncol=2, nrow = 1, widths = c(0.7, 0.3), labels = c("A", "B"))
+ggpubr::ggarrange(p2, p1, ncol=2, nrow = 1, widths = c(0.3, 0.7), labels = c("A", "B"))
 
 ggsave("manuscript/manuscript_figures/figure3.tiff", width = 8.5, height = 5)
 }

### 2023-04-21 manuscript/manuscript_file.Rmd
not sure what changes are in this commitgit add -u

@@
          'N2O saturation ratio' = paste0(post_med_sat, " (", LCI_sat, "-", UCI_sat, ")"),
-         'N2O emission rate' = paste0(post_med_en2o, " (", LCL_en2o, "-", UCL_en2o, ")")) %>% 
+         'N2O emission rate' = paste0(round(post_med_en2o, digits = 1), " (", round(LCL_en2o, digits = 1), "-", round(UCL_en2o, digits = 1), ")")) %>% 
   select(WSA9, 'N2O undersaturation extent', 'N2O saturation ratio', 'N2O emission rate') %>%
@@
 #### Controls on N~2~O
+##### NO3
 Dissolved N~2~O concentration increased with NO~3~ concentration **[need to pick a figure]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the  sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ (Knowles 1982; Richardson et al.
@@
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
-Incomplete denitrification often occurs when NO~3~ is relatively abundant, likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification (Koike and Hattori, 1975;Richardson et al., 2009).
-When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~ thereby functioning as an N~2~O sink.
+Incomplete denitrification often occurs when NO~3~ is relatively abundant (Beaulieu et al Harsha Lake, Beauliue et al USACE), likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification (Koike and Hattori, 1975;Richardson et al., 2009).
+When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~, thereby functioning as an N~2~O sink.
 
@@
 
-The sensitivity of N2O to NO3 increased with water temperature **[need to pick a figure]**.
-This is consistent with N2O production and consumption occurring through a temperature sensitive biological process.
-This relationship is apparent at high NO3 concentrations, where nitrogen biogeochemistry is a net source of N2O, and at low NO3 concentrations where nitrogen biogeochemistry is a net sink for N2O.
-Thus NO3 concentration dictates whether a water body is a source or sink of N2O and elevated water temperature magnifies the effect.
+##### Temperature
+The sensitivity of N2O to NO3 increased with water temperature **[need to pick a figure]**, which is consistent with N2O production and consumption occurring through temperature sensitive biological processes.
+At low NO3 concentrations, greater water temperature exacerbated N2O undersaturation, presumably by stimulating N2O consumption rates.
+At high NO3 concentration, greater water temperature increased N2O supersaturation, likely by stimulating N2O production rates.
+The relative effect of NO3 was greater than that of temperature, thus elevated water temperature serves to magnify the rates of N2O production or consumption dictated by NO3 availability.
 
-**[N2O vs NO3 with different lines representing quantiles of the temperature distribution]**
+##### Waterbody size
 
-#### Controls on N~2~O saturation **[this heading is to organize the .Rmd, not intended to be retained in final MS]**
-
-While NO3 availability determined whether N biogeochemistry was a source or sink of N2O and water temperature affected the rate of N2O production or consumption, waterbody morphology determined the extent to which N2O production or consumption affected dissolved N~2~O concentration at the air-water interface.
-When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases linearly with increasing waterbody size class (Figure 3A).
+While NO3 and water temperature were the primary drivers of N2O production or consumption rates, waterbody morphology determined the extent to which N2O processing affected dissolved N~2~O in surface waters.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases with increasing waterbody size class (Figure 3A).
 Waterbody size can affect dissolved gas concentration through several mechanisms.
@@
 
-The relationship between N~2~O and waterbody size is also evident in a few ecoregions (Figure 3B).
-For example, mean N2O saturation ratio is greater than 1 for all size classed in the Coastal Plains, consistent with the high NO3 concentrations in this ecoregion **[is this true?]**, but decreases with increasing lake size, consistent the effect of waterbody morphology on N2O production and gas exchange.
-By contrast, mean N2O saturation is less than one across all size classes in the low NO3 northern plains, but again, N2O saturation moves toward equilibrium with increasing lake size.
-These relationships are evident to lesser degrees in other ecoregions, likely dependent on the distribution on NO3 and water temperature across lake classes.
-**[I'm beginning to think that Fig.
-3B unnecessarily complicates an otherwise pretty clean story.
-Maybe we just show Fig.
-3B?]**
+The influence of waterbody size is also evident in the statistical interaction between NO3 and surface area (Fig XX).
+At high NO3 concentrations, N2O is highly supersatured in small waterbodies, but moves toward equilibrium in large waterbodies, as seen in the Coastal Plains ecoregion**[is CPL high NO3?]**.
+At low NO3 concentrations, N2O is undersatured in small waterbodies, but increases toward equilibrium in larger waterbodies, as seen in the Northern Plains ecoregion **[is NPL low nitrate?]**(Fig).
+
+The model indicates that NO3, water temperature, and surface area interact to determine the N2O saturation ratio.
+These interactions yield a broad range of N2O saturation values across the population of CONUS waterbodies, but also enable the identifcation of broad scale patterns.
+First, small waterbodies with low NO3 concentration tend to be undersaturated, whereas small waterbodies with high NO3 tend to be highly supersaturated.
+Large waterbodies, regardless of NO3 status, tend to have N2O saturation ratios close to equilibrium.
+Therefore intra and inter regional patterns in N2O saturation are determined by the joint distributions of waterbody size, NO3 availability, and water temperature within an ecoregion.
 
@@
 # EMISSION RATES
+Despite widespread N2O undersaturation in CONUS waterbodies, the national average N2O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N2O m^-2^ d^-1^ and only the Northern Appalachians and Northern Plains ecoregions had mean emission rates less than 0 (Table 1).
+This is accordance with the distribution on N2O saturation values where relatively rare, but highly supersatured, waterbodies pull the mean to a value that far exceeds the median.
+ecoregional means range from `r ecoreg_stats %>% slice_min(n = 1, post_med_en2o) %>% select(post_med_en2o) %>% pull()` to `r ecoreg_stats %>% slice_max(n = 1, post_med_en2o) %>% select(post_med_en2o) %>% pull()` mg N2O m^-2^ d^-1^ (Table 1).
 

### 2023-04-21 manuscript/manuscript_support.R

@@ -140,11 +140,11 @@ pred.k600 <- function(ws, area) {
 ### calculate emission and flux----
 all_predictions <- all_predictions %>%
   mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
-         # 1000 L to m3. 44ng to 1nmol. ng->ug. ug->mg
-         e.n2o.mg.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000 * 1000)) * 
+         # 1000 L to m3. 44ng to 1nmol. ng->ug.
+         e.n2o.ug.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000)) * 
            (k600.cm.h * (24/100)), #h->day. cm->m
-         f.n2o.Mg.y = (e.n2o.mg.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
-           (1000*1000*1000)) #mg-g, g-kg, kg-Mg
+         f.n2o.Mg.y = (e.n2o.ug.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
+           (1000*1000*1000*1000)) #ug-mg, mg-g, g-kg, kg-Mg
 
 
 ## ecoregion and national means----
@@ -154,7 +154,7 @@ ecoreg_stats <- all_predictions %>%
              mean_n2oeq = mean(n2oeq),
              mean_sat = mean(n2osat),
              prop_sat = sum(n2osat < 1) / length(unique(.row)),
-             mean_en2o = mean(e.n2o.mg.m2.d),
+             mean_en2o = mean(e.n2o.ug.m2.d),
              fn2o_Mg = sum(f.n2o.Mg.y)) %>%
   summarise( post_med_n2o = round(mean(mean_n2o), 1),
              LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
@@ -174,7 +174,8 @@ ecoreg_stats <- all_predictions %>%
              post_med_fn2o = round(mean(fn2o_Mg), 3), 
              LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
              UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
-             n = dplyr::n())
+             n = dplyr::n()) %>%
+  ungroup()
 
 national_stats <- all_predictions %>%
   group_by(.draw) %>%
@@ -182,7 +183,7 @@ national_stats <- all_predictions %>%
              mean_n2oeq = mean(n2oeq),
[…]

### 2023-06-21 manuscript/manuscript_file.Rmd
minor updates to emission rate section of manuscript
reconciling merge conflict
manuscript progress

--- 2023-04-14 manuscript/manuscript_file.Rmd
@@
 library(tidyverse) # dplyr, ggplot
+library(flextable)
 
@@
 ## NOTES
-* Table and figure captions should specify if displaying sample vs population estimates.
+
+-   Table and figure captions should specify if displaying sample vs population estimates. "posterior predictive distribution"
+-   Note that right tail is truncated in figure 2.
 
@@
 ### N~2~O undersaturation extent
+
 In this study we present the first national-scale survey of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S.
 (CONUS).
-A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1).
+A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
+compare my sample IDs to NLA IDs]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1).
 Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?).
 `r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks (Figure 1).
-We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897]**.
+We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897.
+See technical support document]**.
 At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`), the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), and >65% of the waterbodies in every ecoregion were undersaturated (Table 1).
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`), the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), and \>65% of the waterbodies in every ecoregion were undersaturated (Table 1).
 
@@
 The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
-This finding  runs counter to the long standing paradigm that nitrogen rich surface waters are a source of N~2~O to the atmosphere (Mosier et al.
-1998) and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
-Webb et al.
-(2019) found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Soued et al (2015) report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020)  but kortelainen and Soued are mostly pristine lakes]**.
+This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere (@mosier_etal_1998) and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
+@Webb_etal_2019 found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and @Soued_etal_2016 report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020) but kortelainen and Soued are mostly pristine lakes]**.
 While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
@@
 ```{r table1, echo = FALSE}
-rbind(national_stats %>% 
-        select(WSA9, estimate, LCL, UCL, post_med_sat, LCI_sat, UCI_sat, post_med_en2o, LCL_en2o, UCL_en2o),
+# Set up data
+table1_dat <- rbind(national_stats %>% 
+        select(WSA9_NAME, 
+               total_sa,
+               estimate, LCL, UCL, 
+               post_med_sat, LCI_sat, UCI_sat, 
+               post_med_en2o, LCL_en2o, UCL_en2o, 
+               post_med_fn2o, LCL_fn2o, UCL_fn2o),
       ecoreg_stats %>% 
-        select(WSA9, estimate, LCL, UCL, post_med_sat, LCI_sat, UCI_sat, post_med_en2o, LCL_en2o, UCL_en2o)) %>%
-  mutate('N2O undersaturation extent' = paste0(estimate*100, " (", LCL*100, "-", UCL*100, ")"),
-         'N2O saturation ratio' = paste0(post_med_sat, " (", LCI_sat, "-", UCI_sat, ")"),
-         'N2O emission rate' = paste0(post_med_en2o, " (", LCL_en2o, "-", UCL_en2o, ")")) %>% 
-  select(WSA9, 'N2O undersaturation extent', 'N2O saturation ratio', 'N2O emission rate') %>%
-  rename('spatial domain' = WSA9) %>%
-knitr::kable(caption = "Table 1. Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, and emission rate at the national and ecoregional scales")
+        select(WSA9_NAME, 
+               total_sa,
+               estimate, LCL, UCL, 
+               post_med_sat, LCI_sat, UCI_sat, 
+               post_med_en2o, LCL_en2o, UCL_en2o,
+               post_med_fn2o, LCL_fn2o, UCL_fn2o)) %>%
+  mutate('N2O undersaturation extent' = paste0(estimate*100, " (", LCL*100, " - ", UCL*100, ")"),
+         'N2O saturation ratio' = paste0(post_med_sat, " (", LCI_sat, " - ", UCI_sat, ")"),
+         'N2O emission rate' = paste0(round(post_med_en2o, digits = 3), " (", round(LCL_en2o, digits = 3), " - ", round(UCL_en2o, digits = 3), ")"),
+         'N2O flux' = paste0(round(post_med_fn2o, digits = 0), " (", round(LCL_fn2o, digits = 0), " - ", round(UCL_fn2o, digits = 0), ")")) %>% 
+  select(WSA9_NAME, total_sa, 'N2O undersaturation extent', 'N2O saturation ratio', 'N2O emission rate', 'N2O flux') %>%
+  rename('spatial domain' = WSA9_NAME,
+         'waterbody surface area' = total_sa)
+
+# create flextable
+flextable(table1_dat) %>%
+  set_header_labels('waterbody surface area' = "waterbody surface area\n(hectares)",
+                    'N2O undersaturation extent' = "N2O undersaturation extent\n(% waterbodies)",
+                    'N2O emission rate' = "N2O emission rate\n(mg N2O m-2 d-1)",
+                    'N2O flux' = "N2O flux\n(Mg N2O y-1)") %>%
+  align(j=c(2:6), align = "center", part = "all") %>% # center align columns 2:4
+  valign(valign = "top", part = "header") %>% # top align column names
+  # caption doesn't show in .Rmd viewer, but renders correctly in .docx
+  set_caption(
+    caption = as_paragraph( # allows for formatting
+      as_chunk("Table 1. Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, emission rate, and flux at the national and ecoregional scales", 
+               props = fp_text_default(bold = TRUE) # here is where we specify formatting (font, size, bold, etc)
+      )
+    ),
+    align_with_table = FALSE) %>% # table is center aligned, ignore that and apply default left align to caption
+  autofit() %>% # convenient column spacing
+  fontsize(size = 8, part = "all") # doesn't apply to caption when as_paragraph is invoked.
 
@@
 
-Here is a citation example: Multilevel models are useful for survey inference [@Kennedy_Gelman_2021]. This paper by @Kennedy_Gelman_2021 provides a nice worked example.
+### N~2~O saturation ratio
 
-### N~2~O saturation ratio
 #### Distribution
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean N~2~O saturation ratio (ratio of measured to equilibrium concentration) was equal to or greater than 1 in seven of the nine ecoregions (Table 1, Figure 2), suggesting that, on average, waterbodies in these ecoregions were supersatured with N~2~O and therefore functioning as a source of N~2~O.
+
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean N~2~O saturation ratio (ratio of measured to equilibrium concentration) was equal to or greater than 1 in seven of the nine ecoregions **[(Table 1, Figure 2)]**, suggesting that, on average, waterbodies in these ecoregions were supersatured with N~2~O and were therefore functioning as a source of N~2~O.
 The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations that exceed equilibrium in most ecoregions can be attributed to the distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies (Figure 2).
 Across the CONUS, and within each of the nine major ecoregions, the distribution of N~2~O saturation ratios had median values that were less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
-Thus most waterbodies in the CONUS are functioning as N~2~O sinks, but a few function as strong N~2~O sources.
+Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources.
 
@@
 Vertical black dashed and solid lines indicated median and mean, respectively.
-Vertical red line indicates a saturation ratio of 1.](manuscript_figures/figure2.tiff)
+Vertical red line indicates a saturation ratio of 1.
+x-axis is truncated at 3 to better show the area where 97% of the observations occur.](manuscript_figures/figure2.tiff)
 
 #### Controls on N~2~O
-Dissolved N~2~O concentration increased with NO~3~ concentration **[need to pick a figure]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the  sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ (Knowles 1982; Richardson et al.
-2009).
+
+##### NO3
+
+Dissolved N~2~O concentration increased with NO~3~ concentration **[Fig.
+3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@knowles_1982, @richardson_etal_2009].
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
-Incomplete denitrification often occurs when NO~3~ is relatively abundant, likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification (Koike and Hattori, 1975;Richardson et al., 2009).
-When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~ thereby functioning as an N~2~O sink.
+Incomplete denitrification often occurs when NO~3~ is relatively abundant [@Beaulieu_etal_2015, @Beaulieu_etal_2014], likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification [@koike_etal_1975, @richardson_etal_2009].
+When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~, thereby functioning as an N~2~O sink.
+N~2~O consumption via denitrification likely explains the N~2~O undersaturation observed at low NO~3~ concentrations, whereas denitrification was likely a source of N~2~O at higher NO~3~ concentrations.
 
-It is also possible that the correlation between N2O and NO3 reflects the simultaneous production of these solutes via nitrification.
-Nitrification, an aerobic process which can produce N2O as a byproduct, is the microbial oxidation of ammonium (NH~4~^+^) to NO~3~ **[citation]**.
+It is also possible that the correlation between N~2~O and NO~3~ reflects the simultaneous production of these solutes via nitrification.
+Nitrification, an aerobic process which can produce N~2~O as a byproduct, is the microbial oxidation of ammonium (NH~4~^+^) to NO~3~ **[citation]**.
 Nitrification might be expected to be more prevalent in the predominantly well oxygenated surface waters where the N~2~O samples were collected than denitrification, which is an anaeorbic process.
-However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column **[citation]** and can be transported from aoxic sediments to the water surface via advection and diffusion (Delsontro no longer a paradox).
-Thus the N~2~O - NO3 relationship may be reflective of some combination of N2O production/consumption via denitrification and N2O production via nitrification.
+However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column **[citation]** and can be transported from aoxic sediments to the water surface via advection and diffusion [@delsontro_etal_2018].
+Thus the N~2~O - NO3 relationship may be reflective of some combination of N~2~O production/consumption via denitrification and N~2~O production via nitrification.
 
-The sensitivity of N2O to NO3 increased with water temperature **[need to pick a figure]**.
-This is consistent with N2O production and consumption occurring through a temperature sensitive biological process.
-This relationship is apparent at high NO3 concentrations, where nitrogen biogeochemistry is a net source of N2O, and at low NO3 concentrations where nitrogen biogeochemistry is a net sink for N2O.
-Thus NO3 concentration dictates whether a water body is a source or sink of N2O and elevated water temperature magnifies the effect.
+![Figure 3.
+A) Relationship between N~2~O saturation and nitrate (NO~3~) concentration, dependent upon surface water temperature.
+The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13.
+B) Relationship between N~2~O saturation and surface water temperature, dependent upon NO~3~ category.
+Horizontal dashed lines in both panels reflects an N~2~O saturation value of 1.](manuscript_figures/sat_cond_surftemp.tiff)
 
-**[N2O vs NO3 with different lines representing quantiles of the temperature distribution]**
+##### Temperature
 
-#### Controls on N~2~O saturation **[this heading is to organize the .Rmd, not intended to be retained in final MS]**
+N~2~O saturation also increased with water temperature **[Fig.
+3]**, which is consistent with N~2~O production and consumption occurring through temperature sensitive biological processes, but the relationship was dependent on NO~3~ availability.
+At low NO3 concentrations, temperature had a relatively minor affect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
+The effect of temperature became more pronounced when NO~3~ availability was greater, possibly reflecting the stimulatory affect of higher temperature on biological N~2~O production.
+The relative effect of NO~3~ was greater than that of temperature, thus elevated water temperature serves to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
 
-While NO3 availability determined whether N biogeochemistry was a source or sink of N2O and water temperature affected the rate of N2O production or consumption, waterbody morphology determined the extent to which N2O production or consumption affected dissolved N~2~O concentration at the air-water interface.
-When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases linearly with increasing waterbody size class (Figure 3A).
-Waterbody size can affect dissolved gas concentration through several mechanisms.
-In lakes with low volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column.
-Since water volume to benthic surface area ratios tend to scale positively with lake surface area, benthic N~2~O biogeochemistry will have a greater affect on dissolved N~2~O concentration in small lakes, regardless of whether the sediments function as a net sink or source of N~2~O **[see Wetzel for references]**.
-Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes **[need a citation for N2O ~ temperature relationship]**.
-Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (Beaulieu et al N2O Harsha Lake and USACE N2O papers).
+##### Waterbody size
+
+While NO~3~ and water temperature were the primary drivers of N~2~O production or consumption rates, waterbody morphology determined the extent to which N~2~O processing affected dissolved N~2~O in surface waters.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases with increasing waterbody size class **[Fig.
+4A]**.
+This pattern holds regardless of whether N~2~O is supersaturated or undersatured **[Fig.
+4B, C]**.
+Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
+In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column (@carpenter_1983) .
+Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes.
+Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (@Beaulieu_etal_2015, @Beaulieu_etal_2014).
+Unfortunately, the morphometric data required to test these hypothesis (e.g.
+mean depth, max depth) are not available for the samples waterbodies.
 
@@
 Regardless of whether a waterbody is superaturated or undersatured in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
-One of the primary drivers of air-water gas exchange in lentic waters is wind speed (Vachon and Prairie reference).
+One of the primary drivers of air-water gas exchange in lentic waters is wind speed (@Vachon_etal_2013).
 Wind speeds tend to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
@@
 
-The relationship between N~2~O and waterbody size is also evident in a few ecoregions (Figure 3B).
-For example, mean N2O saturation ratio is greater than 1 for all size classed in the Coastal Plains, consistent with the high NO3 concentrations in this ecoregion **[is this true?]**, but decreases with increasing lake size, consistent the effect of waterbody morphology on N2O production and gas exchange.
-By contrast, mean N2O saturation is less than one across all size classes in the low NO3 northern plains, but again, N2O saturation moves toward equilibrium with increasing lake size.
-These relationships are evident to lesser degrees in other ecoregions, likely dependent on the distribution on NO3 and water temperature across lake classes.
-**[I'm beginning to think that Fig.
-3B unnecessarily complicates an otherwise pretty clean story.
-Maybe we just show Fig.
-3B?]**
+![Figure 4.
+A) Relationship between $\Delta$ N~2~O (difference between observed and equilibrium N~2~O concentration) and waterbody surface areas.
+B,C) relationship between N~2~O saturation and waterbody size for the Coastal Plains and Northern Plains ecoregions.
+The vertical blue line references a N~2~O saturation value of 1.0.
+The horizontal red arrows highlight $\Delta$ N~2~O for the smallest lake size class.](manuscript_figures/deltaN2ObySize.tiff)
 
-![Figure 3.
-Relationship between delta N~2~O and lake size category across A) the population of CONUS waterbodies and B) within each of the nine major CONUS ecoregions.
-Red arrows in panel A illustrate the magnitude of N~2~O disequilibrium in the smallest lake size class.
-Data are inferences made to the population of CONUS waterbodies.](manuscript_figures/figure3.tiff)
+The influence of waterbody size is also evident in the statistical interaction between NO~3~ and surface area **[Fig.
+5]**.
+At high NO~3~ concentrations, N~2~O is highly supersatured in small waterbodies, but moves toward equilibrium in large waterbodies **[Fig.
+5A]**.
+At low NO3 concentrations, N~2~O is undersatured in small waterbodies, but increases toward equilibrium in larger waterbodies **[Fig.
+5A]**.
 
-![Figure 4.
-Partial dependence plot illustrating the relationship between N~2~O concentration and waterbody surface area as a function of nitrate concentration.](manuscript_figures/N2O vs surface area x NO3 partial dependence plot.png)
+The model indicates that NO~3~, water temperature, and surface area interact to determine the N~2~O saturation ratio.
+These interactions yield a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enable the identifcation of broad scale patterns.
+First, small waterbodies with low NO3 concentration tend to be undersaturated, whereas small waterbodies with high NO3 tend to be highly supersaturated.
+Large waterbodies, regardless of NO~3~ status, tend to have N~2~O saturation ratios closer to equilibrium.
+Therefore intra and inter regional patterns in N~2~O saturation are determined by the joint distributions of waterbody size, NO~3~ availability, and water temperature.
 
-**Not sure where I am going with this paragraph.  Maybe scratch it.  Mostly included as a placeholder for these references**
-While previous surveys have documented relationships between lake size and dissolved methane concentration (Pighini et al.
+![Figure 5.
+Partial dependence plots illustrating the relationship between N~2~O saturation and A) nitrate (NO~3~ availability dependent upon lake size, and B) lake size dependent upon NO~3~ availability.
+The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13.](manuscript_figures/sat_cond_area.tiff)
+
+**Not sure where I am going with this paragraph.
+Maybe scratch it.
+Mostly included as a placeholder for these references** While previous surveys have documented relationships between lake size and dissolved methane concentration (Pighini et al.
 2018, Juutinen et al 2009, Borges et al.
 2011), another greenhouse gas, to the best of our knowledge, this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
-Perhaps because previous surveys have included little variation in waterbody size (Webb et al.
-2019, Kortelainen et al) or water quality (Soued et al.
-2015).
+Perhaps because previous surveys have included little variation in waterbody size (@webb_etal_2019, @kortelainen_etal_2020) or water quality (@soued_etal_2016).
 
-# EMISSION RATES
+## Emission Rates
+
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N2O m^-2^ d^-1^ (Table 1).
+This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N2O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N2O m^-2^ d^-1^) (SI Figure 1).
+At the ecoregional scale, only the Northern Appalachians and Northern Plains ecoregions had mean emission rates less than 0.
+These ecoregions also had the lowest mean NO~3~ availability of all nine ecoregions **[need table or figure.
+See ecoreg_stats$post_med_no3 variable.
+maybe a SI table with summaries of all data from all_predictions?]**.
+
+## Fluxes 
+Total N~2~O flux from the surface of CONUS waterbodies is a function of the N~2~O emission rate and waterbody surface area.
+The largest lake size class (> 50 Ha) had the greatest N~2~O flux in all ecoregions, but flux estimates for this lake size class were also the most uncertain (Fig.
+6).
+Small uncertainties in emission rates translate to large uncertainties in flux when expressed across the surface area of large waterbodies.
+Despite widespread N~2~O undersaturation, all ecoregions were a net source of N~2~O to the atmosphere (Table 1).
+Differences in N~2~O flux across ecoregions are explained by patterns in N~2~O saturation and waterbody surface.
+The Coastal Plains ecoregion had the greatest flux, driven by the high levels of N~2~O saturation and extensive waterbody surface area.
+By contrast, the Northern Plains had the lowest N~2~O saturation, waterbody surface area, and N~2~O flux.
+
+At the national scale, we estimate that CONUS waterbodies emit [problem knitting inline code?
+table1_dat %>% filter(`spatial domain` == "national") %>% select(`N2O flux`) %>% pull()] 1049 (358 - 1581) Mg N~2~O year^{-1}, assuming that the emission rates reported here hold for an entire calendar year.
+This is likely a conservative estimate because our data is based on summer conditions, whereas several studies have reported maximum N~2~O emission rates during the winter (citation, see climate benefits paper) and Fall (harsha n2o paper, LD n2o paper) in lakes and reservoirs.
+While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r ipcc.n2o.surface` (22,767) Mg N~2~O year^{-1} (@epa_2023).
+Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions, with the balance emanating from streams and rivers.
+Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00  Mg N~2~O year^{-1}), stationary combustion (84,000 Mg N~2~O year^{-1}), and manure management (66,000  Mg N~2~O year^{-1}).
+
+![Figure 6: N~2~O flux by ecoregion and lake size category](manuscript_figures/n2oFluxBySizeEco.tiff)
+
+## Conclusion
+* nearly three quarters of CONUS lakes and reservoirs are N2O sinks during the summer!
+* despite widespread undersaturation, CONUS lakes and waterbodies are a small source of N2O
+* our study is limited in temporal scope, would seasonality change the story?
+* greatest uncertainty in flux is for large lakes.  Focused investigations on the nation's largest waterbodies to improve estimate?
+
+![SI Figure 1. Distribution of N~2~O emission rates by ecoregion.  Data are inferences made to the population of CONUS waterbodies.](manuscript_figures/SIfigure1.tiff)
 

### 2023-06-21 scripts/dgIndicatorAnalysis.Rmd
Incremental progress on manuscript.  Added discussion of N2O saturation vs lake size
manuscript progress

@@
 
-all_data <- all_predictions %>%
+all_predictions <- all_predictions %>%
   mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
@@
 ```{r}
-all_predictions %>% 
-  filter(.draw == 1) %>% 
-  ggplot(aes(x=e.n2o.mg.m2.d)) + 
-  geom_density() +
-  geom_rug() + 
-  facet_wrap(~WSA9, scales = "free")
-
-knitr::include_graphics((paste0(Sys.getenv("USERPROFILE"), 
+# Compare to literature
+# Read data from Webb et al. 2019
+webb <- readr::read_csv(paste0(Sys.getenv("USERPROFILE"), 
                                "/Environmental Protection Agency (EPA)/",
                                "ORD NLA17 Dissolved Gas - Documents/",
-                                    "EDA/en2oLiterature.tiff")))
+                               "literature/n2o/Webb et al. 2019 PNAS/",
+                               "SI_data.csv")) %>%
+  mutate(e.n2o.mg.m2.d = N2O.umol.m2.d * (44/1000), # umol->mg
+         source = "webb")
+
+# Read data from Deemer et al. 2016
+deemer <- readxl::read_xls(paste0(Sys.getenv("USERPROFILE"), 
+                               "/Environmental Protection Agency (EPA)/",
+                               "ORD NLA17 Dissolved Gas - Documents/",
+                               "literature/n2o/Deemer et al. 2016 Bioscience SI.xls"),
+                           sheet = "Supplementary_Reservoir_GHG_Dat",
+                           skip = 1) %>%
+  mutate(e.n2o.mg.m2.d = `mg N2O-N m-2 d-1` * (44/28),
+         source = "deemer")
+
+# combine NLA and lit data
+all_dat <- rbind(all_predictions %>% 
+  filter(.draw == 1) %>% # just one iteration
+  mutate(source = "nla") %>%
+  select(source, e.n2o.mg.m2.d),
+  webb %>% select(source, e.n2o.mg.m2.d),
+  deemer%>% select(source, e.n2o.mg.m2.d))
+
+# dot plot of N2O emission rates by source
+ggplot(all_dat, aes(source, e.n2o.mg.m2.d)) +
+  geom_point(position = "jitter")
+
+# summary statistics by source
+all_dat %>%
+  split(f = list(.$source)) %>% # split into named lists
+  map(~ summary(.x)) # apply summary function to each list element
 
@@
 n2o <- data.frame(kt = c(1156,88,84,66,60,34,18,14,7),
-                     source = c("Agricultural Soil Management",
+                     source = c("Agricultural Soil Management", # this includes leaching and runoff
                                 "Wastewater Treatment",
@@
   #group_by(.draw) %>%
-  summarise(mean_f = mean(f.n2o.Mg.y)) %>% # 500 means for each WSA9
+  summarise(mean_f = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9
   # now summarize to 1 statistic per WSA9
@@
   geom_hline(yintercept = 0, color='blue') +
-  ylab("mean N2O saturation ratio") +
+  ylab("N2O flux (Mg N2O)") +
   coord_flip() + 
   theme_bw()
-ggsave("output/figures/n2oSatRatioByEco.tiff")
-
-# point and line range by size and ecoregion
+ggsave("output/figures/n2oFuxByEco.tiff")
+
+# point and line range by size and ecoregion: facetted
 all_predictions %>%
@@
   #group_by(.draw) %>%
-  summarise(mean_f.n2o.Mg.y = mean(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
+  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
   # now summarize to 1 statistic per WSA9
@@
   geom_hline(yintercept = 0, color='blue') +
-  ylab("mean N2O flux") +
+  ylab("mean N2O flux (Mg N2O year-1)") +
   coord_flip() + 
@@
 
+# point and line range by size and ecoregion: single plot
+all_predictions %>%
+  group_by(WSA9, size_cat, .draw) %>% # group by iteration
+  #group_by(.draw) %>%
+  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
+    LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
+    UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
+  mutate(unique = paste(WSA9, size_cat, "_")) %>%
+  # mutate(ecoregion = fct_reorder(ecoregion, )) %>%
+  ggplot( aes( y = size_cat, x = estimate ) ) +
+  geom_point() +
+  geom_linerange( aes( xmin = LCL, xmax = UCL)) +
+  geom_vline(xintercept = 0, color='blue') +
+  xlab(expression(N[2]*O~flux~(Mg~year^{-1}))) +
+  #coord_flip() + 
+  theme_bw() +
+  facet_grid(WSA9~., scales = "free")
+ggsave("output/figures/n2oFluxBySizeEcoVerticalPanel.tiff", height = 5.7, width = 2.5)
+
 # bar plot by size category and ecoregion  
@@
   filter(.draw == 1) %>% 
-  group_by(WSA9, size_cat2) %>% 
+  group_by(WSA9, size_cat) %>% 
   summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
-  ggplot(aes(x=size_cat2, y=total.n2o)) + 
+  ggplot(aes(x=size_cat, y=total.n2o)) + 
   geom_bar(stat = "identity") + 
@@
 ggsave("output/figures/n2oFluxByLakeSizeNoError.tiff", g)
+
+# N2O flux vs total surface area in each ecoregion
+all_predictions %>%
+  group_by(WSA9, WSA9_NAME, .draw) %>%
+  summarise(fn2o_Mg = sum(f.n2o.Mg.y),
+             total_sa = sum(area_ha)) %>%
+  summarise( post_med_fn2o = round(median(fn2o_Mg), 3), 
+             LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
+             UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
+             total_sa = mean(total_sa)) %>%
+  ggplot(aes(total_sa, post_med_fn2o)) +
+  geom_point()
 

### 2023-06-21 manuscript/manuscript_support.R

@@ -66,8 +66,6 @@ states <- USAboundaries::us_states() %>%
 
 
 ## load population data----
-# this object is created in dgIndicatorAnalysis.Rmd.  It is Roy's
-# all_predictions object with emission rates and total flux added.
 #tic()
 load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
             "ORD NLA17 Dissolved Gas - Documents\\",
@@ -82,6 +80,7 @@ wsa9names <- structure(list(WSA9 = c("CPL", "NAP", "NPL", "SAP", "SPL",
                                           "Southern Plains", "Temperate Plains", 
                                           "Upper Midwest", "Western Mountains", "Xeric")), 
                        row.names = c(NA, 9L), class = "data.frame")
+
 
 all_predictions <- full_join(all_predictions, wsa9names)
 
@@ -119,7 +118,7 @@ dim(wind) # 465897
 dim(all_predictions) #232948500, 500 observations for each lake
 465897*500 #232948500
 
-
+summary(wind$ws)
 
 ### estimate gas exchange rate for population----
 
@@ -140,11 +139,13 @@ pred.k600 <- function(ws, area) {
 ### calculate emission and flux----
 all_predictions <- all_predictions %>%
   mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
-         # 1000 L to m3. 44ng to 1nmol. ng->ug.
-         e.n2o.ug.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000)) * 
+         ScN2o = 2055.6 - 137.11*surftemp + 4.3173*surftemp^2 - 0.05435*surftemp^3, # schmidt number (Wanninkhof et al 1992)
+         kn2o.cm.h = k600.cm.h * (ScN2o/600)^-(2/3), #k600 to kn2o
+         # 1000 L to m3. 44ng to 1nmol. ng->ug. ug->mg
+         e.n2o.mg.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000 * 1000)) * 
            (k600.cm.h * (24/100)), #h->day. cm->m
-         f.n2o.Mg.y = (e.n2o.ug.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
-           (1000*1000*1000*1000)) #ug-mg, mg-g, g-kg, kg-Mg
+         f.n2o.Mg.y = (e.n2o.mg.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
[…]

### 2023-06-21 scripts/nlaSampleFrameWind.R

@@ -53,8 +53,7 @@ dim(sampFrm) # 465,897, 71
 # awesome package vignette: 
 # https://cran.r-project.org/web/packages/futureheatwaves/vignettes/starting_from_netcdf.html#:~:text=You%20can%20read%20netCDF%20data,connection%20to%20a%20netCDF%20file.
 # ncdf file downloaded  from Climate Data Center contains hourly wind and lake temperature
-# for US.
-# 10u10v24hour.nc file is too big.  crashed system after 32 hours
+# for US. (https://confluence.ecmwf.int/display/CKB/ERA5-Land?src=contextnavpagetreemode)
 wind <- nc_open(paste0(Sys.getenv("USERPROFILE"), 
                        "/Environmental Protection Agency (EPA)/",
                        "ORD NLA17 Dissolved Gas - Documents/",
@@ -77,7 +76,7 @@ wind.time <- as.POSIXct(wind$dim$time$vals * 3600, # convert hours to seconds, w
 wind.time # looks good
 range(wind.time) # May 1 - Nov. 1, range of NLA sample dates
 
-windU <- ncvar_get(wind, varid = "u10") # pull u10 data. 
+windU <- ncvar_get(wind, varid = "u10") # pull u10 data. in m/s
 
 # This variable is in a 3 dimensional array ordered as longitude, then latitude, then time
 dim(windU) # 701 461 4416

### 2023-06-27 manuscript/manuscript_file.Rmd
Reworked conclusion

@@
 
-While NO~3~ and water temperature were the primary drivers of N~2~O production or consumption rates, waterbody morphology determined the extent to which N~2~O processing affected dissolved N~2~O in surface waters.
+While NO~3~ and water temperature were the primary drivers of N~2~O production or consumption rates, dissolved N~2~O concentration was also related to waterbody size.
 When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases with increasing waterbody size class **[Fig.
@@
 Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
-In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column (@carpenter_1983) .
+In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column (@carpenter_1983).
 Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes.
 Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (@Beaulieu_etal_2015, @Beaulieu_etal_2014).
-Unfortunately, the morphometric data required to test these hypothesis (e.g.
-mean depth, max depth) are not available for the samples waterbodies.
 
-Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequlibrim in smaller waterbodies.
+Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequilibrium in smaller waterbodies.
 Air-water gas exchange will vent excess N~2~O from surface waters or facilitate the invasion of air that can replenish depleted N~2~O in undersaturated waterbodies.
@@
 Despite widespread N~2~O undersaturation, all ecoregions were a net source of N~2~O to the atmosphere (Table 1).
-Differences in N~2~O flux across ecoregions are explained by patterns in N~2~O saturation and waterbody surface.
+Differences in N~2~O flux across ecoregions are explained by patterns in N~2~O saturation and waterbody surface (Table 1).
 The Coastal Plains ecoregion had the greatest flux, driven by the high levels of N~2~O saturation and extensive waterbody surface area.
@@
 While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r ipcc.n2o.surface` (22,767) Mg N~2~O year^{-1} (@epa_2023).
-Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions, with the balance emanating from streams and rivers.
+Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions reported by the IPCC (@epa_2023), with the balance emanating from streams and rivers.
 Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00  Mg N~2~O year^{-1}), stationary combustion (84,000 Mg N~2~O year^{-1}), and manure management (66,000  Mg N~2~O year^{-1}).
@@
 ## Conclusion
-* nearly three quarters of CONUS lakes and reservoirs are N2O sinks during the summer!
-* despite widespread undersaturation, CONUS lakes and waterbodies are a small source of N2O
-* our study is limited in temporal scope, would seasonality change the story?
-* greatest uncertainty in flux is for large lakes.  Focused investigations on the nation's largest waterbodies to improve estimate?
+Our survey is the first to demonstrate that most lakes and reservoirs in the CONUS function as N~2~O sinks during the summer.
+Despite widespread N~2~O uptake, CONUS waterbodies function as a net source of N~2~O due to high emissions from from a subset of the population.
+Dissolved N~2~O in lakes and reservoirs can be predicted from non-linear interactions between water temperature, NO~3~ availability, and waterbody size.
+NO~3~ availability largely determines whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink is exaggerated with increasing water temperature.
+The degree of N~2~O disequilibrium also increases with decreasing waterbody size, possibly because smaller waterbodies have low air-water gas exchange rates.
+
+Although waterbodies in the largest lake size class tend to have N~2~O levels close to equilibrium, they emit more N~2~O than the other size classes due to their expansive surface area.
+The N~2~O flux from these lakes is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved N~2~O translate to large uncertainties in flux when propagated across the surface of large waterbodies.
+Efforts to improve the accuracy of the lentic N~2~O flux estimate for the U.S.
+should therefore focus on investigations of the nations largest lakes and reservoirs.
+
+The smallest lake size classes exhibited both the highest and lowest levels of N~2~O saturation, depending on NO~3~ availability.
+It is unclear whether this variability is related to differences in NO~3~ loading or temporal variation.
+For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted (@beaulieu_etal_2014).
+A better understanding of the temporal evolution of N~2~O production and consumption in small waterbodies is an important research need to determine the extent to which these systems can remove excess nitrogen from surface waters and simulatenously mitigate N~2~O emissions.
 

### 2023-06-29 manuscript/manuscript_file.Rmd
Adding methods to ms file; updated references file; added csl reference style file

@@
 link-citations: yes
+csl: nature_geoscience.csl
 ---
@@
 (CONUS).
-A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
+A generalized random-tessellation survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
 compare my sample IDs to NLA IDs]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1).
@@
 `r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks (Figure 1).
-We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897.
+We modeled the data using a Bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897.
 See technical support document]**.
@@
 
-The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
-This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere (@mosier_etal_1998) and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
+The extent of N~2~O undersaturation in CONUS waterbodies was surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
+This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere [@mosier_etal_1998] and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
 @Webb_etal_2019 found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and @Soued_etal_2016 report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020) but kortelainen and Soued are mostly pristine lakes]**.
@@
 # References
+
+<div id="refs"></div>
+
+# Methods
+
+## Survey design and inference
+
+The population of interest (POI) for the 2017 NLA survey [@USEPA_NLA17_2022] was CONUS waterbodies of at least 1 hectare in area, at least 1 meter deep, with at least 0.1 hectare of open water, and a minimum residence time of one week (Figure 1).
+Inferences from the survey samples were assumed relevant to average conditions during the summer index period between July and early September 2017.
+The POI was sampled according to a generalized random-tessellation design with stratification [@Stevens_Olsen_2004], which prescribed a spatially balanced, probabilistic sample with stratification along three factors: surface area (hectares; 5 ordered levels), ecoregion [@Omernik_1987; @Herlihy_etal_2008; @Omernik_Griffith_2014] (9 levels), and US state (excluding AK and HI; 48 levels).
+The size categories were sampled with unequal probability, such that larger lakes were over-sampled relative to smaller lakes.
+In total, 1091 waterbodies were sampled; and 95 of those waterbodies were re-visited for a second sample.
+The revisit samples were used to assess out-of-sample model performance.
+
+The survey samples were treated statistically as a randomly selected subset of the complete POI, but with a pattern of "missingess" prescribed by the selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
+For inferences, estimates were statistically adjusted to account for the selection mechanism using a Multilevel regression with poststratification (MRP) approach [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020].
+MRP is a model-based approach carried out in two steps.
+In the first step, the sampled responses of interest (_e.g._, N~2~O) are conditioned on the survey design variables (_e.g._, ecoregion, state, size) in a multilevel regression model.
+Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the population of interest [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
+In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (_e.g.,_ US waterbodies) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (_i.e._, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (_e.g._, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
+The fully Bayesian implementation made propagation of uncertainty from the waterbody-level predictions to higher level summaries straightforward.
+
+The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio; and (3) the proportion of under-saturated water bodies (_i.e._, saturation ratio < 1).
+Saturation ratio was calculated as the ratio of dissolved to equilibrium N~2~O.
+Because dissolved and equilibrium N~2~O were each measured on the same sample units (_i.e._, a single waterbody), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, common dependencies on geography.
+The modeled response variable was, therefore, multivariate.
+Although the mean marginal probabilities estimated from separate models may have provided comparable estimates, the joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
+The saturation ratio was not modeled or predicted directly, but was estimated _post hoc_ as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
+Similarly, the proportion of under-saturated waterbodies was estimated from this PPD as the proportion of waterbodies with a saturation ratio less than 1.
+All models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models using Hamiltonian Monte Carlo (HMC).
+
+Fully reproducible code and further results and model workflow details are provided in Modeling **Supplement X**.
+Though our initial models conditioned the response on just the survey design variables, those models consistently resulted in poorer fits to to the observed N~2~O data compared to models with additional covariates.
+Though the initial models seemed to work well at replicating the central tendency of the dissolved N~2~O data, simulations of other properties of the observed data, such as the variance, skewness, and kurtosis, were less satisfactory.
+The relatively poor fit to this aspect of the observed data also resulted in a downward bias in estimates of the observed saturation ratio.
+However, including a measure of NO~3~ concentration as a covariate on dissolved N~2~O resulted in substantially improved posterior predictive model checks.
+In addition, including surface temperature and elevation measures as covariates on equilibrium N~2~O resulted in substantially improved fits, as did adding flexibility in the distributional terms for both dissolved and equilibrium N~2~O.
+Both dissolved and equilibrium N~2~O were allowed to vary according to the survey design variables and environmental covariates in the final model presented here.
+
+Including these helpful covariates, however, led to some complexities in the MRP process.
+To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled.
+For the lake area and elevation covariates, data was available for all lakes in the POI from previously compiled geospatial databases.
+However, surface temperature and NO~3~ were not observed for lakes outside of the sample.
+Therefore, a more complex model was developed to include surface temperature and NO3 as additional responses conditioned on the survey design variables and the fully observed covariates.
+This approach to inference was akin to Bayesian structural equation modeling approaches [@Merkle_etal_2021; @Merkle_Rosseel_2018].
+The dependence structure of the represented by this model may be characterized as:
+
+$$\begin{aligned} 
+{p( \boldsymbol{N_2O_{diss}}}\ &|\ Survey,\ Area,\ {\boldsymbol{NO_3}},\ {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{N_2O_{equi}}}\ &|\ Survey,\ Elev,\ {\boldsymbol{Temp}})\\
+{p( \boldsymbol{NO_3}}\ &|\ Survey,\ Area,\ {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{Temp}}\ &|\ Survey,\ Lat,\ Elev,\ Day)
+\end{aligned}$$
+
+Variables in bold text above were treated as partially observed with respect to the POI (_i.e._, observed only in the sample), whereas variables not in bold were fully observed.
+The partially observed variables, including dissolved and equilibrium N~2~O, NO~3~, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.
+In order to make predictions and propagate uncertainty through the dependency structure, the fitted multivariate model was used to first predict surface temperature in the POI, since it depended only on the fully observed covariates.
+That predictive distribution was then used alongside the relevant fully observed covariates to predict NO~3~ in the POI.
+Finally, the predictive distributions for temperature and NO~3~ were used in predicting both N~2~O responses in the POI.

### 2023-06-30 manuscript/manuscript_file.Rmd
Minor updates to manuscript.

--- 2023-06-27 manuscript/manuscript_file.Rmd
@@
 
--   Table and figure captions should specify if displaying sample vs population estimates. "posterior predictive distribution"
--   Note that right tail is truncated in figure 2.
+- include a better description of full range of lake size: sample and population.  also add NO3 to better support emission rate statement.
+- incorporate in Methods section how/why size is incorporated as categorical variable and continuos variable.
+- maybe discuss upper midwest.  Lots of lake area, but largely at equilibrium, therefore very small fluxes.  Large area doesn't necessarily mean big flux.
+-  last sentence of emission rates.  we should look at summary stats of observed NO3 rather than predicted 
+-  Mike will take a stab at big SI table.
 
@@
 The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
-This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere (@mosier_etal_1998) and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
+This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere (@Mosier_etal_1998) and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
 @Webb_etal_2019 found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and @Soued_etal_2016 report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020) but kortelainen and Soued are mostly pristine lakes]**.
@@
 flextable(table1_dat) %>%
-  set_header_labels('waterbody surface area' = "waterbody surface area\n(hectares)",
+  set_header_labels('waterbody surface area' = "total waterbody surface area\n(hectares)",
                     'N2O undersaturation extent' = "N2O undersaturation extent\n(% waterbodies)",
                     'N2O emission rate' = "N2O emission rate\n(mg N2O m-2 d-1)",
-                    'N2O flux' = "N2O flux\n(Mg N2O y-1)") %>%
+                    'N2O flux' = "N2O flux\n(metric ton N2O y-1)") %>% # 1 megagram (Mg) = 1 1000 kg = 1 metric ton
   align(j=c(2:6), align = "center", part = "all") %>% # center align columns 2:4
@@
     caption = as_paragraph( # allows for formatting
-      as_chunk("Table 1. Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, emission rate, and flux at the national and ecoregional scales", 
+      as_chunk("Table 1. Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, emission rate, and flux at the national and ecoregional scales.", 
                props = fp_text_default(bold = TRUE) # here is where we specify formatting (font, size, bold, etc)
@@
     align_with_table = FALSE) %>% # table is center aligned, ignore that and apply default left align to caption
+
+  footnote(i = 1, j = 3:6,
+           ref_symbols = "a",
+           value = as_paragraph("Mean and 95% credible interval of 500 realizations of posterior predictive distribution"),
+           part = "header"
+  ) %>%
+    footnote(i = 1, j = 3,
+           ref_symbols = ", b",
+           value = as_paragraph("Percent of population with N2O saturation ratio < 1."),
+           part = "header",
+           inline = TRUE
+  ) %>%
+  footnote(i = 1, j = 5:6,
+           ref_symbols = ", c",
+           value = as_paragraph("Negative values indicate N2O uptake"),
+           part = "header",
+           inline = TRUE
+  ) %>%
+  fontsize(size = 8, part = "all") %>% # doesn't apply to caption when as_paragraph is invoked.
+  fontsize(size = 7, part = "footer") %>% #smaller font for footnote
     autofit() %>% # convenient column spacing
-  fontsize(size = 8, part = "all") # doesn't apply to caption when as_paragraph is invoked.
+  fit_to_width(ft_1, max_width = 11) # scale to fill 8x11 paper, landscape
 
@@
 Dissolved N~2~O concentration increased with NO~3~ concentration **[Fig.
-3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@knowles_1982, @richardson_etal_2009].
+3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Knowles_1982, @Richardson_etal_2009].
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
-Incomplete denitrification often occurs when NO~3~ is relatively abundant [@Beaulieu_etal_2015, @Beaulieu_etal_2014], likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification [@koike_etal_1975, @richardson_etal_2009].
+Incomplete denitrification often occurs when NO~3~ is relatively abundant [@Beaulieu_etal_2015, @Beaulieu_etal_2014], likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification [@Koike_etal_1975, @Richardson_etal_2009].
 When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~, thereby functioning as an N~2~O sink.
@@
 Nitrification might be expected to be more prevalent in the predominantly well oxygenated surface waters where the N~2~O samples were collected than denitrification, which is an anaeorbic process.
-However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column **[citation]** and can be transported from aoxic sediments to the water surface via advection and diffusion [@delsontro_etal_2018].
+However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column **[citation]** and can be transported from aoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
 Thus the N~2~O - NO3 relationship may be reflective of some combination of N~2~O production/consumption via denitrification and N~2~O production via nitrification.
@@
 ![Figure 3.
-A) Relationship between N~2~O saturation and nitrate (NO~3~) concentration, dependent upon surface water temperature.
-The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13.
-B) Relationship between N~2~O saturation and surface water temperature, dependent upon NO~3~ category.
+Conditional effects plot illustrating A) the relationship between N~2~O saturation and nitrate (NO~3~) concentration, dependent upon surface water temperature.
+The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13, and B) relationship between N~2~O saturation and surface water temperature, dependent upon NO~3~ category.
 Horizontal dashed lines in both panels reflects an N~2~O saturation value of 1.](manuscript_figures/sat_cond_surftemp.tiff)
@@
 Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
-In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column (@carpenter_1983).
+In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column (@Carpenter_1983).
 Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes.
@@
 The vertical blue line references a N~2~O saturation value of 1.0.
-The horizontal red arrows highlight $\Delta$ N~2~O for the smallest lake size class.](manuscript_figures/deltaN2ObySize.tiff)
+The horizontal red arrows highlight $\Delta$ N~2~O for the smallest lake size class.
+Mean (points) and 95% credible interview (horizontal bars) are derived from 500 realizations of the poterior predictive distribution.
+](manuscript_figures/deltaN2ObySize.tiff)
 
@@
 Large waterbodies, regardless of NO~3~ status, tend to have N~2~O saturation ratios closer to equilibrium.
-Therefore intra and inter regional patterns in N~2~O saturation are determined by the joint distributions of waterbody size, NO~3~ availability, and water temperature.
+Therefore intra and inter regional patterns in N~2~O saturation can be well predicted by the joint distributions of waterbody size, NO~3~ availability, and water temperature.
 
 ![Figure 5.
-Partial dependence plots illustrating the relationship between N~2~O saturation and A) nitrate (NO~3~ availability dependent upon lake size, and B) lake size dependent upon NO~3~ availability.
+Conditional effects plots illustrating the relationship between N~2~O saturation and A) nitrate (NO~3~ availability dependent upon lake size, and B) lake size dependent upon NO~3~ availability.
 The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13.](manuscript_figures/sat_cond_area.tiff)
@@
 2011), another greenhouse gas, to the best of our knowledge, this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
-Perhaps because previous surveys have included little variation in waterbody size (@webb_etal_2019, @kortelainen_etal_2020) or water quality (@soued_etal_2016).
+Perhaps because previous surveys have included little variation in waterbody size (@Webb_etal_2019, @Kortelainen_etal_2020) or water quality (@Soued_etal_2016).
 
@@
 
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N2O m^-2^ d^-1^ (Table 1).
-This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N2O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N2O m^-2^ d^-1^) (SI Figure 1).
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^ (Table 1).
+This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) (SI Figure 1).
 At the ecoregional scale, only the Northern Appalachians and Northern Plains ecoregions had mean emission rates less than 0.
@@
 At the national scale, we estimate that CONUS waterbodies emit [problem knitting inline code?
-table1_dat %>% filter(`spatial domain` == "national") %>% select(`N2O flux`) %>% pull()] 1049 (358 - 1581) Mg N~2~O year^{-1}, assuming that the emission rates reported here hold for an entire calendar year.
-This is likely a conservative estimate because our data is based on summer conditions, whereas several studies have reported maximum N~2~O emission rates during the winter (citation, see climate benefits paper) and Fall (harsha n2o paper, LD n2o paper) in lakes and reservoirs.
-While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r ipcc.n2o.surface` (22,767) Mg N~2~O year^{-1} (@epa_2023).
+table1_dat %>% filter(`spatial domain` == "national") %>% select(`N2O flux`) %>% pull()] 1049 (358 - 1581) metric tons N~2~O year^-1^, assuming that the emission rates reported here hold for an entire calendar year.
+While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ (@epa_2023).
 Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions reported by the IPCC (@epa_2023), with the balance emanating from streams and rivers.
-Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00  Mg N~2~O year^{-1}), stationary combustion (84,000 Mg N~2~O year^{-1}), and manure management (66,000  Mg N~2~O year^{-1}).
+Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^).
 
-![Figure 6: N~2~O flux by ecoregion and lake size category](manuscript_figures/n2oFluxBySizeEco.tiff)
+Our annual flux estimates assume that emission rate estimates based on samples collected during the summer reflect annual conditions.
+The literature contains strong evidence of seasonality in N~2~O emission rates, but the patterns differ among studies.
+Some systems have maximum N~2~O emission rates during the winter (@Musenze_etal_2014, @Kortelainen_etal_2020), others in the Fall (@Beaulieu_etal_2014), and yet others during the summer (@Beaulieu_etal_2010).
+While our assumption of constant emissions across the year adds uncertainty to the annual flux estimate, it is not clear that the estimate is biased high or low.
+
+![Figure 6: N~2~O flux by ecoregion and lake size category.
+Mean (points) and 95% credible interval (horizontal bars) reflect the distribution of 500 realizations of the posterior predictive distribution.](manuscript_figures/n2oFluxBySizeEco.tiff)
 
@@
 It is unclear whether this variability is related to differences in NO~3~ loading or temporal variation.
-For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted (@beaulieu_etal_2014).
-A better understanding of the temporal evolution of N~2~O production and consumption in small waterbodies is an important research need to determine the extent to which these systems can remove excess nitrogen from surface waters and simulatenously mitigate N~2~O emissions.
+For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted (@Beaulieu_etal_2014).
+A better understanding of the temporal evolution of N~2~O production and consumption in small waterbodies is an important research need to determine the extent to which these systems can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
 
-![SI Figure 1. Distribution of N~2~O emission rates by ecoregion.  Data are inferences made to the population of CONUS waterbodies.](manuscript_figures/SIfigure1.tiff)
+![SI Figure 1.
+Distribution of N~2~O emission rates by ecoregion.
+Each blue line (n=500) represents the distribution of values from a single realization of the posterior prediction.](manuscript_figures/SIfigure1.tiff)
 

### 2023-06-30 manuscript/manuscript_support.R

@@ -393,7 +393,7 @@ ggsave("manuscript/manuscript_figures/deltaN2ObySize.tiff", width = 6, height = 
 
 }
 
-## Figure X: N2O flux by waterbody size 
+## Figure X: N2O flux by waterbody size------
 # See 'N2O emission rate and flux: Figures and tables' section of dgIndicatorAnalysis.Rmd
 # if the image is already on computer, then nothing, else create image
 
@@ -412,7 +412,8 @@ all_predictions %>%
   geom_point() +
   geom_linerange(aes(xmin = LCL, xmax = UCL)) +
   geom_vline(xintercept = 0, color='blue') +
-  ylab(expression(N[2]*O~flux~(Mg~year^{-1}))) +
+  xlab(expression(N[2]*O~flux~(metric~tons~year^{-1}))) + # 1Mg = 1 metric ton
+    ylab("Lake size class (ha)") +
   theme_bw() +
   facet_wrap(~ecoregion, scales = "free")
 ggsave("manuscript/manuscript_figures/n2oFluxBySizeEco.tiff", height = 5.4, width = 9.6)
@@ -523,7 +524,7 @@ propN2oSinkUpperNPL <- all_predictions %>%
 # 1 mass of N2O = 265 mass CO2 www.epa.gov/system/files/documents/2023-04/US-GHG-Inventory-2023-Chapter-1-Introduction.pdf
 # 1 MMT (million metric ton) = 1000 kt (kilo ton) www.epa.gov/system/files/documents/2023-04/US-GHG-Inventory-2023-Chapter-1-Introduction.pdf
 # 1 ton = 1000 kg
-# 1000 kg = 1 Mg
+# 1000 kg = 1 Mg = 1 metric ton
 # (15.2+2.9) = 18 MMT CO2 eq * (1MMT N2O / 265 MMT CO2 Eq) * (1000kt N2O/1MMt N2O) * (1000t N2O/1kt N2O) * (1000kg N2O/1t N2O) * (1Mg/1000kg)
 ipcc.n2o.indirect <- (15.2+2.9)*(1/265)*(1000/1)*(1000/1)*(1000/1)*(1/1000) #68,302 Mg N2O
 

### 2023-06-30 manuscript/manuscript_file.Rmd

@@
 (CONUS).
-A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
-compare my sample IDs to NLA IDs]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1).
+A generalized random-tessellation survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
+compare my sample IDs to NLA IDs]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week **[(Fig.
+1)]**.
 Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?).
-`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks (Figure 1).
-We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897.
+`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **[(Fig.
+1)]**.
+We modeled the data using a Bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897.
 See technical support document]**.
@@
 The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
-This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere (@Mosier_etal_1998) and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
-@Webb_etal_2019 found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and @Soued_etal_2016 report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020) but kortelainen and Soued are mostly pristine lakes]**.
+This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
+Webb et al.
+[@Webb_etal_2019] found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Soued et al.
+[@Soued_etal_2016] report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020) but kortelainen and Soued are mostly pristine lakes]**.
 While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
@@
 
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean N~2~O saturation ratio (ratio of measured to equilibrium concentration) was equal to or greater than 1 in seven of the nine ecoregions **[(Table 1, Figure 2)]**, suggesting that, on average, waterbodies in these ecoregions were supersatured with N~2~O and were therefore functioning as a source of N~2~O.
-The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations that exceed equilibrium in most ecoregions can be attributed to the distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies (Figure 2).
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean N~2~O saturation ratio (ratio of measured to equilibrium concentration) was equal to or greater than 1 in seven of the nine ecoregions **[(Table 1, Fig.
+2)]**, suggesting that, on average, waterbodies in these ecoregions were supersatured with N~2~O and were therefore functioning as a source of N~2~O.
+The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations that exceed equilibrium in most ecoregions can be attributed to the distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies **[(Fig.
+2)]**.
 Across the CONUS, and within each of the nine major ecoregions, the distribution of N~2~O saturation ratios had median values that were less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
@@
 Dissolved N~2~O concentration increased with NO~3~ concentration **[Fig.
-3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Knowles_1982, @Richardson_etal_2009].
+3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Knowles_1982; @Richardson_etal_2009].
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
-Incomplete denitrification often occurs when NO~3~ is relatively abundant [@Beaulieu_etal_2015, @Beaulieu_etal_2014], likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification [@Koike_etal_1975, @Richardson_etal_2009].
+Incomplete denitrification often occurs when NO~3~ is relatively abundant [@Beaulieu_etal_2015; @Beaulieu_etal_2014], likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification [@Koike_etal_1975; @Richardson_etal_2009].
 When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~, thereby functioning as an N~2~O sink.
@@
 Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes.
-Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface (@Beaulieu_etal_2015, @Beaulieu_etal_2014).
+Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface [@Beaulieu_etal_2015; @Beaulieu_etal_2014].
 
@@
 2011), another greenhouse gas, to the best of our knowledge, this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
-Perhaps because previous surveys have included little variation in waterbody size (@Webb_etal_2019, @Kortelainen_etal_2020) or water quality (@Soued_etal_2016).
+Perhaps because previous surveys have included little variation in waterbody size [@Webb_etal_2019; @Kortelainen_etal_2020] or water quality [@Soued_etal_2016].
 
@@
 Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^ (Table 1).
-This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) (SI Figure 1).
+This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **[(SI Fig.
+1)]**.
 At the ecoregional scale, only the Northern Appalachians and Northern Plains ecoregions had mean emission rates less than 0.
@@
 Total N~2~O flux from the surface of CONUS waterbodies is a function of the N~2~O emission rate and waterbody surface area.
-The largest lake size class (> 50 Ha) had the greatest N~2~O flux in all ecoregions, but flux estimates for this lake size class were also the most uncertain (Fig.
-6).
+The largest lake size class (> 50 Ha) had the greatest N~2~O flux in all ecoregions, but flux estimates for this lake size class were also the most uncertain **[(Fig.
+6)]**.
 Small uncertainties in emission rates translate to large uncertainties in flux when expressed across the surface area of large waterbodies.
@@
 
-At the national scale, we estimate that CONUS waterbodies emit [problem knitting inline code?
-table1_dat %>% filter(`spatial domain` == "national") %>% select(`N2O flux`) %>% pull()] 1049 (358 - 1581) metric tons N~2~O year^-1^, assuming that the emission rates reported here hold for an entire calendar year.
-While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ (@epa_2023).
-Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions reported by the IPCC (@epa_2023), with the balance emanating from streams and rivers.
+At the national scale, we estimate that CONUS waterbodies emit `r table1_dat %>% janitor::clean_names() %>% filter(spatial_domain == "national") %>% select(n2o_flux) %>% pull()` metric tons N~2~O year^-1^, assuming that the emission rates reported here hold for an entire calendar year.
+While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
+Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions reported by the IPCC [@epa_2023], with the balance emanating from streams and rivers.
 Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^).
@@
 The literature contains strong evidence of seasonality in N~2~O emission rates, but the patterns differ among studies.
-Some systems have maximum N~2~O emission rates during the winter (@Musenze_etal_2014, @Kortelainen_etal_2020), others in the Fall (@Beaulieu_etal_2014), and yet others during the summer (@Beaulieu_etal_2010).
+Some systems have maximum N~2~O emission rates during the winter [@Musenze_etal_2014; @Kortelainen_etal_2020], others in the Fall [@Beaulieu_etal_2014], and yet others during the summer [@Beaulieu_etal_2010].
 While our assumption of constant emissions across the year adds uncertainty to the annual flux estimate, it is not clear that the estimate is biased high or low.
@@
 It is unclear whether this variability is related to differences in NO~3~ loading or temporal variation.
-For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted (@Beaulieu_etal_2014).
+For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted [@Beaulieu_etal_2014].
 A better understanding of the temporal evolution of N~2~O production and consumption in small waterbodies is an important research need to determine the extent to which these systems can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
@@
 # References
+
+<div id="refs"></div>
+
+# Methods
+
+## Survey design
+
+The population of interest (POI) for the 2017 NLA survey [@USEPA_NLA17_2022] was CONUS waterbodies of at least 1 hectare in area, at least 1 meter deep, with at least 0.1 hectare of open water, and a minimum residence time of one week **[(Fig.
+1)]**.
+Inferences from the survey samples were assumed relevant to average conditions during the summer index period between June and end of September 2017.
+The POI was sampled according to a generalized random-tessellation design with stratification [@Stevens_Olsen_2004], which prescribed a spatially balanced, probabilistic sample with stratification along three factors: surface area (hectares; 5 ordered levels), ecoregion [@Omernik_1987; @Herlihy_etal_2008; @Omernik_Griffith_2014] (9 levels), and US state (excluding AK and HI; 48 levels).
+The size categories were sampled with unequal probability, such that larger lakes were over-sampled relative to smaller lakes.
+In total, 1091 waterbodies were sampled; and 95 of those waterbodies were re-visited for a second sample.
+The revisit samples were used to assess out-of-sample model performance.
+
+## Sampling
+Samples were collected from approximately the middle of each waterbody for analysis of algal toxins (microcystins and cylindrospermopsin), benthic macroinvertebrates,  phytoplankton, atrazine, water chemistry, chlorophyll‐a, zooplankton, and dissolved N2O.
+Duplicate dissolved N~2~O  samples were collected approximately 0.05 m below the water surface by drawing 105 mL of water into a 140 mL syringe equipped with a stopcock and containing 35 mL of air.
+The syringes were gently shaken for five minutes and the headspace gas transferred to a 30 mL syringe.
+Finally, the gas sample was transferred into a 12 mL pre-evacuated (<50 mTorr) glass vial (Labco, Wales, UK) sealed with a PTFE/Silicone stacked on top of a grey chlorobutyl septa.
+The temperatue of the water in the syringe was recorded immediately after the gas was transferred into the vial.
+Duplicate 20 mL air samples were collected from approximately 2 m above the water surface and transferred to 12 mL glass vials.
+A vertical profile of water temperatue was measured at the same site.
+
+Nitrous oxide partial pressure in the gas samples was measured via gas chromatography (Bruker 450, Billerica, MA, USA), an electron capture detector, and autosampler (CTC Analytics, PAL-xt, Zwingen, Switzerland).
+A 90:10 mixture of argon and methane was used as the carrier gas.
+The original dissolved N~2~O concentration was calculated from the measured headspace partial pressue, the temperature specific Bunsen coefficient, and a mass balance for the headspace equilibration system as described in Beaulieu et al.
+[@Beaulieu_etal_2012].
+The equilibrium dissolved gas concentration at the time of sampling was calculated from the measured N~2~O partial pressure in the air samples, the measured barometric pressure, and the water temperature.
+Dissolved N~2~O concentration is presented as a saturation ratio, defined as the ratio of measured to equilibrium concentration.
+
+Analytical, quality assurance, and field sampling details for all measurements can be found at the 2017 National Lakes Assessment Laboratory Operations Manual[@epa_lom], 2017 National Lakes Assessment Quality Assurance Project Plan [@epa_qapp], and 2017 National Lakes Assessment Field Operations Manual [@epa_fom], respectively.
+All data can be accessed online [@epa_data].
+
+## Statistical inference
+
+The survey samples were treated statistically as a randomly selected subset of the complete POI, but with a pattern of "missingess" prescribed by the selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
+For inferences, estimates were statistically adjusted to account for the selection mechanism using a Multilevel regression with poststratification (MRP) approach [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020].
+MRP is a model-based approach carried out in two steps.
+In the first step, the sampled responses of interest (_e.g._, N~2~O) are conditioned on the survey design variables (_e.g._, ecoregion, state, size) in a multilevel regression model.
+Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the population of interest [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
+In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (_e.g.,_ US waterbodies) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (_i.e._, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (_e.g._, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
+The fully Bayesian implementation made propagation of uncertainty from the waterbody-level predictions to higher level summaries straightforward.
+
+The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio; and (3) the proportion of under-saturated water bodies (_i.e._, saturation ratio < 1).
+Saturation ratio was calculated as the ratio of dissolved to equilibrium N~2~O.
+Because dissolved and equilibrium N~2~O were each measured on the same sample units (_i.e._, a single waterbody), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, common dependencies on geography.
+The modeled response variable was, therefore, multivariate.
+Although the mean marginal probabilities estimated from separate models may have provided comparable estimates, the joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
+The saturation ratio was not modeled or predicted directly, but was estimated _post hoc_ as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
+Similarly, the proportion of under-saturated waterbodies was estimated from this PPD as the proportion of waterbodies with a saturation ratio less than 1.
+All models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models using Hamiltonian Monte Carlo (HMC).
+
+Fully reproducible code and further results and model workflow details are provided in Modeling **Supplement X**.
+Though our initial models conditioned the response on just the survey design variables, those models consistently resulted in poorer fits to to the observed N~2~O data compared to models with additional covariates.
+Though the initial models seemed to work well at replicating the central tendency of the dissolved N~2~O data, simulations of other properties of the observed data, such as the variance, skewness, and kurtosis, were less satisfactory.
+The relatively poor fit to this aspect of the observed data also resulted in a downward bias in estimates of the observed saturation ratio.
+However, including a measure of NO~3~ concentration as a covariate on dissolved N~2~O resulted in substantially improved posterior predictive model checks.
+In addition, including surface temperature and elevation measures as covariates on equilibrium N~2~O resulted in substantially improved fits, as did adding flexibility in the distributional terms for both dissolved and equilibrium N~2~O.
+Both dissolved and equilibrium N~2~O were allowed to vary according to the survey design variables and environmental covariates in the final model presented here.
+
+Including these helpful covariates, however, led to some complexities in the MRP process.
+To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled.
+For the lake area and elevation covariates, data was available for all lakes in the POI from previously compiled geospatial databases.
+However, surface temperature and NO~3~ were not observed for lakes outside of the sample.
+Therefore, a more complex model was developed to include surface temperature and NO3 as additional responses conditioned on the survey design variables and the fully observed covariates.
+This approach to inference was akin to Bayesian structural equation modeling approaches [@Merkle_etal_2021; @Merkle_Rosseel_2018].
+The dependence structure of the represented by this model may be characterized as:
+
+$$\begin{aligned} 
+{p( \boldsymbol{N_2O_{diss}}}\ &|\ Survey,\ Area,\ {\boldsymbol{NO_3}},\ {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{N_2O_{equi}}}\ &|\ Survey,\ Elev,\ {\boldsymbol{Temp}})\\
+{p( \boldsymbol{NO_3}}\ &|\ Survey,\ Area,\ {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{Temp}}\ &|\ Survey,\ Lat,\ Elev,\ Day)
+\end{aligned}$$
+
+Variables in bold text above were treated as partially observed with respect to the POI (_i.e._, observed only in the sample), whereas variables not in bold were fully observed.
+The partially observed variables, including dissolved and equilibrium N~2~O, NO~3~, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.
+In order to make predictions and propagate uncertainty through the dependency structure, the fitted multivariate model was used to first predict surface temperature in the POI, since it depended only on the fully observed covariates.
+That predictive distribution was then used alongside the relevant fully observed covariates to predict NO~3~ in the POI.
+Finally, the predictive distributions for temperature and NO~3~ were used in predicting both N~2~O responses in the POI.

### 2023-06-30 manuscript/manuscript_file.Rmd
Mostly formatting and citation updates to manuscript

@@
 -  Mike will take a stab at big SI table.
+-  Citation “epa_data” isn’t populating citation 26 in bibliography.  The bibTex entry is classified as a “website”.  Roy, can you attempt to make this citation play nice with nature_geosciences.csl?
 
@@
 At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`), the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), and \>65% of the waterbodies in every ecoregion were undersaturated (Table 1).
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`), the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), and \>65% of the waterbodies in every ecoregion were undersaturated **[(Table 1)]**.
 
@@
 
-##### NO3
+##### NO~3~
 
 Dissolved N~2~O concentration increased with NO~3~ concentration **[Fig.
-3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Knowles_1982; @Richardson_etal_2009].
+3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Kuypers_etal_2018; @Richardson_etal_2009].
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
@@
 It is also possible that the correlation between N~2~O and NO~3~ reflects the simultaneous production of these solutes via nitrification.
-Nitrification, an aerobic process which can produce N~2~O as a byproduct, is the microbial oxidation of ammonium (NH~4~^+^) to NO~3~ **[citation]**.
+Nitrification, an aerobic process which can produce N~2~O as a byproduct, is the microbial oxidation of ammonium (NH~4~^+^) to NO~3~ [@Kuypers_etal_2018].
 Nitrification might be expected to be more prevalent in the predominantly well oxygenated surface waters where the N~2~O samples were collected than denitrification, which is an anaeorbic process.
-However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column **[citation]** and can be transported from aoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
-Thus the N~2~O - NO3 relationship may be reflective of some combination of N~2~O production/consumption via denitrification and N~2~O production via nitrification.
+However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column [@Broman_etal_2021] and can be transported from aoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
+Thus the N~2~O - NO~3~ relationship may be reflective of some combination of N~2~O production/consumption via denitrification and N~2~O production via nitrification.
 
@@
 3]**, which is consistent with N~2~O production and consumption occurring through temperature sensitive biological processes, but the relationship was dependent on NO~3~ availability.
-At low NO3 concentrations, temperature had a relatively minor affect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
+At low NO~3~ concentrations, temperature had a relatively minor affect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
 The effect of temperature became more pronounced when NO~3~ availability was greater, possibly reflecting the stimulatory affect of higher temperature on biological N~2~O production.
@@
 Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
-In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column (@Carpenter_1983).
+In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column [@Carpenter_1983].
 Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes.
@@
 Regardless of whether a waterbody is superaturated or undersatured in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
-One of the primary drivers of air-water gas exchange in lentic waters is wind speed (@Vachon_etal_2013).
+One of the primary drivers of air-water gas exchange in lentic waters is wind speed [@Vachon_etal_2013].
 Wind speeds tend to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
@@
 5A]**.
-At low NO3 concentrations, N~2~O is undersatured in small waterbodies, but increases toward equilibrium in larger waterbodies **[Fig.
+At low NO~3~ concentrations, N~2~O is undersatured in small waterbodies, but increases toward equilibrium in larger waterbodies **[Fig.
 5A]**.
@@
 These interactions yield a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enable the identifcation of broad scale patterns.
-First, small waterbodies with low NO3 concentration tend to be undersaturated, whereas small waterbodies with high NO3 tend to be highly supersaturated.
+First, small waterbodies with low NO~3~ concentration tend to be undersaturated, whereas small waterbodies with high NO~3~ tend to be highly supersaturated.
 Large waterbodies, regardless of NO~3~ status, tend to have N~2~O saturation ratios closer to equilibrium.
@@
 
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^ (Table 1).
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **[(Table 1)]**.
 This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **[(SI Fig.
@@
 Small uncertainties in emission rates translate to large uncertainties in flux when expressed across the surface area of large waterbodies.
-Despite widespread N~2~O undersaturation, all ecoregions were a net source of N~2~O to the atmosphere (Table 1).
-Differences in N~2~O flux across ecoregions are explained by patterns in N~2~O saturation and waterbody surface (Table 1).
+Despite widespread N~2~O undersaturation, all ecoregions were a net source of N~2~O to the atmosphere **[(Table 1)]**.
+Differences in N~2~O flux across ecoregions are explained by patterns in N~2~O saturation and waterbody surface **[(Table 1)]**.
 The Coastal Plains ecoregion had the greatest flux, driven by the high levels of N~2~O saturation and extensive waterbody surface area.
@@
 
-At the national scale, we estimate that CONUS waterbodies emit `r table1_dat %>% janitor::clean_names() %>% filter(spatial_domain == "national") %>% select(n2o_flux) %>% pull()` metric tons N~2~O year^-1^, assuming that the emission rates reported here hold for an entire calendar year.
+At the national scale, we estimate that CONUS waterbodies emit `r table1_dat %>% janitor::clean_names() %>% filter(spatial_domain == "national") %>% select(n2o_flux) %>% pull()` metric tons N~2~O year^-1^.
 While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
 Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions reported by the IPCC [@epa_2023], with the balance emanating from streams and rivers.
-Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^).
+Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
 
@@
 ## Sampling
-Samples were collected from approximately the middle of each waterbody for analysis of algal toxins (microcystins and cylindrospermopsin), benthic macroinvertebrates,  phytoplankton, atrazine, water chemistry, chlorophyll‐a, zooplankton, and dissolved N2O.
+Samples were collected from approximately the middle of each waterbody for analysis of algal toxins (microcystins and cylindrospermopsin), benthic macroinvertebrates,  phytoplankton, atrazine, water chemistry, chlorophyll‐a, zooplankton, and dissolved N~2~O.
 Duplicate dissolved N~2~O  samples were collected approximately 0.05 m below the water surface by drawing 105 mL of water into a 140 mL syringe equipped with a stopcock and containing 35 mL of air.
@@
 
-The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio; and (3) the proportion of under-saturated water bodies (_i.e._, saturation ratio < 1).
+The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N~2~O concentrations; (2) the N~2~O saturation ratio; and (3) the proportion of under-saturated water bodies (_i.e._, saturation ratio < 1).
 Saturation ratio was calculated as the ratio of dissolved to equilibrium N~2~O.
@@
 However, surface temperature and NO~3~ were not observed for lakes outside of the sample.
-Therefore, a more complex model was developed to include surface temperature and NO3 as additional responses conditioned on the survey design variables and the fully observed covariates.
+Therefore, a more complex model was developed to include surface temperature and NO~3~ as additional responses conditioned on the survey design variables and the fully observed covariates.
 This approach to inference was akin to Bayesian structural equation modeling approaches [@Merkle_etal_2021; @Merkle_Rosseel_2018].

### 2023-07-21 manuscript/manuscript_file.Rmd
Most recent updates to manuscript

@@
 
-- include a better description of full range of lake size: sample and population.  also add NO3 to better support emission rate statement.
 - incorporate in Methods section how/why size is incorporated as categorical variable and continuos variable.
-- maybe discuss upper midwest.  Lots of lake area, but largely at equilibrium, therefore very small fluxes.  Large area doesn't necessarily mean big flux.
--  last sentence of emission rates.  we should look at summary stats of observed NO3 rather than predicted 
 -  Mike will take a stab at big SI table.
--  Citation “epa_data” isn’t populating citation 26 in bibliography.  The bibTex entry is classified as a “website”.  Roy, can you attempt to make this citation play nice with nature_geosciences.csl?
+-  Citation “epa_data” isn’t populating citation 11 in bibliography.  The bibTex entry is classified as a “website”.  Roy, can you attempt to make this citation play nice with nature_geosciences.csl?
+
+## Psuedo abstract
+Lakes, ponds, and reservoirs are thought to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016;@Webb_etal_2019;@Kortelainen_etal_2020] have called into question the accuracy of emission estimates.
+Here we measured dissolved N~2~O concentrations and calculated emission rates across `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
+I get 984 if I exclude hand picked sites.
+NLA17 site info file shows 1210 sites sampled.
+I think the differences between these values are related to how hand picked and state intensification sites are handled.
+The sampled we used contains data from 1090 sites, so that is what we should report here.]** waterbodies in the continental U.S.
+(CONUS).
+Nitrous oxide undersaturation was widespread through the CONUS, with `r propN2oSinkMean` % of the waterbodies functioning as sinks for atmospheric N~2~O.
+Despite widespread N~2~O uptake by CONUS waterbodies, a small proportion of the waterbodies in each ecoregion had sufficiently high N~2~O emission rates to offset N~2~O uptake in the majority of waterbodies, such that all ecoregions functioned as net N~2~O sources to the atmosphere.
+Waterbodies with high nitrate (NO~3~) tended to be sources of N~2~O whereas those with low NO~3~ were generally sinks.
+Nitrate interacted with waterbody size, however, with large waterbodies having N~2~O concentrations closer to equilibrium than small waterbodies.
+We estimate that CONUS waterbodies emit `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` )metric tons N~2~O year^-1^, equivalent to only 5% of N~2~O emissions from all CONUS surface waters (e.g.
+zero order streams through river mouths) per IPCC estimates .
+The demonstrated capacity for widespread N~2~O uptake in CONUS waterbodies highlights the need to revisit global N~2~O models which presume surface waters are a N~2~O source [@epa_2023].
 
 ## Introduction
+Lakes, reservoirs, and ponds (waterbodies hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018], a potent greenhouse gas and contributor to stratospheric ozone destruction.
+It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across waterbodies in Canada [@Soued_etal_2016;@Webb_etal_2019] and Finland [@Kortelainen_etal_2020] have called this assumption into question.
+Furthermore, detailed investigations of individual waterbodies have shown that lentic N~2~O biogeochemistry is dynamic, with some waterbodies alternating between N~2~O sources and sinks in response to nitrogen availability, thermal stratification, and other factors [@Beaulieu_etal_2015;@Beaulieu_etal_2014;@Musenze_etal_2014].
+
+While recent data sets demonstrate that waterbodies can function as N~2~O sinks, these studies have been conducted at limited spatial scales and encompassed small ranges in potential environmental drivers (e.g.
+nitrogen loading).
+Most work on aquatic N~2~O biogeochemistry has focused on streams and rivers, but these systems have been consistently found to be N~2~O sources and the controls on N~2~O biogeochmistry likely differ between flowing and standing waters.
+These limitations in data availability have complicated efforts to incorporate N~2~O consumption into freshwater N~2~O budgets.
 
@@
 
-In this study we present the first national-scale survey of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S.
-(CONUS).
-A generalized random-tessellation survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
-compare my sample IDs to NLA IDs]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week **[(Fig.
+In this study we present the first national-scale survey of N~2~O concentrations and emissions from CONUS waterbodies.
+A generalized random-tessellation survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week **[(Fig.
 1)]**.
-Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?).
+Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate [@epa_data].
 `r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **[(Fig.
 1)]**.
-We modeled the data using a Bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population **[where did I get 224,916, Roy has 465,897.
-See technical support document]**.
+We modeled the data using a Bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 465,897 CONUS waterbodies in the target population.
 At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
@@
 Webb et al.
-[@Webb_etal_2019] found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Soued et al.
-[@Soued_etal_2016] report that 40% of rivers, lakes, and ponds sampled during the summer in a boreal region of Quebec, Canada were functioning as N~2~O sinks **[29% of lakes in Finland (kortelainen et al 2020) but kortelainen and Soued are mostly pristine lakes]**.
+[@Webb_etal_2019] found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months.
+An investigation of boreal systems in Canada found that  40% of rivers, lakes, and ponds sampled during the summer were functioning as N~2~O sinks[@Soued_etal_2016].
+Similarly, 29% of surveyed boreal lakes in Finland were found to be N~2~O sinks during the summer [@Kortelainen_etal_2020].
 While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
@@
 Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the affect of biogeochemistry on dissolved N~2~O is more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
+While previous surveys have documented relationships between lake size and dissolved methane concentration [@Pighini_etal_2018;@Juutinen_etal_2009;@Borges_etal_2011], to the best of our knowledge this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
 
@@
 
-The model indicates that NO~3~, water temperature, and surface area interact to determine the N~2~O saturation ratio.
-These interactions yield a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enable the identifcation of broad scale patterns.
+![Figure 5.
+Conditional effects plots illustrating the relationship between N~2~O saturation and A) nitrate (NO~3~ availability dependent upon lake size, and B) lake size dependent upon NO~3~ availability.
+The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13.](manuscript_figures/sat_cond_area.tiff)
+
+The model indicates that NO~3~ exerts a strong control on N~2~O, but the relationship is complicated by interactions with water temperature and lake surface area.
+This finding mirrors reports from a survey of farm ponds in Canada where N~2~O was found to be partially controlled by dissolved inorganic nitrogen (DIN) concentration, but the effect of DIN was conditional upon and thermal stratification strength.
+These complex interactions may explain contrasting literature reports regarding the role of NO~3~ in regulating N~2~O.
+Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020;@Yang_etal_2015] while others have found no relationship [@Soued_etal_2016;@Whitfield_etal_2011].
+It is possible that the NO~3~ ~ N~2~O relationship wasn't detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
+
+Interactions between NO~3~, water temperature, and lake size produced a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enable the identifcation of broad scale patterns.
 First, small waterbodies with low NO~3~ concentration tend to be undersaturated, whereas small waterbodies with high NO~3~ tend to be highly supersaturated.
@@
 
-![Figure 5.
-Conditional effects plots illustrating the relationship between N~2~O saturation and A) nitrate (NO~3~ availability dependent upon lake size, and B) lake size dependent upon NO~3~ availability.
-The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13.](manuscript_figures/sat_cond_area.tiff)
-
-**Not sure where I am going with this paragraph.
-Maybe scratch it.
-Mostly included as a placeholder for these references** While previous surveys have documented relationships between lake size and dissolved methane concentration (Pighini et al.
-2018, Juutinen et al 2009, Borges et al.
-2011), another greenhouse gas, to the best of our knowledge, this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
-Perhaps because previous surveys have included little variation in waterbody size [@Webb_etal_2019; @Kortelainen_etal_2020] or water quality [@Soued_etal_2016].
-
 ## Emission Rates
 
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was `r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **[(Table 1)]**.
+The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which is expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015;@Kortelainen_etal_2020;@Webb_etal_2019].
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was positive (`r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^)) **[(Table 1)]**.
 This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **[(SI Fig.
@@
 At the ecoregional scale, only the Northern Appalachians and Northern Plains ecoregions had mean emission rates less than 0.
-These ecoregions also had the lowest mean NO~3~ availability of all nine ecoregions **[need table or figure.
-See ecoreg_stats$post_med_no3 variable.
-maybe a SI table with summaries of all data from all_predictions?]**.
+These ecoregions also had the lowest mean NO~3~ availability of all nine ecoregions **[(Table S1)]**.
 
@@
 Total N~2~O flux from the surface of CONUS waterbodies is a function of the N~2~O emission rate and waterbody surface area.
-The largest lake size class (> 50 Ha) had the greatest N~2~O flux in all ecoregions, but flux estimates for this lake size class were also the most uncertain **[(Fig.
+The largest lake size class (> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
+This lake size class had the greatest N~2~O flux, but flux estimates for this lake size class were also the most uncertain **[(Fig.
 6)]**.
@@
 
+While total waterbody surface area contributed to differences in N~2~O flux among ecoregions, large cumulative surface area does not necessarily equate to large fluxes.
+The Upper Midwest ecoregion had the second largest waterbody surface area, but only the sixth largest flux.
+This is due to a mean saturation (mean = 1, 95% credible interview: 0.99 - 1.01) indicative of equilibirum with the atmosphere, which translates to very low fluxes.
+
 At the national scale, we estimate that CONUS waterbodies emit `r table1_dat %>% janitor::clean_names() %>% filter(spatial_domain == "national") %>% select(n2o_flux) %>% pull()` metric tons N~2~O year^-1^.
@@
 Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
+
+**[extract n2o data from Lauerwald for CONUS and compare]**
 
@@
 Although waterbodies in the largest lake size class tend to have N~2~O levels close to equilibrium, they emit more N~2~O than the other size classes due to their expansive surface area.
-The N~2~O flux from these lakes is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved N~2~O translate to large uncertainties in flux when propagated across the surface of large waterbodies.
-Efforts to improve the accuracy of the lentic N~2~O flux estimate for the U.S.
-should therefore focus on investigations of the nations largest lakes and reservoirs.
+The N~2~O flux from these waterbodies is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved N~2~O translate to large uncertainties in flux when propagated across the surface of large waterbodies.
+While our sample included some of the largest waterbodies in the CONUS **[(Table S2)]**, efforts to improve the accuracy of the lentic N~2~O flux estimate should include measurements at the nations largest lakes and reservoirs.
 
 The smallest lake size classes exhibited both the highest and lowest levels of N~2~O saturation, depending on NO~3~ availability.
-It is unclear whether this variability is related to differences in NO~3~ loading or temporal variation.
+It is unclear whether this variability is related to differences in NO~3~ loading or is an artifact of temporal variation in NO~3~ processing and N~2~O production/consumption.
 For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted [@Beaulieu_etal_2014].
-A better understanding of the temporal evolution of N~2~O production and consumption in small waterbodies is an important research need to determine the extent to which these systems can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
+Most small waterbodies in the CONUS function as N~2~O sinks during the summer, but a small subset are N~2~O emission hotspots.
+A better understanding of the temporal evolution of N~2~O production and consumption is an important research need to determine the extent to which small waterbodies can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
 

### 2023-07-21 manuscript/manuscript_support.R

@@ -515,6 +515,40 @@ propN2oSinkUpperNPL <- all_predictions %>%
   filter(WSA9 == "NPL") %>% 
   pull()
 
+
+## Lake size distribution----
+# proportion of total SA attributable to lakes <=50 ha
+proportionSmallSA <- all_predictions %>%
+  filter(.draw == 1) %>% # just grab one realization
+  select(area_ha) %>%
+  arrange(area_ha) %>% 
+  mutate(cumulative = cumsum(area_ha), # cumulative surface area from small to large
+         distribution = cumulative / sum(area_ha), # prop of total SA
+         delta50 = abs(50-area_ha)) %>% # how close to 50 Ha
+  filter(delta50 == min(delta50)) %>% # select closest to 50 Ha
+  pull(distribution) # 50Ha lake
+  
+# proportion of population <=50ha
+proportionSmallNumber <- all_predictions %>%
+  filter(.draw == 1) %>% # just grab one realization
+  select(area_ha) %>%
+  arrange(area_ha) %>% 
+  mutate(number = 1:nrow(.), # number each lake from smallest to largest
+         proportion = number / nrow(.)) %>% # proportion of all lakes 
+  filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% # grab 50Ha lake
+  pull(proportion) # extract proportion of all lakes represented by this lake
+  
+
+# Mean and max emission rates-----
+min.max <- all_predictions %>%
+  group_by(.row) %>%
+  summarize(e.n2o.mg.m2.d = mean(e.n2o.mg.m2.d)) %>%
+  ungroup() %>%
+  summarize(min = min(e.n2o.mg.m2.d),
+            max = max(e.n2o.mg.m2.d))
+
+
+
 ## IPCC Indirect N2O Leaching and Runoff----
 # 1990 - 2021 inventory reports 15.2 + 2.9 MMT CO2-Eq per year for CONUS (Table 5-19) in 2021.
 # https://www.epa.gov/system/files/documents/2023-04/US-GHG-Inventory-2023-Chapter-5-Agriculture.pdf

### 2023-08-24 manuscript/manuscript_file.Rmd
manuscript updates after expaning to 2000 realizations

@@
 -   Mike will take a stab at big SI table.
--  Citation “epa_data” isn’t populating citation 11 in bibliography.  The bibTex entry is classified as a “website”.  Roy, can you attempt to make this citation play nice with nature_geosciences.csl?
+-   Citation "epa_data" isn't populating citation 11 in bibliography. The bibTex entry is classified as a "website". Roy, can you attempt to make this citation play nice with nature_geosciences.csl?
+-   netCDF data from Lauerwald
+-   Letters vs article format at Nature Geosciences
+-   Add quantitative NO3 values to Fig.3B and Fig. 5B legend.
+-   Add quantitative size category values to Fig. 5A legend.
 
 ## Psuedo abstract
-Lakes, ponds, and reservoirs are thought to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016;@Webb_etal_2019;@Kortelainen_etal_2020] have called into question the accuracy of emission estimates.
-Here we measured dissolved N~2~O concentrations and calculated emission rates across `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?
-I get 984 if I exclude hand picked sites.
-NLA17 site info file shows 1210 sites sampled.
-I think the differences between these values are related to how hand picked and state intensification sites are handled.
-The sampled we used contains data from 1090 sites, so that is what we should report here.]** waterbodies in the continental U.S.
+
+Lakes, ponds, and reservoirs are thought to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020] have called into question the accuracy of emission estimates[@Lauerwald_etal_2019].
+Here we measured dissolved N~2~O concentrations and calculated emission rates across `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies in the continental U.S.
 (CONUS).
@@
 Nitrate interacted with waterbody size, however, with large waterbodies having N~2~O concentrations closer to equilibrium than small waterbodies.
-We estimate that CONUS waterbodies emit `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` )metric tons N~2~O year^-1^, equivalent to only 5% of N~2~O emissions from all CONUS surface waters (e.g.
+We estimate that CONUS waterbodies emit `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` ) metric tons N~2~O year^-1^, equivalent to only `r (national_stats %>% select(UCL_fn2o) %>% pull() / 22767) * 100`% of N~2~O emissions from all CONUS surface waters (e.g.
 zero order streams through river mouths) per IPCC estimates .
@@
 ## Introduction
-Lakes, reservoirs, and ponds (waterbodies hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018], a potent greenhouse gas and contributor to stratospheric ozone destruction.
-It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across waterbodies in Canada [@Soued_etal_2016;@Webb_etal_2019] and Finland [@Kortelainen_etal_2020] have called this assumption into question.
-Furthermore, detailed investigations of individual waterbodies have shown that lentic N~2~O biogeochemistry is dynamic, with some waterbodies alternating between N~2~O sources and sinks in response to nitrogen availability, thermal stratification, and other factors [@Beaulieu_etal_2015;@Beaulieu_etal_2014;@Musenze_etal_2014].
 
-While recent data sets demonstrate that waterbodies can function as N~2~O sinks, these studies have been conducted at limited spatial scales and encompassed small ranges in potential environmental drivers (e.g.
+Lakes, reservoirs, and ponds (waterbodies hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], a potent greenhouse gas and contributor to stratospheric ozone destruction.
+It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across waterbodies in Canada [@Soued_etal_2016; @Webb_etal_2019] and Finland [@Kortelainen_etal_2020] have called this assumption into question.
+Furthermore, detailed investigations of individual waterbodies have shown that lentic N~2~O biogeochemistry is dynamic, with some waterbodies alternating between N~2~O sources and sinks in response to nitrogen availability, thermal stratification, and other factors [@Beaulieu_etal_2015; @Beaulieu_etal_2014; @Musenze_etal_2014].
+
+While several measurement campaigns have demonstrated that waterbodies can function as N~2~O sinks[@Soued_etal_2016; @Webb_etal_201; @Kortelainen_etal_2020; @Beaulieu_etal_2015; @Beaulieu_etal_2014; @Musenze_etal_2014], these studies have been conducted at limited spatial scales and encompassed small ranges in potential environmental drivers (e.g.
 nitrogen loading).
-Most work on aquatic N~2~O biogeochemistry has focused on streams and rivers, but these systems have been consistently found to be N~2~O sources and the controls on N~2~O biogeochmistry likely differ between flowing and standing waters.
+Most work on aquatic N~2~O biogeochemistry has focused on streams and rivers, but these systems have been consistently found to be N~2~O sources [@Beaulieu_etal_2008; @Beaulieu_etal_2010; @Quick_etal_2019] and the controls on N~2~O biogeochmistry likely differ between flowing and standing waters.
 These limitations in data availability have complicated efforts to incorporate N~2~O consumption into freshwater N~2~O budgets.
@@
 In this study we present the first national-scale survey of N~2~O concentrations and emissions from CONUS waterbodies.
-A generalized random-tessellation survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week **[(Fig.
+A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[Roy has 1079.
+Handpicked and NA sites not used.
+Update this code.]**sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week **[(Fig.
 1)]**.
@@
 1)]**.
-We modeled the data using a Bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 465,897 CONUS waterbodies in the target population.
+We modeled the data using a Bayesian hierarchical approach and predicted the dissolved N~2~O concentration for all 465,897 CONUS waterbodies in the target population.
 At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`), the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), and \>65% of the waterbodies in every ecoregion were undersaturated **[(Table 1)]**.
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`) **[(Table 1)]**.
 
@@
     caption = as_paragraph( # allows for formatting
-      as_chunk("Table 1. Mean and 95% confidence interval of N2O undersaturation extent, saturation ratio, emission rate, and flux at the national and ecoregional scales.", 
+      as_chunk("Table 1. Mean and 95% credible interval of N2O undersaturation extent, saturation ratio, emission rate, and flux at the national and ecoregional scales.", 
                props = fp_text_default(bold = TRUE) # here is where we specify formatting (font, size, bold, etc)
@@
 Across the CONUS, and within each of the nine major ecoregions, the distribution of N~2~O saturation ratios had median values that were less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
-Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources.
+Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources, particularly at high NO~3~ at water temperature **[Fig.
+3]**.
 
@@
 Distribution of N~2~O saturation values for the population of waterbodies in the nine major ecoregions.
-Each blue line (n=500) represents the distribution of values from a single realization of the posterior prediction.
-Vertical black dashed and solid lines indicated median and mean, respectively.
+Each blue line (n=2000) represents the distribution of values from a single realization of the posterior prediction.
+Vertical black dashed and solid lines indicate median and mean, respectively.
 Vertical red line indicates a saturation ratio of 1.
@@
 Conditional effects plot illustrating A) the relationship between N~2~O saturation and nitrate (NO~3~) concentration, dependent upon surface water temperature.
-The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13, and B) relationship between N~2~O saturation and surface water temperature, dependent upon NO~3~ category.
+The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13, and B) relationship between N~2~O saturation and surface water temperature, dependent upon NO~3~ category.
 Horizontal dashed lines in both panels reflects an N~2~O saturation value of 1.](manuscript_figures/sat_cond_surftemp.tiff)
@@
 At low NO~3~ concentrations, temperature had a relatively minor affect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
-The effect of temperature became more pronounced when NO~3~ availability was greater, possibly reflecting the stimulatory affect of higher temperature on biological N~2~O production.
+The effect of temperature became more pronounced when NO~3~ availability was greater, likely reflecting the stimulatory affect of higher temperature on biological N~2~O production.
 The relative effect of NO~3~ was greater than that of temperature, thus elevated water temperature serves to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
@@
 Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the affect of biogeochemistry on dissolved N~2~O is more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
-While previous surveys have documented relationships between lake size and dissolved methane concentration [@Pighini_etal_2018;@Juutinen_etal_2009;@Borges_etal_2011], to the best of our knowledge this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
+While previous surveys have documented relationships between lake size and dissolved methane concentration [@Pighini_etal_2018; @Juutinen_etal_2009; @Borges_etal_2011], to the best of our knowledge this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
 
@@
 The horizontal red arrows highlight $\Delta$ N~2~O for the smallest lake size class.
-Mean (points) and 95% credible interview (horizontal bars) are derived from 500 realizations of the poterior predictive distribution.
-](manuscript_figures/deltaN2ObySize.tiff)
+Mean (points) and 95% credible interview (horizontal bars) are derived from 2000 realizations of the poterior predictive distribution.](manuscript_figures/deltaN2ObySize.tiff)
 
@@
 Conditional effects plots illustrating the relationship between N~2~O saturation and A) nitrate (NO~3~ availability dependent upon lake size, and B) lake size dependent upon NO~3~ availability.
-The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and < 4.09, 3) $\geq$ 4.09 and < 30.20, 4) $\geq$ 30.20 and < 223.13, and 5) $\geq$ 223.13.](manuscript_figures/sat_cond_area.tiff)
+The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13.](manuscript_figures/sat_cond_area.tiff)
 
@@
 These complex interactions may explain contrasting literature reports regarding the role of NO~3~ in regulating N~2~O.
-Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020;@Yang_etal_2015] while others have found no relationship [@Soued_etal_2016;@Whitfield_etal_2011].
-It is possible that the NO~3~ ~ N~2~O relationship wasn't detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
+Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020; @Yang_etal_2015] while others have found no relationship [@Soued_etal_2016; @Whitfield_etal_2011].
+It is possible that the NO~3~ \~ N~2~O relationship wasn't detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
 
 Interactions between NO~3~, water temperature, and lake size produced a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enable the identifcation of broad scale patterns.
-First, small waterbodies with low NO~3~ concentration tend to be undersaturated, whereas small waterbodies with high NO~3~ tend to be highly supersaturated.
+First, small waterbodies with low NO~3~ concentration tend to be undersaturated, whereas small waterbodies with high NO~3~ tend to be highly supersaturated **[Fig.
+5]**.
 Large waterbodies, regardless of NO~3~ status, tend to have N~2~O saturation ratios closer to equilibrium.
@@
 
-The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which is expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015;@Kortelainen_etal_2020;@Webb_etal_2019].
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was positive (`r national_stats$post_med_en2o` (95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^)) **[(Table 1)]**.
+The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which is expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **[(Table 1)]**.
 This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **[(SI Fig.
 1)]**.
-At the ecoregional scale, only the Northern Appalachians and Northern Plains ecoregions had mean emission rates less than 0.
-These ecoregions also had the lowest mean NO~3~ availability of all nine ecoregions **[(Table S1)]**.
+At the ecoregional scale, the Northern Appalachians, Northern Plains, and Upper Midwest ecoregions had mean emission rates less than 0.
+These ecoregions also had the lowest mean NO~3~ availability of all nine ecoregions **[(Table S1)--is this table updated yet?]**.
 
 ## Fluxes
+
 Total N~2~O flux from the surface of CONUS waterbodies is a function of the N~2~O emission rate and waterbody surface area.
-The largest lake size class (> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
+The largest lake size class (\> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
 This lake size class had the greatest N~2~O flux, but flux estimates for this lake size class were also the most uncertain **[(Fig.
@@
 Small uncertainties in emission rates translate to large uncertainties in flux when expressed across the surface area of large waterbodies.
-Despite widespread N~2~O undersaturation, all ecoregions were a net source of N~2~O to the atmosphere **[(Table 1)]**.
+Despite widespread N~2~O undersaturation, all but two ecoregions were a net source of N~2~O to the atmosphere **[(Table 1)]**.
 Differences in N~2~O flux across ecoregions are explained by patterns in N~2~O saturation and waterbody surface **[(Table 1)]**.
@@
 
-**[extract n2o data from Lauerwald for CONUS and compare]**
-
 Our annual flux estimates assume that emission rate estimates based on samples collected during the summer reflect annual conditions.
@@
 ## Conclusion
+
 Our survey is the first to demonstrate that most lakes and reservoirs in the CONUS function as N~2~O sinks during the summer.
@@
 The N~2~O flux from these waterbodies is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved N~2~O translate to large uncertainties in flux when propagated across the surface of large waterbodies.
-While our sample included some of the largest waterbodies in the CONUS **[(Table S2)]**, efforts to improve the accuracy of the lentic N~2~O flux estimate should include measurements at the nations largest lakes and reservoirs.
+While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include measurements at the nations largest lakes and reservoirs.
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 ## Sampling
+
 Samples were collected from approximately the middle of each waterbody for analysis of algal toxins (microcystins and cylindrospermopsin), benthic macroinvertebrates, phytoplankton, atrazine, water chemistry, chlorophyll‐a, zooplankton, and dissolved N~2~O.
@@
 The syringes were gently shaken for five minutes and the headspace gas transferred to a 30 mL syringe.
-Finally, the gas sample was transferred into a 12 mL pre-evacuated (<50 mTorr) glass vial (Labco, Wales, UK) sealed with a PTFE/Silicone stacked on top of a grey chlorobutyl septa.
-The temperatue of the water in the syringe was recorded immediately after the gas was transferred into the vial.
+Finally, the gas sample was transferred into a 12 mL pre-evacuated (\<50 mTorr) glass vial (Labco, Wales, UK) sealed with a PTFE/Silicone stacked on top of a grey chlorobutyl septa.
+The temperature of the water in the syringe was recorded immediately after the gas was transferred into the vial.
 Duplicate 20 mL air samples were collected from approximately 2 m above the water surface and transferred to 12 mL glass vials.
-A vertical profile of water temperatue was measured at the same site.
+A vertical profile of water temperature was measured at the same site.
 
@@
 MRP is a model-based approach carried out in two steps.
-In the first step, the sampled responses of interest (_e.g._, N~2~O) are conditioned on the survey design variables (_e.g._, ecoregion, state, size) in a multilevel regression model.
+In the first step, the sampled responses of interest (*e.g.*, N~2~O) are conditioned on the survey design variables (*e.g.*, ecoregion, state, size) in a multilevel regression model.
 Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the population of interest [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
-In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (_e.g.,_ US waterbodies) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
-In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (_i.e._, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (_e.g._, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
+In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (*e.g.,* US waterbodies) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (*e.g.*, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
 The fully Bayesian implementation made propagation of uncertainty from the waterbody-level predictions to higher level summaries straightforward.
 
-The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N~2~O concentrations; (2) the N~2~O saturation ratio; and (3) the proportion of under-saturated water bodies (_i.e._, saturation ratio < 1).
+The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N~2~O concentrations; (2) the N~2~O saturation ratio; and (3) the proportion of under-saturated water bodies (*i.e.*, saturation ratio \< 1).
 Saturation ratio was calculated as the ratio of dissolved to equilibrium N~2~O.
-Because dissolved and equilibrium N~2~O were each measured on the same sample units (_i.e._, a single waterbody), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, common dependencies on geography.
+Because dissolved and equilibrium N~2~O were each measured on the same sample units (*i.e.*, a single waterbody), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, common dependencies on geography.
 The modeled response variable was, therefore, multivariate.
 Although the mean marginal probabilities estimated from separate models may have provided comparable estimates, the joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
-The saturation ratio was not modeled or predicted directly, but was estimated _post hoc_ as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
+The saturation ratio was not modeled or predicted directly, but was estimated *post hoc* as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
 Similarly, the proportion of under-saturated waterbodies was estimated from this PPD as the proportion of waterbodies with a saturation ratio less than 1.
@@
 
-Variables in bold text above were treated as partially observed with respect to the POI (_i.e._, observed only in the sample), whereas variables not in bold were fully observed.
+Variables in bold text above were treated as partially observed with respect to the POI (*i.e.*, observed only in the sample), whereas variables not in bold were fully observed.
 The partially observed variables, including dissolved and equilibrium N~2~O, NO~3~, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.

### 2023-08-29 manuscript/manuscript_file.Rmd
New figure 6 in manuscript

@@
 localPath <- Sys.getenv("USERPROFILE")
-knitr::opts_knit$set(root.dir = "../") # set wd to root of RStudio project
+#knitr::opts_knit$set(root.dir = "../") # set wd to root of RStudio project
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set wd to root of RStudio project
 
@@
 source("manuscript/manuscript_support.R", local = knitr::knit_global())
+
 ```
@@
 
-While several measurement campaigns have demonstrated that waterbodies can function as N~2~O sinks[@Soued_etal_2016; @Webb_etal_201; @Kortelainen_etal_2020; @Beaulieu_etal_2015; @Beaulieu_etal_2014; @Musenze_etal_2014], these studies have been conducted at limited spatial scales and encompassed small ranges in potential environmental drivers (e.g.
+While several measurement campaigns have demonstrated that waterbodies can function as N~2~O sinks[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020; @Beaulieu_etal_2015; @Beaulieu_etal_2014; @Musenze_etal_2014], these studies have been conducted at limited spatial scales and encompassed small ranges in potential environmental drivers (e.g.
 nitrogen loading).
@@
 In this study we present the first national-scale survey of N~2~O concentrations and emissions from CONUS waterbodies.
-A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[Roy has 1079.
-Handpicked and NA sites not used.
-Update this code.]**sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week **[(Fig.
+A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week[@EPAdesign] **[(Fig.
 1)]**.
@@
 
-## Emission Rates
+## Emission Rates[a]
 
@@
 1)]**.
-At the ecoregional scale, the Northern Appalachians, Northern Plains, and Upper Midwest ecoregions had mean emission rates less than 0.
-These ecoregions also had the lowest mean NO~3~ availability of all nine ecoregions **[(Table S1)--is this table updated yet?]**.
 
@@
 Finally, the predictive distributions for temperature and NO~3~ were used in predicting both N~2~O responses in the POI.
+
+[a]note overlap between all ecoregion.
+all ecoregions overlap 0
+Note relationship with lake size.

### 2023-08-29 manuscript/manuscript_support.R

@@ -9,6 +9,8 @@ library(tidyverse) # dplyr, ggplot
 library(janitor) # clean names
 library(USAboundaries) # state boundaries
 library(tictoc) # processing time
+library(ggpubr) # multiple plots
+library(ggallin) # psuedlolog transformation for negative values
 
 ## Identify local path for each user
 localPath <- Sys.getenv("USERPROFILE")
@@ -16,10 +18,14 @@ localPath <- Sys.getenv("USERPROFILE")
 
 # DATA---------------
 ## load sample data (dg object)----
+if(localPath == "") { # if DMAP, then
+  load("inputData/dg.2021-02-01.RData")
+} else { # if not DMAP, then...
 load(paste0(localPath,  # object name dg
             "/Environmental Protection Agency (EPA)/",
             "ORD NLA17 Dissolved Gas - Documents/",
             "inputData/dg.2021-02-01.RData"))
+}
 
 # To enable spatial analysis of the data, the dataframe will be converted to a 
 # 'simple features' (sf) object.
@@ -33,12 +39,17 @@ dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"),
 
 ## load ecoregions----
 # read in ecoregion polygons
+if(localPath == ""){ # if DMAP, then
+  ecoR <- st_read(dsn = "inputData",
+                  layer = "aggr_ecoregions_simple",
+                  quiet = TRUE)
+  } else { # if not DMAP, then...
 ecoR <- st_read(dsn = paste0(localPath, 
                              "/Environmental Protection Agency (EPA)/",
                              "ORD NLA17 Dissolved Gas - Documents/inputData"),
                 layer = "aggr_ecoregions_simple",
                 quiet = TRUE)
-
+}
 # Check CRS
 st_crs(ecoR) # 3857
 ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
@@ -66,11 +77,23 @@ states <- USAboundaries::us_states() %>%
 
 
 ## load population data----
-#tic()
+#tic() # 7.6min on DMAP
+if(localPath == ""){ # if DMAP, then
+  load("../../shared/jbeaulie/all_predictions.rda")
+} else { # if not DMAP
 load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
[…]

### 2023-08-30 manuscript/manuscript_file.Rmd
Revised Figure 6 and text

@@
 
-## Emission Rates[a]
+## Emission Rates
 
 The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which is expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
+Variability and magnitude of predicted N~2~O emission rates decreased with increasing waterbody size **[(Fig.
+6A)]**, which is consistent with the relationship between N~2~O saturation and waterbody size **[(Fig.
+5A,B)]**.
 Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **[(Table 1)]**.
@@
 
+![Figure 6: A) N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 waterbodies in the target population.
+C) N~2~O flux plotted against the five lake size categories used in the survey design.
+Across all plots, points and bars reflect the mean (points) and 95% credible interval of the 2000 realizations of the posterior predictive distribution.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
+
 ## Fluxes
 
-Total N~2~O flux from the surface of CONUS waterbodies is a function of the N~2~O emission rate and waterbody surface area.
-The largest lake size class (\> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
-This lake size class had the greatest N~2~O flux, but flux estimates for this lake size class were also the most uncertain **[(Fig.
-6)]**.
-Small uncertainties in emission rates translate to large uncertainties in flux when expressed across the surface area of large waterbodies.
-Despite widespread N~2~O undersaturation, all but two ecoregions were a net source of N~2~O to the atmosphere **[(Table 1)]**.
-Differences in N~2~O flux across ecoregions are explained by patterns in N~2~O saturation and waterbody surface **[(Table 1)]**.
-The Coastal Plains ecoregion had the greatest flux, driven by the high levels of N~2~O saturation and extensive waterbody surface area.
-By contrast, the Northern Plains had the lowest N~2~O saturation, waterbody surface area, and N~2~O flux.
+Total N~2~O flux from the surface of CONUS waterbodies is a function of N~2~O emission rates and waterbody surface areas.
+The four smallest size classes (\< 50 Ha) in the survey design contained `r round(proportionSmallNumber, 2) * 100`% of all lakes in the population, but contributed only `r round((FluxBySize %>% filter(size_cat != "50_max") %>% select(estimate) %>% summarise(tot_small = sum(estimate)) %>% pull())/(FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) * 100, 0)`% to the N~2~O flux from CONUS waterbodies **[(Fig.
+6C)]**.
+The reason for this pattern is two-fold.
+First, lakes in the smaller size classes constituted only `r round(proportionSmallSA, 2) * 100`% of total waterbody surface area, thereby limiting their N~2~O flux.
+Secondly, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the population **[(Fig.
+6A)].** Waterbodies in the largest size class (\> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
+This lake size class had the greatest N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year^{-1}), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year^{-1} **[(Fig.
+6C)]**.
+While uncertainty in N~2~O emission rates was smallest for large lakes **(Fig.
+6A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large waterbodies.
 
-While total waterbody surface area contributed to differences in N~2~O flux among ecoregions, large cumulative surface area does not necessarily equate to large fluxes.
-The Upper Midwest ecoregion had the second largest waterbody surface area, but only the sixth largest flux.
-This is due to a mean saturation (mean = 1, 95% credible interview: 0.99 - 1.01) indicative of equilibirum with the atmosphere, which translates to very low fluxes.
+Despite widespread N~2~O undersaturation, only two ecoregions had flux estimates less than zero (Xeric and Coastal Plains), but the credible intervals for those ecoregions were broad and include zero **(Table 1)**.
+Only the Temperate Plains, Southern Plains, and Southern Appalachians had flux estimate credible intervals that did not overlap with zero and these ecoregions functioned as net N~2~O sources.
+Overall, flux estimates overlapped across all nine ecoregions, which is somewhat surprising given the large differences in environmental conditions across the ecoregions.
 
-At the national scale, we estimate that CONUS waterbodies emit `r table1_dat %>% janitor::clean_names() %>% filter(spatial_domain == "national") %>% select(n2o_flux) %>% pull()` metric tons N~2~O year^-1^.
+At the national scale, we estimate that CONUS waterbodies are a source of N~2~O, emitting `r table1_dat %>% janitor::clean_names() %>% filter(spatial_domain == "national") %>% select(n2o_flux) %>% pull()` metric tons N~2~O year^-1^; however, this estimate is not well constrained and the 95% credible interval includes zero, thus we cannot conclude with statistical confidence whether CONUS waterbodies are a net source of sink of N~2~O.
 While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
@@
 Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
+Whether CONUS waterbodies are a net source or sink of N~2~O, their contribution to the US N~2~O budget is small.
 
@@
 While our assumption of constant emissions across the year adds uncertainty to the annual flux estimate, it is not clear that the estimate is biased high or low.
-
-![Figure 6: N~2~O flux by ecoregion and lake size category.
-Mean (points) and 95% credible interval (horizontal bars) reflect the distribution of 500 realizations of the posterior predictive distribution.](manuscript_figures/n2oFluxBySizeEco.tiff)
 
@@
 The N~2~O flux from these waterbodies is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved N~2~O translate to large uncertainties in flux when propagated across the surface of large waterbodies.
-While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include measurements at the nations largest lakes and reservoirs.
+While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nations largest lakes including an assesment of intra-waterbody spatial variation in N~2~O, which can be important [@Salk_etal_2019, @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propogate to large uncertainties in flux estimates.
 
@@
 
-[a]note overlap between all ecoregion.
-all ecoregions overlap 0
-Note relationship with lake size.
+[a]note overlap between all ecoregion. all ecoregions overlap 0 Note relationship with lake size.

### 2023-08-30 manuscript/manuscript_support.R

@@ -494,10 +494,26 @@ b3 <- all_predictions %>%
 #                   ggarrange(b3, ncol = 2, nrow=1, labels = "AUTO"), # single plot in 3rd row, only one column
 #                   nrow=3, labels = "AUTO")
 
-ggpubr::ggarrange(b1, b2, b3, ncol = 3, nrow=1, labels = "AUTO")
+ggpubr::ggarrange(b2, b1, b3, ncol = 3, nrow=1, labels = "AUTO")
 ggsave("manuscript/manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff", width = 8, height = 4)   
 
 }
+
+all_predictions %>%
+  group_by(WSA9_NAME, .draw) %>% # group by iteration
+  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 2000 means for size cat
+  # now summarize to 1 statistic per WSA9
+  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
+             LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
+             UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
+  ggplot(., aes(x = estimate, y = WSA9_NAME)) +
+  geom_point() +
+  geom_linerange(aes(xmin = LCL, xmax = UCL)) +
+  geom_vline(xintercept = 0, color='blue') +
+  xlab(expression(N[2]*O~flux~(metric~tons~year^{-1}))) + # 1Mg = 1 metric ton
+  ylab("Lake size class (ha)") +
+  theme_bw() 
+
 
 
 ## SI Figure 1: N2O emission rate distribution----
@@ -624,7 +640,16 @@ proportionSmallNumber <- all_predictions %>%
   filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% # grab 50Ha lake
   pull(proportion) # extract proportion of all lakes represented by this lake
   
-
+# flux from largest lake size class------
+FluxBySize <- all_predictions %>%
+  group_by(size_cat, .draw) %>% # group by iteration
+  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 2000 means for size cat
+  # now summarize to 1 statistic per size_cat
+  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
[…]

### 2023-08-31 manuscript/manuscript_file.Rmd
Minor tweaks to manuscript files

@@
 -   incorporate in Methods section how/why size is incorporated as categorical variable and continuos variable.
--   Mike will take a stab at big SI table.
 -   Citation "epa_data" isn't populating citation 11 in bibliography. The bibTex entry is classified as a "website". Roy, can you attempt to make this citation play nice with nature_geosciences.csl?
--   netCDF data from Lauerwald
 -   Letters vs article format at Nature Geosciences
 -   Add quantitative NO3 values to Fig.3B and Fig. 5B legend.
--   Add quantitative size category values to Fig. 5A legend.
+-   nitrate in methods
+-   Lester Luan and Jana Compton for internal tech reviews.
 
@@
 6A)].** Waterbodies in the largest size class (\> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
-This lake size class had the greatest N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year^{-1}), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year^{-1} **[(Fig.
+This lake size class had the greatest N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year\^{-1}), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year\^{-1} **[(Fig.
 6C)]**.
@@
 Finally, the predictive distributions for temperature and NO~3~ were used in predicting both N~2~O responses in the POI.
-
-[a]note overlap between all ecoregion. all ecoregions overlap 0 Note relationship with lake size.

### 2023-08-31 manuscript/manuscript_support.R

@@ -1,6 +1,7 @@
 ## This script generates data to be referenced in inline-code in the manuscript.
-## This makes the inline code more succint and the .Rmd easier to read.
+## This makes the inline code more succinct and the .Rmd easier to read.
 
+# 32 minutes to run on DMAP with images already stored locally
 
 # SETUP----------------------
 ## load libraries

### 2023-09-12 manuscript/manuscript_file.Rmd
minor tweaks to MS

@@
            ref_symbols = "a",
-           value = as_paragraph("Mean and 95% credible interval of 500 realizations of posterior predictive distribution"),
+           value = as_paragraph("Mean and 95% credible interval of 2000 realizations of posterior predictive distribution"),
            part = "header"

### 2023-09-12 manuscript/manuscript_support.R

@@ -4,7 +4,19 @@
 # 32 minutes to run on DMAP with images already stored locally
 
 # SETUP----------------------
+## Identify local path for each user
+localPath <- Sys.getenv("USERPROFILE")
+
 ## load libraries
+# DMAP not under renv control.  Should probably specify same
+# versions as used for renv library
+if(localPath == "") { # if DMAP, then
+install.packages(c("sf", "tidyverse", "janitor",
+                   "tictoc", "ggpubr", "ggallin",
+                   "devtools"))
+devtools::install_version("USAboundaries", "0.4.0")
+}
+
 library(sf) # spatial data
 library(tidyverse) # dplyr, ggplot
 library(janitor) # clean names
@@ -13,8 +25,7 @@ library(tictoc) # processing time
 library(ggpubr) # multiple plots
 library(ggallin) # psuedlolog transformation for negative values
 
-## Identify local path for each user
-localPath <- Sys.getenv("USERPROFILE")
+
 
 
 # DATA---------------
@@ -295,7 +306,7 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   # using xlim to zoom in on values close to 1
   
   dummy <- all_predictions %>%
-    group_by(WSA9) %>%
+    group_by(WSA9, .draw) %>%  # test this
     summarize(mean = mean(n2osat),
               median = median(n2osat)) %>%
     mutate(dens.med = c(2, 2.5, 2.3, 2.05, 1.75, 2.15, 2.2, 1.9, 1.9),

### 2023-09-15 manuscript/manuscript_support.R
Minor updates

@@ -27,7 +27,6 @@ library(ggallin) # psuedlolog transformation for negative values
 
 
 
-
 # DATA---------------
 ## load sample data (dg object)----
 if(localPath == "") { # if DMAP, then
@@ -197,6 +196,7 @@ ecoreg_stats <- all_predictions %>%
   summarise( mean_n2o = mean(n2o),
              mean_n2oeq = mean(n2oeq),
              mean_sat = mean(n2osat),
+             median_sat = median(n2osat),
              prop_sat = sum(n2osat < 1) / length(unique(.row)),
              mean_en2o = mean(e.n2o.mg.m2.d),
              median_en2o = median(e.n2o.mg.m2.d),
@@ -211,6 +211,7 @@ ecoreg_stats <- all_predictions %>%
              LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
              UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
              post_med_sat = round(median(mean_sat), 2),
+             post_med_med_sat = round(median(median_sat), 2),
              LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
              UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
              estimate = round(median(prop_sat), 3),
@@ -305,37 +306,33 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   # density plots of predicted distribution, mean, and median
   # using xlim to zoom in on values close to 1
   
-  dummy <- all_predictions %>%
-    group_by(WSA9, .draw) %>%  # test this
-    summarize(mean = mean(n2osat),
-              median = median(n2osat)) %>%
-    mutate(dens.med = c(2, 2.5, 2.3, 2.05, 1.75, 2.15, 2.2, 1.9, 1.9),
-           dens.mean = c(.4, 1.9, 2, .6, .65, .45, 1.5, 1.45, 0.9),
+  dummy <- ecoreg_stats %>%
+                        #CPL   NAP   NPL    SAP    SPL   TPL  UWM    WMT    XER
+    mutate(dens.med = c(2.5,  4,    3.1,   3.5,   2.5,  3,   3,      2,     2.5),
+           dens.mean = c(.5,  3,    2.5,   1.5,   1,    1,   2,      1.5,   1.1),
            .draw = 1) # needed to match grouping aesthetic in ggplot call
   
[…]

### 2023-10-16 manuscript/manuscript_file.Rmd
Minor manuscript updates

@@
 
-Lakes, ponds, and reservoirs are thought to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020] have called into question the accuracy of emission estimates[@Lauerwald_etal_2019].
-Here we measured dissolved N~2~O concentrations and calculated emission rates across `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies in the continental U.S.
+Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020] have called the accuracy of emission estimates into question [@Lauerwald_etal_2019].
+Here we measured dissolved N~2~O concentrations and calculated emission rates for `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies distributed across the nine major ecoregions of the conterminous U.S.
 (CONUS).
-Nitrous oxide undersaturation was widespread through the CONUS, with `r propN2oSinkMean` % of the waterbodies functioning as sinks for atmospheric N~2~O.
-Despite widespread N~2~O uptake by CONUS waterbodies, a small proportion of the waterbodies in each ecoregion had sufficiently high N~2~O emission rates to offset N~2~O uptake in the majority of waterbodies, such that all ecoregions functioned as net N~2~O sources to the atmosphere.
-Waterbodies with high nitrate (NO~3~) tended to be sources of N~2~O whereas those with low NO~3~ were generally sinks.
+We use a Bayesian hierarchical approach to model the data and predicted N~2~O in`r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull()` CONUS waterbodies.
+We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean` % of the waterbodies functioning as sinks for atmospheric N~2~O.
+Despite widespread N~2~O uptake by CONUS waterbodies, a small proportion of the waterbodies in each ecoregion had sufficiently high N~2~O emission rates to offset N~2~O uptake in the majority of waterbodies, and none of the ecoregions are estimated to functione as net N~2~O sinks.
+The model predicts that waterbodies with high nitrate (NO~3~) tend to be sources of N~2~O whereas those with low NO~3~ are generally sinks.
 Nitrate interacted with waterbody size, however, with large waterbodies having N~2~O concentrations closer to equilibrium than small waterbodies.
-We estimate that CONUS waterbodies emit `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` ) metric tons N~2~O year^-1^, equivalent to only `r (national_stats %>% select(UCL_fn2o) %>% pull() / 22767) * 100`% of N~2~O emissions from all CONUS surface waters (e.g.
+Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS waterbodies is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` ) metric tons N~2~O year^-1^.
+The mean estimate is a net emission of  `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^, which is small relative to other known sources of N2O in the CONUS (e.g., wastewater treatment, manure management) and equivalent to only `r (national_stats %>% select(UCL_fn2o) %>% pull() / 22767) * 100`% of N~2~O emissions from all CONUS surface waters (e.g.
 zero order streams through river mouths) per IPCC estimates .
-The demonstrated capacity for widespread N~2~O uptake in CONUS waterbodies highlights the need to revisit global N~2~O models which presume surface waters are a N~2~O source [@epa_2023].
+Emission estimates were most uncertain for large waterbodies, suggesting that efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nation’s largest lakes and reservoirs.
+Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS waterbodies are a net source of sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source [@epa_2023].
 

### 2023-10-16 scripts/NLA_N2O_models.Rmd
Updated files for modeling workflow documentation

@@
 ---
-title: "Modeling workflow: Lake N2O survey data"
+title: "Modeling workflow: NLA N2O survey data"
 author: "Roy Martin, Jake Beaulieu, Michael McManus"
@@
 
-```{r begin, eval=TRUE, include=FALSE}
+```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
 library(ggpubr)
@@
 library(brms)
+library(tictoc)
+library(arrow)
 
@@
 
-# function for DHARMa residual analysis
-check_brms <- function(model,             # brms model
-                       integer = FALSE,   # integer response? (TRUE/FALSE)
-                       plot = TRUE,       # make plot?
-                       resp = NULL,
-                       ...                # further arguments for DHARMa::plotResiduals 
-) {
-
-  mdata <- brms::standata(model)
-  if(!"Y" %in% names(mdata))
-    oResp <- mdata[[paste0("Y_", resp)]]
-  else
-    oResp <- mdata[["Y"]]
-  #  stop("Cannot extract the required information from this brms model")
-
-  dharma.obj <- DHARMa::createDHARMa(
-    simulatedResponse = t(brms::posterior_predict(model, resp = resp, ndraws = 1000)),
-    observedResponse = oResp, 
-    fittedPredictedResponse = apply(
-      t(brms::posterior_epred(model, resp = resp, ndraws = 1000, re.form = NA)),
-      1,
-      median),
-    integerResponse = integer)
-
-  if (isTRUE(plot)) {
-    plot(dharma.obj, ...)
-  }
-
-  invisible(dharma.obj)
-
-}
+# Load model and data files for knitting (if they already exist)
+f1 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda"
+f2 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda"
+f3 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda"
+f4 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda"
+f5 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda"
+f6 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda"
+
+f7 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda"
+f8 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda"
+f9 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda"
+f10 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda"
+f11 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda"
+f12 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda"
+f13 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions_b.rda"
+f14 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda"
+f15 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda"
+f16 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda"
+f17 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda"
+f18 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda"
+f19 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests_size.rda"
+
+if(file.exists(f1))
+  load(f1)
+if(file.exists(f2))
+  load(f2)
+if(file.exists(f3))
+  load(f3)
+if(file.exists(f4))
+  load(f4)
+if(file.exists(f5))
+  load(f5)
+if(file.exists(f6))
+  load(f6)
+if(file.exists(f7))
+  load(f7)
+if(file.exists(f8))
+  load(f8)
+if(file.exists(f9))
+  load(f9)
+if(file.exists(f10))
+  load(f10)
+if(file.exists(f11))
+  load(f11)
+if(file.exists(f12))
+  load(f12)
+if(file.exists(f13))
+  load(f13)
+if(file.exists(f14))
+  load(f14)
+if(file.exists(f15))
+  load(f15)
+if(file.exists(f16))
+  load(f16)
+if(file.exists(f17))
+  load(f17)
+if(file.exists(f18))
+  load(f18)
+if(file.exists(f19))
+  load(f19)
 ```
@@
 # Background and Objectives
-This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O gas concentrations and saturation ratios for all freshwater lakes and reservoirs in the lower 48 US states larger than 4 ha.
-Data for these estimates came from the 2017 Nation Lakes Assessment (NLA) survey, wherein waterbodies were sampled according to a spatially balanced, stratified, and unequal probability design.
-Stratification was among categories of lake size (surface area in hectares), WSA9 ecoregion, and US state (excluding AK and HI); and larger lakes were intentionally over-sampled relative to smaller lakes.
-In order for this sample to be useful for making inferences relevant to the population of interest, any estimates derived from it needed to be adjusted for potential biases arising from these design factors.
-It can be useful to consider the complete data likelihood when imagining how such biases may arise in sampling designs[@Link_Barker_2010; @Zachmann_etal_2022].
-For the 2017 NLA survey, the complete data was considered to be all US lakes and reservoirs in the lower 48 states larger than 4 hectares.
-The survey samples, by comparison, are then considered a subset of the complete data with a known pattern of "missingess" assigned by the survey.
-For example, some lakes in the target population were missing from the sample not at random, but conditional on the design parameters.
-Their "missingness" is not random and, therefore, not ignorable when making inferences from the sample to the population of interest [@Gelman_etal_2014, Ch.
+This document describes the modeling workflow and results from estimating dissolved and equilibrium N2O gas concentrations and saturation ratios for all freshwater waterbodies (natural and man-made lakes and reservoirs) in the lower 48 US states larger than 1 ha.
+Data for these estimates came from the 2017 Nation Lakes Assessment (NLA) survey, wherein a subset of waterbodies were sampled from the broader population of interest (POI) according to a spatially balanced, stratified, and unequal probability design.
+Stratification was among categories of lake size (surface area in hectares), WSA9 ecoregion, and US state (excluding AK and HI).
+In order to better capture variability among the wide range of lake sizes in the POI, larger lakes were intentionally over-sampled relative to smaller lakes.
+
+To make inferences back to the POI from this sample, estimates needed to be adjusted in a manner that addressed potential biases related to the relative representation of waterbody characteristics in the sample compared to the POI, specifically with respect to the survey design variables.
+It can be helpful to define the complete data likelihood (CDL) when considering how to address biases that may arise from sampling designs [@Link_Barker_2010; @Zachmann_etal_2022].
+For the 2017 NLA survey, the complete data would be considered all US lakes and reservoirs in the lower 48 states larger than 1 hectare (_i.e._, the POI).
+The survey samples, by comparison, could then be considered a subset of that complete data with a known pattern of "missingess" assigned by the survey.
+In this case, waterbodies in the POI were missing from the sample not completely at random, but according to a probability assigned by the design parameters.
+This type of missingness is considered not ignorable when making inferences from a sample to a POI [@Gelman_etal_2014, Ch.
 8; @Zachmann_etal_2022].
-
-One model-based strategy to account for this selection bias is to include the survey design variables as predictors in a regression model and then adjust the resulting estimates based on a known target population distribution; a post-fitting adjustment referred to as poststratification [@Gelman_etal_2020, Ch.
+However, _conditional_ on the design variables, the sample could be considered a simple random sample (SRS) from the POI [@Gelman_etal_2014, Ch.
+8; @Zachmann_etal_2022].
+
+A simple, model-based strategy for adjusting a sample to a POI is to include the survey design variables as predictors in a regression model [@Gelman_etal_2020, Ch.
 17].
-A particularly popular version of this approach uses multilevel regression models and is aptly referred to as multilevel regression and poststratification, or MRP [@Park_etal_2004; @Gelman_Little_1997; @Gelman_etal_2020, Ch.
+After fitting the regression, the parameter estimates can be used to make adjustments according to the distribution of the design variables in the POI in a second step referred to as "poststratification" [@Gelman_etal_2020, Ch.
+17].
+A particularly popular version of this approach uses multilevel regression models and is referred to as multilevel regression with poststratification, or MRP [@Park_etal_2004; @Gelman_Little_1997; @Gelman_etal_2020, Ch.
 17; @Kennedy_Gelman_2021].
-Multilevel models are often recommended because they can provide regularized estimates along the design groupings, which can improve out-of-sample inferences [e.g., @Gelman_Little_1997; @Kennedy_Gelman_2021].
+Multilevel regression models are often recommended because they can provide regularized estimates along the design groupings, which can improve out-of-sample inferences [_e.g._, @Gelman_Little_1997; @Kennedy_Gelman_2021].
 Estimates for group levels that may be missing from the sample, but are part of the population of interest, are also straightforward using the multilevel approach [@Gelman_etal_2020 Ch.
 17; @McElreath_2020].
-
-With a stratified, unequal probability design, such as the one used for the 2017 NLA, the selection design prescribes a number of random draws for each of a number of pre-specified groupings (i.e., sampling unit types).
-Within a particular stratification level or grouping, the draws may be considered a simple random sample (SRS) from that group type in the target population.
-A regression model conditioning the sample observations on the design grouping structure could, therefore, be parameterized to provide group-specific estimates that are unbiased with respect to the target population.
-To make inferences to the broader target population, however, those regression estimates would need to be re-weighted by the known proportions of units across groups in the target population.
-If the estimates are not adjusted in this manner, the estimates may be biased, particularly if the response of interest varies importantly across the grouping structure.
-For example, for the 2017 NLA survey, the proportion of samples prescribed to small lakes was much lower, relative to their proportion in the target population, compared to larger lakes, which were sampled in greater proportion relative to the target population.
-Small lakes are so numerous in the US that, if this unequal weighting by lake size weren't prescribed, nearly all the samples from the survey would be from small lakes and inferences regarding larger lakes could be highly uncertain for any reasonably achievable number of samples.
-On the other hand, if lake size had no real influence on the N2O responses, the stratification across size class would make little difference with regard to inferences.
-Sampling any sizes of lakes would result in the same marginal estimates and standard errors.
-
-For a recent applied example and helpful tutorial employing these concepts in a model-based approach using MRP, see @Kennedy_Gelman_2021.
+For a recent applied example and helpful tutorial employing this approach, see @Kennedy_Gelman_2021.
 For an example of a recent application in the context of national surveys of environmental resources, see @Zachmann_etal_2022.
-In this study, a similar model-based approach was used to make population-level estimates from the 2017 NLA dissolved gas data.
-The specific workflow, data, models, and code used to make these estimates are documented in the remainder of this document.
-First, Bayesian multilevel regression models were used to fit the sample data to the design variable structure.
-Next, because eventual flux estimates needed to be estimated from lake-level measures (i.e., surface area), instead of predicting to a typical postratification table, predictions were made to each individual lake in the population of interest.
-This meant predicting to 465,897 individual natural and man made US lakes larger than 4 hectares in the lower 48 states.
+
+In this study, we used a MRP-like approach to provide population-level estimates based on the 2017 NLA dissolved gas survey data.
+The remainder of this document provides the workflow, data, models, and code used to make these estimates.
+To begin, a number of Bayesian multilevel regression models were fit to the sample data.
+These models were fit and checked against both the data used to construct the models as well as hold-out data from future visits.
+After evaluating a number of different models, a final model was selected for use in inference.
+Because total N2O flux estimates needed to be estimated based on individual, waterbody-level measures (_e.g._, surface area), predictions from the model were made to each individual lake in the POI (n = 465,897).
 These predictions were assumed relevant to average conditions during the biological index period for each lake in 2017.
-The specific objective of the modeling effort was to provide population estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1).
-As previously indicated, the gas estimates would also be used to estimate the total flux of N2O gas attributable to the target population of lakes over the index period.
+
+The objective of the modeling effort was to provide estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1) for the National-level POI and a number of sub-populations therein (_e.g._, ecoregion, state, size category).
+The gas estimates were also used to estimate the total flux of N2O gas attributable to the target populations.
 Saturation ratio was calculated as the ratio of dissolved to equilibrium N2O.
-Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), models were fit to their joint distribution and the ratio estimates were assembled as a derived quantity.
-The response variable was, therefore, multivariate in order to account for any potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography.
+Because dissolved and equilibrium N2O were observed on the same sample units (sites in lakes), the regression models were fit to their joint distribution and the ratio estimates were assembled as a derived quantity calculated from the joint posterior predictive distribution of those variables.
+The response variable was, therefore, modeled as multivariate in order to account for any potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography.
 Although predictions of the mean marginal probabilities from separate models may have provided comparable estimates, a joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
@@
 
-Gas samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L.
+Water samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L.
 As part of the survey design, each gas observation was indexed to an individual lake selected with unequal probability from 5 different lake size categories, $j \in j=1,...,J = 5$, according to surface area (ha), and from within a state, $k \in k=1,...,K = 48$, situated within an aggregated, WSA9 or Omernik ecoregion, $l \in l=1,...,L = 9$.
-All 9 WSA9 ecoregions were represented in the sample, including Xeric (XER), Western Mountain (WMT),  Northern Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and Southern Appalachian (SAP) regions.
-As shown below, the data from the initial and revisit samples were separately compiled into data frame objects in $\textbf{R}$, with $n=984$ and $n=95$ rows, respectively, of gas observations indexed to the survey design variables and several potentially relevant covariates.
+All WSA9 ecoregions were represented, including the Xeric (XER), Western Mountain (WMT),  Northern Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and Southern Appalachian (SAP) regions.
+Gas data from the first and second (follow-up) visits were separated, with $n=984$ _vs._ $n=95$ observations each.
+In addition to the design variables, a number of potentially useful covariates were also indexed to the gas observations, including measures of NO3, surface water temperature, elevation, chlorophyll a content, dissolved oxygen content, and waterbody size (hectares).
+These gas data and covariates were previously described and munged at: https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html.
 
 ## Import
-The gas data and covariates were previously described and munged at 
-https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html. That dataset was imported below.
+That originally munged gas dataset was imported below.
 
@@
 
-A new data frame for modeling was constructed from the original file including only the variables of interest: (1) the N2O gas observations; (2) the survey design variables indexed to those observations; and (3) additional covariates considered potentially useful for improving the fit of the model.
+A new data frame for modeling was constructed from the original file, including only the variables of interest: (1) the N2O gas observations; (2) the survey design variables indexed to those observations; and (3) additional covariates considered potentially useful for improving the fit and/or practical interpretation of the model.
 The modeling data frame below excluded the second-visit observations, which would later be used for model checking.
 Some variables from the imported data were renamed for convenience.
-In addition, the NO3 covariate was rounded according to the documented measurement precision and an alternative version was also created by log-transforming and re-coding the variable as an ordered factor with five levels.
+In addition, the NO3 covariate was rounded according to the documented measurement precision and an alternative version was created by log-transforming and re-coding the variable as an ordered factor with five levels.
 The left-most cut point separated observations below the detection limit from the completely observed samples.
@@
 
-```{r model_data, echo=TRUE, paged.print=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda")
-
+```{r all_data_n, echo=TRUE, paged.print=TRUE}
 dg %>%
-  filter(!is.na(dissolved.n2o.nmol)) %>% # 1 obs with missing measurement
+  filter(!is.na(dissolved.n2o.nmol)) %>% # 1 obs with missing N2O measurement
   nrow() # number of observations before filtering
-
+```
+
+```{r model_data, eval=FALSE, include=TRUE}
 df_model <- dg %>%
@@
   filter(sitetype == "PROB") %>% # probability samples only
-  filter(visit.no == 1) %>%
+  filter(visit.no == 1) %>% # first-visit sites only
   mutate(n2o = round(dissolved.n2o.nmol, 2),
@@
          sqrt_bf = sqrt(bf_max),
-         size_cat = recode(area.cat6, 
+         size_cat = recode(area.cat6, # simpler naming conventions for the size categories
                            "(1,4]" = "min_4" ,
@@
 save(df_model, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda") 
-
+```
+
+A preview showing the first 10 rows of the modeling data:
+
+```{r model_data_print, echo=TRUE, paged.print=TRUE}
 nrow(df_model) # number of obs after filtering
@@
 
-A second data frame including only the second visit observations was constructed below.
+A second data frame, including only the second visit observations was constructed below.
 These data were later used as a "test set" to assess the out-of-sample performance of the model developed on the first-visit or "training set".
 
-```{r test_data, echo=TRUE, paged.print=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda")
-
+```{r all_data_n_2, echo=TRUE, paged.print=TRUE}
 # number of observations before filtering probability samples
@@
   nrow()
-
+```
+
+```{r test_data, eval=FALSE, include=TRUE}
 df_test <- dg %>%
@@
   filter(sitetype == "PROB") %>% # probability samples only
-  filter(visit.no == 2) %>%
+  filter(visit.no == 2) %>% # second-visit sites only
   mutate(n2o = round(dissolved.n2o.nmol, 2),
@@
 save(df_test, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda") 
-
+```
+
+A preview of the first 10 rows of the test data:
+
+```{r test_data_print, echo=TRUE, paged.print=TRUE}
 nrow(df_test) # number of obs after filtering for probability samples, first visits, and removing one site missing ecoregion (WSA9) info.
@@
 ## Target population
-Below. the NLA sampling frame was imported and filtered to include all lakes the target population. The resulting target population above included a total of 465,897 waterbodies.
-
-```{r import_sample_frame, echo=TRUE}
+
+Below. the original NLA sampling frame was imported and filtered to include only lakes in the POI. The resulting target population included a total of 465,897 waterbodies.
+
+```{r import_sample_frame, eval=FALSE, include=TRUE}
 df_pop <- read.csv(file = paste0(localPath,
@@
 save(sframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda") 
-
+```
+
+Below is a sample of the first ten rows describing the POI:
+
+```{r sample_frame_print, echo=TRUE, paged.print=TRUE}
 sframe %>%
@@
 
-Cross tabulations below describe the structure of the target population with respect to the survey design variables.
+Cross tabulations below show the structure of the target population with respect to the survey design variables.
 The cross-tabulation shows that each ecoregion does not contain each state.
-Therefore, in the statistical sense, states were nested in ecoregions.
+Therefore, states were nested within ecoregions.
 
 ```{r frame_dimensions_1, echo=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")
-
 sframe %>%
@@
 
-The cross-tablulation below indicates that lake size category was nested in state (which was nested in ecoregion). That is, not every ecoregion:state contained every size category.
+The cross-tablulation below indicates that lake size category was nested within state (which was nested in ecoregion). That is, not every ecoregion:state contained every size category.
 
 ```{r frame_dimensions_4, echo=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")
-
 sframe %>%
@@
 
-Below, the sampling frame was munged to create a post-stratification table.
-There were 536 "types" or groupings of lakes in the population of interest with respect to the sampling design.
+Below, the sampling frame was used to create a typical post-stratification table.
+There were 536 "types" or groupings of lakes in the POI with respect to the sampling design.
 The total counts of those lake types (n_lakes) and their proportions (prop_cell) relative to the counts in the target population were tabulated.
 
-```{r filter_frame, echo=TRUE, message=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")
-
+```{r filter_frame, eval=FALSE, include=TRUE}
 pframe <- sframe %>%
@@
 save(pframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
-
+```
+
+```{r filter_frame_print}
 pframe %>%
@@
 ## Sample vs. population
-Below, the lake distributions in the population of interest were compared to the proportions in the observed sample.
-There were 352 lake types in the sample compared to the 536 in the population of of interest.
+
+Below, the distribution of waterbody types in the POI were compared to their proportions in the observed sample.
+There were 352 waterbody types in the sample compared to the 536 in the POI.
 There were 984 observations distributed across these 352 lake types in the sample; and the number of samples was not distributed evenly across the types.
@@
 
-```{r sample_cell_counts, echo=TRUE, message=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
+```{r sample_cell_counts, eval=FALSE, include=TRUE}
 samp_props <- df_model %>%
@@
 save(samp_props, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
-
+```
+
+```{r sample_cell_count_print, echo=TRUE, paged.print=TRUE}
 samp_props %>%
-  head(10)
-```
-
-Below, a graphical comparison was constructed to depict the distribution of cells in the population of interest _vs._ those in the sample.
-
-```{r compare_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
-
+  print()
+```
+
+Below, a graphical comparison was constructed to depict the distribution of cells in the POI _vs._ those in the sample.
+
+```{r compare_sample_pop_cells, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=10}
 pframe %>%
@@
 
-Another comparison between population and sample was constructed by ecoregion.
-The samples were not balanced across ecoregions.
+Another comparison between population and sample was constructed by ecoregion below.
 Lakes in the Coastal Plains (CPL) ecoregion, for example, were clearly undersampled relative to their proportion of the population.
 
-```{r eco_props_pop, eval=FALSE, include=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
-
+```{r eco_props_pop, fig.align='center', fig.asp=0.5, fig.width=10}
 pframe_eco <- pframe %>%
@@
 
-save(pframe_eco, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_eco.rda")
-```
-
-```{r eco_props_sample, eval=FALSE, include=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
-
 samp_props_eco <- samp_props %>%
@@
   mutate(type = 'sample')
-
-save(samp_props_eco, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_eco.rda")
-```
-
-```{r compare_eco_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=6}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_eco.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_eco.rda")
 
@@
 
-```{r state_props_pop, eval=FALSE, include=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
-
+```{r state_props_pop, fig.align='center', fig.asp=0.5, fig.width=10}
 pframe_state <- pframe %>%
@@
 
-save(pframe_state, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_state.rda")
-```
-
-```{r state_props_sample, eval=FALSE, include=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
-
 samp_props_state <- samp_props %>%
@@
   mutate(type = 'sample')
-
-save(samp_props_state, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_state.rda")
-```
-
-```{r compare_state_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_state.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_state.rda")
 
@@
 
-Finally, a comparison by lake size category is shown below. Small lakes were under-sampled relative to larger lakes.
-
-```{r size_props_pop, eval=FALSE, include=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda")
-
+Finally, a comparison by NLA size category is shown below.
+Waterbodies of the smallest size category were heavily under-sampled relative to the POI and those from the larger size categories were over-sampled.
+
+```{r size_props_pop, fig.align='center', fig.asp=0.5, fig.width=10}
 pframe_size <- pframe %>%
@@
 
-save(pframe_size, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_size.rda")
-```
-
-```{r size_props_sample, eval=FALSE, include=TRUE, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda")
-
 samp_props_size <- samp_props %>%
@@
   mutate(type = 'sample')
-
-save(samp_props_size, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_size.rda")
-```
-
-```{r compare_size_sample_pop_cells, echo=FALSE, fig.align='center', fig.height=4, fig.width=6}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe_size.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props_size.rda")
 
@@
 ## Sample-based estimates
-Below are _naive_ estimates, based only on the sample, for national means for dissolved and equilibrium N2O and the saturation ratio.
+
+Below are (naive) estimates, based only on the sample, for national means for dissolved and equilibrium N2O and the saturation ratio.
+
+Dissolved N2O:
 
 ```{r sample_summary_n2o}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 
+Equilibrium N2O:
+
 ```{r sample_summary_n2oeq}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 
+Saturation ratio:
+
 ```{r sample_summary_n2osat}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 ```{r sample_summary_propsat}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 The shaded areas indicate +/- 1 standard deviation.
-Neither dissolved N2O nor the saturation ratio were clearly structured by ecoregion, but there did appear to be some structure along this variable in the equilibrium N2O observations.
-
-```{r sample_summary_eco, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
+Neither dissolved N2O nor the saturation ratio were clearly structured by ecoregion in the sample, but there did appear to be some potential structure along this variable in the equilibrium N2O observations.
+
+```{r sample_summary_eco, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=10, message=FALSE}
 p1 <- df_model %>%
@@
 
-```{r sample_summary_state, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
+```{r sample_summary_state, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=10, message=FALSE}
 p1 <- df_model %>%
@@
 
-```{r sample_summary_size, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
+```{r sample_summary_size, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=10, message=FALSE}
 p1 <- df_model %>%
@@
 ## Sample data exploration
+
 Below, the distribution N2O concentrations for the sample was summarized using a density and rug plot.
@@
 Both the dissolved and equilibrium N2O data had considerable right skew even after the log transformation.
-This was not unexpected and has been noted in other studies [@Webb_etal_2019].
-The saturation ratio was also skewed since it was derived from the other two observed variables (i.e., sat_ratio = n2o / n2o_eq).
-
-```{r summary_N2O, echo=FALSE, fig.align='center', fig.height=8, fig.width=5, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
+This was not unexpected and has been noted in other studies [_e.g._, @Webb_etal_2019].
+The saturation ratio was also skewed (sat_ratio = n2o / n2o_eq).
+
+```{r summary_N2O, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=10, message=FALSE}
 p1 <- df_model %>%
@@
 ```{r summary_N2O_vs_NO3, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 ```{r summary_N2O_vs_NO3_surftemp, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 ```{r summary_N2O_vs_NO3_logarea, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 ```{r summary_N2O_vs_NO3cat, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 ```{r summary_N2O_vs_NO3cat_surftemp, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 ```{r summary_N2O_vs_NO3cat_logarea, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 
-Below is a plot of log(dissolved N2O) vs. log(NO3) by ecoregion, which suggested that the NO3 effect on N2O may have varied by ecoregion.
+Below is a plot of log(dissolved N2O) vs. log(NO3) by ecoregion, which suggested that the NO3 effect on dissolved N2O may have varied by ecoregion.
 
 ```{r summary_N2O_vs_NO3_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 ```{r summary_N2O_vs_NO3cat_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 
-A plot below shows trends by state within just the Temperate Plains (TPL) ecoregion. Within states, the number of observations were relatively small, but the trends appeared closer to linear.
+A plot below shows the relationships between dissolved N2O and log(NO3) by state, but within just the Temperate Plains (TPL) ecoregion.
+Within states, the number of observations were relatively small, but the trends were perhaps more linear.
 
 ```{r summary_N2O_vs_NO3_wsa9state3, echo=FALSE, fig.align='center', fig.height=4, fig.width=8, message=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 df_model %>%
@@
 # Model fitting
-The first regression model was constructed to estimate the joint distribution of log-transformed dissolved and equilibrium N2O conditional on the the design factors.
+The first regression model was constructed to estimate the joint distribution of log-transformed dissolved and equilibrium N2O conditional only on the the survey design factors.
 Each log-transformed observation, $i \in 1,..,N=984$, for each response, $p \in 1:P=2$, was assumed to be drawn from a multivariate normal distribution with the parameters $\nu$ and $\Sigma$, where $\nu$ is the multivariate mean estimated conditional on the design effects and $\Sigma$ is a covariance matrix containing the observation-level variances and residual correlation:
@@
 Finally, for the correlation matrices, an $LKJ(\eta =2)$ prior was used, which, for a 2-dimensional response, placed most support for correlations between approximately -0.9 and 0.9.
-This prior seemed reasonable as there were no apparant causal mechanisms to ensure a very strong correlation between the N2O measures.
+This prior seemed reasonable as there were no clear causal mechanisms to ensure a very strong correlation between the N2O measures.
 Any potential residual dependence was expected to be indirect due to, for example, a common correlate (e.g., elevation, temperature).
@@
 
-In the $\textbf{brms}$ package, there is additional functionality and syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:
+In the $\textbf{brms}$ package, there is additional functionality and formula syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:
 
@@
 
-The above syntax would indicate that the linear predictor for both responses in the multivariate model have the same group-level varying effects, and that those effects may be correlated between responses.
+The above formula syntax would indicate that the linear predictor for both responses in the multivariate model have the same group-level varying effects, and that those effects may be correlated between responses.
 
@@
 ## Model 1
-The first model fit was as explained above.
+
+The first model fit.
 
 ```{r n2o_mod_mv_1, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 bf_n2o <- bf(log(n2o) ~ 1 + 
@@
 ### Summarize fit
+
 The summaries of the estimated parameters and key HMC convergence diagnostics for the fitted model are printed below.
@@
 ```{r print_mod1, echo=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
-
 print(n2o_mod1, prior = T)
@@
 
-Before investing too much into the inferences from this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch.
+Before investing too much into inferences from this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch.
 11].
@@
 ### Model checks
+
 #### Dissolved N2O
+
 Below are a series of panels illustrating graphical PPCs for the dissolved N2O component of the multivariate model.
@@
 
-```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 
-```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
+```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 
-```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 
-```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 
-```{r ppc_biv1, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r ppc_biv1, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 
-```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 df_test %>%
@@
 
-```{r ppc_sat1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r ppc_sat1, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 
-```{r ppc_sat1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r ppc_sat1_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### R-square
+
 Below, the Bayesian $R^2$ values are reported for each response in the model.
 
-```{r r2_1, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r r2_1, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE}
 round(bayes_R2(n2o_mod1, resp = "logn2o", cores = 1), 3)
@@
 
-```{r r2_1_test, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
+```{r r2_1_test, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE}
 round(bayes_R2(n2o_mod1, resp = "logn2o", newdata = df_test, cores = 1), 3)
@@
 ## Model 2
-In an attempt to better fit the observed data, the next model included a distributional model for each sub-model that allowed for heterogeneous variances.
-The distributional terms were each fit as a function of the survey design structure.
-The same structure as for the models for the mean components.
+
+In an attempt to better fit the observed data, the next model included a distributional sub-model that allowed for heterogeneous variances.
+The distributional terms for both N2O responses were each fit as a function of the survey design structure.
+The same structure as for the models for the submodels for the mean component before.
 
 ```{r n2o_mod_2, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 bf_n2o <- bf(log(n2o) ~ 1 +
@@
 ```{r print_mod2, echo=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
-
 print(n2o_mod2, prior = T)
@@
 ### Model checks
+
 Below the same PPCs were performed as with the initial model (see above for more details on each panel).
@@
 
-```{r ppc_n2o2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
+```{r ppc_n2o2, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Equilibrium N2O
+
 The checks below suggested that this model offered little no improvement upon the initial model for equilibrium N2O.
 
-```{r ppc_n2oeq2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
+```{r ppc_n2oeq2, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Bivariate
+
 This check perhaps suggested an improvement with regard to replicating the joint density. However, the predictions were still clearly over-dispersed relative to the observations.
 
-```{r ppc_biv2, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
+```{r ppc_biv2, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 #### Saturation
+
 The PPCs for the saturation metrics below indicated that including the distributional models was perhaps an improvement on the initial model in some aspects; in particular, the bias in the predicted proportion of under-saturated lakes was substantially decreased.
@@
 
-```{r ppc_sat2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
+```{r ppc_sat2, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### R-square
-Relative to model 1, there was a substantial decrease in the $R^2$ estimate for the dissolved N2O component of this model. The estimate for the equilibrium N2O-eq component was similar to the model 1.
-
-```{r r2_2, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda")
+
+Relative to model 1, there was a substantial decrease in the $R^2$ estimate for the dissolved N2O component of this model. The estimate for the equilibrium N2O-eq component was similar to model 1.
+
+```{r r2_2, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
 round(bayes_R2(n2o_mod2, resp = "logn2o", cores = 1), 3) 
@@
 ## Model 3
-In the next model, the categorical version of the NO3 covariate and an surface temperature covariate were included to try to improve the fit.
+
+In the next model, the categorical version of the NO3 covariate and a surface temperature covariate were included to try to improve the fit.
 The ordinal NO3 variable was used as a monotonic, ordinal effect and only in the dissolved N2O component of the model.
@@
 ```{r n2o_mod_3, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
@@
 ### Summarize fit
+
 The fitted parameters and MCMC diagnostics are below.
@@
 ```{r print_mod3, echo=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
-
 print(n2o_mod3, prior = T)
@@
 ### Model checks
+
 #### Dissolved N2O
+
 The PPCs below indicated a better fit compared to the previous models.
 The central tendency and tail behavior looked to be reasonably replicated by comparison.
-However, the observed _vs._ predicted plot suggested that larger overserved values were likely being systematically underestimated.
-
-```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
+However, the observed _vs._ predicted plot suggested that larger N2O observations were likely being systematically underestimated.
+
+```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Equilibrium N2O
+
 The PPCs below indicated that this model appeared to be an improvement for equilibrium N2O as well. However, some checks (e.g., skewness) suggested some room for additional improvement.
 
-```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
+```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Bivariate
+
 The check for the joint distribution below also suggested an improvement upon the previous models.
 
-```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
+```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 #### Saturation
+
 This model looked to be an improvement with regard to the PPCs for the saturation metrics.
@@
 
-```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
+```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### R-square
-The $R^2$ estimates for this model are below and suggested substantial improvements on the previous models.
-
-```{r r2_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
+
+The $R^2$ estimates for this model are below and suggested substantial improvements over the previous models.
+
+```{r r2_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
 round(bayes_R2(n2o_mod3, resp = "logn2o", cores = 1), 3) 
@@
 ### Covariate effects
+
 Below are plots illustrating the modeled effects of covariates on both N2O and equilibrium N2O.
@@
 #### Dissolved N2O
+
 The conditional effects plots below for dissolved N2O illustrated a positive, monotonic, and non-linear relationship with NO3; and a negative, linear relationship with surface temperature.
 
-```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
-
+```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod3, 
@@
 
-annotate_figure(plt, top = text_grob("N2O: conditional effects", 
+annotate_figure(plt, top = text_grob("Conditional effects: N2O", 
                color = "black", face = "bold", size = 14))
@@
 #### Equilibrium N2O
+
 The modeled effects for the equilibrium N2O component illustrated a negative relationship with both predictors and an interaction such that the surface temperature effect became slightly steeper at lower elevations.
 
-```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=3, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
+```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod3, 
@@
 
-annotate_figure(plt, top = text_grob("Equilibrium N2O: conditional effects", 
+annotate_figure(plt, top = text_grob("Conditional effects: Equilibrium N2O", 
                color = "black", face = "bold", size = 14))
@@
 ## Model 4
+
 In the next model, covariate terms were also included in the $\sigma$ components of both models in order to try to better capture remaining heterogeneity in the variances of both N2O and N2O-eq.
@@
 ```{r n2o_mod_mv_4, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
@@
 ### Summarize fit
+
 Below is a summary of the fitted parameters along with some convergence diagnostics.
@@
 ```{r print_mod4, echo=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
-
 print(n2o_mod4, prior = T, digits = 3)
@@
 ### Model checks
+
 The same PPCs were employed for this model as above.
@@
 #### Dissolved N2O
-This model appeared to be an improvement on the previous model, particularly with regard to the more constant variance indicated in the observed _vs._ predicted plot (bottom, right panel).
-
-```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
+
+This model appeared to be a moderate improvement on the previous models.
+
+```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Equilibrium N2O
+
 This component of the model also seemed to be an improvement over model 3, with better representation in the tails as indicated in the skewness _vs._ kurtosis PPC.
 
-```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
+```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Bivariate
+
 Again, an improvement over the previous model with a tighter fit of the PPC to the observed bivariate density.
 
-```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
+```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 #### Saturation
+
 This check also suggested an improvement over the previous models, with better tail behavior and less bias in the proportion under-saturated measure.
 
-```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
+```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### R-square
+
 The Bayesian $R^2$ estimates below indicated an improvement from the previous models.
 
-```{r r2_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
+```{r r2_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
 round(bayes_R2(n2o_mod4, resp = "logn2o", cores = 1), 3) 
@@
 ### Covariate effects
+
 #### Dissolved N2O
-The conditional effects plots for the covariate effects on dissolved N2O remained largely unchanged from the previous model.
-
-```{r conditional_effects_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
-
+
+The conditional effects plots for the covariate effects on dissolved N2O remained similar to the previous model.
+
+```{r conditional_effects_mod_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod4, 
@@
 
-annotate_figure(plt, top = text_grob(expression(paste("N2O: covariate effects on ", mu)), 
+annotate_figure(plt, top = text_grob(expression(paste("Covariate effects on ", mu, ": N2O")), 
                color = "black", face = "bold", size = 14))
@@
 
-```{r conditional_effects_sigma_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
-
+```{r conditional_effects_sigma_n2o4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod4, 
@@
 annotate_figure(plt, 
-                top = text_grob(expression(paste("N2O: covariate effects on ", sigma)),
+                top = text_grob(expression(paste("Covariate effects on ", sigma, ": N2O")),
                                 color = "black",
@@
 #### Equilibrium N2O
+
 The covariate effects on equilibrium N2O remained largely the same as for the previous model.
 
-```{r conditional_effects_mod_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
-
+```{r conditional_effects_mod_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod4, 
@@
 
-annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", mu)), 
+annotate_figure(plt, top = text_grob(expression(paste("Covariate effects on ", mu, ": Equilibrium N2O")), 
                color = "black", face = "bold", size = 14))
@@
 
-```{r conditional_effects_sigma_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
-
+```{r conditional_effects_sigma_n2oeq4, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod4, 
@@
 
-annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", sigma)), 
+annotate_figure(plt, top = text_grob(expression(paste("Covariate effects on ", sigma, ": Equilibrium N2O")), 
                color = "black", face = "bold", size = 14))
@@
 ## Model 5
+
 In the next model, additional complexity is added to the dissolved N2O component by including a covariate for continuous lake surface area (log scale) as well as interactions between NO3 and log(surface area) and surface temperature.
+Although, some aspect of lake size was likely captured by the categorical size factor from the survey in the models above, the largest category is very coarse and includes all lakes >50 hecatares, which is an extremely wide distribution of sizes from 50 to over 300,000 hectares at max.
+Including the continuous size covariate at the observation level in this manner was expected to do a better job of resolving any trends with size.
+Colinearity between the grouped and observation-level measures were not a major, concern, since the observation-level measure would account for any variation first.
+However, we did later check for any common violations of the Gauss-Markov assumptions, by looking at potential correlations between the model errors and covariates [not shown; @Bafumi_Gelman_2007].
 
 ```{r n2o_mod_mv_5, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
@@
 ### Summarize fit
+
 Below is a summary of the fitted parameters along with MCMC convergence diagnostics.
@@
 ```{r print_mod5, echo=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
-
 print(n2o_mod5, prior = T, digits = 3)
@@
 ### Model checks
+
 Again, the same PPCs as above were performed for this model.
@@
 #### Dissolved N2O
+
 This PPCs for dissolved N2O looked similar to the previous model.
 
-```{r ppc_full_checks_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
+```{r ppc_full_checks_mod_n2o5, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Equilibrium N2O
+
 Again, the PPCs for this model were similar to the previous model, which was unsurprising given that it was the same model for N2O-eq.
 
-```{r ppc_full_checks_mod_n2oeq5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
+```{r ppc_full_checks_mod_n2oeq5, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Bivariate
+
 This PPC was also similar to the previous model.
 
-```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
+```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 #### Saturation
+
 This check was also similar to the prevoius model, with perhaps slightly less bias in the proportion unsaturated estimates.
@@
 
-```{r ppc_sat_check_mod_n2o5, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
+```{r ppc_sat_check_mod_n2o5, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 
-```{r r2_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
+```{r r2_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
 round(bayes_R2(n2o_mod5, resp = "logn2o", cores = 1), 3) 
@@
 ### Covariate effects
+
 #### Dissolved N2O
+
 The conditional effects plots suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature.
@@
 
-```{r conditional_effects_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
-
+```{r conditional_effects_mod_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=1, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod5, 
@@
 
-```{r conditional_effects_sigma_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
-
+```{r conditional_effects_sigma_n2o5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod5, 
@@
 #### Equilibrium N2O
+
 The estimated covariate effects on equilibrium N2O remained largely the same as estimated in the previous model.
 
-```{r conditional_effects_mod_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
-
+```{r conditional_effects_mod_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod5, 
@@
 
-```{r conditional_effects_sigma_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda")
-
+```{r conditional_effects_sigma_n2oeq5, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod5, 
@@
 ## A Final Model
-As demonstrated above, models excluding the NO3 covariate consistently resulted in poorer fits to to the observed dissolved N2O data and potentially strongly biased estimates of the saturation ratio.
+
+As demonstrated above, models excluding the NO3 covariate consistently resulted in poorer fits to to the observed dissolved N2O data and potentially biased estimates of the saturation ratio.
 Including surface temperature and elevation in the equilibrium N2O part of the model also resulted in substantially improved replication of key aspects of the observed data.
-Likewise, added flexibility in the distributional terms for both dissolved and equilibrium N2O led to improvements.
-
-To make inferences to the population of interest from a model including these covariates, however, however, the covariates needed to be (1) fully observed across that population or (2) their missingness needed to be modeled.
+Added flexibility in the distributional terms for both dissolved and equilibrium N2O also led to improvements.
+
+To make inferences to the population of interest from a model including these covariates, however, the covariates needed to be (1) fully observed across that population or (2) their missingness needed to be modeled.
 For the lake area and elevation covariates, data _was_ available for all lakes from previously compiled geospatial databases.
@@
 Therefore, a more complex model was constructed below that included surface temperature and NO3 as additional responses conditioned on the survey design variables and fully observed covariates.
-This approach to inference for N2O was similar to Bayesian structural equation modeling approaches [@Merkle_etal_2021; @Merkle_Rosseel_2018].
+This approach to inference for N2O was similar to a Bayesian structural equation modeling approach [@Merkle_etal_2021; @Merkle_Rosseel_2018].
 The logical dependence structure could be characterized as:
@@
 $$\begin{aligned} 
-{\boldsymbol{N_2O_{diss}}} &\sim Survey + Area + {\boldsymbol{NO_3}} + {\boldsymbol{Temp}} \\ 
-{\boldsymbol{N_2O_{equil}}} &\sim Survey + Elev  + {\boldsymbol{Temp}}\\
-{\boldsymbol{NO_3}} &\sim Survey + Area + {\boldsymbol{Temp}} \\ 
-{\boldsymbol{Temp}} &\sim Survey + Lat + Elev + Day
+{p( \boldsymbol{N_2O_{diss}}} &| Survey, Area, {\boldsymbol{NO_3}}, {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{N_2O_{equil}}} &| Survey, Elev, {\boldsymbol{Temp}})\\
+{p( \boldsymbol{NO_3}} &| Survey, Area, {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{Temp}} &| Survey, Lat, Elev, Day)
 \end{aligned}$$
@@
 The partially observed variables, being dissolved and equilibrium N2O, NO3, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.
-This piece-wise approach required a more complex set of post-processing steps compared to a typical MRP analysis.
-In order to propagate estimates and uncertainty through the dependency structure and make inferences, the fitted model was used to first predict surface temperature in the target population, since it depended only on the fully observed covariates.
+This piece-wise approach to predicting N2O required a more complex set of post-processing steps compared to a typical MRP analysis.
+In order to propagate uncertainty through the dependency structure and make inferences to the POI below, the fitted final model was used to first predict surface temperature in the target population, since it depended only on the fully observed covariates.
 That predictive distribution was then used alongside the relevant fully observed covariates to predict NO3 in the target population.
@@
 ```{r n2o_mod_mv_6, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-
 bf_n2o <- bf(n2o ~ mo(no3_cat) +
@@
 ```{r print_mod6, echo=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-
 print(n2o_mod6, digits = 3, prior = T)
@@
 ### Model checks
+
 Below, the same PPCs for dissolved and equilibrium N2O were used as before.
@@
 #### Dissolved N2O
+
 The PPCs for dissolved N2O were similar to those for models 4 and 5 above.
 
-```{r ppc_full_checks_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
+```{r ppc_full_checks_mod_n2o6, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Equilibrium N2O
+
 The PPCs for eqilibrium N2O were also similar to the same checks in models 4 and 5.
 
-```{r ppc_full_checks_mod_n2oeq6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
+```{r ppc_full_checks_mod_n2oeq6, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### Bivariate
-This model provided a reasonable representation of the bivariate relationship between the two N2O responses.
-
-```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
+
+This final model provided a reasonable representation of the bivariate relationship between the two N2O responses.
+
+```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 #### Saturation
+
 The saturation ratio PPCs below suggested similar behavior as with models 4 and 5 above, but with perhaps slightly less bias in the predictions for the proportion of undersaturated waterbodies and fewer extreme predictions for the means and standard deviations.
@@
 
-```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
+```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 
-```{r ppc_sat_check_testdata_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
+```{r ppc_sat_check_testdata_mod_n2o6, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
@@
 #### R-square
+
 Below are estimates for the Bayesian $R^2$, which were largely similar for the N2O responses as with models 4 and 5 above. The $R^2$ for the surface temperature response also suggested a good fit.
 
-```{r r2_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
+```{r r2_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
 round(bayes_R2(n2o_mod6, resp = "n2o"), 3) 
@@
 
-```{r r2_test_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.align='center', fig.width=4, fig.height=4}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
+```{r r2_test_mod_n2o6, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
 round(bayes_R2(n2o_mod6, resp = "n2o", newdata = df_test), 3) 
@@
 ### Covariate effects
+
 #### Dissolved N2O
+
 The conditional effects plot for the covariate effects on dissolved N2O suggested a similar effect of NO3 as in previous models, but with interesting potential interactions between NO3 and lake area and NO3 and surface temperature.
@@
 
-```{r conditional_effects_mod_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=4, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-
+```{r conditional_effects_mod_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=1, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod6, 
                           resp = "n2o", 
-                          effects = c("no3_cat"), 
+                          effects = c("no3_cat:log_area"), 
                           plot = F)
@@
                           plot = F)
+
 p3 <- conditional_effects(n2o_mod6, 
+                          resp = "n2o", 
+                          effects = c("no3_cat:surftemp"), 
+                          plot = F)
+
+p4 <- conditional_effects(n2o_mod6, 
                           resp = "n2o", 
@@
                  plot(p2, plot = F)[[1]],
-                 plot(p3, plot = F)[[1]])
-
-rm(p1, p2, p3)
-
-annotate_figure(plt, top = text_grob(expression(paste("N2O: covariate effects on ", mu)), 
+                 plot(p3, plot = F)[[1]],
+                 plot(p4, plot = F)[[1]])
+
+rm(p1, p2, p3, p4)
+
+annotate_figure(plt, top = text_grob(expression(paste("Covariate effects on ", mu, ": N2O")), 
                color = "black", face = "bold", size = 14))
@@
 
-```{r conditional_effects_sigma_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-
+```{r conditional_effects_sigma_n2oF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod6, 
@@
 annotate_figure(plt, 
-                top = text_grob(expression(paste("N2O: covariate effects on ", shape)),
+                top = text_grob(expression(paste("Covariate effects on ", shape, ": N2O")),
                                 color = "black",
@@
 #### Equilibrium N2O
+
 The estimated covariate effects on equilibrium N2O remained largely the same as estimated in the previous model.
 
-```{r conditional_effects_mod_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-
+```{r conditional_effects_mod_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod6, 
@@
 
-```{r conditional_effects_sigma_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=2, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-
+```{r conditional_effects_sigma_n2oeqF, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
 p1 <- conditional_effects(n2o_mod6, 
@@
 
-annotate_figure(plt, top = text_grob(expression(paste("Equilibrium N2O: covariate effects on ", shape)), 
+annotate_figure(plt, top = text_grob(expression(paste("Covariate effects on ", shape, ": Equilibrium N2O")), 
                color = "black", face = "bold", size = 14))
@@
 
+#### Saturation ratio
+
+Below are the estimated conditional effects of the covariates on means for both n2o responses as well as the implied effects on the mean saturation ratio.
+
+```{r conditional_effects_mod_n2osat_temp_F, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+effect_no3_surftemp <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               surftemp = round(quantile(df_model$surftemp, 
+                                                   probs = seq(0.10, 0.90, 0.2)), 0),
+                               log_area = mean(df_model$log_area),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         surftemp = factor(surftemp))
+
+effect_surftemp_no3 <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               surftemp = seq(range(df_model$surftemp)[1],
+                                              range(df_model$surftemp)[2],
+                                              length.out = 101),
+                               log_area = mean(df_model$log_area),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         temp = surftemp)
+
+plt_n2oeq_surftemp <-
+  effect_surftemp_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = surftemp,
+                      y = n2oeq_pred),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  xlab("") +
+  ylab(expression(paste("Equilibrium N"[2], "O (nmol/L)"))) +
+  theme_pubr() +
+  theme(axis.text.x = element_blank(),
+        legend.position = "none") +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 20), breaks = seq(0, 20, 5)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_fill_brewer(palette = "Greys", guide = "none")
+
+plt_n2o_surftemp_no3 <-
+  effect_surftemp_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = surftemp,
+                      y = n2o_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  xlab("") +
+  ylab(expression(paste("Dissolved N"[2], "O (nmol/L)"))) +
+  theme_pubr() +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 55), breaks = seq(0, 50, 10)) +
+  scale_fill_brewer(palette = "Greys", guide = "none") + 
+  theme(axis.text.x = element_blank(),
+        legend.position = "none")
+
+plt_n2o_no3_surftemp <-
+  effect_no3_surftemp %>%
+  ggplot(aes(x = no3_cat, 
+             y = n2o_pred, 
+             group = surftemp,
+             color = surftemp)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  xlab("") +
+  ylab("") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 55), breaks = seq(0, 50, 10)) +
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        legend.position = "none")
+
+plt_sat_no3_surftemp <-
+  effect_no3_surftemp %>%
+  ggplot(aes(x = no3_cat, 
+             y = sat_pred, 
+             group = surftemp,
+             color = surftemp)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  xlab(expression(paste("NO"[3], " category"))) +
+  ylab("") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  theme(legend.position = "none",
+        axis.text.y = element_blank())
+
+plt_sat_surftemp_no3 <-
+  effect_surftemp_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = surftemp,
+                      y = sat_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  xlab("Surface temperature (\u00B0C)") +
+  ylab("Saturation ratio") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  scale_fill_brewer(palette = "Greys", guide = "none") +
+  theme(legend.position = "none")
+
+legend1 <- effect_surftemp_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = surftemp,
+                      y = n2o_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  theme_void() +
+  scale_fill_brewer(palette = "Greys", guide = "none") + 
+  theme(legend.position = c(0.5, 0.1),
+        legend.direction = "horizontal",
+        legend.text = element_text(size = 10, angle = 45)) +
+  labs(color = expression(paste("NO"[3], " category"))) +
+  guides(color = guide_legend(title.position = "top", title.hjust = 0.5, title.vjust = 0.5))
+
+legend1 <- get_legend(legend1)
+
+legend2 <- 
+  effect_no3_surftemp %>%
+  ggplot(aes(x = no3_cat, 
+             y = n2o_pred, 
+             group = surftemp,
+             color = surftemp)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  theme_void() +
+  theme(legend.position = c(0.5, 0.5),
+        legend.direction = "horizontal",
+        legend.text = element_text(size = 10, angle = 45)) +
+  labs(color = "Surface temperature (\u00B0C)") +
+  guides(color = guide_legend(title.position = "top", title.hjust = 0.5, title.vjust = 1))
+
+legend2 <- get_legend(legend2)
+
+legendPlot <- ggarrange(legend1, legend2, ncol = 1)
+
+plt_surftemp_effects <- ggarrange(plt_n2oeq_surftemp,
+                                  legendPlot,
+                                  plt_n2o_surftemp_no3,
+                                  plt_n2o_no3_surftemp,
+                                  plt_sat_surftemp_no3,
+                                  plt_sat_no3_surftemp,
+                             ncol = 2,
+                             nrow = 3,
+                             widths = c(1, 1, 1, 1, 1, 1),
+                             heights = c(1, 1, 1, 1, 1, 1),
+                             labels = c("A", "Legend", "B", "C", "D", "E"),
+                             hjust = c(-6, -1, -6, -6, -6, -6),
+                             vjust = c(2, 2, 2, 2, 2, 2)
+                             )
+
+plt_surftemp_effects <- plt_surftemp_effects %>%
+  annotate_figure(plt_surftemp_effects,
+                  left = text_grob("Posterior mean", 
+                                   size = 18, 
+                                   rot = 90)
+                  )
+
+# ggsave only works in R console (i.e., copy + paste to save)
+ggsave(filename = "cond_effect_surftemp_no3.tif",
+       plot = plt_surftemp_effects,
+       device = "tiff",
+       width = 180 / 25.4,
+       height = 210 / 25.4,
+       units = "in")
+```
+
+```{r plot_surftemp_effects, echo=FALSE, fig.align='center', fig.asp=1.166667, fig.width=8}
+plt_surftemp_effects
+```
+
+```{r conditional_effects_mod_n2osat_area_F, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+effect_no3_area <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               log_area = round(quantile(df_model$log_area, 
+                                                   probs = seq(0.10, 0.90, 0.2)), 0),
+                               surftemp = mean(df_model$surftemp),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         log_area = factor(log_area))
+
+effect_area_no3 <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               log_area = seq(range(df_model$log_area)[1],
+                                              range(df_model$log_area)[2],
+                                              length.out = 101),
+                               surftemp = mean(df_model$surftemp),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         log_area = log_area)
+
+plt_n2oeq_area <-
+  effect_area_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = log_area,
+                      y = n2oeq_pred),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  xlab("") +
+  ylab(expression(paste("Equilibrium N"[2], "O (nmol/L)"))) +
+  theme_pubr() +
+  theme(axis.text.x = element_blank(),
+        legend.position = "none") +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 20), breaks = seq(0, 20, 5)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_fill_brewer(palette = "Greys", guide = "none")
+
+plt_n2o_area_no3 <-
+  effect_area_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = log_area,
+                      y = n2o_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  xlab("") +
+  ylab(expression(paste("Dissolved N"[2], "O (nmol/L)"))) +
+  theme_pubr() +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 55), breaks = seq(0, 50, 10)) +
+  scale_fill_brewer(palette = "Greys", guide = "none") + 
+  theme(axis.text.x = element_blank(),
+        legend.position = "none")
+
+plt_n2o_no3_area <-
+  effect_no3_area %>%
+  ggplot(aes(x = no3_cat, 
+             y = n2o_pred, 
+             group = log_area,
+             color = log_area)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  xlab("") +
+  ylab("") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 55), breaks = seq(0, 50, 10)) +
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        legend.position = "none")
+
+plt_sat_no3_area <-
+  effect_no3_area %>%
+  ggplot(aes(x = no3_cat, 
+             y = sat_pred, 
+             group = log_area,
+             color = log_area)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  xlab(expression(paste("NO"[3], " category"))) +
+  ylab("") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  theme(legend.position = "none",
+        axis.text.y = element_blank())
+
+plt_sat_area_no3 <-
+  effect_area_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = log_area,
+                      y = sat_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  xlab("log(surface area (ha))") +
+  ylab("Saturation ratio") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  scale_fill_brewer(palette = "Greys", guide = "none") +
+  theme(legend.position = "none")
+
+legend1 <- effect_area_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = log_area,
+                      y = n2o_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  theme_void() +
+  scale_fill_brewer(palette = "Greys", guide = "none") + 
+  theme(legend.position = c(0.5, 0.1),
+        legend.direction = "horizontal",
+        legend.text = element_text(size = 10, angle = 45)) +
+  labs(color = expression(paste("NO"[3], " category"))) +
+  guides(color = guide_legend(title.position = "top", title.hjust = 0.5, title.vjust = 0.5))
+
+legend1 <- get_legend(legend1)
+
+legend2 <- 
+  effect_no3_area %>%
+  ggplot(aes(x = no3_cat, 
+             y = n2o_pred, 
+             group = log_area,
+             color = log_area)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  theme_void() +
+  theme(legend.position = c(0.5, 0.5),
+        legend.direction = "horizontal",
+        legend.text = element_text(size = 10, angle = 45)) +
+  labs(color = "log(surface area (ha))") +
+  guides(color = guide_legend(title.position = "top", title.hjust = 0.5, title.vjust = 1))
+
+legend2 <- get_legend(legend2)
+
+legendPlot <- ggarrange(legend1, legend2, ncol = 1)
+
+plt_area_effects <- ggarrange(plt_n2oeq_area,
+                                  legendPlot,
+                                  plt_n2o_area_no3,
+                                  plt_n2o_no3_area,
+                                  plt_sat_area_no3,
+                                  plt_sat_no3_area,
+                             ncol = 2,
+                             nrow = 3,
+                             widths = c(1, 1, 1, 1, 1, 1),
+                             heights = c(1, 1, 1, 1, 1, 1),
+                             labels = c("A", "Legend", "B", "C", "D", "E"),
+                             hjust = c(-6, -1, -6, -6, -6, -6),
+                             vjust = c(2, 2, 2, 2, 2, 2)
+                             )
+
+plt_area_effects <- plt_area_effects %>%
+  annotate_figure(plt_area_effects,
+                  left = text_grob("Posterior mean", 
+                                   size = 18, 
+                                   rot = 90)
+                  )
+# ggsave only works in R console (i.e., copy + paste to save)
+ggsave(filename = "cond_effect_area_no3.tif",
+       plot = plt_area_effects,
+       device = "tiff",
+       width = 180 / 25.4,
+       height = 210 / 25.4,
+       units = "in")
+```
+
+```{r plot_area_effects, fig.align='center', fig.asp=1.166667, fig.width=8, }
+plt_area_effects
+```
+
 # Predict to population
-As previously described, in order to make inferences to the population of interest, the final model above was used to, first, predict surface temperature in the target population, since it depended only on the fully observed covariates.
+
+As mentioned above, in order to make inferences to the population of interest, the final model above was used to, first, predict surface temperature in the target population, since it depended only on the fully observed covariates.
 Next, the predictive distribution for surface temperature was used, along with the relevant fully observed covariates, to predict NO3 in the target population.
@@
                       cores =1, 
-                      ndraws = 500) %>%
+                      ndraws = 2000) %>%
   mutate(surftemp = .prediction)
 
-save(predict_temp, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")
+save(predict_temp, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp_b.rda")
 ```
@@
 NO3 was next predicted using the posterior predictions for surface temperature and the other fully observed covariates.
-Note that the posterior predictive distribution for NO3 was subsampled in order to minimize excess simulations.
+Note that only one posterior draw per posterior predictive draw for surface temperature was used in order to reduce excess simulations.
 
 ```{r parallel_predict_draws, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")
-
 temp_X <- predict_temp %>% # select relevant columns as predictors
   ungroup() %>%
-  select(WSA9,
-         state,
-         size_cat,
-         log_area,
-         .row,
-         .draw,
-         surftemp) %>%
-  select(WSA9, state, size_cat, log_area, surftemp)
+  arrange(.draw) %>%
+  select(.row, .draw, WSA9, state, size_cat, log_area, log_elev, surftemp)
 
@@
 # and register the workers
-cl <- parallel::makeCluster(5) 
+cl <- parallel::makeCluster(25)
 doSNOW::registerDoSNOW(cl)
@@
 # make a progress bar
-pb <- txtProgressBar(max = 1500, style = 3)
+pb <- txtProgressBar(max = 2000, 
+                     style = 3)
 progress <- function(n) setTxtProgressBar(pb, n)
@@
 
-system.time( # approx 26 hrs with 5 workers & 500 draws from PPD
-predict_no3 <- foreach(sub_X = isplitRows(temp_X, chunkSize = 155299), 
+system.time( # approx x hrs with 25 workers & 1 draw from PPD
+predict_no3 <- foreach(sub_X = isplitRows(temp_X, chunkSize = 465897), 
                        .combine = 'c',
@@
                        ) %dopar% {
-                         apply(brms::posterior_predict(n2o_mod6,
+                         brms::posterior_predict(n2o_mod6,
                                                  newdata = sub_X,
@@
                                                  allow_new_levels = T,
-                                                 ndraws = 500,
-                                                 cores = 1), 2, sample, 1)
+                                                 ndraws = 1,
+                                                 cores = 1)
                          }
@@
 
-save(predict_no3, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
-```
-
-Finally, dissolved and equilibrium N2O were predicted using the surface temperature and NO3 predictions along with the survey variables and known covariates.
-Again, the posterior was subsampled in order to reduce excess simulations.
+save(predict_no3, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3_b.rda")
+```
+
+Finally, dissolved and equilibrium N2O were predicted using the surface temperature and NO3 predictions, along with the survey variables and other fully observed covariates.
 
 ```{r n2o_covariates_X, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")
-
-# Assemble dataframe containing relevant covariates (known and predicted)
+# re-load the temp predictions into memory
+load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp_b.rda")
+
+# Assemble dataframe containing relevant covariates (known and predicted) for predicting N2O responses.
 n2o_X <- predict_temp %>%
@@
 
-# clear objects to reduce memory overhead
-rm(predict_no3, predict_temp) 
-gc()
-
-# save the predictors for n2o and n2oeq
-save(n2o_X, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_X.rda")
-```
-
-```{r parallel_predict, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_X.rda")
+rm(predict_temp, predict_no3) # reduce memory
 
@@
 # and register the workers
-cl <- parallel::makeCluster(6) 
+cl <- parallel::makeCluster(25) 
 doSNOW::registerDoSNOW(cl)
@@
 # make a progress bar
-pb <- txtProgressBar(max = 1500, style = 3)
+pb <- txtProgressBar(max = 2000, style = 3)
 progress <- function(n) setTxtProgressBar(pb, n)
@@
 system.time(
-predict_n2o <- foreach(sub_X = isplitRows(n2o_X, chunkSize = 155299),
+predict_n2o <- foreach(sub_X = isplitRows(n2o_X, chunkSize = 465897),
                  .combine = rbind,
@@
                           allow_new_levels = T,
-                          ndraws = 500,
-                          cores = 1),
-        2, sample, 1)
+                          ndraws = 1,
+                          cores = 1), 3, t)
                    }
@@
 
-save(predict_n2o, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_n2o.rda")
+save(predict_n2o, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_n2o_b.rda")
 ```
@@
 ```{r assemble_predictions, eval=FALSE, include=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_n2o.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp.rda")
+load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_temp_b.rda")
+load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_no3_b.rda")
+load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/predict_n2o_b.rda")
 
@@
 
-rm(predict_n2o, predict_temp, predict_no3) # clean up workspace for RAM
-gc()
-
-save(all_predictions, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
+save(all_predictions, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions_b.rda")
 ```
@@
 # Population estimates
+
 A number of estimates for the target population were assembled and presented below.
@@
 ## Posterior predictive distributions
+
 Below, a density plot summarized the posterior predictive distributions across the target population of lakes.
-The PPDs consisted of based on 500 simulations for each variable.
+The summaries of the PPDs were based on 2000 draws.
 Note that the x-axis was truncated at 50 nmol/L for a clearer visualization of the bulk of the predictive distribution.
@@
 
-```{r plot_n2o_posterior_preds, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE, cache.lazy=FALSE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2o_posterior_preds, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE, cache.lazy=FALSE}
 all_predictions %>%
@@
 
-```{r plot_n2oeq_posterior_preds, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2oeq_posterior_preds, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_sat_posterior_preds, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_sat_posterior_preds, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 ## Estimated means
+
 ### National
+
 Below are density plots summarizing the posterior distribution of _means_ for N2O concentrations and the saturation ratio for all US lakes and reservoirs > 4ha in the lower 48 states).
 
-```{r plot_n2o_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2o_nat_posterior_mean, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_n2oeq_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2oeq_nat_posterior_mean, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_sat_nat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_sat_nat_posterior_mean, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_sat_nat_posterior_median, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_sat_nat_posterior_median, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_undersat_posterior_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_undersat_posterior_mean, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 ### Ecoregion
+
 Below are posterior estimates of the means for dissolved and equilibrium N2O and the saturation ratio by WSA9 ecoregion.
 
-```{r plot_n2o_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2o_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_n2oeq_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2oeq_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_sat_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_sat_wsa9_posterior_mean, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-A plot of the posterior estimates for the median saturation ratio below indicated, again, that most lakes in each ecoregion were undersaturated (i.e., median << 1).
-
-```{r plot_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+A plot of the posterior estimates for the median saturation ratio below indicated that most lakes in each ecoregion were undersaturated (i.e., median < 1).
+
+```{r plot_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_prop_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.height=8, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_prop_sat_wsa9_posterior_median, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 ### State
-Comparisons of mean estimates by state are below. Density estimates were not included to minimize the vertical plot space.
-
-```{r plot_state_mean_n2o, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+
+Comparisons of mean estimates by state are below. Density polygons were not included to minimize the vertical plot space.
+
+```{r plot_state_mean_n2o, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_state_mean_n2oeq, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_state_mean_n2oeq, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 A horizontal, dashed, black line is shown at ratio = 1, indicating the boundary for under- _vs._ oversaturation.
-Only a few states (e.g., NV, DE) had median estimates that were 1 or greater, suggesting that, for most states, most lakes were undersaturated.
-
-```{r plot_state_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+Only a few states (e.g., NV, DE) had median estimates that were 1 or greater, suggesting that, for most states, most lakes were estimated to be undersaturated.
+
+```{r plot_state_mean_median_sat, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_state_prop_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_state_prop_sat, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 ### Size category
+
 The estimated means and medians by size category are below for dissolved and equilibrium N2O and the saturation ratio.
 
-```{r plot_n2o_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2o_size_posterior_mean, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_n2oeq_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_n2oeq_size_posterior_mean, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_sat_size_posterior_mean, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_sat_size_posterior_mean, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_sat_size_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_sat_size_posterior_median, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_size_cat_mean_median_sat, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_size_cat_mean_median_sat, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_prop_sat_size_cat_posterior_median, echo=FALSE, fig.align='center', fig.height=6, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_prop_sat_size_cat_posterior_median, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 ## Model- _vs._ design-based
+
 Below, estimates from the model-based approach are compared to previously calculated, design-based estimates.
@@
 ### Dissolved N2O
+
 Below, National mean estimates for dissolved N2O from the model and design-based approaches were compared.
 The sample-based estimate was also included as a reference.
-The black, vertical, dashed line represents the mean of the sample.
+The black, vertical, dashed line indicates the mean of the sample.
 
 ```{r n2o_means_national, message=FALSE, cache=TRUE}
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
 all_predictions %>%
@@
 
-```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
-
+```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 ```{r n2o_mean_wsa9, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
-
 all_predictions %>%
@@
 
-```{r plot_mean_n2o_wsa9, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda")
-
+```{r plot_mean_n2o_wsa9, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r table_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda")
-
+```{r table_size_mean_n2o, echo=FALSE,message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_size_mean_n2o, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda")
-
+```{r plot_size_mean_n2o, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 ### Saturation
+
 Below, the same comparisons were made for the saturation estimates.
 
-```{r table_nat_sat_mean, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=8, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
-
+```{r table_nat_sat_mean, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_nat_sat_mean, echo=FALSE, fig.align='center', fig.height=2, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
-
+```{r plot_nat_sat_mean, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_wsa9_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda")
-
+```{r plot_wsa9_sat_mean, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r table_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests_size.rda")
-
+```{r table_size_sat_mean, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
 
-```{r plot_size_sat_mean, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE, cache=TRUE}
-#load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions.rda")
-
+```{r plot_size_sat_mean, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 all_predictions %>%
@@
   scale_colour_manual(values = c("black", "grey")) +
+  geom_hline(yintercept = 1, color = "lightgrey") +
   ylab("mean N2O saturation ratio") +

### 2023-10-18 manuscript/manuscript_file.Rmd
10/18/2023 manuscript revision

@@
 -   Letters vs article format at Nature Geosciences
+    + Letters no longer supported.  Of all the options, Article is best suited to us.
+    + 2200 – 3000 words (excluding abstract, Methods, references and figure legends)
+    + Pseudo-abstract, 200 words, no citations
+    + Subheadings are OK
+    + Instructions to authors list separate Results and Discussion, but article in current issue do not follow this format.
+    + Figures: article in current issue have 2-5 figures, mostly multi-panel.  Instructions specify 4-6 figures + tables.
 -   Add quantitative NO3 values to Fig.3B and Fig. 5B legend.
--   nitrate in methods
 -   Lester Luan and Jana Compton for internal tech reviews.
@@
 
-Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020] have called the accuracy of emission estimates into question [@Lauerwald_etal_2019].
-Here we measured dissolved N~2~O concentrations and calculated emission rates for `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies distributed across the nine major ecoregions of the conterminous U.S.
+**Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020] have called the accuracy of emission estimates into question [@Lauerwald_etal_2019].
+Here we measured N~2~O concentrations and calculated emission rates for `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies distributed across the nine major ecoregions of the conterminous U.S.
 (CONUS).
 We use a Bayesian hierarchical approach to model the data and predicted N~2~O in`r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull()` CONUS waterbodies.
-We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean` % of the waterbodies functioning as sinks for atmospheric N~2~O.
-Despite widespread N~2~O uptake by CONUS waterbodies, a small proportion of the waterbodies in each ecoregion had sufficiently high N~2~O emission rates to offset N~2~O uptake in the majority of waterbodies, and none of the ecoregions are estimated to functione as net N~2~O sinks.
+We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean` % (95% credible interval:`r propN2oSinkLower` - `r propN2oSinkUpper`% ) of waterbodies functioning as sinks for atmospheric N~2~O.
+Despite widespread N~2~O uptake by CONUS waterbodies, a small proportion of the waterbodies in each ecoregion had sufficiently high N~2~O emission rates to offset N~2~O uptake in the majority of waterbodies, and none of the ecoregions are estimated to function as net N~2~O sinks.
 The model predicts that waterbodies with high nitrate (NO~3~) tend to be sources of N~2~O whereas those with low NO~3~ are generally sinks.
@@
 Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS waterbodies is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` ) metric tons N~2~O year^-1^.
-The mean estimate is a net emission of  `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^, which is small relative to other known sources of N2O in the CONUS (e.g., wastewater treatment, manure management) and equivalent to only `r (national_stats %>% select(UCL_fn2o) %>% pull() / 22767) * 100`% of N~2~O emissions from all CONUS surface waters (e.g.
+The mean estimate is a net emission of `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^, which is small relative to other known sources of N2O in the CONUS (e.g., wastewater treatment, manure management) and equivalent to only `r (national_stats %>% select(UCL_fn2o) %>% pull() / 22767) * 100 %>% round(., 1)`% of N~2~O emissions from all CONUS surface waters (e.g.
 zero order streams through river mouths) per IPCC estimates .
 Emission estimates were most uncertain for large waterbodies, suggesting that efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nation’s largest lakes and reservoirs.
-Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS waterbodies are a net source of sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source [@epa_2023].
+Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS waterbodies are a net source or sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source.**
 
@@
 nitrogen loading).
-Most work on aquatic N~2~O biogeochemistry has focused on streams and rivers, but these systems have been consistently found to be N~2~O sources [@Beaulieu_etal_2008; @Beaulieu_etal_2010; @Quick_etal_2019] and the controls on N~2~O biogeochmistry likely differ between flowing and standing waters.
+Nitrous oxide dynamics have been investigated in streams and rivers spanning a broad range of environemtal conditions, but these systems have been consistently found to be N~2~O sources [@Beaulieu_etal_2008; @Beaulieu_etal_2010; @Quick_etal_2019] and the controls on N~2~O biogeochmistry likely differ between flowing and standing waters.
 These limitations in data availability have complicated efforts to incorporate N~2~O consumption into freshwater N~2~O budgets.
@@
 In this study we present the first national-scale survey of N~2~O concentrations and emissions from CONUS waterbodies.
-A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week[@EPAdesign] **[(Fig.
+A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that were at least 1 hectare in area, at least 1 meter deep, had at least 0.1 hectare of open water, and had a minimum residence time of one week[@EPAdesign] **[(Fig.
 1)]**.
@@
 1)]**.
-We modeled the data using a Bayesian hierarchical approach and predicted the dissolved N~2~O concentration for all 465,897 CONUS waterbodies in the target population.
-At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`) **[(Table 1)]**.
+These data, along with the survey design information, were used to develop a Bayesian hierarchical model to predict N~2~O concentrations for all 465,897 CONUS waterbodies in the target population.
+At the population level, the majority of CONUS waterbodies (mean = `r propN2oSinkMean`%, 95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) were predicted to be a net sink for N~2~O.
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), although the 95% credible intervals overlap across all ecoregions **[(Table 1)]**.
 
@@
 
-The extent of N~2~O undersaturation in CONUS waterbodies is surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
-This finding runs counter to the long standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lentic waterbodies can function as an N~2~O sink.
+The estimated extent of N~2~O undersaturation in CONUS waterbodies was surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
+This finding runs counter to the long-standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lentic waterbodies can function as N~2~O sinks.
 Webb et al.
@@
 While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
-This pattern holds across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use across the continental U.S..
+This pattern held across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use.
 
@@
 
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean N~2~O saturation ratio (ratio of measured to equilibrium concentration) was equal to or greater than 1 in seven of the nine ecoregions **[(Table 1, Fig.
-2)]**, suggesting that, on average, waterbodies in these ecoregions were supersatured with N~2~O and were therefore functioning as a source of N~2~O.
-The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations that exceed equilibrium in most ecoregions can be attributed to the distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies **[(Fig.
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean predicted N~2~O saturation ratio (ratio of measured to equilibrium concentration) was statistically indistinguishable from one in all ecoregions **[(Table 1, Fig.
 2)]**.
-Across the CONUS, and within each of the nine major ecoregions, the distribution of N~2~O saturation ratios had median values that were less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
-Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources, particularly at high NO~3~ at water temperature **[Fig.
+The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations close to equilibrium in all ecoregions can be attributed to the predicted distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies **[(Fig.
+2)]**.
+Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had median values less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
+Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources, particularly at high NO~3~ at high water temperature **[Fig.
 3]**.
@@
 ![Figure 2.
-Distribution of N~2~O saturation values for the population of waterbodies in the nine major ecoregions.
-Each blue line (n=2000) represents the distribution of values from a single realization of the posterior prediction.
-Vertical black dashed and solid lines indicate median and mean, respectively.
+Posterior predictive distribution of N~2~O saturation values for waterbodies in the nine major ecoregions.
+Each blue line (n=2000) represents the distribution of values from a single realization of the posterior predictive distribution.
+Vertical black dashed and solid lines indicate point estimates of the median and mean, respectively (Table 1).
 Vertical red line indicates a saturation ratio of 1.
@@
 Nitrification, an aerobic process which can produce N~2~O as a byproduct, is the microbial oxidation of ammonium (NH~4~^+^) to NO~3~ [@Kuypers_etal_2018].
-Nitrification might be expected to be more prevalent in the predominantly well oxygenated surface waters where the N~2~O samples were collected than denitrification, which is an anaeorbic process.
+Nitrification may be more prevalent in the predominantly well oxygenated surface waters where the N~2~O samples were collected because denitrification is an anaeorbic process.
 However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column [@Broman_etal_2021] and can be transported from aoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
@@
 ![Figure 3.
-Conditional effects plot illustrating A) the relationship between N~2~O saturation and nitrate (NO~3~) concentration, dependent upon surface water temperature.
+Predicted interactive effects of surface water temperature and NO~3~ category the N~2~O saturation ratio.
+A) The relationship between N~2~O saturation and nitrate (NO~3~) concentration, dependent upon surface water temperature.
 The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13, and B) relationship between N~2~O saturation and surface water temperature, dependent upon NO~3~ category.
@@
 
-N~2~O saturation also increased with water temperature **[Fig.
-3]**, which is consistent with N~2~O production and consumption occurring through temperature sensitive biological processes, but the relationship was dependent on NO~3~ availability.
-At low NO~3~ concentrations, temperature had a relatively minor affect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
-The effect of temperature became more pronounced when NO~3~ availability was greater, likely reflecting the stimulatory affect of higher temperature on biological N~2~O production.
-The relative effect of NO~3~ was greater than that of temperature, thus elevated water temperature serves to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
+The model detected a pattern of increasing N~2~O saturation with water temperature **[Fig.
+3]**, which is consistent with nitrogen transformations occurring through temperature sensitive biological processes, but the relationship was dependent on NO~3~ concentration.
+At low NO~3~ concentrations, temperature was estimated to have a relatively minor effect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
+The effect of temperature was more pronounced at greater NO~3~ concentrations, potentially reflecting the stimulatory effect of higher temperature on biological N~2~O production.
+The relative effect of NO~3~ was estimated to be greater than that of temperature, thus elevated water temperature may serve to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
 
@@
 
-While NO~3~ and water temperature were the primary drivers of N~2~O production or consumption rates, dissolved N~2~O concentration was also related to waterbody size.
-When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreases with increasing waterbody size class **[Fig.
-4A]**.
-This pattern holds regardless of whether N~2~O is supersaturated or undersatured **[Fig.
-4B, C]**.
+While NO~3~ and water temperature were associated with the N~2~O saturation ratio, N~2~O was also related to waterbody size.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreased with increasing waterbody size class **[Fig.
+4]**.
+This pattern holds regardless of whether N~2~O was supersaturated or undersatured.
 Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
@@
 Air-water gas exchange will vent excess N~2~O from surface waters or facilitate the invasion of air that can replenish depleted N~2~O in undersaturated waterbodies.
-Regardless of whether a waterbody is superaturated or undersatured in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
+Regardless of whether a waterbody is supersaturated or undersaturated in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
 One of the primary drivers of air-water gas exchange in lentic waters is wind speed [@Vachon_etal_2013].
 Wind speeds tend to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
-Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the affect of biogeochemistry on dissolved N~2~O is more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
+Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the effect of biogeochemistry on dissolved N~2~O may be more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
 While previous surveys have documented relationships between lake size and dissolved methane concentration [@Pighini_etal_2018; @Juutinen_etal_2009; @Borges_etal_2011], to the best of our knowledge this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
@@
 ![Figure 4.
-A) Relationship between $\Delta$ N~2~O (difference between observed and equilibrium N~2~O concentration) and waterbody surface areas.
-B,C) relationship between N~2~O saturation and waterbody size for the Coastal Plains and Northern Plains ecoregions.
-The vertical blue line references a N~2~O saturation value of 1.0.
-The horizontal red arrows highlight $\Delta$ N~2~O for the smallest lake size class.
-Mean (points) and 95% credible interview (horizontal bars) are derived from 2000 realizations of the poterior predictive distribution.](manuscript_figures/deltaN2ObySize.tiff)
+Relationship between $\Delta$ N~2~O (difference between predicted dissolved and equilibrium N~2~O concentration) and waterbody surface areas.
+Mean (points) and 95% credible interview (horizontal bars) are derived from 2000 realizations of the posterior predictive distribution.](manuscript_figures/deltaN2ObySize.tiff)
 
-The influence of waterbody size is also evident in the statistical interaction between NO~3~ and surface area **[Fig.
+The influence of waterbody size is also suggested in the interaction between NO~3~ and surface area **[Fig.
 5]**.
@@
 5A]**.
-At low NO~3~ concentrations, N~2~O is undersatured in small waterbodies, but increases toward equilibrium in larger waterbodies **[Fig.
+At low NO~3~ concentrations, N~2~O is undersaturated in small waterbodies, but increases toward equilibrium in larger waterbodies **[Fig.
 5A]**.
@@
 
-The model indicates that NO~3~ exerts a strong control on N~2~O, but the relationship is complicated by interactions with water temperature and lake surface area.
-This finding mirrors reports from a survey of farm ponds in Canada where N~2~O was found to be partially controlled by dissolved inorganic nitrogen (DIN) concentration, but the effect of DIN was conditional upon and thermal stratification strength.
+The model results suggest that NO~3~ may exert a strong control on N~2~O, but the relationship is complicated by interactions with water temperature and lake surface area.
+This finding mirrors reports from a survey of farm ponds in Canada where N~2~O was found to be associated with dissolved inorganic nitrogen (DIN) concentration, but the effect of DIN was conditional upon thermal stratification strength.
 These complex interactions may explain contrasting literature reports regarding the role of NO~3~ in regulating N~2~O.
@@
 
-Interactions between NO~3~, water temperature, and lake size produced a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enable the identifcation of broad scale patterns.
-First, small waterbodies with low NO~3~ concentration tend to be undersaturated, whereas small waterbodies with high NO~3~ tend to be highly supersaturated **[Fig.
+Interactions between NO~3~, water temperature, and lake size produced a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enabled identifcation of broad scale patterns.
+First, small waterbodies with low NO~3~ concentration tended to be undersaturated, whereas small waterbodies with high NO~3~ tended to be highly supersaturated **[Fig.
 5]**.
-Large waterbodies, regardless of NO~3~ status, tend to have N~2~O saturation ratios closer to equilibrium.
-Therefore intra and inter regional patterns in N~2~O saturation can be well predicted by the joint distributions of waterbody size, NO~3~ availability, and water temperature.
+Large waterbodies, regardless of NO~3~ status, tended to have N~2~O saturation ratios closer to equilibrium.
+Therefore intra and inter regional patterns in N~2~O saturation may be well predicted by the joint distributions of waterbody size, NO~3~ availability, and water temperature.
 
@@
 
-The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which is expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
+The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
 Variability and magnitude of predicted N~2~O emission rates decreased with increasing waterbody size **[(Fig.
@@
 5A,B)]**.
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **[(Table 1)]**.
-This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeds the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **[(SI Fig.
+Despite predicted widespread N~2~O undersaturation in CONUS waterbodies, the estimated national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **[(Table 1)]**.
+This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeded the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **[(SI Fig.
 1)]**.
@@
 C) N~2~O flux plotted against the five lake size categories used in the survey design.
-Across all plots, points and bars reflect the mean (points) and 95% credible interval of the 2000 realizations of the posterior predictive distribution.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
+Across all plots, points and bars reflect the mean (points) and 95% credible interval, respectively, of the 2000 realizations of the posterior predictive distribution.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
 
@@
 Total N~2~O flux from the surface of CONUS waterbodies is a function of N~2~O emission rates and waterbody surface areas.
-The four smallest size classes (\< 50 Ha) in the survey design contained `r round(proportionSmallNumber, 2) * 100`% of all lakes in the population, but contributed only `r round((FluxBySize %>% filter(size_cat != "50_max") %>% select(estimate) %>% summarise(tot_small = sum(estimate)) %>% pull())/(FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) * 100, 0)`% to the N~2~O flux from CONUS waterbodies **[(Fig.
+The four smallest size classes (\< 50 Ha) in the survey design contained `r round(proportionSmallNumber, 2) * 100`% of all lakes in the population, but were estimated to have contributed only `r round((FluxBySize %>% filter(size_cat != "50_max") %>% select(estimate) %>% summarise(tot_small = sum(estimate)) %>% pull())/(FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) * 100, 0)`% to the N~2~O flux from CONUS waterbodies **[(Fig.
 6C)]**.
@@
 First, lakes in the smaller size classes constituted only `r round(proportionSmallSA, 2) * 100`% of total waterbody surface area, thereby limiting their N~2~O flux.
-Secondly, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the population **[(Fig.
+Second, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the population **[(Fig.
 6A)].** Waterbodies in the largest size class (\> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
-This lake size class had the greatest N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year\^{-1}), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year\^{-1} **[(Fig.
+This lake size class had the greatest N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year^-1^), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year^-1^ **[(Fig.
 6C)]**.
 While uncertainty in N~2~O emission rates was smallest for large lakes **(Fig.
-6A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large waterbodies.
+6A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large waterbodies **(Fig.
+6B,C)**.
 
-Despite widespread N~2~O undersaturation, only two ecoregions had flux estimates less than zero (Xeric and Coastal Plains), but the credible intervals for those ecoregions were broad and include zero **(Table 1)**.
-Only the Temperate Plains, Southern Plains, and Southern Appalachians had flux estimate credible intervals that did not overlap with zero and these ecoregions functioned as net N~2~O sources.
-Overall, flux estimates overlapped across all nine ecoregions, which is somewhat surprising given the large differences in environmental conditions across the ecoregions.
+Despite widespread N~2~O undersaturation, the credible interval for the predicted fluxes overlapped with zero for all ecoregions but the Temperate Plains, Southern Plains, and Southern Appalachians ecoregions, and the credible intervals were positive for these ecoregions **(Table 1)**, suggesting they were more likely to have functioned as net N~2~O sources.
+Overall, the credible intervals of the predicted mean flux overlapped across all nine ecoregions, which is somewhat surprising given the large differences in environmental conditions across the ecoregions.
 
-At the national scale, we estimate that CONUS waterbodies are a source of N~2~O, emitting `r table1_dat %>% janitor::clean_names() %>% filter(spatial_domain == "national") %>% select(n2o_flux) %>% pull()` metric tons N~2~O year^-1^; however, this estimate is not well constrained and the 95% credible interval includes zero, thus we cannot conclude with statistical confidence whether CONUS waterbodies are a net source of sink of N~2~O.
-While we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change estimates that CONUS drainage networks, inclusive of zero order streams through major river mouths, receiving run off from croplands and grasslands emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
-Emissions from lakes and reservoirs represent only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of total surface water emissions reported by the IPCC [@epa_2023], with the balance emanating from streams and rivers.
-Emissions from CONUS lakes and reservoirs are also small when compared to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
-Whether CONUS waterbodies are a net source or sink of N~2~O, their contribution to the US N~2~O budget is small.
+The mean flux estimate for CONUS waterbodies was positive (`r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval was quite broad and included zero (`r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^) so we could not conclude with statistical confidence whether CONUS waterbodies were a net source or sink of N~2~O overall.
+Though we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change (IPCC) estimated that CONUS drainage networks, inclusive of zero order streams through major river mouths, emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
+The IPCC estimates total N~2~O emissions from aquatic ecosystems as a fraction (emission factor) of the anthropogenic nitrogen that leaches from soils into surface water systems and does not distinguish between components of the aquatic continuum, thus a direct comparison between the IPCC estimate and the value reported here is not possible; however, the mean N~2~O  flux from lakes and reservoirs reported here comprises only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of the IPCC estimate, suggesting the IPCC estimate is too high, possibly because the IPCC model does not allow for N~2~O consumption in lakes, ponds, and reservoirs.
+This limitation extends to most approaches used to estimate aquatic N~2~O fluxes [@Lauerwald_etal_2023], inlcuding more recent mechanistic models [@Lauerwald_etal_2019].
+Given our estimate of widespread N~2~O consumption in CONUS waterbodies, failure to include this dynamic in aquatic N~2~O models may yield biased emission estimates.
 
-Our annual flux estimates assume that emission rate estimates based on samples collected during the summer reflect annual conditions.
+Not only is our estimate of N~2~O emissions from CONUS lakes and reservoirs small when compared to the IPCC estimate, it is also small relative to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
+Whether CONUS waterbodies were a net source or sink of N~2~O, their contribution to the US N~2~O budget was likely small by comparison.
+
+Our annual flux estimates are derived from a model of single samples collected from individual waterbodies during the summer.
+Our approach therefore implicitly assumes that our summertime observations are reflective of annual conditions.
 The literature contains strong evidence of seasonality in N~2~O emission rates, but the patterns differ among studies.
 Some systems have maximum N~2~O emission rates during the winter [@Musenze_etal_2014; @Kortelainen_etal_2020], others in the Fall [@Beaulieu_etal_2014], and yet others during the summer [@Beaulieu_etal_2010].
-While our assumption of constant emissions across the year adds uncertainty to the annual flux estimate, it is not clear that the estimate is biased high or low.
+While our assumption of constant emissions across the year adds uncertainty to our annual flux estimate, it is difficult to determine, based on the literature, if this assumption has biased our estimates.
 
@@
 
-Our survey is the first to demonstrate that most lakes and reservoirs in the CONUS function as N~2~O sinks during the summer.
-Despite widespread N~2~O uptake, CONUS waterbodies function as a net source of N~2~O due to high emissions from from a subset of the population.
-Dissolved N~2~O in lakes and reservoirs can be predicted from non-linear interactions between water temperature, NO~3~ availability, and waterbody size.
-NO~3~ availability largely determines whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink is exaggerated with increasing water temperature.
+Our research is the first to suggest that most lakes and reservoirs in the CONUS function as N~2~O sinks during the summer.
+Our observations of dissolved N~2~O were predicted reasonably well using non-linear interactions between observed water temperature, NO~3~ concentration, and waterbody size (see SI).
+The model form and estimated parameters suggest that NO~3~ concentration plays a role in determining whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink is exaggerated with increasing water temperature.
 The degree of N~2~O disequilibrium also increases with decreasing waterbody size, possibly because smaller waterbodies have low air-water gas exchange rates.
+Our observations of dissolved N~2~O was predicted  due to high emissions from from a subset of the population.
 
 Although waterbodies in the largest lake size class tend to have N~2~O levels close to equilibrium, they emit more N~2~O than the other size classes due to their expansive surface area.
-The N~2~O flux from these waterbodies is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved N~2~O translate to large uncertainties in flux when propagated across the surface of large waterbodies.
-While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nations largest lakes including an assesment of intra-waterbody spatial variation in N~2~O, which can be important [@Salk_etal_2019, @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propogate to large uncertainties in flux estimates.
+The N~2~O flux from these waterbodies is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved and equilibrium N~2~O concentrations translate to large uncertainties in flux when propagated across the surface of large waterbodies.
+Uncertainty in the national flux estimate was too large to determine if CONUS waterbodies functioned as a net source or sink of N~2~O.
+While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nations largest lakes including an assesment of intra-waterbody spatial variation in N~2~O, which can be important [@Salk_etal_2019, @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propagate to large uncertainties in flux estimates.
 
-The smallest lake size classes exhibited both the highest and lowest levels of N~2~O saturation, depending on NO~3~ availability.
-It is unclear whether this variability is related to differences in NO~3~ loading or is an artifact of temporal variation in NO~3~ processing and N~2~O production/consumption.
+The smallest lake size classes were estimated to have both the highest and lowest levels of N~2~O saturation, depending largely on NO~3~ availability.
+It is unclear whether this variability is related to differences in NO~3~ loading or is an artifact of temporal variation in NO~3~ processing.
 For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted [@Beaulieu_etal_2014].
-Most small waterbodies in the CONUS function as N~2~O sinks during the summer, but a small subset are N~2~O emission hotspots.
+Most small waterbodies in the CONUS were estimated to have functioned as N~2~O sinks during the summer, but a small subset were also N~2~O emission hotspots.
+It is unknown if the high emitting waterbodies maintain those emissions throughout the summer, or if they convert to sinks later in the season.
 A better understanding of the temporal evolution of N~2~O production and consumption is an important research need to determine the extent to which small waterbodies can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.

### 2023-10-18 manuscript/manuscript_support.R

@@ -9,12 +9,21 @@ localPath <- Sys.getenv("USERPROFILE")
 
 ## load libraries
 # DMAP not under renv control.  Should probably specify same
-# versions as used for renv library
+# versions as used for renv library, but causing issues.  Just 
+# install newest version.
 if(localPath == "") { # if DMAP, then
-install.packages(c("sf", "tidyverse", "janitor",
-                   "tictoc", "ggpubr", "ggallin",
-                   "devtools"))
-devtools::install_version("USAboundaries", "0.4.0")
+  # require returns FALSE and gives a warning 
+  # (rather than an error as library() does by default) 
+  # if the package does not exist.
+  if(!require("ggallin")) install.packages("ggallin")
+  if(!require("sf")) install.packages("sf")
+  if(!require("tidyverse")) install.packages("tidyverse")
+  if(!require("janitor")) install.packages("janitor")
+  if(!require("tictoc")) install.packages("tictoc")
+  if(!require("pbkrtest")) install.packages("pbkrtest") 
+  if(!require("ggpubr")) install.packages("ggpubr")
+  if(!require("ggallin")) install.packages("ggallin")
+  if(!require("USAboundaries")) install.packages("USAboundaries")
 }
 
 library(sf) # spatial data
@@ -553,12 +562,6 @@ ggsave("manuscript/manuscript_figures/SIfigure1.tiff", width = 8.5, height = 5)
 undersatN2oPercent <- round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% 
                          {nrow(.)} / dg %>% distinct(site.id) %>% 
                          {nrow(.)}) * 100), 1)
-
-propN2oSinkMean <- all_predictions %>% 
-  group_by(.draw) %>% 
-  summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% 
-  summarise(estimate = round(median(prop_sat), 3) * 100) %>% 
-  pull()
 
 propN2oSinkMean <- all_predictions %>% 
   group_by(.draw) %>% 

### 2023-11-29 manuscript/manuscript_file.Rmd
Most recent updates to manuscript

@@
 
-## NOTES
-
--   incorporate in Methods section how/why size is incorporated as categorical variable and continuos variable.
--   Citation "epa_data" isn't populating citation 11 in bibliography. The bibTex entry is classified as a "website". Roy, can you attempt to make this citation play nice with nature_geosciences.csl?
--   Letters vs article format at Nature Geosciences
-    + Letters no longer supported.  Of all the options, Article is best suited to us.
-    + 2200 – 3000 words (excluding abstract, Methods, references and figure legends)
-    + Pseudo-abstract, 200 words, no citations
-    + Subheadings are OK
-    + Instructions to authors list separate Results and Discussion, but article in current issue do not follow this format.
-    + Figures: article in current issue have 2-5 figures, mostly multi-panel.  Instructions specify 4-6 figures + tables.
--   Add quantitative NO3 values to Fig.3B and Fig. 5B legend.
--   Lester Luan and Jana Compton for internal tech reviews.
-
 ## Psuedo abstract
 
-**Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but the relative lack of published N~2~O emission rates and recent evidence of N~2~O uptake across a broad range of waterbodies[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020] have called the accuracy of emission estimates into question [@Lauerwald_etal_2019].
-Here we measured N~2~O concentrations and calculated emission rates for `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies distributed across the nine major ecoregions of the conterminous U.S.
+**Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but recent evidence of N~2~O uptake across a broad range of waterbodies have called the accuracy of emission estimates into question.
+Here we use a new national-scale dataset on dissolved N2O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull()` waterbodies in the conterminous U.S.
 (CONUS).
-We use a Bayesian hierarchical approach to model the data and predicted N~2~O in`r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull()` CONUS waterbodies.
-We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean` % (95% credible interval:`r propN2oSinkLower` - `r propN2oSinkUpper`% ) of waterbodies functioning as sinks for atmospheric N~2~O.
-Despite widespread N~2~O uptake by CONUS waterbodies, a small proportion of the waterbodies in each ecoregion had sufficiently high N~2~O emission rates to offset N~2~O uptake in the majority of waterbodies, and none of the ecoregions are estimated to function as net N~2~O sinks.
-The model predicts that waterbodies with high nitrate (NO~3~) tend to be sources of N~2~O whereas those with low NO~3~ are generally sinks.
-Nitrate interacted with waterbody size, however, with large waterbodies having N~2~O concentrations closer to equilibrium than small waterbodies.
+We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean` % (95% credible interval:`r propN2oSinkLower` - `r propN2oSinkUpper`% ) of waterbodies functioning as N~2~O sinks.
+The model predicts dissolved N2O concentrations reasonably well based partly on interactions between nitrate (NO~3~) concentration, waterbody size, and water temperature.
 Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS waterbodies is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` ) metric tons N~2~O year^-1^.
-The mean estimate is a net emission of `r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^, which is small relative to other known sources of N2O in the CONUS (e.g., wastewater treatment, manure management) and equivalent to only `r (national_stats %>% select(UCL_fn2o) %>% pull() / 22767) * 100 %>% round(., 1)`% of N~2~O emissions from all CONUS surface waters (e.g.
-zero order streams through river mouths) per IPCC estimates .
-Emission estimates were most uncertain for large waterbodies, suggesting that efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nation’s largest lakes and reservoirs.
 Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS waterbodies are a net source or sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source.**
@@
 nitrogen loading).
-Nitrous oxide dynamics have been investigated in streams and rivers spanning a broad range of environemtal conditions, but these systems have been consistently found to be N~2~O sources [@Beaulieu_etal_2008; @Beaulieu_etal_2010; @Quick_etal_2019] and the controls on N~2~O biogeochmistry likely differ between flowing and standing waters.
+Nitrous oxide dynamics have been investigated in streams and rivers spanning a broad range of environmental conditions, but these systems have been consistently found to be N~2~O sources [@Beaulieu_etal_2008; @Beaulieu_etal_2010; @Quick_etal_2019] and the controls on N~2~O biogeochemistry likely differ between flowing and standing waters.
 These limitations in data availability have complicated efforts to incorporate N~2~O consumption into freshwater N~2~O budgets.
@@
 In this study we present the first national-scale survey of N~2~O concentrations and emissions from CONUS waterbodies.
-A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that were at least 1 hectare in area, at least 1 meter deep, had at least 0.1 hectare of open water, and had a minimum residence time of one week[@EPAdesign] **[(Fig.
-1)]**.
+A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that were at least 1 hectare in area, at least 1 meter deep, had at least 0.1 hectare of open water, and had a minimum residence time of one week[@EPAdesign] **(Fig.
+1)**.
 Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate [@epa_data].
-`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **[(Fig.
-1)]**.
+`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
+1)**.
 These data, along with the survey design information, were used to develop a Bayesian hierarchical model to predict N~2~O concentrations for all 465,897 CONUS waterbodies in the target population.
 At the population level, the majority of CONUS waterbodies (mean = `r propN2oSinkMean`%, 95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) were predicted to be a net sink for N~2~O.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), although the 95% credible intervals overlap across all ecoregions **[(Table 1)]**.
+The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), although the 95% credible intervals overlap across all ecoregions **(Table 1)**.
 
@@
 
-##### NO~3~
+The model results suggest that NO~3~ may exert a strong control on N~2~O, but the relationship is complicated by interactions with water temperature and lake surface area **(Fig.
+3)**.
+This finding mirrors reports from a survey of farm ponds in Canada where N~2~O was found to be associated with dissolved inorganic nitrogen (DIN) concentration, but the effect of DIN was conditional upon thermal stratification strength[@Webb_etal_2019].
+These complex interactions may explain contrasting literature reports regarding the role of NO~3~ in regulating N~2~O.
+Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020; @Yang_etal_2015] while others have found no relationship [@Soued_etal_2016; @Whitfield_etal_2011].
+It is possible that the NO~3~ \~ N~2~O relationship wasn't detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
 
-Dissolved N~2~O concentration increased with NO~3~ concentration **[Fig.
-3]**, likely because NO~3~ concentration partially controls denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Kuypers_etal_2018; @Richardson_etal_2009].
+The influence of NO~3~ levels on N~2~O saturation **(Fig.
+3)** may reflect the relationship between NO~3~ levels and denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Kuypers_etal_2018; @Richardson_etal_2009].
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
@@
 When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~, thereby functioning as an N~2~O sink.
-N~2~O consumption via denitrification likely explains the N~2~O undersaturation observed at low NO~3~ concentrations, whereas denitrification was likely a source of N~2~O at higher NO~3~ concentrations.
+N~2~O consumption via denitrification may explain the N~2~O undersaturation predicted at low NO~3~ levels, whereas denitrification may be a source of N~2~O at higher NO~3~ levels.
 
@@
 Nitrification, an aerobic process which can produce N~2~O as a byproduct, is the microbial oxidation of ammonium (NH~4~^+^) to NO~3~ [@Kuypers_etal_2018].
-Nitrification may be more prevalent in the predominantly well oxygenated surface waters where the N~2~O samples were collected because denitrification is an anaeorbic process.
+Nitrification may be more prevalent than denitrification in the predominantly well oxygenated surface waters where the N~2~O samples were collected because denitrification is an anaerobic process.
 However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column [@Broman_etal_2021] and can be transported from aoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
@@
 ![Figure 3.
-Predicted interactive effects of surface water temperature and NO~3~ category the N~2~O saturation ratio.
-A) The relationship between N~2~O saturation and nitrate (NO~3~) concentration, dependent upon surface water temperature.
-The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13, and B) relationship between N~2~O saturation and surface water temperature, dependent upon NO~3~ category.
-Horizontal dashed lines in both panels reflects an N~2~O saturation value of 1.](manuscript_figures/sat_cond_surftemp.tiff)
+Estimated interactive effects of surface water temperature (A and B) and waterbody size (B and D) with NO~3~ levels on the N~2~O saturation ratio.
+A) The relationship between N~2~O saturation and surface water temperature, dependent upon nitrate (NO~3~) level.
+The five, ordered, categorical levels for NO~3~ correspond to concentration ($\mu$g L^-1^) ranges in the observed data of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13.
+B) The relationship between N~2~O saturation and NO~3~ level conditional on surface water temperature.
+The NO~3~ effect was estimated at 5 quantiles of the surface temperatures observed (10%, 25%, 50%, 75%, 90%).
+C) The relationship between N~2~O saturation and waterbody surface area (log-transformed hectares) conditional on NO~3~ level.
+D) The relationship between N~2~O saturation and NO~3~ level conditional on (log) waterbody surface area (5 quantiles).
+Horizontal dashed lines in each panel reflect an N~2~O saturation value of 1.](manuscript_figures/cond_effect_no3.tiff)
 
-##### Temperature
-
-The model detected a pattern of increasing N~2~O saturation with water temperature **[Fig.
-3]**, which is consistent with nitrogen transformations occurring through temperature sensitive biological processes, but the relationship was dependent on NO~3~ concentration.
-At low NO~3~ concentrations, temperature was estimated to have a relatively minor effect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
-The effect of temperature was more pronounced at greater NO~3~ concentrations, potentially reflecting the stimulatory effect of higher temperature on biological N~2~O production.
+The estimated pattern of increasing N~2~O saturation with water temperature **(Fig.
+3)** is  consistent with nitrogen transformations occurring through temperature sensitive biological processes, but the relationship was dependent on NO~3~ levels.
+At low NO~3~ levels, temperature was estimated to have a relatively minor effect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
+The effect of temperature was more pronounced at greater NO~3~ levels, potentially reflecting the stimulatory effect of higher temperature on biological N~2~O production.
 The relative effect of NO~3~ was estimated to be greater than that of temperature, thus elevated water temperature may serve to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
 
-##### Waterbody size
-
-While NO~3~ and water temperature were associated with the N~2~O saturation ratio, N~2~O was also related to waterbody size.
+Nitrous oxide saturation ratio was also related to waterbody size **(Fig.
+3)**.
+Small waterbodies with low NO~3~ levels were predicted to be undersaturated, whereas small waterbodies with high NO~3~ were predicted to be highly supersaturated.
+Large waterbodies, regardless of NO~3~ status, were predicted to have N~2~O saturation ratios closer to equilibrium.
 When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreased with increasing waterbody size class **[Fig.
@@
 
-The influence of waterbody size is also suggested in the interaction between NO~3~ and surface area **[Fig.
-5]**.
-At high NO~3~ concentrations, N~2~O is highly supersatured in small waterbodies, but moves toward equilibrium in large waterbodies **[Fig.
-5A]**.
-At low NO~3~ concentrations, N~2~O is undersaturated in small waterbodies, but increases toward equilibrium in larger waterbodies **[Fig.
-5A]**.
-
-![Figure 5.
-Conditional effects plots illustrating the relationship between N~2~O saturation and A) nitrate (NO~3~ availability dependent upon lake size, and B) lake size dependent upon NO~3~ availability.
-The five NO~3~ categories correspond to concentration ($\mu$g L^-1^) ranges of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13.](manuscript_figures/sat_cond_area.tiff)
-
-The model results suggest that NO~3~ may exert a strong control on N~2~O, but the relationship is complicated by interactions with water temperature and lake surface area.
-This finding mirrors reports from a survey of farm ponds in Canada where N~2~O was found to be associated with dissolved inorganic nitrogen (DIN) concentration, but the effect of DIN was conditional upon thermal stratification strength.
-These complex interactions may explain contrasting literature reports regarding the role of NO~3~ in regulating N~2~O.
-Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020; @Yang_etal_2015] while others have found no relationship [@Soued_etal_2016; @Whitfield_etal_2011].
-It is possible that the NO~3~ \~ N~2~O relationship wasn't detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
-
-Interactions between NO~3~, water temperature, and lake size produced a broad range of N~2~O saturation values across the population of CONUS waterbodies, but also enabled identifcation of broad scale patterns.
-First, small waterbodies with low NO~3~ concentration tended to be undersaturated, whereas small waterbodies with high NO~3~ tended to be highly supersaturated **[Fig.
-5]**.
-Large waterbodies, regardless of NO~3~ status, tended to have N~2~O saturation ratios closer to equilibrium.
-Therefore intra and inter regional patterns in N~2~O saturation may be well predicted by the joint distributions of waterbody size, NO~3~ availability, and water temperature.
-
 ## Emission Rates
@@
 The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
-Variability and magnitude of predicted N~2~O emission rates decreased with increasing waterbody size **[(Fig.
-6A)]**, which is consistent with the relationship between N~2~O saturation and waterbody size **[(Fig.
-5A,B)]**.
-Despite predicted widespread N~2~O undersaturation in CONUS waterbodies, the estimated national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **[(Table 1)]**.
-This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeded the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **[(SI Fig.
-1)]**.
+Variability and magnitude of predicted N~2~O emission rates decreased with increasing waterbody size **(Fig.
+5A)**, which is consistent with the relationship between N~2~O saturation and waterbody size **(Fig.
+3C,D)**.
+Despite predicted widespread N~2~O undersaturation in CONUS waterbodies, the estimated national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)**.
+This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeded the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **(SI Fig.
+1)**.
 
-![Figure 6: A) N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 waterbodies in the target population.
-C) N~2~O flux plotted against the five lake size categories used in the survey design.
+![Figure 5: A) N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 waterbodies in the target population.
 Across all plots, points and bars reflect the mean (points) and 95% credible interval, respectively, of the 2000 realizations of the posterior predictive distribution.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
@@
 Total N~2~O flux from the surface of CONUS waterbodies is a function of N~2~O emission rates and waterbody surface areas.
-The four smallest size classes (\< 50 Ha) in the survey design contained `r round(proportionSmallNumber, 2) * 100`% of all lakes in the population, but were estimated to have contributed only `r round((FluxBySize %>% filter(size_cat != "50_max") %>% select(estimate) %>% summarise(tot_small = sum(estimate)) %>% pull())/(FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) * 100, 0)`% to the N~2~O flux from CONUS waterbodies **[(Fig.
-6C)]**.
+The four smallest size classes (\< 50 Ha) in the survey design represented `r round(proportionSmallNumber, 2) * 100`% of all lakes in the population but were estimated to have contributed only `r round((FluxBySize %>% filter(size_cat != "50_max") %>% select(estimate) %>% summarise(tot_small = sum(estimate)) %>% pull())/(FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) * 100, 0)`% to the N~2~O flux from CONUS waterbodies.
 The reason for this pattern is two-fold.
 First, lakes in the smaller size classes constituted only `r round(proportionSmallSA, 2) * 100`% of total waterbody surface area, thereby limiting their N~2~O flux.
-Second, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the population **[(Fig.
-6A)].** Waterbodies in the largest size class (\> 50 Ha) contained only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population, but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
-This lake size class had the greatest N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year^-1^), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year^-1^ **[(Fig.
-6C)]**.
+Second, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the population.
+Waterbodies in the largest size class (\> 50 Ha) represented only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
+This lake size class was the source of `r ((FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) / (national_stats %>% select(post_med_fn2o) %>% pull()) * 100) %>% round(., 0)`% of the total N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year^-1^), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year^-1^.
 While uncertainty in N~2~O emission rates was smallest for large lakes **(Fig.
 6A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large waterbodies **(Fig.
-6B,C)**.
+6B)**.
 
@@
 
-The mean flux estimate for CONUS waterbodies was positive (`r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval was quite broad and included zero (`r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^) so we could not conclude with statistical confidence whether CONUS waterbodies were a net source or sink of N~2~O overall.
+The mean flux estimate for CONUS waterbodies was positive (`r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval was quite broad and included zero (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^) so we could not conclude with statistical confidence whether CONUS waterbodies were a net source or sink of N~2~O overall.
 Though we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change (IPCC) estimated that CONUS drainage networks, inclusive of zero order streams through major river mouths, emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
-The IPCC estimates total N~2~O emissions from aquatic ecosystems as a fraction (emission factor) of the anthropogenic nitrogen that leaches from soils into surface water systems and does not distinguish between components of the aquatic continuum, thus a direct comparison between the IPCC estimate and the value reported here is not possible; however, the mean N~2~O  flux from lakes and reservoirs reported here comprises only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of the IPCC estimate, suggesting the IPCC estimate is too high, possibly because the IPCC model does not allow for N~2~O consumption in lakes, ponds, and reservoirs.
-This limitation extends to most approaches used to estimate aquatic N~2~O fluxes [@Lauerwald_etal_2023], inlcuding more recent mechanistic models [@Lauerwald_etal_2019].
-Given our estimate of widespread N~2~O consumption in CONUS waterbodies, failure to include this dynamic in aquatic N~2~O models may yield biased emission estimates.
+The IPCC estimates total N~2~O emissions from aquatic ecosystems as a fraction of the anthropogenic nitrogen that leaches from soils into surface water systems but does not distinguish standing from flowing waters, thus a direct comparison between the IPCC estimate and the value reported here is not possible; however, the mean N~2~O  flux from lakes and reservoirs reported here comprises only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of the IPCC estimate, suggesting the IPCC estimate is too high, possibly because the IPCC model does not allow for N~2~O consumption in lakes, ponds, and reservoirs.
+This limitation extends to many mechanistic models used to estimate N~2~O fluxes from inland waters [@Lauerwald_etal_2019;@Marzadri_etal_2021;@Marzadri_etal_2017].
+Given our estimate of widespread N~2~O consumption in CONUS waterbodies, failure to include this dynamic in aquatic N~2~O models may yield low-biased emission estimates [@Lauerwald_etal_2023].
+Recent model parameterizations that allow for N~2~O consumption in inland waters are promising developments in this regard [@Yao_etal_2019; @Wang_etal_2023;@Maavara_etal_2019].
 
@@
 The degree of N~2~O disequilibrium also increases with decreasing waterbody size, possibly because smaller waterbodies have low air-water gas exchange rates.
-Our observations of dissolved N~2~O was predicted  due to high emissions from from a subset of the population.
 
@@
 Uncertainty in the national flux estimate was too large to determine if CONUS waterbodies functioned as a net source or sink of N~2~O.
-While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nations largest lakes including an assesment of intra-waterbody spatial variation in N~2~O, which can be important [@Salk_etal_2019, @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propagate to large uncertainties in flux estimates.
+While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nations largest lakes including an assessment of intra-waterbody spatial variation in N~2~O, which can be important [@Salk_etal_2019, @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propagate to large uncertainties in flux estimates.
 
@@
 
-The population of interest (POI) for the 2017 NLA survey [@USEPA_NLA17_2022] was CONUS waterbodies of at least 1 hectare in area, at least 1 meter deep, with at least 0.1 hectare of open water, and a minimum residence time of one week **[(Fig.
-1)]**.
+The population of interest (POI) for the 2017 NLA survey [@USEPA_NLA17_2022] was CONUS waterbodies of at least 1 hectare in area, at least 1 meter deep, with at least 0.1 hectare of open water, and a minimum residence time of one week **(Fig.
+1)**.
 Inferences from the survey samples were assumed relevant to average conditions during the summer index period between June and end of September 2017.
-The POI was sampled according to a generalized random-tessellation design with stratification [@Stevens_Olsen_2004], which prescribed a spatially balanced, probabilistic sample with stratification along three factors: surface area (hectares; 5 ordered levels), ecoregion [@Omernik_1987; @Herlihy_etal_2008; @Omernik_Griffith_2014] (9 levels), and US state (excluding AK and HI; 48 levels).
+The POI was sampled according to a generalized random-tessellation design[@Stevens_Olsen_2004] with stratification along three factors: surface area (hectares; 5 ordered levels), ecoregion [@Omernik_1987; @Herlihy_etal_2008; @Omernik_Griffith_2014] (9 levels), and US state (excluding AK and HI; 48 levels).
 The size categories were sampled with unequal probability, such that larger lakes were over-sampled relative to smaller lakes.
@@
 A 90:10 mixture of argon and methane was used as the carrier gas.
-The original dissolved N~2~O concentration was calculated from the measured headspace partial pressue, the temperature specific Bunsen coefficient, and a mass balance for the headspace equilibration system as described in Beaulieu et al.
+The original dissolved N~2~O concentration was calculated from the measured headspace partial pressure, the temperature specific Bunsen coefficient, and a mass balance for the headspace equilibration system as described in Beaulieu et al.
 [@Beaulieu_etal_2012].
@@
 In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (*e.g.,* US waterbodies) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
-In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (*e.g.*, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
+In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (*e.g.*, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI (See SI) and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
 The fully Bayesian implementation made propagation of uncertainty from the waterbody-level predictions to higher level summaries straightforward.

### 2023-11-29 manuscript/manuscript_support.R

@@ -11,20 +11,27 @@ localPath <- Sys.getenv("USERPROFILE")
 # DMAP not under renv control.  Should probably specify same
 # versions as used for renv library, but causing issues.  Just 
 # install newest version.
-if(localPath == "") { # if DMAP, then
-  # require returns FALSE and gives a warning 
-  # (rather than an error as library() does by default) 
-  # if the package does not exist.
-  if(!require("ggallin")) install.packages("ggallin")
-  if(!require("sf")) install.packages("sf")
-  if(!require("tidyverse")) install.packages("tidyverse")
-  if(!require("janitor")) install.packages("janitor")
-  if(!require("tictoc")) install.packages("tictoc")
-  if(!require("pbkrtest")) install.packages("pbkrtest") 
-  if(!require("ggpubr")) install.packages("ggpubr")
-  if(!require("ggallin")) install.packages("ggallin")
-  if(!require("USAboundaries")) install.packages("USAboundaries")
-}
+# if(localPath == "") { # if DMAP, then
+#   # require returns FALSE and gives a warning 
+#   # (rather than an error as library() does by default) 
+#   # if the package does not exist.
+#   
+#   # older version of dplyr already installed on instance.  The only way I could
+#   # figure out how to update was to use devtools::install_version
+#   if(!require("devtools")) install.packages("devtools")
+#   devtools::install_version("dplyr", "1.1.4", 
+#                             repos = "https://cran.r-project.org") # get newest version
+#   if(!require("ggallin")) install.packages("ggallin")
+#   if(!require("sf")) install.packages("sf")
+#   if(!require("tidyverse")) install.packages("tidyverse")
+#   if(!require("janitor")) install.packages("janitor")
+#   if(!require("tictoc")) install.packages("tictoc")
+#   if(!require("pbkrtest")) install.packages("pbkrtest") 
+#   if(!require("ggpubr")) install.packages("ggpubr")
+#   if(!require("ggallin")) install.packages("ggallin")
[…]

### 2023-11-30 manuscript/manuscript_file.Rmd
Minor updates, passing the torch to Roy

@@
 **Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but recent evidence of N~2~O uptake across a broad range of waterbodies have called the accuracy of emission estimates into question.
-Here we use a new national-scale dataset on dissolved N2O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull()` waterbodies in the conterminous U.S.
+Here we use a new national-scale dataset on dissolved N~2~O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` waterbodies in the conterminous U.S.
 (CONUS).
-We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean` % (95% credible interval:`r propN2oSinkLower` - `r propN2oSinkUpper`% ) of waterbodies functioning as N~2~O sinks.
-The model predicts dissolved N2O concentrations reasonably well based partly on interactions between nitrate (NO~3~) concentration, waterbody size, and water temperature.
-Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS waterbodies is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` ) metric tons N~2~O year^-1^.
+We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean`% (95% credible interval: `r propN2oSinkLower` - `r propN2oSinkUpper`%) of waterbodies functioning as N~2~O sinks.
+The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate (NO~3~) concentration, waterbody size, and water temperature.
+Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS waterbodies is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
 Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS waterbodies are a net source or sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source.**
@@
 1)**.
-These data, along with the survey design information, were used to develop a Bayesian hierarchical model to predict N~2~O concentrations for all 465,897 CONUS waterbodies in the target population.
+These data, along with the survey design information, were used to develop a Bayesian hierarchical model to predict N~2~O concentrations for all `r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS waterbodies in the target population.
 At the population level, the majority of CONUS waterbodies (mean = `r propN2oSinkMean`%, 95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) were predicted to be a net sink for N~2~O.
@@
 
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean predicted N~2~O saturation ratio (ratio of measured to equilibrium concentration) was statistically indistinguishable from one in all ecoregions **[(Table 1, Fig.
-2)]**.
-The apparent contradiction between widespread N~2~O undersaturation throughout the CONUS and average dissolved N~2~O concentrations close to equilibrium in all ecoregions can be attributed to the predicted distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies **[(Fig.
-2)]**.
+Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean predicted N~2~O saturation ratio (ratio of measured to equilibrium concentration) was statistically indistinguishable from one in all ecoregions **(Table 1, Fig.
+2)**.
+The apparent contradiction between widespread N~2~O undersaturation and average dissolved N~2~O concentrations close to equilibrium can be attributed to the predicted distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies **(Fig.
+2)**.
 Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had median values less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
-Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources, particularly at high NO~3~ at high water temperature **[Fig.
-3]**.
+Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources, particularly at high NO~3~ at high water temperature **(Fig.
+3)**.
 
@@
 D) The relationship between N~2~O saturation and NO~3~ level conditional on (log) waterbody surface area (5 quantiles).
-Horizontal dashed lines in each panel reflect an N~2~O saturation value of 1.](manuscript_figures/cond_effect_no3.tiff)
+Horizontal dashed lines in each panel reflect an N~2~O saturation value of 1.](manuscript_figures/cond_effect_no3.tif)
 
@@
 The effect of temperature was more pronounced at greater NO~3~ levels, potentially reflecting the stimulatory effect of higher temperature on biological N~2~O production.
-The relative effect of NO~3~ was estimated to be greater than that of temperature, thus elevated water temperature may serve to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
+It appears that elevated water temperature may serve to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
 
@@
 Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
-In lakes with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column [@Carpenter_1983].
-Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep lakes, resulting in greater rates of N~2~O production or consumption in small lakes.
+In waterbodies with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column [@Carpenter_1983].
+Furthermore, sediment may warm faster in small and shallow waterbodies, whereas thermal stratification may substantially delay sediment warming in large and deep waterbodies, resulting in greater rates of N~2~O production or consumption in small lakes.
 Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface [@Beaulieu_etal_2015; @Beaulieu_etal_2014].
@@
 Relationship between $\Delta$ N~2~O (difference between predicted dissolved and equilibrium N~2~O concentration) and waterbody surface areas.
-Mean (points) and 95% credible interview (horizontal bars) are derived from 2000 realizations of the posterior predictive distribution.](manuscript_figures/deltaN2ObySize.tiff)
+Mean (points) and 95% credible interview (horizontal bars) are derived from 2000 realizations of the posterior predictive distribution.](manuscript_figures/n2oStarBySize.tiff)
 
@@
 
-The emission rates reported here span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
+The emission rates estimated across CONUS waterbodies span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
 Variability and magnitude of predicted N~2~O emission rates decreased with increasing waterbody size **(Fig.
@@
 Despite predicted widespread N~2~O undersaturation in CONUS waterbodies, the estimated national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)**.
-This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeded the median (`r national_stats$post_median_en2o` mg N~2~O m^-2^ d^-1^) **(SI Fig.
+This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeded the median (`r national_stats$post_median_en2o` 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^) **(SI Fig.
 1)**.
@@
 The IPCC estimates total N~2~O emissions from aquatic ecosystems as a fraction of the anthropogenic nitrogen that leaches from soils into surface water systems but does not distinguish standing from flowing waters, thus a direct comparison between the IPCC estimate and the value reported here is not possible; however, the mean N~2~O  flux from lakes and reservoirs reported here comprises only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of the IPCC estimate, suggesting the IPCC estimate is too high, possibly because the IPCC model does not allow for N~2~O consumption in lakes, ponds, and reservoirs.
-This limitation extends to many mechanistic models used to estimate N~2~O fluxes from inland waters [@Lauerwald_etal_2019;@Marzadri_etal_2021;@Marzadri_etal_2017].
+This limitation extends to many mechanistic models used to estimate N~2~O fluxes from inland waters [@Lauerwald_etal_2019; @Marzadri_etal_2021; @Marzadri_etal_2017].
 Given our estimate of widespread N~2~O consumption in CONUS waterbodies, failure to include this dynamic in aquatic N~2~O models may yield low-biased emission estimates [@Lauerwald_etal_2023].
-Recent model parameterizations that allow for N~2~O consumption in inland waters are promising developments in this regard [@Yao_etal_2019; @Wang_etal_2023;@Maavara_etal_2019].
+Recent model parameterizations that allow for N~2~O consumption in inland waters are promising developments in this regard [@Yao_etal_2019; @Wang_etal_2023; @Maavara_etal_2019].
 
@@
 Uncertainty in the national flux estimate was too large to determine if CONUS waterbodies functioned as a net source or sink of N~2~O.
-While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include additional measurements at the nations largest lakes including an assessment of intra-waterbody spatial variation in N~2~O, which can be important [@Salk_etal_2019, @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propagate to large uncertainties in flux estimates.
+While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O, which can be important in large waterbodies[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propagate to uncertainties in flux estimates.
 
@@
 
-Fully reproducible code and further results and model workflow details are provided in Modeling **Supplement X**.
+Fully reproducible code and further results and model workflow details are provided in the Modeling Supplement Document.
 Though our initial models conditioned the response on just the survey design variables, those models consistently resulted in poorer fits to to the observed N~2~O data compared to models with additional covariates.

### 2023-11-30 manuscript/manuscript_support.R

@@ -235,6 +235,8 @@ ecoreg_stats <- all_predictions %>%
              UCL = round(quantile(prop_sat, probs = 0.975), 3),
              post_med_en2o = round(median(mean_en2o), 3),
              post_median_en2o = round(median(median_en2o), 3),
+             LCL_median_en2o = round(quantile(median_en2o, probs = 0.025), 3),
+             UCL_median_en2o = round(quantile(median_en2o, probs = 0.975), 3),
              LCL_en2o = round(quantile(mean_en2o, probs = 0.025), 3),
              UCL_en2o = round(quantile(mean_en2o, probs = 0.975), 3),
              post_med_fn2o = round(median(fn2o_Mg), 3), 
@@ -273,6 +275,8 @@ national_stats <- all_predictions %>%
              LCL = round(quantile(prop_sat, probs = 0.025), 3),
              UCL = round(quantile(prop_sat, probs = 0.975), 3),
              post_median_en2o = round(median(median_en2o), 3),
+             LCL_median_en2o = round(quantile(median_en2o, probs = 0.025), 3),
+             UCL_median_en2o = round(quantile(median_en2o, probs = 0.975), 3),
              post_med_en2o = round(median(mean_en2o), 3),
              LCL_en2o = round(quantile(mean_en2o, probs = 0.025), 3),
              UCL_en2o = round(quantile(mean_en2o, probs = 0.975), 3),

### 2023-12-14 manuscript/manuscript_file.Rmd
Updated manuscript file and support file with revised results

@@
 
-```{r source_scripts, echo=FALSE, warning=FALSE, message=FALSE}
+```{r source_scripts, message=FALSE, warning=FALSE, include=FALSE}
+# Load results objects from disk for knitting (if they already exist) ----------
+# vector of object names
+objs <- c("dg",
+          "national_stats",
+          "ecoreg_stats",
+          "em_min_max",
+          "FluxBySize",
+          "ipcc.n2o"
+          )
+
+N <- length(objs)
+
+# list of file paths for objects saved
+f <- list()
+for(i in 1:N){
+    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/manuscript/manuscript_files/",
+    objs[i], ".rda")
+    }
+
+# load files if they exist
+for(i in 1:N){
+  if(file.exists(f[[i]]))
+    load(f[[i]])
+  else
   # source scripts------
@@
     source("manuscript/manuscript_support.R", local = knitr::knit_global())
+}
 
+# clean up workspace
+rm(f, i, objs, N)
+
+# load sample frame from model files folder
+load("modelFiles/sframe.rda")
+
+# remove the Great Salt Lake from sample frame
+sframe <- sframe[-258807,]
+
+# Check for figures. If not available, load and run source script --------
+
+# vector of figure names
+objs <- c("cond_effect_no3.tif",
+          "figure1.png",
+          "figure2.tiff",
+          "n2oFluxAndEmissionRateVsContinuousArea.tiff",
+          "n2oStarBySize.tiff",
+          "SIfigure1.tiff"
+          )
+
+N <- length(objs)
+
+# list of file paths for objects saved
+f <- list()
+for(i in 1:N){
+    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/manuscript/manuscript_figures/",
+    objs[i])
+    }
+
+# load files if they exist
+for(i in 1:N){
+  if(file.exists(f[[i]]))
+    print("Good! Figure already created")
+  else
+    print(paste0("missing figure: ", f[[1]]))
+ }
+
+# clean up workspace
+rm(f, i, objs, N)
 ```
@@
 **Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but recent evidence of N~2~O uptake across a broad range of waterbodies have called the accuracy of emission estimates into question.
-Here we use a new national-scale dataset on dissolved N~2~O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` waterbodies in the conterminous U.S.
+Here we use a new national-scale dataset on dissolved N~2~O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` waterbodies in the conterminous U.S.
 (CONUS).
-We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r propN2oSinkMean`% (95% credible interval: `r propN2oSinkLower` - `r propN2oSinkUpper`%) of waterbodies functioning as N~2~O sinks.
+We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of waterbodies functioning as N~2~O sinks.
 The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate (NO~3~) concentration, waterbody size, and water temperature.
@@
 Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate [@epa_data].
-`r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
+`r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
 1)**.
-These data, along with the survey design information, were used to develop a Bayesian hierarchical model to predict N~2~O concentrations for all `r all_predictions %>% distinct(.row) %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS waterbodies in the target population.
-At the population level, the majority of CONUS waterbodies (mean = `r propN2oSinkMean`%, 95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) were predicted to be a net sink for N~2~O.
-The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`), although the 95% credible intervals overlap across all ecoregions **(Table 1)**.
+These data, along with the survey design information, were used to develop a Bayesian hierarchical model to predict N~2~O concentrations for all `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS waterbodies in the target population.
+At the population level, the majority of CONUS waterbodies (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% CI: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`) were predicted to be a net sink for N~2~O.
+The Western Mountains ecoregion had the smallest proportion of waterbodies predicted as N~2~O sinks (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(UCL_prop_sat) * 100`) and the Northern Plains had the greatest (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(UCL_prop_sat) * 100`), although the 95% credible intervals over the predictive distributions overlap across all ecoregions **(Table 1)**.
 
@@
                total_sa,
-               estimate, LCL, UCL, 
-               post_med_sat, LCI_sat, UCI_sat, 
-               post_med_en2o, LCL_en2o, UCL_en2o, 
-               post_med_fn2o, LCL_fn2o, UCL_fn2o),
+               post_m_mean_sat, LCL_mean_sat, UCL_mean_sat, 
+               post_m_prop_sat, LCL_prop_sat, UCL_prop_sat, 
+               post_m_mean_en2o, LCL_mean_en2o, UCL_mean_en2o, 
+               post_m_fn2o, LCL_fn2o, UCL_fn2o),
       ecoreg_stats %>% 
@@
                total_sa,
-               estimate, LCL, UCL, 
-               post_med_sat, LCI_sat, UCI_sat, 
-               post_med_en2o, LCL_en2o, UCL_en2o,
-               post_med_fn2o, LCL_fn2o, UCL_fn2o)) %>%
-  mutate('N2O undersaturation extent' = paste0(estimate*100, " (", LCL*100, " - ", UCL*100, ")"),
-         'N2O saturation ratio' = paste0(post_med_sat, " (", LCI_sat, " - ", UCI_sat, ")"),
-         'N2O emission rate' = paste0(round(post_med_en2o, digits = 3), " (", round(LCL_en2o, digits = 3), " - ", round(UCL_en2o, digits = 3), ")"),
-         'N2O flux' = paste0(round(post_med_fn2o, digits = 0), " (", round(LCL_fn2o, digits = 0), " - ", round(UCL_fn2o, digits = 0), ")")) %>% 
-  select(WSA9_NAME, total_sa, 'N2O undersaturation extent', 'N2O saturation ratio', 'N2O emission rate', 'N2O flux') %>%
+               post_m_mean_sat, LCL_mean_sat, UCL_mean_sat, 
+               post_m_prop_sat, LCL_prop_sat, UCL_prop_sat, 
+               post_m_mean_en2o, LCL_mean_en2o, UCL_mean_en2o,
+               post_m_fn2o, LCL_fn2o, UCL_fn2o)) %>%
+  mutate('Mean N2O saturation ratio' = paste0(post_m_mean_sat,
+                                         " (", 
+                                         LCL_mean_sat,
+                                         " - ", 
+                                         UCL_mean_sat, ")"),
+         'N2O undersaturation extent' = paste0(post_m_prop_sat * 100,
+                                               " (", 
+                                               LCL_prop_sat * 100, 
+                                               " - ", 
+                                               UCL_prop_sat * 100,
+                                               ")"),
+         'N2O emission rate' = paste0(round(post_m_mean_en2o, digits = 3),
+                                      " (", 
+                                      round(LCL_mean_en2o, digits = 3), 
+                                      " - ", 
+                                      round(UCL_mean_en2o, digits = 3), 
+                                      ")"),
+         'N2O flux' = paste0(round(post_m_fn2o, digits = 0), 
+                             " (", 
+                             round(LCL_fn2o, digits = 0), 
+                             " - ",
+                             round(UCL_fn2o, digits = 0),
+                             ")")) %>% 
+  select(WSA9_NAME, total_sa, 'Mean N2O saturation ratio', 'N2O undersaturation extent', 'N2O emission rate', 'N2O flux') %>%
   rename('spatial domain' = WSA9_NAME,
@@
     caption = as_paragraph( # allows for formatting
-      as_chunk("Table 1. Mean and 95% credible interval of N2O undersaturation extent, saturation ratio, emission rate, and flux at the national and ecoregional scales.", 
+      as_chunk("Table 1. Mean saturation ratio, undersaturation extent, emission rate, and flux estimates for N2O at the national and ecoregional scales.", 
                props = fp_text_default(bold = TRUE) # here is where we specify formatting (font, size, bold, etc)
@@
            ref_symbols = "a",
-           value = as_paragraph("Mean and 95% credible interval of 2000 realizations of posterior predictive distribution"),
+           value = as_paragraph("Mean and 95% credible intervals based on 2000 draws from the posterior distribution"),
            part = "header"
   ) %>%
-    footnote(i = 1, j = 3,
+    footnote(i = 1, j = 4,
            ref_symbols = ", b",
@@
   fit_to_width(ft_1, max_width = 11) # scale to fill 8x11 paper, landscape
-
 ```
@@
 
-Despite widespread N~2~O undersaturation in CONUS waterbodies, the mean predicted N~2~O saturation ratio (ratio of measured to equilibrium concentration) was statistically indistinguishable from one in all ecoregions **(Table 1, Fig.
+Despite widespread N~2~O undersaturation across CONUS waterbodies, the estimated means for N~2~O saturation ratio (ratio of measured to equilibrium concentration) were indistinguishable from one in all ecoregions **(Table 1, Fig.
 2)**.
-The apparent contradiction between widespread N~2~O undersaturation and average dissolved N~2~O concentrations close to equilibrium can be attributed to the predicted distribution of dissolved N~2~O concentrations across the population of CONUS waterbodies **(Fig.
+The apparent contradiction of widespread N~2~O undersaturation with average dissolved N~2~O concentrations closer to equilibrium can be attributed to the prominent right skew predicted for dissolved N~2~O concentrations across the population of CONUS waterbodies **(Fig.
 2)**.
-Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had median values less than 1, indicating undersaturation, but a long right tail composed of relatively rare waterbodies with levels of N~2~O supersaturation that are sufficiently high to drive the population mean to values close to, or even greater than 1.
-Thus most waterbodies in the CONUS are functioning as N~2~O sinks during the summer months, but a few function as strong N~2~O sources, particularly at high NO~3~ at high water temperature **(Fig.
+Across the CONUS, and within each of the nine major ecoregions, each predicted distribution of N~2~O saturation ratios had a median value less than 1, indicating that most waterbodies were undersaturated in these populations.
+However, the long right tails on these distributions, comprised of relatively rare waterbodies with very high levels of N~2~O supersaturation, were sufficient to drive the population means to values close to, or even greater than 1.
+Thus, most waterbodies in the CONUS were functioning as N~2~O sinks during the summer months, but a few functioned as strong N~2~O sources, particularly at high NO~3~ and high water temperature **(Fig.
 3)**.
@@
 ![Figure 2.
-Posterior predictive distribution of N~2~O saturation values for waterbodies in the nine major ecoregions.
+Predicted distribution of N~2~O saturation values across waterbodies in the nine major ecoregions.
 Each blue line (n=2000) represents the distribution of values from a single realization of the posterior predictive distribution.
-Vertical black dashed and solid lines indicate point estimates of the median and mean, respectively (Table 1).
+Vertical black dashed and solid lines indicate point estimates of the median and mean, respectively, across all realizations (Table 1).
 Vertical red line indicates a saturation ratio of 1.
-x-axis is truncated at 3 to better show the area where 97% of the observations occur.](manuscript_figures/figure2.tiff)
+x-axis is truncated at 3 to better show the area where 97% of the observations occurred.](manuscript_figures/figure2.tiff)
 
@@
 
-The model results suggest that NO~3~ may exert a strong control on N~2~O, but the relationship is complicated by interactions with water temperature and lake surface area **(Fig.
+The model results suggest that NO~3~ may be an important predictor for N~2~O, but the relationship is complicated by interactions with water temperature and lake surface area **(Fig.
 3)**.
@@
 These complex interactions may explain contrasting literature reports regarding the role of NO~3~ in regulating N~2~O.
-Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020; @Yang_etal_2015] while others have found no relationship [@Soued_etal_2016; @Whitfield_etal_2011].
+Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020; @Yang_etal_2015] whereas others have found no relationship [@Soued_etal_2016; @Whitfield_etal_2011].
 It is possible that the NO~3~ \~ N~2~O relationship wasn't detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
@@
 The influence of NO~3~ levels on N~2~O saturation **(Fig.
-3)** may reflect the relationship between NO~3~ levels and denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Kuypers_etal_2018; @Richardson_etal_2009].
+3)** may reflect a relationship between NO~3~ levels and denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Kuypers_etal_2018; @Richardson_etal_2009].
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
@@
 Nitrification may be more prevalent than denitrification in the predominantly well oxygenated surface waters where the N~2~O samples were collected because denitrification is an anaerobic process.
-However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column [@Broman_etal_2021] and can be transported from aoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
+However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column [@Broman_etal_2021] and can be transported from anoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
 Thus the N~2~O - NO~3~ relationship may be reflective of some combination of N~2~O production/consumption via denitrification and N~2~O production via nitrification.
@@
 The effect of temperature was more pronounced at greater NO~3~ levels, potentially reflecting the stimulatory effect of higher temperature on biological N~2~O production.
-It appears that elevated water temperature may serve to magnify rates of N~2~O production that are largely dictated by NO~3~ availability.
+Elevated water temperature may serve to magnify rates of N~2~O production that may be largely dictated by NO~3~ availability.
 
@@
 Large waterbodies, regardless of NO~3~ status, were predicted to have N~2~O saturation ratios closer to equilibrium.
-When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreased with increasing waterbody size class **[Fig.
-4]**.
-This pattern holds regardless of whether N~2~O was supersaturated or undersatured.
+When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreased with increasing waterbody size class **(Fig.
+4)**.
+This pattern holds regardless of whether N~2~O was supersaturated or undersaturated.
 Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
@@
 ![Figure 4.
-Relationship between $\Delta$ N~2~O (difference between predicted dissolved and equilibrium N~2~O concentration) and waterbody surface areas.
-Mean (points) and 95% credible interview (horizontal bars) are derived from 2000 realizations of the posterior predictive distribution.](manuscript_figures/n2oStarBySize.tiff)
+Estimated mean $\Delta$ N~2~O (difference between dissolved and equilibrium N~2~O concentrations) versus waterbody surface area (ha) size categories.
+Posterior mean (points) and 95% credible intervals (horizontal bars) were based on 2000 draws from the posterior distribution.](manuscript_figures/n2oStarBySize.tiff)
 
@@
 
-The emission rates estimated across CONUS waterbodies span a broad range (`r min.max$min %>% round(.,2)` - `r min.max$max %>% round(.,2)` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of waterbodies, but are within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
+The predicted emission rates for CONUS waterbodies spanned a broad range (95% CI min: `r em_min_max$min_LCL` - `r em_min_max$min_UCL` mg N~2~O m^-2^ d^-1^, 95% CI max: `r em_min_max$max_LCL` - `r em_min_max$max_UCL` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of waterbodies, but within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
 Variability and magnitude of predicted N~2~O emission rates decreased with increasing waterbody size **(Fig.
-5A)**, which is consistent with the relationship between N~2~O saturation and waterbody size **(Fig.
+5A)**, which reflects the relationship between N~2~O saturation and waterbody size **(Fig.
 3C,D)**.
-Despite predicted widespread N~2~O undersaturation in CONUS waterbodies, the estimated national average N~2~O emission rate was positive (mean = `r national_stats$post_med_en2o`, 95% CI: `r national_stats$LCL_en2o` - `r national_stats$UCL_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)**.
-This contradiction is due to relatively rare waterbodies with very high emission rates that pull the mean (`r national_stats$post_med_en2o` mg N~2~O m^-2^ d^-1^) to a value that far exceeded the median (`r national_stats$post_median_en2o` 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^) **(SI Fig.
+Despite widespread N~2~O undersaturation predicted across CONUS waterbodies, the estimate for the national average N~2~O emission rate was centered over positive values (posterior mean = `r national_stats$post_m_mean_en2o`, 95% CI: `r national_stats$LCL_mean_en2o` - `r national_stats$UCL_mean_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)** and clearly exceeded the negative estimate for the national median (posterior mean = `r national_stats$post_m_median_en2o`, 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^).
+This departure from the median was due to relatively rare waterbodies with very high emission rates pulling the mean towards the direction of positive skew **(SI Fig.
 1)**.
 
-![Figure 5: A) N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 waterbodies in the target population.
-Across all plots, points and bars reflect the mean (points) and 95% credible interval, respectively, of the 2000 realizations of the posterior predictive distribution.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
+![Figure 5: A) Predicted N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 waterbodies in the target population.
+Points and bars reflect the posterior means (points) and 95% credible intervals, respectively, based on 2000 draws from the posterior predictive distribution for each waterbody.
+C) A comparison of the estimated relative flux contributions from waterbodies in each size class to the total CONUS flux.
+The red line connects the mean of the posterior estimate of the ratio, at each size class, based on 2000 draws.
+The black polygon indicates the 95% credible intervals for the ratio estimates across the size classes](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
 
@@
 Total N~2~O flux from the surface of CONUS waterbodies is a function of N~2~O emission rates and waterbody surface areas.
-The four smallest size classes (\< 50 Ha) in the survey design represented `r round(proportionSmallNumber, 2) * 100`% of all lakes in the population but were estimated to have contributed only `r round((FluxBySize %>% filter(size_cat != "50_max") %>% select(estimate) %>% summarise(tot_small = sum(estimate)) %>% pull())/(FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) * 100, 0)`% to the N~2~O flux from CONUS waterbodies.
+The four smallest size classes (\< 50 Ha) in the survey design represented `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(number = 1:nrow(.), proportion = number / nrow(.)) %>% filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% pull(proportion) %>% round(2) * 100`% of all lakes in the population but were estimated to have contributed less, on average, to the total N~2~O flux from CONUS waterbodies **(Fig 5C)**.
 The reason for this pattern is two-fold.
-First, lakes in the smaller size classes constituted only `r round(proportionSmallSA, 2) * 100`% of total waterbody surface area, thereby limiting their N~2~O flux.
+First, lakes in the smaller size classes constituted only `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(cumulative = cumsum(area_ha), distribution = cumulative / sum(area_ha), delta50 = abs(50-area_ha)) %>% filter(delta50 == min(delta50)) %>% pull(distribution) %>% round(3) * 100`% of total waterbody surface area, thereby limiting their N~2~O flux.
 Second, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the population.
-Waterbodies in the largest size class (\> 50 Ha) represented only `r round(1 - proportionSmallNumber, 2) * 100`% of the total number of lakes in the population but had `r round(1-proportionSmallSA, 2) * 100`% of total lake surface area.
-This lake size class was the source of `r ((FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull()) / (national_stats %>% select(post_med_fn2o) %>% pull()) * 100) %>% round(., 0)`% of the total N~2~O flux (mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(estimate) %>% pull() %>% round(0)` metric tons N~2~O year^-1^), but the flux estimate is poorly constrained with a 95% credible interval of `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull()` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull()` metric tons N~2~O year^-1^.
+Waterbodies in the largest size class (\> 50 Ha) represented only `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(number = 1:nrow(.), proportion = number / nrow(.)) %>% filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% pull(proportion) %>% round(2) * 100 * -1 + 100`% of the total number of lakes in the population but had `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(cumulative = cumsum(area_ha), distribution = cumulative / sum(area_ha), delta50 = abs(50-area_ha)) %>% filter(delta50 == min(delta50)) %>% pull(distribution) %>% round(3) * 100 * -1 +100`% of total lake surface area.
+This lake size class was estimated to be, on average, the largest contributor to the total N~2~O flux across CONUS **(Fig 5C)**, though the flux estimate is poorly constrained (posterior mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(post_m_flux) %>% pull() %>% round(1)`, 95%CI: `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull() %>% round(1)` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull() %>% round(1)` metric tons N~2~O year^-1^).
 While uncertainty in N~2~O emission rates was smallest for large lakes **(Fig.
-6A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large waterbodies **(Fig.
-6B)**.
+5A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large waterbodies **(Fig.
+5B)**.
 
-Despite widespread N~2~O undersaturation, the credible interval for the predicted fluxes overlapped with zero for all ecoregions but the Temperate Plains, Southern Plains, and Southern Appalachians ecoregions, and the credible intervals were positive for these ecoregions **(Table 1)**, suggesting they were more likely to have functioned as net N~2~O sources.
-Overall, the credible intervals of the predicted mean flux overlapped across all nine ecoregions, which is somewhat surprising given the large differences in environmental conditions across the ecoregions.
+Despite the estimated widespread N~2~O undersaturation among CONUS waterbodies, the credible intervals for ecoregion-level flux estimates included zero for all but three ecoregions **(Table 1)**.
+Thus, status as a net source versus sink of N~2~O was uncertain for most ecoregions.
+However, the estimates for the Southern Appalachians, Southern Plains, and Temperate Plains included only positive values, suggesting they were likely to have functioned as net N~2~O sources.
+We were also unable to infer any clear differences in total flux between ecoregions.
+This result may sound surprising given the large differences in environmental conditions across these ecoregions, but it should not be mistaken as inferring "no differences" as it primarily emphasizes the challenges in making sufficiently precise estimates at this scale; even with unprecedented data coverage, a robust survey methodology, and relatively advanced modeling techniques.
 
-The mean flux estimate for CONUS waterbodies was positive (`r national_stats %>% select(post_med_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval was quite broad and included zero (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^) so we could not conclude with statistical confidence whether CONUS waterbodies were a net source or sink of N~2~O overall.
-Though we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change (IPCC) estimated that CONUS drainage networks, inclusive of zero order streams through major river mouths, emit `r format(ipcc.n2o.surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
-The IPCC estimates total N~2~O emissions from aquatic ecosystems as a fraction of the anthropogenic nitrogen that leaches from soils into surface water systems but does not distinguish standing from flowing waters, thus a direct comparison between the IPCC estimate and the value reported here is not possible; however, the mean N~2~O  flux from lakes and reservoirs reported here comprises only `r round((national_stats %>% select(post_med_fn2o) %>% pull() / ipcc.n2o.surface) * 100, digits = 1)` % of the IPCC estimate, suggesting the IPCC estimate is too high, possibly because the IPCC model does not allow for N~2~O consumption in lakes, ponds, and reservoirs.
+The estimate for flux across all CONUS waterbodies was centered over positive values (posterior mean = `r national_stats %>% select(post_m_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval suggested a good deal of uncertainty and included zero (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
+Again, we could not confidently infer the extent to which CONUS waterbodies, overall, were a net source or sink of N~2~O.
+Though we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change (IPCC) estimated that CONUS drainage networks, inclusive of zero order streams through major river mouths, emit `r format(ipcc.n2o$surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
+The IPCC estimates total N~2~O emissions from aquatic ecosystems as a fraction of the anthropogenic nitrogen that leaches from soils into surface water systems, but does not distinguish standing from flowing waters, thus a direct comparison between the IPCC estimate and the value reported here is not possible.
+However, the upper 95% credible interval of the N~2~O flux estimate for lakes and reservoirs reported here comprises only `r round((national_stats %>% select(UCL_fn2o) %>% pull() / ipcc.n2o$surface) * 100, digits = 1)` % of the IPCC estimate, suggesting the IPCC estimate is too high, possibly because the IPCC model does not allow for N~2~O consumption in lakes, ponds, and reservoirs.
 This limitation extends to many mechanistic models used to estimate N~2~O fluxes from inland waters [@Lauerwald_etal_2019; @Marzadri_etal_2021; @Marzadri_etal_2017].
 Given our estimate of widespread N~2~O consumption in CONUS waterbodies, failure to include this dynamic in aquatic N~2~O models may yield low-biased emission estimates [@Lauerwald_etal_2023].
-Recent model parameterizations that allow for N~2~O consumption in inland waters are promising developments in this regard [@Yao_etal_2019; @Wang_etal_2023; @Maavara_etal_2019].
+Recent models allowing for N~2~O consumption in inland waters are promising developments in this regard [@Yao_etal_2019; @Wang_etal_2023; @Maavara_etal_2019].
 
-Not only is our estimate of N~2~O emissions from CONUS lakes and reservoirs small when compared to the IPCC estimate, it is also small relative to other anthropogenic N~2~O sources including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
+Not only is our estimate of N~2~O emissions from CONUS lakes and reservoirs small compared to the IPCC estimate, it is also small relative to other anthropogenic N~2~O sources, including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
 Whether CONUS waterbodies were a net source or sink of N~2~O, their contribution to the US N~2~O budget was likely small by comparison.
@@
 Our research is the first to suggest that most lakes and reservoirs in the CONUS function as N~2~O sinks during the summer.
-Our observations of dissolved N~2~O were predicted reasonably well using non-linear interactions between observed water temperature, NO~3~ concentration, and waterbody size (see SI).
-The model form and estimated parameters suggest that NO~3~ concentration plays a role in determining whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink is exaggerated with increasing water temperature.
-The degree of N~2~O disequilibrium also increases with decreasing waterbody size, possibly because smaller waterbodies have low air-water gas exchange rates.
+Our observations of dissolved N~2~O were predicted reasonably well using a regression model with interactions between observed water temperature, NO~3~ concentration, and waterbody size (see SI).
+These modeling results suggested that NO~3~ levels may play a role in determining whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink may be exaggerated with increasing water temperature.
+The degree of N~2~O disequilibrium also increased with decreasing waterbody size, possibly because smaller waterbodies have low air-water gas exchange rates.
 
@@
 Uncertainty in the national flux estimate was too large to determine if CONUS waterbodies functioned as a net source or sink of N~2~O.
-While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O, which can be important in large waterbodies[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty which can propagate to uncertainties in flux estimates.
+While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O, which can be important in large waterbodies[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty, which can propagate to uncertainties in flux estimates.
 

### 2023-12-14 manuscript/manuscript_support.R

@@ -63,6 +63,7 @@ dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"),
                   crs = 4269) %>% # standard for lat/lon
   st_transform(5070) # project to CONUS ALBERS for plotting
 
+save(national_stats, file = "manuscript/manuscript_files/dg.rda")
 
 ## load ecoregions----
 # read in ecoregion polygons
@@ -132,8 +133,9 @@ wsa9names <- structure(list(WSA9 = c("CPL", "NAP", "NPL", "SAP", "SPL",
                        row.names = c(NA, 9L), class = "data.frame")
 
 
-all_predictions <- full_join(all_predictions, wsa9names)
+all_predictions_ms <- full_join(all_predictions, wsa9names)
 
+rm(all_predictions) # free up RAM
 
 ## emission rates----
 # code copied from dgIndicatorAnalysis.Rmd
@@ -153,7 +155,7 @@ wind <- read.table(paste0(Sys.getenv("USERPROFILE"),
 wind %>% head()
 
 # print first 6 .row values of all_predictions
-all_predictions %>%
+all_predictions_ms %>%
   filter(.row %in% 1:6) %>% # each row is a unique waterbody
   group_by(.row) %>%
   slice(1) # grabs first record
@@ -168,10 +170,10 @@ wind <- wind %>% mutate(.id = row_number())
 # Merge wind and all_predictions
 #tic() # 21 seconds on VM
 # left join needed to exclude great salt lake in wind data.  see line 88-93
-all_predictions <- left_join(all_predictions, wind, by = c(".row" = ".id")) # 40 seconds on laptop
+all_predictions_ms <- left_join(all_predictions_ms, wind, by = c(".row" = ".id")) # 40 seconds on laptop
 #toc()
 dim(wind) # 465897
-dim(all_predictions) #931792000, 2000 observations for each lake
+dim(all_predictions_ms) #931792000, 2000 observations for each lake
 465896*2000 #931792000
 
 summary(wind$ws)
@@ -193,7 +195,7 @@ pred.k600 <- function(ws, area) {
 }
 
 ### calculate emission and flux----
-all_predictions <- all_predictions %>%
+all_predictions_ms <- all_predictions_ms %>%
   mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
[…]

### 2023-12-14 manuscript/manuscript_support.R
comments for Roy to address

@@ -31,6 +31,7 @@ localPath <- Sys.getenv("USERPROFILE")
 #   if(!require("ggallin")) install.packages("ggallin")
 #   if(!require("USAboundaries")) devtools::install_version("USAboundaries", "0.4.0", 
 #                                                 repos = "https://cran.r-project.org/")
+#   if(!require("tidybayes")) install.packages("tidybayes")
 # }
 
 library(sf) # spatial data
@@ -40,8 +41,8 @@ library(USAboundaries) # state boundaries
 library(tictoc) # processing time
 library(ggpubr) # multiple plots
 library(ggallin) # psuedlolog transformation for negative values
-
-
+library(USAboundaries) # state boundaries for maps
+library(tidybayes) # prop flux by size figure
 
 # DATA---------------
 ## load sample data (dg object)----
@@ -63,7 +64,7 @@ dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"),
                   crs = 4269) %>% # standard for lat/lon
   st_transform(5070) # project to CONUS ALBERS for plotting
 
-save(national_stats, file = "manuscript/manuscript_files/dg.rda")
+save(national_stats, file = "manuscript/manuscript_files/dg.rda")  # this object not created yet
 
 ## load ecoregions----
 # read in ecoregion polygons
@@ -555,8 +556,12 @@ PropFluxBySize <- all_predictions_ms %>%
 save(PropFluxBySize, file = "manuscript/manuscript_files/PropFluxBySize.rda")
 
 ## Figure on proportion of flux (ratio size class / CONUS)
+############## ROY
+##############
+# OBJECT PropFluxBySize DOES NOT HAVE size_f and size_p_f variables.
+# they were lost during the final summarize above (551-553)
 b3 <- PropFluxBySize %>% 
-  mutate(total_f = size_f / size_p_f) %>% 
+  mutate(total_f = size_f / size_p_f) %>% # CODE BREAKS HERE
   select(size_cat, .draw, total_f, size_f, size_p_f) %>%
   mutate(
     size_cat = fct_recode(as.factor(size_cat),

### 2024-03-18 scripts/NLA_N2O_models.Rmd
Updated modeling documentation

@@
       engine: webtex
-      url: https://latex.codecogs.com/svg.image?
+    #  url: https://latex.codecogs.com/svg.image?
     html_preview: TRUE
@@
   chunk_output_type: inline
+  markdown: 
+    wrap: 72
 ---
@@
 
-# Load model and data files for knitting (if they already exist)
-f1 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda"
-f2 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda"
-f3 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda"
-f4 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda"
-f5 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda"
-f6 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/samp_props.rda"
-
-f7 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda"
-f8 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod2.rda"
-f9 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda"
-f10 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda"
-f11 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod5.rda"
-f12 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod6.rda"
-f13 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/all_predictions_b.rda"
-f14 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda"
-f15 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/pframe.rda"
-f16 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests.rda"
-f17 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_survey_ests_size.rda"
-f18 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests.rda"
-f19 <- "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sat_survey_ests_size.rda"
-
-if(file.exists(f1))
-  load(f1)
-if(file.exists(f2))
-  load(f2)
-if(file.exists(f3))
-  load(f3)
-if(file.exists(f4))
-  load(f4)
-if(file.exists(f5))
-  load(f5)
-if(file.exists(f6))
-  load(f6)
-if(file.exists(f7))
-  load(f7)
-if(file.exists(f8))
-  load(f8)
-if(file.exists(f9))
-  load(f9)
-if(file.exists(f10))
-  load(f10)
-if(file.exists(f11))
-  load(f11)
-if(file.exists(f12))
-  load(f12)
-if(file.exists(f13))
-  load(f13)
-if(file.exists(f14))
-  load(f14)
-if(file.exists(f15))
-  load(f15)
-if(file.exists(f16))
-  load(f16)
-if(file.exists(f17))
-  load(f17)
-if(file.exists(f18))
-  load(f18)
-if(file.exists(f19))
-  load(f19)
-```
-
-# Background and Objectives
-This document describes the modeling workflow and results from estimating dissolved and equilibrium N2O gas concentrations and saturation ratios for all freshwater waterbodies (natural and man-made lakes and reservoirs) in the lower 48 US states larger than 1 ha.
-Data for these estimates came from the 2017 Nation Lakes Assessment (NLA) survey, wherein a subset of waterbodies were sampled from the broader population of interest (POI) according to a spatially balanced, stratified, and unequal probability design.
-Stratification was among categories of lake size (surface area in hectares), WSA9 ecoregion, and US state (excluding AK and HI).
-In order to better capture variability among the wide range of lake sizes in the POI, larger lakes were intentionally over-sampled relative to smaller lakes.
-
-To make inferences back to the POI from this sample, estimates needed to be adjusted in a manner that addressed potential biases related to the relative representation of waterbody characteristics in the sample compared to the POI, specifically with respect to the survey design variables.
-It can be helpful to define the complete data likelihood (CDL) when considering how to address biases that may arise from sampling designs [@Link_Barker_2010; @Zachmann_etal_2022].
-For the 2017 NLA survey, the complete data would be considered all US lakes and reservoirs in the lower 48 states larger than 1 hectare (_i.e._, the POI).
-The survey samples, by comparison, could then be considered a subset of that complete data with a known pattern of "missingess" assigned by the survey.
-In this case, waterbodies in the POI were missing from the sample not completely at random, but according to a probability assigned by the design parameters.
-This type of missingness is considered not ignorable when making inferences from a sample to a POI [@Gelman_etal_2014, Ch.
-8; @Zachmann_etal_2022].
-However, _conditional_ on the design variables, the sample could be considered a simple random sample (SRS) from the POI [@Gelman_etal_2014, Ch.
-8; @Zachmann_etal_2022].
-
-A simple, model-based strategy for adjusting a sample to a POI is to include the survey design variables as predictors in a regression model [@Gelman_etal_2020, Ch.
-17].
-After fitting the regression, the parameter estimates can be used to make adjustments according to the distribution of the design variables in the POI in a second step referred to as "poststratification" [@Gelman_etal_2020, Ch.
-17].
-A particularly popular version of this approach uses multilevel regression models and is referred to as multilevel regression with poststratification, or MRP [@Park_etal_2004; @Gelman_Little_1997; @Gelman_etal_2020, Ch.
-17; @Kennedy_Gelman_2021].
-Multilevel regression models are often recommended because they can provide regularized estimates along the design groupings, which can improve out-of-sample inferences [_e.g._, @Gelman_Little_1997; @Kennedy_Gelman_2021].
-Estimates for group levels that may be missing from the sample, but are part of the population of interest, are also straightforward using the multilevel approach [@Gelman_etal_2020 Ch.
-17; @McElreath_2020].
-For a recent applied example and helpful tutorial employing this approach, see @Kennedy_Gelman_2021.
-For an example of a recent application in the context of national surveys of environmental resources, see @Zachmann_etal_2022.
-
-In this study, we used a MRP-like approach to provide population-level estimates based on the 2017 NLA dissolved gas survey data.
-The remainder of this document provides the workflow, data, models, and code used to make these estimates.
-To begin, a number of Bayesian multilevel regression models were fit to the sample data.
-These models were fit and checked against both the data used to construct the models as well as hold-out data from future visits.
-After evaluating a number of different models, a final model was selected for use in inference.
-Because total N2O flux estimates needed to be estimated based on individual, waterbody-level measures (_e.g._, surface area), predictions from the model were made to each individual lake in the POI (n = 465,897).
-These predictions were assumed relevant to average conditions during the biological index period for each lake in 2017.
-
-The objective of the modeling effort was to provide estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1) for the National-level POI and a number of sub-populations therein (_e.g._, ecoregion, state, size category).
-The gas estimates were also used to estimate the total flux of N2O gas attributable to the target populations.
-Saturation ratio was calculated as the ratio of dissolved to equilibrium N2O.
-Because dissolved and equilibrium N2O were observed on the same sample units (sites in lakes), the regression models were fit to their joint distribution and the ratio estimates were assembled as a derived quantity calculated from the joint posterior predictive distribution of those variables.
-The response variable was, therefore, modeled as multivariate in order to account for any potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography.
-Although predictions of the mean marginal probabilities from separate models may have provided comparable estimates, a joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
-All of the models fit were constructed using the `brms` package [@Burkner_2017] in `R` [@R_Core_Team_2021] as an interface to Stan, a software package for fitting fully Bayesian models via Hamiltonian Monte Carlo [HMC, @Stan_Development_Team_2018_a; @Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c].
-More specific details on the model structures is provided throughout the "Model fitting" section below.
+# Load model and data objects from drive for knitting (if they already exist)
+## vector of object names
+objs <- c("dg",
+          "df_model",
+          "df_test",
+          "sframe",
+          "pframe",
+          "samp_props",
+          "n2o_mod1",
+          "n2o_mod2",
+          "n2o_mod3",
+          "n2o_mod4",
+          "n2o_mod5",
+          "n2o_mod6",
+          #"all_predictions_b",
+          "n2o_survey_ests",
+          "n2o_survey_ests_size",
+          "sat_survey_ests",
+          "sat_survey_ests_size"
+          )
+
+N <- length(objs)
+
+# list of file paths for objects saved
+f <- list()
+for(i in 1:N){
+    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/modelFiles/",
+    objs[i], ".rda")
+    }
+
+# load files if they exist
+for(i in 1:N){
+  if(file.exists(f[[i]]))
+    load(f[[i]])
+}
+
+# clean up workspace
+rm(f, i, objs, N)
+```
+
+# Introduction
+
+This supplement documents the modeling workflow used for estimating
+dissolved and equilibrium N2O gas concentrations and saturation ratios
+for freshwater lakes (natural and man-made) across the continental
+United States (CONUS). Data for the estimates came from the 2017 Nation
+Lakes Assessment (NLA) survey, which aimed to provide representative
+samples for describing environmental conditions across CONUS lakes
+during the summer index period of 2017 [@USEPA_NLA_DESIGN_2017]. The
+population of interest (POI) for this study was defined as all lakes and
+reservoirs in the CONUS that were (1) larger than 1 hectare, (2) with at
+least 0.1 hectares of open water, (3) at least 1 meter deep, and (4)
+with a residence time of less than 1 week. The 2017 NLA sampling frame
+consisted of 465,897 lakes delineated from georeferenced data documented
+in the National Hydrography Dataset (NHD). The samples were drawn
+according to a spatially balanced, stratified, and unequal probability
+design. Sampling was stratified among categories of lake size (surface
+area in hectares), WSA9 ecoregion, and US state (excluding AK and HI).
+The probability of selection was greater for larger lakes relative to
+smaller ones in order to sufficiently represent larger lakes in a sample
+of practical size. Small lakes vastly outnumber large ones in the POI.
+
+The POI was defined in this study explicitly by the NLA 2017 sampling
+frame ([@USEPA_NLA_DESIGN_2017]). Therefore, the effective and stated
+POI may depart in important ways that would be difficult to confirm in
+practice. For example, it would be impractical to visit all lakes in the
+sampling frame to determine which contained less than 0.1 ha of open
+water or were less than 1 meter deep. These conditions were evaluated in
+the field for sites that were sampled. If a field crew arrived at a site
+that did not meet the conditions for inclusion, another site could be
+chosen from an oversample frame. In effect, the size of the stated POI
+is likely smaller than the 465,897 lakes described in the sampling frame
+and, therefore, the possibility of some remaining biases in these
+estimates should be considered, if only to emphasize that there remains
+uncertainty in the population-level estimates beyond what was estimated
+from modeling. Though it may be practically impossible to determine the
+precise magnitude or direction in which any particular estimate could be
+biased, it is clear that estimates could vary systematically given any
+substantial mismatches between the POI and the sampling frame.
+
+When making inferences to a POI from a sample, estimates should be
+adjusted to address potential biases resulting from known differences
+between the POI and sample. For this study, lakes in the POI were
+missing from the sample not completely at random, but systematically and
+according to the probability assigned by the survey design. This type of
+"missingness" is not ignorable for making inferences from a sample, but
+may be considered ignorable if the inferences are adjusted such that
+they are conditional on the survey design [@Gelman_etal_2014, Ch. 8;
+@Gelman_etal_2020 Ch. 17]. A straightforward approach to adjustment is
+to simply include the design variables as predictors in a regression
+model for the surveyed responses of interest [@Gelman_etal_2020, Ch.
+17]. Then, in a second step referred to as "poststratification"
+[@Gelman_etal_2020, Ch. 17], the model parameter estimates can be used
+to make population-level estimates by summarizing over the known
+distribution of the design variables in the POI or sub-populations
+therein. Multilevel regression models are often preferred in this
+approach, which is why it is often referred to as multilevel regression
+with poststratification, or MRP [@Park_etal_2004; @Gelman_Little_1997;
+@Gelman_etal_2020, Ch. 17; @Kennedy_Gelman_2021]. Multilevel regression
+models are recommended because they naturally provide regularized
+estimates along the design groupings, which can improve out-of-sample
+inferences [@Gelman_Little_1997]. Likewise, estimates for group levels
+that may be missing from the sample (e.g., an unsampled U.S. state), but
+are part of the POI, are straightforward using the multilevel approach
+[@Gelman_etal_2020 Ch. 17; @McElreath_2020]. For a worked and narrated
+example using this approach, see a recent paper by Kennedy and Gelman
+[@Kennedy_Gelman_2021]. For an example in the context of national
+surveys of environmental resources, see a recent paper from Zachmann and
+others [@Zachmann_etal_2022].
+
+In this study, we used the 2017 NLA survey data and an adjustment
+approach similar to MRP to provide population-level estimates for (1)
+dissolved and equilibrium N2O concentrations; (2) the N2O saturation
+ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of
+under-saturated water bodies (i.e., saturation ratio \< 1). The
+resulting estimates were later combined with lake-level attributes
+(e.g., surface area) to provide estimates of N2O emission rates and
+flux. A number of Bayesian multilevel regression models were fit to the
+survey data and predictions from those models were checked against both
+the data used in their fitting as well as a subset of held-out data from
+re-visited sites. Because dissolved and equilibrium N2O were each
+observed on the same sample units (a single site within each waterbody),
+our regression models were constructed to estimate their joint
+probability distribution. The multivariate response allowed for
+conditional dependencies between dissolved and equilibrium N2O due to,
+for example, physiography. Although predictions from separate models for
+each N2O response may have provided comparable estimates, a joint model
+was expected to better capture uncertainty and potentially improve
+out-of-sample predictions, should the data be conditionally correlated
+[@Warton_etal_2015; @Poggiato_etal_2021]. The saturation ratio
+(saturation = dissolved N2O/equilibrium N2O) was estimated as a derived
+quantity calculated from the joint posterior predictive distribution of
+the two N2O variables.
+
+Our initial regression models were focused on conditioning the N2O data
+on the survey design factors only, which included multilevel groupings
+of the sampled lakes according to WSA9 ecoregion, states within
+ecoregions, and size classes within states. Although these initial
+models performed well at replicating the observed group-level means,
+further model checking indicated that they did a relatively poor job of
+replicating some tail properties, such as the long right tail of the
+observed dissolved N2O data. As a result of this misfit, the predictive
+distribution for the saturation ratio was also consistently biased low
+relative to the observed data. In subsequent models, we found that
+including an NO3 covariate was particularly helpful for better
+replicating these extreme values. The addition of predictors for surface
+temperature and lake surface area, along with interaction terms for
+these predictors with NO3, also lead to apparent improvements. Likewise,
+the inclusion of surface temperature and elevation covariates in the
+equilibrium N2O response resulted in improvements over the initial
+models, as did added flexibility in the distributional terms
+(heterogeneous errors) for both dissolved and equilibrium N2O.
+
+A number of additional, potentially relevant covariates were evaluated
+in exploratory analyses and models, but are not covered in this
+document. These variables are, however, available to the public should
+the reader wish to further explore their utility. These measures, which
+were taken at the time of N2O sampling or constructed from those
+measures, included chlorophyll a concentration, dissolved oxygen (DO)
+content at the surface, and a buoyancy metric. Some of these variables
+are also covered in the data munging document at:
+<https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html>.
+An exploratory analysis using classification trees in that document
+suggested that NO3 as well as both mean depth and chlorophyll a
+concentration were associated with dissolved N2O concentration. The
+decision to not include these variables in the final model was based on
+several issues. First, our primary goal was to make population-level
+estimates for N2O while considering the survey design structure. As a
+result, we were less concerned with exploring the relative contributions
+of environmental correlates to N2O variation. However, our initial
+models excluding all covariates proved to result in relatively poor fits
+to important aspects of the observed data. In an attempt to improving
+the fit, the NO3 covariate was then included due to the relatively large
+effects implied in exploratory analyses (see "Sample Data Exploration"
+to follow). Conversely, our evaluations of exploratory plots and models
+with chlorophyll a, surface DO, and the BF covariates did not result in
+similarly substantial improvements. Second, these other covariates were
+only available in the sample and, therefore, were not fully observed in
+the POI. As such, using them to make population-level predictions would
+have come with additional complexity (discussed below) that didn't
+appear to be reciprocated by critical improvements in model fit.
+Finally, several observations were missing from the sample for these
+covariates. To include them in our model would have required sub-models
+for the missing data, resulting in added complexity and computational
+burden that, again, did not appear to be worth the added complexity and
+computational burden. Hence, the progression of models presented in this
+supplement was intended to efficiently convey the improvements in fit
+following the addition of key covariates and parameterizations that
+ultimately comprised the final model.
+
+The first regression model we fit was constructed to estimate the joint
+distribution of log-transformed dissolved and equilibrium N2O
+conditional only on the the survey design factors, such that:
+
+$$Y \sim MVN(\nu, \Sigma)$$
+
+Indexing to each response, $p$:
+
+$$Y = [y_{p=1}, y_{p=2}] \sim MVN(\nu = [\mu_{p=1}, \mu_{p=2}], \Sigma)$$
+
+Indexing to the observation level, $i$, within each response:
+
+$$\mu_{pijkl} = \alpha_{0(pi)} + \alpha_{1(pij)} + \alpha_{2(pijk)} + \alpha_{3(pijkl)}$$
+
+$$\alpha_1 \sim MVN(0, \Lambda_1)$$ 
+$$\alpha_2 \sim MVN(0, \Lambda_2)$$
+$$\alpha_3 \sim MVN(0, \Lambda_3)$$
+
+$$\Lambda_{\alpha = 1,..,3} = \begin{pmatrix} 1 & \tau ^2_{p=1} \\ \tau^2_{p=2} & 1 \end{pmatrix} \chi_{\alpha = 1,..,3} \begin{pmatrix} 1 & \tau^2_{p=1} \\ \tau^2_{p=2} & 1 \end{pmatrix}$$
+
+$$\chi_{\alpha = 1,..,3} = \begin{pmatrix} 1 & \varrho_{p1,p2} \\ \varrho_{p1,p2} & 1 \end{pmatrix}$$
+
+$$\Sigma = \begin{pmatrix} 1 & \sigma^2_{p=1} \\ \sigma^2_{p=2} & 1 \end{pmatrix} \Omega \begin{pmatrix} 1 & \sigma^2_{p=1} \\ \sigma^2_{p=2} & 1 \end{pmatrix}$$
+$$\Omega = \begin{pmatrix} 1 & \rho_{p1,p2} \\ \rho_{p1,p2} & 1 \end{pmatrix}$$
+
+Each log-transformed observation, $i \in 1,..,N=984$, for each N2O
+response variable, $p \in 1:P=2$, was assumed to be drawn from a
+multivariate normal (MVN) distribution with the parameters $\nu$ and
+$\Sigma$, where $\nu$ is the multivariate mean and $\Sigma$ is a
+covariance matrix containing the observation-level variances and
+(residual) correlations.
+
+The multivariate mean was defined by a vector of location parameters,
+$\nu = [\mu_{p=1}, \mu_{p=2}]$, for each response, $p$. As noted above,
+each location parameter was also defined according to a linear
+combination of parameters indexed to each response $p$ and observation
+$i$. This linear combination included a fixed global intercept,
+$\alpha_0$, conditioned on the data, and three latent group-level
+effects matrices, $\alpha_1, \alpha_2, \alpha_3$. The effects in each of
+these matrices were also drawn from MVN probability distributions
+centered on zero in $p$-dimensional space. The spread of the effects
+around zero were determined by a corresponding variance-covariance
+matrix, $\Lambda_1, \Lambda_2, \text{or } \Lambda_3$. The standard
+deviation parameters, $\tau$ , of those matrices were conditioned on the
+data and captured group-level standard deviations that controlled the
+spread of the group-level effects for each response. The $\chi$
+parameter contained the group-level residual correlation matrix, wherein
+$\varrho$ denotes the group-level residual correlations (i.e., between
+responses).
+
+The explicit indexing in the notation above conveys the relationship
+between the parameters and each observation, $i$, and emphasizes the
+nested structure of the observations within the group effects.
+Specifically, each observation, $i$, was nested in a lake size category,
+$l$, which was nested in a state, $k$, and ecoregion, $j$. The
+group-level effects in $\alpha_1$, therefore, accounted for
+ecoregion-level deviations from the global mean; $\alpha_2$ accounted
+for state-level group effects nested in ecoregions; and $\alpha_3$
+accounted for lake size category effects within states and ecoregions.
+
+Finally, the observation-level covariance term, $\Sigma$, comprised of
+the observation-level standard deviations, $\sigma$, for each response,
+and $\Omega$ or the observation-level residual correlation matrix,
+wherein $\rho$ captured residual correlation between responses.
+
+For model fitting, priors were needed for all parameters conditioned
+directly on the data, which included the global intercepts, $\alpha_0$;
+the scale parameters of the latent group effects, $\tau$ and $\sigma$;
+and the correlation parameters, $\varrho$ and $\rho$. A normal or
+Gaussian prior, $N(\mu = 2, \sigma = 1)$ centered near the (log-scale)
+data means, was used as the prior for the global mean parameter for each
+response. This prior was considered minimally informative as it placed
+most (\~80%) of the prior mass over values between about 2 and 27 ng/L
+for median N2O and median N2O equilibrium concentration; and included
+support in the tails for values approaching 0 ng/L on the lower end and
+80 ng/L on the high end. We placed $Exp(2)$ priors over all scale
+parameters, which placed most of the support between values very close
+to 0 and values near 1 (central 80% density interval from approximately
+0.005 to 1.15). Finally, for the correlation matrices, an $LKJ(\eta =2)$
+prior was used, which, for a 2-dimensional response, placed most support
+for correlations between approximately -0.9 and 0.9. This prior seemed
+reasonable as there is no direct causal link between these responses
+that would warrant a very strong correlation . Any residual dependence
+was expected to be conditional or indirect due to, for example, a common
+correlate (e.g., elevation, temperature). For more information on prior
+choice recommendations in Stan, see:
+<https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations>
+
+The $\textbf{brms}$ package [@Burkner_2017] for $\textbf{R}$
+[@R_Core_Team_2021] was used to fit all of the models in a fully
+Bayesian setting. The formula syntax of the $\textbf{brms}$ package is
+similar to the syntax used in the $\textbf{lme4}$ package that is widely
+used to fit mixed effects models in frequentist settings. In either
+package, the linear predictor for $\mu$ described above, but ignoring
+multivariate dependencies, would be expressed as:
+
+$$\sim 1 + (1|WSA9) + (1|WSA9:state) + (1|WSA9:state:size)$$
+
+In the $\textbf{brms}$ package, there is additional functionality and
+formula syntax for multivariate responses, which allows for the varying
+intercepts of the separate linear predictors in a multivariate model to
+be correlated across responses, e.g.,:
+
+$$
+\begin{aligned} 
+  N_2O_{diss} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\
+  N_2O_{equi} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) 
+\end{aligned}
+$$
+
+The above formula syntax indicates that the linear predictor for both
+N2O responses have the same group-level varying effects terms, and that
+the effects associated with each of those terms can be correlated across
+responses. The relevant sections of $\textbf{R}$ code that are helpful
+for describing the first model above using $\textbf{brms}$ syntax are
+below:
+
+```{r model_1_text, eval=FALSE, include=TRUE}
+# Assign separate linear predictors for the dissolved and equilibrium N2O responses
+bf_n2o <- bf(log(n2o) ~ 1 + 
+               (1 | a | WSA9) + 
+               (1 | b | WSA9:state) + 
+               (1 | c | WSA9:state:size_cat),
+             family = gaussian())
+
+bf_n2oeq <- bf(log(n2o_eq) ~ 1 + 
+               (1 | a | WSA9) + 
+               (1 | b | WSA9:state) +
+               (1 | c | WSA9:state:size_cat),
+             family = gaussian())
+
+# Set priors for parameters
+priors <- c(
+  prior(normal(2, 1), class = "Intercept", resp = "logn2o"), # alpha_0
+  prior(exponential(2), class = "sd", resp = "logn2o"), # sd alpha_1..3
+  prior(exponential(2), class = "sigma", resp = "logn2o"), # observation level sd
+  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), 
+  prior(exponential(2), class = "sd", resp = "logn2oeq"),
+  prior(exponential(2), class = "sigma", resp = "logn2oeq"),
+  prior(lkj(2), class = "rescor"), # observation level residual correlation
+  prior(lkj(2), class = "cor") # correlation for group effects
+  )
+
+# Fit the multivariate model
+n2o_mod1 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE), ...)
+```
+
+The first two objects assigned in the above code example are the linear
+predictors assigned to the dissolved and equilibrium N2O responses. The
+second object is a list of priors as defined above. The third object
+results from the `brm()` call to fit the model. In this syntax, the two
+linear predictor objects are "added" along with the
+`set_rescor(rescor = TRUE)` option, which defines whether the residual
+correlation is to be modeled (TRUE) or not (FALSE). Note that, with
+`set_rescor(rescor = FALSE)` and without the correlated group effects
+(i.e., 'a', 'b', 'c'), $\textbf{brms}$ would effectively treat the two
+responses as entirely independent in fitting.
+
+For the remainder of this document, only this simplified syntax is
+presented to describe the model structure. For more information on
+$\textbf{brms}$ functionality and syntax with multivariate response
+models, see the package vignette:
+<https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html>.
+
+The relevant code for the final model used for inference and described
+in the main paper is below:
+
+```{r model_final_text, eval=FALSE, include=TRUE}
+# Assign separate linear predictors for each response: N2O, N2o-eq, Surface temp, and NO3 level.
+
+# N2O
+bf_n2o <- bf(n2o ~ mo(no3_cat) +
+               log_area +
+               surftemp + 
+               mo(no3_cat):log_area +
+               mo(no3_cat):surftemp +
+               (mo(no3_cat) | a | WSA9) + 
+               (mo(no3_cat) | b | WSA9:state) + 
+               (1 | c | WSA9:state:size_cat),
+             shape ~ log_area +
+               mo(no3_cat) +
+               (1 | WSA9) + 
+               (1 | WSA9:state) + 
+               (1 | WSA9:state:size_cat),
+             family = Gamma(link = "log"))
+
+# Equilibrium N2O
+bf_n2oeq <- bf(n2o_eq ~ surftemp +
+                 log_elev +
+                 surftemp:log_elev +
+                 (1 | a | WSA9) + 
+                 (1 | b | WSA9:state) +
+                 (1 | c | WSA9:state:size_cat),
+             shape ~ surftemp +
+               log_elev +
+               (1 | WSA9) + 
+               (1 | WSA9:state) + 
+               (1 | WSA9:state:size_cat),
+             family = Gamma(link = "log"))
+
+# Surface temperature
+bf_temp <- bf(surftemp ~ lat +
+                s(log_elev) +
+                s(jdate) +
+                (1 | a | WSA9) + 
+                (1 | b | WSA9:state) +
+                (1 | c | WSA9:state:size_cat),
+              shape ~ lat,
+              family = Gamma(link = "log"))
+
+# NO3 level
+bf_no3 <- bf(no3_cat ~ surftemp +
+               log_area +
+               (1 | a | WSA9) +
+               (1 | b | WSA9:state) +
+               (1 | c | WSA9:state:size_cat),
+             family = cumulative(link = "logit", threshold="flexible"))
+
+# Assign priors 
+priors <- c( # N2O
+  prior(normal(2, 1), class = "Intercept", resp = "n2o"),
+  prior(normal(0, 1), class = "b", resp = "n2o"),
+  prior(exponential(2), class = "sd", resp = "n2o"),
+  prior(normal(5, 4), class = "Intercept", dpar = "shape", resp = "n2o"),
+  prior(normal(0, 1), class = "b", dpar = "shape", resp = "n2o"),
+  prior(exponential(2), class = "sd", dpar = "shape", resp = "n2o"),
+
+  # Equil N2O
+  prior(normal(2, 1), class = "Intercept", resp = "n2oeq"), 
+  prior(normal(0, 1), class = "b", resp = "n2oeq"),  
+  prior(exponential(2), class = "sd", resp = "n2oeq"),
+  prior(normal(5, 4), class = "Intercept", dpar = "shape", resp = "n2oeq"),
+  prior(normal(0, 1), class = "b", dpar = "shape", resp = "n2oeq"),
+  prior(exponential(2), class = "sd", dpar = "shape", resp = "n2oeq"),
+
+  # Surface temp
+  prior(normal(3, 1), class = "Intercept", resp = "surftemp"), 
+  prior(normal(0, 1), class = "b", resp = "surftemp"), 
+  prior(exponential(0.5), class = "sds", resp = "surftemp"),
+  prior(exponential(2), class = "sd", resp = "surftemp"),
+  prior(normal(5, 4), class = "Intercept", dpar = "shape", resp = "surftemp"),
+  prior(normal(0, 1), class = "b", dpar = "shape", resp = "surftemp"),
+
+  # NO3
+  prior(normal(0, 3), class = "Intercept", resp = "no3cat"),
+  prior(normal(0, 1), class = "b", resp = "no3cat"),
+  prior(exponential(1), class = "sd", resp = "no3cat"),
+
+  prior(lkj(2), class = "cor")
+  )
+
+# Fit the multivariate model
+n2o_mod6 <- brm(bf_n2o + bf_n2oeq + bf_temp + bf_no3 + set_rescor(rescor = FALSE), ...)
+```
+
+In this final model, the dissolved and equilibrium N2O responses were
+each modeled with Gamma distributed errors. The Gamma error structure
+appeared to result in slightly better performance in the predictive
+checks compared to the Gaussian errors. This was primarily apparent in
+the saturation ratio checks, which may have been more sensitive to model
+performance in the tails of the N2O responses. Others have indicated
+that the Gamma error distribution can work well for modeling dissolved
+N2O data [@Webb_etal_2019]. The linear predictor for dissolved N2O
+included the survey design effects along with covariate terms for lake
+area (ha), NO3 level, and surface water temperature. To accomodate
+heterogeneous errors, the shape parameter of the Gamma distribution was
+modeled as a function of NO3 level, lake area, and the survey design
+factors. The equilibrium N2O response was modeled as a function of the
+survey design factors, lake elevation, and lake surface temperature. The
+shape parameter was, again, allowed to vary conditional on the survey
+factors, surface temperature, and elevation.
+
+The sub-model for the surface temperature response also assumed a Gamma
+distributed error distribution and covariates included the survey design
+variables, latitude, elevation, and Julian date. The shape parameter was
+also modeled as a function of latitude to address increasing temperature
+variance along the latitudinal gradient.
+
+The NO3 sub-model was a cumulative logit model and the linear predictor
+included all of the survey factors as well as surface temperature and
+lake area.
+
+Note that there was no observation-level residual correlation term for
+this final model, since the residuals are undefined for the Gamma and
+cumulative logit models. Dropping the observation-level residual
+correlation term was deemed a reasonable compromise that enabled the
+inclusion of NO3 as a covariate on dissolved N2O. The random intercepts,
+however, still allowed for potential correlations between the four
+responses at the group levels. As mentioned previously, including
+covariates that were not available in the full POI necessitated some
+additional steps in the poststratification process. To make
+population-level inferences from a model with covariates, those
+covariates need to be (1) fully observed across the POI or (2) their
+"missingness" needs to be addressed with modeling. For the lake area and
+elevation covariates, for example, data was available for the entire POI
+from previously compiled geospatial databases. However, neither surface
+temperature or NO3 were observed for lakes outside of the sample. That
+is, they were partially observed with respect to the POI. Therefore, the
+final multivariate model, coded above included these as additional
+responses, with each also being conditioned on the survey design
+structure and the fully observed covariates. The dependence structure
+among the fully and partially observed variables in this model could be
+described such that:
+
+$$\begin{aligned} 
+{p( \boldsymbol{N_2O_{diss}}} &| Survey, Area, {\boldsymbol{NO_3}}, {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{N_2O_{equil}}} &| Survey, Elev, {\boldsymbol{Temp}})\\
+{p( \boldsymbol{NO_3}} &| Survey, Area, {\boldsymbol{Temp}}) \\ 
+{p( \boldsymbol{Temp}} &| Survey, Lat, Elev, Day)
+\end{aligned}$$
+
+Variables in bold text above were partially observed with respect to the
+POI (i.e., observed only in the sample), whereas variables not in bold
+were considered fully observed. The partially observed variables, being
+dissolved and equilibrium N2O, NO3, and surface temperature, were each
+modeled conditional on the survey design variables and other partially
+and/or fully observed covariates. This piece-wise approach then required
+a more complex set of post-processing steps compared to a typical MRP
+analysis. In order to propagate uncertainty through to the
+population-level inferences, the fitted model was first used to predict
+surface temperature in the target population, since it depended only on
+the fully observed covariates. That predictive distribution was then
+used alongside the relevant fully observed covariates to predict NO3 in
+the target population. Finally, these predictive distributions for
+temperature and NO3 were used to predict the N2O responses. These steps
+were carried out in the "Predict to population" section below. This
+approach is similar to MRP approaches with "non-census" variables
+[@Kastellec_etal_2015]. It could also be described as a form of Bayesian
+structural equation model [@Merkle_etal_2021; @Merkle_Rosseel_2018].
+
+Finally, because N2O flux was to be later estimated as a function of
+lake-level measures (*e.g.*, surface area), predictions for N2O were
+made to each of the 465,897 individual lakes in the POI. For population
+level estimates of dissolved N2O, equilibrium N2O, and N2O saturation,
+the lake-level predictions were summarized across the relevant units
+(*e.g.*, CONUS, ecoregion). Predictions were assumed relevant for
+average conditions during the biological index period of 2017.
+
+All of the models were constructed and assessed within in `R`
+[@R_Core_Team_2021] using the `brms` package [@Burkner_2017] as an
+interface to Stan, a software package for fitting fully Bayesian models
+via Hamiltonian Monte Carlo [HMC, @Stan_Development_Team_2018_a;
+@Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c]. More
+specific details on the model structures and fits is provided in the
+"Model fitting" section below.
 
 # Data
-As explained in a previous data munging document document
-(https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html), duplicate dissolved gas samples were collected at a depth of ~0.1m at designated index sites distributed across 1091 lakes nationwide, of which 95 were sampled twice as repeat visits.
-This subset of revisit sites was used as a test set for assessing model fit and out-of-sample performance.
-
-Water samples were analyzed via gas chromotography and concentrations were recorded to the nearest 0.001 nmol/L.
-As part of the survey design, each gas observation was indexed to an individual lake selected with unequal probability from 5 different lake size categories, $j \in j=1,...,J = 5$, according to surface area (ha), and from within a state, $k \in k=1,...,K = 48$, situated within an aggregated, WSA9 or Omernik ecoregion, $l \in l=1,...,L = 9$.
-All WSA9 ecoregions were represented, including the Xeric (XER), Western Mountain (WMT),  Northern Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and Southern Appalachian (SAP) regions.
-Gas data from the first and second (follow-up) visits were separated, with $n=984$ _vs._ $n=95$ observations each.
-In addition to the design variables, a number of potentially useful covariates were also indexed to the gas observations, including measures of NO3, surface water temperature, elevation, chlorophyll a content, dissolved oxygen content, and waterbody size (hectares).
-These gas data and covariates were previously described and munged at: https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html.
+
+As explained in the data munging document document
+(<https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dgIndicatorAnalysis.html>),
+duplicate dissolved gas samples were collected at a depth of \~0.1m at
+designated index sites distributed across 1091 lakes nationwide, of
+which 95 were sampled twice as repeat visits. This subset of revisit
+sites was used as a test set for assessing model fit and out-of-sample
+performance.
+
+Water samples were analyzed via gas chromotography and concentrations
+were recorded to the nearest 0.001 nmol/L. As part of the survey design,
+each gas observation was indexed to an individual lake selected with
+unequal probability from 5 different lake size categories,
+$j \in j=1,...,J = 5$, according to surface area (ha), and from within a
+state, $k \in k=1,...,K = 48$, situated within an aggregated, WSA9 or
+Omernik ecoregion, $l \in l=1,...,L = 9$. All WSA9 ecoregions were
+represented, including the Xeric (XER), Western Mountain (WMT), Northern
+Plains (NPL), Southern Plains (SPL), Temperate Plains (TPL), Coastal
+Plains (CPL), Upper Midwest (UMW), Northern Appalachian (NAP), and
+Southern Appalachian (SAP) regions. Gas data from the first and second
+(follow-up) visits were separated, with $n=984$ *vs.* $n=95$
+observations each. In addition to the design variables, a number of
+potentially useful covariates were also indexed to the gas observations,
+including measures of NO3, surface water temperature, elevation,
+chlorophyll a content, dissolved oxygen content, and waterbody size
+(hectares). These gas data and covariates were previously described and
+munged at:
+<https://github.com/USEPA/DissolvedGasNla/blob/master/scripts/dataMunge.html>.
 
 ## Import
+
 That originally munged gas dataset was imported below.
@@
 
-A new data frame for modeling was constructed from the original file, including only the variables of interest: (1) the N2O gas observations; (2) the survey design variables indexed to those observations; and (3) additional covariates considered potentially useful for improving the fit and/or practical interpretation of the model.
-The modeling data frame below excluded the second-visit observations, which would later be used for model checking.
-Some variables from the imported data were renamed for convenience.
-In addition, the NO3 covariate was rounded according to the documented measurement precision and an alternative version was created by log-transforming and re-coding the variable as an ordered factor with five levels.
-The left-most cut point separated observations below the detection limit from the completely observed samples.
-The remaining cut points were drawn at approximately equal distances in the positive direction along the log scale.
-Finally, one lake had no N2O gas information and was removed from the data frame.
+A new data frame for modeling was constructed from the original file,
+including only the variables of interest: (1) the N2O gas observations;
+(2) the survey design variables indexed to those observations; and (3)
+measures and metrics considered potentially useful as covariates for
+improving the fit and/or practical interpretation of the model. The
+modeling data frame below excluded the second-visit observations, which
+would later be used for model checking. Some variables from the imported
+data were renamed for convenience. In addition, the NO3 covariate was
+rounded according to the documented measurement precision and an
+alternative version was created by log-transforming and re-coding the
+variable as an ordered factor with five levels. The left-most cut point
+separated observations below the detection limit from the completely
+observed samples. The remaining cut points were drawn at equal distances
+in the positive direction along the log scale.
 
@@
 dg %>%
-  filter(!is.na(dissolved.n2o.nmol)) %>% # 1 obs with missing N2O measurement
-  nrow() # number of observations before filtering
+  filter(sample.source == "DG") %>%
+  nrow() # number of dissolved gas observations before filtering out probability sites and second visits
 ```
@@
 df_model <- dg %>%
-  filter(!is.na(dissolved.n2o.nmol)) %>%
-  filter(sitetype == "PROB") %>% # probability samples only
+  filter(sample.source == "DG", # dissolved gas info
+         sitetype == "PROB") %>% # probability samples only
   filter(visit.no == 1) %>% # first-visit sites only
@@
 
-A preview showing the first 10 rows of the modeling data:
+A preview of the `r nrow(df_model)` probability samples to be used in
+modeling:
 
 ```{r model_data_print, echo=TRUE, paged.print=TRUE}
-nrow(df_model) # number of obs after filtering
-
 df_model %>%
-  head(10)
-```
-
-A second data frame, including only the second visit observations was constructed below.
-These data were later used as a "test set" to assess the out-of-sample performance of the model developed on the first-visit or "training set".
-
-```{r all_data_n_2, echo=TRUE, paged.print=TRUE}
-# number of observations before filtering probability samples
-dg %>%
-  filter(!is.na(dissolved.n2o.nmol)) %>% # remove obs with missing response measurements
-  nrow()
-```
+  head()
+```
+
+A second data frame, including only the second visit observations was
+constructed below. These data were later used as a "test set" to assess
+the out-of-sample performance of the model developed on the first-visit
+or "training set".
 
@@
 df_test <- dg %>%
-  filter(!is.na(dissolved.n2o.nmol)) %>%
-  filter(sitetype == "PROB") %>% # probability samples only
+  filter(sample.source == "DG", # dissolved gas info
+         sitetype == "PROB") %>% # probability samples only
   filter(visit.no == 2) %>% # second-visit sites only
@@
 
-A preview of the first 10 rows of the test data:
+A preview of the re-visit data with `r nrow(df_test)` observations:
 
 ```{r test_data_print, echo=TRUE, paged.print=TRUE}
-nrow(df_test) # number of obs after filtering for probability samples, first visits, and removing one site missing ecoregion (WSA9) info.
-
 df_test %>%
-  head(10)
+  head()
 ```
@@
 
-Below. the original NLA sampling frame was imported and filtered to include only lakes in the POI. The resulting target population included a total of 465,897 waterbodies.
+Below, the 2017 NLA sampling frame was imported and filtered to include
+only lakes in the assumed POI. The resulting target population included
+a total of 465,897 waterbodies.
 
@@
 
-Below is a sample of the first ten rows describing the POI:
+Below is a sample of the first ten rows of the POI:
 
@@
 
-Cross tabulations below show the structure of the target population with respect to the survey design variables.
-The cross-tabulation shows that each ecoregion does not contain each state.
-Therefore, states were nested within ecoregions.
+Cross tabulations below show the structure of the target population with
+respect to the survey design variables. The cross-tabulation shows that,
+of course, each ecoregion does not contain each state. Therefore, states
+were nested within ecoregions.
 
@@
   summarise(n = n(), .groups = "drop") %>%
-  spread(state, n) %>%
-  head(10)
-```
-
-The cross-tablulation below indicates that lake size category was nested within state (which was nested in ecoregion). That is, not every ecoregion:state contained every size category.
+  spread(state, n) 
+```
+
+The cross-tablulation below indicates that lake size category was nested
+within state (which was nested in ecoregion). That is, not every
+ecoregion:state contained every size category.
 
@@
   summarise(n = n(), .groups = "drop") %>%
-  spread(size_cat, n) %>%
-  head(10)
-```
-
-Below, the sampling frame was used to create a typical post-stratification table.
-There were 536 "types" or groupings of lakes in the POI with respect to the sampling design.
-The total counts of those lake types (n_lakes) and their proportions (prop_cell) relative to the counts in the target population were tabulated.
+  spread(size_cat, n) 
+```
+
+Below, the sampling frame was used to create a typical
+post-stratification table. There were 536 "types" or groupings of lakes
+in the POI with respect to the sampling design. The total counts of
+those lake types (n_lakes) and their proportions (prop_cell) relative to
+the counts in the target population were tabulated.
 
@@
 
-Below, the distribution of waterbody types in the POI were compared to their proportions in the observed sample.
-There were 352 waterbody types in the sample compared to the 536 in the POI.
-There were 984 observations distributed across these 352 lake types in the sample; and the number of samples was not distributed evenly across the types.
-Some cells were represented by as few as 1 lake.
-In total, 536-352 = 184 lake types in the population of interest were not represented in the sample.
+Below, the distribution of waterbody types in the POI were compared to
+their proportions in the observed sample. There were 352 waterbody types
+in the sample compared to the 536 in the POI. There were 984
+observations distributed across these 352 lake types in the sample; and
+the number of samples was not distributed evenly across the types. Some
+cells were represented by as few as 1 lake. In total, 536-352 = 184 lake
+types in the population of interest were not represented in the sample.
 
@@
 samp_props %>%
-  print()
-```
-
-Below, a graphical comparison was constructed to depict the distribution of cells in the POI _vs._ those in the sample.
+  print(n = 10)
+```
+
+Below, a graphical comparison was constructed to depict the distribution
+of cells in the POI *vs.* those in the sample.
 
@@
 
-Another comparison between population and sample was constructed by ecoregion below.
-Lakes in the Coastal Plains (CPL) ecoregion, for example, were clearly undersampled relative to their proportion of the population.
+Another comparison between population and sample was constructed by
+ecoregion below. Lakes in the Coastal Plains (CPL) ecoregion, for
+example, were clearly undersampled relative to their proportion of the
+population.
 
@@
 
-Finally, a comparison by NLA size category is shown below.
-Waterbodies of the smallest size category were heavily under-sampled relative to the POI and those from the larger size categories were over-sampled.
+Finally, a comparison by NLA size category is shown below. Waterbodies
+of the smallest size category were heavily under-sampled relative to the
+POI and those from the larger size categories were over-sampled.
 
@@
 
-Below are (naive) estimates, based only on the sample, for national means for dissolved and equilibrium N2O and the saturation ratio.
+Below are (naive) estimates, based only on the sample, for national
+means for dissolved and equilibrium N2O and the saturation ratio.
 
@@
   summarise(mean = mean(n2o),
-             sd = sd(n2o)) %>%
-  print()
+             sd = sd(n2o)) 
 ```
@@
   summarise(mean = mean(n2o_eq),
-             sd = sd(n2o_eq)) %>%
-  print()
+             sd = sd(n2o_eq))
 ```
@@
   summarise(mean = mean(n2o / n2o_eq),
-             sd = sd(n2o / n2o_eq)) %>%
-  print()
-```
-
-Roughly 67% of lakes in the sample were under-saturated (i.e., saturation ratio < 1):
+             sd = sd(n2o / n2o_eq)) 
+```
+
+Roughly 67% of lakes in the sample were under-saturated (i.e.,
+saturation ratio \< 1):
 
@@
 df_model %>%
-  summarise(prop_undersat = sum((n2o / n2o_eq) < 1) / 984) %>%
-  print()
-```
-
-Using only the sample observations again, a plot was constructed of the overall mean (dashed line) along with the ecoregion-specific means (black circles).
-The shaded areas indicate +/- 1 standard deviation.
-Neither dissolved N2O nor the saturation ratio were clearly structured by ecoregion in the sample, but there did appear to be some potential structure along this variable in the equilibrium N2O observations.
+  summarise(prop_undersat = sum((n2o / n2o_eq) < 1) / 984) 
+```
+
+Using only the sample observations again, a plot was constructed of the
+overall mean (dashed line) along with the ecoregion-specific means
+(black circles). The shaded areas indicate +/- 1 standard deviation.
+Neither dissolved N2O nor the saturation ratio were clearly structured
+by ecoregion in the sample, but there did appear to be some potential
+structure along this variable in the equilibrium N2O observations.
 
@@
 
-Below, the distribution N2O concentrations for the sample was summarized using a density and rug plot.
-Note the natural log scale of the x-axis.
-Both the dissolved and equilibrium N2O data had considerable right skew even after the log transformation.
-This was not unexpected and has been noted in other studies [_e.g._, @Webb_etal_2019].
-The saturation ratio was also skewed (sat_ratio = n2o / n2o_eq).
+Below, the distribution N2O concentrations for the sample was summarized
+using a density and rug plot. Note the natural log scale of the x-axis.
+Both the dissolved and equilibrium N2O data had considerable right skew
+even after the log transformation. This was not unexpected and has been
+noted in other studies [*e.g.*, @Webb_etal_2019]. The saturation ratio
+was also skewed (sat_ratio = n2o / n2o_eq).
 
@@
 
-Below are plots of dissolved N2O vs.
-NO3.
-The first plot shows log(N2O) vs.
-log(NO3), as well as the ordinal categories assigned to NO3 (vertical lines).
-The leftmost vertical line is dashed and separates the NO3 observations below the detection limit.
-The trend is increasing and nonlinear on the log scale, with increasing variance in N2O as NO3 increased.
-
-```{r summary_N2O_vs_NO3, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
+Below are plots of dissolved N2O vs. NO3 and some other potential
+covariates. This first plot shows log(N2O) vs. log(NO3), as well as the
+ordinal categories assigned to NO3 (vertical lines). The leftmost
+vertical line is dashed and separates the NO3 observations below the
+detection limit. The trend is increasing and nonlinear on the log scale,
+with increasing variance in N2O as NO3 increased.
+
+```{r summary_N2O_vs_NO3, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=5, message=FALSE, warning=FALSE}
 df_model %>%
@@
 
-Below are plots of dissolved N2O vs.
-NO3 for 6 quantiles of the surface temperature measurements (quantiles increasing from 1 to 6).
-This plot suggested that the NO3 effect on N2O may have been stronger in lakes with higher observed temperatures.
-
-```{r summary_N2O_vs_NO3_surftemp, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
+Below is a plot of N2O vs. the surface temperature observations.
+
+```{r summary_N2O_vs_surftemp, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  ggplot(aes(x = surftemp, y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  theme_bw()
+```
+
+Below are plots of dissolved N2O vs. NO3 for 3 quantiles of the surface
+temperature measurements (quantiles increasing from 1 to 3). This plot
+suggested that the NO3 effect on N2O may have been stronger in lakes
+with higher observed temperatures.
+
+```{r summary_N2O_vs_NO3_surftemp, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE}
 df_model %>%
@@
   theme(text = element_text(size=12)) +
-  facet_wrap(~ as.factor(ntile(surftemp, 6))) +
+  facet_wrap(~ as.factor(ntile(surftemp, 3)), ncol = 1) +
   theme_bw()
@@
 
-The next plot below shows the relationship between dissolved N2O and NO3 at 6 different quantiles (increasing 1 to 6) of the log-scaled lake surface area estimates.
-
-```{r summary_N2O_vs_NO3_logarea, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
+Below is a plot of N2O vs. the natural log of the surface area data for
+each lake from the NHD. The plot doesn't suggest a strong effect, but
+the pattern of variation suggests more variability in dissolved N2O in
+smaller lakes.
+
+```{r summary_N2O_vs_logarea, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  ggplot(aes(x = log_area, y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  theme_bw()
+```
+
+The next plot below shows the relationship between dissolved N2O and NO3
+at 3 different quantiles (increasing 1 to 3) of the log-scaled lake
+surface area estimates. This plot perhaps suggests a weaker NO3 effect
+in the largest lakes.
+
+```{r summary_N2O_vs_NO3_logarea, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE}
 df_model %>%
@@
   theme(text = element_text(size=12)) +
-  facet_wrap(~ as.factor(ntile(log_area, 6))) +
+  facet_wrap(~ as.factor(ntile(log_area, 3)), ncol = 1) +
   theme_bw()
@@
 
-Similar plots are below, but with NO3 expressed as an ordered categorical variable with 5 levels.
-The positive and monotonic trends area similar to the previous plots where NO3 was treated as continuous.
-Note the large number of observations in the first NO3 category (no3_cat = 1).
-This category represented all of the observations for NO3 that were below the detection limit, which was most of the data.
-
-```{r summary_N2O_vs_NO3cat, echo=FALSE, fig.align='center', fig.height=4, fig.width=6, message=FALSE}
+Below is a plot of N2O vs. the natural log of the dissolved oxygen
+measurements. Measurements for DO were missing from 4 samples and these
+observations were excluded from the plot below. The plot suggests a
+positive effect of DO, but the pattern may be dominated by extreme
+values at the upper and lower end of log(DO).
+
+```{r summary_N2O_vs_logDO, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  filter(!is.na(log_do)) %>%
+  ggplot(aes(x = log_do, y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  theme_bw()
+```
+
+The next plot below shows the relationship between dissolved N2O and NO3
+at 3 different quantiles (increasing 1 to 3) of the log-scaled dissolved
+oxygen measurements. Measurements for DO were missing from 4 samples and
+these observations were excluded from the plot below. This plot doesn't
+suggest a clear interaction of DO with NO3.
+
+```{r summary_N2O_vs_NO3_logDO, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  filter(!is.na(log_do)) %>%
+  ggplot(aes(x = log(no3), y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  facet_wrap(~ as.factor(ntile(log_do, 3)), ncol =1) +
+  theme_bw()
+```
+
+Below is a plot of N2O vs. the natural log of the chlorophyll a
+measurements. Measurements were missing from 4 samples and these
+observations were excluded from the plot. The plot suggests perhaps a
+slight negative effect of chl-a.
+
+```{r summary_N2O_vs_logChla, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  filter(!is.na(log_chla)) %>%
+  ggplot(aes(x = log_chla, y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  theme_bw()
+```
+
+The next plot below shows the relationship between dissolved N2O and NO3
+at 3 different quantiles (increasing 1 to 3) of the log-scaled
+chlorophyll a measurements. Measurements for chl-a were missing from 4
+samples and these observations were excluded from the plot below. There
+is no clear interaction indicated in the plot below, but the NO3 effect
+may be stronger at high chl-a levels.
+
+```{r summary_N2O_vs_NO3_logChla, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  filter(!is.na(log_chla)) %>%
+  ggplot(aes(x = log(no3), y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  facet_wrap(~ as.factor(ntile(log_chla, 3)), ncol = 1) +
+  theme_bw()
+```
+
+Below is a plot of N2O vs. a square root transformation of the bouyancy
+factor (BF) metric. Again, measurements were missing from 4 samples and
+these observations were excluded from the plot. The plot suggests
+perhaps a slight negative effect of the BF metric.
+
+```{r summary_N2O_vs_sqrt_bf, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  filter(!is.na(sqrt_bf)) %>%
+  ggplot(aes(x = sqrt_bf, y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  theme_bw()
+```
+
+The next plot below shows the relationship between dissolved N2O and NO3
+at 3 different quantiles (increasing 1 to 3) of the BF metric. There is
+no clear interaction indicated in the plot below.
+
+```{r summary_N2O_vs_NO3_BF_max, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE}
+df_model %>%
+  filter(!is.na(sqrt_bf)) %>%
+  ggplot(aes(x = log(no3), y = log(n2o))) +
+  geom_point(show.legend = F) +
+  geom_smooth(method = "loess", span = 2) +
+  theme(text = element_text(size=12)) +
+  facet_wrap(~ as.factor(ntile(sqrt_bf, 3)), ncol = 1) +
+  theme_bw()
+```
+
+Similar plots are below, but with NO3 expressed as an ordered
+categorical variable with 5 levels. The positive and monotonic trends
+area similar to the previous plots where NO3 was treated as continuous.
+Note the large number of observations in the first NO3 category (no3_cat
+= 1). This category represented all of the observations for NO3 that
+were below the detection limit, which was most of the data.
+
+```{r summary_N2O_vs_NO3cat, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=5, message=FALSE}
 df_model %>%
@@
 
-```{r summary_N2O_vs_NO3cat_surftemp, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE}
+The plot below is N2O vs NO3 for three quantiles (increasing) of surface
+temperature.
+
+```{r summary_N2O_vs_NO3cat_surftemp, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE}
 df_model %>%
@@
 
-```{r summary_N2O_vs_NO3cat_logarea, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, message=FALSE}
+The plot below is N2O vs NO3 for three quantiles (increasing) of (log)
+surface area.
+
+```{r summary_N2O_vs_NO3cat_logarea, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE}
 df_model %>%
@@
 
-Below is a plot of log(dissolved N2O) vs. log(NO3) by ecoregion, which suggested that the NO3 effect on dissolved N2O may have varied by ecoregion.
-
-```{r summary_N2O_vs_NO3_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
+Below is a plot of log(dissolved N2O) vs. log(NO3) by ecoregion, which
+suggested that the NO3 effect on dissolved N2O may have varied by
+ecoregion.
+
+```{r summary_N2O_vs_NO3_ecoregion, echo=FALSE, fig.align='center', fig.asp = 0.75, fig.width = 8, message=FALSE}
 df_model %>%
@@
 
-Below is the same plot as above but for the ordered categorical version of NO3.
-
-```{r summary_N2O_vs_NO3cat_ecoregion, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, message=FALSE}
+Below is the same plot as above but for the ordered categorical version
+of NO3.
+
+```{r summary_N2O_vs_NO3cat_ecoregion, echo=FALSE, fig.align='center', fig.asp = 0.75, fig.width = 8, message=FALSE}
 df_model %>%
@@
 
-A plot below shows the relationships between dissolved N2O and log(NO3) by state, but within just the Temperate Plains (TPL) ecoregion.
-Within states, the number of observations were relatively small, but the trends were perhaps more linear.
-
-```{r summary_N2O_vs_NO3_wsa9state3, echo=FALSE, fig.align='center', fig.height=4, fig.width=8, message=FALSE}
+A plot below shows the relationships between dissolved N2O and log(NO3)
+by state, but within just the Temperate Plains (TPL) ecoregion. Within
+states, the number of observations were relatively small, but the trends
+were perhaps more linear.
+
+```{r summary_N2O_vs_NO3_wsa9state3, echo=FALSE, fig.align='center', fig.asp = 0.75, fig.width = 8, message=FALSE}
 df_model %>%
@@
 
-# Model fitting
-The first regression model was constructed to estimate the joint distribution of log-transformed dissolved and equilibrium N2O conditional only on the the survey design factors.
-Each log-transformed observation, $i \in 1,..,N=984$, for each response, $p \in 1:P=2$, was assumed to be drawn from a multivariate normal distribution with the parameters $\nu$ and $\Sigma$, where $\nu$ is the multivariate mean estimated conditional on the design effects and $\Sigma$ is a covariance matrix containing the observation-level variances and residual correlation:
-$$Y \sim MVN(\nu, \Sigma)$$
-
-The multivariate mean is a vector of location parameters, $\nu:[\mu_{p=1}, \mu_{p=2}]$, for each response.
-Each location parameter was further defined by a linear combination of parameters where, for each response $p$ and observation $i$:
-
-$$\mu_{pi} = \alpha_{0(pi)} + \alpha_{1(pij)} + \alpha_{2(pijk)} + \alpha_{3(pijkl)} \\
-\alpha_1 \sim MVN(0, \Lambda_1) \\
-\alpha_2 \sim MVN(0, \Lambda_2) \\
-\alpha_3 \sim MVN(0, \Lambda_3)$$
-
-The linear combination included a fixed global intercept, $a_0$, estimated directly from the data, and three separate, latent group-level effects matrices, $\alpha_1, \alpha_2, \alpha_3$.
-The group effects were assumed to be multivariate normal and were centered on zero in the two-dimensional multivariate space.
-The spread of the effects around zero were determined by a covariance matrix, $\Lambda_1, \Lambda_2, \text{or } \Lambda_3$, which were estimated directly from the data.
-The covariance terms were further defined where:
-
-$$\Lambda = \begin{pmatrix} 1 & \tau^2_{p=1} \\ \tau^2_{p=2} & 1 \end{pmatrix} \chi \begin{pmatrix} 1 & \tau^2_{p=1} \\ \tau^2_{p=2} & 1 \end{pmatrix}$$
-
-The $\tau$ parameters captured the group-level standard deviations, which constrain the spread of group-level effects for each response, and $\chi$ was the group-level residual correlation matrix:
-
-$$\chi = \begin{pmatrix} 1 & \varrho \\ \varrho & 1 \end{pmatrix}$$
-
-wherein $\varrho$ captured the group-level residual correlation between effects.
-
-The explicit indexing in the notation above conveys the relationship between the parameters and each observation, $i$, and emphasizes the nested structure of the observations within the group effects.
-Specifically, each observation, $i$, was nested in a lake size category, $l$, which was nested in a state, $k$, and ecoregion, $j$.
-The parameter $\alpha_1$, therefore, accounted for ecoregion-scale group effects or deviations from the global mean; $\alpha_2$ accounted for state-level group effects nested in ecoregions; and $\alpha_3$ accounted for lake size group effects within states and ecoregions.
-
-Finally, the observation-level covariance term, $\Sigma$, was parameterized as:
-$$\Sigma = \begin{pmatrix} 1 & \sigma^2_{p=1} \\ \sigma^2_{p=2} & 1 \end{pmatrix} \Omega \begin{pmatrix} 1 & \sigma^2_{p=1} \\ \sigma^2_{p=2} & 1 \end{pmatrix}$$
-
-wherein the $\sigma$ parameters were the observation-level standard deviations for each response and $\Omega$ was the observation-level residual correlation matrix:
-$$\Omega = \begin{pmatrix} 1 & \rho \\ \rho & 1 \end{pmatrix}$$
-wherein $\rho$ captured the residual correlation between responses.
-
-For model fitting, priors were needed for all parameters conditioned directly on the data, which included the global intercept, the scale parameters, and the correlation matrices.
-A normal or Gaussian prior, $N(\mu = 2, \sigma = 1)$ centered near the (log-scale) data means, was used for the global mean parameter for each response.
-This prior was considered minimally informative as it placed most (~80%) of the prior mass over values between approximately 2 and 27 ng/L for median N2O or N2O equilibrium concentration and included support in the tails for values approaching 0 ng/L on the lower end and 80 ng/L on the high end.
-We placed $Exp(2)$ priors over all scale parameters, which placed most of the support between values very close to 0 and values near 1 (central 80% density interval from approximately 0.005 to 1.15).
-Finally, for the correlation matrices, an $LKJ(\eta =2)$ prior was used, which, for a 2-dimensional response, placed most support for correlations between approximately -0.9 and 0.9.
-This prior seemed reasonable as there were no clear causal mechanisms to ensure a very strong correlation between the N2O measures.
-Any potential residual dependence was expected to be indirect due to, for example, a common correlate (e.g., elevation, temperature).
-For more information on prior choice recommendations in Stan, see: https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
-
-The $\textbf{brms}$ package [@Burkner_2017] for $\textbf{R}$ [@R_Core_Team_2021] was used to fit all of the models in a fully Bayesian setting.
-The formula syntax of the $\textbf{brms}$ package is similar to the syntax used in the $\textbf{lme4}$ package that is widely used to fit mixed effects models in frequentist settings.
-In either package, the linear predictor for $\mu$ described above could be expressed as:
-
-$$\sim 1 + (1|WSA9) + (1|WSA9:state) + (1|WSA9:state:size)$$
-
-In the $\textbf{brms}$ package, there is additional functionality and formula syntax for multivariate responses and for allowing the varying intercepts in a multivariate model to be correlated, e.g.,:
-
-$$
-\begin{aligned} 
-  N_2O_{diss} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\
-  N_2O_{equi} \sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) 
-\end{aligned}
-$$
-
-The above formula syntax would indicate that the linear predictor for both responses in the multivariate model have the same group-level varying effects, and that those effects may be correlated between responses.
-
-For the remainder of this document, only this simplified syntax is presented to describe the model structure.
-For more information on $\textbf{brms}$ functionality and syntax with multivariate response models, the package vignette may be helpful, and can be found at: https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html.
+# Model Fitting
 
@@
 
-The first model fit.
+The first model for the two N2O responses, conditioned only on the study design factors.
 
@@
 
-The summaries of the estimated parameters and key HMC convergence diagnostics for the fitted model are printed below.
-There were no obvious issues with the HMC sampling.
-All $\hat{R}$ values were less than 1.01 and effective sample size ($ESS$) calculations suggested that the posterior contained a sufficient number of effective samples for conducting inference.
+The summaries of the estimated parameters and key HMC convergence
+diagnostics for the fitted model are printed below. There were no
+apparent issues with the HMC sampling. All $\hat{R}$ values were less
+than 1.01 and effective sample size ($ESS$) calculations suggested that
+the posterior contained a sufficient number of effective samples for
+conducting inference.
 
@@
 
-In the summary above, the estimated standard deviations for the varying group effects on the mean behavior of the dissolved N2O response suggested fairly low, but non-zero variability across each of the three levels.
-The standard deviations estimated for the same varying effects for equilibrium N2O were also relatively small.
-Finally, note the relatively small, but positive residual correlation between the two N2O responses.
-
-Before investing too much into inferences from this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch.
-11].
+In the summary above, the estimated standard deviations for the varying
+group effects on the mean behavior of the dissolved N2O response
+suggested fairly low, but non-zero variability across each of the three
+levels. The standard deviations estimated for the same varying effects
+for equilibrium N2O were also relatively small. Finally, note the
+relatively small, but positive residual correlation between the two N2O
+responses.
+
+Before investing too much into inferences from this model, however, the
+model fit was evaluated below using a series of graphical posterior
+predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch. 11].
 
@@
 
-Below are a series of panels illustrating graphical PPCs for the dissolved N2O component of the multivariate model.
-The top left panel compares a density plot of the observed data (black line) to density estimates drawn for 200 samples from the posterior predictive distribution (PPD; blue lines) of the fitted model.
-The top right panel compares the cumulative density distributions in the same manner.
-The left middle panel compares means _vs._ standard deviations for 1000 draws from the PPD (blue dots) to the sample mean _vs._ standard deviation (black dot).
-The right middle panel compares skewness _vs._ kurtosis for 1000 draws from the PPD to the skewness _vs._ kurtosis estimates for the observed data.
-The bottom left panel compares max _vs._ min values for 1000 draws from the PPD to the max _vs._ min values of the sample data.
-Finally, the bottom right panel shows the observed _vs._ average predicted values for each observation in the sample.
-The average predicted values were calculated as the mean prediction for each observation in the PPD based on 1000 draws.
+Below are a series of panels illustrating graphical PPCs for the
+dissolved N2O component of the multivariate model. The top left panel
+compares a density plot of the observed data (black line) to density
+estimates drawn for 200 samples from the posterior predictive
+distribution (PPD; blue lines) of the fitted model. The top right panel
+compares the cumulative density distributions in the same manner. The
+left middle panel compares means *vs.* standard deviations for 1000
+draws from the PPD (blue dots) to the sample mean *vs.* standard
+deviation (black dot). The right middle panel compares skewness *vs.*
+kurtosis for 1000 draws from the PPD to the skewness *vs.* kurtosis
+estimates for the observed data. The bottom left panel compares max
+*vs.* min values for 1000 draws from the PPD to the max *vs.* min values
+of the sample data. Finally, the bottom right panel shows the observed
+*vs.* average predicted values for each observation in the sample. The
+average predicted values were calculated as the mean prediction for each
+observation in the PPD based on 1000 draws.
 
@@
          type = "dens_overlay",
-         ndraws = 200,
-         cores = 1) + 
+         ndraws = 200) + 
   theme_tidybayes() +
@@
          type = "ecdf_overlay",
-         ndraws = 200, 
-         cores = 1) + 
+         ndraws = 200) + 
   theme_tidybayes() +
@@
          stat = c("mean", "sd"),
-         ndraws = 1000,
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
          stat = c("kurtosis", "skewness"), 
-         ndraws = 1000,
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
          stat = c("min", "max"), 
-         ndraws = 1000,
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
          type = "scatter_avg",
-         ndraws = 1000,
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
 
-The general takeaway from the PPCs above was that the model replicated the central tendency of the observed data fairly well, but failed to sufficiently replicate other important aspects of the distribution, such as skewness and kurtosis.
-The observed _vs._ average predictions scatterplot also suggested substantial heteroscedasticity in the errors.
-
-The same checks were run below, but for the test set of 95 held-out, second-visit data points. The patterns in misfit were similar to the patterns indicated in the PPCs with the training data above.
+The general takeaway from the PPCs above was that the model replicated
+the central tendency of the observed data fairly well, but failed to
+sufficiently replicate other important aspects of the distribution, such
+as skewness and kurtosis. The observed *vs.* average predictions
+scatterplot also suggested substantial heteroscedasticity in the errors.
+
+The same checks were run below, but for the test set of 95 held-out,
+second-visit data points. The patterns in misfit were similar to the
+patterns indicated in the PPCs with the training data above.
 
@@
 #### Equilibrium N2O
-Below are PPCs for the equilibrium N2O component of the model.
-As with the dissolved N2O PPCs, the model seemed to do an OK job at replicating the central tendency, but performed less well at replicating other important aspects of the overall distribution.
+
+Below are PPCs for the equilibrium N2O component of the model. As with
+the dissolved N2O PPCs, the model seemed to do an OK job at replicating
+the central tendency, but performed less well at replicating other
+important aspects of the overall distribution.
 
@@
          stat = c("kurtosis", "skewness"),
-         ndraws = 1000,
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
          type = "scatter_avg", 
-         ndraws = 1000,
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
 #### Bivariate
-The graphical check below compared bivariate density contours estimated from the observed data (black lines) to density contours estimated for each of 20 draws from the PPD.
-The model appeared to do a good job of replicating the bivariate mean, but was poor at representing the overall joint distribution.
-
-```{r ppc_biv1, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
+
+The graphical check below compared bivariate density contours estimated
+from the observed data (black lines) to density contours estimated for
+each of 20 draws from the PPD. The model appeared to do a good job of
+replicating the bivariate mean, but was poor at representing the overall
+joint distribution.
+
+```{r ppc_biv1, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5,  message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 #### Saturation
-The graphical PPCs below were aimed at evaluating how well the multivariate model did at representing the observed saturation ratio:
+
+The graphical PPCs below were aimed at evaluating how well the
+multivariate model did at representing the observed saturation ratio:
 
@@
 
-This quantity was estimated as a derived variable by dividing the dissolved N2O PPD by the equilibrium N2O PPD.
-The proportion of under-saturated lakes in the sample was estimated by summing the number of lakes from each posterior predictive draw wherein the ratio was < 1 and dividing that number by the total number of lakes in the sample, which was 984.
-Overall, these checks indicated that properly representing the tails of the dissolved N2O and N2O-eq observations would likely be necessary in order to better replicate the observed saturation metrics.
-The observed proportion of under-saturated lakes was underestimated by more than 10 percentage points, on average.
-
-The top left panel, below, is a density plot of the observed saturation ratio (black line) compared to an estimate using 50 draws from the derived PPD (blue lines).
-The top right panel shows the observed proportion of under-saturated lakes compared to a model estimate based on 1000 draws from the PPD.
-The left middle panel shows the mean _vs._ standard deviation of the saturation ratio for the observed data compared to the same estimates for 500 draws from the PPD.
-The right middle panel shows the max _vs._ min for the sample compared to 500 draws from the PPD.
-Finally, the bottom left panel shows the observed _vs._ average predicted saturation ratio for all 984 lakes sampled in the dataset.
+This quantity was estimated as a derived variable by dividing the
+dissolved N2O PPD by the equilibrium N2O PPD. The proportion of
+under-saturated lakes in the sample was estimated by summing the number
+of lakes from each posterior predictive draw wherein the ratio was \< 1
+and dividing that number by the total number of lakes in the sample,
+which was 984. Overall, these checks indicated that properly
+representing the tails of the dissolved N2O and N2O-eq observations
+would likely be necessary in order to better replicate the observed
+saturation metrics. The observed proportion of under-saturated lakes was
+underestimated by more than 10 percentage points, on average.
+
+The top left panel, below, is a density plot of the observed saturation
+ratio (black line) compared to an estimate using 50 draws from the
+derived PPD (blue lines). The top right panel shows the observed
+proportion of under-saturated lakes compared to a model estimate based
+on 1000 draws from the PPD. The left middle panel shows the mean *vs.*
+standard deviation of the saturation ratio for the observed data
+compared to the same estimates for 500 draws from the PPD. The right
+middle panel shows the max *vs.* min for the sample compared to 500
+draws from the PPD. Finally, the bottom left panel shows the observed
+*vs.* average predicted saturation ratio for all 984 lakes sampled in
+the dataset.
 
@@
   add_predicted_draws(n2o_mod1, 
-                      ndraws = 50) %>%
+                      ndraws = 200) %>%
   rename(n2o_obs = n2o,
@@
   add_predicted_draws(n2o_mod1, 
-                      ndraws = 500) %>%
+                      ndraws = 1000) %>%
   rename(n2o_obs = n2o,
@@
   group_by(.draw) %>%
-  summarise(prop_pred = sum(usat_pred) / 984) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_model)) %>%
   ggplot(aes(x = prop_pred)) +
   geom_histogram(binwidth = 0.001, fill = "lightblue") +
-  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
+  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_model))) +
   xlab("Proportion undersaturated waterbodies") +
@@
 
-The same PPCs are shown below for the revisit data. These checks indicated that the model did a similarly underwhelming job of replicating the re-visit data.
+The same PPCs are shown below for the revisit data. These checks
+indicated that the model did a similarly underwhelming job of
+replicating the re-visit data.
 
@@
   group_by(.draw) %>%
-  summarise(prop_pred = sum(usat_pred) / 95) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_test)) %>%
   ggplot(aes(x = prop_pred)) +
   geom_histogram(binwidth = 0.001, fill = "lightblue") +
-  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/95)) +
+  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_test))) +
   xlab("Proportion undersaturated waterbodies") +
@@
 
-Below, the Bayesian $R^2$ values are reported for each response in the model.
+Below, the Bayesian $R^2$ values are reported for each response in the
+model.
 
@@
 
-In an attempt to better fit the observed data, the next model included a distributional sub-model that allowed for heterogeneous variances.
-The distributional terms for both N2O responses were each fit as a function of the survey design structure.
-The same structure as for the models for the submodels for the mean component before.
+In an attempt to better fit the observed data, the next model included a
+distributional sub-model that allowed for heterogeneous variances. The
+distributional terms for both N2O responses were each fit as a function
+of the survey design structure. The same structure as for the models for
+the submodels for the mean component before.
 
@@
 
-The summaries of the estimated parameters and key HMC convergence diagnostics for the fitted model are printed below.
+The summaries of the estimated parameters and key HMC convergence
+diagnostics for the fitted model are printed below.
 
@@
 
-Below the same PPCs were performed as with the initial model (see above for more details on each panel).
-Though the checks below suggested some improvement in replicating the tails of the observed data, the overall fit again suggested room for improvement.
+Below the same PPCs were performed as with the initial model (see above
+for more details on each panel). Though the checks below suggested some
+improvement in replicating the tails of the observed data, the overall
+fit again suggested room for improvement.
 
@@
 
+The same check is below for the 95 re-visit sites.
+
+```{r ppc_n2o2_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "dens_overlay",
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlim(-5, 5) +
+  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "ecdf_overlay",
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlim(-5, 5) +
+  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("mean", "sd"),
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness"), 
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("min", "max"), 
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "scatter_avg", 
+         ndraws = 1000) + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### Equilibrium N2O
 
-The checks below suggested that this model offered little no improvement upon the initial model for equilibrium N2O.
+The checks below suggested that this model offered little improvement
+upon the initial model for equilibrium N2O.
 
@@
          type = "ecdf_overlay",
-         ndraws = 200,
-         cores = 1) + 
+         ndraws = 200) + 
   theme_tidybayes() +
@@
          stat = c("mean", "sd"), 
-         ndraws = 1000,
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
+         ndraws = 1000) + 
   theme_tidybayes()
@@
          stat = c("kurtosis", "skewness"),
-         ndraws = 1000,
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
+         ndraws = 1000) + 
   theme_tidybayes()
@@
 
+The same check with the re-visit data is below.
+
+```{r ppc_n2oeq2_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "dens_overlay",
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "ecdf_overlay",
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("mean", "sd"), 
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness"),
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("min", "max"), 
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod2, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "scatter_avg", 
+         ndraws = 1000) + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### Bivariate
 
-This check perhaps suggested an improvement with regard to replicating the joint density. However, the predictions were still clearly over-dispersed relative to the observations.
-
-```{r ppc_biv2, echo=FALSE, fig.align='center', fig.asp=0.75, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+This check perhaps suggested an improvement with regard to replicating
+the joint density. However, the predictions were still clearly
+over-dispersed relative to the observations.
+
+```{r ppc_biv2, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 
+The same check with the re-visit data is below.
+
+```{r ppc_biv2_test, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+df_test %>%
+  add_predicted_draws(n2o_mod2, 
+                      ndraws = 20) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
+  geom_density_2d(aes(x = logn2o, 
+                      y = logn2oeq, 
+                      group = .draw),
+                  bins = 10,
+                  color = "lightblue", 
+                  alpha = 0.4) +
+  geom_density_2d(color = "black", bins = 10) +
+  xlim(1.25, 2.75) +
+  ylim(1.25, 2.75) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+```
+
 #### Saturation
 
-The PPCs for the saturation metrics below indicated that including the distributional models was perhaps an improvement on the initial model in some aspects; in particular, the bias in the predicted proportion of under-saturated lakes was substantially decreased.
-However, there appeared to still be issues in replicating the tails as well as issues with central tendency.
+The PPCs for the saturation metrics below indicated that including the
+distributional models was perhaps an improvement on the initial model in
+some aspects; in particular, the bias in the predicted proportion of
+under-saturated lakes was substantially decreased. However, there
+appeared to still be issues in replicating the tails as well as issues
+with central tendency.
 
@@
   add_predicted_draws(n2o_mod2, 
-                      ndraws = 50) %>%
+                      ndraws = 200) %>%
   rename(n2o_obs = n2o,
@@
   add_predicted_draws(n2o_mod2, 
-                      ndraws = 500) %>%
+                      ndraws = 1000) %>%
   rename(n2o_obs = n2o,
@@
 
-#### R-square
-
-Relative to model 1, there was a substantial decrease in the $R^2$ estimate for the dissolved N2O component of this model. The estimate for the equilibrium N2O-eq component was similar to model 1.
-
-```{r r2_2, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
-round(bayes_R2(n2o_mod2, resp = "logn2o", cores = 1), 3) 
-round(bayes_R2(n2o_mod2, resp = "logn2oeq", cores = 1), 3)
-```
-
-## Model 3
-
-In the next model, the categorical version of the NO3 covariate and a surface temperature covariate were included to try to improve the fit.
-The ordinal NO3 variable was used as a monotonic, ordinal effect and only in the dissolved N2O component of the model.
-For the equlibrium N2O component, surface temperature and log-transformed elevation were used, along with their interaction.
-The model also retained the same distributional specifications included in model 2 above.
-
-```{r n2o_mod_3, eval=FALSE, include=TRUE}
-bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
-               surftemp +
-               (mo(no3_cat) | a | WSA9) + 
-               (mo(no3_cat) | b | WSA9:state) + 
-               (1 | c | WSA9:state:size_cat),
-             sigma ~ 1 +
-               (1 | WSA9) + 
-               (1 | WSA9:state) + 
-               (1 | WSA9:state:size_cat), 
-             family = gaussian())
-
-bf_n2oeq <- bf(log(n2o_eq) ~ surftemp +
-                 log_elev +
-                 surftemp:log_elev +
-                 (1 | a | WSA9) + 
-                 (1 | b | WSA9:state) +
-                 (1 | c | WSA9:state:size_cat),
-             sigma ~ 1 +
-               (1 | WSA9) + 
-               (1 | WSA9:state) + 
-               (1 | WSA9:state:size_cat),
-             family = gaussian())
-
-priors <- c(
-  prior(normal(2, 1), class = "Intercept", resp = "logn2o"),
-  prior(normal(0, 1), class = "b", resp = "logn2o"),
-  prior(exponential(2), class = "sd", resp = "logn2o"),
-  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2o"),
-  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2o"),
-  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), 
-  prior(normal(0, 1), class = "b", resp = "logn2oeq"), 
-  prior(exponential(2), class = "sd", resp = "logn2oeq"), 
-  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2oeq"),
-  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2oeq"),
-
-  prior(lkj(2), class = "rescor"),
-  prior(lkj(2), class = "cor")
-  )
-
-n2o_mod3 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
-                data = df_model, 
-                prior = priors,
-  control = list(adapt_delta = 0.975, max_treedepth = 12),
-  #sample_prior = "only",
-  save_pars = save_pars(all = TRUE),
-  seed = 98456,
-  chains=4, 
-  iter=5000, 
-  cores=4)
-
-save(n2o_mod3, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
-```
-
-### Summarize fit
-
-The fitted parameters and MCMC diagnostics are below.
-
-```{r print_mod3, echo=FALSE}
-print(n2o_mod3, prior = T)
-```
-
-### Model checks
-
-#### Dissolved N2O
-
-The PPCs below indicated a better fit compared to the previous models.
-The central tendency and tail behavior looked to be reasonably replicated by comparison.
-However, the observed _vs._ predicted plot suggested that larger N2O observations were likely being systematically underestimated.
-
-```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+The check for the revisit data is below.
+
+```{r ppc_sat2_revisit, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
-pp_check(n2o_mod3, 
-         resp = "logn2o",
-         type = "dens_overlay",
-         ndraws = 200,
-         cores = 1) + 
-  theme_tidybayes() +
-  xlim(0, 5) +
-  xlab(expression(paste("log(N"[2],"O)"))) + 
-  ylab("density")
-,
-pp_check(n2o_mod3, 
-         resp = "logn2o",
-         type = "ecdf_overlay",
-         ndraws = 200,
-         cores = 1) + 
-  theme_tidybayes() +
-  xlim(0, 5) +
-  xlab(expression(paste("log(N"[2],"O)"))) + 
-  ylab("cumulative density")
-,
-pp_check(n2o_mod3, 
-         resp = "logn2o",
-         type = "stat_2d", 
-         stat = c("mean", "sd"), 
-         ndraws = 1000,
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod3, 
-         resp = "logn2o",
-         type = "stat_2d", 
-         stat = c("kurtosis", "skewness"),  
-         ndraws = 1000,
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod3, 
-         resp = "logn2o",
-         type = "stat_2d", 
-         stat = c("min", "max"),  
-         ndraws = 1000,
-         cores = 1) + 
-  theme_tidybayes()
-,
-pp_check(n2o_mod3, 
-         resp = "logn2o",
-         type = "scatter_avg",  
-         ndraws = 1000,
-         cores = 1) + 
-  theme_tidybayes()
-, ncol = 2)
-```
-
-#### Equilibrium N2O
-
-The PPCs below indicated that this model appeared to be an improvement for equilibrium N2O as well. However, some checks (e.g., skewness) suggested some room for additional improvement.
-
-```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
-grid.arrange(
-pp_check(n2o_mod3, 
-         resp = "logn2oeq",
-         type = "dens_overlay", 
-         ndraws = 200,
-         cores = 1) + 
-  theme_tidybayes() +
-  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
-  ylab("density")
-,
-pp_check(n2o_mod3, 
-         resp = "logn2oeq",
-         type = "ecdf_overlay", 
-         ndraws = 200,
-         cores = 1) + 
-  theme_tidybayes() +
-  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
-  ylab("cumulative density")
-,
-pp_check(n2o_mod3, 
-         resp = "logn2oeq",
-         type = "stat_2d", 
-         stat = c("mean", "sd"),  
-         ndraws = 1000,
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod3, 
-         resp = "logn2oeq",
-         type = "stat_2d", 
-         stat = c("kurtosis", "skewness"), 
-         ndraws = 1000, 
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod3, 
-         resp = "logn2oeq",
-         type = "stat_2d", 
-         stat = c("min", "max"),  
-         ndraws = 1000,
-         cores = 1) + 
-  theme_tidybayes()
-,
-pp_check(n2o_mod3, 
-         resp = "logn2oeq",
-         type = "scatter_avg",  
-         ndraws = 1000,
-         cores = 1) + 
-  theme_tidybayes()
-, ncol = 2)
-```
-
-#### Bivariate
-
-The check for the joint distribution below also suggested an improvement upon the previous models.
-
-```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
-df_model %>%
-  add_predicted_draws(n2o_mod3, 
-                      ndraws = 20) %>%
-  rename(n2o_obs = n2o,
-         n2oeq_obs = n2o_eq) %>%
-  tidyr::pivot_wider(names_from = .category,
-              values_from = .prediction) %>%
-  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
-  geom_density_2d(aes(x = logn2o, 
-                      y = logn2oeq, 
-                      group = .draw),
-                  bins = 10,
-                  color = "lightblue", 
-                  alpha = 0.4) +
-  geom_density_2d(color = "black", bins = 10) +
-  xlim(1.25, 2.75) +
-  ylim(1.25, 2.75) +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
-  theme_tidybayes()
-```
-
-#### Saturation
-
-This model looked to be an improvement with regard to the PPCs for the saturation metrics.
-However, the proportion of under-saturated lakes remained biased low and other checks indicated that further improvements would be ideal.
-
-```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
-grid.arrange(
-df_model %>%
-  add_predicted_draws(n2o_mod3, 
-                      ndraws = 50) %>%
+df_test %>%
+  add_predicted_draws(n2o_mod2, 
+                      ndraws = 200) %>%
   rename(n2o_obs = n2o,
@@
 ,
-df_model %>%
-  add_predicted_draws(n2o_mod3, 
-                      ndraws = 500) %>%
+df_test %>%
+  add_predicted_draws(n2o_mod2, 
+                      ndraws = 1000) %>%
   rename(n2o_obs = n2o,
@@
               values_from = .prediction) %>%
-  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
+  mutate(usat_pred = ifelse(logn2o < logn2oeq, 1, 0)) %>%
   group_by(.draw) %>%
-  summarise(prop_pred = sum(usat_pred) / 984) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_test)) %>%
   ggplot(aes(x = prop_pred)) +
   geom_histogram(binwidth = 0.001, fill = "lightblue") +
-  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
+  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_test))) +
   xlab("Proportion undersaturated waterbodies") +
@@
 
-The $R^2$ estimates for this model are below and suggested substantial improvements over the previous models.
-
-```{r r2_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
-round(bayes_R2(n2o_mod3, resp = "logn2o", cores = 1), 3) 
-round(bayes_R2(n2o_mod3, resp = "logn2oeq", cores = 1), 3)
-```
-
-### Covariate effects
-
-Below are plots illustrating the modeled effects of covariates on both N2O and equilibrium N2O.
-
-#### Dissolved N2O
-
-The conditional effects plots below for dissolved N2O illustrated a positive, monotonic, and non-linear relationship with NO3; and a negative, linear relationship with surface temperature.
-
-```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
-p1 <- conditional_effects(n2o_mod3, 
-                          resp = "logn2o", 
-                          effects = c("no3_cat"), 
-                          plot = F)
-p2 <- conditional_effects(n2o_mod3, 
-                          resp = "logn2o", 
-                          effects = c("surftemp"), 
-                          plot = F)
-
-plt <- ggarrange(plot(p1, plot = F)[[1]], 
-          plot(p2, plot = F)[[1]],
-          ncol = 2)
-
-rm(p1, p2)
-
-annotate_figure(plt, top = text_grob("Conditional effects: N2O", 
-               color = "black", face = "bold", size = 14))
-```
-
-#### Equilibrium N2O
-
-The modeled effects for the equilibrium N2O component illustrated a negative relationship with both predictors and an interaction such that the surface temperature effect became slightly steeper at lower elevations.
-
-```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
-p1 <- conditional_effects(n2o_mod3, 
-                          resp = "logn2oeq", 
-                          effects = c("log_elev:surftemp"), 
-                          plot = F)
-p2 <- conditional_effects(n2o_mod3, 
-                          resp = "logn2oeq", 
-                          effects = c("surftemp:log_elev"), 
-                          plot = F)
-
-plt <- ggarrange(plot(p1, plot = F)[[1]],
-                 plot(p2, plot = F)[[1]])
-
-rm(p1, p2)
-
-annotate_figure(plt, top = text_grob("Conditional effects: Equilibrium N2O", 
-               color = "black", face = "bold", size = 14))
-```
-
-## Model 4
-
-In the next model, covariate terms were also included in the $\sigma$ components of both models in order to try to better capture remaining heterogeneity in the variances of both N2O and N2O-eq.
-
-```{r n2o_mod_mv_4, eval=FALSE, include=TRUE}
+Relative to model 1, there was a substantial decrease in the $R^2$
+estimate for the dissolved N2O component of this model. The estimate for
+the equilibrium N2O-eq component was similar to model 1.
+
+```{r r2_2, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+round(bayes_R2(n2o_mod2, resp = "logn2o", cores = 1), 3) 
+round(bayes_R2(n2o_mod2, resp = "logn2oeq", cores = 1), 3)
+```
+
+And for the re-visit data:
+
+```{r r2_2_test, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+round(bayes_R2(n2o_mod2, resp = "logn2o", newdata = df_test, cores = 1), 3) 
+round(bayes_R2(n2o_mod2, resp = "logn2oeq", newdata = df_test, cores = 1), 3)
+```
+
+## Model 3
+
+In the next model, the categorical version of the NO3 covariate and a
+surface temperature covariate were included to try to improve the fit.
+The ordinal NO3 variable was used as a monotonic, ordinal effect and
+only in the dissolved N2O component of the model. For the equlibrium N2O
+component, surface temperature and log-transformed elevation were used,
+along with their interaction. The model also retained the same
+distributional specifications included in model 2 above.
+
+```{r n2o_mod_3, eval=FALSE, include=TRUE}
 bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
@@
                (1 | c | WSA9:state:size_cat),
-             sigma ~ mo(no3_cat) +
-               surftemp +
+             sigma ~ 1 +
                (1 | WSA9) + 
@@
                  (1 | c | WSA9:state:size_cat),
-             sigma ~ surftemp +
-               log_elev +
+             sigma ~ 1 +
                (1 | WSA9) + 
@@
   prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2o"),
-  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2o"),
   prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2o"),
@@
   prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2oeq"),
-  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2oeq"),
   prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2oeq"),
@@
 
-n2o_mod4 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
+n2o_mod3 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
                 data = df_model, 
@@
   save_pars = save_pars(all = TRUE),
-  seed = 15851,
+  seed = 98456,
   chains=4, 
@@
 
-save(n2o_mod4, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
+save(n2o_mod3, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod3.rda")
 ```
@@
 
-Below is a summary of the fitted parameters along with some convergence diagnostics.
-
-```{r print_mod4, echo=FALSE}
-print(n2o_mod4, prior = T, digits = 3)
+The fitted parameters and MCMC diagnostics are below.
+
+```{r print_mod3, echo=FALSE}
+print(n2o_mod3, prior = T)
 ```
@@
 
-The same PPCs were employed for this model as above.
-
 #### Dissolved N2O
 
-This model appeared to be a moderate improvement on the previous models.
-
-```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+The PPCs below indicated a better fit compared to the previous models.
+The central tendency and tail behavior looked to be reasonably
+replicated by comparison. However, the observed *vs.* predicted plot
+suggested that larger N2O observations were likely being systematically
+underestimated.
+
+```{r ppc_full_checks_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
-pp_check(n2o_mod4, 
-         ndraws = 200,
+pp_check(n2o_mod3, 
          resp = "logn2o",
          type = "dens_overlay",
-         nsamples = 40,
-         cores = 1) + 
+         ndraws = 200) + 
   theme_tidybayes() +
+  xlim(0, 5) +
   xlab(expression(paste("log(N"[2],"O)"))) + 
@@
 ,
-pp_check(n2o_mod4, 
-         ndraws = 200,
+pp_check(n2o_mod3, 
          resp = "logn2o",
          type = "ecdf_overlay",
-         nsamples = 40, 
-         cores = 1) + 
+         ndraws = 200) + 
   theme_tidybayes() +
+  xlim(0, 5) +
   xlab(expression(paste("log(N"[2],"O)"))) + 
@@
 ,
-pp_check(n2o_mod4, 
-         ndraws = 1000,
+pp_check(n2o_mod3, 
          resp = "logn2o",
@@
          stat = c("mean", "sd"), 
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod4, 
-         ndraws = 1000,
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
          resp = "logn2o",
@@
          stat = c("kurtosis", "skewness"),  
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod4, 
-         ndraws = 1000,
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
          resp = "logn2o",
@@
          stat = c("min", "max"),  
-         cores = 1) + 
-  theme_tidybayes()
-,
-pp_check(n2o_mod4, 
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
          resp = "logn2o",
          type = "scatter_avg",  
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
 
+Below are the same PPCs for the second-visit data.
+
+```{r ppc_full_checks_mod_n2o3_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "dens_overlay",
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlim(0, 5) +
+  xlab(expression(paste("log(N"[2],"O)"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "ecdf_overlay",
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlim(0, 5) +
+  xlab(expression(paste("log(N"[2],"O)"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("mean", "sd"), 
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness"),  
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("min", "max"),  
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "scatter_avg",  
+         ndraws = 1000) + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### Equilibrium N2O
 
-This component of the model also seemed to be an improvement over model 3, with better representation in the tails as indicated in the skewness _vs._ kurtosis PPC.
-
-```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+The PPCs below indicated that this model appeared to be an improvement
+for equilibrium N2O as well. However, some checks (e.g., skewness)
+suggested some room for additional improvement.
+
+```{r ppc_full_checks_mod_n2oeq3, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
-pp_check(n2o_mod4, 
-         ndraws = 200,
+pp_check(n2o_mod3, 
          resp = "logn2oeq",
          type = "dens_overlay", 
-         cores = 1) + 
+         ndraws = 200) + 
   theme_tidybayes() +
@@
 ,
-pp_check(n2o_mod4, 
-         ndraws = 200,
+pp_check(n2o_mod3, 
          resp = "logn2oeq",
          type = "ecdf_overlay", 
-         cores = 1) + 
+         ndraws = 200) + 
   theme_tidybayes() +
@@
 ,
-pp_check(n2o_mod4, 
-         ndraws = 1000,
+pp_check(n2o_mod3, 
          resp = "logn2oeq",
@@
          stat = c("mean", "sd"),  
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod4, 
-         ndraws = 1000,
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
          resp = "logn2oeq",
@@
          stat = c("kurtosis", "skewness"), 
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
-  theme_tidybayes()
-,
-pp_check(n2o_mod4, 
-         ndraws = 1000,
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
          resp = "logn2oeq",
@@
          stat = c("min", "max"),  
-         cores = 1) + 
-  theme_tidybayes()
-,
-pp_check(n2o_mod4, 
-         ndraws = 1000,
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
          resp = "logn2oeq",
          type = "scatter_avg",  
-         cores = 1) + 
+         ndraws = 1000) + 
   theme_tidybayes()
@@
 
+Checks for the 2nd visit data are below.
+
+```{r ppc_full_checks_mod_n2oeq3_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "dens_overlay", 
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "ecdf_overlay", 
+         ndraws = 200) + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("mean", "sd"),  
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness"), 
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("min", "max"),  
+         ndraws = 1000) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod3, 
+         newdata = df_test,
+         resp = "logn2oeq",
+         type = "scatter_avg",  
+         ndraws = 1000) + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### Bivariate
 
-Again, an improvement over the previous model with a tighter fit of the PPC to the observed bivariate density.
-
-```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+The check for the joint distribution below also suggested an improvement
+upon the previous models.
+
+```{r ppc_bv_check_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
-  add_predicted_draws(n2o_mod4, 
+  add_predicted_draws(n2o_mod3, 
                       ndraws = 20) %>%
@@
 
+The same check below for the second-visit sites.
+
+```{r ppc_bv_check_mod_n2o3_test, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+df_test %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 20) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
+  geom_density_2d(aes(x = logn2o, 
+                      y = logn2oeq, 
+                      group = .draw),
+                  bins = 10,
+                  color = "lightblue", 
+                  alpha = 0.4) +
+  geom_density_2d(color = "black", bins = 10) +
+  xlim(1.25, 2.75) +
+  ylim(1.25, 2.75) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+```
+
 #### Saturation
 
-This check also suggested an improvement over the previous models, with better tail behavior and less bias in the proportion under-saturated measure.
-
-```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+This model looked to be an improvement with regard to the PPCs for the
+saturation metrics. However, the proportion of under-saturated lakes
+remained biased low and other checks indicated that further improvements
+would be ideal.
+
+```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
 grid.arrange(
 df_model %>%
-  add_predicted_draws(n2o_mod4, 
-                      ndraws = 50) %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 200) %>%
   rename(n2o_obs = n2o,
@@
 df_model %>%
-  add_predicted_draws(n2o_mod4, 
-                      ndraws = 500) %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 1000) %>%
   rename(n2o_obs = n2o,
@@
 
+The same check for the second-visit sites are next.
+
+```{r ppc_sat_check_mod_n2o3_test, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10,  message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+df_test %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 200) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  ggplot(aes(x = sat_ratio)) +
+  geom_density(aes(x = sat_pred, group = .draw), 
+               n = 1024, 
+               adjust = 1,
+               color = "lightblue") +
+  geom_density(n = 1024, adjust = 2) +
+  xlim(0, 5) +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
+  group_by(.draw) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_test)) %>%
+  ggplot(aes(x = prop_pred)) +
+  geom_histogram(binwidth = 0.001, fill = "lightblue") +
+  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_test))) +
+  xlab("Proportion undersaturated waterbodies") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(mean_yrep = mean(sat_pred),
+         sd_yrep = sd(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(mean_y = mean(sat_ratio),
+         sd_y = sd(sat_ratio)) %>%
+  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(min_yrep = min(sat_pred),
+         max_yrep = max(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(min_y = min(sat_ratio),
+         max_y = max(sat_ratio)) %>%
+  ggplot(aes(x = min_yrep, y = max_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod3, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.row) %>%
+  mutate(mean_yrep = mean(sat_pred)) %>% 
+  filter(.draw == 1) %>% 
+  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
+  geom_point(color = "lightblue") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### R-square
 
-The Bayesian $R^2$ estimates below indicated an improvement from the previous models.
+The $R^2$ estimates for this model are below and suggested substantial
+improvements over the previous models.
+
+```{r r2_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+round(bayes_R2(n2o_mod3, resp = "logn2o", cores = 1), 3) 
+round(bayes_R2(n2o_mod3, resp = "logn2oeq", cores = 1), 3)
+```
+
+Next, $R^2$ for the 2nd visit sites.
+
+```{r r2_mod_n2o3_test, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+round(bayes_R2(n2o_mod3, newdata = df_test, resp = "logn2o", cores = 1), 3) 
+round(bayes_R2(n2o_mod3, newdata = df_test, resp = "logn2oeq", cores = 1), 3)
+```
+
+### Covariate effects
+
+Below are plots illustrating the modeled effects of covariates on both
+N2O and equilibrium N2O.
+
+#### Dissolved N2O
+
+The conditional effects plots below for dissolved N2O illustrated a
+positive, monotonic, and non-linear relationship with NO3; and a
+negative, linear relationship with surface temperature.
+
+```{r conditional_effects_mod_n2o3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
+p1 <- conditional_effects(n2o_mod3, 
+                          resp = "logn2o", 
+                          effects = c("no3_cat"), 
+                          plot = F)
+p2 <- conditional_effects(n2o_mod3, 
+                          resp = "logn2o", 
+                          effects = c("surftemp"), 
+                          plot = F)
+
+plt <- ggarrange(plot(p1, plot = F)[[1]], 
+          plot(p2, plot = F)[[1]],
+          ncol = 2)
+
+rm(p1, p2)
+
+annotate_figure(plt, top = text_grob("Conditional effects: N2O", 
+               color = "black", face = "bold", size = 14))
+```
+
+#### Equilibrium N2O
+
+The modeled effects for the equilibrium N2O component illustrated a
+negative relationship with both predictors and an interaction such that
+the surface temperature effect became slightly steeper at lower
+elevations.
+
+```{r conditional_effects_mod_n2oeq3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, cache=TRUE}
+p1 <- conditional_effects(n2o_mod3, 
+                          resp = "logn2oeq", 
+                          effects = c("log_elev:surftemp"), 
+                          plot = F)
+p2 <- conditional_effects(n2o_mod3, 
+                          resp = "logn2oeq", 
+                          effects = c("surftemp:log_elev"), 
+                          plot = F)
+
+plt <- ggarrange(plot(p1, plot = F)[[1]],
+                 plot(p2, plot = F)[[1]])
+
+rm(p1, p2)
+
+annotate_figure(plt, top = text_grob("Conditional effects: Equilibrium N2O", 
+               color = "black", face = "bold", size = 14))
+```
+
+## Model 4
+
+In the next model, covariate terms were also included in the $\sigma$
+components of both models in order to try to better capture remaining
+heterogeneity in the variances of both N2O and N2O-eq.
+
+```{r n2o_mod_mv_4, eval=FALSE, include=TRUE}
+bf_n2o <- bf(log(n2o) ~ mo(no3_cat) +
+               surftemp +
+               (mo(no3_cat) | a | WSA9) + 
+               (mo(no3_cat) | b | WSA9:state) + 
+               (1 | c | WSA9:state:size_cat),
+             sigma ~ mo(no3_cat) +
+               surftemp +
+               (1 | WSA9) + 
+               (1 | WSA9:state) + 
+               (1 | WSA9:state:size_cat), 
+             family = gaussian())
+
+bf_n2oeq <- bf(log(n2o_eq) ~ surftemp +
+                 log_elev +
+                 surftemp:log_elev +
+                 (1 | a | WSA9) + 
+                 (1 | b | WSA9:state) +
+                 (1 | c | WSA9:state:size_cat),
+             sigma ~ surftemp +
+               log_elev +
+               (1 | WSA9) + 
+               (1 | WSA9:state) + 
+               (1 | WSA9:state:size_cat),
+             family = gaussian())
+
+priors <- c(
+  prior(normal(2, 1), class = "Intercept", resp = "logn2o"),
+  prior(normal(0, 1), class = "b", resp = "logn2o"),
+  prior(exponential(2), class = "sd", resp = "logn2o"),
+  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2o"),
+  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2o"),
+  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2o"),
+  prior(normal(2, 1), class = "Intercept", resp = "logn2oeq"), 
+  prior(normal(0, 1), class = "b", resp = "logn2oeq"), 
+  prior(exponential(2), class = "sd", resp = "logn2oeq"), 
+  prior(normal(-1, 2), class = "Intercept", dpar = "sigma", resp = "logn2oeq"),
+  prior(normal(0, 1), class = "b", dpar = "sigma", resp = "logn2oeq"),
+  prior(exponential(2), class = "sd", dpar = "sigma", resp = "logn2oeq"),
+
+  prior(lkj(2), class = "rescor"),
+  prior(lkj(2), class = "cor")
+  )
+
+n2o_mod4 <- brm(bf_n2o + bf_n2oeq + set_rescor(rescor = TRUE),
+                data = df_model, 
+                prior = priors,
+  control = list(adapt_delta = 0.975, max_treedepth = 12),
+  #sample_prior = "only",
+  save_pars = save_pars(all = TRUE),
+  seed = 15851,
+  chains=4, 
+  iter=5000, 
+  cores=4)
+
+save(n2o_mod4, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod4.rda")
+```
+
+### Summarize fit
+
+Below is a summary of the fitted parameters along with some convergence
+diagnostics.
+
+```{r print_mod4, echo=FALSE}
+print(n2o_mod4, prior = T, digits = 3)
+```
+
+### Model checks
+
+The same PPCs were employed for this model as above.
+
+#### Dissolved N2O
+
+This model appeared to be a moderate improvement on the previous models.
+
+```{r ppc_full_checks_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod4, 
+         ndraws = 200,
+         resp = "logn2o",
+         type = "dens_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(N"[2],"O)"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod4, 
+         ndraws = 200,
+         resp = "logn2o",
+         type = "ecdf_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(N"[2],"O)"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod4, 
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("mean", "sd")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("min", "max")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         resp = "logn2o",
+         type = "scatter_avg") + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
+This same checks for the 2nd visit sites is below.
+
+```{r ppc_full_checks_mod_n2o4_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2o",
+         type = "dens_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(N"[2],"O)"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2o",
+         type = "ecdf_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(N"[2],"O)"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("mean", "sd")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("min", "max")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         resp = "logn2o",
+         type = "scatter_avg") + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
+#### Equilibrium N2O
+
+This component of the model also seemed to be an improvement over model
+3, with better representation in the tails as indicated in the skewness
+*vs.* kurtosis PPC.
+
+```{r ppc_full_checks_mod_n2oeq4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod4, 
+         ndraws = 200,
+         resp = "logn2oeq",
+         type = "dens_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod4, 
+         ndraws = 200,
+         resp = "logn2oeq",
+         type = "ecdf_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod4, 
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("mean", "sd")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("min", "max")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "scatter_avg") + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
+For the 2nd visit sites.
+
+```{r ppc_full_checks_mod_n2oeq4_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod4,
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2oeq",
+         type = "dens_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2oeq",
+         type = "ecdf_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O)"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("mean", "sd")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("min", "max")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod4, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "scatter_avg") + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
+#### Bivariate
+
+Again, an improvement over the previous model with a tighter fit of the
+PPC to the observed bivariate density.
+
+```{r ppc_bv_check_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+df_model %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 20) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
+  geom_density_2d(aes(x = logn2o, 
+                      y = logn2oeq, 
+                      group = .draw),
+                  bins = 10,
+                  color = "lightblue", 
+                  alpha = 0.4) +
+  geom_density_2d(color = "black", bins = 10) +
+  xlim(1.25, 2.75) +
+  ylim(1.25, 2.75) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+```
+
+For the 2nd visit sites.
+
+```{r ppc_bv_check_mod_n2o4_test, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+df_test %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 20) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
+  geom_density_2d(aes(x = logn2o, 
+                      y = logn2oeq, 
+                      group = .draw),
+                  bins = 10,
+                  color = "lightblue", 
+                  alpha = 0.4) +
+  geom_density_2d(color = "black", bins = 10) +
+  xlim(1.25, 2.75) +
+  ylim(1.25, 2.75) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+```
+
+#### Saturation
+
+This check also suggested an improvement over the previous models, with
+better tail behavior and less bias in the proportion under-saturated
+measure.
+
+```{r ppc_sat_check_mod_n2o4, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+df_model %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 200) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  ggplot(aes(x = sat_ratio)) +
+  geom_density(aes(x = sat_pred, group = .draw), 
+               n = 1024, 
+               adjust = 1,
+               color = "lightblue") +
+  geom_density(n = 1024, adjust = 2) +
+  xlim(0, 5) +
+  theme_tidybayes()
+,
+df_model %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
+  group_by(.draw) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_model)) %>%
+  ggplot(aes(x = prop_pred)) +
+  geom_histogram(binwidth = 0.001, fill = "lightblue") +
+  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_model))) +
+  xlab("Proportion undersaturated waterbodies") +
+  theme_tidybayes()
+,
+df_model %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(mean_yrep = mean(sat_pred),
+         sd_yrep = sd(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(mean_y = mean(sat_ratio),
+         sd_y = sd(sat_ratio)) %>%
+  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_model %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(min_yrep = min(sat_pred),
+         max_yrep = max(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(min_y = min(sat_ratio),
+         max_y = max(sat_ratio)) %>%
+  ggplot(aes(x = min_yrep, y = max_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_model %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.row) %>%
+  mutate(mean_yrep = mean(sat_pred)) %>% 
+  filter(.draw == 1) %>% 
+  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
+  geom_point(color = "lightblue") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+, ncol = 2)
+```
+
+For the second visit sites:
+
+```{r ppc_sat_check_mod_n2o4_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+df_test %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 200) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  ggplot(aes(x = sat_ratio)) +
+  geom_density(aes(x = sat_pred, group = .draw), 
+               n = 1024, 
+               adjust = 1,
+               color = "lightblue") +
+  geom_density(n = 1024, adjust = 2) +
+  xlim(0, 5) +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
+  group_by(.draw) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_test)) %>%
+  ggplot(aes(x = prop_pred)) +
+  geom_histogram(binwidth = 0.001, fill = "lightblue") +
+  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_test))) +
+  xlab("Proportion undersaturated waterbodies") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(mean_yrep = mean(sat_pred),
+         sd_yrep = sd(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(mean_y = mean(sat_ratio),
+         sd_y = sd(sat_ratio)) %>%
+  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(min_yrep = min(sat_pred),
+         max_yrep = max(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(min_y = min(sat_ratio),
+         max_y = max(sat_ratio)) %>%
+  ggplot(aes(x = min_yrep, y = max_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod4, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.row) %>%
+  mutate(mean_yrep = mean(sat_pred)) %>% 
+  filter(.draw == 1) %>% 
+  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
+  geom_point(color = "lightblue") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+, ncol = 2)
+```
+
+#### R-square
+
+The Bayesian $R^2$ estimates below indicated an improvement from the
+previous models.
 
@@
 
+For the 2nd visit data.
+
+```{r r2_mod_n2o4_test, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+round(bayes_R2(n2o_mod4, newdata = df_test, resp = "logn2o", cores = 1), 3) 
+round(bayes_R2(n2o_mod4, newdata = df_test, resp = "logn2oeq", cores = 1), 3)
+```
+
 ### Covariate effects
@@
 
-The conditional effects plots for the covariate effects on dissolved N2O remained similar to the previous model.
+The conditional effects plots for the covariate effects on dissolved N2O
+remained similar to the previous model.
 
@@
 
-Below are estimates of the conditional effects of the covariates on $\sigma$ for N2O.
-These plots suggested a large effect of NO3 on the variance of N2O, but little to no effect of surface temperature.
+Below are estimates of the conditional effects of the covariates on
+$\sigma$ for N2O. These plots suggested a large effect of NO3 on the
+variance of N2O, but little to no effect of surface temperature.
 
@@
 
-The covariate effects on equilibrium N2O remained largely the same as for the previous model.
+The covariate effects on equilibrium N2O remained largely the same as
+for the previous model.
 
@@
 
-The covariate effects on $\sigma$ for N2O-eq suggested a negative effect of surface temperature and litte to no effect of elevation.
+The covariate effects on $\sigma$ for N2O-eq suggested a negative effect
+of surface temperature and litte to no effect of elevation.
 
@@
 
-In the next model, additional complexity is added to the dissolved N2O component by including a covariate for continuous lake surface area (log scale) as well as interactions between NO3 and log(surface area) and surface temperature.
-Although, some aspect of lake size was likely captured by the categorical size factor from the survey in the models above, the largest category is very coarse and includes all lakes >50 hecatares, which is an extremely wide distribution of sizes from 50 to over 300,000 hectares at max.
-Including the continuous size covariate at the observation level in this manner was expected to do a better job of resolving any trends with size.
-Colinearity between the grouped and observation-level measures were not a major, concern, since the observation-level measure would account for any variation first.
-However, we did later check for any common violations of the Gauss-Markov assumptions, by looking at potential correlations between the model errors and covariates [not shown; @Bafumi_Gelman_2007].
+In the next model, additional complexity is added to the dissolved N2O
+component by including a covariate for continuous lake surface area (log
+scale) as well as interactions between NO3 and log(surface area) and
+surface temperature. Although some variation due to lake size was likely
+captured by the categorical survey variable for size in the models
+above, the categorical distinction may not well represent the true
+continuous nature of waterbody size and its effects on the response. For
+example, the largest size category represented all lakes \>50 hecatares,
+which included an extremely wide range of waterbody sizes ranging from
+50 to over 300,000 hectares at max. Including the continuous size
+covariate at the observation level in this manner was expected to do a
+better job of resolving trends with size. Colinearity between the
+grouped and observation-level measures were not a major, concern, since
+the observation-level measure would account for any variation first.
+Though not presented in this document, violations of the Gauss-Markov
+assumptions were checked by looking for potential correlations between
+model errors and the covariates [@Bafumi_Gelman_2007].
 
@@
 
-Below is a summary of the fitted parameters along with MCMC convergence diagnostics.
+Below is a summary of the fitted parameters along with MCMC convergence
+diagnostics.
 
@@
 
-This PPCs for dissolved N2O looked similar to the previous model.
+The PPC for dissolved N2O below suggested minimal improvements upon the
+the previous model.
 
@@
          resp = "logn2o",
-         type = "dens_overlay",
-         cores = 1) + 
+         type = "dens_overlay") + 
   theme_tidybayes() +
@@
          resp = "logn2o",
-         type = "ecdf_overlay",
-         cores = 1) + 
+         type = "ecdf_overlay") + 
   theme_tidybayes() +
@@
          type = "stat_2d", 
-         stat = c("mean", "sd"), 
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
+         stat = c("mean", "sd")) + 
   theme_tidybayes()
@@
          type = "stat_2d", 
-         stat = c("kurtosis", "skewness"), 
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
+         stat = c("kurtosis", "skewness")) + 
   theme_tidybayes()
@@
          type = "stat_2d", 
-         stat = c("min", "max"), 
-         cores = 1) + 
-  scale_x_continuous(trans = "log") +
-  scale_y_continuous(trans = "log") +
+         stat = c("min", "max")) + 
   theme_tidybayes()
@@
 
+The same check for the 2nd visit data is below.
+
+```{r ppc_full_checks_mod_n2o5_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2o",
+         type = "dens_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2o",
+         type = "ecdf_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("mean", "sd")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "stat_2d", 
+         stat = c("min", "max")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2o",
+         type = "scatter_avg") + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### Equilibrium N2O
 
-Again, the PPCs for this model were similar to the previous model, which was unsurprising given that it was the same model for N2O-eq.
+The PPCs for this model were similar to the previous model, which was
+unsurprising given that it was the same model for N2O-eq.
 
@@
          resp = "logn2oeq",
-         type = "dens_overlay",
-         cores = 1) + 
+         type = "dens_overlay") + 
   theme_tidybayes() +
@@
          resp = "logn2oeq",
-         type = "ecdf_overlay",
-         cores = 1) + 
+         type = "ecdf_overlay") + 
   theme_tidybayes() +
@@
 pp_check(n2o_mod5, 
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "scatter_avg") + 
+  theme_tidybayes()
+, ncol = 2)
+```
+
+Below, is the same check using the 2nd visit data.
+
+```{r ppc_full_checks_mod_n2oeq5_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2oeq",
+         type = "dens_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 200,
+         resp = "logn2oeq",
+         type = "ecdf_overlay") + 
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("mean", "sd")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness")) +
+  theme_tidybayes()
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "logn2oeq",
+         type = "stat_2d", 
+         stat = c("min", "max")) + 
+  theme_tidybayes()
+,
+pp_check(n2o_mod5, 
+         newdata = df_test,
          ndraws = 1000,
@@
 
-```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.asp=1, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+```{r ppc_bv_check_mod_n2o5, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
 
+And the check with the 2nd visit data.
+
+```{r ppc_bv_check_mod_n2o5_test, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+df_test %>%
+  add_predicted_draws(n2o_mod5, 
+                      ndraws = 20) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
+  geom_density_2d(aes(x = logn2o, 
+                      y = logn2oeq, 
+                      group = .draw),
+                  bins = 10,
+                  color = "lightblue", 
+                  alpha = 0.4) +
+  geom_density_2d(color = "black", bins = 10) +
+  xlim(1.25, 2.75) +
+  ylim(1.25, 2.75) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+```
+
 #### Saturation
 
-This check was also similar to the prevoius model, with perhaps slightly less bias in the proportion unsaturated estimates.
-There was also a potentially concerning extreme prediction in the observed _vs_ predicted PPC.
+This check was also similar to the prevoius model, with perhaps slightly
+less bias in the proportion unsaturated estimates. There was also a
+potentially concerning extreme prediction in the observed *vs* predicted
+PPC.
 
@@
   add_predicted_draws(n2o_mod5, 
-                      ndraws = 50) %>%
+                      ndraws = 200) %>%
   rename(n2o_obs = n2o,
@@
   add_predicted_draws(n2o_mod5, 
-                      ndraws = 500) %>%
+                      ndraws = 1000) %>%
   rename(n2o_obs = n2o,
@@
 
+The same check for the revisit sites below:
+
+```{r ppc_sat_check_mod_n2o5_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+df_test %>%
+  add_predicted_draws(n2o_mod5, 
+                      ndraws = 200) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  ggplot(aes(x = sat_ratio)) +
+  geom_density(aes(x = sat_pred, group = .draw), 
+               n = 1024, 
+               adjust = 1,
+               color = "lightblue") +
+  geom_density(n = 1024, adjust = 2) +
+  xlim(0, 5) +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod5, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(usat_pred = ifelse(exp(logn2o) < exp(logn2oeq), 1, 0)) %>%
+  group_by(.draw) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_test)) %>%
+  ggplot(aes(x = prop_pred)) +
+  geom_histogram(binwidth = 0.001, fill = "lightblue") +
+  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_test))) +
+  xlab("Proportion undersaturated waterbodies") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod5, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(mean_yrep = mean(sat_pred),
+         sd_yrep = sd(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(mean_y = mean(sat_ratio),
+         sd_y = sd(sat_ratio)) %>%
+  ggplot(aes(x = mean_yrep, y = sd_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = mean_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = sd_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod5, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.draw) %>%
+  mutate(min_yrep = min(sat_pred),
+         max_yrep = max(sat_pred)) %>% 
+  ungroup() %>%
+  mutate(min_y = min(sat_ratio),
+         max_y = max(sat_ratio)) %>%
+  ggplot(aes(x = min_yrep, y = max_yrep)) +
+  geom_point(color = "lightblue") +
+  geom_vline(aes(xintercept = min_y), linetype = "dashed") +
+  geom_hline(aes(yintercept = max_y), linetype = "dashed") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+df_test %>%
+  add_predicted_draws(n2o_mod5, 
+                      ndraws = 1000) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  mutate(sat_ratio = n2o_obs / n2oeq_obs,
+         sat_pred = exp(logn2o) / exp(logn2oeq)) %>%
+  group_by(.row) %>%
+  mutate(mean_yrep = mean(sat_pred)) %>% 
+  filter(.draw == 1) %>% 
+  ggplot(aes(x = mean_yrep, y = sat_ratio)) +
+  geom_point(color = "lightblue") +
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### R-square
@@
 
+For the revisit sites:
+
+```{r r2_mod_n2o5_test, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+round(bayes_R2(n2o_mod5, resp = "logn2o", newdata = df_test, cores = 1), 3) 
+round(bayes_R2(n2o_mod5, resp = "logn2oeq", newdata = df_test, cores = 1), 3)
+```
+
 ### Covariate effects
@@
 
-The conditional effects plots suggested a similar effect of NO3, but interesting interactions between NO3 and lake area and NO3 and surface temperature.
-For lake area, the effect was estimated to be larger and more negative at the highest levels of NO3; and slightly negative at the lowest level of NO3.
-For surface temperature, the effect was estimated to be largest and positive at the highest level of NO3; and negative at the lowest level of NO3.
+The conditional effects plots suggested a similar effect of NO3, but
+interesting interactions between NO3 and lake area and NO3 and surface
+temperature. For lake area, the effect was estimated to be larger and
+more negative at the highest levels of NO3; and slightly negative at the
+lowest level of NO3. For surface temperature, the effect was estimated
+to be largest and positive at the highest level of NO3; and negative at
+the lowest level of NO3.
 
@@
 
-The estimated covariate effects on $\sigma$ suggested a negative relationship with log(area) and a positive relationship, again, with NO3.
+The estimated covariate effects on $\sigma$ suggested a negative
+relationship with log(area) and a positive relationship, again, with
+NO3.
 
@@
 
-The estimated covariate effects on equilibrium N2O remained largely the same as estimated in the previous model.
+The estimated covariate effects on equilibrium N2O remained largely the
+same as estimated in the previous model.
 
@@
 
-As demonstrated above, models excluding the NO3 covariate consistently resulted in poorer fits to to the observed dissolved N2O data and potentially biased estimates of the saturation ratio.
-Including surface temperature and elevation in the equilibrium N2O part of the model also resulted in substantially improved replication of key aspects of the observed data.
-Added flexibility in the distributional terms for both dissolved and equilibrium N2O also led to improvements.
-
-To make inferences to the population of interest from a model including these covariates, however, the covariates needed to be (1) fully observed across that population or (2) their missingness needed to be modeled.
-For the lake area and elevation covariates, data _was_ available for all lakes from previously compiled geospatial databases.
-However, neither surface temperature or NO3 were observed for lakes outside of the sample.
-That is, they were partially observed with respect to the target population.
-Therefore, a more complex model was constructed below that included surface temperature and NO3 as additional responses conditioned on the survey design variables and fully observed covariates.
-This approach to inference for N2O was similar to a Bayesian structural equation modeling approach [@Merkle_etal_2021; @Merkle_Rosseel_2018].
-The logical dependence structure could be characterized as:
-
-$$\begin{aligned} 
-{p( \boldsymbol{N_2O_{diss}}} &| Survey, Area, {\boldsymbol{NO_3}}, {\boldsymbol{Temp}}) \\ 
-{p( \boldsymbol{N_2O_{equil}}} &| Survey, Elev, {\boldsymbol{Temp}})\\
-{p( \boldsymbol{NO_3}} &| Survey, Area, {\boldsymbol{Temp}}) \\ 
-{p( \boldsymbol{Temp}} &| Survey, Lat, Elev, Day)
-\end{aligned}$$
-
-Variables in bold text above were treated as partially observed with respect to the population of interest (i.e., observed only in the sample), whereas variables not in bold were considered fully observed.
-The partially observed variables, being dissolved and equilibrium N2O, NO3, and surface temperature, were each modeled conditional on the survey design variables and other partially and/or fully observed covariates.
-This piece-wise approach to predicting N2O required a more complex set of post-processing steps compared to a typical MRP analysis.
-In order to propagate uncertainty through the dependency structure and make inferences to the POI below, the fitted final model was used to first predict surface temperature in the target population, since it depended only on the fully observed covariates.
-That predictive distribution was then used alongside the relevant fully observed covariates to predict NO3 in the target population.
-Finally, the predictive distributions for temperature and NO3 were used to predict the N2O responses.
-These steps were carried out in the "Predict to population" section to follow.
-
-In the final model below, the sub-model for surface temperature assumed a Gamma distributed error distribution and the linear predictor included the survey design variables, latitude, elevation, and Julian date.
-The shape parameter was also modeled as a function of latitude to address increasing response variance along the latitudinal gradient.
-The NO3 sub-model was a cumulative logit model and the linear predictor included all of the survey factors as well as surface temperature and lake area.
-
-The dissolved and equilibrium N2O responses were each modeled with Gamma distributed errors, but with the same covariate structure as in model 5.
-The same structure was also employed for the shape terms in these responses, corresponding to the $\sigma$ terms in the previous models.
-Though not shown in this document, the Gamma error structure appeared to result in slightly better performance in the predictive checks compared to the Gaussian errors in previous models.
-This was primarily apparent in the saturation ratio checks, which may have been more sensitive to model performance in the tails of the N2O responses.
-Others have also indicated that the Gamma error distribution can work well for dissolved N2O data [@Webb_etal_2019].
-
-Note that there was no observation-level residual correlation term for this model, since the residuals are undefined for the Gamma and cumulative logit models.
-Dropping the observation-level residual correlation term was deemed a reasonable compromise that enabled the inclusion of NO3 as a covariate on dissolved N2O.
-The random intercepts, however, still allowed for potential correlations between the four responses at the group levels.
+As demonstrated above, models excluding the NO3 covariate consistently
+resulted in poorer fits to to the observed dissolved N2O data and
+potentially biased estimates of the saturation ratio. Including surface
+temperature and elevation in the equilibrium N2O part of the model also
+resulted in substantially improved replication of key aspects of the
+observed data. Added flexibility in the distributional terms for both
+dissolved and equilibrium N2O also led to improvements. To make
+inferences to the population of interest from a model including these
+covariates, however, they needed to be (1) fully observed across that
+population or (2) their missingness needed to be modeled. For the lake
+area and elevation covariates, data *was* available for all lakes from
+previously compiled geospatial databases. However, neither surface
+temperature or NO3 were observed for lakes outside of the sample.
+Therefore, the final model below included surface temperature and NO3 as
+additional responses conditioned on the survey design variables and
+fully observed covariates.
+
+In the final model below, the sub-model for surface temperature assumed
+a Gamma distributed error distribution and the linear predictor included
+the survey design variables, latitude, elevation, and Julian date. The
+shape parameter was also modeled as a function of latitude to address
+increasing response variance along the latitudinal gradient. The NO3
+response was modeled with a cumulative logit likelihood and the linear
+predictor included all of the survey factors as well as surface
+temperature and lake area. The dissolved and equilibrium N2O responses
+were each modeled with Gamma distributed errors, but with the same
+covariate structure as in model 5. The same structure was also employed
+for the shape terms in these responses, corresponding to the $\sigma$
+terms in the previous models. Note that there was no observation-level
+residual correlation term for this model, since the residuals are
+undefined for the Gamma and cumulative logit models. Dropping the
+observation-level residual correlation term was deemed a reasonable
+compromise that enabled the inclusion of NO3 and surface temperature in
+the N2O model response. The random intercepts for the group terms,
+however, still allowed for potential correlations between responses at
+the group levels.
 
@@
 ### Summarize fit
+
 Below is a summary of the fitted parameters and MCMC diagnostics.
+
 ```{r print_mod6, echo=FALSE}
@@
 
-Below, the same PPCs for dissolved and equilibrium N2O were used as before.
+Below, the same PPCs for dissolved and equilibrium N2O were used as
+before.
 
@@
 
-The PPCs for dissolved N2O were similar to those for models 4 and 5 above.
+The PPCs for dissolved N2O were similar to those for models 4 and 5
+above.
 
@@
 pp_check(n2o_mod6, 
+         ndraws = 200,
          resp = "n2o",
          type = "dens_overlay",
-         nsamples = 40,
          cores = 1) + 
@@
 pp_check(n2o_mod6, 
+         ndraws = 200, 
          resp = "n2o",
          type = "ecdf_overlay",
-         nsamples = 40, 
          cores = 1) + 
@@
 pp_check(n2o_mod6, 
+         ndraws = 1000,
          resp = "n2o",
@@
 
+The same check for the re-visit data:
+
+```{r ppc_full_checks_mod_n2o6_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod6,
+         newdata = df_test,
+         ndraws = 200,
+         resp = "n2o",
+         type = "dens_overlay",
+         cores = 1) + 
+  theme_tidybayes() +
+  scale_x_continuous(trans = "log") +
+  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod6, 
+         ndraws = df_test,
+         nsamples = 200, 
+         resp = "n2o",
+         type = "ecdf_overlay",
+         cores = 1) + 
+  theme_tidybayes() +
+  scale_x_continuous(trans = "log") +
+  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod6, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "n2o",
+         type = "stat_2d", 
+         stat = c("mean", "sd"), 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+pp_check(n2o_mod6, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "n2o",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness"), 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+pp_check(n2o_mod6, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "n2o",
+         type = "stat_2d", 
+         stat = c("min", "max"), 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+pp_check(n2o_mod6, 
+         newdata = df_test,
+         ndraws = 1000,
+         resp = "n2o",
+         type = "scatter_avg", 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### Equilibrium N2O
 
-The PPCs for eqilibrium N2O were also similar to the same checks in models 4 and 5.
+The PPCs for eqilibrium N2O were also similar to the same checks in
+models 4 and 5.
 
@@
 pp_check(n2o_mod6, 
+         ndraws = 200,
          resp = "n2oeq",
          type = "dens_overlay",
-         nsamples = 40,
          cores = 1) + 
@@
 pp_check(n2o_mod6, 
+         ndraws = 200,
          resp = "n2oeq",
          type = "ecdf_overlay",
-         nsamples = 40, 
          cores = 1) + 
@@
 pp_check(n2o_mod6, 
+         ndraws = 1000,
          resp = "n2oeq",
@@
 
+The same check with the re-visit data:
+
+```{r ppc_full_checks_mod_n2oeq6_test, echo=FALSE, fig.align='center', fig.asp=1.25, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+grid.arrange(
+pp_check(n2o_mod6, 
+         ndraws = 200,
+         newdata = df_test,
+         resp = "n2oeq",
+         type = "dens_overlay",
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
+  ylab("density")
+,
+pp_check(n2o_mod6, 
+         ndraws = 200,
+         newdata = df_test,
+         resp = "n2oeq",
+         type = "ecdf_overlay",
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  theme_tidybayes() +
+  xlab(expression(paste("log(Equilibrium N"[2],"O) concentration"))) + 
+  ylab("cumulative density")
+,
+pp_check(n2o_mod6, 
+         ndraws = 1000,
+         newdata = df_test,
+         resp = "n2oeq",
+         type = "stat_2d", 
+         stat = c("mean", "sd"), 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+pp_check(n2o_mod6, 
+         ndraws = 1000,
+         newdata = df_test,
+         resp = "n2oeq",
+         type = "stat_2d", 
+         stat = c("kurtosis", "skewness"), 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+pp_check(n2o_mod6, 
+         ndraws = 1000,
+         newdata = df_test,
+         resp = "n2oeq",
+         type = "stat_2d", 
+         stat = c("min", "max"), 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+,
+pp_check(n2o_mod6, 
+         ndraws = 1000,
+         newdata = df_test,
+         resp = "n2oeq",
+         type = "scatter_avg", 
+         cores = 1) + 
+  scale_x_continuous(trans = "log") +
+  scale_y_continuous(trans = "log") +
+  theme_tidybayes()
+, ncol = 2)
+```
+
 #### Bivariate
 
-This final model provided a reasonable representation of the bivariate relationship between the two N2O responses.
-
-```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
+This final model provided a reasonable representation of the bivariate
+relationship between the two N2O responses.
+
+```{r ppc_bv_check_mod_n2o6, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
 df_model %>%
@@
                       resp = c("n2o","n2oeq"),
-                      ndraws = 50) %>%
+                      ndraws = 20) %>%
   rename(n2o_obs = n2o,
@@
 
+The check with the re-visit data:
+
+```{r ppc_bv_check_mod_n2o6_revisit, echo=FALSE, fig.align='center', fig.asp=1, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
+df_test %>%
+  add_predicted_draws(n2o_mod6,
+                      resp = c("n2o","n2oeq"),
+                      ndraws = 20) %>%
+  rename(n2o_obs = n2o,
+         n2oeq_obs = n2o_eq) %>%
+  tidyr::pivot_wider(names_from = .category,
+              values_from = .prediction) %>%
+  ggplot(aes(x = log(n2o_obs), y = log(n2oeq_obs))) +
+  geom_density_2d(aes(x = log(n2o), 
+                      y = log(n2oeq), 
+                      group = .draw),
+                  bins = 10,
+                  color = "lightblue", 
+                  alpha = 0.4) +
+  geom_density_2d(color = "black", bins = 10) +
+  xlim(1.25, 2.75) +
+  ylim(1.25, 2.75) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_tidybayes()
+```
+
 #### Saturation
 
-The saturation ratio PPCs below suggested similar behavior as with models 4 and 5 above, but with perhaps slightly less bias in the predictions for the proportion of undersaturated waterbodies and fewer extreme predictions for the means and standard deviations.
-The observed _vs._ predicted PPC also appears to have a better behaved variance and no extreme predictions, compared to models 4 and 5.
+The saturation ratio PPCs below suggested similar behavior as with
+models 4 and 5 above, but with perhaps slightly less bias in the
+predictions for the proportion of undersaturated waterbodies and fewer
+extreme predictions for the means and standard deviations. The observed
+*vs.* predicted PPC also appears to have a better behaved variance and
+no extreme predictions, compared to models 4 and 5.
 
@@
                       resp = c("n2o", "n2oeq"),
-                      ndraws = 50) %>%
+                      ndraws = 200) %>%
   rename(n2o_obs = n2o,
@@
                       resp = c("n2o", "n2oeq"), 
-                      ndraws = 500) %>%
+                      ndraws = 1000) %>%
   rename(n2o_obs = n2o,
@@
 
-The plot below shows the same PPC, but for the second-vist data.
+The check for the second-vist data:
 
@@
   group_by(.draw) %>%
-  summarise(prop_pred = sum(usat_pred) / 95) %>%
+  summarise(prop_pred = sum(usat_pred) / nrow(df_test)) %>%
   ggplot(aes(x = prop_pred)) +
   geom_histogram(binwidth = 0.001, fill = "lightblue") +
-  geom_vline(data = df_model, mapping = aes(xintercept = sum(n2o < n2o_eq)/984)) +
+  geom_vline(data = df_test, mapping = aes(xintercept = sum(n2o < n2o_eq)/nrow(df_test))) +
   xlab("Proportion undersaturated waterbodies") +
@@
 
-Below are estimates for the Bayesian $R^2$, which were largely similar for the N2O responses as with models 4 and 5 above. The $R^2$ for the surface temperature response also suggested a good fit.
+Below are estimates for the Bayesian $R^2$, which were largely similar
+for the N2O responses as with models 4 and 5 above. The $R^2$ for the
+surface temperature response also suggested a good fit.
 
@@
 
-Below are the $R^2$ estimates for the second-visit data.
-That these estimates were similar to those for the data used to fit the model was encouraging and suggested that the model may perform reasonably well out-of-sample.
+Below are the $R^2$ estimates for the second-visit data. That these
+estimates were similar to those for the data used to fit the model was
+encouraging and suggested that the model may perform reasonably well
+out-of-sample.
 
@@
 
-The conditional effects plot for the covariate effects on dissolved N2O suggested a similar effect of NO3 as in previous models, but with interesting potential interactions between NO3 and lake area and NO3 and surface temperature.
-The lake area effect was estimated to be larger and more negative at the highest levels of NO3 and slightly negative at the lowest level of NO3.
-The surface temperature effect was estimated to be largest and positive at the highest level of NO3 and negative at the lowest level of NO3.
+The conditional effects plot for the covariate effects on dissolved N2O
+suggested a similar effect of NO3 as in previous models, but with
+interesting potential interactions between NO3 and lake area and NO3 and
+surface temperature. The lake area effect was estimated to be larger and
+more negative at the highest levels of NO3 and slightly negative at the
+lowest level of NO3. The surface temperature effect was estimated to be
+largest and positive at the highest level of NO3 and negative at the
+lowest level of NO3.
 
@@
 
-The estimated covariate effects on $\sigma$ for dissolved N2O suggested a negative relationship with log(area) and a positive relationship with NO3.
+The estimated covariate effects on $\sigma$ for dissolved N2O suggested
+a negative relationship with log(area) and a positive relationship with
+NO3.
 
@@
 
-Below are the estimated conditional effects of the covariates on means for both n2o responses as well as the implied effects on the mean saturation ratio.
+Below are the estimated conditional effects of the covariates on means
+for both n2o responses as well as the implied effects on the mean
+saturation ratio.
 
@@
 
+```{r conditional_effects_mod_n2osat_temp_area_F, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+effect_no3_surftemp <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               surftemp = round(quantile(df_model$surftemp, 
+                                                   probs = seq(0.10, 0.90, 0.2)), 0),
+                               log_area = mean(df_model$log_area),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         surftemp = factor(surftemp))
+
+effect_surftemp_no3 <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               surftemp = seq(range(df_model$surftemp)[1],
+                                              range(df_model$surftemp)[2],
+                                              length.out = 101),
+                               log_area = mean(df_model$log_area),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         surftemp = surftemp)
+
+effect_no3_area <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               log_area = round(quantile(df_model$log_area, 
+                                                   probs = seq(0.10, 0.90, 0.2)), 0),
+                               surftemp = mean(df_model$surftemp),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         log_area = factor(log_area))
+
+effect_area_no3 <- expand.grid(no3_cat = levels(df_model$no3_cat),
+                               log_area = seq(range(df_model$log_area)[1],
+                                              range(df_model$log_area)[2],
+                                              length.out = 101),
+                               surftemp = mean(df_model$surftemp),
+                               log_elev = mean(df_model$log_elev)) %>%
+  add_epred_draws(n2o_mod6, resp = c("n2o", "n2oeq"), re_formula = ~ 1) %>%
+  pivot_wider(names_from = .category, values_from = .epred) %>%
+  mutate(n2o_pred = n2o,
+         n2oeq_pred = n2oeq,
+         sat_pred = n2o / n2oeq,
+         log_area = log_area)
+
+plt_sat_no3_temp <-
+  effect_no3_surftemp %>%
+  ggplot(aes(x = no3_cat, 
+             y = sat_pred, 
+             group = surftemp,
+             color = surftemp)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  scale_color_brewer(palette = "PuBu") +
+  xlab(expression(paste("NO"[3], " category"))) +
+  ylab("") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  theme(legend.position = c(0.4, 0.6),
+        axis.text.y = element_blank()) +
+  labs(color = "Surface temperature (\u00B0C)")
+
+plt_sat_temp_no3 <-
+  effect_surftemp_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = surftemp,
+                      y = sat_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  scale_color_brewer(palette = "OrRd") +
+  xlab("Surface temperature (\u00B0C)") +
+  ylab("Saturation ratio") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  scale_fill_brewer(palette = "Greys", guide = "none") +
+  theme(legend.position = c(0.4, 0.6)) + 
+  labs(color = expression(paste("NO"[3], " category")))
+
+plt_sat_no3_area <-
+  effect_no3_area %>%
+  ggplot(aes(x = no3_cat, 
+             y = sat_pred, 
+             group = log_area,
+             color = log_area)) +
+  stat_pointinterval(shape = 95,
+                     point_interval = "median_qi",
+                     position = position_dodge(width = 1.4, preserve = "single")) +
+  scale_color_brewer(palette = "Greys") +
+  xlab(expression(paste("NO"[3], " category"))) +
+  ylab("") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  theme(legend.position = c(0.4, 0.6),
+        axis.text.y = element_blank()) +
+  labs(color = "log(surface area (ha))")
+
+plt_sat_area_no3 <-
+  effect_area_no3 %>%
+  ggplot() +
+  stat_lineribbon(aes(x = log_area,
+                      y = sat_pred, 
+                      group = no3_cat,
+                      color = no3_cat),
+                  point_interval = "median_qi",
+                  show.legend = TRUE,
+                  size = 0.7,
+                  .width = 0.95) +
+  scale_color_brewer(palette = "OrRd") +
+  xlab("log(surface area (ha))") +
+  ylab("Saturation ratio") +
+  theme_pubr() +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 7), breaks = seq(0, 7, 1)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  scale_fill_brewer(palette = "Greys", guide = "none") +
+  theme(legend.position = c(0.6, 0.6)) + 
+  labs(color = expression(paste("NO"[3], " category")))
+
+plt_temp_area_effects <- ggarrange(plt_sat_temp_no3,
+                                  plt_sat_no3_temp,
+                                  plt_sat_area_no3,
+                                  plt_sat_no3_area,
+                             ncol = 2,
+                             nrow = 2,
+                             widths = c(0.9, 0.9, 0.9, 0.9),
+                             heights = c(0.9, 0.9, 0.9, 0.9),
+                             labels = c("A", "B", "C", "D"),
+                             hjust = -7,
+                             vjust = 3
+                             )
+
+plt_temp_area_effects <- plt_temp_area_effects %>%
+  annotate_figure(plt_temp_area_effects,
+                  left = text_grob("Derived posterior mean", 
+                                   size = 18, 
+                                   rot = 90)
+                  )
+
+# ggsave only works in R console (i.e., copy + paste to save)
+ggsave(filename = "cond_effect_no3.tif",
+       plot = plt_temp_area_effects,
+       device = "tiff",
+       width = 210 / 25.4,
+       height = 210 / 25.4,
+       units = "in",
+       bg = "white")
+```
+
+```{r plot_temp_area_effects, fig.align='center', fig.asp=1.166667, fig.width=8, }
+plt_temp_area_effects
+```
+
 # Predict to population
 
-As mentioned above, in order to make inferences to the population of interest, the final model above was used to, first, predict surface temperature in the target population, since it depended only on the fully observed covariates.
-Next, the predictive distribution for surface temperature was used, along with the relevant fully observed covariates, to predict NO3 in the target population.
-Finally, the predictive distributions for temperature and NO3 were used to predict the N2O responses.
-The code for these steps is outlined in the following.
-
-The first step used the final model to predict surface temperature to the population:
+As mentioned above, in order to make inferences to the population of
+interest, the final model above was used to, first, predict surface
+temperature in the target population, since it depended only on the
+fully observed covariates. Next, the predictive distribution for surface
+temperature was used, along with the relevant fully observed covariates,
+to predict NO3 in the target population. Finally, the predictive
+distributions for temperature and NO3 were used to predict the N2O
+responses. The code for these steps is outlined in the following.
+
+The first step used the final model to predict surface temperature to
+the population:
 
@@
 
-NO3 was next predicted using the posterior predictions for surface temperature and the other fully observed covariates.
-Note that only one posterior draw per posterior predictive draw for surface temperature was used in order to reduce excess simulations.
+NO3 was next predicted using the posterior predictions for surface
+temperature and the other fully observed covariates. Note that only one
+posterior draw per posterior predictive draw for surface temperature was
+used in order to reduce excess simulations.
 
@@
 
-Finally, dissolved and equilibrium N2O were predicted using the surface temperature and NO3 predictions, along with the survey variables and other fully observed covariates.
+Finally, dissolved and equilibrium N2O were predicted using the surface
+temperature and NO3 predictions, along with the survey variables and
+other fully observed covariates.
 
@@
 
-Finally, the predictions for all four partially observed responses were assembled into a new dataframe for use in inference:
+Finally, the predictions for all four partially observed responses were
+assembled into a new dataframe for use in inference:
 
@@
 
-A number of estimates for the target population were assembled and presented below.
-First, the full posterior predictive distributions for dissolved N2O, equilibrium N2O, and the saturation ratio were assessed.
-These distributions summarized the predicted distribution of concentrations or ratios for all lakes in the population of interest and included parameter uncertainty propagated through the model.
-Next, population means were assessed, followed by comparisons of some model-based estimates to previously calculated design-based estimates.
+A number of estimates for the target population were assembled and
+presented below. First, the full posterior predictive distributions for
+dissolved N2O, equilibrium N2O, and the saturation ratio were assessed.
+These distributions summarized the predicted distribution of
+concentrations or ratios for all lakes in the population of interest and
+included parameter uncertainty propagated through the model. Next,
+population means were assessed, followed by comparisons of some
+model-based estimates to previously calculated design-based estimates.
 
@@
 
-Below, a density plot summarized the posterior predictive distributions across the target population of lakes.
-The summaries of the PPDs were based on 2000 draws.
-Note that the x-axis was truncated at 50 nmol/L for a clearer visualization of the bulk of the predictive distribution.
-For reference, the max predicted value was 4403.2 nmol/L for dissolved N2O, 20.4 nmol/L for dissolved N2O, and 793.5 for the saturation ratio.
+Below, a density plot summarized the posterior predictive distributions
+across the target population of lakes. The summaries of the PPDs were
+based on 2000 draws. Note that the x-axis was truncated at 50 nmol/L for
+a clearer visualization of the bulk of the predictive distribution. For
+reference, the max predicted value was 4403.2 nmol/L for dissolved N2O,
+20.4 nmol/L for dissolved N2O, and 793.5 for the saturation ratio.
 
@@
 
-Below are density plots summarizing the posterior distribution of _means_ for N2O concentrations and the saturation ratio for all US lakes and reservoirs > 4ha in the lower 48 states).
+Below are density plots summarizing the posterior distribution of
+*means* for N2O concentrations and the saturation ratio for all US lakes
+and reservoirs \> 4ha in the lower 48 states).
 
@@
 
-To illustrate the skewness in the predictive distribution for the saturation ratio, an estimate for the median ratio is shown below.
-The entire posterior distribution of the mean above was larger than 1, which represents the boundary of under- _vs._ oversaturation.
-By comparison, the posterior estimate of the median below only included values less than one, suggesting that though the mean saturation ratio was likely greater than 1, most lakes in the national population were undersaturated (i.e., ratio less than 1).
-In distributions with strong right-skew, the mean can often be considerably larger than the median.
+To illustrate the skewness in the predictive distribution for the
+saturation ratio, an estimate for the median ratio is shown below. The
+entire posterior distribution of the mean above was larger than 1, which
+represents the boundary of under- *vs.* oversaturation. By comparison,
+the posterior estimate of the median below only included values less
+than one, suggesting that though the mean saturation ratio was likely
+greater than 1, most lakes in the national population were
+undersaturated (i.e., ratio less than 1). In distributions with strong
+right-skew, the mean can often be considerably larger than the median.
 
@@
 
-Below is a plot of the posterior mean estimate for the proportion of unsaturated lakes at the national scale.
+Below is a plot of the posterior mean estimate for the proportion of
+unsaturated lakes at the national scale.
 
@@
 
-Below are posterior estimates of the means for dissolved and equilibrium N2O and the saturation ratio by WSA9 ecoregion.
+Below are posterior estimates of the means for dissolved and equilibrium
+N2O and the saturation ratio by WSA9 ecoregion.
 
@@
 
-A plot of the posterior estimates for the median saturation ratio below indicated that most lakes in each ecoregion were undersaturated (i.e., median < 1).
+A plot of the posterior estimates for the median saturation ratio below
+indicated that most lakes in each ecoregion were undersaturated (i.e.,
+median \< 1).
 
@@
 
-A plot of the estimates of the proportion of under-saturated lakes by ecoregion is below.
-These summaries again suggested that most lakes in each ecoregion were likely undersaturated (i.e., median << 1).
+A plot of the estimates of the proportion of under-saturated lakes by
+ecoregion is below. These summaries again suggested that most lakes in
+each ecoregion were likely undersaturated (i.e., median \<\< 1).
 
@@
 
-Comparisons of mean estimates by state are below. Density polygons were not included to minimize the vertical plot space.
+Comparisons of mean estimates by state are below. Density polygons were
+not included to minimize the vertical plot space.
 
@@
 
-Below is a plot of estimates for the mean (black circles) and median (grey circles) saturation ratio by state.
-A horizontal, dashed, black line is shown at ratio = 1, indicating the boundary for under- _vs._ oversaturation.
-Only a few states (e.g., NV, DE) had median estimates that were 1 or greater, suggesting that, for most states, most lakes were estimated to be undersaturated.
+Below is a plot of estimates for the mean (black circles) and median
+(grey circles) saturation ratio by state. A horizontal, dashed, black
+line is shown at ratio = 1, indicating the boundary for under- *vs.*
+oversaturation. Only a few states (e.g., NV, DE) had median estimates
+that were 1 or greater, suggesting that, for most states, most lakes
+were estimated to be undersaturated.
 
@@
 
-Finally, a plot of the estimated proportion of undersaturated lakes for each state in the target population.
-Point estimates are the posterior median of the proportion and bars are the upper and lower boundaries of the central 95th percentile of the posterior distributions of proportions.
+Finally, a plot of the estimated proportion of undersaturated lakes for
+each state in the target population. Point estimates are the posterior
+median of the proportion and bars are the upper and lower boundaries of
+the central 95th percentile of the posterior distributions of
+proportions.
 
@@
 
-The estimated means and medians by size category are below for dissolved and equilibrium N2O and the saturation ratio.
+The estimated means and medians by size category are below for dissolved
+and equilibrium N2O and the saturation ratio.
 
@@
 
-Mean _vs._ median below.
+Mean *vs.* median below.
 
@@
 
-And, finally, the estimated proportion of undersaturated lakes in the target population by size category.
+And, finally, the estimated proportion of undersaturated lakes in the
+target population by size category.
 
@@
 
-## Model- _vs._ design-based
-
-Below, estimates from the model-based approach are compared to previously calculated, design-based estimates.
-In general, the model-based estimates were similar to the design-based estimates.
-The model-based estimates were typically within the confidence bounds of the design-based estimates, but with much greater precision.
-Improved precision was expected due to the "shrinkage" induced by the multilevel parameterization, which affords some "borrowing" of information across the various levels of the survey factors.
+## Model- *vs.* design-based
+
+Below, estimates from the model-based approach are compared to
+previously calculated, design-based estimates. In general, the
+model-based estimates were similar to the design-based estimates. The
+model-based estimates were typically within the confidence bounds of the
+design-based estimates, but with much greater precision. Improved
+precision was expected due to the "shrinkage" induced by the multilevel
+parameterization, which affords some "borrowing" of information across
+the various levels of the survey factors.
 
@@
 
-Below, National mean estimates for dissolved N2O from the model and design-based approaches were compared.
-The sample-based estimate was also included as a reference.
-The black, vertical, dashed line indicates the mean of the sample.
+Below, National mean estimates for dissolved N2O from the model and
+design-based approaches were compared. The sample-based estimate was
+also included as a reference. The black, vertical, dashed line indicates
+the mean of the sample.
 

### 2024-03-20 manuscript/manuscript_support.R
minor tweaks while attempting to add legend to fig2

@@ -347,31 +347,40 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   # using xlim to zoom in on values close to 1
   
   dummy <- ecoreg_stats %>%
-    #CPL   NAP   NPL    SAP    SPL   TPL  UWM    WMT    XER
+                    #CPL   NAP   NPL    SAP    SPL   TPL  UWM    WMT    XER
     mutate(dens.med = c(2.5,  4,    3.1,   3.5,   2.5,  3,   3,      2,     2.5),
            dens.mean = c(.5,  3,    2.5,   1.5,   1,    1,   2,      1.5,   1.1),
+           # dens.equ = rep(1, 9),
+           # `median legend` = "median",
+           # `mean legend` = "mean",
+           # `eq legend` = "equilibrium",
            .draw = 1) # needed to match grouping aesthetic in ggplot call
   
   tic() # 60 seconds for 100 draws, and 16 minutes for 2000 draws, on memory intensive DMAP, 2797 secs for 2000 draws on Precision workstation
+  # After spending many hours trying to get the legend right, I gave up.  Write
+  # image to disk, then add legend in powerpoint.
   all_predictions_ms %>%
     #filter(.draw %in% 1:100) %>% # subset for practice
-    ggplot(aes(x = n2osat, group = .draw, color = .draw ) ) +
-    geom_density(show.legend = FALSE) +
+    ggplot(aes(x = n2osat, group = .draw, color = .draw )) +
+    geom_density(show.legend = FALSE) + # this argument doesn't work when faceted, hence guides call below
     geom_vline(xintercept = 1, color = "red") +
     geom_segment(data = dummy,
-                 aes(x = post_m_mean_sat, xend = post_m_mean_sat, 
+                 aes(x = post_m_mean_sat, xend = post_m_mean_sat,
                      y = 0, yend = dens.mean),
-                 linetype = "solid") +
+                 color = "black", linetype = "solid") +
     geom_segment(data = dummy,
-                 aes(x = post_m_median_sat, xend = post_m_median_sat, 
+                 aes(x = post_m_median_sat, xend = post_m_median_sat,
                      y = 0, yend = dens.med),
-                 linetype = "dashed") +
-    ylab ("density") + 
+                     color = "black", linetype = "dotted") +
+    # geom_segment(#data = dummy,
+    #              aes(x = dens.equ, xend = dens.equ,
[…]

### 2024-03-20 manuscript/manuscript_support.R
stashing minor changes to manuscript_support.R to facilitate merge with master

@@ -106,7 +106,7 @@ states <- USAboundaries::us_states() %>%
 
 
 ## load population data----
-#tic() # 7.6min on DMAP
+#tic() # 7.6min on DMAP.  5 minutes on Dell Precision workstation
 if(localPath == ""){ # if DMAP, then
   load("../../shared/jbeaulie/all_predictions.rda")
 } else { # if not DMAP
@@ -328,7 +328,7 @@ if(!("figure1.png" %in% list.files("manuscript/manuscript_figures"))) {
     # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
     scale_fill_manual("Ecoregion", values = cols,
                       guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("white", "black"), name = "source/sink") +
+    scale_color_manual(values = c("grey", "black"), name = "source/sink") +
     scale_size(name = expression(N[2]*O~(nM)),
                range = c(0.1, 10), # custom size range
                breaks = c(1, 10, 25, 50, 100)) + # custom breaks
@@ -354,7 +354,7 @@ if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   
   tic() # 60 seconds for 100 draws, and 16 minutes for 2000 draws, on memory intensive DMAP, 2797 secs for 2000 draws on Precision workstation
   all_predictions_ms %>%
-    #filter(.draw %in% 1:100) %>% # subset for practice
+    filter(.draw %in% 1:100) %>% # subset for practice
     ggplot(aes(x = n2osat, group = .draw, color = .draw ) ) +
     geom_density(show.legend = FALSE) +
     geom_vline(xintercept = 1, color = "red") +

### 2024-03-21 manuscript/manuscript_file.Rmd
Updates to manuscript methods text and support file Figure 5

@@
 
-The survey samples were treated statistically as a randomly selected subset of the complete POI, but with a pattern of "missingess" prescribed by the selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
+The survey samples were treated as a randomly selected subset of the POI, but with a pattern of "missingess" prescribed by the selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
 For inferences, estimates were statistically adjusted to account for the selection mechanism using a Multilevel regression with poststratification (MRP) approach [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020].
 MRP is a model-based approach carried out in two steps.
-In the first step, the sampled responses of interest (*e.g.*, N~2~O) are conditioned on the survey design variables (*e.g.*, ecoregion, state, size) in a multilevel regression model.
-Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the population of interest [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
-In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (*e.g.,* US waterbodies) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
-In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (*e.g.*, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI (See SI) and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
+In the first step, the sampled response (*e.g.*, N~2~O) is conditioned on the survey design variables (*e.g.*, ecoregion, state, size) in a multilevel regression model.
+Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the POI [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
+In the second step of MRP, the fitted regression parameters are used to "poststratify" or re-weight the response based on a known or assumed distribution of units (*e.g.,* US lakes) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+Although it is typical to postratify estimates at a grouping level above the observational unit (*e.g.*, ecoregion), in this study, N~2~O flux needed to be calculated as a function of lake-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$).
+Therefore, predictions were made to each of the 465,897 individual lakes in the assumed POI.
+For group level inferences, the posterior predictive distribution of the response was summarized across all lakes in the groupings of interest.
 The fully Bayesian implementation made propagation of uncertainty from the waterbody-level predictions to higher level summaries straightforward.
 
-The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N~2~O concentrations; (2) the N~2~O saturation ratio; and (3) the proportion of under-saturated water bodies (*i.e.*, saturation ratio \< 1).
+The multilevel regression model was used to make probabilistic predictions for (1) dissolved and equilibrium N~2~O concentrations; (2) the N~2~O saturation ratio; and (3) the proportion of under-saturated water bodies (*i.e.*, saturation ratio \< 1).
 Saturation ratio was calculated as the ratio of dissolved to equilibrium N~2~O.
-Because dissolved and equilibrium N~2~O were each measured on the same sample units (*i.e.*, a single waterbody), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, common dependencies on geography.
+Because dissolved and equilibrium N~2~O were each measured on the same sample units (*i.e.*, individual lakes), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, conditional dependencies on geography.
 The modeled response variable was, therefore, multivariate.
-Although the mean marginal probabilities estimated from separate models may have provided comparable estimates, the joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
-The saturation ratio was not modeled or predicted directly, but was estimated *post hoc* as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
-Similarly, the proportion of under-saturated waterbodies was estimated from this PPD as the proportion of waterbodies with a saturation ratio less than 1.
-All models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models using Hamiltonian Monte Carlo (HMC).
+Although the mean marginal probabilities estimated from separate models may have provided comparable estimates, the joint model allowing correlated responses was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
 
-Fully reproducible code and further results and model workflow details are provided in the Modeling Supplement Document.
-Though our initial models conditioned the response on just the survey design variables, those models consistently resulted in poorer fits to to the observed N~2~O data compared to models with additional covariates.
-Though the initial models seemed to work well at replicating the central tendency of the dissolved N~2~O data, simulations of other properties of the observed data, such as the variance, skewness, and kurtosis, were less satisfactory.
-The relatively poor fit to this aspect of the observed data also resulted in a downward bias in estimates of the observed saturation ratio.
-However, including a measure of NO~3~ concentration as a covariate on dissolved N~2~O resulted in substantially improved posterior predictive model checks.
-In addition, including surface temperature and elevation measures as covariates on equilibrium N~2~O resulted in substantially improved fits, as did adding flexibility in the distributional terms for both dissolved and equilibrium N~2~O.
-Both dissolved and equilibrium N~2~O were allowed to vary according to the survey design variables and environmental covariates in the final model presented here.
+The saturation ratio was not modeled or predicted directly, but was estimated *post hoc* as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
+Similarly, the proportion of under-saturated lakes was calculated from this PPD as the proportion of lakes with a saturation ratio less than 1.
+All models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models using Hamiltonian Monte Carlo (HMC).Fully reproducible code, results, and model workflow details are provided in the Modeling Supplement Document.
 
-Including these helpful covariates, however, led to some complexities in the MRP process.
-To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled.
+Our initial models conditioned the response on just the survey design variables, however, those models consistently resulted in poor fits to some aspects of the observed N~2~O data.
+These models worked reasonably well at replicating the central tendency of the observed N~2~O data, but other properties of the observed distributions, such as the variance, skewness, and kurtosis, proved difficult to replicate.
+A consequence of the relatively poor fit to the tail properties of the data was a downward bias in estimates of the saturation ratio.
+In subsequent models, a measure of NO~3~ levels was included as a covariate on the dissolved N~2~O response, which substantially improved the fit according to the posterior predictive model checks.
+In addition, including surface temperature and elevation measures as covariates on equilibrium N~2~O resulted in improvements, as did allowing for heterogeneous errors for both dissolved and equilibrium N~2~O.
+Both dissolved and equilibrium N~2~O were allowed to vary according to the survey design variables and environmental covariates.
+
+Including covariates, however, necessitated additional steps in the MRP process.
+To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled as with the N2O responses.
 For the lake area and elevation covariates, data was available for all lakes in the POI from previously compiled geospatial databases.
@@
 Therefore, a more complex model was developed to include surface temperature and NO~3~ as additional responses conditioned on the survey design variables and the fully observed covariates.
-This approach to inference was akin to Bayesian structural equation modeling approaches [@Merkle_etal_2021; @Merkle_Rosseel_2018].
-The dependence structure of the represented by this model may be characterized as:
+This approach was similar to MRP approaches with "non-census" variables [@Kastellec_etal_2015].
+It could also be described as a form of Bayesian structural equation model [@Merkle_etal_2021; @Merkle_Rosseel_2018].The dependence structure of this final model can be characterized where:
 

### 2024-03-21 manuscript/manuscript_support.R

@@ -525,7 +525,7 @@ if(!("n2oFluxAndEmissionRateVsContinuousArea.tiff" %in% list.files("manuscript/m
   b2 <- ggplot(b2.dat, aes(x = area_ha, y = post_m_pred)) +
     geom_errorbar(aes(xmin = area_ha, ymin = LCL_pred,
                       xmax = area_ha, ymax = UCL_pred),
-                  size = 0.1) + # size with ggplot 3.3.3.  linewidth with other versions?
+                  linewidth = 0.1) + # size with ggplot 3.3.3.  linewidth with other versions?
     geom_point(size = 0.1, color = "red") +
     scale_x_log10(labels=scales::comma) + 
     scale_y_continuous(trans = ggallin::pseudolog10_trans, 
@@ -592,12 +592,15 @@ if(!("n2oFluxAndEmissionRateVsContinuousArea.tiff" %in% list.files("manuscript/m
     group_by(.row) %>% 
     mean_qi()
 
-  b3 <-  ggplot(b3.dat, aes(x = area_ha, y = p_total_f, ymin = p_total_f.lower, ymax = p_total_f.upper)) +
-    tidybayes::geom_pointinterval(point_color = 'red', point_size = 0.75, linewidth = 0.5) + 
+  b3 <-  ggplot(b3.dat, aes(x = area_ha, y = p_total_f)) +
+    geom_errorbar(aes(xmin = area_ha, ymin = p_total_f.lower,
+                      xmax = area_ha, ymax = p_total_f.upper),
+                  linewidth = 0.1) + 
+    geom_point(size = 0.1, color = "red") +
     scale_x_log10(labels = scales::comma) +
     coord_cartesian(ylim = c(-3, 3)) +
     xlab("Waterbody size (ha)") +
-    ylab(expression(atop(Proportion~of, CONUS~flux))) +
+    ylab(expression(atop(Proportion, of~CONUS~flux))) +
     theme_bw()  +
     theme(panel.grid.major = element_blank(), 
           panel.grid.minor = element_blank())

### 2024-03-21 manuscript/manuscript_file.Rmd
response to technical reviews

--- 2023-12-14 manuscript/manuscript_file.Rmd
@@
 
-## Psuedo abstract
-
-**Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere [@DelSontro_etal_2018a], but recent evidence of N~2~O uptake across a broad range of waterbodies have called the accuracy of emission estimates into question.
+**Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere but recent evidence of N~2~O uptake across a broad range of lakes have called the accuracy of emission estimates into question.
 Here we use a new national-scale dataset on dissolved N~2~O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` waterbodies in the conterminous U.S.
 (CONUS).
-We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of waterbodies functioning as N~2~O sinks.
-The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate (NO~3~) concentration, waterbody size, and water temperature.
-Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS waterbodies is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
-Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS waterbodies are a net source or sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source.**
+We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of lakes functioning as N~2~O sinks.
+The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate concentration, waterbody size, and water temperature.
+Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS lakes is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
+Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS lakes are a net source or sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source.**
 
@@
 
-Lakes, reservoirs, and ponds (waterbodies hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], a potent greenhouse gas and contributor to stratospheric ozone destruction.
-It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across waterbodies in Canada [@Soued_etal_2016; @Webb_etal_2019] and Finland [@Kortelainen_etal_2020] have called this assumption into question.
-Furthermore, detailed investigations of individual waterbodies have shown that lentic N~2~O biogeochemistry is dynamic, with some waterbodies alternating between N~2~O sources and sinks in response to nitrogen availability, thermal stratification, and other factors [@Beaulieu_etal_2015; @Beaulieu_etal_2014; @Musenze_etal_2014].
+Lakes, reservoirs, and ponds (lakes hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], a potent greenhouse gas and contributor to stratospheric ozone destruction.
+It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across lakes in Canada [@Soued_etal_2016; @Webb_etal_2019], the United States [Aho_et_al_2023], and Finland [@Kortelainen_etal_2020] have called this assumption into question.
+These recent studies have been conducted at relatively small spatial scales, or with a relatively small number of lakes, and the prevalence of N~2~O uptake in lakes at national to continental scales is unknown.
 
-While several measurement campaigns have demonstrated that waterbodies can function as N~2~O sinks[@Soued_etal_2016; @Webb_etal_2019; @Kortelainen_etal_2020; @Beaulieu_etal_2015; @Beaulieu_etal_2014; @Musenze_etal_2014], these studies have been conducted at limited spatial scales and encompassed small ranges in potential environmental drivers (e.g.
-nitrogen loading).
-Nitrous oxide dynamics have been investigated in streams and rivers spanning a broad range of environmental conditions, but these systems have been consistently found to be N~2~O sources [@Beaulieu_etal_2008; @Beaulieu_etal_2010; @Quick_etal_2019] and the controls on N~2~O biogeochemistry likely differ between flowing and standing waters.
-These limitations in data availability have complicated efforts to incorporate N~2~O consumption into freshwater N~2~O budgets.
+Nitrous oxide can be produced and consumed via microbially mediated nitrogen transformations occurring in the sediment and water column of lakes.
+The balance between N~2~O production and consumption determines whether a lake functions as a N~2~O source or sink to the atmosphere.
+Nitrous oxide production and consumption rates are determined by a broad range of factors including inorganic nitrogen, dissolved organic carbon, and oxygen concentrations, among others, making it difficult to predict the balance between these processes[@Beaulieu_etal_2015;@Soued_etal_2016;@Webb_etal_2019].
+Furthermore, N~2~O production and consumption rates can vary spatially and temporally within a lake, with some waterbodies even alternating between N~2~O sources and sinks through time[@Beaulieu_etal_2014;@Musenze_etal_2014], adding additional challenges to predicting the source/sink status of lakes.
+
+Here we present results from the first national-scale survey of N~2~O concentrations and emissions from lakes in the continental U.S.
+(CONUS).
+Samples sites were selected using a generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] and the observational data were used to develop a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS lakes.
+Combining the largest freshwater N~2~O dataset with a hierarchical model enabled us to estimate the prevalence of N~2~O uptake in CONUS lakesfor the first time and to identify important environmental controls on N~2~O biogeochemistry.
 
@@
 
-In this study we present the first national-scale survey of N~2~O concentrations and emissions from CONUS waterbodies.
-A generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] was used to select a spatially-balanced and representative sample of `r dg %>% filter(sitetype == "PROB") %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` sites from CONUS waterbodies that were at least 1 hectare in area, at least 1 meter deep, had at least 0.1 hectare of open water, and had a minimum residence time of one week[@EPAdesign] **(Fig.
+During the summer of 2017 the United States Environmental Protection Agency conducted the first national-scale survey of N~2~O concentrations and emissions in CONUS lakes as part of the National Lakes Assessment (NLA) [@USEPA_NLA17_2022].
+The population of interest (POI) for the survey was `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS lakes that had a surface area of at least 1 ha.
+A representative sample (n=`r dg %>% filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>% nrow()`) was selected from the POI using a generalized random-tessellation stratified survey design [@Stevens_Olsen_2004]**(Fig.
 1)**.
-Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate [@epa_data].
-`r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
+The survey design was stratified by state and used unequal probability categories based on lake size to ensure that large lakes, which are far less common than small lakes, were included in the study.
+Sample sites were distributed across all nine major CONUS ecoregions [@Herlihy_etal_2008] and varied widely in water chemistry, morphometry, watershed land-use, and climate [@epa_data].
+The sample data and survey design information (e.g., stratification variables and unequal probability categories) were used to develop a Bayesian hierarchical model to predict dissolved and equilibrium N~2~O concentrations for the POI.
+Predictor variables for the model were restricted to those required to generate predictions that reproduced the observed N~2~O measures in an out-of-sample dataset and were available for the POI.
+See Methods and SI for a detailed discussion of variable selection.
+
+Nearly `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
 1)**.
-These data, along with the survey design information, were used to develop a Bayesian hierarchical model to predict N~2~O concentrations for all `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS waterbodies in the target population.
-At the population level, the majority of CONUS waterbodies (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% CI: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`) were predicted to be a net sink for N~2~O.
-The Western Mountains ecoregion had the smallest proportion of waterbodies predicted as N~2~O sinks (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(UCL_prop_sat) * 100`) and the Northern Plains had the greatest (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(UCL_prop_sat) * 100`), although the 95% credible intervals over the predictive distributions overlap across all ecoregions **(Table 1)**.
+The Bayesian hierarchical model predicted that the majority lakes in the POI (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% credible interval (CI): `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`) functioned as net N~2~O sinks during the summer months.
+The Western Mountains ecoregion had the smallest proportion of lakes predicted as N~2~O sinks (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(UCL_prop_sat) * 100`) and the Northern Plains had the greatest (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(UCL_prop_sat) * 100`), although the 95% credible intervals over the predictive distributions overlap across all ecoregions **(Table 1)**.
 
 ![Figure 1.
-Location of sampling sites.
+Location of sampling sites with an underlay of the nine aggregated level III US ecoregions.
 Color of the points indicates whether the waterbody was functioning as a source or sink of nitrous oxide (N~2~O).
@@
 
-The estimated extent of N~2~O undersaturation in CONUS waterbodies was surprisingly high, particularly for waterbodies in nitrogen rich ecoregions that are managed for agricultural production.
-This finding runs counter to the long-standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lentic waterbodies can function as N~2~O sinks.
+The estimated extent of N~2~O undersaturation in CONUS lakes was surprisingly high, particularly for nitrogen rich ecoregions that are managed for agricultural production.
+This finding runs counter to the long-standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lakes can function as N~2~O sinks.
 Webb et al.
@@
 Similarly, 29% of surveyed boreal lakes in Finland were found to be N~2~O sinks during the summer [@Kortelainen_etal_2020].
-While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lentic waterbodies function as N~2~O sinks during the summer months.
+While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lakes function as N~2~O sinks during the summer months.
 This pattern held across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use.
@@
   set_header_labels('waterbody surface area' = "total waterbody surface area\n(hectares)",
-                    'N2O undersaturation extent' = "N2O undersaturation extent\n(% waterbodies)",
+                    'N2O undersaturation extent' = "N2O undersaturation extent\n(% lakes)",
                     'N2O emission rate' = "N2O emission rate\n(mg N2O m-2 d-1)",
@@
            ref_symbols = ", b",
-           value = as_paragraph("Percent of population with N2O saturation ratio < 1."),
+           value = as_paragraph("Percent of population of interest with N2O saturation ratio < 1."),
            part = "header",
@@
 
-Despite widespread N~2~O undersaturation across CONUS waterbodies, the estimated means for N~2~O saturation ratio (ratio of measured to equilibrium concentration) were indistinguishable from one in all ecoregions **(Table 1, Fig.
+Despite widespread N~2~O undersaturation across the POI, estimated means for N~2~O saturation ratio (ratio of measured to equilibrium concentration) were indistinguishable from a value of one in all ecoregions **(Table 1, Fig.
 2)**.
-The apparent contradiction of widespread N~2~O undersaturation with average dissolved N~2~O concentrations closer to equilibrium can be attributed to the prominent right skew predicted for dissolved N~2~O concentrations across the population of CONUS waterbodies **(Fig.
+The apparent contradiction of widespread N~2~O undersaturation and average dissolved N~2~O concentrations close to equilibrium can be attributed to the prominent right skew in the predicted distribution of dissolved N~2~O concentrations across the POI **(Fig.
 2)**.
-Across the CONUS, and within each of the nine major ecoregions, each predicted distribution of N~2~O saturation ratios had a median value less than 1, indicating that most waterbodies were undersaturated in these populations.
-However, the long right tails on these distributions, comprised of relatively rare waterbodies with very high levels of N~2~O supersaturation, were sufficient to drive the population means to values close to, or even greater than 1.
-Thus, most waterbodies in the CONUS were functioning as N~2~O sinks during the summer months, but a few functioned as strong N~2~O sources, particularly at high NO~3~ and high water temperature **(Fig.
+Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had a median value less than 1, indicating that most lakes were undersaturated in these populations.
+However, the long right tails on these distributions, comprised of relatively rare lakes with very high levels of N~2~O supersaturation, were sufficient to drive the population means to values close to, or even greater than 1.
+Thus, most lakes in the CONUS were functioning as N~2~O sinks during the summer months, but a few functioned as strong N~2~O sources, particularly at high NO~3~ and high water temperature **(Fig.
 3)**.
@@
 ![Figure 2.
-Predicted distribution of N~2~O saturation values across waterbodies in the nine major ecoregions.
+Predicted distribution of N~2~O saturation values across lakes in the nine major ecoregions.
 Each blue line (n=2000) represents the distribution of values from a single realization of the posterior predictive distribution.
@@
 A) The relationship between N~2~O saturation and surface water temperature, dependent upon nitrate (NO~3~) level.
-The five, ordered, categorical levels for NO~3~ correspond to concentration ($\mu$g L^-1^) ranges in the observed data of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13.
+The five, ordered, categorical levels for NO~3~ correspond to concentration ($\mu$g N L^-1^) ranges in the observed data of 1) below detection, 2) $\geq$ 0.55 and \< 4.09, 3) $\geq$ 4.09 and \< 30.20, 4) $\geq$ 30.20 and \< 223.13, and 5) $\geq$ 223.13.
 B) The relationship between N~2~O saturation and NO~3~ level conditional on surface water temperature.
@@
 
-Nitrous oxide saturation ratio was also related to waterbody size **(Fig.
+Lake size was also related to N~2~O saturation ratio, but the relationship was dependent on NO~3~ concentration **(Fig.
 3)**.
-Small waterbodies with low NO~3~ levels were predicted to be undersaturated, whereas small waterbodies with high NO~3~ were predicted to be highly supersaturated.
-Large waterbodies, regardless of NO~3~ status, were predicted to have N~2~O saturation ratios closer to equilibrium.
-When pooled across all ecoregions, the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O, decreased with increasing waterbody size class **(Fig.
+Small lakes with low NO~3~ levels were predicted to be undersaturated, whereas small lakes with high NO~3~ were predicted to be highly supersaturated.
+Large lakes, regardless of NO~3~ status, were predicted to have N~2~O saturation ratios closer to equilibrium.
+.
+Thus N~2~O saturation ratio tended to deviate from a value of one more strongly in small lakes as compared to larger lakes.
+This pattern can also be seen by calculating the absolute magnitude of the difference between the measured and equilibrium N~2~O concentration, known as $\Delta$ N~2~O.
+When pooled across all ecoregions, $\Delta$ N~2~O is predicted to decline as lake size increases **(Fig.
 4)**.
 This pattern holds regardless of whether N~2~O was supersaturated or undersaturated.
-Waterbody size and shape can affect dissolved gas concentration through several mechanisms.
-In waterbodies with low water volume to benthic surface area ratios, the mass of N~2~O consumed or produced in the sediments can be a substantial fraction of the mass of N~2~O dissolved in the water column [@Carpenter_1983].
-Furthermore, sediment may warm faster in small and shallow waterbodies, whereas thermal stratification may substantially delay sediment warming in large and deep waterbodies, resulting in greater rates of N~2~O production or consumption in small lakes.
+Lake size and shape can affect dissolved gas concentration through several mechanisms.
+For example, N~2~O inputs from streams or groundwater[@Wang_etal_2023] may constitute a substantial N~2~O subsidy to small lakes but are more likely be diluted in larger lakes with greater volumes.
+Small lakes also tend to have lower water volume to benthic surface area ratios,and therefore stronger connections between sediment, an area of active N~2~O processing, and the air-water interface where our samples were collected.
+Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep waterbodies, resulting in greater rates of N~2~O production or consumption in small lakes.
 Thermal stratification may also prevent the exchange of water between the lake surface and the hypolimnion, thereby preventing N~2~O depleted or enriched water from advecting to the lake surface [@Beaulieu_etal_2015; @Beaulieu_etal_2014].
 
-Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequilibrium in smaller waterbodies.
-Air-water gas exchange will vent excess N~2~O from surface waters or facilitate the invasion of air that can replenish depleted N~2~O in undersaturated waterbodies.
-Regardless of whether a waterbody is supersaturated or undersaturated in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
+Persistent patterns in the air-water gas exchange rate across lake size classes may also contribute to greater N~2~O disequilibrium in smaller lakes.
+Air-water gas exchange will vent excess N~2~O from surface waters or facilitate the invasion of air that can replenish depleted N~2~O in undersaturated lakes.
+Regardless of whether a lake is supersaturated or undersaturated in N~2~O, air-water gas exchange will function to move the dissolved N~2~O concentration toward equilibrium.
 One of the primary drivers of air-water gas exchange in lentic waters is wind speed [@Vachon_etal_2013].
-Wind speeds tend to be greater in large lakes with long fetches and lower in small waterbodies, particularly if they are sheltered by shoreline vegetation.
-Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the effect of biogeochemistry on dissolved N~2~O may be more pronounced in small waterbodies because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
+Wind speeds tend to be greater in large lakes with long fetches and lower in small lakes, particularly if they are sheltered by shoreline vegetation.
+Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the effect of biogeochemistry on dissolved N~2~O may be more pronounced in small lakes because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
 While previous surveys have documented relationships between lake size and dissolved methane concentration [@Pighini_etal_2018; @Juutinen_etal_2009; @Borges_etal_2011], to the best of our knowledge this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
@@
 
-The predicted emission rates for CONUS waterbodies spanned a broad range (95% CI min: `r em_min_max$min_LCL` - `r em_min_max$min_UCL` mg N~2~O m^-2^ d^-1^, 95% CI max: `r em_min_max$max_LCL` - `r em_min_max$max_UCL` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of waterbodies, but within the range of values reported in the literature for lentic ecosystems [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
-Variability and magnitude of predicted N~2~O emission rates decreased with increasing waterbody size **(Fig.
-5A)**, which reflects the relationship between N~2~O saturation and waterbody size **(Fig.
+The predicted emission rates for the POI spanned a broad range (95% CI min: `r em_min_max$min_LCL` - `r em_min_max$min_UCL` mg N~2~O m^-2^ d^-1^, 95% CI max: `r em_min_max$max_LCL` - `r em_min_max$max_UCL` mg N~2~O m^-2^ d^-1^), which was expected for this large and diverse collection of lakes, but is within the range of values reported in the literature [@Yang_etal_2015; @Kortelainen_etal_2020; @Webb_etal_2019].
+Variability and magnitude of predicted N~2~O emission rates decreased with increasing lake size **(Fig.
+5A)**, which reflects the relationship between N~2~O saturation and lake size **(Fig.
 3C,D)**.
-Despite widespread N~2~O undersaturation predicted across CONUS waterbodies, the estimate for the national average N~2~O emission rate was centered over positive values (posterior mean = `r national_stats$post_m_mean_en2o`, 95% CI: `r national_stats$LCL_mean_en2o` - `r national_stats$UCL_mean_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)** and clearly exceeded the negative estimate for the national median (posterior mean = `r national_stats$post_m_median_en2o`, 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^).
-This departure from the median was due to relatively rare waterbodies with very high emission rates pulling the mean towards the direction of positive skew **(SI Fig.
+Despite widespread N~2~O undersaturation predicted across the POI, the estimate for the national average N~2~O emission rate was centered over positive values (posterior mean = `r national_stats$post_m_mean_en2o`, 95% CI: `r national_stats$LCL_mean_en2o` - `r national_stats$UCL_mean_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)** and clearly exceeded the negative estimate for the national median (posterior mean = `r national_stats$post_m_median_en2o`, 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^).
+This departure from the median was due to relatively rare lakes with very high emission rates pulling the mean towards the direction of positive skew **(SI Fig.
 1)**.
 
-![Figure 5: A) Predicted N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 waterbodies in the target population.
+![Figure 5: A) Predicted N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 lakes in the population of interest.
 Points and bars reflect the posterior means (points) and 95% credible intervals, respectively, based on 2000 draws from the posterior predictive distribution for each waterbody.
-C) A comparison of the estimated relative flux contributions from waterbodies in each size class to the total CONUS flux.
+C) A comparison of the estimated relative flux contributions from lakes in each size class to the total CONUS flux.
 The red line connects the mean of the posterior estimate of the ratio, at each size class, based on 2000 draws.
-The black polygon indicates the 95% credible intervals for the ratio estimates across the size classes](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
+The black polygon indicates the 95% credible intervals for the ratio estimates across the size classes.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
 
@@
 
-Total N~2~O flux from the surface of CONUS waterbodies is a function of N~2~O emission rates and waterbody surface areas.
-The four smallest size classes (\< 50 Ha) in the survey design represented `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(number = 1:nrow(.), proportion = number / nrow(.)) %>% filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% pull(proportion) %>% round(2) * 100`% of all lakes in the population but were estimated to have contributed less, on average, to the total N~2~O flux from CONUS waterbodies **(Fig 5C)**.
+Total N~2~O flux from the surface of a lake is calculated as the product of the N~2~O emission rate and lake surface area.
+The four smaller size classes (\< 50 Ha) in the survey design represented `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(number = 1:nrow(.), proportion = number / nrow(.)) %>% filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% pull(proportion) %>% round(2) * 100`% of all lakes in the POI but were estimated to have contributed less, on average, to the total N~2~O flux from CONUS lakes than the largest lake size class **(Fig 5C)**.
 The reason for this pattern is two-fold.
 First, lakes in the smaller size classes constituted only `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(cumulative = cumsum(area_ha), distribution = cumulative / sum(area_ha), delta50 = abs(50-area_ha)) %>% filter(delta50 == min(delta50)) %>% pull(distribution) %>% round(3) * 100`% of total waterbody surface area, thereby limiting their N~2~O flux.
-Second, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the population.
-Waterbodies in the largest size class (\> 50 Ha) represented only `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(number = 1:nrow(.), proportion = number / nrow(.)) %>% filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% pull(proportion) %>% round(2) * 100 * -1 + 100`% of the total number of lakes in the population but had `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(cumulative = cumsum(area_ha), distribution = cumulative / sum(area_ha), delta50 = abs(50-area_ha)) %>% filter(delta50 == min(delta50)) %>% pull(distribution) %>% round(3) * 100 * -1 +100`% of total lake surface area.
-This lake size class was estimated to be, on average, the largest contributor to the total N~2~O flux across CONUS **(Fig 5C)**, though the flux estimate is poorly constrained (posterior mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(post_m_flux) %>% pull() %>% round(1)`, 95%CI: `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull() %>% round(1)` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull() %>% round(1)` metric tons N~2~O year^-1^).
+Second, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the POI.
+Lakes in the largest size class (\> 50 Ha) represented only `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(number = 1:nrow(.), proportion = number / nrow(.)) %>% filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% pull(proportion) %>% round(2) * 100 * -1 + 100`% of the total number of lakes in the population but had `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(cumulative = cumsum(area_ha), distribution = cumulative / sum(area_ha), delta50 = abs(50-area_ha)) %>% filter(delta50 == min(delta50)) %>% pull(distribution) %>% round(3) * 100 * -1 +100`% of total lake surface area.
+This lake size class was estimated to be, on average, the largest contributor to the total N~2~O flux across the POI **(Fig 5C)**, though the flux estimate is poorly constrained (posterior mean = `r FluxBySize %>% filter(size_cat == "50_max") %>% select(post_m_flux) %>% pull() %>% round(1)`, 95%CI: `r FluxBySize %>% filter(size_cat == "50_max") %>% select(LCL) %>% pull() %>% round(1)` - `r FluxBySize %>% filter(size_cat == "50_max") %>% select(UCL) %>% pull() %>% round(1)` metric tons N~2~O year^-1^).
 While uncertainty in N~2~O emission rates was smallest for large lakes **(Fig.
-5A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large waterbodies **(Fig.
+5A)**, these small uncertainties propagated to large uncertainties in flux when expressed across the surface area of large lakes **(Fig.
 5B)**.
 
-Despite the estimated widespread N~2~O undersaturation among CONUS waterbodies, the credible intervals for ecoregion-level flux estimates included zero for all but three ecoregions **(Table 1)**.
+Despite the estimated widespread N~2~O undersaturation among CONUS lakes, the credible intervals for ecoregion-level flux estimates included zero for all but three ecoregions **(Table 1)**.
 Thus, status as a net source versus sink of N~2~O was uncertain for most ecoregions.
@@
 
-The estimate for flux across all CONUS waterbodies was centered over positive values (posterior mean = `r national_stats %>% select(post_m_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval suggested a good deal of uncertainty and included zero (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
-Again, we could not confidently infer the extent to which CONUS waterbodies, overall, were a net source or sink of N~2~O.
-Though we are unaware of independent estimates of N~2~O emissions from CONUS lakes and reservoirs, the Intergovernmental Panel on Climate Change (IPCC) estimated that CONUS drainage networks, inclusive of zero order streams through major river mouths, emit `r format(ipcc.n2o$surface, scientific = FALSE, big.mark = ",")` metric tons N~2~O year^-1^ [@epa_2023].
-The IPCC estimates total N~2~O emissions from aquatic ecosystems as a fraction of the anthropogenic nitrogen that leaches from soils into surface water systems, but does not distinguish standing from flowing waters, thus a direct comparison between the IPCC estimate and the value reported here is not possible.
-However, the upper 95% credible interval of the N~2~O flux estimate for lakes and reservoirs reported here comprises only `r round((national_stats %>% select(UCL_fn2o) %>% pull() / ipcc.n2o$surface) * 100, digits = 1)` % of the IPCC estimate, suggesting the IPCC estimate is too high, possibly because the IPCC model does not allow for N~2~O consumption in lakes, ponds, and reservoirs.
-This limitation extends to many mechanistic models used to estimate N~2~O fluxes from inland waters [@Lauerwald_etal_2019; @Marzadri_etal_2021; @Marzadri_etal_2017].
-Given our estimate of widespread N~2~O consumption in CONUS waterbodies, failure to include this dynamic in aquatic N~2~O models may yield low-biased emission estimates [@Lauerwald_etal_2023].
-Recent models allowing for N~2~O consumption in inland waters are promising developments in this regard [@Yao_etal_2019; @Wang_etal_2023; @Maavara_etal_2019].
+The estimate for flux across the POI was centered over positive values (posterior mean = `r national_stats %>% select(post_m_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval suggested a good deal of uncertainty and included zero (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
+Again, we could not confidently infer the extent to which CONUS lakes, overall, were a net source or sink of N~2~O.
+Lauerwald et al [@Lauerwald_etal_2019] used a mechanistic model to estimate that CONUS lakes emit 8,425 metric tons N~2~O year^-1^.
+That estimate is nearly 25-fold greater than the posterior mean of our estimate (344 metric tons N~2~O year^-1^), despite our estimate including an order of magnitude more lakes than Lauerwald et al.[@Lauerwald_etal_2019]  A possible explanation for the large difference between these estimates is that the Lauerwald model does not allow for lakes to behave as N~2~O sinks, a This limitation that extends to many mechanistic models used to estimate N~2~O fluxes from inland waters[@Marzadri_etal_2021;@Marzadri_etal_2017;@Mosier_etal_1998]].
+However, even models that allow for net N~2~O uptake can overpredict N~2~O flux.
+For example, Wang et al.[@Wang_etal_2023] used a model that allows for N~2~O consumption via denitrification to predict that CONUS lakes emitted an average of 47,614 metric tons N~2~O year^-1^ from 2011-2015.
+This estimate is two orders of magnitude greater than the posterior mean of our estimate and suggests the mechanistic does not properly describe N~2~O consumption.
+Given our estimate of widespread N~2~O consumption in CONUS lakes, failure to properly describe this dynamic in aquatic N~2~O models may yield high-biased emission estimates.
 
-Not only is our estimate of N~2~O emissions from CONUS lakes and reservoirs small compared to the IPCC estimate, it is also small relative to other anthropogenic N~2~O sources, including wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
-Whether CONUS waterbodies were a net source or sink of N~2~O, their contribution to the US N~2~O budget was likely small by comparison.
+Not only is our estimate of N~2~O emissions from CONUS lakes small compared to other published estimates [@Wang_etal_2023;@Lauerwald_etal_2019], it is also small relative to other N~2~O sources including CONUS streams and rivers[@Yao_etal_2019] (55,490 metric tons N~2~O year^-1^), wastewater treatment (88,00 metric tons N~2~O year^-1^), stationary combustion (84,000 metric tons N~2~O year^-1^), and manure management (66,000 metric tons N~2~O year^-1^) [@epa_2023].
+Whether CONUS lakes were a net source or sink of N~2~O, their contribution to the US N~2~O budget was likely small by comparison.
 
-Our annual flux estimates are derived from a model of single samples collected from individual waterbodies during the summer.
+Our annual flux estimates are derived from a model of single samples collected from individual lakes during the summer.
 Our approach therefore implicitly assumes that our summertime observations are reflective of annual conditions.
@@
 These modeling results suggested that NO~3~ levels may play a role in determining whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink may be exaggerated with increasing water temperature.
-The degree of N~2~O disequilibrium also increased with decreasing waterbody size, possibly because smaller waterbodies have low air-water gas exchange rates.
+These results suggest that the convergence of the ongoing warming of CONUS lakes [@Smucker_etal_2021;@OReilly_etal_2015;@Winslow_etal_2017] and projected increases in freshwater nitrogen loading [@Sinha_etal_2017] may increase lentic N~2~O emissions over the next century[@Wang_etal_2023].
 
-Although waterbodies in the largest lake size class tend to have N~2~O levels close to equilibrium, they emit more N~2~O than the other size classes due to their expansive surface area.
-The N~2~O flux from these waterbodies is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved and equilibrium N~2~O concentrations translate to large uncertainties in flux when propagated across the surface of large waterbodies.
-Uncertainty in the national flux estimate was too large to determine if CONUS waterbodies functioned as a net source or sink of N~2~O.
-While our sample included some of the largest waterbodies in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O, which can be important in large waterbodies[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement an analytical uncertainty, which can propagate to uncertainties in flux estimates.
+Although lakes in the largest lake size class tend to have N~2~O levels close to equilibrium, they emit more N~2~O than the other size classes due to their expansive surface area.
+The N~2~O flux from these lakes is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved and equilibrium N~2~O concentrations translate to large uncertainties in flux when propagated across the surface of large lakes.
+Uncertainty in the national flux estimate was too large to determine if CONUS lakes functioned as a net source or sink of N~2~O.
+While our sample included some of the largest lakes in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O, which can be important in large waterbodies[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement and analytical uncertainty, which can propagate to uncertainties in flux estimates.
 
@@
 For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted [@Beaulieu_etal_2014].
-Most small waterbodies in the CONUS were estimated to have functioned as N~2~O sinks during the summer, but a small subset were also N~2~O emission hotspots.
-It is unknown if the high emitting waterbodies maintain those emissions throughout the summer, or if they convert to sinks later in the season.
-A better understanding of the temporal evolution of N~2~O production and consumption is an important research need to determine the extent to which small waterbodies can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
+Most small lakes in the CONUS were estimated to have functioned as N~2~O sinks during the summer, but a small subset were also N~2~O emission hotspots.
+It is unknown if the high emitting lakes maintain those emissions throughout the summer, or if they convert to sinks later in the season.
+A better understanding of the temporal evolution of N~2~O production and consumption is an important research need to determine the extent to which small lakes can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
 
@@
 
-The population of interest (POI) for the 2017 NLA survey [@USEPA_NLA17_2022] was CONUS waterbodies of at least 1 hectare in area, at least 1 meter deep, with at least 0.1 hectare of open water, and a minimum residence time of one week **(Fig.
+The population of interest (POI) for the 2017 NLA survey [@USEPA_NLA17_2022] was CONUS lakes of at least 1 hectare in area, at least 1 meter deep, with at least 0.1 hectare of open water, and a minimum residence time of one week **(Fig.
 1)**.
@@
 The size categories were sampled with unequal probability, such that larger lakes were over-sampled relative to smaller lakes.
-In total, 1091 waterbodies were sampled; and 95 of those waterbodies were re-visited for a second sample.
+In total, `r  dg %>% filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>% nrow()` lakes were sampled; and `r  dg %>% filter(sample.source == "DG", visit.no == 2, sitetype == "PROB") %>% nrow()` of those lakes were re-visited for a second sample.
 The revisit samples were used to assess out-of-sample model performance.
 
-## Sampling
+## Sampling and analysis
 
-Samples were collected from approximately the middle of each waterbody for analysis of algal toxins (microcystins and cylindrospermopsin), benthic macroinvertebrates, phytoplankton, atrazine, water chemistry, chlorophyll‐a, zooplankton, and dissolved N~2~O.
+Samples were collected from approximately the middle of each lake for analysis of NO~3~ and dissolved N~2~O, among other other analytes[@USEPA_NLA17_2022].
 Duplicate dissolved N~2~O samples were collected approximately 0.05 m below the water surface by drawing 105 mL of water into a 140 mL syringe equipped with a stopcock and containing 35 mL of air.
@@
 Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the population of interest [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
-In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (*e.g.,* US waterbodies) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
-In this study, because N~2~O flux was to be calculated as a function of waterbody-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (*e.g.*, ecoregions), predictions were made to each of the 465,897 individual waterbodies in the assumed POI (See SI) and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
+In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (*e.g.,* CONUS lakes) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+In this study, because N~2~O flux was to be calculated as a function of lake-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (*e.g.*, ecoregions), predictions were made to each of the `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` lakes in the assumed POI (See SI) and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
 The fully Bayesian implementation made propagation of uncertainty from the waterbody-level predictions to higher level summaries straightforward.
@@
 The saturation ratio was not modeled or predicted directly, but was estimated *post hoc* as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
-Similarly, the proportion of under-saturated waterbodies was estimated from this PPD as the proportion of waterbodies with a saturation ratio less than 1.
+Similarly, the proportion of under-saturated lakes was estimated from this PPD as the proportion of lakes with a saturation ratio less than 1.
 All models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models using Hamiltonian Monte Carlo (HMC).

### 2024-03-21 manuscript/manuscript_support.R

@@ -41,7 +41,6 @@ library(USAboundaries) # state boundaries
 library(tictoc) # processing time
 library(ggpubr) # multiple plots
 library(ggallin) # psuedlolog transformation for negative values
-library(USAboundaries) # state boundaries for maps
 library(tidybayes) # prop flux by size figure
 
 # DATA---------------
@@ -489,7 +488,7 @@ if(!("n2oStarBySize.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   
 }
 
-## Figure X:  Flux by lake, emission rate by lake, and nation flux.  VS continuous size-------
+## Figure X:  Flux by lake, emission rate by lake, and national flux VS continuous size-------
 if(!("n2oFluxAndEmissionRateVsContinuousArea.tiff" %in% list.files("manuscript/manuscript_figures"))) {
   
   # Predicted (mean, L95CI, U95CI of PPD) waterbody N2O emission rates vs waterbody size continuous 

### 2024-03-21 manuscript/manuscript_file.Rmd

@@
 
-The survey samples were treated statistically as a randomly selected subset of the complete POI, but with a pattern of "missingess" prescribed by the selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
+The survey samples were treated as a randomly selected subset of the POI, but with a pattern of "missingess" prescribed by the selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
 For inferences, estimates were statistically adjusted to account for the selection mechanism using a Multilevel regression with poststratification (MRP) approach [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020].
 MRP is a model-based approach carried out in two steps.
-In the first step, the sampled responses of interest (*e.g.*, N~2~O) are conditioned on the survey design variables (*e.g.*, ecoregion, state, size) in a multilevel regression model.
-Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the population of interest [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
-In the second step of MRP, the fitted regression parameters are "poststratified" or re-weighted based on a known or assumed distribution of units (*e.g.,* CONUS lakes) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
-In this study, because N~2~O flux was to be calculated as a function of lake-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$), instead of postratifying parameter estimates across groups of units (*e.g.*, ecoregions), predictions were made to each of the `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` lakes in the assumed POI (See SI) and then scaled up for group level inferences by summarizing the full predictive distributions across the groupings of interest.
+In the first step, the sampled response (*e.g.*, N~2~O) is conditioned on the survey design variables (*e.g.*, ecoregion, state, size) in a multilevel regression model.
+Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the POI [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
+In the second step of MRP, the fitted regression parameters are used to "poststratify" or re-weight the response based on a known or assumed distribution of units (*e.g.,* US lakes) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+Although it is typical to postratify estimates at a grouping level above the observational unit (*e.g.*, ecoregion), in this study, N~2~O flux needed to be calculated as a function of lake-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$).
+Therefore, predictions were made to each of the `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` lakes in the assumed POI.
+For group level inferences, the posterior predictive distribution of the response was summarized across all lakes in the groupings of interest.
 The fully Bayesian implementation made propagation of uncertainty from the waterbody-level predictions to higher level summaries straightforward.
 
-The final model was used to make probabilistic predictions for (1) dissolved and equilibrium N~2~O concentrations; (2) the N~2~O saturation ratio; and (3) the proportion of under-saturated water bodies (*i.e.*, saturation ratio \< 1).
+The multilevel regression model was used to make probabilistic predictions for (1) dissolved and equilibrium N~2~O concentrations; (2) the N~2~O saturation ratio; and (3) the proportion of under-saturated water bodies (*i.e.*, saturation ratio \< 1).
 Saturation ratio was calculated as the ratio of dissolved to equilibrium N~2~O.
-Because dissolved and equilibrium N~2~O were each measured on the same sample units (*i.e.*, a single waterbody), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, common dependencies on geography.
+Because dissolved and equilibrium N~2~O were each measured on the same sample units (*i.e.*, individual lakes), their joint distribution was modeled in order to account for potential statistical dependencies due to, for example, conditional dependencies on geography.
 The modeled response variable was, therefore, multivariate.
-Although the mean marginal probabilities estimated from separate models may have provided comparable estimates, the joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
-The saturation ratio was not modeled or predicted directly, but was estimated *post hoc* as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014; @McElreath_2020] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
-Similarly, the proportion of under-saturated lakes was estimated from this PPD as the proportion of lakes with a saturation ratio less than 1.
-All models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models using Hamiltonian Monte Carlo (HMC).
+Although the mean marginal probabilities estimated from separate models may have provided comparable estimates, the joint model allowing correlated responses was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021].
 
-Fully reproducible code and further results and model workflow details are provided in the Modeling Supplement Document.
-Though our initial models conditioned the response on just the survey design variables, those models consistently resulted in poorer fits to to the observed N~2~O data compared to models with additional covariates.
-Though the initial models seemed to work well at replicating the central tendency of the dissolved N~2~O data, simulations of other properties of the observed data, such as the variance, skewness, and kurtosis, were less satisfactory.
-The relatively poor fit to this aspect of the observed data also resulted in a downward bias in estimates of the observed saturation ratio.
-However, including a measure of NO~3~ concentration as a covariate on dissolved N~2~O resulted in substantially improved posterior predictive model checks.
-In addition, including surface temperature and elevation measures as covariates on equilibrium N~2~O resulted in substantially improved fits, as did adding flexibility in the distributional terms for both dissolved and equilibrium N~2~O.
-Both dissolved and equilibrium N~2~O were allowed to vary according to the survey design variables and environmental covariates in the final model presented here.
+The saturation ratio was not modeled or predicted directly, but was estimated *post hoc* as a "derived" quantity [@Link_Barker_2010; @Gelman_etal_2014] by dividing the posterior predictive distribution (PPD) for dissolved N~2~O by the PPD for equilibrium N~2~O.
+Similarly, the proportion of under-saturated lakes was calculated from this PPD as the proportion of lakes with a saturation ratio less than 1.
+All models were fit using the `brms` [@Burkner_2017] package in `R` [@R_Core_Team] as an interface to Stan [@Stan_Development_Team], a software for fitting fully Bayesian models using Hamiltonian Monte Carlo (HMC).Fully reproducible code, results, and model workflow details are provided in the Modeling Supplement Document.
 
-Including these helpful covariates, however, led to some complexities in the MRP process.
-To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled.
+Our initial models conditioned the response on just the survey design variables, however, those models consistently resulted in poor fits to some aspects of the observed N~2~O data.
+These models worked reasonably well at replicating the central tendency of the observed N~2~O data, but other properties of the observed distributions, such as the variance, skewness, and kurtosis, proved difficult to replicate.
+A consequence of the relatively poor fit to the tail properties of the data was a downward bias in estimates of the saturation ratio.
+In subsequent models, a measure of NO~3~ levels was included as a covariate on the dissolved N~2~O response, which substantially improved the fit according to the posterior predictive model checks.
+In addition, including surface temperature and elevation measures as covariates on equilibrium N~2~O resulted in improvements, as did allowing for heterogeneous errors for both dissolved and equilibrium N~2~O.
+Both dissolved and equilibrium N~2~O were allowed to vary according to the survey design variables and environmental covariates.
+
+Including covariates, however, necessitated additional steps in the MRP process.
+To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled as with the N2O responses.
 For the lake area and elevation covariates, data was available for all lakes in the POI from previously compiled geospatial databases.

### 2024-03-21 manuscript/manuscript_file.Rmd
more changes in response to tech reviews

@@
 Lakes, reservoirs, and ponds (lakes hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], a potent greenhouse gas and contributor to stratospheric ozone destruction.
-It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across lakes in Canada [@Soued_etal_2016; @Webb_etal_2019], the United States [Aho_et_al_2023], and Finland [@Kortelainen_etal_2020] have called this assumption into question.
+It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across lakes in Canada [@Soued_etal_2016; @Webb_etal_2019], the United States [@Aho_etal_2023], and Finland [@Kortelainen_etal_2020] have called this assumption into question.
 These recent studies have been conducted at relatively small spatial scales, or with a relatively small number of lakes, and the prevalence of N~2~O uptake in lakes at national to continental scales is unknown.
@@
 Samples sites were selected using a generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] and the observational data were used to develop a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS lakes.
-Combining the largest freshwater N~2~O dataset with a hierarchical model enabled us to estimate the prevalence of N~2~O uptake in CONUS lakesfor the first time and to identify important environmental controls on N~2~O biogeochemistry.
+Combining the largest freshwater N~2~O dataset with a hierarchical model enabled us to estimate the prevalence of N~2~O uptake in CONUS lakes for the first time and to identify important environmental controls on N~2~O biogeochemistry.
 
@@
 
-Nearly `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
+Nearly `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent of the sampled lakes were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
 1)**.
-The Bayesian hierarchical model predicted that the majority lakes in the POI (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% credible interval (CI): `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`) functioned as net N~2~O sinks during the summer months.
-The Western Mountains ecoregion had the smallest proportion of lakes predicted as N~2~O sinks (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(UCL_prop_sat) * 100`) and the Northern Plains had the greatest (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(UCL_prop_sat) * 100`), although the 95% credible intervals over the predictive distributions overlap across all ecoregions **(Table 1)**.
+The Bayesian hierarchical model predicted that the majority lakes in the POI (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% credible interval (CI): `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) functioned as net N~2~O sinks during the summer months.
+The Western Mountains ecoregion had the smallest proportion of lakes predicted as N~2~O sinks (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(UCL_prop_sat) * 100`%) and the Northern Plains had the greatest (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(UCL_prop_sat) * 100`%), although the 95% credible intervals of the predicted distributions overlap across all ecoregions **(Table 1)**.
 
@@
 2)**.
-Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had a median value less than 1, indicating that most lakes were undersaturated in these populations.
-However, the long right tails on these distributions, comprised of relatively rare lakes with very high levels of N~2~O supersaturation, were sufficient to drive the population means to values close to, or even greater than 1.
+Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had a median value less than one, indicating that most lakes were undersaturated in these populations.
+However, the long right tails on these distributions, comprised of relatively rare lakes with very high levels of N~2~O supersaturation, were sufficient to drive the population means to values close to, or even greater than one.
 Thus, most lakes in the CONUS were functioning as N~2~O sinks during the summer months, but a few functioned as strong N~2~O sources, particularly at high NO~3~ and high water temperature **(Fig.
@@
 Vertical black dashed and solid lines indicate point estimates of the median and mean, respectively, across all realizations (Table 1).
-Vertical red line indicates a saturation ratio of 1.
-x-axis is truncated at 3 to better show the area where 97% of the observations occurred.](manuscript_figures/figure2.tiff)
+Vertical red line indicates a saturation ratio of one.
+x-axis is truncated at a value of three to better show the area where 97% of the observations occurred.](manuscript_figures/figure2.tiff)
 
@@
 The influence of NO~3~ levels on N~2~O saturation **(Fig.
-3)** may reflect a relationship between NO~3~ levels and denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Kuypers_etal_2018; @Richardson_etal_2009].
+3)** may reflect a relationship between NO~3~ concentration and denitrification, a type of anaerobic metabolism that couples the oxidation of organic matter with the sequential reduction of NO~3~ to nitrite (NO~2~), nitric oxide (NO), N~2~O, and N~2~ [@Kuypers_etal_2018; @Richardson_etal_2009].
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
@@
 D) The relationship between N~2~O saturation and NO~3~ level conditional on (log) waterbody surface area (5 quantiles).
-Horizontal dashed lines in each panel reflect an N~2~O saturation value of 1.](manuscript_figures/cond_effect_no3.tif)
+Horizontal dashed lines in each panel reflect an N~2~O saturation value of one.](manuscript_figures/cond_effect_no3.tif)
 
@@
 Lauerwald et al [@Lauerwald_etal_2019] used a mechanistic model to estimate that CONUS lakes emit 8,425 metric tons N~2~O year^-1^.
-That estimate is nearly 25-fold greater than the posterior mean of our estimate (344 metric tons N~2~O year^-1^), despite our estimate including an order of magnitude more lakes than Lauerwald et al.[@Lauerwald_etal_2019]  A possible explanation for the large difference between these estimates is that the Lauerwald model does not allow for lakes to behave as N~2~O sinks, a This limitation that extends to many mechanistic models used to estimate N~2~O fluxes from inland waters[@Marzadri_etal_2021;@Marzadri_etal_2017;@Mosier_etal_1998]].
+That estimate is nearly 25-fold greater than the posterior mean of our estimate (344 metric tons N~2~O year^-1^), despite our estimate including an order of magnitude more lakes than Lauerwald et al.[@Lauerwald_etal_2019]  A possible explanation for the large difference between these estimates is that the Lauerwald model does not allow for lakes to behave as N~2~O sinks, a limitation that extends to many mechanistic models used to estimate N~2~O fluxes from inland waters[@Marzadri_etal_2021;@Marzadri_etal_2017;@Mosier_etal_1998]].
 However, even models that allow for net N~2~O uptake can overpredict N~2~O flux.
 For example, Wang et al.[@Wang_etal_2023] used a model that allows for N~2~O consumption via denitrification to predict that CONUS lakes emitted an average of 47,614 metric tons N~2~O year^-1^ from 2011-2015.
-This estimate is two orders of magnitude greater than the posterior mean of our estimate and suggests the mechanistic does not properly describe N~2~O consumption.
+This estimate is two orders of magnitude greater than the posterior mean of our estimate and suggests the mechanistic model does not properly describe N~2~O consumption.
 Given our estimate of widespread N~2~O consumption in CONUS lakes, failure to properly describe this dynamic in aquatic N~2~O models may yield high-biased emission estimates.
@@
 Multilevel models are often recommended for this adjustment because they naturally provide (1) regularized estimates along the design groupings; and (2) estimates for design levels that may be missing from the sample, but are part of the POI [@Gelman_Little_1997; @Gelman_etal_2020; @Kennedy_Gelman_2021].
-In the second step of MRP, the fitted regression parameters are used to "poststratify" or re-weight the response based on a known or assumed distribution of units (*e.g.,* US lakes) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
+In the second step of MRP, the fitted regression parameters are used to "poststratify" or re-weight the response based on a known or assumed distribution of units (*e.g.,* CONUS lakes) in the POI [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020; @Kennedy_Gelman_2021; @Zachmann_etal_2022].
 Although it is typical to postratify estimates at a grouping level above the observational unit (*e.g.*, ecoregion), in this study, N~2~O flux needed to be calculated as a function of lake-level properties (*i.e.*, $N_2O_{flux} \propto \text{surface area}$).

### 2024-03-21 manuscript/manuscript_support.R

@@ -319,7 +319,9 @@ save(national_stats, file = "manuscript/manuscript_files/national_stats.rda")
 if(!("figure1.png" %in% list.files("manuscript/manuscript_figures"))) {
   ggplot() +
     geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
-    geom_sf(data = dg.sf %>% filter(!is.na(n2o.src.snk)),
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk)), # grey on top of black
             aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
             show.legend = "point") +
     geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +

### 2024-03-28 scripts/DG_sensitivity_to_measurement_error.Rmd
Updated uncertainty calculations for dissolved N2O

@@
 output:
+  html_document:
+    toc: true
+    df_print: paged
+  pdf_document:
+    toc: true
   html_notebook:
     depth: 4
-    fig_caption: yes
-    number_sections: yes
-    toc: yes
-    toc_float: yes
-  pdf_document:
-    toc: yes
-#bibliography: RWM_Endnote_Library.bib
-link-citations: yes
+    fig_caption: true
+    number_sections: true
+    toc: true
+    toc_float: true
+link-citations: true
 ---
@@
 library(knitr)
-opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
-
-#set directories and R package library
-setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
-.libPaths("C:/Users/rmartin/Desktop/R/library")
+# opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
+# 
+# #set directories and R package library
+# setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
+# .libPaths("C:/Users/rmartin/Desktop/R/library")
 ```
@@
 library(gridExtra)
+library(readxl)
+library(janitor)
+library(sf)
 
@@
 
-$$C = 10^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}$$
+$$C = 1^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}$$
 
@@
 ```{r setvars_and_check}
+load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
+            "\\ORD NLA17 Dissolved Gas - Documents\\",
+            "inputData\\dg.2021-02-01.RData"))
+
+# summary of measured N2O ppm in equilibrated headspace gas
+dg %>%
+  group_by(sample.source) %>%
+  summarize(mean_n2O = mean(n2o.ppm.gc, na.rm = TRUE),
+          median_n2o = median(n2o.ppm.gc, na.rm = TRUE)) #mean = 0.36, median = 0.31
+
+#    mean_n2o  median_n2o
+#AIR	0.3093312	0.3104941		
+#DG	 0.3618681	0.3067609
+
 B <- 99 # barometric pressure (kPa)
-V_g <- 20 # volume of reference gas in headspace
-V_w <- 40 # volume of water below headspace
+V_g <- 35 # volume of reference gas in headspace
+V_w <- 105 # volume of water below headspace
 T_c <- 23 # temperature ( C )
-C_e <- 0.055 # gas mixture ratio in headspace
-C_a <- 0.091 # gas mixture ratio in reference
+C_e <- 0.307 # gas mixture ratio in headspace
+C_a <- 0.310 # gas mixture ratio in reference
 H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
 
-C <- 10e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+C <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 
@@
 ```
+# Define GC measurement error
+Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.
+
+```{r get_gc_error}
+# READ GC DATA----------------
+##########################################################################
+#       Code is commented out because collaborators may not have access  #
+#       to the file paths. The calculated standard deviation is hard     #
+#       coded below.
+##########################################################################
+
+# Below we read raw GC data into the environment, extract repeat analysis of
+# an analytical standard, then calculate the standard deviation of the replicate
+# measurements.
+
+# gc.1 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_06_29_STD_ECD_FID_UNK.xlsx", 
+#                    trim_ws = TRUE, skip=59) 
+# gc.2 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_07_25_UNK_STD_ECD_FID.xlsx", 
+#                    trim_ws = TRUE, skip=59)
+# gc.3 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_02_UNK_STD_ECD_FID.xlsx", 
+#                    trim_ws = TRUE, skip=50)
+# gc.4 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_04_UNK_STD_ECD_FID.xlsx", 
+#                    trim_ws = TRUE, skip=50)
+# gc.5 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_16_Dgas_UNK_STD_ECD_FID.xlsx", 
+#                    trim_ws = TRUE, skip=62)
+# gc.6 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_21_UNK_STD_ECD_FID.xlsx", 
+#                    trim_ws = TRUE, skip=62)
+# gc.7 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_31_UNK_STD_ECD_FID2.xlsx", 
+#                    trim_ws = TRUE, skip=65)
+# gc.8 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_14_UNK_STD_ECD_FID.xlsx", 
+#                    trim_ws = TRUE, skip=62)
+# gc.9 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_19_STD_UNK_ECD_FID_TCD.xlsx", 
+#                    trim_ws = TRUE, skip=102)
+# gc.10 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_28_UNK_STD_ECD_FID.xlsx", 
+#                    trim_ws = TRUE, skip=50)
+# gc.11 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_10_11_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.12 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_10_17_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.13 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_12_20_STD_UNK_ECD_FID.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.14 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_03_AIR_UNK_STD_ECD_FID.xlsx", 
+#                     trim_ws = TRUE, skip=48)
+# gc.15 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_09_FID_ECD_STD.xlsx", 
+#                     trim_ws = TRUE, skip=48)
+# gc.16 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_16_UNK_STD_ECD_FID.xlsx", 
+#                     trim_ws = TRUE, skip=48)
+# gc.17 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_22_UNK_STD_ECD_FID.xlsx", 
+#                     trim_ws = TRUE, skip=60)
+# gc.18 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_31_STD_UNK_ECD_FID_TCD.xlsx", 
+#                     trim_ws = TRUE, skip=102)
+# gc.19 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_05_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.20 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_09_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.21 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_12_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.22 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_16_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.23 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_20_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.24 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_04_25_UNK_STD_ECD_FID.xlsx", 
+#                     trim_ws = TRUE, skip=72)
+# gc.25 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_23_UNK_STD_ECD_FID_Dgas.xlsx", 
+#                     trim_ws = TRUE, skip=67)
+# gc.26 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_27_UNK_STD_ECD_FID_DP.xlsx", 
+#                     trim_ws = TRUE, skip=72)
+# gc.27 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_03_05_Trap_6ml.xlsx", 
+#                     trim_ws = TRUE, skip=104)
+# # Merge gas data
+# gc.all <- Reduce(function(...) merge(..., all=T),
+#                   list(gc.1, gc.2, gc.3, gc.4, gc.5, gc.6, gc.7, 
+#                        gc.8, gc.9, gc.10, gc.11, gc.12, gc.13, gc.14, gc.15, 
+#                        gc.16, gc.17, gc.18, gc.19, gc.20, gc.21, gc.22, gc.23, 
+#                        gc.24, gc.25, gc.26, gc.27)) %>%
+#   janitor::clean_names()
+# 
+# 
+# # Format gas data
+# gc.std <- gc.all %>% 
+#   select( sample, contains("n2O")) %>%
+#   filter(grepl("std", sample, ignore.case = TRUE)) %>% # only keep standards
+#   mutate(sample = tolower(sample), # simplify formatting
+#          # we have multiple injections of this 1.8 Prax Air standard
+#          sample = case_when(grepl("1.8 prax", sample, ignore.case = TRUE) ~ "1.8prax",
+#                             TRUE ~ sample))
+# 
+# # with some invenstigating we could pull out multiple standards used
+# # across and within runs to calculate standard deviation, but the 27
+# # replicate injections of the 1.8 prax standard is low hanging fruit
+# gc.prax <- gc.std %>%
+#   filter(sample == "1.8prax")
+# 
+# gc.prax %>%
+#   ggplot(aes(x = n2o_ppm)) +
+#   geom_histogram()
+# 
+# n2o_sd <- sd(gc.prax$n2o_ppm) # 0.007872511ppm
+n2o_sd <- 0.007872511
+
+# # begin digging for other standards to work with
+# gc.std %>%
+#   filter(grepl("3p", sample, ignore.case = TRUE)) %>%
+#   ggplot(aes(sample,n2o_ppm)) +
+#   geom_point()
+# 
+# # list items 17 and 18 contain "m3p chk std"
+# # see if these are same standards
+# map(list(gc.1, gc.2, gc.3, gc.4, gc.5, gc.6, gc.7, 
+#          gc.8, gc.9, gc.10, gc.11, gc.12, gc.13, gc.14, gc.15, 
+#          gc.16, gc.17, gc.18, gc.19, gc.20, gc.21, gc.22, gc.23, 
+#          gc.24, gc.25, gc.26, gc.27),
+#     function(x) {
+#       x %>% janitor::clean_names(.) %>% 
+#         filter(grepl("m3p chk std", sample, ignore.case = TRUE))
+#     })
+# 
+# table(gc.std$sample)
+
+```
 
@@
 ```{r simulate_me_Ce, message=F}
-C_me <- rep( NA, 1e4 )
+C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector
 
-C_e_me <- rnorm( 1e4, C_e, 0.001 ) #m.e. in measured gas in headspace
+# this simulates measurement error on GC.
+# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
+C_e_me <- rnorm( 1e4, C_e, n2o_sd ) # me in measured gas in headspace.
+                                   # add random error to measured concentration
 
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
 }
 
-qplot( x = C_e_me ) +  
+# combine into dataframe for ggplot. qplot, which acceptes vectors,
+# was deprecated in ggplot 3.4.0
+C_df <- tibble(C_me = C_me, C_e_me = C_e_me)
+
+ggplot(C_df, aes(x = C_e_me )) +  
   geom_histogram( binwidth = 0.0001 ) + 
   geom_vline( xintercept = C_e , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in water gas concentration") +
-  xlab( "concentration (?)")
+  ggtitle( "Simulated measurement error in equilibrated gas concentration") +
+  xlab( "concentration (nmol L-1)")
 
-qplot( x = C_me * 1e9 ) +  
+ggplot(C_df, aes(x = C_me * 1e9 )) +  
   geom_histogram( binwidth = 0.01 ) + 
   geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
   xlab( "concentration (nMol)")
 
-qplot( x = C_me * 1e9 - C ) +  
+ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
   geom_histogram( binwidth = 0.01 ) + 
   geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
   xlab( "absolute error (nMol)")
@@
 
-C_a_me <- rnorm( 1e4, C_a, 0.001 ) #m.e. in measured gas in air
+C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air
 
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
 
-qplot( x = C_a_me ) +  
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+Ca_df <- tibble(C_me = C_me, C_a_me = C_a_me)
+
+ggplot(Ca_df, aes(x = C_a_me )) +  
   geom_histogram( binwidth = 0.0001 ) + 
@@
 
-qplot( x = C_me * 1e9 ) +  
-  geom_histogram( binwidth = 0.01 ) + 
+ggplot(Ca_df, aes(x = C_me * 1e9 )) +  
+  geom_histogram( binwidth = 0.001 ) + 
   geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in air") +
   xlab( "concentration (nMol)")
 
-qplot( x = C_me * 1e9 - C ) +  
+ggplot(Ca_df, aes(x = C_me * 1e9 - C )) +  
   geom_histogram( binwidth = 0.01 ) + 
   geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in N2O concentration due to GC measurment error \nof N2O in air") +
   xlab( "absolute error (nMol)")
@@
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
 }
 
-qplot( x = C_me * 1e9 ) +  
+# combine into dataframe for ggplot. qplot, which acceptes vectors,
+# was deprecated in ggplot 3.4.0
+C_df <- tibble(C_me = C_me)
+
+ggplot(C_df, aes(x = C_me * 1e9 )) +  
   geom_histogram( binwidth = 0.01 ) + 
   geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof 
+           N2O in equilibrated headspace and air") +
   xlab( "concentration (nMol)")
 
-qplot( x = C_me * 1e9 - C ) +  
+ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
   geom_histogram( binwidth = 0.01 ) + 
   geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in equilibrated headspace and air") +
   xlab( "absolute error (nMol)")
@@
 
-T_c_me <- rnorm( 1e4, T_c, 0.1 ) #m.e. in measured water temp
+# How to get standard deviation of temperature measurement?
+# calibration certificate reports an uncertainty of +/-0.058C
+# take standard deviation of simulated measurements with specified uncertainty
+T_c_sd <- sd(runif(n = 1e4, min = T_c - 0.058, max = T_c + 0.058)) # calibration certificate reports an uncertainty of +/-0.058C
+T_c_me <- rnorm( 1e4, T_c, T_c_sd ) #m.e. in measured water temp
 
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
 
-qplot( x = T_c_me ) +  
+# combine into dataframe for ggplot. qplot, which acceptes vectors,
+# was deprecated in ggplot 3.4.0
+C_df <- tibble(C_me = C_me, T_c_me = T_c_me)
+
+ggplot(C_df, aes(x = T_c_me )) +  
   geom_histogram( binwidth = 0.01 ) + 
@@
 
-qplot( x = C_me * 1e9 ) +  
-  geom_histogram( binwidth = 0.01 ) + 
+ggplot(C_df, aes( x = C_me * 1e9 )) +  
+  geom_histogram( binwidth = 0.001 ) + 
   geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
   xlab( "concentration (nMol)")
 
-qplot( x = C_me * 1e9 - C ) +  
-  geom_histogram( binwidth = 0.01 ) + 
+ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
+  geom_histogram( binwidth = 0.001 ) + 
   geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
   xlab( "absolute error (nMol)")
@@
 
-B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured water temp
+# How to get standard deviation of barometric pressure?
+# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa)
+# take standard deviation of simulated measurements with specified uncertainty
+B_sd <- sd(runif(n = 1e4, min = B - 0.4, max = B + 0.4)) # calibration certificate reports an uncertainty of +/-0.058C
+B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured barometric pressure
 
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 10e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
 
-qplot( x = B_me ) +  
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_df <- tibble(C_me = C_me, B_me = B_me, B_ru = runif(n = 1e4, min = B - 0.4, max = B + 0.4))
+
+ggplot(C_df, aes( x = B_me )) +  
   geom_histogram( binwidth = 0.01 ) + 
@@
 
-qplot( x = C_me * 1e9 ) +  
+ggplot(C_df, aes( x = B_ru )) +  
   geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = B , color = 'red' ) + 
+  ggtitle( "Simulated measurement error in barometric pressure") +
+  xlab( "barometric pressure (kPa)")
+
+ggplot(C_df, aes( x = C_me * 1e9 )) +  
+  geom_histogram( binwidth = 0.001 ) + 
   geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
   xlab( "concentration (nMol)")
 
-qplot( x = C_me * 1e9 - C ) +  
-  geom_histogram( binwidth = 0.01 ) + 
+ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
+  geom_histogram( binwidth = 0.001 ) + 
   geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
   xlab( "absolute error (nMol)")
@@
 V_w_me <- rnorm( 1e4, V_w, 1 ) #m.e. in measured water volume
-V_g_me <- 60 - V_w_me
+V_g_me <- 140 - V_w_me
 
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 10e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
 
-qplot( x = V_w_me ) +  
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_df <- tibble(C_me = C_me, V_w_me = V_w_me)
+
+ggplot(C_df, aes( x = V_w_me )) +  
   geom_histogram( binwidth = 0.01 ) + 
@@
 
-qplot( x = C_me * 1e9 ) +  
-  geom_histogram( binwidth = 0.01 ) + 
+ggplot(C_df, aes( x = C_me * 1e9 )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
   geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissollved N2O concentration due to water volume \nmeasurment error") +
   xlab( "concentration (nMol)")
 
-qplot( x = C_me * 1e9 - C ) +  
-  geom_histogram( binwidth = 0.01 ) + 
+ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
   geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in N2O concentration due to measurment error") +
   xlab( "absolute error (nMol)")
@@
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 10e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
 
-qplot( x = C_me * 1e9 ) +  
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_df <- tibble(C_me = C_me)
+
+ggplot(C_df, aes( x = C_me * 1e9 )) +  
   geom_histogram( binwidth = 0.01 ) + 
   geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
   xlab( "concentration (nMol)")
 
-qplot( x = C_me * 1e9 - C ) +  
+ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
   geom_histogram( binwidth = 0.01 ) + 
   geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
+  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
   xlab( "absolute error (nMol)")
+
+# standard deviation of a measurement is:
+sd(C_me* 1e9) # 0.15 nmol L-1, as compared to 1.25 nmol L-1 in Webb et al.
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me* 1e9) * 2 # +/- 0.302 nmol N2O
 ```
+
+# Incorporate measurement uncertainty into source/sink status
+The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.
+The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 2`nM.
+Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 2`nM of the equilibrium value cannot be determined as a source or a sink.
+
+```{r map_table, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+dg <- dg %>%
+  mutate(n2o.src.snk.with.error = case_when(
+    abs(dissolved.n2o.nmol - sat.n2o.nmol) > sd(C_me* 1e9) * 2 ~ n2o.src.snk,
+    TRUE ~ "undetermined"))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.error),
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.1, 10), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```

### 2024-04-05 scripts/DG_sensitivity_to_measurement_error.Rmd
adding uncertainty in denominator

@@
 ```
-# Define GC measurement error
+## Define GC measurement error
 Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.
@@
 
-# Simulate measurement error for gas measured in water
+## Simulate measurement error in GC measurement of N2O in equilibrated headspace
 ```{r simulate_me_Ce, message=F}
@@
 
-# Simulate measurement error for gas measured in air
+## Simulate measurement error for GC measurement of N2O in air
 ```{r simulate_me_Ca, message=F}
@@
 
-# Simulate measurement error in $C_e$ and $C_a$
+## Simulate measurement error in $C_e$ and $C_a$
 ```{r simulate_me_Ca_Ce, message=F}
@@
 
-# Simulate measurement error in water temperature
+## Simulate measurement error in water temperature
 ```{r simulate_me_T_c, message=F}
@@
 
-# Simulate measurement error in barometric pressure
+## Simulate measurement error in barometric pressure
 ```{r simulate_me_B, message=F}
@@
 
-# Simulate measurement error in water volume
+## Simulate measurement error in water volume
 ```{r simulate_me_V_w, message=F}
@@
 
-# Simulate measurement error for all observables
+## Simulate measurement error for all observables
 ```{r simulate_me_all, message=F}
@@
 # the 95% CI is therefore the computed value +/- 2 standard deviations
-sd(C_me* 1e9) * 2 # +/- 0.302 nmol N2O
+sd(C_me* 1e9) * 1.96 # +/- 0.302 nmol N2O
 ```
 
-# Incorporate measurement uncertainty into source/sink status
+## Incorporate measurement uncertainty into source/sink status
 The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.
-The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 2`nM.
-Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 2`nM of the equilibrium value cannot be determined as a source or a sink.
+The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 1.96`nM.
+Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 1.96`nM of the equilibrium value cannot be determined as a source or a sink.
 
@@
   mutate(n2o.src.snk.with.error = case_when(
-    abs(dissolved.n2o.nmol - sat.n2o.nmol) > sd(C_me* 1e9) * 2 ~ n2o.src.snk,
+    abs(dissolved.n2o.nmol - sat.n2o.nmol) > sd(C_me* 1e9) * 1.96 ~ n2o.src.snk,
     TRUE ~ "undetermined"))
@@
 ```
+# Set up model for simulating true saturation concentration
+
+The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water that is in equilibrium with the atmosphere is:
+
+$$C = 1^{-6}B*C_{air}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15})$$
+```{r c_sat}
+C_sat <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+
+C_sat <- C_sat * 1e9 #convert mol to nmol
+
+```
+
+## Simulate measurement error in GC measurement of N2O in air
+```{r simulate_me_sat_C_a, message=F}
+C_me_sat <- rep( NA, 1e4 )
+
+C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air
+
+for( i in 1:length( C_me_sat ) ){
+    C_me_sat[i] <- C_a_me[i] * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_me = C_a_me)
+
+ggplot(C_sat_df, aes(x = C_a_me )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
+  geom_vline( xintercept = C_a , color = 'red' ) + 
+  ggtitle( "Simulated measurement error in air gas concentration") +
+  xlab( "concentration (ppm)")
+
+ggplot(C_sat_df, aes(x = C_me_sat * 1e9 )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = C_sat , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
+  xlab( "concentration (nMol)")
+
+ggplot(C_sat_df, aes(x = C_me_sat * 1e9 - C_sat )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
+  xlab( "absolute error (nMol)")
+```
+
+## Simulate measurement error in water temperature
+```{r simulate_me_sat_T_c, message=F}
+C_me_sat <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_sat ) ){
+    C_me_sat[i] <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
+}
+
+# combine into dataframe for ggplot. qplot, which acceptes vectors,
+# was deprecated in ggplot 3.4.0
+C_sat_df <- tibble(C_me_sat = C_me_sat, T_c_me = T_c_me)
+
+ggplot(C_sat_df, aes(x = T_c_me )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = T_c , color = 'red' ) + 
+  ggtitle( "Simulated measurement error in temperature") +
+  xlab( "temperature (C)")
+
+ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline(xintercept = C_sat , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error") +
+  xlab( "concentration (nMol)")
+
+ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error" ) +
+  xlab( "absolute error (nMol)")
+```
+
+## Simulate measurement error in barometric pressure
+```{r simulate_me_sat_B, message=F}
+C_me_sat <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_sat ) ){
+    C_me_sat[i] <- C_a * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_sat_df <- tibble(C_me_sat = C_me_sat, B_me = B_me)
+
+ggplot(C_sat_df, aes( x = B_me )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = B , color = 'red' ) + 
+  ggtitle( "Simulated measurement error in barometric pressure") +
+  xlab( "barometric pressure (kPa)")
+
+ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline( xintercept = C_sat , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
+  xlab( "concentration (nMol)")
+
+ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
+  xlab( "absolute error (nMol)")
+```
+
+## Simulate measurement error for all observables
+```{r simulate_me_sat_all, message=F}
+C_me_sat <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_sat ) ){
+    C_me_sat[i] <- C_a_me[ i ] * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_sat_df <- tibble(C_me_sat = C_me_sat)
+
+ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = C_sat , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
+  xlab( "concentration (nMol)")
+
+ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
+  xlab( "absolute error (nMol)")
+
+# standard deviation of a measurement is:
+sd(C_me_sat* 1e9) # 0.20 nmol L-1
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me_sat * 1e9) * 1.96 # +/- 0.395 nmol N2O
+```

### 2024-04-05 scripts/DG_sensitivity_to_measurement_error.Rmd
added simulation of saturation ratio uncertatinty

@@
 
-$$C = 1^{-6}B*C_{air}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15})$$
+$$C_{sat} = 1^{-6}B*C_{air}*(H^{\theta}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15}))$$
 ```{r c_sat}
@@
 ```
+
+# Set up model for simulating true saturation ratio
+
+The full equation for estimating the saturation ratio of $N_20$ gas dissolved in water is:
+
+$$C_ratio = \frac{1^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}}{1^{-6}B*C_{air}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15})}$$
+
+``` {r}
+
+C_ratio <- (1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
+  (C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))
+
+```
+## Simulate measurement error for all observables
+```{r simulate_me_ratio_all, message=F}
+C_me_ratio <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_ratio ) ){
+C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - C_a_me[i] ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
+  (C_a_me[i] * B_me[i] * 1e-6 * (H  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_ratio_df <- tibble(C_me_ratio = C_me_ratio)
+
+ggplot(C_ratio_df, aes( x = C_me_ratio )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = C_ratio , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors") +
+  xlab( "saturation ratio")
+
+ggplot(C_ratio_df, aes( x = C_me_ratio - C_ratio )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors") +
+  xlab( "absolute error (saturation ratio)")
+
+# standard deviation of a saturation ratio value s:
+sd(C_me_ratio) # 0.0.05479
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me_ratio) * 1.96 # +/- 0.1073
+
+# so any ratio within 0.1073 of 1.0 cannot be discerned as a source or sink?
+```

### 2024-04-22 scripts/DG_sensitivity_to_measurement_error.Rmd
Added more precise methods to simulations

@@
 
+```{=html}
 <style>
@@
 </style>
+```
+# Background
 
-# Background
 In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).
@@
 # Setup R
+
 ```{r setup_1, include=FALSE, echo=FALSE}
@@
 
-# Set up model for simulating true value
+# Set up model for simulating true dissolved N2O value
 
@@
 C_e <- 0.307 # gas mixture ratio in headspace
-C_a <- 0.310 # gas mixture ratio in reference
-H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+C_a_n2o <- 0.310 # gas mixture ratio in reference
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant
 
-C <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+C <- 1e-6 * B * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 
@@
 ```
+
 ## Define GC measurement error
+
 Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.
@@
 # n2o_sd <- sd(gc.prax$n2o_ppm) # 0.007872511ppm
-n2o_sd <- 0.007872511
+n2o_sd_epa <- 0.007872511 # from AWBERC GC
 
@@
 ## Simulate measurement error in GC measurement of N2O in equilibrated headspace
+
 ```{r simulate_me_Ce, message=F}
@@
 # assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
-C_e_me <- rnorm( 1e4, C_e, n2o_sd ) # me in measured gas in headspace.
+C_e_me <- rnorm( 1e4, C_e, n2o_sd_epa ) # me in measured gas in headspace.
                                    # add random error to measured concentration
@@
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
 }
@@
 ## Simulate measurement error for GC measurement of N2O in air
+
 ```{r simulate_me_Ca, message=F}
@@
 
-C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air
+C_a_n2o_me <- rnorm( 1e4, C_a_n2o, n2o_sd_epa ) #m.e. in measured gas in air
 
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_n2o_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
@@
 # was deprecated in ggplot 3.4.0
-Ca_df <- tibble(C_me = C_me, C_a_me = C_a_me)
+Ca_df <- tibble(C_me = C_me, C_a_n2o_me = C_a_n2o_me)
 
-ggplot(Ca_df, aes(x = C_a_me )) +  
+ggplot(Ca_df, aes(x = C_a_n2o_me )) +  
   geom_histogram( binwidth = 0.0001 ) + 
-  geom_vline( xintercept = C_a , color = 'red' ) + 
+  geom_vline( xintercept = C_a_n2o , color = 'red' ) + 
   ggtitle( "Simulated measurement error in air gas concentration") +
@@
 
-## Simulate measurement error in $C_e$ and $C_a$
+## Simulate measurement error in $C_e$ and $C_a_n2o$
+
 ```{r simulate_me_Ca_Ce, message=F}
@@
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_n2o_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
 }
@@
 ## Simulate measurement error in water temperature
+
 ```{r simulate_me_T_c, message=F}
@@
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
 
-# combine into dataframe for ggplot. qplot, which acceptes vectors,
+# combine into dataframe for ggplot. qplot, which accepts vectors,
 # was deprecated in ggplot 3.4.0
@@
 ## Simulate measurement error in barometric pressure
+
 ```{r simulate_me_B, message=F}
@@
 # How to get standard deviation of barometric pressure?
-# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa)
-# take standard deviation of simulated measurements with specified uncertainty
-B_sd <- sd(runif(n = 1e4, min = B - 0.4, max = B + 0.4)) # calibration certificate reports an uncertainty of +/-0.058C
-B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured barometric pressure
+# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa), but EPA calibration
+# checks performed in the AWBERC metrology lab demonstrate much greater precision.
+# below we calculat sd of replicate measurements made with 15 different barometers.
+# we then take the mean of the 15 sd values.
+
+#2023 barometer verification in AWBERC metrology lab
+B_sd <- mean(
+  sd(c(999,998.6)*0.1), # YSI ProDSS
+  sd(c(1000.848, 1000.848)*0.1), # EXO 21G104022
+  sd(c(1000.982, 1000.982)*0.1), # EXO 18H111572
+  sd(c(1000.982, 1000.982)*0.1), # EXO 22F105975
+  sd(c(1001.382, 1001.382)*0.1), # EXO 21K100162
+  sd(c(1000.98, 1000.85)*0.1), # EXO 21F103579
+  sd(c(1000.31, 1000.44)*0.1), # MDS 16D100327
+  sd(c(1001.25, 1001.25)*0.1), # MDS 12F102618
+  sd(c(1000, 1000)*0.1), # MDS 16D100328
+  sd(c(1002, 1002)*0.1), # MDS 02K599 AF
+  sd(c(1003,1003)*0.1), # MDS 12J0793 AF
+  sd(c(1000.85,1000.85)*0.1), # MDS 16E100831
+  sd(c(1000.32, 1000.32)*0.1), # MDS 12F102619
+  sd(c(1002, 1002)*0.1)) # MDS 15F104725
+
+B_me <- rnorm( 1e4, B, B_sd ) #m.e. in measured barometric pressure
 
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
@@
 # was deprecated in ggplot 3.4.0
-C_df <- tibble(C_me = C_me, B_me = B_me, B_ru = runif(n = 1e4, min = B - 0.4, max = B + 0.4))
+C_df <- tibble(C_me = C_me, B_me = B_me)
 
 ggplot(C_df, aes( x = B_me )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = B , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in barometric pressure") +
-  xlab( "barometric pressure (kPa)")
-
-ggplot(C_df, aes( x = B_ru )) +  
   geom_histogram( binwidth = 0.01 ) + 
@@
 ## Simulate measurement error in water volume
+
 ```{r simulate_me_V_w, message=F}
@@
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
@@
 ## Simulate measurement error for all observables
+
 ```{r simulate_me_all, message=F}
@@
 for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
+  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_n2o_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
 }
@@
 ## Incorporate measurement uncertainty into source/sink status
+
 The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.
@@
 ```
+
 # Set up model for simulating true saturation concentration
@@
 $$C_{sat} = 1^{-6}B*C_{air}*(H^{\theta}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15}))$$
+
 ```{r c_sat}
-C_sat <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+C_sat <- C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
 
@@
 ## Simulate measurement error in GC measurement of N2O in air
-```{r simulate_me_sat_C_a, message=F}
+
+```{r simulate_me_sat_C_a_n2o, message=F}
 C_me_sat <- rep( NA, 1e4 )
 
-C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air
+C_a_n2o_me <- rnorm( 1e4, C_a_n2o, n2o_sd_epa ) #m.e. in measured gas in air
 
 for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a_me[i] * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+    C_me_sat[i] <- C_a_n2o_me[i] * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
 }
@@
 # was deprecated in ggplot 3.4.0
-C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_me = C_a_me)
+C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_n2o_me = C_a_n2o_me)
 
-ggplot(C_sat_df, aes(x = C_a_me )) +  
+ggplot(C_sat_df, aes(x = C_a_n2o_me )) +  
   geom_histogram( binwidth = 0.0001 ) + 
-  geom_vline( xintercept = C_a , color = 'red' ) + 
+  geom_vline( xintercept = C_a_n2o , color = 'red' ) + 
   ggtitle( "Simulated measurement error in air gas concentration") +
@@
 ## Simulate measurement error in water temperature
+
 ```{r simulate_me_sat_T_c, message=F}
@@
 for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
+    C_me_sat[i] <- C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
 }
 
-# combine into dataframe for ggplot. qplot, which acceptes vectors,
+# combine into dataframe for ggplot. qplot, which accepts vectors,
 # was deprecated in ggplot 3.4.0
@@
 ## Simulate measurement error in barometric pressure
+
 ```{r simulate_me_sat_B, message=F}
@@
 for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+    C_me_sat[i] <- C_a_n2o * B_me[ i ] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
 }
@@
 ## Simulate measurement error for all observables
+
 ```{r simulate_me_sat_all, message=F}
@@
 for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a_me[ i ] * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
+    C_me_sat[i] <- C_a_n2o_me[ i ] * B_me[ i ] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
 }
@@
 
-``` {r}
+```{r}
 
-C_ratio <- (1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
-  (C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))
+C_ratio <- (1e-6 * B * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
+  (C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))
 
 ```
-## Simulate measurement error for all observables
+
+## Simulate measurement error for all observables when air is used as headspace gas
+
 ```{r simulate_me_ratio_all, message=F}
@@
 for( i in 1:length( C_me_ratio ) ){
-C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - C_a_me[i] ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
-  (C_a_me[i] * B_me[i] * 1e-6 * (H  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
+C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - C_a_n2o_me[i] ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
+  (C_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
 }
@@
 # standard deviation of a saturation ratio value s:
-sd(C_me_ratio) # 0.0.05479
+sd(C_me_ratio) # 0.026
 
 # the 95% CI is therefore the computed value +/- 2 standard deviations
-sd(C_me_ratio) * 1.96 # +/- 0.1073
+sd(C_me_ratio) * 1.96 # +/- 0.05258214
 
-# so any ratio within 0.1073 of 1.0 cannot be discerned as a source or sink?
+# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
 ```
+
+## Map of saturation ratio values with uncertainty when air is used as headspace gas
+
+```{r map_sat_ratio, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.1, 10), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+## Simulate measurement error for all observables when a pure gas (e.g. helium, nitrogen) is used as headspace gas
+
+```{r simulate_me_ratio_pure_gas_all, message=F}
+# set C_e to a lower value to generate a saturation ratio close to that used above
+# C_e will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
+C_e <- 0.2
+
+# this simulates measurement error on GC.
+# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
+C_e_me <- rnorm( 1e4, C_e, n2o_sd_epa ) # me in measured gas in headspace.
+                                   # add random error to measured concentration
+
+# under this scenario we don't need to subtract C_a_n2o (N2O concentration in air) from C_e (N2O concentration in equilibrated headspace)
+C_ratio <- (1e-6 * B * ( V_g * ( C_e - 0 ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
+  (C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))
+
+C_me_ratio <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_ratio ) ){
+C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - 0 ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
+  (C_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_ratio_df <- tibble(C_me_ratio = C_me_ratio)
+
+ggplot(C_ratio_df, aes( x = C_me_ratio )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = C_ratio , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors") +
+  xlab( "saturation ratio")
+
+ggplot(C_ratio_df, aes( x = C_me_ratio - C_ratio )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors") +
+  xlab( "absolute error (saturation ratio)")
+
+# standard deviation of a saturation ratio value s:
+sd(C_me_ratio) # 0.048
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me_ratio) * 1.96 # +/- 0.094
+
+# so any ratio within 0.094 of 1.0 cannot be discerned as a source or sink
+```
+
+## Simulate measurement error for all observables when a pure gas (e.g. helium, nitrogen) is used as headspace gas combined with high precision GC
+
+The GC used in the current study is optimized to quantify several analytes across a broad range of concentrations but greater precision can be achieved by optimizing the instrument for a single analyte across a narrower range of concentrations.
+For example,  laboratories participating in an international analysis of N2O standards used for atmsopheric monitoring reported a mean standard deviation of `r mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015))`, which is 10-fold lower than the standard deviation of N2O measurements in this study (`r n2o_sd_epa`).
+Utilizing more precise N2O measurements for the equilibrated headspace and/or air will reduce uncertainty in the estimated dissolved N2O concentration, dissolved N2O equilibrium value, and N2O saturation ratio.
+
+```{r simulate_me_ratio_gc_pure_gas_all, message=F}
+# set C_e to a lower value to generate a saturation ratio close to that used above
+# C_e will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
+C_e <- 0.2
+
+# two standard deviations reported from each of three labs participating in
+# a laboratory intercomparison study
+n2o_sd_lit <- mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015)) # from literature
+
+# simulate improved measurement of N2O in equilibrated headspace gas 
+# assume TRUE value is C_e and error is simulated with with sd from literature
+C_e_me <- rnorm( 1e4, C_e, n2o_sd_lit )
+
+# under this scenario we don't need to subtract C_a_n2o (N2O concentration in air) from C_e (N2O concentration in equilibrated headspace)
+C_ratio <- (1e-6 * B * ( V_g * ( C_e - 0 ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
+  (C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))
+
+C_me_ratio <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_ratio ) ){
+C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - 0 ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
+  (C_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_ratio_df <- tibble(C_me_ratio = C_me_ratio)
+
+ggplot(C_ratio_df, aes( x = C_me_ratio )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = C_ratio , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors") +
+  xlab( "saturation ratio")
+
+ggplot(C_ratio_df, aes( x = C_me_ratio - C_ratio )) +  
+  geom_histogram( binwidth = 0.001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors") +
+  xlab( "absolute error (saturation ratio)")
+
+# standard deviation of a saturation ratio value s:
+sd(C_me_ratio) # 0.029
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me_ratio) * 1.96 # +/- 0.0569
+
+# so any ratio within 0.0569 of 1.0 cannot be discerned as a source or sink
+```
+
+# Set up model for MIMS
+
+In Membrane Inlet Mass Spectrometry (MIMS), dissolved gases (typically N2, Ar, O2, N2O) are extracted from a water sample as it passes through a semi-permeable membrane.
+The abundance of the extracted gases are then measured using a quadrapole mass spectrometer.
+One advantage of this approach relative to the traditional headspace equilibrium method is that dissolved gases are extracted without contamination and/or dilution from a headspace gas (e.g.
+air, nitrogen, helium).
+Dilution by a pure gas isn't much of a concern because N2O detection limits are very low via GC, but correction for the presence of N2O in the headspace gas reduces measurement precision.
+
+Dissolved N2O concentration in a MIMS sample can be calculated from the N2O signal strength and a calibration factor.
+The calibration factor is determined by analyzing a water sample with known N2O concentration.
+This is most frequently accomplished by equilibrating laboratory water with air at constant temperature, barometric pressure and N2O partial pressure.
+The dissolved N2O concentration of the equilibrated water is calculated from solubility laws (see [Set up model for simulating true saturation concentration]) and the calibration factor is:
+
+$$N_2O \text{ calibration factor} =\frac{[N_2O_{\text{dissolved standard}}]}{N_2O_{\text{standard}} \text{ signal strength}}$$
+
+The N2O concentration of an unknown sample is then calculated as:
+
+$$C = [N_2O_{sample}] = N_2O_{\text{sample}} \text{ signal strength} * N_2O \text{ calibration factor}$$
+
+The equation simply scales the concentration of the N2O standard to the ratio of the MIMS signal strength of the unknown and standard.
+Via substitution and rearrangement, the equation can be written as:
+
+$$C =[N_2O_{sample}] = \frac{N_2O_{\text{sample}} \text{ signal strength}}{N_2O_{\text{standard}} \text{ signal strength}} * [N_2O_{\text{dissolved standard}}]$$
+MIMS provides greater precision when measuring gas ratios as compared to measuring a single gas.
+High precision gas ratio data can be converted to concentration data with the independent measurement of concentration of one of the component gases.
+N2O:Ar ratios are particularly useful for environmental samples because Ar is biologically inert, therefore its concentration can be calculated from solubility, water temperature, barometric pressure, and atmospheric partial pressure (essentially constant across the globe).
+As with single gas measurements, N2O:Ar ratios are converted to N2O concentration using a calibration factor derived from measurement of laboratory water equilibrated with air:
+$$N_2O \text{ calibration factor} =\frac{[N_2O:Ar]_{\text{standard}}}{N_2O_{\text{standard}} \text{ signal strength }\text{: }{Ar_{\text{standard}} \text{ signal strength}}}$$ 
+The N2O concentration of an unknown sample is then calculated as: 
+$$C = [N_2O_{sample}] =  [N_2O_\text{sample} \text{ signal strength : }Ar_\text{sample}\text{signal strength}] * N_2O \text{ calibration factor} * [Ar_{\text{dissolved standard}}]$$
+As above, the equation simply scales the N2O:Ar ratio measured in the sample to that of the standard.  Via substitution and rearrangement, the equation can be written as: 
+$$C = [N_2O_{sample}] = \frac{\frac{N_2O_{\text{sample}} \text{ signal strength}}{Ar_{\text{sample}} \text{ signal strength}}}{\frac{N_2O_{\text{standard}} \text{ signal strength}}{Ar_{\text{standard}} \text{ signal strength}}} * [N_2O:Ar]_{\text{standard}} * [Ar]_{\text{standard}}$$
+Concentration of N2O and Ar in the standard is calculated from temperature of the equilibrated water (T_C_standard), laboratory barometric pressure (B), N2O and Ar partial pressure in the atmosphere (C_a_N2O, C_a_Ar, respectively)(see [Set up model for simulating true saturation concentration]).
+By substitution, the equation becomes:
+$$C = [N_2O_{sample}] = \frac{\frac{N_2O_{\text{sample}} \text{ signal strength}}{Ar_{\text{sample}} \text{ signal strength}}}{\frac{N_2O_{\text{standard}} \text{ signal strength}}{Ar_{\text{standard}} \text{ signal strength}}} *\frac{ 1^{-6}B*C_{a\text_N2O}*(H_{N2O}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15}))}{1^{-6}B*C_{a\text_Ar}*(H_{Ar}*exp(1500* \frac{1}{Tc} -  \frac{1}{298.15}))} * 1^{-6}B*C_{a\text_Ar}*(H_{Ar}*exp(1500* \frac{1}{Tc} -  \frac{1}{298.15}))$$
+N2O concentration in laboratory air (Ca_N2O) is derived from direct measurement of N2O via gas chromatography and has a standard deviation of `r n2o_sd_epa` using the method employed in this study [Define GC measurement error], although more precise methods are available (see below).
+Ar in the lab atmosphere (Ca_Ar) is 9340ppm and is assumed to be constant and perfectly known (e.g.
+zero error).
+Genther et al (2013) report a standard deviation of 0.022 for the N2O:Ar signal strength ratio but Speir et al (2023) report a CV of 0.05% (sd ~ 0.000096).
+Here we provisionally use the Speir value pending confirmation.
+MIMS uses a very precise thermister  (+/-0.01 C) to control the temperature of the water used as the analytical standard (simulated below).
+Measurement error in B quantified above.
+
+## Simulate measurement error for all observables
+
+```{r simulate_me_pure_gas_all_MIMS, message=F}
+
+n2o_ar_standard_signal <- 0.27 # Speir et al. 0.27 pers comm. Genther et al. section 3.4 reports 0.192
+n2o_ar_sample_signal <- 0.23 # just a guess
+C_a_ar <- 9340 #ppm.  invariant
+T_c_mims <- T_c # set up for simulation of MIMS temperature bath
+
+# this simulates MIMS measurement.
+# assume TRUE value is n2o_ar_standard_signal and error is simulated with CV from Speier et al. 2023
+# Speier et al reports "...our precision was good among all sample and standard replicated (average CV = 0.05%)."
+# Using that CV to model measurement error in the N2O:Ar ratio yields a sd of dissolved N2O that exceeds
+# the sd when measured via GC, which is weird.  Speier reported that the CV is for the N2O:Ar ratio.
+# Genther et al reports a n2o:Ar sd of 0.022, but this also generated a sd for dissolved N2O which exceeds that 
+# measured with GC!  Here we use the Speir data
+n2o_ar_signal_sd <- 0.0005 * n2o_ar_standard_signal # sd = CV * mean.  0.05% = 0.0005
+
+n2o_ar_standard_signal_me <- rnorm( 1e4, n2o_ar_standard_signal, n2o_ar_signal_sd ) # me in N2O:Ar of standard  
+n2o_ar_sample_signal_me <- rnorm( 1e4, n2o_ar_sample_signal, n2o_ar_signal_sd ) # me in N2O:Ar of sample
+
+# High precision water batch used to control temperature of standards
+T_c_mims_me <- rnorm(1e4, T_c_mims, 0.005) # thermostated bath precision is +/-0.01 Celsius
+# hist(T_c_mims_me) # sd of 0.005 gives expected distribution for the specified precision
+
+# True dissolved N2O concentration
+C <- ((n2o_ar_sample_signal / n2o_ar_standard_signal) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B * C_a_n2o * (H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) )) / #N2O concentration in standard
+  (1e-6 * B * C_a_ar * (H_ar * exp( 1500 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) )) ) * #Ar concentration in standard
+  (1e-6 * B * C_a_ar*(H_ar  * exp(1500 *(1/(T_c + 273.15) - 1/298.15)))))
+
+C <- C * 1e9 #convert mol to nmol
+
+# variation in measured N2O concentration
+C_me <- rep(NA, 1e4)
+
+for( i in 1:length( C_me ) ){
+C_me[ i ] <- (n2o_ar_sample_signal_me[ i ] / n2o_ar_standard_signal_me[ i ]) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B_me[ i ] * C_a_n2o_me[i] * (H_n2o * exp( 2700 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) / #N2O concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar * exp( 1500 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) ) * #Ar concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar  * exp(1500 *(1/(T_c_mims_me[ i ] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot.
+C_me_df <- tibble(C_me = C_me * 1e9)
+
+ggplot(C_me_df, aes( x = C_me)) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = C , color = 'red' ) + 
+  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors using MIMS") +
+  xlab( "dissolved N2O (nmol L-1)")
+
+ggplot(C_me_df, aes( x = C_me - C )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors using MIMS") +
+  xlab( "absolute error (nmol L-1)")
+
+# standard deviation of dissolved N2O via MIMS:
+sd(C_me * 1e9) # 0.169 using Speir CV to calculate N2O:Ar sd.   0.15 with GC and headspace equilibrium!
+# CV of dissolved N2O
+sd(C_me * 1e9) / mean(C_me * 1e9) # 0.025 = 2.5%  12% using Genther sd.
+
+```
+## Simulate measurement error for all observables with improved GC precision
+
+The GC used in the current study is optimized to quantify several analytes across a broad range of concentrations but greater precision can be achieved by optimizing the instrument for a single analyte across a narrower range of concentrations.
+For example,  laboratories participating in an international analysis of N2O standards used for atmsopheric monitoring reported a mean standard deviation of `r n2o_sd_lit`, which is 10-fold lower than the standard deviation of N2O measurements in this study (`r n2o_sd_epa`).
+Utilizing more precise N2O measurements for the equilibrated headspace and/or air will reduce uncertainty in the estimated dissolved N2O concentration, dissolved N2O equilibrium value, and N2O saturation ratio.
+
+```{r simulate_me_gc_pure_gas_all_MIMS, message=F}
+# simulate improved measurement precision of GC measurement of N2O in air
+C_a_n2o_lit_me <- rnorm(1e4, C_a_n2o, n2o_sd_lit)
+
+# variation in measured N2O concentration
+C_me <- rep(NA, 1e4)
+
+for( i in 1:length( C_me ) ){
+C_me[ i ] <- (n2o_ar_sample_signal_me[ i ] / n2o_ar_standard_signal_me[ i ]) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B_me[ i ] * C_a_n2o_lit_me[i] * (H_n2o * exp( 2700 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) / #N2O concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar * exp( 1500 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) ) * #Ar concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar  * exp(1500 *(1/(T_c_mims_me[ i ] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot.
+C_me_df <- tibble(C_me = C_me * 1e9)
+
+ggplot(C_me_df, aes( x = C_me)) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = C , color = 'red' ) + 
+  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors using MIMS") +
+  xlab( "dissolved N2O (nmol L-1)")
+
+ggplot(C_me_df, aes( x = C_me - C )) +  
+  geom_histogram( binwidth = 0.01 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors using MIMS") +
+  xlab( "absolute error (nmol L-1)")
+
+# standard deviation of dissolved N2O via MIMS:
+sd(C_me * 1e9) # 0.0168 using Speir CV to calculate N2O:Ar sd. 0.15 with GC and headspace equilibrium!
+
+# CV of dissolved N2O
+sd(C_me * 1e9) / mean(C_me * 1e9) # 0.002 = 0.25% CV
+
+```
+
+## Simulate ratio with MIMS and standard GC
+```{r simulate_me_ratio_pure_gas_all_MIMS, message=F}
+# True ratio
+C_ratio <- ((n2o_ar_sample_signal / n2o_ar_standard_signal) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B * C_a_n2o * (H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) )) / #N2O concentration in standard
+  (1e-6 * B * C_a_ar * (H_ar * exp( 1500 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) )) ) * #Ar concentration in standard
+  (1e-6 * B * C_a_ar*(H_ar  * exp(1500 *(1/(T_c + 273.15) - 1/298.15))))) / # Ar concentration in standard
+  (C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))) # N2o saturation value in waterbody
+
+C_me_ratio <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_ratio ) ){
+C_me_ratio[ i ] <- (n2o_ar_sample_signal_me[ i ] / n2o_ar_standard_signal_me[ i ]) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B_me[ i ] * C_a_n2o_me[ i ] * (H_n2o * exp( 2700 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) / #N2O concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar * exp( 1500 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) ) * #Ar concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar  * exp(1500 *(1/(T_c_mims_me[ i ] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
+  (C_a_n2o_me[ i ] * B_me[ i ] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody 
+}
+
+# combine into dataframe for ggplot.
+C_ratio_df <- tibble(C_me_ratio = C_me_ratio)
+
+ggplot(C_ratio_df, aes( x = C_me_ratio )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
+  geom_vline( xintercept = C_ratio , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors using MIMS and standard GC analytics") +
+  xlab( "saturation ratio")
+
+ggplot(C_ratio_df, aes( x = C_me_ratio - C_ratio )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors using MIMS and standard GC analytics") +
+  xlab( "absolute error (saturation ratio)")
+
+# standard deviation of a saturation ratio value s:
+sd(C_me_ratio) # 0.0011
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me_ratio) * 1.96 # +/- 0.00216
+
+# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink
+
+```
+
+## Simulate ratio with MIMS and high precision GC
+```{r simulate_me_ratio_gc_pure_gas_all_MIMS, message=F}
+
+C_me_ratio <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_ratio ) ){
+C_me_ratio[ i ] <- (n2o_ar_sample_signal_me[ i ] / n2o_ar_standard_signal_me[ i ]) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B_me[ i ] * C_a_n2o_lit_me[ i ] * (H_n2o * exp( 2700 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) / #N2O concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar * exp( 1500 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) ) * #Ar concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar  * exp(1500 *(1/(T_c_mims_me[ i ] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
+  (C_a_n2o_lit_me[ i ] * B_me[ i ] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody 
+}
+
+# combine into dataframe for ggplot.
+C_ratio_df <- tibble(C_me_ratio = C_me_ratio)
+
+ggplot(C_ratio_df, aes( x = C_me_ratio )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
+  geom_vline( xintercept = C_ratio , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors using MIMS and high precision GC analytics") +
+  xlab( "saturation ratio")
+
+ggplot(C_ratio_df, aes( x = C_me_ratio - C_ratio )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors using MIMS and standard GC analytics") +
+  xlab( "absolute error (saturation ratio)")
+
+# standard deviation of a saturation ratio value s:
+sd(C_me_ratio) # 0.0011
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me_ratio) * 1.96 # +/- 0.00216
+
+# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink
+
+```
+
+## Simulate ratio with MIMS, high precision GC, and high precision temperature in field
+
+The thermometers used to measure water temperature in this study have an uncertainty of +/- 0.058C but electronic thermisters are commonly available (e.g.
+YSI sondes) with an uncertainty of +/-0.01 Celsius.
+More precise water temperature measurements will reduce uncertainty in the equilibrium N2O conentration and the saturation ratio.
+```{r simulate_me_ratio_gc_temp_all_MIMS, message=F}
+
+C_me_ratio <- rep( NA, 1e4 )
+
+for( i in 1:length( C_me_ratio ) ){
+C_me_ratio[ i ] <- (n2o_ar_sample_signal_me[ i ] / n2o_ar_standard_signal_me[ i ]) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B_me[ i ] * C_a_n2o_lit_me[ i ] * (H_n2o * exp( 2700 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) / #N2O concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar * exp( 1500 * ( 1 / ( T_c_mims_me[ i ] + 273.15 ) - 1 / 298.15 ) ) )) ) * #Ar concentration in standard
+  (1e-6 * B_me[ i ] * C_a_ar * (H_ar  * exp(1500 *(1/(T_c_mims_me[ i ] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
+  (C_a_n2o_lit_me[ i ] * B_me[ i ] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_mims_me[ i ] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody 
+}
+
+# combine into dataframe for ggplot.
+C_ratio_df <- tibble(C_me_ratio = C_me_ratio)
+
+ggplot(C_ratio_df, aes( x = C_me_ratio )) +  
+  geom_histogram( binwidth = 0.0001 ) + 
+  geom_vline( xintercept = C_ratio , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors using MIMS, high precision GC analytics, and high precision \nin situ temperature measurements") +
+  xlab( "saturation ratio")
+
+ggplot(C_ratio_df, aes( x = C_me_ratio - C_ratio )) +  
+  geom_histogram( binwidth = 0.00001 ) + 
+  geom_vline( xintercept = 0 , color = 'red' ) + 
+  ggtitle( "Uncertainty in N2O saturation ratio due to cumulative \nmeasurment errors using MIMS, high precision GC analytics, and high precision \nin situ temperature measurements") +
+  xlab( "absolute error (saturation ratio)")
+
+# standard deviation of a saturation ratio value s:
+sd(C_me_ratio) # 0.0006603624
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(C_me_ratio) * 1.96 # +/- 0.00129431
+
+# so any ratio within 0.00129 of 1.0 cannot be discerned as a source or sink
+
+```
+# Summary of results
+
+Precision of the N2O saturation ratio can be improved using high precision GC analytics and N2O-free headspace gas (e.g.
+helium, nitrogen) but the greatest improvement is seen when dissolved N2O is directly measured using MIMS.
+```{r final_figure, message=FALSE}
+
+results <- tibble(ratio_sd = c(0.055,
+                               0.048,
+                               0.029,
+                               0.001207747,
+                               0.001207747,
+                               0.000690139),
+                  method = c("GC",
+                             "GC + helium",
+                             "High precision GC + helium",
+                             "MIMS",
+                             "MIMS + high precision GC",
+                             "MIMS + high precision GC + \nhigh precision water temp"))
+
+results %>%
+  mutate(lower.95 = C_ratio - (1.96 * ratio_sd),
+         upper.95 = C_ratio + (1.96 * ratio_sd),
+         C_ratio = C_ratio) %>%
+  ggplot(aes(C_ratio, method)) +
+  geom_point() +
+  geom_segment(aes(x=lower.95, xend = upper.95, y=method, yend=method)) +
+  xlab("True saturation ratio (black dot) and 95% confidence interval") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 1))
+
+```

### 2024-05-24 scripts/DG_sensitivity_to_measurement_error.Rmd
Updated formulas and figures; Added scenarios

--- 2024-04-05 scripts/DG_sensitivity_to_measurement_error.Rmd
@@
 ---
-title: "Evaluating sensitivity to measurement error in dissolved gas measurements"
+title: "Sensitivity to measurement error in dissolved gas measurements"
 author: "Roy Martin, Jake Beaulieu, Michael McManus"
@@
 output:
-  html_document:
+  github_document:
     toc: true
-    df_print: paged
-  pdf_document:
-    toc: true
-  html_notebook:
-    depth: 4
-    fig_caption: true
+    toc_depth: 5
     number_sections: true
-    toc: true
-    toc_float: true
-link-citations: true
+    keep_html: true
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
 
-<style>
-.vscroll-plot {
-    width: 800px;
-    height: 600px;
-    overflow-y: scroll;
-    overflow-x: hidden;
-}
-</style>
-
-# Background
-In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).
-
-# Setup R
-```{r setup_1, include=FALSE, echo=FALSE}
+```{r setup, warning=FALSE, message=FALSE}
 library(knitr)
-# opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
-# 
-# #set directories and R package library
-# setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
-# .libPaths("C:/Users/rmartin/Desktop/R/library")
-```
-
-```{r load_packages, warning=FALSE, message=FALSE}
-library(brms)
+library(rmarkdown)
 library(bayesplot)
 library(kableExtra)
+library(ggpubr)
 library(tidyverse)
+library(tidybayes)
 library(ggplot2)
@@
 
-options(mc.cores = parallel::detectCores())
-options( max.print = 2000 )
-```
-
-```{r utility_functions, include=FALSE}
 # Identify local path for each user
@@
 
-#Calc geometric mean
-gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
-  if(any(x < 0, na.rm = TRUE)){
-    return(NaN)
-  }
-  if(zero.propagate){
-    if(any(x == 0, na.rm = TRUE)){
-      return(0)
-    }
-    exp(mean(log(x), na.rm = na.rm))
-  } else {
-    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
-  }
-}
-
-gm_sd = function(x, na.rm=TRUE, zero.propagate = FALSE){
-  if(any(x < 0, na.rm = TRUE)){
-    return(NaN)
-  }
-  if(zero.propagate){
-    if(any(x == 0, na.rm = TRUE)){
-      return(0)
-    }
-    exp(sd(log(x), na.rm = na.rm))
-  } else {
-    exp(sd(log(x[x > 0]), na.rm=na.rm) / length(x))
-  }
-}
-
-#calculate skew
-MySkew <- function(x) {
-     n <- length(x)
-     v <- var(x)
-     m <- mean(x)
-     third.moment <- (1/(n - 2)) * sum((x - m)^3)
-     third.moment/(var(x)^(3/2))
- }
-
-#Center and scale by (1) sd
-MyStd <- function(x){ (x-mean(x))/(sd(x)*2)}
-
-#First letter in string to uppercase
-firstup <- function(x) {
-  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
-  x}
-```
-
-# Dissolved gas concentration calculation
-
-The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water is:
-
-$$C = 1^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}$$
-
-# Set up model for simulating true value
-
-```{r setvars_and_check}
+options(max.print = 2000)
+```
+
+# Background
+
+In this document we explore a range of scenarios wherein measurement errors 
+could arise in methods for determining the concentration of $\text{N}_2\text{O}$ 
+from field and laboratory measurements observed as part of the 2017 National 
+Lakes Assessment (NLA).
+
+# Dissolved concentration
+
+The equation for estimating the concentration (nM) of $\text{N}_2\text{O}$ gas
+dissolved in water is:
+
+$$C = B^{-6} \cdot \left(\frac{V_g(M_h - M_r)}{(8.3144598 \cdot K \cdot V_w) + (H^{\theta} \cdot e^{2700 \cdot ( \frac{1}{K} - \frac{1}{298.15})} \cdot M_h)} \right)$$
+
+where:
+
+-   $C$ = concentration (nM)
+-   $B$ = barometric pressure (kPa)
+-   $V_g$ = volume of reference gas in headspace (mL)
+-   $M_h$ = gas mixture ratio in the headspace (C_e below)
+-   $M_r$ = gas mixture ratio in the reference (C_a_n2o below)
+-   $V_w$ = volume of water below the headspace (mL)
+-   $K$ = temperature (K) (temp celsius or T_c + 273 below)
+-   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ or $H_{Ar}$ below)
+
+and:
+
+$$\begin{split}
+ V & = V_g + V_w \\
+  V_g & = V \cdot R_M \\
+  V_w & = V \cdot (1 - R_M)
+  \end{split}$$
+
+where:
+
+-   $V$ = Total volume of the mixing vessel (e.g. 140mL syringe)
+-   $V_g$ = Volume of gas (mL) in the mixing vessel 
+-   $V_w$ = Volume of water (mL) in the mixing vessel
+-   $R_M$ = Mixing ratio (gas:water)
+
+## Simulate $\text{N}_2\text{O}$
+
+```{r summary_empirical, echo=TRUE}
 load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
@@
           median_n2o = median(n2o.ppm.gc, na.rm = TRUE)) #mean = 0.36, median = 0.31
-
-#    mean_n2o  median_n2o
-#AIR	0.3093312	0.3104941		
-#DG	 0.3618681	0.3067609
+```
+
+Simulate a single concentration without measurment error to represent the "true" 
+or "known" value for dissolved $\text{N}_2\text{O}$.
+```{r dg_calc, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel
+R_M <- 0.25
 
 B <- 99 # barometric pressure (kPa)
-V_g <- 35 # volume of reference gas in headspace
-V_w <- 105 # volume of water below headspace
+V_g <- V * R_M # volume of reference gas in headspace
+V_w <- V * (1 - R_M) # volume of water below headspace
 T_c <- 23 # temperature ( C )
-C_e <- 0.307 # gas mixture ratio in headspace
-C_a <- 0.310 # gas mixture ratio in reference
-H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
-
-C <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+M_h <- 0.307 # gas mixture ratio in headspace (Mr_h) * Changes as function of R_M at constant C
+M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant
+
+C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )
 
@@
 round( C, 1 ) 
-
-```
-## Define GC measurement error
-Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.
-
-```{r get_gc_error}
+```
+
+```{r dg_calc_vary_r, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel
+R_M <- 0.25
+
+B <- 99 # barometric pressure (kPa)
+V_g <- V * R_M # volume of reference gas in headspace
+V_w <- V * (1 - R_M) # volume of water below headspace
+T_c <- 23 # temperature ( C )
+M_h <- 0.307 # gas mixture ratio in headspace (Mr_h)
+M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant
+
+C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )
+
+C <- C * 1e9 #convert mol to nmol
+
+round( C, 1 ) 
+```
+
+## Empricial variation
+Estimate variation due to measurement uncertainty using empirical data from
+laboratory replicates.
+```{r get_gc_error, echo=FALSE}
 # READ GC DATA----------------
@@
 
-## Simulate measurement error in GC measurement of N2O in equilibrated headspace
-```{r simulate_me_Ce, message=F}
-C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector
-
-# this simulates measurement error on GC.
-# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
-C_e_me <- rnorm( 1e4, C_e, n2o_sd ) # me in measured gas in headspace.
-                                   # add random error to measured concentration
-
-for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
+Measurement error is defined here as the standard deviation of repeated
+analysis of analytical standards throughout the period of time that the
+NLA2017 samples were analyzed, which was $\sigma$ = `r n2o_sd_epa`.
+
+## Simulate measurment error
+
+Below, potential errors associated with dissolved $\text{N}_2\text{O}$ 
+measurement are simulated and assessed separately for GC observations in:
+
+-
+
+    (1) the equilibrated headspace,
+
+-
+
+    (2) the air, and
+
+-
+
+    (3) the headspace + air
+
+Measurement error is then simulated and assessed separately for:
+
+-
+
+    (4) water temperature,
+
+-
+
+    (5) barometric pressure, and
+
+-
+
+    (5) water volume
+
+Finally, all of the sources (1, 2, 4, 5, and 6) are simulated
+simultaneously and assessed as the total variation in dissolved
+$\text{N}_2\text{O}$. This total error is then added to the point estimates for 
+$\text{N}_2\text{O}$ observed in this study, which results in an estimate of
+uncertainty due to measurement error. These error from these simulations was also
+carried through to estimated of the saturation ratio. The source-sink status for
+each site included in this study was then mapped to present a source-sink 
+categorization that reflected these estimated uncertainties. The categorization 
+was made such that any observation with simulated error that was within the 
+central 95th percentile of
+
+Set the number of simulations for error:
+
+```{r set_n_sims}
+nsim <- 1e5
+```
+
+### Equilibrated headspace
+Simulate measurment error in dissolved $\text{N}_2\text{O}$ due solely to variation in 
+measurements within the equlibrated headspace.
+```{r simulate_C_h_me, echo=TRUE, message=FALSE}
+# Create empty vector to fill below with simulated values for N2O observed with 
+# error due to variation in the GC measurments of gas in the equilibrated 
+# headspace. The subscript, "me", denotes an observed or measured value, which 
+# is contrasted with the known value, C, set above.
+C_h_me <- rep(NA, nsim)
+
+# The true value for the concentration of equilibrated gas in the headspace, 
+# C_e, was set above. This step simulates additive error in this measurement 
+# that is centered over the true value. The observed concentration of
+# equilibrated gas is simulated as normally distributed variable, with a mean
+# C_e, and a standard deviation specified as the observed sd of replicate GC 
+# measurements in our lab. 
+M_h_me <- rnorm(nsim, M_h, n2o_sd_epa)
+
+for(i in 1:nsim){
+  C_h_me[i] <- 1e-6 * B * (V_g * (M_h_me[i] - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_me[i])
 }
 
-# combine into dataframe for ggplot. qplot, which acceptes vectors,
-# was deprecated in ggplot 3.4.0
-C_df <- tibble(C_me = C_me, C_e_me = C_e_me)
-
-ggplot(C_df, aes(x = C_e_me )) +  
-  geom_histogram( binwidth = 0.0001 ) + 
-  geom_vline( xintercept = C_e , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in equilibrated gas concentration") +
-  xlab( "concentration (nmol L-1)")
-
-ggplot(C_df, aes(x = C_me * 1e9 )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error for GC measurement of N2O in air
+# combine into dataframe for ggplot. 
+C_h_df <- tibble(C_h_me = C_h_me, M_h_me = M_h_me)
+# save absolute error for plot
+C_h_me_abs <- (C_h_me * 1e9) - C
+```
+
+The results of the simulations are summarized with density histograms
+below.
+
+```{r plot_C_h_me, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_h_df, aes(x = M_h_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_h , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurements") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_h_df, aes(x = C_h_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_h_df, aes(x = C_h_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+### Ambient air
+Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) is simulated 
+below. The result of the simulation is plotted as a histogram to illustrate the
+variation in the measurement. THe resulting effect on the uncertainty of the 
+dissolved N2O measurement is also plotted along with the absolute error in 
+N2O as a result of error in this measurement.
 ```{r simulate_me_Ca, message=F}
-C_me <- rep( NA, 1e4 )
-
-C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air
-
-for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+C_a_me <- rep(NA, nsim)
+
+M_a_n2o_me <- rnorm(nsim, M_a_n2o, n2o_sd_epa) # m.e. in measured gas in air
+
+for(i in 1:nsim){
+  C_a_me[i] <- 1e-6 * B * (V_g * (M_h - M_a_n2o_me[i]) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
 }
@@
 # was deprecated in ggplot 3.4.0
-Ca_df <- tibble(C_me = C_me, C_a_me = C_a_me)
-
-ggplot(Ca_df, aes(x = C_a_me )) +  
-  geom_histogram( binwidth = 0.0001 ) + 
-  geom_vline( xintercept = C_a , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in air gas concentration") +
-  xlab( "concentration (?)")
-
-ggplot(Ca_df, aes(x = C_me * 1e9 )) +  
+C_a_df <- tibble(C_a_me = C_a_me, M_a_n2o_me = M_a_n2o_me)
+# save absolute error for plot
+C_a_me_abs<- (C_a_me * 1e9) - C
+```
+
+```{r plot_me_Ca, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_a_df, aes(x = M_a_n2o_me, y = after_stat(density))) +  
   geom_histogram( binwidth = 0.001 ) +
-  geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in air") +
-  xlab( "concentration (nMol)")
-
-ggplot(Ca_df, aes(x = C_me * 1e9 - C )) +  
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = M_a_n2o , color = 'red' ) +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Simulated variation in \nreference gas (Ma) GC \nmeasurements") +
+  xlab( "Concentration (nmol L-1)"),
+
+ggplot(C_a_df, aes(x = C_a_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_a_df, aes(x = C_a_me_abs, y = after_stat(density))) +  
   geom_histogram( binwidth = 0.01 ) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
   geom_vline( xintercept = 0 , color = 'red' ) +  
-  ggtitle( "Uncertainty in N2O concentration due to GC measurment error \nof N2O in air") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error in $C_e$ and $C_a$
-```{r simulate_me_Ca_Ce, message=F}
-C_me <- rep( NA, 1e4 )
-
-for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
-}
-
-# combine into dataframe for ggplot. qplot, which acceptes vectors,
-# was deprecated in ggplot 3.4.0
-C_df <- tibble(C_me = C_me)
-
-ggplot(C_df, aes(x = C_me * 1e9 )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof 
-           N2O in equilibrated headspace and air") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in equilibrated headspace and air") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error in water temperature
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1)
+```
+
+### Water temperature
+
 ```{r simulate_me_T_c, message=F}
-C_me <- rep( NA, 1e4 )
+C_tc_me <- rep(NA, nsim)
 
@@
 # take standard deviation of simulated measurements with specified uncertainty
-T_c_sd <- sd(runif(n = 1e4, min = T_c - 0.058, max = T_c + 0.058)) # calibration certificate reports an uncertainty of +/-0.058C
-T_c_me <- rnorm( 1e4, T_c, T_c_sd ) #m.e. in measured water temp
-
-for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
-}
-
-# combine into dataframe for ggplot. qplot, which acceptes vectors,
-# was deprecated in ggplot 3.4.0
-C_df <- tibble(C_me = C_me, T_c_me = T_c_me)
-
-ggplot(C_df, aes(x = T_c_me )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = T_c , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in temperature") +
-  xlab( "temperature (C)")
-
-ggplot(C_df, aes( x = C_me * 1e9 )) +  
-  geom_histogram( binwidth = 0.001 ) + 
-  geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
-  geom_histogram( binwidth = 0.001 ) + 
-  geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error in barometric pressure
-```{r simulate_me_B, message=F}
-C_me <- rep( NA, 1e4 )
-
-# How to get standard deviation of barometric pressure?
-# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa)
-# take standard deviation of simulated measurements with specified uncertainty
-B_sd <- sd(runif(n = 1e4, min = B - 0.4, max = B + 0.4)) # calibration certificate reports an uncertainty of +/-0.058C
-B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured barometric pressure
-
-for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+T_c_sd <- sd(runif(n = nsim, min = T_c - 0.058, max = T_c + 0.058)) # calibration certificate reports an uncertainty of +/-0.058C
+T_c_me <- rnorm(nsim, T_c, T_c_sd) #m.e. in measured water temp
+
+for(i in 1:nsim){
+  C_tc_me[i] <- 1e-6 * B * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c_me[i] + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h)
 }
@@
 # was deprecated in ggplot 3.4.0
-C_df <- tibble(C_me = C_me, B_me = B_me, B_ru = runif(n = 1e4, min = B - 0.4, max = B + 0.4))
-
-ggplot(C_df, aes( x = B_me )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = B , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in barometric pressure") +
-  xlab( "barometric pressure (kPa)")
-
-ggplot(C_df, aes( x = B_ru )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = B , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in barometric pressure") +
-  xlab( "barometric pressure (kPa)")
-
-ggplot(C_df, aes( x = C_me * 1e9 )) +  
-  geom_histogram( binwidth = 0.001 ) + 
-  geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
-  geom_histogram( binwidth = 0.001 ) + 
-  geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error in water volume
-```{r simulate_me_V_w, message=F}
-C_me <- rep( NA, 1e4 )
-
-V_w_me <- rnorm( 1e4, V_w, 1 ) #m.e. in measured water volume
-V_g_me <- 140 - V_w_me
-
-for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
+C_tc_df <- tibble(C_tc_me = C_tc_me, T_c_me = T_c_me)
+# save absolute error for plot
+C_tc_me_abs <- (C_tc_me * 1e9) - C
+```
+
+```{r plot_me_T_c, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_tc_df, aes(x = T_c_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = T_c , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in water \ntemperature measurements \n") +
+  xlab("Temperature (C)"),
+
+ggplot(C_tc_df, aes(x = C_tc_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_tc_df, aes(x = C_tc_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1)
+```
+
+### Barometric pressure
+
+```{r simulate_me_B, message=F}
+C_b_me <- rep(NA, nsim)
+
+# How to get standard deviation of barometric pressure?
+# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa), but EPA calibration
+# checks performed in the AWBERC metrology lab demonstrate much greater precision.
+# below we calculat sd of replicate measurements made with 15 different barometers.
+# we then take the mean of the 15 sd values.
+
+#2023 barometer verification in AWBERC metrology lab
+B_sd <- mean(
+  sd(c(999,998.6)*0.1), # YSI ProDSS
+  sd(c(1000.848, 1000.848)*0.1), # EXO 21G104022
+  sd(c(1000.982, 1000.982)*0.1), # EXO 18H111572
+  sd(c(1000.982, 1000.982)*0.1), # EXO 22F105975
+  sd(c(1001.382, 1001.382)*0.1), # EXO 21K100162
+  sd(c(1000.98, 1000.85)*0.1), # EXO 21F103579
+  sd(c(1000.31, 1000.44)*0.1), # MDS 16D100327
+  sd(c(1001.25, 1001.25)*0.1), # MDS 12F102618
+  sd(c(1000, 1000)*0.1), # MDS 16D100328
+  sd(c(1002, 1002)*0.1), # MDS 02K599 AF
+  sd(c(1003,1003)*0.1), # MDS 12J0793 AF
+  sd(c(1000.85,1000.85)*0.1), # MDS 16E100831
+  sd(c(1000.32, 1000.32)*0.1), # MDS 12F102619
+  sd(c(1002, 1002)*0.1)) # MDS 15F104725
+
+B_me <- rnorm(nsim, B, B_sd) #m.e. in measured barometric pressure
+
+for(i in 1:nsim){
+  C_b_me[i] <- 1e-6 * B_me[i] * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
 }
@@
 # was deprecated in ggplot 3.4.0
-C_df <- tibble(C_me = C_me, V_w_me = V_w_me)
-
-ggplot(C_df, aes( x = V_w_me )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = V_w , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in water volume") +
-  xlab( "water vol (mL)")
-
-ggplot(C_df, aes( x = C_me * 1e9 )) +  
-  geom_histogram( binwidth = 0.0001 ) + 
-  geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissollved N2O concentration due to water volume \nmeasurment error") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
-  geom_histogram( binwidth = 0.0001 ) + 
-  geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in N2O concentration due to measurment error") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error for all observables
-```{r simulate_me_all, message=F}
-C_me <- rep( NA, 1e4 )
-
-for( i in 1:length( C_me ) ){
-  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
+C_b_df <- tibble(C_b_me = C_b_me, B_me = B_me)
+# save absolute error for plot
+C_b_me_abs <- (C_b_me * 1e9) - C
+```
+
+```{r plot_me_B, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_b_df, aes(x = B_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = B , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nbarometric pressure \nmeasurements") +
+  xlab("Barometric pressure (kPa)"),
+
+ggplot(C_b_df, aes(x = C_b_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_b_df, aes(x = C_b_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1)
+```
+
+### Water volume
+
+```{r simulate_me_V_w, message=F}
+C_w_me <- rep(NA, nsim)
+
+V_w_me <- rnorm(nsim, V_w, 1) #m.e. in measured water volume
+V_g_me <- 140 - V_w_me
+
+for(i in 1:nsim){
+  C_w_me[i] <- 1e-6 * B * (V_g_me[i] * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w_me[i] ) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
 }
@@
 # was deprecated in ggplot 3.4.0
-C_df <- tibble(C_me = C_me)
-
-ggplot(C_df, aes( x = C_me * 1e9 )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = C , color = 'red' ) + 
-  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
-  geom_histogram( binwidth = 0.01 ) + 
+C_w_df <- tibble(C_w_me = C_w_me, V_w_me = V_w_me)
+# save absolute error for plot
+C_w_me_abs <- (C_w_me * 1e9) - C
+```
+
+```{r plot_me_V_w, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_w_df, aes(x = V_w_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.1) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = V_w , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nwater volume \nmeasurements") +
+  xlab("Water volume (mL)"),
+
+ggplot(C_w_df, aes(x = C_w_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.00025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_w_df, aes(x = C_w_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.00025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1)
+```
+
+### All observables
+Simulate measurement error for all observables (_i.e._, Equilibrated headspace, 
+air, water temperature, and barometric pressure ) and the resulting effects on
+dissolved N2O observations
+```{r simulate_me_all, message=F}
+C_all_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+  C_all_me[i] <- 1e-6 * B_me[i] * (V_g_me[i] * (M_h_me[i] - M_a_n2o_me[i] ) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h_me[i])
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_all_df <- tibble(C_all_me = C_all_me)
+# save absolute error for plot
+C_all_me_abs <- (C_all_me * 1e9) - C
+```
+
+```{r plot_me_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(C_all_df, aes(x = C_all_me * 1e9, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.05 ) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in dissolved N2O \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(C_all_df, aes(x = C_all_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.05) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
   geom_vline( xintercept = 0 , color = 'red' ) +
-  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
-  xlab( "absolute error (nMol)")
-
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+Print the estimated standard deviations of dissolved N2O measurements
+and the resulting 95% credible interval for error in measurements.
+
+```{r calc_sd_me_all}
 # standard deviation of a measurement is:
-sd(C_me* 1e9) # 0.15 nmol L-1, as compared to 1.25 nmol L-1 in Webb et al.
+sd(C_all_me* 1e9) # 0.20 nmol L-1
+
+# 2 standard deviations
+sd(C_all_me * 1e9) * 1.96 # +/- 0.395 nmol N2O
+```
+
+```{r calc_sd_me_eq_all_2}
+# standard deviation of a measurement is:
+round(sd(C_all_me_abs), 3) # 0.20 nmol L-1
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((C + C_all_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+## Summary of results
+
+```{r final_figure_N2O_diss, message=FALSE, fig.align='center', fig.asp=0.6, fig.width=6}
+results_dissolved_N2O <- tibble(Eh_me= C_h_me_abs,
+                                Air_me = C_a_me_abs,
+                                Tc_me = C_tc_me_abs,
+                                B_me = C_b_me_abs,
+                                W_me = C_w_me_abs,
+                                All_me = C_all_me_abs) %>%
+  pivot_longer(cols = ends_with("me"), names_to = "Source", values_to = "Error")
+
+results_dissolved_N2O %>%
+  mutate(Source = factor(Source)) %>%
+  mutate(Source = fct_relevel(Source, c("B_me",
+                                        "W_me",
+                                        "Tc_me",
+                                        "Air_me",
+                                        "Eh_me",
+                                        "All_me"
+                                        ))) %>%
+  ggplot(aes(x = Error, y = Source)) +
+  stat_pointinterval() +
+  scale_y_discrete(labels=c("Eh_me" = "Equilibrated headspace",
+                            "Air_me" = "Ambient air",
+                            "Tc_me" = "Water temperature",
+                            "B_me" = "Barometric pressure",
+                            "W_me" = "Water volume",
+                            "All_me" = "All sources")) +
+  xlab("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  ylab("") 
+```
+
+# Equilibrium concentration
+
+The full equation for estimating the concentration (nM) of ($\text{N}_2\text{O}$) gas
+dissolved in water that is in equilibrium with the atmosphere is:
+
+$$C_{eq} = 1^{-6} \cdot B \cdot C_a \cdot H^{\theta} \cdot e^{2700 \cdot \frac{1}{Tc} -  \frac{1}{298.15}}$$
+where:
+
+-   $C_{eq}$ = The equilibrium concentration in nM
+-   $B$ = barometric pressure (kPa)
+-   $C_a$ = The air concentration in nM
+-   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ below)
+-   $T_c$ = temperature celsius
+
+## Simulate equilibrium $\text{N}_2\text{O}$
+
+A "known" concentration is simulated:
+
+```{r c_sat}
+C_eq <- M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+
+C_eq <- C_eq * 1e9 #convert mol to nmol
+```
+
+## Simulate measurement error
+
+### Ambient air
+
+Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) is simulated 
+below. The resulting effects on the uncertainty of the equilibrium N2O 
+measurement is plotted along with the absolute error in N2O as a result of this
+measurement.
+
+```{r simulate_me_sat_C_a_n2o, message=F}
+C_a_eq_me <- rep( NA, nsim )
+
+M_a_n2o_me <- rnorm( nsim, M_a_n2o, n2o_sd_epa ) #m.e. in measured gas in air
+
+for(i in 1:nsim){
+    C_a_eq_me[i] <- M_a_n2o_me[i] * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+}
+
+C_a_eq_df <- tibble(C_a_eq_me = C_a_eq_me, M_a_n2o_me = M_a_n2o_me)
+# save absolute error for plot
+C_a_eq_me_abs <- (C_a_eq_me * 1e9) - C_eq
+```
+
+```{r plot_me_sat_C_a_n2o, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+ggarrange(
+ggplot(C_a_eq_df, aes(x = M_a_n2o_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_a_n2o , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in air \ngas (M_a_n2o) GC \nmeasurements") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_a_eq_df, aes(x = C_a_eq_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C_eq , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \nequilibrium N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_a_eq_df, aes(x = C_a_eq_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+### Water temperature
+
+```{r simulate_me_sat_T_c, message=F}
+C_tc_eq_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+    C_tc_eq_me[i] <- M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1/298.15)))
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_tc_eq_df <- tibble(C_tc_eq_me = C_tc_eq_me, T_c_me = T_c_me)
+# save absolute error for plot
+C_tc_eq_me_abs <- (C_tc_eq_me * 1e9) - C_eq
+```
+
+```{r plot_me_sat_T_c, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_tc_eq_df, aes(x = T_c_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = T_c , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in water \ntemperature measurements") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_tc_eq_df, aes(x = C_tc_eq_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C_eq , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \nequilibrium N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_tc_eq_df, aes(x = C_tc_eq_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+### Barometric pressure
+
+```{r simulate_me_sat_B, message=F}
+C_b_eq_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+    C_b_eq_me[i] <- M_a_n2o * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_b_eq_df <- tibble(C_b_eq_me = C_b_eq_me, B_me = B_me)
+# save absolute error for plot
+C_b_eq_me_abs <- (C_b_eq_me * 1e9) - C_eq
+```
+
+```{r plot_me_sat_B, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_b_eq_df, aes(x = B_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = B , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nbarometic pressure \nmeasurements") +
+  xlab("Barometric pressure (kPa)"),
+
+ggplot(C_b_eq_df, aes(x = C_b_eq_me * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.0005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C_eq , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \nequilibrium N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_b_eq_df, aes(x = C_b_eq_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.00025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+### All observables
+Simulate measurement error for all observables (_i.e._, air, water temperature,
+and barometric pressure ) and the resulting effects on dissolved N2O 
+observations.
+```{r simulate_me_sat_all, message=F}
+C_all_eq_me <- rep( NA, nsim )
+
+for(i in 1:nsim){
+    C_all_eq_me[i] <- M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)))
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+C_all_eq_df <- tibble(C_all_eq_me = C_all_eq_me)
+# save absolute error for plot
+C_all_eq_me_abs <- (C_all_eq_me * 1e9) - C_eq
+```
+
+```{r plot_me_sat_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(C_all_eq_df, aes(x = C_all_eq_me * 1e9, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.025 ) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C_eq , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in equilibrium N2O \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(C_all_eq_df, aes(x = C_all_eq_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_sd_me_sat_all}
+# standard deviation of a measurement is:
+sd(C_all_eq_me* 1e9) # 0.20 nmol L-1
+
+# 2 standard deviations
+sd(C_all_eq_me * 1e9) * 1.96 # +/- 0.395 nmol N2O
+```
+
+```{r calc_sd_me_sat_all_2}
+# standard deviation of a measurement is:
+round(sd(C_all_eq_me_abs), 3) # 0.20 nmol L-1
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((C_eq + C_all_eq_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+## Summary of results
+
+```{r final_figure_N2O_eq, message=FALSE, fig.align='center', fig.asp=0.6, fig.width=6}
+results_equilibrium_N2O <- tibble(Air_me = C_a_eq_me_abs,
+                                Tc_me = C_tc_eq_me_abs,
+                                B_me = C_b_eq_me_abs,
+                                All_me = C_all_eq_me_abs) %>%
+  pivot_longer(cols = ends_with("me"), names_to = "Source", values_to = "Error")
+
+results_equilibrium_N2O %>%
+  mutate(Source = factor(Source)) %>%
+  mutate(Source = fct_relevel(Source, c("B_me",
+                                        "Tc_me",
+                                        "Air_me",
+                                        "All_me"
+                                        ))) %>%
+  ggplot(aes(x = Error, y = Source)) +
+  stat_pointinterval() +
+  scale_y_discrete(labels=c("Air_me" = "Ambient air",
+                            "Tc_me" = "Water temperature",
+                            "B_me" = "Barometric pressure",
+                            "All_me" = "All sources")) +
+  xlab("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  ylab("") 
+```
+
+# Saturation ratio
+(Check this equation in text! Particularly the description of variables)
+The full equation for estimating the saturation ratio of ($N_20$) gas
+dissolved in water is:
+
+$$S = \left(\frac{(1^{-6} \cdot B ) \biggl( \frac{V_a (Mr_h - Mr_r)}{(8.3144598 \cdot K \cdot V_w) + H^{\theta} \cdot e^{2700 \cdot ( \frac{1}{K} - \frac{1}{298.15})} \cdot Mr_h} \biggr)}{1^{-6} \cdot B \cdot C_a \cdot e^{2700 \cdot \frac{1}{Tc} -  \frac{1}{298.15}}} \right)$$
+
+where:
+
+-   $S$ = Saturatio ratio
+-   $B$ = barometric pressure (kPa)
+-   $C_a$ = the air concentration in nM
+-   $V_a$ = volume of reference gas in air
+-   $M_h$ = gas mixture ratio in the headspace
+-   $M_r$ = gas mixture ratio in the reference (C_a_n2o below)
+-   $V_w$ = volume of water below the headspace
+-   $T_c$ = temperature celsius
+-   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ below)
+
+Simulate a "true" saturation ratio.
+```{r sat_ratio_true}
+S <- (1e-6 * B * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)) /
+  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) 
+S
+```
+
+## Standard GC
+Simulate uncertainty in the saturation ratio due to measurement error associated
+with the standard GC, using ambient air as the headspace gas, and the standard 
+field thermometer.
+
+### Ambient air, standard thermometer
+
+```{r simulate_me_ratio_all_air, message=F}
+S_a_me <- rep( NA, nsim )
+
+for(i in 1:nsim){
+S_a_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_me[i] - M_a_n2o_me[i]) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h_me[i])) /
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+S_a_df <- tibble(S_a_me = S_a_me)
+# save absolute error for plot
+S_a_me_abs <- S_a_me - S
+```
+
+```{r plot_me_ratio_all_air, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_a_df, aes(x = S_a_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_a_df, aes(x = S_a_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_sd_me_ratio_all_air}
+# standard deviation of a saturation ratio value s:
+sd(S_a_me) # 0.026
 
 # the 95% CI is therefore the computed value +/- 2 standard deviations
-sd(C_me* 1e9) * 1.96 # +/- 0.302 nmol N2O
-```
-
-## Incorporate measurement uncertainty into source/sink status
-The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.
-The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 1.96`nM.
-Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 1.96`nM of the equilibrium value cannot be determined as a source or a sink.
-
-```{r map_table, message=F}
+sd(S_a_me) * 1.96 # +/- 0.05258214
+
+# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
+```
+
+```{r calc_sd_me_ratio_all_air_2}
+# standard deviation of a measurement is:
+round(sd(S_a_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_a_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S + S_a_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+#### Map source-sink status
+
+```{r map_ratio_all_air, message=F}
 # two standard deviations = 0.3019449
@@
 
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
 dg <- dg %>%
-  mutate(n2o.src.snk.with.error = case_when(
-    abs(dissolved.n2o.nmol - sat.n2o.nmol) > sd(C_me* 1e9) * 1.96 ~ n2o.src.snk,
-    TRUE ~ "undetermined"))
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_a_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
 
@@
   filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
-  {table(.$n2o.src.snk.with.error)} %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
   kable(digits = 2, caption = "Number of lakes classified as source/sink")
@@
   filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
-  {prop.table(table(.$n2o.src.snk.with.error))} %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
 
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_ratio_all_air, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
 ## load ecoregions----
@@
 ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
     geom_sf(data = dg.sf %>% 
               filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.error),
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
             show.legend = "point") +
@@
     # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", values = cols,
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
                       guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
     scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.1, 10), # custom size range
+               range = c(0.5, 8), # custom size range
                breaks = c(1, 10, 25, 50, 100)) + # custom breaks
@@
 ```
-# Set up model for simulating true saturation concentration
-
-The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water that is in equilibrium with the atmosphere is:
-
-$$C = 1^{-6}B*C_{air}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15})$$
-```{r c_sat}
-C_sat <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
-
-C_sat <- C_sat * 1e9 #convert mol to nmol
-
-```
-
-## Simulate measurement error in GC measurement of N2O in air
-```{r simulate_me_sat_C_a, message=F}
-C_me_sat <- rep( NA, 1e4 )
-
-C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air
-
-for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a_me[i] * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+
+### Pure gas, standard thermometer
+
+Simulate uncertainty in the saturation ratio due to measurement error associated
+with the standard GC, using a pure gas (helium) as the headspace gas, 
+and the standard field thermometer.
+
+```{r simulate_me_ratio_pg_all, message=F}
+# set M_h to a lower value to generate a saturation ratio close to that used above
+# M_h will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
+M_h_pg <- 0.2
+
+# this simulates measurement error on GC.
+# assume TRUE value is M_h and error is simulated with with sd of replicate GC measurements
+M_h_pg_me <- rnorm(nsim, M_h_pg, n2o_sd_epa) # me in measured gas in headspace.
+                                   # add random error to measured concentration
+
+# under this scenario we don't need to subtract M_a_n2o (N2O concentration in air) from M_h (N2O concentration in equilibrated headspace)
+S_pg <- (1e-6 * B * (V_g * (M_h_pg - 0) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15) ) * M_h_pg)) /
+  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))
+
+S_pg_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+S_pg_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_me[i] - 0) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h_pg_me[i])) /
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
 }
@@
 # was deprecated in ggplot 3.4.0
-C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_me = C_a_me)
-
-ggplot(C_sat_df, aes(x = C_a_me )) +  
-  geom_histogram( binwidth = 0.0001 ) + 
-  geom_vline( xintercept = C_a , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in air gas concentration") +
-  xlab( "concentration (ppm)")
-
-ggplot(C_sat_df, aes(x = C_me_sat * 1e9 )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = C_sat , color = 'red' ) + 
-  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_sat_df, aes(x = C_me_sat * 1e9 - C_sat )) +  
-  geom_histogram( binwidth = 0.01 ) + 
+S_pg_df <- tibble(S_pg_me = S_pg_me)
+# save absolute error for plot
+S_pg_me_abs <- S_pg_me - S_pg
+```
+
+```{r plot_me_ratio_pg_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_pg_df, aes(x = S_pg_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_pg_df, aes(x = S_pg_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
   geom_vline( xintercept = 0 , color = 'red' ) +
-  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error in water temperature
-```{r simulate_me_sat_T_c, message=F}
-C_me_sat <- rep( NA, 1e4 )
-
-for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
-}
-
-# combine into dataframe for ggplot. qplot, which acceptes vectors,
-# was deprecated in ggplot 3.4.0
-C_sat_df <- tibble(C_me_sat = C_me_sat, T_c_me = T_c_me)
-
-ggplot(C_sat_df, aes(x = T_c_me )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = T_c , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in temperature") +
-  xlab( "temperature (C)")
-
-ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
-  geom_histogram( binwidth = 0.001 ) + 
-  geom_vline(xintercept = C_sat , color = 'red' ) + 
-  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
-  geom_histogram( binwidth = 0.001 ) + 
-  geom_vline( xintercept = 0 , color = 'red' ) + 
-  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error" ) +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error in barometric pressure
-```{r simulate_me_sat_B, message=F}
-C_me_sat <- rep( NA, 1e4 )
-
-for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_sd_me_ratio_pg_all}
+# standard deviation of a saturation ratio value s:
+sd(S_pg_me) # 0.048
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_pg_me) * 1.96 # +/- 0.094
+
+# so any ratio within 0.094 of 1.0 cannot be discerned as a source or sink
+```
+
+```{r calc_sd_me_ratio_all_pg_2}
+# standard deviation of a measurement is:
+round(sd(S_pg_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_pg_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_pg + S_pg_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+#### Map source-sink status
+
+```{r map_ratio_pg, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_pg_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_ratio_pg, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+### Ambient air, high precision thermometer
+Simulate uncertainty in the saturation ratio due to measurement error associated
+with the standard GC, ambient air as the headspace gas, and using a high 
+precision thermometer (e.g., MIMS).
+
+```{r simulate_tc_mims_air_hpt, message=F}
+# High precision water batch used to control temperature of standards
+T_c_mims_me <- rnorm(nsim, T_c, 0.005) # thermo stated bath precision is +/-0.01 Celsius
+# hist(T_c_mims_me) # sd of 0.005 gives approx distribution for the specified precision
+
+T_c_df <- tibble(T_c_me = T_c_me, 
+                 T_c_me_abs = T_c_me - T_c,  
+                 T_c_mims_me = T_c_mims_me,
+                 T_c_mims_me_abs = T_c_mims_me - T_c)
+```
+
+The precision gain from the MIMS thermometer relative to the standard field 
+thermometer is roughly one order of magnitude.
+```{r plot_me_T_c_vs_T_c_MIMS, echo=FALSE, fig.align='center', fig.asp=0.9, fig.width=6}
+ggarrange(
+ggplot(T_c_df, aes(x = T_c_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = T_c , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in water \ntemperature measurements \nusing standard field thermometer") +
+  xlab("Temperature (C)"),
+ggplot(T_c_df, aes(x = T_c_mims_me, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = T_c , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in water \ntemperature measurements \n using MIMS thermometer") +
+  xlab("Temperature (C)"),
+ggplot(T_c_df, aes(x = T_c_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Absolute error in water \ntemperature measurements \nusing standard field thermometer") +
+  xlab("Error"),
+ggplot(T_c_df, aes(x = T_c_mims_me_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Absolute error in water \ntemperature measurements \nusing standard field thermometer") +
+  xlab("Error"),
+nrow = 2,
+ncol = 2)
+```
+
+Simulate uncertainty in the saturation ratio using ambient air in the 
+headspace and high precision temperature measurements
+```{r simulate_me_ratio_all_air_hpt, message=F}
+S_a_hpt_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+S_a_hpt_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_me[i] - M_a_n2o_me[i]) / (8.3144598 * (T_c_mims_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)) * M_h_me[i])) /
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) 
 }
@@
 # was deprecated in ggplot 3.4.0
-C_sat_df <- tibble(C_me_sat = C_me_sat, B_me = B_me)
-
-ggplot(C_sat_df, aes( x = B_me )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = B , color = 'red' ) + 
-  ggtitle( "Simulated measurement error in barometric pressure") +
-  xlab( "barometric pressure (kPa)")
-
-ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
-  geom_histogram( binwidth = 0.001 ) + 
-  geom_vline( xintercept = C_sat , color = 'red' ) + 
-  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
-  geom_histogram( binwidth = 0.001 ) + 
+S_a_hpt_df <- tibble(S_a_hpt_me = S_a_hpt_me)
+# save absolute error for plot
+S_a_hpt_me_abs <- S_a_hpt_me - S
+```
+
+```{r plot_me_ratio_all_air_hpt, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_a_hpt_df, aes(x = S_a_hpt_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_a_hpt_df, aes(x = S_a_hpt_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
   geom_vline( xintercept = 0 , color = 'red' ) +
-  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
-  xlab( "absolute error (nMol)")
-```
-
-## Simulate measurement error for all observables
-```{r simulate_me_sat_all, message=F}
-C_me_sat <- rep( NA, 1e4 )
-
-for( i in 1:length( C_me_sat ) ){
-    C_me_sat[i] <- C_a_me[ i ] * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_sd_tc_mims_air_hpt}
+# standard deviation of a saturation ratio value s:
+sd(S_a_hpt_me) # 0.026
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_a_hpt_me) * 1.96 # +/- 0.05258214
+
+# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
+```
+
+```{r calc_sd_tc_mims_air_hpt_2}
+# standard deviation of a measurement is:
+round(sd(S_a_hpt_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_a_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S + S_a_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+#### Map source-sink status
+
+```{r map_mims_air_hpt, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_a_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_mims_air_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+### Pure gas, high precision thermometer
+Simulate uncertainty in the saturation ratio due to measurement error associated
+with the standard GC, a pure gas (helium) as the headspace gas, and using a high 
+precision thermometer (e.g., MIMS).
+
+```{r simulate_me_ratio_all_pg_hpt, message=F}
+S_pg_hpt_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+S_pg_hpt_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_me[i] - 0) / (8.3144598 * (T_c_mims_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)) * M_h_pg_me[i])) /
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) 
 }
@@
 # was deprecated in ggplot 3.4.0
-C_sat_df <- tibble(C_me_sat = C_me_sat)
-
-ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
-  geom_histogram( binwidth = 0.01 ) + 
-  geom_vline( xintercept = C_sat , color = 'red' ) + 
-  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
-  xlab( "concentration (nMol)")
-
-ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
-  geom_histogram( binwidth = 0.001 ) + 
+S_pg_hpt_df <- tibble(S_pg_hpt_me = S_pg_hpt_me)
+# save absolute error for plot
+S_pg_hpt_me_abs <- S_pg_hpt_me - S_pg
+```
+
+```{r plot_me_ratio_all_pg_hpt, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_pg_hpt_df, aes(x = S_pg_hpt_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S_pg , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_pg_hpt_df, aes(x = S_pg_hpt_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
   geom_vline( xintercept = 0 , color = 'red' ) +
-  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
-  xlab( "absolute error (nMol)")
-
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_sd_me_ratio_all_pg_hpt}
+# standard deviation of a saturation ratio value s:
+sd(S_pg_hpt_me) # 0.026
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_pg_hpt_me) * 1.96 # +/- 0.05258214
+
+# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
+```
+
+```{r calc_sd_me_ratio_all_pg_hpt_2}
 # standard deviation of a measurement is:
-sd(C_me_sat* 1e9) # 0.20 nmol L-1
+round(sd(S_pg_hpt_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_pg_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_pg + S_pg_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+#### Map source-sink status
+
+```{r map_sat_ratio_pg_hpt, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_pg_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_sat_ratio_pg_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+## High precision GC
+
+The GC used in the current study is optimized to quantify several
+analytes across a broad range of concentrations but greater precision
+can be achieved by optimizing the instrument for a single analyte across
+a narrower range of concentrations. For example, laboratories
+participating in an international analysis of N2O standards used for
+atmsopheric monitoring reported a mean standard deviation of
+`r mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015))`, which is
+10-fold lower than the standard deviation of N2O measurements in this
+study (`r n2o_sd_epa`). Utilizing more precise N2O measurements for the
+equilibrated headspace and/or air will reduce uncertainty in the
+estimated dissolved N2O concentration, dissolved N2O equilibrium value,
+and N2O saturation ratio.
+
+### Pure gas, standard thermometer
+```{r simulate_me_ratio_gc_pure_gas_all, message=F}
+# set M_h to a lower value to generate a saturation ratio close to that used above
+# M_h will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
+M_h_pg_hpgc <- 0.2
+
+# two standard deviations reported from each of three labs participating in
+# a laboratory intercomparison study
+n2o_sd_lit <- mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015)) # from literature
+
+# simulate improved measurement of N2O in equilibrated headspace gas 
+# assume TRUE value is C_e and error is simulated with with sd from literature
+M_h_pg_hpgc_me <- rnorm(nsim, M_h_pg_hpgc, n2o_sd_lit)
+
+# under this scenario we don't need to subtract C_a_n2o (N2O concentration in air) from C_e (N2O concentration in equilibrated headspace)
+S_pg_hpgc <- (1e-6 * B * (V_g * (M_h_pg_hpgc - 0) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15) ) * M_h_pg_hpgc)) /
+  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))
+
+S_pg_hpgc_me <- rep(NA, 1e4)
+
+for(i in 1:nsim){
+S_pg_hpgc_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_hpgc_me[i] - 0) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15) ) * M_h_pg_hpgc_me[i])) /
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1 / (T_c_me[i] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+S_pg_hpgc_df <- tibble(S_pg_hpgc_me = S_pg_hpgc_me)
+# save absolute error for plot
+S_pg_hpgc_me_abs <- S_pg_hpgc_me - S_pg_hpgc
+```
+
+```{r plot_me_ratio_gc_pure_gas_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_pg_hpgc_df, aes(x = S_pg_hpgc_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_pg_hpgc_df, aes(x = S_pg_hpgc_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_sd_ratio_gc_pure_gas_all}
+# standard deviation of a saturation ratio value s:
+sd(S_pg_hpgc_me) # 0.029
 
 # the 95% CI is therefore the computed value +/- 2 standard deviations
-sd(C_me_sat * 1e9) * 1.96 # +/- 0.395 nmol N2O
-```
+sd(S_pg_hpgc_me) * 1.96 # +/- 0.0569
+
+# so any ratio within 0.0569 of 1.0 cannot be discerned as a source or sink
+```
+
+```{r calc_sd_me_ratio_gc_pure_gas_all_2}
+# standard deviation of a measurement is:
+round(sd(S_pg_hpgc_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_pg_hpgc_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_pg_hpgc + S_pg_hpgc_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+#### Map source-sink status
+
+```{r map_sat_ratio_gc_pure_gas, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_pg_hpgc_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_sat_ratiogc_pure_gas, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+### Pure gas, high precision thermometer
+
+```{r simulate_me_ratio_hpgc_pg_hpt_all, message=F}
+# set M_h to a lower value to generate a saturation ratio close to that used above
+# M_h will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
+M_h_pg_hpgc_hpt <- 0.2
+
+# two standard deviations reported from each of three labs participating in
+# a laboratory intercomparison study
+n2o_sd_lit <- mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015)) # from literature
+
+# simulate improved measurement of N2O in equilibrated headspace gas 
+# assume TRUE value is C_e and error is simulated with with sd from literature
+M_h_pg_hpgc_hpt_me <- rnorm(nsim, M_h_pg_hpgc, n2o_sd_lit)
+
+S_pg_hpgc_hpt_me <- rep(NA, 1e4)
+
+for(i in 1:nsim){
+S_pg_hpgc_hpt_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_hpgc_hpt_me[i] - 0) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15) ) * M_h_pg_hpgc_hpt_me[i])) /
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1 / (T_c_me[i] + 273.15) - 1/298.15)))) 
+}
+
+# combine into dataframe for ggplot. qplot, which accepts vectors,
+# was deprecated in ggplot 3.4.0
+S_pg_hpgc_hpt_df <- tibble(S_pg_hpgc_hpt_me = S_pg_hpgc_hpt_me)
+# save absolute error for plot
+S_pg_hpgc_hpt_me_abs <- S_pg_hpgc_hpt_me - S_pg_hpgc
+```
+
+```{r plot_me_ratio_hpgc_pg_hpt_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_pg_hpgc_hpt_df, aes(x = S_pg_hpgc_hpt_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S_pg, color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_pg_hpgc_hpt_df, aes(x = S_pg_hpgc_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.005) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_sd_ratio_hpgc_pg_hpt_all}
+# standard deviation of a saturation ratio value s:
+sd(S_pg_hpgc_hpt_me) # 0.029
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_pg_hpgc_hpt_me) * 1.96 # +/- 0.0569
+
+# so any ratio within 0.0569 of 1.0 cannot be discerned as a source or sink
+```
+
+```{r calc_sd_me_ratio_hpgc_pg_hpt_all_2}
+# standard deviation of a measurement is:
+round(sd(S_pg_hpgc_hpt_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_pg_hpgc_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_pg_hpgc + S_pg_hpgc_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+#### Map source-sink status
+
+```{r map_ratio_hpgc_pg_hpt, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_pg_hpgc_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_ratio_hpgc_pg_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+# MIMS
+
+In Membrane Inlet Mass Spectrometry (MIMS), dissolved gases (typically
+N2, Ar, O2, N2O) are extracted from a water sample as it passes through
+a semi-permeable membrane. The abundance of the extracted gases are then
+measured using a quadrapole mass spectrometer. One advantage of this
+approach relative to the traditional headspace equilibrium method is
+that dissolved gases are extracted without contamination and/or dilution
+from a headspace gas (e.g. air, nitrogen, helium). Dilution by a pure
+gas isn't much of a concern because N2O detection limits are very low
+via GC, but correction for the presence of N2O in the headspace gas
+reduces measurement precision.
+
+Dissolved N2O concentration in a MIMS sample can be calculated from the
+N2O signal strength and a calibration factor. The calibration factor is
+determined by analyzing a water sample with known N2O concentration.
+This is most frequently accomplished by equilibrating laboratory water
+with air at constant temperature, barometric pressure and N2O partial
+pressure. The dissolved N2O concentration of the equilibrated water is
+calculated from solubility laws (see [Equilibrium concentration]) and the 
+calibration factor, $f$, is:
+
+$$f_{N_2O} = \frac{C^{N_2O}_s}{X^{N_2O}_s}$$
+where:
+
+-   $f_{N_2O} = \text{N}_2\text{O} \text{ calibration factor}$
+-   $C^{N_2O}_s = \text{Concentration of } \text{N}_2\text{O} \text{ in calibration standard}$
+-   $X^{N_2O}_s = \text{Signal strength of } \text{N}_2\text{O} \text{ in calibration standard}$
+
+The N2O concentration of an unknown sample is then calculated as:
+
+$$C^{N_2O}_{\theta} = \frac{X^{N_2O}}{f_{N_2O}}$$
+
+where:
+
+-   $C^{N_2O}_{\theta} = \text{N}_2\text{O} \text{ concentration in sample}$
+-   $X^{N_2O}_{\theta} = \text{Signal strength of } \text{N}_2\text{O} \text{ in sample}$ 
+-   $f_{N_2O} = \text{N}_2\text{O} \text{ calibration factor}$
+
+The equation simply scales the concentration of the N2O standard to the
+ratio of the MIMS signal strength of the unknown and standard. Via
+substitution and rearrangement, the equation can also be written as:
+
+$$C^{N_2O}_{\theta} =\frac{X^{N_2O}_{\theta}}{X^{N_2O}_s} \cdot C^{N_2O}_s$$
+
+MIMS provides greater precision when measuring gas ratios as compared to
+measuring a single gas. High precision gas ratio data can be converted
+to concentration data with the independent measurement of concentration
+of one of the component gases. N2O:Ar ratios are particularly useful for
+environmental samples because Ar is biologically inert, therefore its
+concentration can be calculated from solubility, water temperature,
+barometric pressure, and atmospheric partial pressure (essentially
+constant across the globe). As with single gas measurements, N2O:Ar
+ratios are converted to N2O concentration using a calibration factor
+derived from measurement of laboratory water equilibrated with air:
+
+$$\begin{aligned}
+f^\prime_{N_2O} =\frac{R^{C}_s}{R^{X}_s} \\
+R^{C}_s = \frac{C^{N_2O}_s}{C^{Ar}_s} \\
+R^{X}_s = \frac{X^{N_2O}_s}{X^{Ar}_s}
+\end{aligned}$$
+
+where:
+
+-   $f^\prime_{N_2O} = \text{N}_2\text{O} \text{ (ratio) calibration factor}$
+-   $R^{C}_s = \text{Concetration ratio of } \text{N}_2\text{O} \text{ to Ar in calibration standard}$
+-   $R^{X}_s = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in calibration standard}$
+
+The N2O concentration of an unknown sample is then calculated as:
+
+$$C^{N_2O}_{\theta} = R^{X}_s \cdot f^\prime_{N_2O} \cdot C^{Ar}_s$$
+where:
+
+-   $C^{N_2O}_{\theta} = \text{Concentration of } \text{N}_2\text{O} \text{ in the sample}$
+-   $C^{Ar}_s = \text{Concentration of } Ar \text{ in the calibration standard}$
+
+As before, the equation above scales the N2O:Ar ratio measured in the
+sample to that of the standard. Via substitution and rearrangement, the
+equation can be written as:
+$$C^{N_2O}_{\theta} = \frac{R^{X}_{\theta}}{R^{X}_s} \cdot R^{C}_s \cdot C^{Ar}_s$$
+where:
+
+-   $R^{X}_\theta = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in the sample}$
+
+Concentration of N2O and Ar in the standard, $C^{N_2O}_s$ is calculated from
+temperature of the equilibrated water, $Tc$, laboratory
+barometric pressure, $B$, and N2O and Ar partial pressure in the atmosphere, 
+$C^{N_2O}_a$, $C^{Ar}_a$, respectively (see [Equilibrium concentration]). By 
+substitution, the equation to estimate the concentration of N2O in the unknown
+sample becomes:
+
+$$\begin{align}
+C^{N_2O}_{\theta} = \frac{R^{X}_{\theta}}{R^{X}_s} \times \\ \frac{ 1^{-6} \cdot B \cdot C^{N_2O}_a \cdot (H_{N_2O} \cdot e^{2700 \cdot \frac{1}{Tc} -  \frac{1}{298.15}})}{1^{-6} \cdot B \cdot C^{Ar}_a \cdot (H_{Ar} \cdot e^{1500 \cdot \frac{1}{Tc} -  \frac{1}{298.15}})} \times \\ 1^{-6}B \cdot C^{Ar}_a \cdot (H_{Ar} \cdot e^{1500 \cdot  \frac{1}{Tc} -  \frac{1}{298.15}})
+\end{align}$$
+
+where (repeating some variables from above for clarity):
+
+-   $C^{N_2O}_{\theta} = \text{concentration of } \text{N}_2\text{O} \text{ in the sample}$
+-   $R^{X}_\theta = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in the sample}$
+-   $R^{X}_s = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in calibration standard}$
+-   $B = \text{barometric pressure (kPa)}$
+-   $Tc = \text{temperature}$
+-   $C^{N_2O}_a = \text{concentration of } \text{N}_2\text{O} \text{ in the atmosphere}$
+-   $H_{N_2O} = \text{Henry's law constant for } \text{N}_2\text{O}$
+-   $C^{Ar}_a = \text{concentration of } Ar \text{ in the atmosphere}$
+-   $H_{Ar} = \text{Henry's law constant for } Ar$
+-   $Tc = \text{Temperature of the equilibrated water}$
+
+N2O concentration in laboratory air, ($C^{N_2O}_a$), is derived from direct
+measurement of N2O via gas chromatography and has a standard deviation
+of `r n2o_sd_epa` using the method employed in this study (see "Empirical 
+variation" under [Dissolved concentration]) although more precise methods are
+available (see [Dissolved $\text{N}_2\text{O}$: high precision GC]). Ar in the lab 
+atmosphere ($C^{Ar}_a$) is 9340ppm and is assumed to be constant and perfectly 
+known (e.g. zero error). Genther et al (2013) report a standard deviation of
+0.022 for the N2O:Ar signal strength ratio but Speir et al (2023) report a CV 
+of 0.05% (sd \~ 0.000096). Here we provisionally use the Speir value pending 
+confirmation. MIMS uses a very precise thermometer (+/-0.01 C) to control the 
+temperature of the water used as the analytical standard (simulated below).
+Measurement error in $B$ was quantified above.
+
+## Dissolved concentration
+
+### Standard GC
+
+```{r simulate_me_hpt_MIMS, message=F}
+#n2o_ar_standard_signal <- 0.27 # Speir et al. 0.27 pers comm. Genther et al. section 3.4 reports 0.192
+#n2o_ar_sample_signal <- 0.23 # just a guess
+#C_a_ar <- 9340 # ppm.  invariant
+#T_c_mims <- T_c # set up for simulation of MIMS temperature bath (from GC simulations above)
+
+R_X_s <- 0.27 # Speir et al. 0.27 pers comm. Genther et al. section 3.4 reports 0.192 (signal ratio of N2O:Ar in standard)
+R_X_theta <- 0.23 # just a guess (signal ratio of N2O:Ar in sample)
+C_Ar_a <- 9340 # ppm.  invariant (Concentration of Ar in atmospher)
+# T_c_mims <- T_c # set up for simulation of MIMS temperature bath (from GC simulations above)
+
+# this simulates MIMS measurement.
+# assume TRUE value is n2o_ar_standard_signal and error is simulated with CV from Speier et al. 2023
+# Speier et al reports "...our precision was good among all sample and standard replicated (average CV = 0.05%)."
+# Using that CV to model measurement error in the N2O:Ar ratio yields a sd of dissolved N2O that exceeds
+# the sd when measured via GC, which is weird.  Speier reported that the CV is for the N2O:Ar ratio.
+# Genther et al reports a n2o:Ar sd of 0.022, but this also generated a sd for dissolved N2O which exceeds that 
+# measured with GC!  Here we use the Speir data
+R_X_s_sd <- 0.0005 * R_X_s # sd = CV * mean. 0.05% = 0.0005
+R_X_theta_sd <- 0.0005 * R_X_theta # sd = CV * mean. 0.05% = 0.0005
+
+R_X_s_me <- rnorm(nsim, R_X_s, R_X_s_sd) # me in N2O:Ar of standard  
+R_X_theta_me <- rnorm(nsim, R_X_theta, R_X_theta_sd) # me in N2O:Ar of sample
+
+# High precision water batch used to control temperature of standards
+T_c_mims_me <- rnorm(nsim, T_c, 0.005) # thermo stated bath precision is +/-0.01 Celsius
+# hist(T_c_mims_me) # sd of 0.005 gives approx distribution for the specified precision
+
+# True dissolved N2O concentration
+C_mims <- ((R_X_theta / R_X_s) * # N2O:Ar ratios for sample and standard
+  ((1e-6 * B * M_a_n2o * (H_n2o * exp( 2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B * C_Ar_a * (H_ar * exp( 1500 * (1 / (T_c + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
+  (1e-6 * B * C_Ar_a*(H_ar  * exp(1500 *(1/(T_c + 273.15) - 1/298.15)))))
+
+C_mims <- C_mims * 1e9 #convert mol to nmol
+
+# variation in measured N2O concentration
+C_mims_hpt_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+C_mims_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
+  ((1e-6 * B_me[i] * M_a_n2o_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) 
+}
+
+# combine into dataframe for ggplot.
+C_mims_hpt_df <- tibble(C_mims_hpt_me = C_mims_hpt_me * 1e9)
+# save absolute error for plot
+C_mims_hpt_me_abs <- (C_mims_hpt_me * 1e9) - C_mims
+```
+
+```{r plot_me_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(C_mims_hpt_df, aes(x = C_mims_hpt_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C_mims , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in dissolved N2O \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(C_mims_hpt_df, aes(x = C_mims_hpt_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.01) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_me_hpt_MIMS}
+# standard deviation of dissolved N2O via MIMS:
+sd(C_mims_hpt_me * 1e9) # 0.169 using Speir CV to calculate N2O:Ar sd.   0.15 with GC and headspace equilibrium!
+
+# 1.96 x sd
+sd(C_mims_hpt_me * 1e9) * 1.96
+
+# CV of dissolved N2O
+(sd(C_mims_hpt_me * 1e9) / mean(C_mims_hpt_me * 1e9)) * 100 # 0.025 = 2.5% vs 12% using Genther sd.
+```
+
+```{r calc_me_hpt_MIMS_2}
+# standard deviation of a measurement is:
+round(sd(C_mims_hpt_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(C_mims_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((C_mims + C_mims_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+### High precision GC
+
+```{r hpgc_M_a_MIMS}
+# simulate improved measurement precision of GC measurement of N2O in air
+M_a_n2o_lit_me <- rnorm(nsim, M_a_n2o, n2o_sd_lit) # from intn'l labs = 0.00073 vs epa sd = 0.00787
+```
+
+```{r simulate_me_hpgc_hpt_MIMS, message=F}
+# variation in measured N2O concentration
+C_mims_hpgc_hpt_me <- rep(NA, nsim)
+
+for(i in 1:nsim){
+C_mims_hpgc_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
+  ((1e-6 * B_me[i] * M_a_n2o_lit_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) 
+}
+
+# combine into dataframe for ggplot.
+C_mims_hpgc_hpt_df <- tibble(C_mims_hpgc_hpt_me = C_mims_hpgc_hpt_me * 1e9)
+# save absolute error for plot
+C_mims_hpgc_hpt_me_abs <- (C_mims_hpgc_hpt_me * 1e9) - C_mims
+```
+
+```{r plot_me_hpgc_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(C_mims_hpgc_hpt_df, aes(x = C_mims_hpgc_hpt_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C_mims , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in dissolved N2O \nestimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(C_mims_hpgc_hpt_df, aes(x = C_mims_hpgc_hpt_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_me_hpgc_hpt_MIMS}
+# standard deviation of dissolved N2O via MIMS:
+sd(C_mims_hpgc_hpt_me * 1e9) # 0.0166 using Speir CV to calculate N2O:Ar sd. 0.15 with GC and headspace equilibrium!
+
+sd(C_mims_hpgc_hpt_me * 1e9) * 1.96 # sd +/- 1.96 (for conf intervals on error)
+
+# CV of dissolved N2O
+sd(C_mims_hpgc_hpt_me * 1e9) / mean(C_mims_hpgc_hpt_me * 1e9) * 100 # 0.0025 = 0.25% CV 
+```
+
+```{r calc_me_hpgc_hpt_MIMS_2}
+# standard deviation of a measurement is:
+round(sd(C_mims_hpgc_hpt_me_abs), 3)
+
+# central 95th percentile of absolute error
+round(quantile(C_mims_hpgc_hpt_me_abs, probs = c(0.025, 0.975)), 3)
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((C_mims + C_mims_hpgc_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+## Saturation ratio
+
+Calculate the "true" ratio via MIMS
+```{r true_ratio_MIMS, message=F}
+# True ratio
+S_mims <- ((R_X_theta / R_X_s) * # N2O:Ar ratios for sample and standard
+  ( (1e-6 * B * M_a_n2o * (H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c + 273.15) - 1 / 298.15)) ))) * #Ar concentration in standard
+  (1e-6 * B * C_Ar_a *(H_ar  * exp(1500 * (1 / (T_c + 273.15) - 1 / 298.15))))) / # Ar concentration in standard
+  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) # N2o saturation value in waterbody 
+```
+
+### Standard GC
+
+#### Standard field temperature
+This simulation is for field measurment of temperature for the equilibrium concentration with the standard glass thermometer.
+```{r simulate_me_ratio_MIMS, message=F}
+# ratio simulated with measurement error in mims
+S_mims_me <- rep(NA, nsim)
+
+for( i in 1:nsim){
+S_mims_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
+  ((1e-6 * B_me[i] * M_a_n2o_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using regular field temp
+}
+
+# combine into dataframe for ggplot.
+S_mims_df <- tibble(S_mims_me = S_mims_me)
+# save absolute error for plot
+S_mims_me_abs <- S_mims_me - S_mims
+```
+
+```{r plot_me_ratio_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_mims_df, aes(x = S_mims_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S_mims , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_mims_df, aes(x = S_mims_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_me_ratio_MIMS}
+# standard deviation of a saturation ratio value s:
+sd(S_mims_me) # 0.0011
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_mims_me) * 1.96 # +/- 0.00216
+
+# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink
+
+# CV of sat
+sd(S_mims_me) / mean(S_mims_me) * 100 # 0.13% CV 
+```
+
+```{r calc_me_ratio_MIMS_2}
+# standard deviation of a measurement is:
+round(sd(S_mims_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_mims_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_mims + S_mims_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O 
+```
+
+##### Map source-sink status
+
+```{r map_me_ratio_MIMS, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_mims_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_me_ratio_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+#### High precision temperature
+
+```{r simulate_me_ratio_hpt_MIMS, message=F}
+# ratio simulated with measurement error in mims
+S_mims_hpt_me <- rep(NA, nsim)
+
+for( i in 1:nsim){
+S_mims_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
+  ((1e-6 * B_me[i] * M_a_n2o_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
+  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using regular field temp
+}
+
+# combine into dataframe for ggplot.
+S_mims_hpt_df <- tibble(S_mims_hpt_me = S_mims_hpt_me)
+# save absolute error for plot
+S_mims_hpt_me_abs <- S_mims_hpt_me - S_mims
+```
+
+```{r plot_me_ratio_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_mims_hpt_df, aes(x = S_mims_hpt_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S_mims , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_mims_hpt_df, aes(x = S_mims_hpt_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_me_ratio_hpt_MIMS}
+# standard deviation of a saturation ratio value s:
+sd(S_mims_hpt_me) # 0.0011
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_mims_hpt_me) * 1.96 # +/- 0.00216
+
+# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink
+
+# CV of sat
+sd(S_mims_hpt_me) / mean(S_mims_hpt_me) * 100 # 0.13% CV 
+```
+
+```{r calc_me_ratio_hpt_MIMS_2}
+# standard deviation of a measurement is:
+round(sd(S_mims_hpt_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_mims_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_mims + S_mims_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O 
+```
+
+##### Map source-sink status
+
+```{r map_me_ratio_hpt_MIMS, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_mims_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_me_ratio_hpt_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+### High precision GC
+
+#### Standard temperature
+
+This simulation is for field measurment of temperature for the equilibrium concentration with the standard glass thermometer.
+```{r simulate_me_ratio_hpgc_MIMS, message=F}
+S_mims_hpgc_me <- rep(NA, nsim)
+
+for( i in 1:nsim){
+S_mims_hpgc_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
+  ((1e-6 * B_me[i] * M_a_n2o_lit_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
+  (M_a_n2o_lit_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using regular field temp
+}
+
+# combine into dataframe for ggplot.
+S_mims_hpgc_df <- tibble(S_mims_hpgc_me = S_mims_hpgc_me)
+# save absolute error for plot
+S_mims_hpgc_me_abs <- S_mims_hpgc_me - S_mims
+```
+
+```{r plot_me_ratio_hpgc_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_mims_hpgc_df, aes(x = S_mims_hpgc_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S_mims , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_mims_hpgc_df, aes(x = S_mims_hpgc_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_me_ratio_hpgc_MIMS}
+# standard deviation of a saturation ratio value s:
+sd(S_mims_hpgc_me) # 0.0011
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_mims_hpgc_me) * 1.96 # +/- 0.00216
+
+# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink
+
+# CV of sat
+sd(S_mims_hpgc_me) / mean(S_mims_hpgc_me) * 100 # 0.07% CV 
+```
+
+```{r calc_me_ratio_hpgc_MIMS_2}
+# standard deviation of a measurement is:
+round(sd(S_mims_hpgc_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_mims_hpgc_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_mims + S_mims_hpgc_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+##### Map source-sink status
+
+```{r map_me_ratio_hpgc_MIMS, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_mims_hpgc_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_me_ratio_hpgc_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+#### High precision temperature
+
+The thermometers used to measure water temperature in this study have an
+uncertainty of +/- 0.058C but electronic thermisters are commonly
+available (e.g. YSI sondes) with an uncertainty of +/-0.01 Celsius. More
+precise water temperature measurements will reduce uncertainty in the
+equilibrium N2O conentration and the saturation ratio.
+
+```{r simulate_me_ratio_hpgc_hpt_MIMS, message=F}
+S_mims_hpgc_hpt_me <- rep(NA, nsim)
+
+for( i in 1:nsim){
+S_mims_hpgc_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
+  ((1e-6 * B_me[i] * M_a_n2o_lit_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
+  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
+  (M_a_n2o_lit_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using high precision field temp
+}
+
+# combine into dataframe for ggplot.
+S_mims_hpgc_hpt_df <- tibble(S_mims_hpgc_hpt_me = S_mims_hpgc_hpt_me)
+
+# save absolute error for plot
+S_mims_hpgc_hpt_me_abs <- S_mims_hpgc_hpt_me - S_mims
+```
+
+```{r plot_me_ratio_hpgc_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
+ggarrange(
+ggplot(S_mims_hpgc_hpt_df, aes(x = S_mims_hpgc_hpt_me, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = S_mims , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
+  xlab( "Concentration (nMol)"),
+
+ggplot(S_mims_hpgc_hpt_df, aes(x = S_mims_hpgc_hpt_me_abs, y = after_stat(density))) +  
+  geom_histogram( binwidth = 0.0001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline( xintercept = 0 , color = 'red' ) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) + 
+  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  xlab( "Error (nMol)"),
+nrow = 1
+)
+```
+
+```{r calc_me_ratio_hpgc_hpt_MIMS}
+# standard deviation of a saturation ratio value s:
+sd(S_mims_hpgc_hpt_me) # 0.0006603624
+
+# the 95% CI is therefore the computed value +/- 2 standard deviations
+sd(S_mims_hpgc_hpt_me) * 1.96 # +/- 0.00129431
+
+# so any ratio within 0.00129 of 1.0 cannot be discerned as a source or sink
+```
+
+```{r calc_me_ratio_hpgc_hpt_MIMS_2}
+# standard deviation of a measurement is:
+round(sd(S_mims_hpgc_hpt_me_abs), 3) # 0.20 nmol L-1
+
+# central 95th percentile of absolute error
+round(quantile(S_mims_hpgc_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11
+
+# the 95% CI for an observation with measurement error can be computed 
+# approximately as the value +/- 2 standard deviations; or the central 95th 
+# percentile of the absolute error distribution.
+round(quantile((S_mims + S_mims_hpgc_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
+```
+
+##### Map source-sink status
+
+```{r map_me_ratio_hpgc_hpt_MIMS, message=F}
+# two standard deviations = 0.3019449
+# 95% CI is therefore x+/-0.3019449
+
+#dg <- dg %>%
+#  mutate(n2o.src.snk.with.all.error = case_when(
+#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
+#    TRUE ~ n2o.src.snk)) 
+dg <- dg %>%
+  mutate(n2o.src.snk.with.all.error = case_when(
+    abs(n2o.sat.ratio - 1) < quantile(S_mims_hpgc_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
+    TRUE ~ n2o.src.snk))
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {table(.$n2o.src.snk.with.all.error)} %>%
+  kable(digits = 2, caption = "Number of lakes classified as source/sink")
+
+dg %>% 
+  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
+  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
+  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
+
+# sf for plotting
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+```
+
+```{r plot_map_me_ratio_hpgc_hpt_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
+## load ecoregions----
+# read in ecoregion polygons
+ecoR <- st_read(dsn = paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
+                layer = "aggr_ecoregions_simple",
+                quiet = TRUE)
+
+# Check CRS
+#st_crs(ecoR) # 3857
+ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
+#st_crs(ecoR) # 5070
+
+# SET UP CUSTOM COLORS FOR ECOREGIONS
+# Custom color pallette for ecoregion polygons. Attempted to mirror
+# https://www.epa.gov/national-aquatic-resource-surveys/
+# ecoregional-results-national-lakes-assessment-2012
+cols <- c("Coastal Plains" = "orange1",
+          "Northern Appalachians" = "lightpink1",
+          "Northern Plains" = "darksalmon",
+          "Southern Appalachians" = "mediumturquoise",
+          "Southern Plains" = "khaki4",
+          "Temperate Plains" = "forestgreen", 
+          "Upper Midwest" = "deepskyblue4",
+          "Western Mountains" = "saddlebrown",
+          "Xeric" = "lightskyblue4")
+
+## Load state boundaries----
+states <- USAboundaries::us_states() %>%
+  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  st_transform(5070) # convert to CONUS Albers
+
+# To enable spatial analysis of the data, the dataframe will be converted to a 
+# 'simple features' (sf) object.
+# Define coordinates
+coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
+
+dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
+                  crs = 4269) %>% # standard for lat/lon
+  st_transform(5070) # project to CONUS ALBERS for plotting
+
+## Map
+ggplot() +
+    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
+    geom_sf(data = dg.sf %>% 
+              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
+            shape = 1,
+            show.legend = "point") +
+    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
+    # guide argument below removes points from the boxes in the ecoregion legend
+    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
+    scale_fill_manual("Ecoregion", 
+                      values = cols,
+                      guide = guide_legend(override.aes = list(shape = NA))) +
+    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
+    scale_size(name = expression(N[2]*O~(nM)),
+               range = c(0.5, 8), # custom size range
+               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
+    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
+          legend.title = element_text(size = 8))
+```
+
+# Summary of ratio results
+
+Precision of the N2O saturation ratio can be improved using high
+precision GC analytics and N2O-free headspace gas (e.g. helium,
+nitrogen) but the greatest improvement is seen when dissolved N2O is
+directly measured using MIMS.
+
+```{r final_summary, message=FALSE}
+#results <- tibble(ratio_sd = c(0.055,
+#                               0.048,
+#                               0.029,
+#                               0.001207747,
+#                               0.001207747,
+#                               0.000690139),
+#                  method = c("GC",
+#                             "GC + helium",
+#                             "High precision GC + helium",
+#                             "MIMS",
+#                             "MIMS + high precision GC",
+#                             "MIMS + high precision GC + \nhigh precision water temp"))
+
+#results %>%
+#  mutate(lower.95 = C_ratio - (1.96 * ratio_sd),
+#         upper.95 = C_ratio + (1.96 * ratio_sd),
+#         C_ratio = C_ratio) %>%
+#  ggplot(aes(C_ratio, method)) +
+#  geom_point() +
+#  geom_segment(aes(x=lower.95, xend = upper.95, y=method, yend=method)) +
+#  xlab("True saturation ratio (black dot) +/- measurement error (1.96 standard deviations))") +
+#  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 1))
+
+results <- tibble(GC_me= S_a_me_abs,
+                  GC_He_me = S_pg_me_abs,
+                  GC_hpt_me = S_a_hpt_me_abs,
+                  GC_He_hpt_me = S_pg_hpt_me_abs,
+                  HPGC_He_me = S_pg_hpgc_me_abs,
+                  HPGC_He_hpt_me = S_pg_hpgc_hpt_me_abs,
+                  MIMS_me = S_mims_me_abs,
+                  MIMS_hpt_me = S_mims_hpt_me_abs,
+                  MIMS_HPGC_me = S_mims_hpgc_me_abs,
+                  MIMS_HPGC_HPT_me = S_mims_hpgc_hpt_me_abs) %>%
+  pivot_longer(cols = ends_with("me"), names_to = "Method", values_to = "Error")
+
+results %>%
+  mutate(Method = factor(Method)) %>%
+  mutate(Method = fct_relevel(Method, c("MIMS_HPGC_HPT_me",
+                                        "MIMS_HPGC_me",
+                                        "MIMS_hpt_me",
+                                        "MIMS_me",
+                                        "HPGC_He_hpt_me",
+                                        "HPGC_He_me",
+                                        "GC_He_hpt_me",
+                                        "GC_hpt_me",
+                                        "GC_He_me",
+                                        "GC_me"
+                                        ))) %>%
+  ggplot(aes(x = Error, y = Method)) +
+  stat_pointinterval() +
+  scale_y_discrete(labels=c("GC_me" = "Standard GC, ambient air, \n standard field temperature",
+                            "GC_He_me" = "Standard GC, helium, \n standard field temperature",
+                            "GC_hpt_me" = "Standard GC, ambient air, \n high precision temperature",
+                            "GC_He_hpt_me" = "Standard GC, helium, \nhigh precision temperature",
+                            "HPGC_He_me" = "High precision GC, helium, \nstandard field temperature",
+                            "HPGC_He_hpt_me" = "High precision GC, helium, \nhigh precision temperature",
+                            "MIMS_me" = "MIMS: standard GC, \nstandard field temperature",
+                            "MIMS_hpt_me" = "MIMS: standard GC, \nhigh precision temperature",
+                            "MIMS_HPGC_me" = "MIMS: high precision GC, \nstandard field temperature",
+                            "MIMS_HPGC_HPT_me" = "MIMS: high precision GC, \nhigh precision temperature")) +
+  xlab("Absolute error in saturation ratio") +
+  ylab("")
+```
+
+Because the MIMS method is so much more precise than the traditional method, 
+the differences between the MIMS-based variations above are difficult to see. 
+Below, just the MIMS-based variations are compared.
+
+Precision of the N2O saturation ratio can be improved using high
+precision GC analytics and N2O-free headspace gas (e.g. helium,
+nitrogen) but the greatest improvement is seen when dissolved N2O is
+directly measured using MIMS.
+
+```{r final_summary_MIMS, message=FALSE}
+results_MIMS <- tibble(MIMS_me = S_mims_me_abs,
+                       MIMS_hpt_me = S_mims_hpt_me_abs,
+                       MIMS_HPGC_me = S_mims_hpgc_me_abs,
+                       MIMS_HPGC_HPT_me = S_mims_hpgc_hpt_me_abs) %>%
+  pivot_longer(cols = ends_with("me"), names_to = "Method", values_to = "Error")
+
+results_MIMS %>%
+  mutate(Method = factor(Method)) %>%
+  mutate(Method = fct_relevel(Method, c("MIMS_HPGC_HPT_me",
+                                        "MIMS_HPGC_me",
+                                        "MIMS_hpt_me",
+                                        "MIMS_me"
+                                        ))) %>%
+  ggplot(aes(x = Error, y = Method)) +
+  stat_pointinterval() +
+  scale_y_discrete(labels=c("MIMS_me" = "MIMS: standard GC, \nstandard field temperature",
+                            "MIMS_hpt_me" = "MIMS: standard GC, \nhigh precision temperature",
+                            "MIMS_HPGC_me" = "MIMS: high precision GC, \nstandard field temperature",
+                            "MIMS_HPGC_HPT_me" = "MIMS: high precision GC, \nhigh precision temperature")) +
+  xlab("Absolute error in saturation ratio") +
+  ylab("")
+```

### 2024-06-18 scripts/DG_sensitivity_to_measurement_error.Rmd
Added mixing scenarios. Added .md file to repo

@@
 ```{r setup, warning=FALSE, message=FALSE}
+#library(devtools) # needed to install USAboundaries package from
 library(knitr)
@@
 
+# Install USAboundaries package from GitHub
+#devtools::install_github("ropensci/USAboundaries")
+
 # Identify local path for each user
@@
 
--   $C$ = concentration (nM)
+-   $C$ = concentration (nM) dissolved in water
 -   $B$ = barometric pressure (kPa)
 -   $V_g$ = volume of reference gas in headspace (mL)
--   $M_h$ = gas mixture ratio in the headspace (C_e below)
--   $M_r$ = gas mixture ratio in the reference (C_a_n2o below)
+-   $M_h$ = gas measured in the equilibrated headspace
+-   $M_r$ = gas measured in the reference
 -   $V_w$ = volume of water below the headspace (mL)
--   $K$ = temperature (K) (temp celsius or T_c + 273 below)
+-   $K$ = temperature (K) (T_c + 273 below)
 -   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ or $H_{Ar}$ below)
@@
 -   $V$ = Total volume of the mixing vessel (e.g. 140mL syringe)
+-   $R_M$ = Gas mixing proportion (gas volume / mixing vessel volume)
 -   $V_g$ = Volume of gas (mL) in the mixing vessel 
 -   $V_w$ = Volume of water (mL) in the mixing vessel
--   $R_M$ = Mixing ratio (gas:water)
-
-## Simulate $\text{N}_2\text{O}$
-
+
+## Calculate $\text{N}_2\text{O}$
+Below is a summary of the empirical measurements of N2O (ppm) in the
+equilibrated headspace.
 ```{r summary_empirical, echo=TRUE}
@@
 
-Simulate a single concentration without measurment error to represent the "true" 
-or "known" value for dissolved $\text{N}_2\text{O}$.
+Calculate a single concentration without measurement error to represent a 
+"true" value for dissolved $\text{N}_2\text{O}$. The median of the empirical 
+measurements was used for values of $M_h$ and $M_r$ (M_a_n2o).
 ```{r dg_calc, echo=TRUE}
@@
 T_c <- 23 # temperature ( C )
-M_h <- 0.307 # gas mixture ratio in headspace (Mr_h) * Changes as function of R_M at constant C
-M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
+M_h <- 0.307 # gas measured in headspace; Changes as function of R_M at constant C
+M_a_n2o <- 0.310 # gas measured in reference (air)
 H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
@@
 
-C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )
+C <- 1e-6 * B * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
 
@@
 
-round( C, 1 ) 
-```
-
-```{r dg_calc_vary_r, echo=TRUE}
-# Set volume of syringe/vessel
-V <- 140 # 140mL syringe
-
-# Set ratio of gas volume to water volume in the mixing vessel
-R_M <- 0.25
-
-B <- 99 # barometric pressure (kPa)
-V_g <- V * R_M # volume of reference gas in headspace
-V_w <- V * (1 - R_M) # volume of water below headspace
-T_c <- 23 # temperature ( C )
-M_h <- 0.307 # gas mixture ratio in headspace (Mr_h)
-M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
-H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
-H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant
-
-C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )
-
-C <- C * 1e9 #convert mol to nmol
-
-round( C, 1 ) 
+round(C, 1)
+
+#### Re-arrange for Mh
+M_h <- (((C / 1e9) * 8.3144598 * (T_c + 273.15) * V_w) + (1e-6 * B * V_g * M_a_n2o)) / (1e-6 * B * (V_g + 8.3144598 * (T_c + 273.15) * V_w * H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15))))
+
+M_h
 ```
@@
 ## Empricial variation
-Estimate variation due to measurement uncertainty using empirical data from
-laboratory replicates.
 ```{r get_gc_error, echo=FALSE}
@@
 
-Measurement error is defined here as the standard deviation of repeated
-analysis of analytical standards throughout the period of time that the
-NLA2017 samples were analyzed, which was $\sigma$ = `r n2o_sd_epa`.
+Measurement error for this analysis is assumed to be the standard deviation of 
+repeated analyses of the analytical standards throughout the period of time that
+the NLA2017 samples were analyzed, which was $\sigma$ = `r n2o_sd_epa`.
 
 ## Simulate measurment error
-
-Below, potential errors associated with dissolved $\text{N}_2\text{O}$ 
-measurement are simulated and assessed separately for GC observations in:
+Below, potential errors in GC measurements for are simulated and assessed s
+eparately for measurements of $\text{N}_2\text{O}$ in:
 
@@
 
-Measurement error is then simulated and assessed separately for:
+Error is then simulated and assessed separately for measurements of:
 
@@
 
-    (5) water volume
-
-Finally, all of the sources (1, 2, 4, 5, and 6) are simulated
-simultaneously and assessed as the total variation in dissolved
-$\text{N}_2\text{O}$. This total error is then added to the point estimates for 
-$\text{N}_2\text{O}$ observed in this study, which results in an estimate of
-uncertainty due to measurement error. These error from these simulations was also
-carried through to estimated of the saturation ratio. The source-sink status for
-each site included in this study was then mapped to present a source-sink 
-categorization that reflected these estimated uncertainties. The categorization 
-was made such that any observation with simulated error that was within the 
-central 95th percentile of
-
-Set the number of simulations for error:
-
+    (6) water volume
+
+Finally, all sources of error (1, 2, 4, 5, and 6) were simulated
+simultaneously and assessed as an estimate of the total variation in $\text{N}_2\text{O}$ 
+measurements. The estimated absolute error was added to each of the empirical 
+point measurements recorded in this study, which resulted in an estimate of 
+uncertainty due to measurement error for each empirical observation. The errors 
+were also carried through for estimates of the saturation ratio and a 
+categorical "source" or "sink" status was assigned for each site. This status 
+was assigned such that any observations where the central 95th percentile of the 
+measurement plus simulated error included 1.0 were deemed to be of 
+"undeterminded" status. The source/sink/undetermined status for each site was 
+then mapped.
+
+Set the number of simulations for estimating error:
 ```{r set_n_sims}
@@
 ### Equilibrated headspace
-Simulate measurment error in dissolved $\text{N}_2\text{O}$ due solely to variation in 
-measurements within the equlibrated headspace.
+Simulate measurement error in dissolved $\text{N}_2\text{O}$ due solely to 
+variation in GC measurements of gas in the equlibrated headspace ($M_h$ above).
 ```{r simulate_C_h_me, echo=TRUE, message=FALSE}
@@
 # The true value for the concentration of equilibrated gas in the headspace, 
-# C_e, was set above. This step simulates additive error in this measurement 
+# M_h, was set above. This step simulates additive error in this measurement 
 # that is centered over the true value. The observed concentration of
-# equilibrated gas is simulated as normally distributed variable, with a mean
-# C_e, and a standard deviation specified as the observed sd of replicate GC 
+# equilibrated gas, M_h_me, is simulated as normally distributed variable, with a mean
+# M_h, and a standard deviation specified as the observed sd of replicate GC 
 # measurements in our lab. 
@@
 
-The results of the simulations are summarized with density histograms
-below.
-
+The results of the simulations are summarized using density histograms below.
 ```{r plot_C_h_me, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
@@
         axis.ticks.y = element_blank()) +
-  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurements") +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurement") +
   xlab("Concentration (nmol L-1)"),
@@
         axis.title.y = element_blank()) +
-  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
   xlab("Error (nMol)"),
@@
 ### Ambient air
-Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) is simulated 
-below. The result of the simulation is plotted as a histogram to illustrate the
-variation in the measurement. THe resulting effect on the uncertainty of the 
-dissolved N2O measurement is also plotted along with the absolute error in 
-N2O as a result of error in this measurement.
+Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air 
+($M_r$ above) reference is simulated below. The result of the simulation is
+plotted as a histogram to illustrate the variation in the measurement. The
+resulting effect on the uncertainty in the simulated dissolved N2O observation 
+is plotted along with the absolute error in the simulated measurement.
 ```{r simulate_me_Ca, message=F}
@@
 
-M_a_n2o_me <- rnorm(nsim, M_a_n2o, n2o_sd_epa) # m.e. in measured gas in air
+M_a_n2o_me <- rnorm(nsim, M_a_n2o, n2o_sd_epa) # m.e. measured gas in air
 
@@
         axis.title.y = element_blank()) +
-  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  ggtitle( "Absolute error in dissolved \nN2O estimate (measured - true) \n") +
   xlab( "Error (nMol)"),
@@
 ### Water temperature
-
+Next, error in water temperature measurements is simulated and its potential
+effect on the final dissolved N2O concentration assessed.
 ```{r simulate_me_T_c, message=F}
@@
         axis.title.y = element_blank()) + 
-  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
   xlab("Error (nMol)"),
@@
 ### Barometric pressure
-
+Next, error in barometric pressure measurements is simulated and its potential
+effect on the final dissolved N2O concentration assessed.
 ```{r simulate_me_B, message=F}
@@
 ### Water volume
-
+Next, error in the volume of water (relative to volume of gas) in the sample 
+vessel is simulated and its potential effect on the final dissolved N2O 
+concentration assessed.
 ```{r simulate_me_V_w, message=F}
@@
 
+### Mixing ratio
+To check the effect of different mixing ratios, we re-arrange the equation to 
+solve for $M_h$, given a known $C$ from above where $C =$ `r round(C, 1)`. For 
+both ambient air and a pure gas (e.g., helium) as the headspace gas, we evaluate 
+three scenarios for the mixing ratio, $R_M$. The $R_M$ value indicates the 
+proportion of gas relative to water in the headspace mixture. In the simulations 
+below, we set $R_M$ to be (1) very low ($R_{M(low)} = 0.1$), (2) at equal mixing 
+proportions ($R_{M(equal)} = 0.5$), and (3) very high ($R_{M(high)} = 0.9$). For 
+reference, in the simulations above, and in the empirical field measurments, 
+$R_M$ was set at `r R_M`. Also note that the total volume of the mixing vessel 
+in the field study was `r V`mL, so the volume of gas in the mixing vessel was 
+intended to be $R_M \cdot V_g =$ `r V_g`mL and the volume of water was 
+$R_M \cdot V_w =$ `r V_w`mL. Finally, note that varying the total volume of the 
+mixing vessel has no consequence for $C$, only the ratio of gas to water.
+
+#### Ambient air
+For the same fixed concentration, $C =$ `r round(C, 1)`, using a low proportion 
+of gas relative to water ($R_{M(low)} = 0.1$) in the equilibrated headspace 
+results in a smaller value of $M_h$ relative to the simulations above with
+$R_M =$ `r R_M`. This simulation assumes ambient air is used as the headspace
+gas.
+```{r dg_calc_vary_rm_low, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel to something
+# smaller than the scenario above
+R_M_low <- 0.1
+
+C <- C # known N2O concentration in sample from simulation above
+B <- 99 # barometric pressure (kPa)
+V_g_low <- V * R_M_low # volume of reference gas in headspace
+V_w_low <- V * (1 - R_M_low) # volume of water below headspace
+T_c <- 23 # temperature ( C )
+M_a_n2o <- 0.310 # gas measured in reference (air)
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+
+#### Re-arranged previous equation to solve for Mh (with low proportion of gas to water)
+M_h_low <- (((C / 1e9) * 8.3144598 * (T_c + 273.15) * V_w_low) + (1e-6 * B * V_g_low * M_a_n2o)) / (1e-6 * B * (V_g_low + 8.3144598 * (T_c + 273.15) * V_w_low * H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15))))
+
+round(M_h_low, 3)
+```
+
+Next, we calculate $M_h$ for equal proportions of gas and water in 
+the headspace, such that $R_{M(equal)} = 0.5$. In this case we get a slightly 
+higher value for $M_h$ relative to the low mixing proportion above.
+```{r dg_calc_vary_rm_equal, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel to something
+# smaller than the scenario above
+R_M_equal <- 0.5
+
+C <- C # known N2O concentration in sample from simulation above
+B <- 99 # barometric pressure (kPa)
+V_g_equal <- V * R_M_equal # volume of reference gas in headspace
+V_w_equal <- V * (1 - R_M_equal) # volume of water below headspace
+T_c <- 23 # temperature ( C )
+M_a_n2o <- 0.310 # gas measured in reference (air)
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+
+#### Re-arranged previous equation to solve for Mh
+M_h_equal <- (((C / 1e9) * 8.3144598 * (T_c + 273.15) * V_w_equal) + (1e-6 * B * V_g_equal * M_a_n2o)) / (1e-6 * B * (V_g_equal + 8.3144598 * (T_c + 273.15) * V_w_equal * H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15))))
+
+round(M_h_equal, 3)
+```
+
+Next, we calculate $M_h$ for a higher proportion of gas relative to water in 
+the headspace, such that $R_{M(high)} = 0.9$. In this case we again get a higher 
+value for $M_h$.
+```{r dg_calc_vary_rm_high, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel to something
+# smaller than the scenario above
+R_M_high <- 0.9
+
+C <- C # known N2O concentration in sample from simulation above
+B <- 99 # barometric pressure (kPa)
+V_g_high <- V * R_M_high # volume of reference gas in headspace
+V_w_high <- V * (1 - R_M_high) # volume of water below headspace
+T_c <- 23 # temperature ( C )
+M_a_n2o <- 0.310 # gas measured in reference (air)
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+
+#### Re-arranged previous equation to solve for Mh
+M_h_high <- (((C / 1e9) * 8.3144598 * (T_c + 273.15) * V_w_high) + (1e-6 * B * V_g_high * M_a_n2o)) / (1e-6 * B * (V_g_high + 8.3144598 * (T_c + 273.15) * V_w_high * H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15))))
+
+round(M_h_high, 3)
+```
+
+Finally, we simulate $C$ again with measurment error on $M_h$, but using the
+values of $M_h$ obtained above using the 3 different values of $R_M$.
+
+##### $R_{M(low)}$
+```{r simulate_C_h_me_low, echo=TRUE, message=FALSE}
+C_h_me_low <- rep(NA, nsim)
+
+M_h_me_low <- rnorm(nsim, M_h_low, n2o_sd_epa)
+
+for(i in 1:nsim){
+  C_h_me_low[i] <- 1e-6 * B * (V_g_low * (M_h_me_low[i] - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w_low) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_me_low[i])
+}
+
+# combine into dataframe for ggplot. 
+C_h_low_df <- tibble(C_h_me_low = C_h_me_low, M_h_me_low = M_h_me_low)
+# save absolute error for plot
+C_h_me_low_abs <- (C_h_me_low * 1e9) - C
+```
+
+```{r plot_C_h_me_low, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_h_low_df, aes(x = M_h_me_low, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_h_low , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurement") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_h_low_df, aes(x = C_h_me_low * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_h_low_df, aes(x = C_h_me_low_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+##### $R_{M(equal)}$
+```{r simulate_C_h_me_equal, echo=TRUE, message=FALSE}
+C_h_me_equal <- rep(NA, nsim)
+
+M_h_me_equal <- rnorm(nsim, M_h_equal, n2o_sd_epa)
+
+for(i in 1:nsim){
+  C_h_me_equal[i] <- 1e-6 * B * (V_g_equal * (M_h_me_equal[i] - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w_equal) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_me_equal[i])
+}
+
+# combine into dataframe for ggplot. 
+C_h_equal_df <- tibble(C_h_me_equal = C_h_me_equal, M_h_me_equal = M_h_me_equal)
+# save absolute error for plot
+C_h_me_equal_abs <- (C_h_me_equal * 1e9) - C
+```
+
+```{r plot_C_h_me_equal, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_h_equal_df, aes(x = M_h_me_equal, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_h_equal , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurement") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_h_equal_df, aes(x = C_h_me_equal * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_h_equal_df, aes(x = C_h_me_equal_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+##### $R_{M(high)}$
+```{r simulate_C_h_me_high, echo=TRUE, message=FALSE}
+C_h_me_high <- rep(NA, nsim)
+
+M_h_me_high <- rnorm(nsim, M_h_high, n2o_sd_epa)
+
+for(i in 1:nsim){
+  C_h_me_high[i] <- 1e-6 * B * (V_g_high * (M_h_me_high[i] - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w_high) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_me_high[i])
+}
+
+# combine into dataframe for ggplot. 
+C_h_high_df <- tibble(C_h_me_high = C_h_me_high, M_h_me_high = M_h_me_high)
+# save absolute error for plot
+C_h_me_high_abs <- (C_h_me_high * 1e9) - C
+```
+
+```{r plot_C_h_me_high, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_h_high_df, aes(x = M_h_me_high, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_h_high , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurement") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_h_high_df, aes(x = C_h_me_high * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_h_high_df, aes(x = C_h_me_high_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+#### Pure gas
+To check the effect of different mixing ratios using a pure gas such as helium 
+as the headspace, we again re-arrange the equation to solve for $M_h$, given 
+known $C$. In this case with a pure gas, we set $M_a = 0$.
+
+For the same fixed concentration, $C$, using a low proportion of pure gas 
+relative to water ($R_{M(low)} = 0.1$) in the equilibrated headspace results in 
+a much smaller value of $M_h$ relative to the original simulations (and field 
+method) where $R_M =$ `r R_M`. This is because the mixing ratio is very low and 
+also because there is no need to account for ambient air in the headspace.
+```{r dg_calc_vary_rm_pg_low, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel to something
+# smaller than the scenario above
+R_M_low <- 0.1
+
+C <- C # known N2O concentration in sample from simulation above
+B <- 99 # barometric pressure (kPa)
+V_g_low <- V * R_M_low # volume of reference gas in headspace
+V_w_low <- V * (1 - R_M_low) # volume of water below headspace
+T_c <- 23 # temperature ( C )
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+
+#### Re-arranged previous equation to solve for Mh
+M_h_pg_low <- (((C / 1e9) * 8.3144598 * (T_c + 273.15) * V_w_low) + (1e-6 * B * V_g_low * 0)) / (1e-6 * B * (V_g_low + 8.3144598 * (T_c + 273.15) * V_w_low * H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15))))
+
+round(M_h_pg_low, 3)
+```
+
+Next, we calculate $M_h$ for equal proportions of pure gas and water in 
+the headspace, such that $R_{M(equal)} = 0.5$. In this case we get a 
+substantially lower value for $M_h$ relative to the low mixing proportion above.
+```{r dg_calc_vary_rm_pg_equal, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel to something
+# smaller than the scenario above
+R_M_equal <- 0.5
+
+C <- C # known N2O concentration in sample from simulation above
+B <- 99 # barometric pressure (kPa)
+V_g_equal <- V * R_M_equal # volume of reference gas in headspace
+V_w_equal <- V * (1 - R_M_equal) # volume of water below headspace
+T_c <- 23 # temperature ( C )
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+
+#### Re-arranged previous equation to solve for Mh
+M_h_pg_equal <- (((C / 1e9) * 8.3144598 * (T_c + 273.15) * V_w_equal) + (1e-6 * B * V_g_equal * 0)) / (1e-6 * B * (V_g_equal + 8.3144598 * (T_c + 273.15) * V_w_equal * H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15))))
+
+round(M_h_pg_equal, 3)
+```
+
+Next, we calculate $M_h$ for a higher proportion of pure gas relative to water 
+in the headspace, such that $R_{M(high)} = 0.9$. In this case we again get a 
+very low value for $M_h$.
+```{r dg_calc_vary_rm_pg_high, echo=TRUE}
+# Set volume of syringe/vessel
+V <- 140 # 140mL syringe
+
+# Set ratio of gas volume to water volume in the mixing vessel to something
+# smaller than the scenario above
+R_M_high <- 0.9
+
+C <- C # known N2O concentration in sample from simulation above
+B <- 99 # barometric pressure (kPa)
+V_g_high <- V * R_M_high # volume of reference gas in headspace
+V_w_high <- V * (1 - R_M_high) # volume of water below headspace
+T_c <- 23 # temperature ( C )
+H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
+
+#### Re-arranged previous equation to solve for Mh
+M_h_pg_high <- (((C / 1e9) * 8.3144598 * (T_c + 273.15) * V_w_high) + (1e-6 * B * V_g_high * 0)) / (1e-6 * B * (V_g_high + 8.3144598 * (T_c + 273.15) * V_w_high * H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15))))
+
+round(M_h_pg_high, 3)
+```
+
+Below, we again simulate $C$ with measurment error on $M_h$, using the
+values of $M_h$ obtained above using the 3 different values of $R_M$ with the 
+hypothetical pure gas headspace mixture.
+
+##### $R_{M(low)}$
+```{r simulate_C_h_me_pg_low, echo=TRUE, message=FALSE}
+C_h_pg_me_low <- rep(NA, nsim)
+
+M_h_pg_me_low <- rnorm(nsim, M_h_pg_low, n2o_sd_epa)
+
+for(i in 1:nsim){
+  C_h_pg_me_low[i] <- 1e-6 * B * (V_g_low * (M_h_pg_me_low[i] - 0) / (8.3144598 * (T_c + 273.15) * V_w_low) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_pg_me_low[i])
+}
+
+# combine into dataframe for ggplot. 
+C_h_pg_low_df <- tibble(C_h_pg_me_low = C_h_pg_me_low, M_h_pg_me_low = M_h_pg_me_low)
+# save absolute error for plot
+C_h_pg_me_low_abs <- (C_h_pg_me_low * 1e9) - C
+```
+
+```{r plot_C_h_me_pg_low, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_h_pg_low_df, aes(x = M_h_pg_me_low, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_h_pg_low , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurement") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_h_pg_low_df, aes(x = C_h_pg_me_low * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_h_pg_low_df, aes(x = C_h_pg_me_low_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+##### $R_{M(equal)}$
+```{r simulate_C_h_me_pg_equal, echo=TRUE, message=FALSE}
+C_h_pg_me_equal <- rep(NA, nsim)
+
+M_h_pg_me_equal <- rnorm(nsim, M_h_pg_equal, n2o_sd_epa)
+
+for(i in 1:nsim){
+  C_h_pg_me_equal[i] <- 1e-6 * B * (V_g_equal * (M_h_pg_me_equal[i] - 0) / (8.3144598 * (T_c + 273.15) * V_w_equal) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_pg_me_equal[i])
+}
+
+# combine into dataframe for ggplot. 
+C_h_pg_equal_df <- tibble(C_h_pg_me_equal = C_h_pg_me_equal, M_h_pg_me_equal = M_h_pg_me_equal)
+# save absolute error for plot
+C_h_pg_me_equal_abs <- (C_h_pg_me_equal * 1e9) - C
+```
+
+```{r plot_C_h_me_pg_equal, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_h_pg_equal_df, aes(x = M_h_pg_me_equal, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_h_pg_equal , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurement") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_h_pg_equal_df, aes(x = C_h_pg_me_equal * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_h_pg_equal_df, aes(x = C_h_pg_me_equal_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
+##### $R_{M(high)}$
+```{r simulate_C_h_me_pg_high, echo=TRUE, message=FALSE}
+C_h_pg_me_high <- rep(NA, nsim)
+
+M_h_pg_me_high <- rnorm(nsim, M_h_pg_high, n2o_sd_epa)
+
+for(i in 1:nsim){
+  C_h_pg_me_high[i] <- 1e-6 * B * (V_g_high * (M_h_pg_me_high[i] - 0) / (8.3144598 * (T_c + 273.15) * V_w_high) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_pg_me_high[i])
+}
+
+# combine into dataframe for ggplot. 
+C_h_pg_high_df <- tibble(C_h_pg_me_high = C_h_pg_me_high, M_h_pg_me_high = M_h_pg_me_high)
+# save absolute error for plot
+C_h_pg_me_high_abs <- (C_h_pg_me_high * 1e9) - C
+```
+
+```{r plot_C_h_me_pg_high, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
+ggarrange(
+ggplot(C_h_pg_high_df, aes(x = M_h_pg_me_high, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.001) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = M_h_pg_high , color = 'red') + 
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurement") +
+  xlab("Concentration (nmol L-1)"),
+
+ggplot(C_h_pg_high_df, aes(x = C_h_pg_me_high * 1e9, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = C , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
+  xlab("Concentration (nMol)"),
+
+ggplot(C_h_pg_high_df, aes(x = C_h_pg_me_high_abs, y = after_stat(density))) +  
+  geom_histogram(binwidth = 0.025) + 
+  geom_density(aes(y = after_stat(density)), color = 'green') +
+  geom_vline(xintercept = 0 , color = 'red') +  
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
+  xlab("Error (nMol)"),
+nrow = 1
+)
+```
+
 ### All observables
-Simulate measurement error for all observables (_i.e._, Equilibrated headspace, 
-air, water temperature, and barometric pressure ) and the resulting effects on
-dissolved N2O observations
+In order to get a sense of the total error in a measurement resulting from the
+sources above, we simulate measurement error for all of the observables (_i.e._, 
+GC measure of gas in equilibrated headspace, GC measure of gas in air, water
+temperature, and barometric pressure) and assess the resulting effects on the 
+dissolved N2O concentration measure. In this scenario, it is assumed that, as 
+with the empirical field measurements, air is used as the headspace gas and the 
+mixing ratio is $R_M = $ `r R_M`.
 ```{r simulate_me_all, message=F}
@@
         axis.title.y = element_blank()) + 
-  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  ggtitle( "Absolute error in dissolved \nN2O estimate (measured - true) \n") +
   xlab( "Error (nMol)"),
@@
 
-Print the estimated standard deviations of dissolved N2O measurements
-and the resulting 95% credible interval for error in measurements.
-
+The estimated standard deviations of dissolved N2O measurements
+and the resulting 95% credible interval for error in measurements is printed 
+below.
 ```{r calc_sd_me_all}
@@
 
+Alternatively, calculate the central 95th percentile of the simulated measures.
 ```{r calc_sd_me_eq_all_2}
@@
 
-## Summary of results
-
+## Summary of contributions to error
+Below, we compare the total error estimated for all observables in the last 
+simulation, to the error estimated from individual sources.
 ```{r final_figure_N2O_diss, message=FALSE, fig.align='center', fig.asp=0.6, fig.width=6}
@@
                             "All_me" = "All sources")) +
-  xlab("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
+  xlab("Absolute error in dissolved \nN2O estimate (measured - true) \n") +
   ylab("") 
@@
 # Equilibrium concentration
-
-The full equation for estimating the concentration (nM) of ($\text{N}_2\text{O}$) gas
+The equation for estimating the concentration (nM) of ($\text{N}_2\text{O}$) gas
 dissolved in water that is in equilibrium with the atmosphere is:
@@
 -   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ below)
--   $T_c$ = temperature celsius
+-   $T_c$ = temperature (K) = T_c (Celsius) + 273.15
 
 ## Simulate equilibrium $\text{N}_2\text{O}$
-
-A "known" concentration is simulated:
-
+A "known" concentration is calculated:
 ```{r c_sat}
-C_eq <- M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
-
-C_eq <- C_eq * 1e9 #convert mol to nmol
+C_eq <- M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1 / (T_c + 273.15) - 1 / 298.15)))
+
+C_eq <- C_eq * 1e9 # convert mol to nmol
 ```
@@
 ## Simulate measurement error
-
 ### Ambient air
-
-Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) is simulated 
-below. The resulting effects on the uncertainty of the equilibrium N2O 
-measurement is plotted along with the absolute error in N2O as a result of this
-measurement.
-
+Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) 
+is simulated below. The resulting effects on the uncertainty of the equilibrium 
+N2O measurement is plotted along with the absolute error in N2O as a result of 
+this measurement.
 ```{r simulate_me_sat_C_a_n2o, message=F}
@@
 for(i in 1:nsim){
-    C_a_eq_me[i] <- M_a_n2o_me[i] * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
+    C_a_eq_me[i] <- M_a_n2o_me[i] * B * 1e-6 * (H_n2o  * exp(2700*(1 / (T_c + 273.15) - 1 / 298.15)))
 }
@@
         axis.title.y = element_blank()) +
-  ggtitle("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  ggtitle("Absolute error in equilibrium \nN2O estimate (measured - true) \n") +
   xlab("Error (nMol)"),
@@
 ### Water temperature
-
 ```{r simulate_me_sat_T_c, message=F}
@@
 ### Barometric pressure
-
 ```{r simulate_me_sat_B, message=F}
@@
 Simulate measurement error for all observables (_i.e._, air, water temperature,
-and barometric pressure ) and the resulting effects on dissolved N2O 
+and barometric pressure) and the resulting effects on equilibrium N2O 
 observations.
@@
         axis.title.y = element_blank()) + 
-  ggtitle( "Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  ggtitle( "Absolute error in equilibrium \nN2O estimate (measured - true) \n") +
   xlab( "Error (nMol)"),
@@
 
-## Summary of results
-
+## Summary of contributions to error
 ```{r final_figure_N2O_eq, message=FALSE, fig.align='center', fig.asp=0.6, fig.width=6}
@@
                             "All_me" = "All sources")) +
-  xlab("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
+  xlab("Absolute error in equilibrium \nN2O estimate (measured - true) \n") +
   ylab("") 
@@
 # Saturation ratio
-(Check this equation in text! Particularly the description of variables)
-The full equation for estimating the saturation ratio of ($N_20$) gas
-dissolved in water is:
+The equation for estimating the saturation ratio for ($N_20$) gas in water is:
 
@@
 -   $V_a$ = volume of reference gas in air
--   $M_h$ = gas mixture ratio in the headspace
--   $M_r$ = gas mixture ratio in the reference (C_a_n2o below)
+-   $M_h$ = gas measured in the headspace
+-   $M_r$ = gas measured in the reference (air)
 -   $V_w$ = volume of water below the headspace
@@
 
+Uncertainty in the saturation ratio, $S$, is simulated below for both a
+standard GC and a high precision GC. The GC used for the current study is 
+optimized to quantify several analytes across a broad range of concentrations 
+but greater precision can be achieved by optimizing the instrument for a single 
+analyte across a narrower range of concentrations. For example, laboratories
+participating in an international analysis of N2O standards used for
+atmsopheric monitoring reported a mean standard deviation of
+`r mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015))`, which is
+10-fold lower than the standard deviation of N2O measurements in this
+study (`r n2o_sd_epa`). Utilizing more precise N2O measurements for the
+equilibrated headspace and/or air will reduce uncertainty in the
+estimated dissolved N2O concentration, dissolved N2O equilibrium value,
+and N2O saturation ratio.
+
+In addition to the two different GC setups, uncertainty in $S$ is also simulated
+for under additional scenarios using combinations of ambient air vs a pure gas 
+as the headspace gas along with a standard _vs._ high precision thermometer for 
+field temperature (C) measurements.
+
+All of these simulations use the mixing ratio used for the empirical field 
+measurements, where $R_M =$ `r R_M`.
+
 ## Standard GC
+### Ambient air, standard thermometer
 Simulate uncertainty in the saturation ratio due to measurement error associated
-with the standard GC, using ambient air as the headspace gas, and the standard 
-field thermometer.
-
-### Ambient air, standard thermometer
-
+with the standard GC, using ambient air as the headspace gas, and a standard 
+field thermometer for temperature. This scenario represents the approach used
+for the empirical observations used in this study.
 ```{r simulate_me_ratio_all_air, message=F}
@@
         axis.title.y = element_blank()) + 
-  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
+  ggtitle( "Absolute error in saturation \nratio estimate (measured - true) \n") +
   xlab( "Error (nMol)"),
@@
 #### Map source-sink status
-
+Below, the source-sink status of the empirical measurments used for this study
+are estimated, assuming the measurement errors estimated above. The proportion 
+of observations assumed to be sinks or sources of $N_2O$ is also estimated along
+with the estimate of the proportion of the lakes sampled for which the status 
+was assumed to be "undetermined" due to uncertainty arising from the estimated
+measurement error. A map of each of the sampled lakes' estimated status is also 
+shown.
 ```{r map_ratio_all_air, message=F}
@@
 ### Pure gas, standard thermometer
-
 Simulate uncertainty in the saturation ratio due to measurement error associated
@@
 
-#### Map source-sink status
-
+#### Source-sink status
 ```{r map_ratio_pg, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_ratio_pg, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 Simulate uncertainty in the saturation ratio due to measurement error associated
-with the standard GC, ambient air as the headspace gas, and using a high 
-precision thermometer (e.g., MIMS).
-
+with the standard GC and ambient air used as the headspace gas, and a high 
+precision thermometer (e.g., MIMS) used for temperature measurements.
 ```{r simulate_tc_mims_air_hpt, message=F}
@@
 
-Simulate uncertainty in the saturation ratio using ambient air in the 
+Simulate uncertainty in the saturation ratio when using ambient air in the 
 headspace and high precision temperature measurements
@@
 
-#### Map source-sink status
-
+#### Source-sink status
 ```{r map_mims_air_hpt, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_mims_air_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 precision thermometer (e.g., MIMS).
-
 ```{r simulate_me_ratio_all_pg_hpt, message=F}
@@
 
-#### Map source-sink status
-
+#### Source-sink status
 ```{r map_sat_ratio_pg_hpt, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_sat_ratio_pg_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 ## High precision GC
-
 The GC used in the current study is optimized to quantify several
@@
 estimated dissolved N2O concentration, dissolved N2O equilibrium value,
-and N2O saturation ratio.
+and N2O saturation ratio. Again, the simulations below assume the mixing 
+ratio used in the empirical field observations, where $R_M =$ `r R_M`.
 
@@
 # simulate improved measurement of N2O in equilibrated headspace gas 
-# assume TRUE value is C_e and error is simulated with with sd from literature
+# assume TRUE value is M_h_pg_hpgc and error is simulated with with sd from literature
 M_h_pg_hpgc_me <- rnorm(nsim, M_h_pg_hpgc, n2o_sd_lit)
 
-# under this scenario we don't need to subtract C_a_n2o (N2O concentration in air) from C_e (N2O concentration in equilibrated headspace)
+# Set true value
+# under this scenario we don't need to subtract M_a_n2o (N2O concentration in air) from M)h (N2O concentration in equilibrated headspace)
 S_pg_hpgc <- (1e-6 * B * (V_g * (M_h_pg_hpgc - 0) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15) ) * M_h_pg_hpgc)) /
@@
 
+# Container for simulations
 S_pg_hpgc_me <- rep(NA, 1e4)
@@
 
-#### Map source-sink status
-
+#### Source-sink status
 ```{r map_sat_ratio_gc_pure_gas, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_sat_ratiogc_pure_gas, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 ### Pure gas, high precision thermometer
-
 ```{r simulate_me_ratio_hpgc_pg_hpt_all, message=F}
@@
 
-#### Map source-sink status
-
+#### Source-sink status
 ```{r map_ratio_hpgc_pg_hpt, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_ratio_hpgc_pg_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 # MIMS
-
 In Membrane Inlet Mass Spectrometry (MIMS), dissolved gases (typically
@@
 ## Dissolved concentration
-
 ### Standard GC
-
 ```{r simulate_me_hpt_MIMS, message=F}
@@
   (1e-6 * B * C_Ar_a * (H_ar * exp( 1500 * (1 / (T_c + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
-  (1e-6 * B * C_Ar_a*(H_ar  * exp(1500 *(1/(T_c + 273.15) - 1/298.15)))))
+  (1e-6 * B * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c + 273.15) - 1/298.15)))))
 
@@
 ### High precision GC
-
 ```{r hpgc_M_a_MIMS}
@@
 ## Saturation ratio
-
 Calculate the "true" ratio via MIMS
@@
 ### Standard GC 
-
 #### Standard field temperature
@@
 
-##### Map source-sink status
-
+##### Source-sink status
 ```{r map_me_ratio_MIMS, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_me_ratio_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 #### High precision temperature
-
 ```{r simulate_me_ratio_hpt_MIMS, message=F}
@@
 
-##### Map source-sink status
-
+##### Source-sink status
 ```{r map_me_ratio_hpt_MIMS, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_me_ratio_hpt_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 ### High precision GC 
-
 #### Standard temperature
-
 This simulation is for field measurment of temperature for the equilibrium concentration with the standard glass thermometer.
@@
 
-##### Map source-sink status
-
+##### Source-sink status
 ```{r map_me_ratio_hpgc_MIMS, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_me_ratio_hpgc_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 #### High precision temperature
-
 The thermometers used to measure water temperature in this study have an
@@
 
-##### Map source-sink status
-
+##### Source-sink status
 ```{r map_me_ratio_hpgc_hpt_MIMS, message=F}
@@
   kable(digits = 2, caption = "Proportion of lakes classified as source/sink")
-
-# sf for plotting
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-```
-
-```{r plot_map_me_ratio_hpgc_hpt_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
-## load ecoregions----
-# read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
-
-# Check CRS
-#st_crs(ecoR) # 3857
-ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
-#st_crs(ecoR) # 5070
-
-# SET UP CUSTOM COLORS FOR ECOREGIONS
-# Custom color pallette for ecoregion polygons. Attempted to mirror
-# https://www.epa.gov/national-aquatic-resource-surveys/
-# ecoregional-results-national-lakes-assessment-2012
-cols <- c("Coastal Plains" = "orange1",
-          "Northern Appalachians" = "lightpink1",
-          "Northern Plains" = "darksalmon",
-          "Southern Appalachians" = "mediumturquoise",
-          "Southern Plains" = "khaki4",
-          "Temperate Plains" = "forestgreen", 
-          "Upper Midwest" = "deepskyblue4",
-          "Western Mountains" = "saddlebrown",
-          "Xeric" = "lightskyblue4")
-
-## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
-
-# To enable spatial analysis of the data, the dataframe will be converted to a 
-# 'simple features' (sf) object.
-# Define coordinates
-coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)
-
-dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
-                  crs = 4269) %>% # standard for lat/lon
-  st_transform(5070) # project to CONUS ALBERS for plotting
-
-## Map
-ggplot() +
-    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
-    geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
-            shape = 1,
-            show.legend = "point") +
-    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
-    # guide argument below removes points from the boxes in the ecoregion legend
-    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
-    scale_fill_manual("Ecoregion", 
-                      values = cols,
-                      guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
-    scale_size(name = expression(N[2]*O~(nM)),
-               range = c(0.5, 8), # custom size range
-               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
-    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
-          legend.title = element_text(size = 8))
 ```
@@
 # Summary of ratio results
-
 Precision of the N2O saturation ratio can be improved using high
@@
 directly measured using MIMS.
-
 ```{r final_summary, message=FALSE}
@@
 directly measured using MIMS.
-
 ```{r final_summary_MIMS, message=FALSE}

### 2025-02-18 manuscript/manuscript_file.Rmd
Completed revisions in response to Lester and Jana's tech review.

@@
 We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of lakes functioning as N~2~O sinks.
-The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate concentration, waterbody size, and water temperature.
+The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate concentration, waterbody surface area, and water temperature.
 Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS lakes is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
@@
 1)**.
-The survey design was stratified by state and used unequal probability categories based on lake size to ensure that large lakes, which are far less common than small lakes, were included in the study.
+The survey design was stratified by state and used unequal probability categories based on lake size (surface area) to ensure that large lakes, which are far less common than small lakes, were included in the study.
 Sample sites were distributed across all nine major CONUS ecoregions [@Herlihy_etal_2008] and varied widely in water chemistry, morphometry, watershed land-use, and climate [@epa_data].
@@
 
-Nearly `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent of the sampled lakes were undersaturated in N~2~O and were therefore functioning as N~2~O sinks **(Fig.
+`r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0) %>% xfun::numbers_to_words(., cap = TRUE)` percent of the sampled lakes had dissolved N~2~O concentrations that were lower than the concentration expected if the lake were in equilibrium with the atmosphere, suggesting those lakes were functioning as N~2~O sinks **(Fig.
 1)**.
+N~2~O concentrations were close to equilibrium values in some of these lakes, however, and after accounting for measurement uncertainty in both dissolved and equilibrium N~2~O concentration (See SI), we can confidently report that a minimum of `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent of the sampled lakes were functioning as sinks, but the number could be as high as `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent.
+When accounting for measurement error, `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "undetermined") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent had a dissolved N~2~O concentration that was indistinguishable from the equilibrium concentration and only `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "source") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent were supersaturated **(Fig.
+1)**.
+
 The Bayesian hierarchical model predicted that the majority lakes in the POI (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% credible interval (CI): `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) functioned as net N~2~O sinks during the summer months.
@@
 This finding runs counter to the long-standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lakes can function as N~2~O sinks.
-Webb et al.
-[@Webb_etal_2019] found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months.
-An investigation of boreal systems in Canada found that 40% of rivers, lakes, and ponds sampled during the summer were functioning as N~2~O sinks[@Soued_etal_2016].
-Similarly, 29% of surveyed boreal lakes in Finland were found to be N~2~O sinks during the summer [@Kortelainen_etal_2020].
-While these latter surveys were conducted at regional scales, the national-scale results presented here provide the strongest evidence to date that most lakes function as N~2~O sinks during the summer months.
+Webb et al.[@Webb_etal_2019] found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Kortelainen et al.[@Kortelainen_etal_2020] found that 29% of surveyed boreal lakes in Finland were N~2~O sinks during the summer.
+An investigation of boreal systems in Canada found that 40% of rivers, lakes, and ponds sampled during the summer were functioning as N~2~O sinks[@Soued_etal_2016] and all 34 lotic and lentic sites in the US National Ecological Observatory Network (NEON) functioned as N~2~O sinks at least once during 1.5 - 6 years of monitoring[@Aho_etal_2023].
+While these latter surveys were conducted at regional scales or had relatively few observtion sites, the national-scale results presented here provide the strongest evidence to date that most lakes function as N~2~O sinks during the summer months.
 This pattern held across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use.
@@
 2)**.
-Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had a median value less than one, indicating that most lakes were undersaturated in these populations.
+Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had a median value less than one, indicating that most lakes in these populations were undersaturated .
 However, the long right tails on these distributions, comprised of relatively rare lakes with very high levels of N~2~O supersaturation, were sufficient to drive the population means to values close to, or even greater than one.
@@
 Large lakes, regardless of NO~3~ status, were predicted to have N~2~O saturation ratios closer to equilibrium.
-.
 Thus N~2~O saturation ratio tended to deviate from a value of one more strongly in small lakes as compared to larger lakes.
@@
 For example, N~2~O inputs from streams or groundwater[@Wang_etal_2023] may constitute a substantial N~2~O subsidy to small lakes but are more likely be diluted in larger lakes with greater volumes.
-Small lakes also tend to have lower water volume to benthic surface area ratios,and therefore stronger connections between sediment, an area of active N~2~O processing, and the air-water interface where our samples were collected.
+Small lakes also tend to have lower water volume to benthic surface area ratios[@Cael_etal_2017], and therefore stronger connections between sediment, an area of active N~2~O processing, and the air-water interface where our samples were collected.
 Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep waterbodies, resulting in greater rates of N~2~O production or consumption in small lakes.
@@
 Thus, greater air-water gas exchange rates in large lakes may maintain dissolved N~2~O concentrations close to equilibrium, whereas the effect of biogeochemistry on dissolved N~2~O may be more pronounced in small lakes because low air-water gas exchange limits the rate at which excess N~2~O can be ventilated or depleted N~2~O can be replenished.
-While previous surveys have documented relationships between lake size and dissolved methane concentration [@Pighini_etal_2018; @Juutinen_etal_2009; @Borges_etal_2011], to the best of our knowledge this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
+While previous surveys have documented negative relationships between lake size and dissolved methane concentration [@Pighini_etal_2018; @Juutinen_etal_2009; @Borges_etal_2011], to the best of our knowledge this is the first study to document a relationship between lake size and $\Delta$ N~2~O.
 
@@
 ![Figure 5: A) Predicted N~2~O emission rate and B) flux plotted against waterbody size for all 465,896 lakes in the population of interest.
-Points and bars reflect the posterior means (points) and 95% credible intervals, respectively, based on 2000 draws from the posterior predictive distribution for each waterbody.
-C) A comparison of the estimated relative flux contributions from lakes in each size class to the total CONUS flux.
-The red line connects the mean of the posterior estimate of the ratio, at each size class, based on 2000 draws.
-The black polygon indicates the 95% credible intervals for the ratio estimates across the size classes.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
+Points (red) and bars (black) reflect the posterior means and 95% credible intervals, respectively, based on 2000 draws from the posterior predictive distribution for each waterbody.
+C) Proportion of total CONUS N~2~O flux attributed to each of the 465,896 lakes in the POI.
+Points (red) and bars (black) reflect the posterior means and 95% credible intervals, respectively, based on 2000 draws from the posterior predictive distribution for each waterbody](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
 
@@
 
-Despite the estimated widespread N~2~O undersaturation among CONUS lakes, the credible intervals for ecoregion-level flux estimates included zero for all but three ecoregions **(Table 1)**.
+Despite the estimated widespread N~2~O undersaturation among CONUS lakes, credible intervals for ecoregion-level flux estimates included zero for all but three ecoregions **(Table 1)**.
 Thus, status as a net source versus sink of N~2~O was uncertain for most ecoregions.
@@
 Lauerwald et al [@Lauerwald_etal_2019] used a mechanistic model to estimate that CONUS lakes emit 8,425 metric tons N~2~O year^-1^.
-That estimate is nearly 25-fold greater than the posterior mean of our estimate (344 metric tons N~2~O year^-1^), despite our estimate including an order of magnitude more lakes than Lauerwald et al.[@Lauerwald_etal_2019]  A possible explanation for the large difference between these estimates is that the Lauerwald model does not allow for lakes to behave as N~2~O sinks, a limitation that extends to many mechanistic models used to estimate N~2~O fluxes from inland waters[@Marzadri_etal_2021;@Marzadri_etal_2017;@Mosier_etal_1998]].
+That estimate is nearly 25-fold greater than the posterior mean of our estimate (344 metric tons N~2~O year^-1^), despite our estimate including an order of magnitude more lakes than Lauerwald et al.[@Lauerwald_etal_2019]  A possible explanation for the large difference between these estimates is that the Lauerwald model does not allow for lakes to behave as N~2~O sinks, a limitation that extends to many mechanistic models used to estimate N~2~O fluxes from inland waters[@Marzadri_etal_2021;@Marzadri_etal_2017;@Mosier_etal_1998].
 However, even models that allow for net N~2~O uptake can overpredict N~2~O flux.
@@
 
-Although lakes in the largest lake size class tend to have N~2~O levels close to equilibrium, they emit more N~2~O than the other size classes due to their expansive surface area.
-The N~2~O flux from these lakes is also more uncertain than the smaller lake size classes because relatively small uncertainties in dissolved and equilibrium N~2~O concentrations translate to large uncertainties in flux when propagated across the surface of large lakes.
-Uncertainty in the national flux estimate was too large to determine if CONUS lakes functioned as a net source or sink of N~2~O.
+Although lakes in the largest lake size class tend to have N~2~O levels close to equilibrium, they make a large contribution to CONUS N~2~O flux due to their expansive surface area.
+The N~2~O flux from large lakes is poorly constrained, however, because relatively small uncertainties in dissolved and equilibrium N~2~O concentrations translate to large uncertainties in flux when propagated across the surface of large lakes.
 While our sample included some of the largest lakes in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O, which can be important in large waterbodies[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement and analytical uncertainty, which can propagate to uncertainties in flux estimates.
+The measurement methods employed in this study have a 95% credible interval of 0.11 (unitless) on the N~2~O saturation ratio.
+The 95% credible interval can be halved by employing gas chromatography optimized for quantifying N~2~O in the range of values observed in this study [@Hall_2024] and decreased by an order of magnitude using membrane inlet mass spectrometry[@Genthner_etal_2013; @Speir_etal_2023].
 
@@
 A better understanding of the temporal evolution of N~2~O production and consumption is an important research need to determine the extent to which small lakes can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
-
-![SI Figure 1.
-Distribution of N~2~O emission rates by ecoregion.
-Each blue line (n=500) represents the distribution of values from a single realization of the posterior prediction.](manuscript_figures/SIfigure1.tiff)
 

### 2025-02-18 scripts/dataMunge.Rmd
Completed revisions in response to Lester and Jana's tech review.

@@
 
+Roy Martin modeled the uncertainty in the N2O saturation ratio in DG_sensitivity_to_measurement_error.Rmd (Saturation ratio/Standard GC sections).
+The absolute error in the simulated values were written to desk and read in below.
+These error estimates are used to define uncertainty in the N2O saturation ratio.
+
 # Data
@@
   filter(!grepl("AK", site.id)) # omit Alaska sites
+
+# Read simulated saturation ratio values (See DG_sensitivity_to_measurement_error.Rmd, Saturation ratio/Standard GC)
+S_a_me_abs <- readRDS(paste0(localPath, 
+                             "/Environmental Protection Agency (EPA)/",
+                             "ORD NLA17 Dissolved Gas - Documents/",
+                             "inputData/S_a_me_abs.RDS"))
+
 ```
@@
          ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
-         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink")) %>%
+         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"),
+         # recompute using error from Roy's uncertainty simulations
+         n2o.src.snk.error = case_when(
+           abs(n2o.sat.ratio - 1) < quantile(S_a_me_abs, probs = c(0.975)) ~ "undetermined",
+           TRUE ~ n2o.src.snk)) %>%
   # remove fields no longer needed

### 2025-02-18 manuscript/manuscript_support.R

@@ -46,15 +46,15 @@ library(tidybayes) # prop flux by size figure
 # DATA---------------
 ## load sample data (dg object)----
 if(localPath == "") { # if DMAP, then
-  load("inputData/dg.2021-02-01.RData")
+  load("inputData/dg.2025-02-14.RData")
 } else { # if not DMAP, then...
   load(paste0(localPath,  # object name dg
               "/Environmental Protection Agency (EPA)/",
               "ORD NLA17 Dissolved Gas - Documents/",
-              "inputData/dg.2021-02-01.RData"))
+              "inputData/dg.2025-02-14.RData"))
 }
 
-save(dg, file = "manuscript/manuscript_files/dg.rda")  # save dg fle to ms folder for faster knitting
+save(dg, file = "manuscript/manuscript_files/dg.rda")  # save dg file to ms folder for faster knitting
 
 # To enable spatial analysis of the data, the dataframe will be converted to a 
 # 'simple features' (sf) object.
@@ -320,21 +320,23 @@ if(!("figure1.png" %in% list.files("manuscript/manuscript_figures"))) {
   ggplot() +
     geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
     geom_sf(data = dg.sf %>% 
-              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
-              arrange(desc(n2o.src.snk)), # grey on top of black
-            aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
+              filter(!is.na(n2o.src.snk.error), sitetype == "PROB", visit.no == 1) %>%
+              arrange(desc(n2o.src.snk.error)), # grey on top of black
+            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.error),
             show.legend = "point") +
     geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
     # guide argument below removes points from the boxes in the ecoregion legend
     # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
     scale_fill_manual("Ecoregion", values = cols,
                       guide = guide_legend(override.aes = list(shape = NA))) +
-    scale_color_manual(values = c("grey", "black"), name = "source/sink") +
[…]

### 2025-03-03 manuscript/manuscript_file.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

@@
           "em_min_max",
-          "FluxBySize",
-          "ipcc.n2o"
+          "FluxBySize"
           )
@@
 for(i in 1:N){
-    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/manuscript/manuscript_files/",
+    f[i] <- paste0("/manuscript/manuscript_files/",
     objs[i], ".rda")
@@
 for(i in 1:N){
-    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/manuscript/manuscript_figures/",
+    f[i] <- paste0("/manuscript/manuscript_figures/",
     objs[i])

### 2025-03-03 scripts/DG_sensitivity_to_measurement_error.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

@@
 ```{r setup, warning=FALSE, message=FALSE}
-#library(devtools) # needed to install USAboundaries package from
 library(knitr)
@@
 library(sf)
-
-# Install USAboundaries package from GitHub
-#devtools::install_github("ropensci/USAboundaries")
+library(usmap)
 
@@
 ```{r summary_empirical, echo=TRUE}
-load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
-            "\\ORD NLA17 Dissolved Gas - Documents\\",
-            "inputData\\dg.2021-02-01.RData"))
+load("./../inputData/dg.RData")
 
@@
 ## Load state boundaries----
-states <- USAboundaries::us_states() %>%
-  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-  st_transform(5070) # convert to CONUS Albers
+states <- usmap::us_map() %>%
+  dplyr::filter(!full %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
+  sf::st_transform(5070) # convert to CONUS Albers
 
@@
 ```
+
+# Update data with uncertainty info
+Finally, we update the dg data frame to include information on estimated measurement error
+
+``` {r update_dg_data}
+dg <- dg %>%
+  mutate(# recompute using error from uncertainty simulations
+         n2o.src.snk.error = case_when(
+           abs(n2o.sat.ratio - 1) < quantile(S_a_me_abs, probs = c(0.975)) ~ "undetermined",
+           TRUE ~ n2o.src.snk))
+
+save(dg, file = "./../inputData/dg.RData")
+```
+
+# Session Info
+
+```{r session}
+sessionInfo()
+```

### 2025-03-03 scripts/Design_Based_Estimates_N2O.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

---
title: "Design-based estimates for N2O from the NLA 2017 Survey"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  github_document:
    number_sections: TRUE
    df_print: "tibble"
    math_method: 
      engine: webtex
    #  url: https://latex.codecogs.com/svg.image?
    html_preview: TRUE
    keep_html: TRUE
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# Purpose
Calculate estimates for the NLA 2017 survey with spsurvey and local mean variance estimator for comparison to Bayesian MRP Estimates.
Tom Kincaid created a .csv file to be used for population estimates, which is loaded and munged below.

# Import and munge survey data
```{r setup_data, echo=FALSE}

# Attach the Hmisc and spsurvey packages
library(tidyverse)
library(Hmisc)
library(spsurvey)

# Setup file path
localPath <- Sys.getenv("USERPROFILE")
```

```{r import_data} 
# modified file to include N2O emission rates.  See "Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" for details
# updated file path
df_survey <- read.csv(file = "./../inputData/populationEstimates/NLA17_DissolvedGases_forPopEst.csv")
df_survey <- df_survey %>% 
  data.frame() %>%
  filter(!is.na(DISSOLVED_N2O))
nr <- nrow(df_survey) # 984

# Reorder levels for the lake size class (AREA_CAT6) variable
levels(df_survey$AREA_CAT6) <- list(
   "(1,4]" = "(1,4]",
   "(4,10]" = "(4,10]",
   "(10,20]" = "(10,20]",
   "(20,50]" = "(20,50]",
   ">50" = ">50")

head(df_survey)

tail(df_survey)

# Create data frame for spsurvey estimates
dframe <- data.frame(
  siteID = df_survey$SITE_ID,
  N2O_SAT_RATIO=df_survey$N2O_SAT_RATIO,
  DISSOLVED_N2O=df_survey$DISSOLVED_N2O,
  National=rep("National", nr),
  LAKE_ORGN=df_survey$LAKE_ORGN,
  AREA_CAT6=df_survey$AREA_CAT6,
  AG_ECO9_NM=df_survey$AG_ECO9_NM,
  State = df_survey$STRATUM,
  wgt = df_survey$WGT_TP,
  xcoord = df_survey$XCOORD,
  ycoord = df_survey$YCOORD)
```

# Survey estimates with spsurvey package
Estimates made using the local mean variance estimator
```{r spsurvey_ests, echo=FALSE, results="hide", include=TRUE}

# Calculate the estimates
#cat("\n\nCalculate CDF estimates using the local mean variance estimator\n")
if(exists("warn.df")) rm("warn.df")
time.initial <- Sys.time()
Design_estimates <- cont_analysis(dframe = dframe, 
                               vars = c("N2O_SAT_RATIO", "DISSOLVED_N2O"),
                               subpops = c("National", "LAKE_ORGN", "AREA_CAT6", "AG_ECO9_NM", "State"), 
                               weight = "wgt", 
                               xcoord = "xcoord", 
                               ycoord = "ycoord",
                               vartype = "Local") # 30 minutes on VM
Sys.time() - time.initial
```

# Summarize estimates and export
```{r summarize_mean_estimates_dissolved, echo=FALSE}
mean_dissolved <- Design_estimates$Mean %>% 
  tibble() %>%
  filter(Indicator == "DISSOLVED_N2O") %>%
  mutate(Estimate = round(Estimate * 1e9, 2),
         StdError = round(StdError * 1e9, 2),
         MarginofError = round(MarginofError * 1e9, 2),
         LCB95Pct = round(LCB95Pct * 1e9, 2),
         UCB95Pct = round(UCB95Pct * 1e9, 2),
         Subpopulation = recode(Subpopulation, # simpler naming conventions for the size categories
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  select(!Indicator)

mean_dissolved %>% print()
```

```{r save_to_drive_1, eval=FALSE, include=TRUE}
write.csv(mean_dissolved, file = "./../inputData/populationEstimates/Survey_Ests_N20_dissolved.csv")
```

```{r summarize_mean_estimates_saturation, echo=FALSE}
mean_sat_ratio <- Design_estimates$Mean %>% 
  tibble() %>%
  filter(Indicator == "N2O_SAT_RATIO") %>%
  mutate(Estimate = round(Estimate, 2),
         StdError = round(StdError, 2),
         MarginofError = round(MarginofError, 2),
         LCB95Pct = round(LCB95Pct, 2),
         UCB95Pct = round(UCB95Pct, 2),
         Subpopulation = recode(Subpopulation, # simpler naming conventions for the size categories
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  select(!Indicator)

mean_sat_ratio %>% print()
```

```{r save_to_drive_2, eval=FALSE, include=TRUE}
write.csv(mean_sat_ratio, file = "./../inputData/populationEstimates/Survey_Ests_N20_sat.csv")
```

# Session Info

```{r session}
sessionInfo()
```

### 2025-03-03 scripts/NLA_N2O_models.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

@@
 library(arrow)
+library(archive)
 
@@
           "n2o_mod6",
-          #"all_predictions_b",
-          "n2o_survey_ests",
-          "n2o_survey_ests_size",
-          "sat_survey_ests",
-          "sat_survey_ests_size"
+          "all_predictions_b"
           )
@@
 for(i in 1:N){
-    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/modelFiles/",
+    f[i] <- paste0("./../modelFiles/",
     objs[i], ".rda")
@@
 
-$$\alpha_1 \sim MVN(0, \Lambda_1)$$ 
-$$\alpha_2 \sim MVN(0, \Lambda_2)$$
+$$\alpha_1 \sim MVN(0, \Lambda_1)$$ $$\alpha_2 \sim MVN(0, \Lambda_2)$$
 $$\alpha_3 \sim MVN(0, \Lambda_3)$$
@@
 
-That originally munged gas dataset was imported below.
+That originally munged gas dataset is imported below.
 
 ```{r import_data, eval=FALSE, include=TRUE}
-load( file = paste0( localPath,
-              "/Environmental Protection Agency (EPA)/",
-              "ORD NLA17 Dissolved Gas - Documents/",
-              "inputData/dg.2021-02-01.RData")
-      )
-
-save(dg, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda") 
+load("./inputData/dg.RData")
 ```
@@
 
-save(df_model, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda") 
+save(df_model, file = "/modelFiles/df_model.rda") 
 ```
@@
 
-save(df_test, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda") 
+save(df_test, file = "modelFiles/df_test.rda") 
 ```
@@
 ```{r import_sample_frame, eval=FALSE, include=TRUE}
-df_pop <- read.csv(file = paste0(localPath,
-              "/Environmental Protection Agency (EPA)/",
-              "ORD NLA17 Dissolved Gas - Documents/",
-              "inputData/NLA_Sample_Frame.csv"), header = T)
+df_pop <- read.csv(archive::archive_read("inputData/NLA_Sample_Frame.7z")) # This takes a bit
 
@@
 
-The first model for the two N2O responses, conditioned only on the study design factors.
+The first model for the two N2O responses, conditioned only on the study
+design factors.
 
@@
 Below, National mean estimates for dissolved N2O from the model and
-design-based approaches were compared. The sample-based estimate was
-also included as a reference. The black, vertical, dashed line indicates
-the mean of the sample.
+design-based approaches were compared. The code and results for the 
+design-based approach are located at: 
+https://github.com/USEPA/DissolvedGasNla/tree/master/scripts
+
+The sample-based estimates were also included as a reference and are
+represented by the black, vertical, dashed lines.
+
+```{r import_design_ests, message=FALSE, cache=TRUE}
+### Dissolved N2O estimates
+
+n2o_survey_ests <- read.csv(file = "./../inputData/populationEstimates/Survey_Ests_N20_dissolved.csv") %>%
+  tibble() %>%
+  rename(estimate = Estimate,
+         LCL = LCB95Pct,
+         UCL = UCB95Pct) %>%
+  select(Type, Subpopulation, estimate, LCL, UCL)
+
+# ests by subpopulations
+n2o_survey_ecoregion <- n2o_survey_ests %>%
+  filter(Type == "AG_ECO9_NM") %>%
+  rename(ecoregion = Subpopulation) %>%
+  select(ecoregion, estimate, LCL, UCL) %>%
+  mutate(ecoregion = recode(ecoregion,
+                            "Coastal Plains" = "CPL",
+                            "Northern Appalachians" = "NAP",
+                            "Northern Plains" = "NPL",
+                            "Southern Appalachians" = "SAP",
+                            "Southern Plains" = "SPL",
+                            "Temperate Plains" = "TPL",
+                            "Upper Midwest" = "UMW",
+                            "Western Mountains" = "WMT",
+                            "Xeric" = "XER"))
+
+n2o_survey_size <- n2o_survey_ests %>%
+  filter(Type == "AREA_CAT6") %>%
+  rename(size_cat = Subpopulation) %>%
+  select(size_cat, estimate, LCL, UCL)
+
+### Saturation ratio estimates
+sat_survey_ests <- read.csv(file = "./../inputData/populationEstimates/Survey_Ests_N20_sat.csv") %>%
+  tibble() %>%
+  rename(estimate = Estimate,
+         LCL = LCB95Pct,
+         UCL = UCB95Pct) %>%
+  select(Type, Subpopulation, estimate, LCL, UCL)
+```
 
@@
   mutate(type = "model") %>%
-  bind_rows(cbind(n2o_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
+  bind_rows(cbind(n2o_survey_ests[1, 3:5], type = rep("survey", 1))) %>%
   add_row(estimate = round(mean(df_model$n2o), 2),
@@
   mutate(type = "model") %>%
-  bind_rows(cbind(n2o_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
+  bind_rows(cbind(n2o_survey_ests[1, 3:5], type = rep("survey", 1))) %>%
   mutate(cl_width = round(UCL - LCL, 2)) %>%
@@
   mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
-  bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
+  bind_rows(cbind(n2o_survey_ecoregion, type = rep("survey", 9))) %>%
+  #bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
   arrange(ecoregion) %>%
@@
   mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
-  bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
+  bind_rows(cbind(n2o_survey_ecoregion, type = rep("survey", 9))) %>%
   ggplot(aes(x = ecoregion, y = estimate, group = type, color = type)) +
@@
   mutate(type = "model") %>%
-  bind_rows(cbind(n2o_survey_ests_size, type = rep("survey", 5))) %>%
+  bind_rows(cbind(n2o_survey_size, type = rep("survey", 5))) %>%
   mutate(size = factor(size_cat)) %>%
@@
   select(size_cat, estimate, LCL, UCL, type) %>%
-  bind_rows(cbind(n2o_survey_ests_size, type = rep("survey", 5))) %>%
+  bind_rows(cbind(n2o_survey_size, type = rep("survey", 5))) %>%
   mutate(size_cat = forcats::fct_reorder(size_cat, estimate)) %>%

### 2025-03-03 scripts/dataMunge.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

--- 2021-02-16 scripts/dataMunge.Rmd
@@
 output:
-  html_document:
-    df_print: paged
+  github_document:
     toc: true
-    toc_float: true
-    fig_caption: yes
-    depth: 2
+    toc_depth: 5
     number_sections: true
+    keep_html: true
 editor_options: 
@@
 # Purpose
-The purpose of this .rmd is to prepare the data set for subsequent modeling.
-The data will be written as .RData object and stored at the shared documents library associated with this project: https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas
+The purpose of this .rmd was to prepare the data set for subsequent modeling.
+The data were written as .RData object that is stored in the repository associated with this project: https://github.com/USEPA/DissolvedGasNla
 
-The dissolved gas data were curated in the 'NLA' dissolved gas project in RStudio.
+The dissolved gas data were originally curated in the 'NLA' dissolved gas project in RStudio.
 The code can be found at a private github repository (https://github.com/USEPA/NLA).
-After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2021-01-13.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).
-When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.
+After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2021-01-13.txt, which is also avaialble in the GitHub repository for this project: https://github.com/USEPA/DissolvedGasNla
+
+The data object ('dg.Rdata') munged in this document was subsequently used to estimate measurement uncertainty in the N2O observations, which is documented in the 'DG_sensitivity_to_measurement_error.Rmd' (Saturation ratio/Standard GC sections) file that is also available in the repository.
+From there, an updated 'dg.RData' object including this new information on uncertainty was imported into additional RStudio document workflows for (1) making design-based estimates from the survey data (Design_Based_Estimates.Rmd) and (2) generating population estimates using a Bayesian hierarchical model (NLA_N2O_models.Rmd).
+The object is also imported for the files used to generate the manuscript (manuscript_file.Rmd and manuscript_support.Rmd).
+All of these files (and more) are included in the GitHub repository for this project in order to ensure the results were reproducible starting from this base file of the observed data.
 
@@
 
-```{r}
+```{r read_data}
 # Read dissolved gas data file
-dg <- read.table(file = paste0(localPath, 
-                               "/Environmental Protection Agency (EPA)/",
-                               "ORD NLA17 Dissolved Gas - Documents/",
-                               "inputData/nla17gasDataAggregated_2021-02-01.txt"),
+dg <- read.table(file = "./../inputData/nla17gasDataAggregated_2021-02-01.txt",
                  header = TRUE, sep = "\t", as.is = TRUE) %>%
   filter(!grepl("AK", site.id)) # omit Alaska sites
+
+# Read simulated saturation ratio values (See DG_sensitivity_to_measurement_error.Rmd, Saturation ratio/Standard GC)
+#S_a_me_abs <- readRDS("inputData/S_a_me_abs.RDS")
 ```
@@
 ### Implement unit conversion and create source/sink column:
-``` {r}
+``` {r mutate_data}
 dg <- dg %>%
@@
          ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
-         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink")) %>%
+         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink")
+         ) %>%
   # remove fields no longer needed
@@
 
-```{r}
+```{r mutate_data_2}
 dg <- dg %>% 
@@
 
-## Emission rate for population estimate
-Tom Kincaid created a file to be used for population estimates.  Lets add our emission rates estimates to this file.  We can then use Tom's code to construct population estimates of emission rates.
-```{r}
-cdf <- read.csv(paste0(localPath,
-            "/Environmental Protection Agency (EPA)/",
-            "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/populationEstimates/",
-            "NLA17_DissolvedGases_forPopEst.csv"))
-
-# although data I provided Tom (nla17gasAggregatedForPopEst2019-10-23.txt) included data from both visits,
-# data used for population estimates included only visit one.  I don't know rationale for this, but will
-# do the same with the revised file here.  Also need to strip out air samples.
-unique(cdf$VISIT_NO) # 1
-dg.1 <- dg %>% filter(visit.no == 1, sample.source == "DG")
-
-# even after filtering, dg has 85 more observations than cdf
-dim(cdf) # 1005, 81
-dim(dg.1) # 1090, 306
-
-# which observations in filtered dg, not in cdf?
-# many of these are handpicked (HP) sites.  Others not sure
-# why are excluded
-dg.1 %>% filter(!(dg.1$site.id %in% cdf$SITE_ID)) %>%
-  select(site.id) %>%
-  as_tibble() %>% print(n=Inf)
-
-# using left_join to exclude sites from dg.1 that Tom did
-# not include in cdf (see above).  Presumably the excluded sites
-# are not probability sites
-cdf.dg <- left_join(cdf, 
-                    select(dg.1, site.id, visit.no, 
-                           e.ch4.umol.d , e.co2.umol.d, e.n2o.nmol.d, 
-                           f.ch4.km.d, f.co2.km.d, f.n2o.m.d) %>%
-                      # rename for consistency with Tom's data
-                      rename_all(~stringr::str_to_upper(.)) %>% 
-                      rename_all(~stringr::str_replace_all(., '\\.', '_')))
-
-dim(cdf.dg) # 1005, 87 same as cdf above, so good
-
-write.csv(cdf.dg, 
-          paste0(localPath,
-            "/Environmental Protection Agency (EPA)/",
-            "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/populationEstimates/",
-            "NLA17_DissolvedGases_forPopEst.jb.csv"))
-
-```
-
 ## Write out data
 Write out final .RData object.
-```{r}
-save(dg, file = paste0(localPath,
-                       "/Environmental Protection Agency (EPA)/",
-                       "ORD NLA17 Dissolved Gas - Documents/",
-                       "inputData/dg.", Sys.Date(), ".RData"))
-
+```{r save_data_munged}
+save(dg, file = "./../inputData/dg.RData")
 ```

### 2025-03-03 manuscript/manuscript_support.R

@@ -5,7 +5,7 @@
 
 # SETUP----------------------
 ## Identify local path for each user
-localPath <- Sys.getenv("USERPROFILE")
+#localPath <- Sys.getenv("USERPROFILE")
 
 ## load libraries
 # DMAP not under renv control.  Should probably specify same
@@ -37,7 +37,8 @@ localPath <- Sys.getenv("USERPROFILE")
 library(sf) # spatial data
 library(tidyverse) # dplyr, ggplot
 library(janitor) # clean names
-library(USAboundaries) # state boundaries
+#library(USAboundaries) # state boundaries (removed from CRAN)
+library(usmap) # state boundaries 
 library(tictoc) # processing time
 library(ggpubr) # multiple plots
 library(ggallin) # psuedlolog transformation for negative values
@@ -45,16 +46,9 @@ library(tidybayes) # prop flux by size figure
 
 # DATA---------------
 ## load sample data (dg object)----
-if(localPath == "") { # if DMAP, then
-  load("inputData/dg.2025-02-14.RData")
-} else { # if not DMAP, then...
-  load(paste0(localPath,  # object name dg
-              "/Environmental Protection Agency (EPA)/",
-              "ORD NLA17 Dissolved Gas - Documents/",
-              "inputData/dg.2025-02-14.RData"))
-}
+load("./inputData/dg.RData")
 
-save(dg, file = "manuscript/manuscript_files/dg.rda")  # save dg file to ms folder for faster knitting
+save(dg, file = "./manuscript/manuscript_files/dg.rda")  # save dg file to ms folder for faster knitting
 
 # To enable spatial analysis of the data, the dataframe will be converted to a 
 # 'simple features' (sf) object.
@@ -67,17 +61,7 @@ dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"),
 
 ## load ecoregions----
 # read in ecoregion polygons
-if(localPath == ""){ # if DMAP, then
-  ecoR <- st_read(dsn = "inputData",
-                  layer = "aggr_ecoregions_simple",
-                  quiet = TRUE)
-} else { # if not DMAP, then...
-  ecoR <- st_read(dsn = paste0(localPath, 
-                               "/Environmental Protection Agency (EPA)/",
[…]

### 2025-03-03 scripts/dgIndicatorAnalysis.Rmd
Updated uncertainty calculations for dissolved N2O
Updates to dgindicatorAnalysis

@@
 output:
-  html_document:
-    df_print: paged
-    toc: true
-    toc_float: true
-    fig_caption: yes
-    depth: 2
-    number_sections: true
-    code_folding:  hide
+  github_document:
+    number_sections: TRUE
+    df_print: "tibble"
+    math_method: 
+      engine: webtex
+    #  url: https://latex.codecogs.com/svg.image?
+    html_preview: TRUE
+    keep_html: TRUE
 editor_options:
-  chunk_output_type: console
+  chunk_output_type: inline
+  markdown: 
+    wrap: 72
 ---
 
-```{r results='hide', message=FALSE, warning=FALSE}
+```{r setup, results='hide', message=FALSE, warning=FALSE}
 # load libraries
@@
 # Identify local path for each user
-localPath <- Sys.getenv("USERPROFILE")
+#localPath <- Sys.getenv("USERPROFILE")
 
@@
 
-```{r}
+```{r load_data}
 # Load data
- load(paste0(localPath,
-            "/Environmental Protection Agency (EPA)/",
-            "ORD NLA17 Dissolved Gas - Documents/",
-            "inputData/dg.2021-02-01.RData"))
-
-```
+load(file = "../inputData/dg.RData")
+```
+
+## Analytical uncertainty
 
@@
 
-```{r}
+```{r responses}
 # These are the main response variables.
@@
 # (12/12/2019), then modified to accomdate the gas emission rate and flux variables.  See CDF_Estimates_Jb.Rmd.
-pop <- read.csv(file = paste0(localPath, 
-                              "/Environmental Protection Agency (EPA)/",
-                              "ORD NLA17 Dissolved Gas - Documents/",
-                              "inputData/populationEstimates/",
-                              "NLA_2017_Percentile_Estimates_Local_Mean_JB20201202.csv"),
+pop <- read.csv(file = "../inputData/populationEstimates/NLA_2017_Percentile_Estimates_Local_Mean_JB20201202.csv",
                 as.is = TRUE) %>%

### 2025-03-03 scripts/n2oSat.R
Removing files from Git control

@@ -1,28 +1,0 @@
-n2oSat <- function(inputFile, waterTemp, airN2O, baro) {
-
-# Input variables
-# inputFile: Name of the data fram containing the information needed to calculate the dissolved gas concentrations. 
-# waterTemp: Temperature of the waterbody when sampled [celsius]
-# airN2O: Concentration of nitrous oxide in atmosphere [ppmv]
-# baro: Barometric pressure at the time of equilibration [kPa]
-
-##### Constants #####
-cGas<-8.3144598 #universal gas constant (J K-1 mol-1)
-cKelvin <- 273.15 #Conversion factor from Kelvin to Celsius
-cPresConv <- 0.000001 # Constant to convert mixing ratio from umol/mol (ppmv) to mol/mol. Unit conversions from kPa to Pa, m^3 to L, cancel out.
-cT0 <- 298.15#Henry's law constant T0
-
-#Henry's law constants and temperature dependence from Sander (2015) DOI: 10.5194/acp-15-4399-2015
-ckHN2O <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025
-cdHdTN2O <- 2700 #K, range: 2600 - 3600
-
-
-##### Calculate dissolved gas concentration at 100% saturation ##### 
-
-# 100% saturation occurs when the dissolved gas concentration is in equilibrium
-# with the atmosphere.
-satN2O <- (ckHN2O * exp(cdHdTN2O*(1/(waterTemp + cKelvin) - 1/cT0))) * airN2O * baro * cPresConv
-
-return(satN2O)
-}
-

### 2025-03-03 scripts/nlaSampleFrameWind.R

@@ -1,167 +1,0 @@
-# THIS SCRIPT IS TO CALCULATE WIND DATA FOR ALL ~600,000 WATERBODIES IN THE 
-# NLA SAMPLE FRAME.  THESE DATA WILL BE USED TO CALCULATE A K VALUE FOR
-# EACH SITE.  WHEN COMBINED WITH ROY'S MODEL PREDICTIONS OF DISSOLVED N2O 
-# WILL ENABLE EMISSION RATE PREDICTION FOR EACH SITE.  THESE PREDICTIONS CAN
-# BE USED TO ESTIMATE POPULATION LEVEL MEAN AND VARIANCE.  
-# THIS CODE SHOULD BE RUN ON VM.  SCRIPT IS NOT YET COMPLETE.
-# SEE 'readWind.R' IN THE NLA PROJECT FOR A VERSION TAILORED TO THE 1200
-# NLA OBSERVATIONS.  THE NLA VERSION UTILIZES LARGE DATA FRAMES, BUT CAN
-# BE RUN ON LAPTOP.  VERSION BELOW HAS A STRONGER RELIANCE ON LISTS, BUT
-# THE DATAFRAME APPROACH MIGHT WORK ON VM.  NEED TO DECIDE HOW WE ARE GOING
-# TO AGGREGATE DATA.  OPTIONS INCLUDE 1) INDEX PERIOD MEAN PER SITE, 2) 
-# INDEX PERIOD MEAN OVER DAYLIGHT HOURS, AND 3) MEAN DAYLIGHT HOURS PER 
-# DAY.  OPTION 3 WOULD ENABLE CALCULATING A K AND EMISSION RATE PER DAY
-# FOR EACH SITE.
-
-
-#R 4.0.3
-library(tidyverse)
-library(sf)
-library(ncdf4)
-library(ncdf4.helpers)
-library(suncalc)
-library(lubridate)
-library(pbapply)
-
-# ENFORCE EPA FORMAT ON NLA NAMES
-toEPA <- function(X1){
-  names(X1) = tolower(names(X1))
-  names(X1) = gsub(pattern = "_", replacement = ".", x = names(X1))
-  X1
-}
-
-
-# NLA17 sample frame--------------------------
-sampFrm <- read.csv(file = paste0(Sys.getenv("USERPROFILE"), 
-                                    "/Environmental Protection Agency (EPA)/",
-                                    "ORD NLA17 Dissolved Gas - Documents/",
[…]

### 2025-03-04 manuscript/manuscript_support.R
Updated renv

@@ -29,15 +29,12 @@
 #   if(!require("pbkrtest")) install.packages("pbkrtest") 
 #   if(!require("ggpubr")) install.packages("ggpubr")
 #   if(!require("ggallin")) install.packages("ggallin")
-#   if(!require("USAboundaries")) devtools::install_version("USAboundaries", "0.4.0", 
-#                                                 repos = "https://cran.r-project.org/")
 #   if(!require("tidybayes")) install.packages("tidybayes")
 # }
 
 library(sf) # spatial data
 library(tidyverse) # dplyr, ggplot
 library(janitor) # clean names
-#library(USAboundaries) # state boundaries (removed from CRAN)
 library(usmap) # state boundaries 
 library(tictoc) # processing time
 library(ggpubr) # multiple plots
@@ -81,11 +78,6 @@ cols <- c("Coastal Plains" = "orange1",
           "Western Mountains" = "saddlebrown",
           "Xeric" = "lightskyblue4")
 
-
-## Load state boundaries----
-#states <- USAboundaries::us_states() %>%
-#  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
-#  st_transform(5070) # convert to CONUS Albers
 
 states <- usmap::us_map() %>%
   dplyr::filter(!full %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%

### 2025-03-06 manuscript/manuscript_file.Rmd
Edits made in DMAP repo required to knit manuscript

@@
 for(i in 1:N){
-    f[i] <- paste0("/manuscript/manuscript_files/",
+    f[i] <- paste0("./manuscript/manuscript_files/",
     objs[i], ".rda")
@@
   # see above for .Rmd working directory
-    source("manuscript/manuscript_support.R", local = knitr::knit_global())
+    source("./manuscript/manuscript_support.R", local = knitr::knit_global())
 }
@@
 for(i in 1:N){
-    f[i] <- paste0("/manuscript/manuscript_figures/",
+    f[i] <- paste0("./manuscript/manuscript_figures/",
     objs[i])

### 2025-03-06 scripts/DG_sensitivity_to_measurement_error.Rmd
Edits made in DMAP repo required to knit manuscript

@@
 # read in ecoregion polygons
-ecoR <- st_read(dsn = paste0(localPath, 
-                             "/Environmental Protection Agency (EPA)/",
-                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
-                layer = "aggr_ecoregions_simple",
-                quiet = TRUE)
+ecoR <- st_read("/vsizip/./../inputData/aggr_ecoregions_simple.zip", quiet = TRUE)
 

### 2025-03-06 scripts/Design_Based_Estimates_N2O.Rmd
Edits made in DMAP repo required to knit manuscript

@@
 
-```{r save_to_drive_1, eval=FALSE, include=TRUE}
+```{r save_to_drive_1, eval=TRUE, include=TRUE}
 write.csv(mean_dissolved, file = "./../inputData/populationEstimates/Survey_Ests_N20_dissolved.csv")
@@
 
-```{r save_to_drive_2, eval=FALSE, include=TRUE}
+```{r save_to_drive_2, eval=TRUE, include=TRUE}
 write.csv(mean_sat_ratio, file = "./../inputData/populationEstimates/Survey_Ests_N20_sat.csv")

### 2025-03-07 manuscript/manuscript_file.Rmd
Cut 500 words from manuscript to meet journal word limit

@@
 ---
-title: "Widespread nitrous oxide undersaturation in U.S. lakes and reservoirs"
+title: "Pervasive nitrous oxide undersaturation in U.S. lakes and reservoirs"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 (CONUS).
-We found that N~2~O undersaturation was widespread through the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of lakes functioning as N~2~O sinks.
+We found that N~2~O undersaturation was pervasive through the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of lakes functioning as N~2~O sinks.
 The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate concentration, waterbody surface area, and water temperature.
 Despite working with the largest aquatic N~2~O dataset to date, our national-scale estimate of N~2~O emissions from CONUS lakes is poorly constrained, with a 95% credible interval ranging from net uptake to net emission (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
-Widespread N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS lakes are a net source or sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source.**
+Pervasive N~2~O undersaturation in CONUS waterbodies and a national-scale emission estimate that is too uncertain to determine if CONUS lakes are a net source or sink of N~2~O highlight the need to revisit N~2~O models which presume surface waters are a N~2~O source.**
 
@@
 
-Lakes, reservoirs, and ponds (lakes hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], a potent greenhouse gas and contributor to stratospheric ozone destruction.
+Lakes, reservoirs, and ponds (lakes hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], the dominant ozone-depleting substance emitted in the 21st century[@Ravishankara_etal_2009].
 It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across lakes in Canada [@Soued_etal_2016; @Webb_etal_2019], the United States [@Aho_etal_2023], and Finland [@Kortelainen_etal_2020] have called this assumption into question.
@@
 1)**.
-N~2~O concentrations were close to equilibrium values in some of these lakes, however, and after accounting for measurement uncertainty in both dissolved and equilibrium N~2~O concentration (See SI), we can confidently report that a minimum of `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent of the sampled lakes were functioning as sinks, but the number could be as high as `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent.
+N~2~O concentrations were close to equilibrium values in some of these lakes, however, and after accounting for measurement uncertainty in both dissolved and equilibrium N~2~O concentration (See SI), we can confidently report that a minimum of `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent of the sampled lakes were functioning as sinks.
 When accounting for measurement error, `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "undetermined") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent had a dissolved N~2~O concentration that was indistinguishable from the equilibrium concentration and only `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "source") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent were supersaturated **(Fig.
@@
 
-The Bayesian hierarchical model predicted that the majority lakes in the POI (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% credible interval (CI): `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) functioned as net N~2~O sinks during the summer months.
-The Western Mountains ecoregion had the smallest proportion of lakes predicted as N~2~O sinks (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "WMT") %>% pull(UCL_prop_sat) * 100`%) and the Northern Plains had the greatest (posterior mean = `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(post_m_prop_sat) * 100`%, 95% CI: `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(LCL_prop_sat) * 100` - `r ecoreg_stats %>% filter(WSA9 == "NPL") %>% pull(UCL_prop_sat) * 100`%), although the 95% credible intervals of the predicted distributions overlap across all ecoregions **(Table 1)**.
+The Bayesian hierarchical model predicted that the majority lakes in the POI (posterior mean = `r national_stats$post_m_prop_sat * 100`%, 95% credible interval (CI): `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) functioned as net N~2~O sinks during the summer months, a surprisingly high estimate, particularly for nitrogen rich ecoregions that are managed for agricultural production.
+This finding runs counter to the long-standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lakes, ponds, streams, and rivers can function as N~2~O sinks[@Webb_etal_2019;@Kortelainen_etal_2020;@Soued_etal_2016;@Aho_etal_2023].
+While these latter surveys were conducted at regional scales or had relatively few observtion sites, the national-scale results presented here provide the strongest evidence to date that most lakes function as N~2~O sinks during the summer months.
 
@@
 Point size indicates the dissolved N~2~O concentration.](manuscript_figures/figure1.png)
-
-The estimated extent of N~2~O undersaturation in CONUS lakes was surprisingly high, particularly for nitrogen rich ecoregions that are managed for agricultural production.
-This finding runs counter to the long-standing paradigm that surface waters are a source of N~2~O to the atmosphere [@Mosier_etal_1998] and joins other recent reports that lakes can function as N~2~O sinks.
-Webb et al.[@Webb_etal_2019] found that 67% of sampled reservoirs in an agricultural region of Canada were undersaturated with N~2~O during the summer months and Kortelainen et al.[@Kortelainen_etal_2020] found that 29% of surveyed boreal lakes in Finland were N~2~O sinks during the summer.
-An investigation of boreal systems in Canada found that 40% of rivers, lakes, and ponds sampled during the summer were functioning as N~2~O sinks[@Soued_etal_2016] and all 34 lotic and lentic sites in the US National Ecological Observatory Network (NEON) functioned as N~2~O sinks at least once during 1.5 - 6 years of monitoring[@Aho_etal_2023].
-While these latter surveys were conducted at regional scales or had relatively few observtion sites, the national-scale results presented here provide the strongest evidence to date that most lakes function as N~2~O sinks during the summer months.
-This pattern held across all 9 major ecoregions in the CONUS, despite the tremendous variation in climate, geology, waterbody morphology, and land use.
 
@@
 
-Despite widespread N~2~O undersaturation across the POI, estimated means for N~2~O saturation ratio (ratio of measured to equilibrium concentration) were indistinguishable from a value of one in all ecoregions **(Table 1, Fig.
+Despite pervasive N~2~O undersaturation across the POI, estimated means for N~2~O saturation ratio (ratio of measured to equilibrium concentration) were indistinguishable from a value of one in all ecoregions **(Table 1, Fig.
 2)**.
-The apparent contradiction of widespread N~2~O undersaturation and average dissolved N~2~O concentrations close to equilibrium can be attributed to the prominent right skew in the predicted distribution of dissolved N~2~O concentrations across the POI **(Fig.
+The apparent contradiction of pervasive N~2~O undersaturation and average dissolved N~2~O concentrations close to equilibrium can be attributed to the prominent right skew in the predicted distribution of dissolved N~2~O concentrations across the POI **(Fig.
 2)**.
@@
 3)**.
-This finding mirrors reports from a survey of farm ponds in Canada where N~2~O was found to be associated with dissolved inorganic nitrogen (DIN) concentration, but the effect of DIN was conditional upon thermal stratification strength[@Webb_etal_2019].
 These complex interactions may explain contrasting literature reports regarding the role of NO~3~ in regulating N~2~O.
@@
 The reduction of NO~3~ to N~2~ can be incomplete, however, resulting in the release of N~2~O to the water column.
-Incomplete denitrification often occurs when NO~3~ is relatively abundant [@Beaulieu_etal_2015; @Beaulieu_etal_2014], likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway, yielding less than 20% of the total energy yield of complete denitrification [@Koike_etal_1975; @Richardson_etal_2009].
+Incomplete denitrification often occurs when NO~3~ is relatively abundant [@Beaulieu_etal_2015; @Beaulieu_etal_2014], likely because the final reduction of N~2~O to N~2~ is the least energetic reaction in the denitrification pathway[@Koike_etal_1975; @Richardson_etal_2009].
 When NO~3~ is limiting, however, denitrification can resort to scavenging N~2~O from the water column and reducing it to N~2~, thereby functioning as an N~2~O sink.
@@
 Nitrification may be more prevalent than denitrification in the predominantly well oxygenated surface waters where the N~2~O samples were collected because denitrification is an anaerobic process.
-However, denitrification can occur in anoxic microsites within particulate matter suspended in an aerobic water column [@Broman_etal_2021] and can be transported from anoxic sediments to the water surface via advection and diffusion [@DelSontro_etal_2018].
 Thus the N~2~O - NO~3~ relationship may be reflective of some combination of N~2~O production/consumption via denitrification and N~2~O production via nitrification.
@@
 At low NO~3~ levels, temperature was estimated to have a relatively minor effect on N~2~O saturation, possibly because N~2~O production/consumption was limited by substrate rather than temperature.
-The effect of temperature was more pronounced at greater NO~3~ levels, potentially reflecting the stimulatory effect of higher temperature on biological N~2~O production.
+The effect of temperature was more pronounced at greater NO~3~ levels when substrate limitation was less severe.
 Elevated water temperature may serve to magnify rates of N~2~O production that may be largely dictated by NO~3~ availability.
 
-Lake size was also related to N~2~O saturation ratio, but the relationship was dependent on NO~3~ concentration **(Fig.
+Lake size (surface area) was also related to N~2~O saturation ratio, but the relationship was dependent on NO~3~ concentration **(Fig.
 3)**.
@@
 For example, N~2~O inputs from streams or groundwater[@Wang_etal_2023] may constitute a substantial N~2~O subsidy to small lakes but are more likely be diluted in larger lakes with greater volumes.
-Small lakes also tend to have lower water volume to benthic surface area ratios[@Cael_etal_2017], and therefore stronger connections between sediment, an area of active N~2~O processing, and the air-water interface where our samples were collected.
 Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep waterbodies, resulting in greater rates of N~2~O production or consumption in small lakes.
@@
 3C,D)**.
-Despite widespread N~2~O undersaturation predicted across the POI, the estimate for the national average N~2~O emission rate was centered over positive values (posterior mean = `r national_stats$post_m_mean_en2o`, 95% CI: `r national_stats$LCL_mean_en2o` - `r national_stats$UCL_mean_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)** and clearly exceeded the negative estimate for the national median (posterior mean = `r national_stats$post_m_median_en2o`, 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^).
+Despite pervasive N~2~O undersaturation predicted across the POI, the estimate for the national average N~2~O emission rate was centered over positive values (posterior mean = `r national_stats$post_m_mean_en2o`, 95% CI: `r national_stats$LCL_mean_en2o` - `r national_stats$UCL_mean_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)** and clearly exceeded the negative estimate for the national median (posterior mean = `r national_stats$post_m_median_en2o`, 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^).
 This departure from the median was due to relatively rare lakes with very high emission rates pulling the mean towards the direction of positive skew **(SI Fig.
@@
 First, lakes in the smaller size classes constituted only `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(cumulative = cumsum(area_ha), distribution = cumulative / sum(area_ha), delta50 = abs(50-area_ha)) %>% filter(delta50 == min(delta50)) %>% pull(distribution) %>% round(3) * 100`% of total waterbody surface area, thereby limiting their N~2~O flux.
-Second, widespread N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the POI.
+Second, pervasive N~2~O uptake in these size classes largely offset the very high emission rates predicted to occur in a minority of the POI.
 Lakes in the largest size class (\> 50 Ha) represented only `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(number = 1:nrow(.), proportion = number / nrow(.)) %>% filter(abs(50-area_ha) == min(abs(50-area_ha))) %>% pull(proportion) %>% round(2) * 100 * -1 + 100`% of the total number of lakes in the population but had `r sframe %>% select(area_ha) %>% arrange(area_ha) %>% mutate(cumulative = cumsum(area_ha), distribution = cumulative / sum(area_ha), delta50 = abs(50-area_ha)) %>% filter(delta50 == min(delta50)) %>% pull(distribution) %>% round(3) * 100 * -1 +100`% of total lake surface area.
@@
 
-Despite the estimated widespread N~2~O undersaturation among CONUS lakes, credible intervals for ecoregion-level flux estimates included zero for all but three ecoregions **(Table 1)**.
-Thus, status as a net source versus sink of N~2~O was uncertain for most ecoregions.
-However, the estimates for the Southern Appalachians, Southern Plains, and Temperate Plains included only positive values, suggesting they were likely to have functioned as net N~2~O sources.
-We were also unable to infer any clear differences in total flux between ecoregions.
-This result may sound surprising given the large differences in environmental conditions across these ecoregions, but it should not be mistaken as inferring "no differences" as it primarily emphasizes the challenges in making sufficiently precise estimates at this scale; even with unprecedented data coverage, a robust survey methodology, and relatively advanced modeling techniques.
-
-The estimate for flux across the POI was centered over positive values (posterior mean = `r national_stats %>% select(post_m_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^); however, the credible interval suggested a good deal of uncertainty and included zero (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^).
-Again, we could not confidently infer the extent to which CONUS lakes, overall, were a net source or sink of N~2~O.
+The estimate for flux across the POI was centered over positive values (posterior mean = `r national_stats %>% select(post_m_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^) but the credible interval included zero (`r national_stats %>% select(LCL_fn2o) %>% pull() %>% round(., 0)` - `r national_stats %>% select(UCL_fn2o) %>% pull() %>% round(., 0)` metric tons N~2~O year^-1^), thus we are unable to infer the extent to which CONUS lakes, overall, were a net source or sink of N~2~O.
 Lauerwald et al [@Lauerwald_etal_2019] used a mechanistic model to estimate that CONUS lakes emit 8,425 metric tons N~2~O year^-1^.
@@
 This estimate is two orders of magnitude greater than the posterior mean of our estimate and suggests the mechanistic model does not properly describe N~2~O consumption.
-Given our estimate of widespread N~2~O consumption in CONUS lakes, failure to properly describe this dynamic in aquatic N~2~O models may yield high-biased emission estimates.
+Given our estimate of pervasive N~2~O consumption in CONUS lakes, failure to properly describe this dynamic in aquatic N~2~O models may yield high-biased emission estimates.
 
@@
 
-Our research is the first to suggest that most lakes and reservoirs in the CONUS function as N~2~O sinks during the summer.
-Our observations of dissolved N~2~O were predicted reasonably well using a regression model with interactions between observed water temperature, NO~3~ concentration, and waterbody size (see SI).
+Our observations of dissolved N~2~O were well predicted using a regression model with interactions between observed water temperature, NO~3~ concentration, and waterbody size (see SI).
 These modeling results suggested that NO~3~ levels may play a role in determining whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink may be exaggerated with increasing water temperature.
@@
 The N~2~O flux from large lakes is poorly constrained, however, because relatively small uncertainties in dissolved and equilibrium N~2~O concentrations translate to large uncertainties in flux when propagated across the surface of large lakes.
-While our sample included some of the largest lakes in the CONUS, efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O, which can be important in large waterbodies[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement and analytical uncertainty, which can propagate to uncertainties in flux estimates.
+Efforts to improve the accuracy of the lentic N~2~O flux estimate should include an assessment of intra-waterbody spatial variation in N~2~O[@Salk_etal_2019; @Borges_etal_2023], as well as careful accounting of measurement uncertainty.
 The measurement methods employed in this study have a 95% credible interval of 0.11 (unitless) on the N~2~O saturation ratio.
-The 95% credible interval can be halved by employing gas chromatography optimized for quantifying N~2~O in the range of values observed in this study [@Hall_2024] and decreased by an order of magnitude using membrane inlet mass spectrometry[@Genthner_etal_2013; @Speir_etal_2023].
+The 95% CI can be halved by employing gas chromatography optimized for quantifying N~2~O in the range of values observed in this study [@Hall_2024] and decreased by an order of magnitude using membrane inlet mass spectrometry[@Genthner_etal_2013; @Speir_etal_2023] **(SI)**.
 
@@
 For example, reservoirs have been shown to function as N~2~O sources in early summer when NO~3~ is replete, but switch to N~2~O sinks in later summer when the NO~3~ supply has been exhausted [@Beaulieu_etal_2014].
-Most small lakes in the CONUS were estimated to have functioned as N~2~O sinks during the summer, but a small subset were also N~2~O emission hotspots.
-It is unknown if the high emitting lakes maintain those emissions throughout the summer, or if they convert to sinks later in the season.
 A better understanding of the temporal evolution of N~2~O production and consumption is an important research need to determine the extent to which small lakes can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.

### 2025-03-07 manuscript/manuscript_file.Rmd
Updates to manuscript files for more efficient knit

@@
 # vector of figure names
-objs <- c("cond_effect_no3.tif",
+objs <- c("cond_effect_no3.tiff",
           "figure1.png",
@@
   else
-    print(paste0("missing figure: ", f[[1]]))
+    print(paste0("missing figure: running manuscript_support.R ", f[[1]]))
+    source("./manuscript/manuscript_support.R", local = knitr::knit_global())
  }
@@
 D) The relationship between N~2~O saturation and NO~3~ level conditional on (log) waterbody surface area (5 quantiles).
-Horizontal dashed lines in each panel reflect an N~2~O saturation value of one.](manuscript_figures/cond_effect_no3.tif)
+Horizontal dashed lines in each panel reflect an N~2~O saturation value of one.](manuscript_figures/cond_effect_no3.tiff)
 

### 2025-03-07 manuscript/manuscript_support.R

@@ -41,12 +41,14 @@ library(ggpubr) # multiple plots
 library(ggallin) # psuedlolog transformation for negative values
 library(tidybayes) # prop flux by size figure
 
+
+if(!("figure1.png" %in% list.files("manuscript/manuscript_figures"))) {
+
 # DATA---------------
 ## load sample data (dg object)----
 load("./inputData/dg.RData")
-
 save(dg, file = "./manuscript/manuscript_files/dg.rda")  # save dg file to ms folder for faster knitting
-
+  
 # To enable spatial analysis of the data, the dataframe will be converted to a 
 # 'simple features' (sf) object.
 # Define coordinates
@@ -82,8 +84,17 @@ cols <- c("Coastal Plains" = "orange1",
 states <- usmap::us_map() %>%
   dplyr::filter(!full %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
   sf::st_transform(5070) # convert to CONUS Albers
+} # fig 1
 
-
+if(!("ecoreg_stats.rda" %in% list.files("manuscript/manuscript_files"))) {
+  if(!("national_stats.rda" %in% list.files("manuscript/manuscript_files"))) {
+    if(!("figure2.tiff" %in% list.files("manuscript/manuscript_figures"))) {
+      if(!("n2oFluxAndEmissionRateVsContinuousArea.tiff" %in% list.files("manuscript/manuscript_figures"))) {
+        if(!("n2oStarBySize.tiff" %in% list.files("manuscript/manuscript_figures"))) {
+          if(!("SIfigure1.tiff" %in% list.files("manuscript/manuscript_figures"))) {
+            if(!("FluxBySize.rda" %in% list.files("manuscript/manuscript_files"))) {
+              if(!("em_min_max.rda" %in% list.files("manuscript/manuscript_files"))) {
+      
 ## load posterior predictions from modeling----
 #tic() # 7.6min on DMAP.  5 minutes on Dell Precision workstation
 load("./inputData/all_predictions.rda")
@@ -189,7 +200,16 @@ all_predictions_ms  <- all_predictions_ms %>%
          n2osat,
          e.n2o.mg.m2.d,
          f.n2o.Mg.y)
+              } # ecoreg_stats
+            } # national_stats
+          } #fig2
+        } # flux
+      } # n2o star
+    } # SI figure
+  } # FluxBySize
+} # emissions min/max
 
[…]

### 2025-03-17 manuscript/manuscript_file.Rmd
Minor revisions in response to Matt Heberlings BC review.

@@
 title: "Pervasive nitrous oxide undersaturation in U.S. lakes and reservoirs"
+author:
+  - Jake J. Beaulieu, USEPA, Office of Research and Development
+  - Roy Martin, USEPA, Office of Research and Development
+  - Michael G. McManus, USEPA, Office of Research and Development
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-**Lakes, ponds, and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere but recent evidence of N~2~O uptake across a broad range of lakes have called the accuracy of emission estimates into question.
+**Lakes and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere but recent evidence of N~2~O uptake across a broad range of lakes have called the accuracy of emission estimates into question.
 Here we use a new national-scale dataset on dissolved N~2~O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` waterbodies in the conterminous U.S.
@@
 
-Lakes, reservoirs, and ponds (lakes hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], the dominant ozone-depleting substance emitted in the 21st century[@Ravishankara_etal_2009].
+Lakes and reservoirs (lakes hereafter) are considered globally significant sources of nitrous oxide (N~2~O)[@DelSontro_etal_2018a], the dominant ozone-depleting substance emitted in the 21st century[@Ravishankara_etal_2009].
 It is generally assumed that N~2~O emissions from surface waters scale with inorganic nitrogen loading [@Mosier_etal_1998] or productivity [@DelSontro_etal_2018], but recent reports of N~2~O uptake across lakes in Canada [@Soued_etal_2016; @Webb_etal_2019], the United States [@Aho_etal_2023], and Finland [@Kortelainen_etal_2020] have called this assumption into question.
@@
 (CONUS).
-Samples sites were selected using a generalized random-tessellation stratified survey design [@Stevens_Olsen_2004] and the observational data were used to develop a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS lakes.
-Combining the largest freshwater N~2~O dataset with a hierarchical model enabled us to estimate the prevalence of N~2~O uptake in CONUS lakes for the first time and to identify important environmental controls on N~2~O biogeochemistry.
+We used the data to train a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` CONUS lakes.
+The data and model enabled us to estimate the prevalence of N~2~O uptake in CONUS lakes for the first time and to identify important environmental controls on N~2~O biogeochemistry.
 
@@
 Predictor variables for the model were restricted to those required to generate predictions that reproduced the observed N~2~O measures in an out-of-sample dataset and were available for the POI.
-See Methods and SI for a detailed discussion of variable selection.
+See Methods and **SI_1** for a detailed discussion of variable selection.
 
@@
 1)**.
-N~2~O concentrations were close to equilibrium values in some of these lakes, however, and after accounting for measurement uncertainty in both dissolved and equilibrium N~2~O concentration (See SI), we can confidently report that a minimum of `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent of the sampled lakes were functioning as sinks.
+Nitrous oxide concentrations were close to equilibrium values in some of these lakes, however, and after accounting for measurement uncertainty in both dissolved and equilibrium N~2~O concentration (**SI_2**), we can confidently report that a minimum of `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent of the sampled lakes were functioning as sinks.
 When accounting for measurement error, `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "undetermined") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent had a dissolved N~2~O concentration that was indistinguishable from the equilibrium concentration and only `r round(((dg %>% dplyr::filter(n2o.src.snk.error == "source") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 0)` percent were supersaturated **(Fig.
@@
   select(WSA9_NAME, total_sa, 'Mean N2O saturation ratio', 'N2O undersaturation extent', 'N2O emission rate', 'N2O flux') %>%
-  rename('spatial domain' = WSA9_NAME,
+  rename('Spatial domain' = WSA9_NAME,
          'waterbody surface area' = total_sa)
@@
 flextable(table1_dat) %>%
-  set_header_labels('waterbody surface area' = "total waterbody surface area\n(hectares)",
+  set_header_labels('waterbody surface area' = "Total waterbody surface area\n(hectares)",
                     'N2O undersaturation extent' = "N2O undersaturation extent\n(% lakes)",
@@
 Several studies have reported positive correlations between NO~3~ and N~2~O in lakes [@Kortelainen_etal_2020; @Yang_etal_2015] whereas others have found no relationship [@Soued_etal_2016; @Whitfield_etal_2011].
-It is possible that the NO~3~ \~ N~2~O relationship wasn't detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
+It is possible that the NO~3~ \~ N~2~O relationship was not detected in the latter studies because it was confounded by interactions with temperature, lake size, thermal stratification, or other environmental variables.
 
@@
 B) The relationship between N~2~O saturation and NO~3~ level conditional on surface water temperature.
-The NO~3~ effect was estimated at 5 quantiles of the surface temperatures observed (10%, 25%, 50%, 75%, 90%).
+The NO~3~ effect was estimated at five quantiles of the surface temperatures observed (10%, 25%, 50%, 75%, 90%).
 C) The relationship between N~2~O saturation and waterbody surface area (log-transformed hectares) conditional on NO~3~ level.
-D) The relationship between N~2~O saturation and NO~3~ level conditional on (log) waterbody surface area (5 quantiles).
+D) The relationship between N~2~O saturation and NO~3~ level conditional on (log) waterbody surface area (five quantiles).
 Horizontal dashed lines in each panel reflect an N~2~O saturation value of one.](manuscript_figures/cond_effect_no3.tiff)
@@
 Lake size and shape can affect dissolved gas concentration through several mechanisms.
-For example, N~2~O inputs from streams or groundwater[@Wang_etal_2023] may constitute a substantial N~2~O subsidy to small lakes but are more likely be diluted in larger lakes with greater volumes.
+For example, N~2~O inputs from streams or groundwater[@Wang_etal_2023] may constitute a substantial N~2~O subsidy to small lakes but are more likely to be diluted in larger lakes with greater volumes.
 Furthermore, sediment may warm faster in small and shallow lakes, whereas thermal stratification may substantially delay sediment warming in large and deep waterbodies, resulting in greater rates of N~2~O production or consumption in small lakes.
@@
 Despite pervasive N~2~O undersaturation predicted across the POI, the estimate for the national average N~2~O emission rate was centered over positive values (posterior mean = `r national_stats$post_m_mean_en2o`, 95% CI: `r national_stats$LCL_mean_en2o` - `r national_stats$UCL_mean_en2o` mg N~2~O m^-2^ d^-1^) **(Table 1)** and clearly exceeded the negative estimate for the national median (posterior mean = `r national_stats$post_m_median_en2o`, 95% CI: `r national_stats$LCL_median_en2o` - `r national_stats$UCL_median_en2o` mg N~2~O m^-2^ d^-1^).
-This departure from the median was due to relatively rare lakes with very high emission rates pulling the mean towards the direction of positive skew **(SI Fig.
+This departure from the median was due to relatively rare lakes with very high emission rates pulling the mean towards the direction of positive skew **(SI_1 Fig.
 1)**.
@@
 C) Proportion of total CONUS N~2~O flux attributed to each of the 465,896 lakes in the POI.
-Points (red) and bars (black) reflect the posterior means and 95% credible intervals, respectively, based on 2000 draws from the posterior predictive distribution for each waterbody](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
+Points (red) and bars (black) reflect the posterior means and 95% credible intervals, respectively, based on 2000 draws from the posterior predictive distribution for each waterbody.](manuscript_figures/n2oFluxAndEmissionRateVsContinuousArea.tiff)
 
@@
 
-Our observations of dissolved N~2~O were well predicted using a regression model with interactions between observed water temperature, NO~3~ concentration, and waterbody size (see SI).
+Our observations of dissolved N~2~O were well predicted using a regression model with interactions between observed water temperature, NO~3~ concentration, and waterbody size (**SI_1**).
 These modeling results suggested that NO~3~ levels may play a role in determining whether a waterbody functions as a sink or source of N~2~O, and the magnitude of the source or sink may be exaggerated with increasing water temperature.
@@
 The measurement methods employed in this study have a 95% credible interval of 0.11 (unitless) on the N~2~O saturation ratio.
-The 95% CI can be halved by employing gas chromatography optimized for quantifying N~2~O in the range of values observed in this study [@Hall_2024] and decreased by an order of magnitude using membrane inlet mass spectrometry[@Genthner_etal_2013; @Speir_etal_2023] **(SI)**.
+The 95% CI can be halved by employing gas chromatography optimized for quantifying N~2~O in the range of values observed in this study [@Hall_2024] and decreased by an order of magnitude using membrane inlet mass spectrometry[@Genthner_etal_2013; @Speir_etal_2023] **(SI_2)**.
 
@@
 A better understanding of the temporal evolution of N~2~O production and consumption is an important research need to determine the extent to which small lakes can remove excess nitrogen from surface waters and simultaneously mitigate N~2~O emissions.
+
+# Disclaimer
+The views expressed in this article are those of the authors and do not necessarily represent the views or policies of the U.S. Environmental Protection Agency.
 
@@
 
+The following is an abbreviated presented of the statistical analysis; see **SI_1** for more details.
 The survey samples were treated as a randomly selected subset of the POI, but with a pattern of "missingess" prescribed by the selection algorithm [@Gelman_etal_2014; @Zachmann_etal_2022].
-For inferences, estimates were statistically adjusted to account for the selection mechanism using a Multilevel regression with poststratification (MRP) approach [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020].
+For inferences, estimates were statistically adjusted to account for the selection mechanism using a multilevel regression with poststratification (MRP) approach [@Gelman_Little_1997; @Park_etal_2004; @Gelman_etal_2020].
 MRP is a model-based approach carried out in two steps.
@@
 Including covariates, however, necessitated additional steps in the MRP process.
-To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled as with the N2O responses.
+To make inferences to the POI, the covariates needed to be (1) fully observed across the POI or (2) their missingness needed to be modeled as with the N~2~O responses.
 For the lake area and elevation covariates, data was available for all lakes in the POI from previously compiled geospatial databases.

### 2025-03-17 manuscript/manuscript_support.R

@@ -293,7 +293,7 @@ national_stats <- all_predictions_ms %>%
              UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
              total_sa = round(mean(total_sa), digits = 0),
              n = dplyr::n()) %>%
-  mutate(WSA9 = "national", WSA9_NAME = "national")
+  mutate(WSA9 = "National", WSA9_NAME = "National")
 
 save(national_stats, file = "manuscript/manuscript_files/national_stats.rda")
 } # national_stats

### 2025-03-18 scripts/NLA_N2O_models.Rmd
Updated modeling script and documentation

@@
 
-```{r conditional_effects_mod_n2osat_temp_F, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+```{r conditional_effects_mod_n2osat_temp_F, echo=FALSE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
 effect_no3_surftemp <- expand.grid(no3_cat = levels(df_model$no3_cat),
@@
 
-legend1 <- get_legend(legend1)
+legend1 <- ggpubr::get_legend(legend1)
 
@@
 
-legend2 <- get_legend(legend2)
+legend2 <- ggpubr::get_legend(legend2)
 
@@
 
-```{r conditional_effects_mod_n2osat_area_F, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+```{r conditional_effects_mod_n2osat_area_F, echo=FALSE, message=FALSE, warning=FALSE, cache.lazy = FALSE}
 effect_no3_area <- expand.grid(no3_cat = levels(df_model$no3_cat),
@@
 
-```{r plot_area_effects, fig.align='center', fig.asp=1.166667, fig.width=8, }
+```{r plot_area_effects, fig.align='center', fig.asp=1.166667, fig.width=8}
 plt_area_effects
@@
 
-```{r conditional_effects_mod_n2osat_temp_area_F, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+```{r conditional_effects_mod_n2osat_temp_area_F, echo=FALSE, message=FALSE, warning=FALSE, cache.lazy=FALSE}
 effect_no3_surftemp <- expand.grid(no3_cat = levels(df_model$no3_cat),
@@
 
-```{r plot_temp_area_effects, fig.align='center', fig.asp=1.166667, fig.width=8, }
+```{r plot_temp_area_effects, fig.align='center', fig.asp=1.166667, fig.width=8}
 plt_temp_area_effects

### 2025-04-05 manuscript/manuscript_file.Rmd
Minor grammar corrections following Susan Cormier's review

@@
 **Lakes and reservoirs are estimated to be globally important sources of nitrous oxide (N~2~O) to the atmosphere but recent evidence of N~2~O uptake across a broad range of lakes have called the accuracy of emission estimates into question.
-Here we use a new national-scale dataset on dissolved N~2~O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` waterbodies in the conterminous U.S.
+Here we use a new national-scale dataset of dissolved N~2~O and a Bayesian hierarchical model to predict N~2~O concentration and emission rates in `r sframe %>% summarize(nlake = n()) %>% pull() %>% format(., big.mark=",", scientific=FALSE)` waterbodies in the conterminous U.S.
 (CONUS).
-We found that N~2~O undersaturation was pervasive through the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of lakes functioning as N~2~O sinks.
+We found that N~2~O undersaturation was pervasive throughout the CONUS, with an estimated `r national_stats$post_m_prop_sat * 100`% (95% credible interval: `r national_stats$LCL_prop_sat * 100` - `r national_stats$UCL_prop_sat * 100`%) of lakes functioning as N~2~O sinks.
 The model predicts dissolved N~2~O concentrations reasonably well based partly on interactions between nitrate concentration, waterbody surface area, and water temperature.
@@
 2)**.
-Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had a median value less than one, indicating that most lakes in these populations were undersaturated .
+Across the CONUS, and within each of the nine major ecoregions, the predicted distribution of N~2~O saturation ratios had a median value less than one, indicating that most lakes in these populations were undersaturated.
 However, the long right tails on these distributions, comprised of relatively rare lakes with very high levels of N~2~O supersaturation, were sufficient to drive the population means to values close to, or even greater than one.

### 2025-04-07 manuscript/manuscript_support.R
Adding code to read all_predictions.rda from Zenodo if not available locally

@@ -97,7 +97,16 @@ if(!("ecoreg_stats.rda" %in% list.files("manuscript/manuscript_files"))) {
       
 ## load posterior predictions from modeling----
 #tic() # 7.6min on DMAP.  5 minutes on Dell Precision workstation
-load("./inputData/all_predictions.rda")
+if(!("all_predictions.rda" %in% list.files("inputData"))){
+  inUrl1  <- "https://zenodo.org/records/15159394/files/all_predictions.rda?download=1"
+  infile1 <- tempfile()
+  try(download.file(inUrl1,infile1,method="curl")) 
+  if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")
+  
+  save(all_predictions, file = "inputData/all_predictions.rda")
+} else {
+  load("inputData/all_predictions.rda")
+}
 
 #toc()
 
